### 2009-02-13 docs/icccn2009/paper.tex
Paper skeleton for ICCCN 2009
git-svn-id: svn://localhost/netbeams/trunk@256 559edde2-a4d4-44e0-96fc-4035e40d7396

\documentclass[conference]{IEEEtran}

\usepackage{epsfig}
\usepackage{fancyvrb}

\DefineVerbatimEnvironment%
  {code}{Verbatim}{numbers=left,numbersep=3pt,frame=lines,%
                   xleftmargin=7pt,fontsize=\footnotesize}

\begin{document}

\title{An OSGi-based Sensor Network for \\ Environmental Monitoring}

\author{\authorblockN{XXX}
\authorblockA{San Francisco State University \\
Computer Science Department \\
1600 Holloway Avenue \\
San Francisco, CA 94132 \\
EMail: XXX@sfsu.edu}}

\maketitle
\begin{abstract}
  Abstract.
\end{abstract}

\section{Introduction}

The introduction.

%\bibliography{../literature/lit}
%\bibliographystyle{plain}

\end{document}

### 2009-02-13 versions/v2/apps/osgi-bundles/dsp/DSPJSwingMouseClient/src/org/netbeams/dsp/demo/mouseactions/osgi/MouseActionsActivator.java
- Updates the xml samples message for the DSP Wire Transport to have the bootstrap   message delivered; The name of the DSP component must be given to the bootstrap message; - Updates to the DSP Wire Transport Client to enable the Marshall/Unmarshall of   messages (adding ObjectFactory from DSProperties); - Fixing the Mouse Actions to open only one client.
git-svn-id: svn://localhost/netbeams/trunk@258 559edde2-a4d4-44e0-96fc-4035e40d7396

@@ -34,9 +34,7 @@ public class MouseActionsActivator implements BundleActivator {
                 
         log.debug("Instantiating the mouse actions component...");
         this.producer = new MouseActionDSPComponent();
-        this.producer.startComponent();
         this.serviceRegistration = ActivatorHelper.registerOSGIService(bc, this.producer);
-
     }
 
     public void stop(BundleContext bc) throws Exception {

### 2009-02-13 versions/v2/apps/osgi-bundles/dsp/DSPWireTransportClient/src/org/netbeams/dsp/wiretransport/client/controller/DSPWireTransportHttpClient.java

@@ -74,8 +74,6 @@ public class DSPWireTransportHttpClient implements DSPComponent {
      */
     public DSPWireTransportHttpClient() {
         log.info("Starting DSP Transport Client...");
-        log.info("Scheduling the transport senders to wake up at every 60 seconds...");
-        scheduler.scheduleWithFixedDelay(new DspTransportSender(), 0, 60, TimeUnit.SECONDS);
     }
 
     /**
@@ -349,6 +347,9 @@ public class DSPWireTransportHttpClient implements DSPComponent {
     public void startComponent() throws DSPException {
         log.info("Starting component");
 
+        log.info("Scheduling the transport senders to wake up at every 60 seconds...");
+        scheduler.scheduleWithFixedDelay(new DspTransportSender(), 0, 60, TimeUnit.SECONDS);
+        
         String destIpAddress = null, hostName = null;
         try {
             destIpAddress = NetworkUtil.getCurrentEnvironmentNetworkIp();

### 2009-02-13 versions/v2/apps/osgi-bundles/dsp/DSPWireTransportServer/src/org/netbeams/dsp/wiretransport/server/osgi/DSPWireTransportServerActivator.java

@@ -32,7 +32,7 @@ public class DSPWireTransportServerActivator implements BundleActivator {
     /**
      * The service reference to the DSP consumer 
      */
-    private DSPWireTransportServer consumer;
+    private DSPWireTransportServer serverConsumer;
 
     /*
      * (non-Javadoc)
@@ -44,8 +44,8 @@ public class DSPWireTransportServerActivator implements BundleActivator {
         
         this.bundleContext = bc;
 
-        this.consumer = new DSPWireTransportServer(this.bundleContext);
-        this.consumer.startComponent();
+        this.serverConsumer = new DSPWireTransportServer(this.bundleContext);
+        this.serviceRegistration = ActivatorHelper.registerOSGIService(bundleContext, this.serverConsumer);
     }
     
     /*
@@ -55,7 +55,7 @@ public class DSPWireTransportServerActivator implements BundleActivator {
      */
     public void stop(BundleContext bc) throws Exception {
         log.info("WireTransport.Activator.stop()");
-        this.consumer.stopComponent();
+        this.serverConsumer.stopComponent();
         ActivatorHelper.unregisterOSGIService(this.bundleContext, this.serviceRegistration);
     }
 }

### 2009-02-13 versions/v2/apps/osgi-bundles/dsp/DSPJSwingMouseClient/src/org/netbeams/dsp/demo/mouseactions/controller/DSPMouseActionsProducer.java
- Updating the DSPMouseActionsProducer to send messages to the correct client; - This is the way that components need to write messages (indicating the destination the DSP component name);
This strategy should be just reflected on the match_config.xml rules instead.

- Removing the old type of messages.

git-svn-id: svn://localhost/netbeams/trunk@259 559edde2-a4d4-44e0-96fc-4035e40d7396

@@ -16,7 +16,6 @@ import org.netbeams.dsp.demo.mouseactions.MouseAction;
 import org.netbeams.dsp.demo.mouseactions.MouseActionsContainer;
 import org.netbeams.dsp.demo.mouseactions.ObjectFactory;
 import org.netbeams.dsp.demo.mouseactions.model.NetBeamsMouseInfo;
-import org.netbeams.dsp.demo.mouseactions.model.dsp.MouseActionDSPComponent;
 import org.netbeams.dsp.message.ComponentIdentifier;
 import org.netbeams.dsp.message.DSPMessagesFactory;
 import org.netbeams.dsp.message.Header;
@@ -62,7 +61,7 @@ public class DSPMouseActionsProducer implements NetBeamsMouseListener {
     public void trackMouseActionUpdate(NetBeamsMouseInfo netBeamsMouseInfo) {
         log.debug("Mouse Actions Local Memory size: " + this.localMemory.size());
         this.localMemory.add(netBeamsMouseInfo);
-        if (this.localMemory.size() <= 30) {
+        if (this.localMemory.size() < 30) {
             log.debug("Tracking updates since there are less than 30 items on the producer local memory...");
         } else {
             log.debug("Sending the Items to the Messages Queues...");
@@ -80,10 +79,10 @@ public class DSPMouseActionsProducer implements NetBeamsMouseListener {
         log.debug("Total number of Mouse Actions: " + data.getMouseAction().size());
         
         ComponentIdentifier producer = DSPMessagesFactory.INSTANCE.makeDSPComponentIdentifier(
-                MouseActionDSPComponent.class.getName(), localIPAddress, data.getContentContextForJAXB());
+                "manual-producer-id", localIPAddress, data.getContentContextForJAXB());
         
-        ComponentIdentifier consumer = DSPMessagesFactory.INSTANCE.makeDSPComponentIdentifier("DSPWebLogger",
-                System.getProperty("WIRE_TRANSPORT_SERVER_IP"), null);
+        ComponentIdentifier consumer = DSPMessagesFactory.INSTANCE.makeDSPComponentIdentifier(
+                "DSPWireTransportClient", System.getProperty("WIRE_TRANSPORT_SERVER_IP"), "org.netbeams.dsp.wiretransport.client");
         // Note that the consumer has the Wire Transport default DEMO IP as the destination.
         
         log.debug("Packaging Mouse Actions to be sent to "

### 2009-02-13 versions/v2/apps/osgi-bundles/dsp/DSPJSwingMouseClient/src/org/netbeams/dsp/demo/mouseactions/controller/DSPMouseActionsProducer.java
- Updates to the DSP Wire Transport Client/Server to have them working; - Adding the value that specifies the delay of how often the DSP Wire   Transport Client transfers Messages to the server on the bootstrap messages;   See DSPWireTransportClient_001.xml - Updating the MANIFEST.MF to have everything running again.
git-svn-id: svn://localhost/netbeams/trunk@260 559edde2-a4d4-44e0-96fc-4035e40d7396

@@ -79,7 +79,7 @@ public class DSPMouseActionsProducer implements NetBeamsMouseListener {
         log.debug("Total number of Mouse Actions: " + data.getMouseAction().size());
         
         ComponentIdentifier producer = DSPMessagesFactory.INSTANCE.makeDSPComponentIdentifier(
-                "manual-producer-id", localIPAddress, data.getContentContextForJAXB());
+                "DSPJSwingMouseProducer", localIPAddress, data.getContentContextForJAXB());
         
         ComponentIdentifier consumer = DSPMessagesFactory.INSTANCE.makeDSPComponentIdentifier(
                 "DSPWireTransportClient", System.getProperty("WIRE_TRANSPORT_SERVER_IP"), "org.netbeams.dsp.wiretransport.client");

### 2009-02-13 versions/v2/apps/osgi-bundles/dsp/DSPWireTransportClient/src/org/netbeams/dsp/wiretransport/client/controller/DSPWireTransportHttpClient.java

@@ -123,15 +123,15 @@ public class DSPWireTransportHttpClient implements DSPComponent {
             URL destinationURL = null;
             try {
                 MessagesQueues messagesQueue = MessagesQueues.INSTANCE;
-                MessagesContainer[] msgContainers = messagesQueue.retrieveAllQueuedMessagesForTransmission();
                 thLog.debug("Retrieving all messages from the Messages Queue for transmission...");
-                thLog.debug("Messages Container size is " + msgContainers.length + ".");
+                MessagesContainer[] msgContainers = messagesQueue.retrieveAllQueuedMessagesForTransmission();
+                thLog.debug("There are " + msgContainers.length + " Messages Containers...");
 
                 if (msgContainers.length > 0) {
                     thLog.debug("Sending each Messages Container to their own destinition...");
                 } else {
                     thLog.debug("No messages registered in the Messages Queue...");
-                    thLog.debug("Stopping the Wire Transport for 60 seconds...");
+                    thLog.debug("Stopping the Wire Transport for " + getDelayForTransportSender() + " seconds...");
                 }
 
                 for (MessagesContainer messagesForRequest : msgContainers) {
@@ -144,8 +144,9 @@ public class DSPWireTransportHttpClient implements DSPComponent {
                         String messagesForRequestInXml = serializeMessagesContainer(messagesForRequest);
                         thLog.trace("HTTP Request BODY: " + messagesForRequestInXml);
 
-                        destinationURL = new URL("http://" + messagesForRequest.getHost() + ":"
-                                + System.getProperty("WIRE_TRANSPORT_SERVER_PORT") + System.getProperty("BASE_URI"));
+                        destinationURL = new URL("http://" + messagesForRequest.getHost() + ":" + 
+                                System.getProperty("WIRE_TRANSPORT_SERVER_PORT") + 
+                                System.getProperty("HTTP_SERVER_BASE_URI"));
 
                         thLog.trace("Setting the messages in the container " + messagesForRequest.getUudi()
                                 + " to the 'transmitted' state...");
@@ -193,7 +194,7 @@ public class DSPWireTransportHttpClient implements DSPComponent {
          * @param messagesQueue is the current MessagesQueues singleton instance.
          * @param messagesForRequest is the messages used for the request
          * @param messagesFromResponseInXml is the response
[…]

### 2009-02-13 versions/v2/apps/osgi-bundles/dsp/DSPWireTransportClient/src/org/netbeams/dsp/wiretransport/client/model/MessagesQueues.java

@@ -76,11 +76,9 @@ public enum MessagesQueues {
      */
     public synchronized MessagesContainer retrieveQueuedMessagesForTransmission(String destinitionIp) {
         MessagesContainer container = DSPMessagesFactory.INSTANCE.makeDSPMessagesContainer(destinitionIp);
-        log.debug("The Messages queue is being queried for the ip address " + destinitionIp);
         Queue<QueueMessageData> outboutQueue = this.outboundQueue.get(destinitionIp);
         if (outboutQueue != null) {
             log.debug("The size of the messages in the outbound queue is " + outboutQueue.size());
-
             for (QueueMessageData data : outboutQueue) {
                 if (data.getState().equals(QueueMessageState.QUEUED)) {
                     container.getMessage().add(data.getMessage());
@@ -99,6 +97,7 @@ public enum MessagesQueues {
         MessagesContainer[] msgContainers = new MessagesContainer[ipAddresses.size()];
         int i = -1;
         for (String ipAddr : ipAddresses) {
+            log.debug("The Messages queue is being queried for the ip address " + ipAddr);
             msgContainers[++i] = this.retrieveQueuedMessagesForTransmission(ipAddr);
         }
         return msgContainers;

### 2009-02-13 versions/v2/apps/osgi-bundles/dsp/DSPWireTransportServer/src/org/netbeams/dsp/wiretransport/server/controller/DSPWireTransportHttpReceiverServlet.java

@@ -23,6 +23,7 @@ import org.netbeams.dsp.message.Message;
 import org.netbeams.dsp.message.MessagesContainer;
 import org.netbeams.dsp.message.content.AckType;
 import org.netbeams.dsp.message.content.AcksContainer;
+import org.netbeams.dsp.util.NetworkUtil;
 import org.netbeams.dsp.wiretransport.client.controller.DSPWireTransportHttpClient;
 import org.netbeams.dsp.wiretransport.client.model.MessagesQueues;
 
@@ -67,6 +68,7 @@ public class DSPWireTransportHttpReceiverServlet extends HttpServlet {
         if (messagesContainerXml != null && !"".equals(messagesContainerXml)) {
             try {
                 responseMessagesContainer = this.deliverMessagesToLocalDSP(messagesContainerXml);
+                log.debug("Messages Container for the response is ready to be serialized..." + responseMessagesContainer);
             } catch (JAXBException e) {
                 log.error(e.getMessage(), e);
                 throw new IllegalArgumentException(e.getMessage());
@@ -82,17 +84,15 @@ public class DSPWireTransportHttpReceiverServlet extends HttpServlet {
                     + System.getProperty("HTTP_SERVER_REQUEST_VARIABLE") + "is missing");
         }
 
-        // The following generates a page showing all the request parameters
-        PrintWriter out = resp.getWriter();
-        resp.setContentType("text/xml");
-
         try {
+            // The following generates a page showing all the request parameters
+            PrintWriter out = resp.getWriter();
+            resp.setContentType("text/xml");
             out.println(DSPWireTransportHttpClient.serializeMessagesContainer(responseMessagesContainer));
-        } catch (JAXBException e) {
+            out.close();
+        } catch (Exception e) {
             log.error(e.getMessage(), e);
         }
-
-        out.close();
     }
 
     /**
@@ -111,25 +111,29 @@ public class DSPWireTransportHttpReceiverServlet extends HttpServlet {
     private MessagesContainer deliverMessagesToLocalDSP(String messagesContainerRequestXML) throws JAXBException,
             DSPException {
 
+        log.debug("Retrieved messages container from the HTTP request... Deserializing XML from the client...");
         // Deliver all the messages to the DSP Broker component
         MessagesContainer requestMessagesContainer = DSPWireTransportHttpClient
                 .deserializeMessagesContainer(messagesContainerRequestXML);
-        log.debug("Retrieved messages container from the HTTP request");
 
[…]

### 2009-02-16 docs/icccn2009/paper.tex
First stab at an outline
git-svn-id: svn://localhost/netbeams/trunk@296 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 
-The introduction.
+The introduction should emphasize that we interpret sensor networks in
+a different way: our focus is on off-the-shelf, highly specialized
+sensors for environmental monitoring. Our focus is end-to-end: how to
+collect data from a remote location with no stationary power or
+Internet connection and deliver the data in near-realtime to a web
+browser. Length: first page (including title and abstract).
+
+\section{Use Case}
+
+Start with some related work. Who else has done an OSGi-based
+infrastructure for sensor networks? Kleber: I believe you have a few
+references. This section should be a combination of a use case (using
+the RTC and their YSI sonde as an example) and a requirements
+analysis: requirements such as remote management and monitoring
+capabilities, time-delayed communication, etc. In this section should
+be no mentioned of DSP or OSGi. Length: 1 page.
+
+\section{DSP}
+
+Should describe the OSGi-based NetBEAMS architecture. Total length of
+this section: 2 pages.
+
+\subsection{Data Sensor Framework and Platform}
+
+should include a picture of a standalone DSP showing a DC and a DP
+(general description only). Explain difference between DSP Framework
+and DSP Platform. This section should give a top-level overview.
+Subsequent sub-sections should explain certain features of the DSP in
+more detail.
+
+\subsection{DP and DC}
+
+Should give some API sniplets for DP and DC.
+
+\subsection{DSP Message}
+
+Should explain the way DSP Messages are defined: XML Schema and JAXB.
+
+\subsection{Matcher}
+
+\section{NetBEAMS}
+
+Should explain some specific DC and DP we have implemented for
+NetBEAMS. Length: 1 - 1.5 pages.
+
+\subsection{YSI Sonde DP}
+
+\subsection{Wire transport DC/DP}
+
+\subsection{Web Managements}
+
+\section{Conclusions and Outlook}
+
+Length (including bibliography): 0.5 pages.
 

### 2009-02-25 docs/icccn2009/paper.tex
Started to work on diagrams
git-svn-id: svn://localhost/netbeams/trunk@315 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 
+Researchers in the geosciences rely on sensor data collected by highly
+specialized sensors such as the Seabird or YSI. These sensors are
+usually deployed in remote locations that not only make their
+maintenance difficult, but also access to their measurements.
+Typically the sensors are autonomous in the sense that they can run
+for a certain time on battery power and store the measurement data in
+internal memory. Only when the sensors are serviced, the measurement
+data they have collected since the last servicing are uploaded to a
+portable storage device. From there the data can be uploaded to
+server. This manual procedure to retrieve sensor data is not only
+error prone, but also results in a serious time lag between the time
+the sensor performs the measurement and when it becomes available for
+further processing. For many applications it would be beneficial to
+have near-realtime access to the sensor measurement. Being able to
+access data when they become available on the sensor device also
+reduces the maintenance overhead.
+
 Start with some related work. Who else has done an OSGi-based
@@
 more detail.
+
+\begin{figure}
+\centering
+\epsfig{file=dsp, width=7cm}
+\caption{\label{FIG_DSP} Data Sensor Platform (DSP).}
+\end{figure}
 

### 2009-02-25 versions/v2/apps/osgi-bundles/dsp/DSPWebDataLogger/src/org/netbeams/dsp/weblogger/DataCollector.java
First version of a generic UI for DSP Properties. The UI can be built at runtime via PropertyUI. Currently supported are buttons, labels, and input fields. Changes are pushed to the browser where they are rendered. See DataCollector.testPropertyUI() for an example.
git-svn-id: svn://localhost/netbeams/trunk@316 559edde2-a4d4-44e0-96fc-4035e40d7396

@@ -2,6 +2,8 @@ package org.netbeams.dsp.weblogger;
 
 
 
+import java.util.ArrayList;
+
 import javax.servlet.ServletException;
 
 import org.apache.log4j.Logger;
@@ -99,6 +101,8 @@ public class DataCollector
                     new GetDataServlet(theBuffer), null, null);
             httpService.registerServlet(WebAppServlet.BASE_URI,
                     new WebAppServlet(), null, null);
+            httpService.registerServlet(PropertyUIServlet.BASE_URI,
+                    new PropertyUIServlet(), null, null);
         }
         catch (ServletException e) {
             log.error(e.getMessage(), e);
@@ -106,9 +110,23 @@ public class DataCollector
         catch (NamespaceException e) {
             log.error(e.getMessage(), e);
         }
+        testPropertyUI();
     }
 
 
+    private void testPropertyUI()
+    {
+        PropertyUI.createButton(0, 0, "Button @ 0, 0");
+        PropertyUI.createInput(1, 0, "Input @ 1, 0");
+        PropertyUI.createLabel(2, 0, "Label @ 2, 0");
+        PropertyUI.createLabel(2, 1, "Label @ 2, 1");
+        ArrayList<String> components = new ArrayList<String>();
+        components.add("YSI Sonde");
+        components.add("Remote Communication");
+        PropertyUI.setComponents(components);
+        //PropertyUI.clear();
+
+    }
 
     //@Override
     public void stopComponent()

### 2009-02-25 versions/v2/apps/osgi-bundles/dsp/DSPWebDataLogger/src/org/netbeams/dsp/weblogger/PropertyUIServlet.java

@@ -1,0 +1,54 @@
+
+package org.netbeams.dsp.weblogger;
+
+import java.io.IOException;
+
+import javax.servlet.ServletException;
+import javax.servlet.ServletOutputStream;
+import javax.servlet.http.HttpServlet;
+import javax.servlet.http.HttpServletRequest;
+import javax.servlet.http.HttpServletResponse;
+
+import org.apache.log4j.Logger;
+
+
+
+public class PropertyUIServlet
+    extends HttpServlet
+{
+
+    private static final Logger log = Logger.getLogger(PropertyUIServlet.class);
+    
+    final public static String BASE_URI     = "/property-ui";
+
+
+    public PropertyUIServlet()
+    {
+    }
+
+
+
+    @Override
+    protected void doGet(HttpServletRequest request,
+            HttpServletResponse response)
+        throws ServletException, IOException
+    {
+        String uri = request.getRequestURI();
+        log.debug("PropertyUIServlet Processing URI: " + uri);
+        if (uri.indexOf("/PULL") != -1) {
+        	response.setContentType("text/plain");
+        	ServletOutputStream out = response.getOutputStream();
+        	String data = PropertyUI.getData();
+        	out.println(data);
+        }
+        else {
+        	// User pressed button on Property UI
+        }
+    }
+
+
+
+    private static final long serialVersionUID = ("urn:" + PropertyUIServlet.class
+                                                       .getName()).hashCode();
+
+}

### 2009-02-25 versions/v2/apps/osgi-bundles/dsp/DSPWebDataLogger/web/DSP Management/source/class/dsp_management/Application.js

@@ -1,0 +1,239 @@
+
+/* ************************************************************************
+
+#asset(dsp_management/*)
+
+************************************************************************ */
+
+function DSPDataFetcher()
+{
+  DSPDataFetcher.list = null;
+  
+
+  DSPDataFetcher.callback = function(e)
+  {
+    var items = e.getContent().split("$");
+    for (var i = 0; i < items.length; i++) {
+      var item = new qx.ui.form.ListItem(items[i]);
+      DSPDataFetcher.list.add(item);
+      DSPDataFetcher.list.scrollChildIntoView(item);
+    }
+    DSPDataFetcher.loop();
+  }
+  
+  DSPDataFetcher.loop = function()
+  {
+    var req = new qx.io.remote.Request("/get-data", "GET", "text/plain");
+    req.setTimeout(600000);
+    req.addListener("completed", DSPDataFetcher.callback);
+    //req.addListener("receiving", function(e) { alert("receiving");});
+    //req.addListener("failed", function(e) { alert("failed");});
+    //req.addListener("timeout", function(e) { alert("timeout");});
+    //req.addListener("aborted", function(e) { alert("aborted");});
+    req.send();
+  }
+}
+DSPDataFetcher();
+
+
+function PropertyUI()
+{
+	  PropertyUI.components = null;
+	  PropertyUI.grid = null;
+	  PropertyUI.offset = 0;
+	  
+	  PropertyUI.widgets = new Array();
+	  
+	  PropertyUI.processInstruction = function(inst)
+	  {
+		  if (inst.action == "CREATE")
+			  PropertyUI.doCreate(inst);
+		  if (inst.action == "CLEAR")
+			  PropertyUI.doClear(inst);
+		  if (inst.action == "SET_COMPONENTS")
+			  PropertyUI.doSetComponents(inst);
+	  }
+	  
+	  PropertyUI.doCreate = function(inst)
+	  {
+		  var r = inst.row + PropertyUI.offset;
+		  var c = inst.column;
+		  
+		  var widget = null;
+		  if (inst.type == "BUTTON") {
+			  widget = new qx.ui.form.Button(inst.text);
+			  widget.addListener("execute", function(e) {
+				  PropertyUI.buttonPressed(e);
+			  });
+		  }
+		  if (inst.type == "LABEL")
+			  widget = new qx.ui.basic.Label(inst.text);
+		  if (inst.type == "INPUT")
+			  widget = new qx.ui.form.TextField();
[…]

### 2009-02-25 versions/v2/apps/osgi-bundles/dsp/DSPWebDataLogger/web/DSP Management/source/class/dsp_management/test/DemoTest.js

@@ -1,0 +1,55 @@
+/* ************************************************************************
+
+   Copyright:
+
+   License:
+
+   Authors:
+
+************************************************************************ */
+
+/**
+ * This class demonstrates how to define unit tests for your application.
+ *
+ * Execute <code>generate.py test</code> to generate a testrunner application 
+ * and open it from <tt>test/index.html</tt>
+ *
+ * The methods that contain the tests are instance methods with a 
+ * <code>test</code> prefix. You can create an arbitrary number of test 
+ * classes like this one. They can be organized in a regular class hierarchy, 
+ * i.e. using deeper namespaces and a corresponding file structure within the 
+ * <tt>test</tt> folder.
+ */
+qx.Class.define("dsp_management.test.DemoTest",
+{
+  extend : qx.dev.unit.TestCase,
+
+  members :
+  {
+    /*
+    ---------------------------------------------------------------------------
+      TESTS
+    ---------------------------------------------------------------------------
+    */
+  
+    /**
+     * Here are some simple tests
+     */
+    testSimple : function()
+    {
+      this.assertEquals(4, 3+1, "This should never fail!");
+      this.assertFalse(false, "Can false be true?!");
+    },
+
+    /**
+     * Here are some more advanced tests
+     */
+    testAdvanced: function () 
+    {
+      var a = 3;
+      var b = a;
+      this.assertIdentical(a, b, "A rose by any other name is still a rose");
+      this.assertInRange(3, 1, 10, "You must be kidding, 3 can never be outside [1,10]!");
+    }
+  }
+});

### 2009-02-25 versions/v2/apps/osgi-bundles/dsp/DSPWebDataLogger/web/DSP Management/source/class/dsp_management/theme/Color.js

@@ -1,0 +1,18 @@
+/* ************************************************************************
+
+   Copyright:
+
+   License:
+
+   Authors:
+
+************************************************************************ */
+
+qx.Theme.define("dsp_management.theme.Color",
+{
+  extend : qx.theme.modern.Color,
+
+  colors :
+  {
+  }
+});

### 2009-02-25 versions/v2/apps/osgi-bundles/dsp/DSPWebDataLogger/web/DSP Web Logger/build/script/dsp_web_logger.js

@@ -1,314 +1,0 @@
-if(!window.qxsettings)qxsettings={};
-qxsettings["qx.application"]="dsp_web_logger.Application";
-qxsettings["qx.theme"]="qx.theme.Modern";
-qxsettings["qx.version"]="0.8";
-
-if(!window.qxvariants)qxvariants={};
-qxvariants["qx.debug"]="off";
-
-if(!window.qxlibraries)qxlibraries={};qxlibraries["qx"]={"resourceUri":"./resource"};qxlibraries["dsp_web_logger"]={"resourceUri":"./resource"};
-
-if(!window.qxresources)qxresources={"qx/decoration/Modern/shadow/shadow-small-r.png": [5, 136, "png", "qx", "qx/decoration/Modern/shadow-small-lr-combined.png", 0, 0], "qx/decoration/Modern/window/captionbar-inactive-br.png": [6, 6, "png", "qx", "qx/decoration/Modern/window-captionbar-tb-inactive-combined.png", 0, 0], "qx/icon/Tango/16/actions/view-refresh.png": [16, 16, "png", "qx"], "qx/decoration/Modern/form/button-t.png": [4, 4, "png", "qx", "qx/decoration/Modern/button-tb-combined.png", 0, -176], "qx/decoration/Modern/tabview-pane-tb-combined.png": [30, 180, "png", "qx"], "qx/decoration/Modern/form/radiobutton-disabled.png": [14, 14, "png", "qx", "qx/decoration/Modern/inputcheckradio-combined.png", -196, 0], "qx/decoration/Modern/tabview/tab-button-bottom-inactive-b.png": [3, 3, "png", "qx", "qx/decoration/Modern/tabview-button-bottom-inactive-b-combined.png", 0, 0], "qx/decoration/Modern/form/button-focused-tl.png": [4, 4, "png", "qx", "qx/decoration/Modern/button-tb-combined.png", 0, -144], "qx/icon/Tango/22/mimetypes/office-document.png": [22, 22, "png", "qx"], "qx/decoration/Modern/shadow/shadow-r.png": [15, 382, "png", "qx", "qx/decoration/Modern/shadow-lr-combined.png", 0, 0], "qx/decoration/Modern/shadow/shadow-tl.png": [15, 15, "png", "qx", "qx/decoration/Modern/shadow-tb-combined.png", 0, -30], "qx/decoration/Modern/window/minimize-active-hovered.png": [9, 9, "png", "qx", "qx/decoration/Modern/window-captionbar-buttons-combine
[…]

### 2009-02-25 versions/v2/apps/osgi-bundles/dsp/DSPWebDataLogger/web/DSP Web Logger/generate.py

@@ -1,50 +1,0 @@
-#!/usr/bin/env python
-################################################################################
-#
-#  qooxdoo - the new era of web development
-#
-#  http://qooxdoo.org
-#
-#  Copyright:
-#    2008 1&1 Internet AG, Germany, http://www.1und1.de
-#
-#  License:
-#    LGPL: http://www.gnu.org/licenses/lgpl.html
-#    EPL: http://www.eclipse.org/org/documents/epl-v10.php
-#    See the LICENSE file in the project's top-level directory for details.
-#
-#  Authors:
-#    * Thomas Herchenroeder (thron7)
-#
-################################################################################
-
-##
-# This is a stub proxy for the real generator.py
-##
-
-import sys, os, subprocess
-
-CMD_PYTHON = 'python'
-QOOXDOO_PATH = '../../../../../../../../../../home/arno/src/qooxdoo-0.8-sdk'
-REAL_GENERATOR = os.path.normpath(
-                    os.path.join(os.path.abspath(os.path.dirname(sys.argv[0])),
-                                 QOOXDOO_PATH, 'tool', 'bin', 'generator.py'))
-                                 
-os.chdir(os.path.dirname(sys.argv[0]))  # switch to skeleton dir
-
-argList = []
-argList.append(CMD_PYTHON)
-argList.append(REAL_GENERATOR)
-argList.extend(sys.argv[1:])
-if sys.platform == "win32":
-    argList1=[]
-    for arg in argList:
-        if arg.find(' ')>-1:
-            argList1.append('"%s"' % arg)
-        else:
-            argList1.append(arg)
-    argList = argList1
-else:
-    argList = ['"%s"' % x for x in argList]  # quote argv elements
-cmd = " ".join(argList)
-subprocess.call(cmd, shell=True)

### 2009-02-25 versions/v2/apps/osgi-bundles/dsp/DSPWebDataLogger/web/DSP Web Logger/source/class/dsp_web_logger/Application.js

@@ -1,97 +1,0 @@
-
-/* ************************************************************************
-
-#asset(dsp_web_logger/*)
-
-************************************************************************ */
-
-function DSPDataFetcher()
-{
-  DSPDataFetcher.list = null;
-  
-
-  DSPDataFetcher.callback = function(e)
-  {
-    var items = e.getContent().split("$");
-    for (var i = 0; i < items.length; i++) {
-      var item = new qx.ui.form.ListItem(items[i]);
-      DSPDataFetcher.list.add(item);
-      DSPDataFetcher.list.scrollChildIntoView(item);
-    }
-    DSPDataFetcher.loop();
-  }
-  
-  DSPDataFetcher.loop = function()
-  {
-    var req = new qx.io.remote.Request("/get-data", "GET", "text/plain");
-    req.setTimeout(600000);
-    req.addListener("completed", DSPDataFetcher.callback);
-    //req.addListener("receiving", function(e) { alert("receiving");});
-    //req.addListener("failed", function(e) { alert("failed");});
-    //req.addListener("timeout", function(e) { alert("timeout");});
-    //req.addListener("aborted", function(e) { alert("aborted");});
-    req.send();
-  }
-}
-DSPDataFetcher();
-
-
-
-/**
- * This is the main application class of your custom application "DSP Web Logger"
- */
-qx.Class.define("dsp_web_logger.Application",
-{
-  extend : qx.application.Standalone,
-
-
-
-  /*
-  *****************************************************************************
-     MEMBERS
-  *****************************************************************************
-  */
-
-  members :
-  {
-    /**
-     * This method contains the initial application code and gets called 
-     * during startup of the application
-     */
-    main : function()
-    {
-      // Call super class
-      this.base(arguments);
-
-      // Enable logging in debug variant
-      if (qx.core.Variant.isSet("qx.debug", "on"))
-      {
-        // support native logging capabilities, e.g. Firebug for Firefox
-        qx.log.appender.Native;
-        // support additional cross-browser console. Press F7 to toggle visibility
-        qx.log.appender.Console;
-      }
-
-      var doc = this.getRoot();
-
-      var label = new qx.ui.basic.Label("DSP Web Logger");
-      var font = new qx.bom.Font(18, ["Arial"]);
-      font.setBold(true);
-      label.setFont(font);
-      doc.add(label, {top: 30, left: 50});
-      
[…]

### 2009-02-25 versions/v2/apps/osgi-bundles/dsp/DSPWebDataLogger/web/DSP Web Logger/source/class/dsp_web_logger/test/DemoTest.js

@@ -1,55 +1,0 @@
-/* ************************************************************************
-
-   Copyright:
-
-   License:
-
-   Authors:
-
-************************************************************************ */
-
-/**
- * This class demonstrates how to define unit tests for your application.
- *
- * Execute <code>generate.py test</code> to generate a testrunner application 
- * and open it from <tt>test/index.html</tt>
- *
- * The methods that contain the tests are instance methods with a 
- * <code>test</code> prefix. You can create an arbitrary number of test 
- * classes like this one. They can be organized in a regular class hierarchy, 
- * i.e. using deeper namespaces and a corresponding file structure within the 
- * <tt>test</tt> folder.
- */
-qx.Class.define("dsp_web_logger.test.DemoTest",
-{
-  extend : qx.dev.unit.TestCase,
-
-  members :
-  {
-    /*
-    ---------------------------------------------------------------------------
-      TESTS
-    ---------------------------------------------------------------------------
-    */
-  
-    /**
-     * Here are some simple tests
-     */
-    testSimple : function()
-    {
-      this.assertEquals(4, 3+1, "This should never fail!");
-      this.assertFalse(false, "Can false be true?!");
-    },
-
-    /**
-     * Here are some more advanced tests
-     */
-    testAdvanced: function () 
-    {
-      var a = 3;
-      var b = a;
-      this.assertIdentical(a, b, "A rose by any other name is still a rose");
-      this.assertInRange(3, 1, 10, "You must be kidding, 3 can never be outside [1,10]!");
-    }
-  }
-});

### 2009-02-25 versions/v2/apps/xml/jaxb/message/generated-src/org/netbeams/dsp/message/AbstractMessage.java
- Change properties to Mac machine - Remove unneeded cache directory
git-svn-id: svn://localhost/netbeams/trunk@317 559edde2-a4d4-44e0-96fc-4035e40d7396

@@ -2,7 +2,7 @@
 // This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vhudson-jaxb-ri-2.1-661 
 // See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
 // Any modifications to this file will be lost upon recompilation of the source schema. 
-// Generated on: 2009.02.21 at 02:31:02 AM PST 
+// Generated on: 2009.02.25 at 01:49:31 AM PST 
 //
 
 
@@ -45,15 +45,15 @@ import javax.xml.bind.annotation.XmlType;
     "body"
 })
 @XmlSeeAlso({
-    DeleteMessage.class,
-    CreateMessage.class,
-    MeasureMessage.class,
-    QueryMessage.class,
-    EventMessage.class,
     AcknowledgementMessage.class,
     ActionMessage.class,
     UpdateMessage.class,
-    InsertMessage.class
+    EventMessage.class,
+    CreateMessage.class,
+    InsertMessage.class,
+    DeleteMessage.class,
+    MeasureMessage.class,
+    QueryMessage.class
 })
 public class AbstractMessage {
 

### 2009-02-25 versions/v2/apps/xml/jaxb/message/generated-src/org/netbeams/dsp/message/ObjectFactory.java

@@ -2,7 +2,7 @@
 // This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vhudson-jaxb-ri-2.1-661 
 // See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
 // Any modifications to this file will be lost upon recompilation of the source schema. 
-// Generated on: 2009.02.21 at 02:31:02 AM PST 
+// Generated on: 2009.02.25 at 01:49:31 AM PST 
 //
 
 
@@ -31,133 +31,21 @@ import javax.xml.namespace.QName;
 @XmlRegistry
 public class ObjectFactory {
 
+    private final static QName _InsertMessage_QNAME = new QName("http://www.netbeams.dsp.org/message.xsd", "InsertMessage");
     private final static QName _DeleteMessage_QNAME = new QName("http://www.netbeams.dsp.org/message.xsd", "DeleteMessage");
-    private final static QName _EventMessage_QNAME = new QName("http://www.netbeams.dsp.org/message.xsd", "EventMessage");
-    private final static QName _ActionMessage_QNAME = new QName("http://www.netbeams.dsp.org/message.xsd", "ActionMessage");
+    private final static QName _CreateMessage_QNAME = new QName("http://www.netbeams.dsp.org/message.xsd", "CreateMessage");
     private final static QName _UpdateMessage_QNAME = new QName("http://www.netbeams.dsp.org/message.xsd", "UpdateMessage");
     private final static QName _AcknowledgementMessage_QNAME = new QName("http://www.netbeams.dsp.org/message.xsd", "AcknowledgementMessage");
-    private final static QName _CreateMessage_QNAME = new QName("http://www.netbeams.dsp.org/message.xsd", "CreateMessage");
+    private final static QName _ActionMessage_QNAME = new QName("http://www.netbeams.dsp.org/message.xsd", "ActionMessage");
     private final static QName _QueryMessage_QNAME = new QName("http://www.netbeams.dsp.org/message.xsd", "QueryMessage");
     private final static QName _MeasureMessage_QNAME = new QName("http://www.netbeams.dsp.org/message.xsd", "MeasureMessage");
-    private final static QName _InsertMessage_QNAME = new QName("http://www.netbeams.dsp.org/message.xsd", "InsertMessage");
+    private final static QName _EventMessage_QNAME = new QName("http://www.netbeams.dsp.org/message.xsd", "EventMessage");
 
     /**
[…]

### 2009-02-25 versions/v2/apps/xml/jaxb/message/generated-src/org/netbeams/dsp/message/UpdateMessage.java

@@ -2,7 +2,7 @@
 // This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vhudson-jaxb-ri-2.1-661 
 // See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
 // Any modifications to this file will be lost upon recompilation of the source schema. 
-// Generated on: 2009.02.21 at 02:31:02 AM PST 
+// Generated on: 2009.02.25 at 01:49:31 AM PST 
 //
 
 

### 2009-02-25 versions/v2/apps/xml/jaxb/message/generated-src/org/netbeams/dsp/message/package-info.java

@@ -2,7 +2,7 @@
 // This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vhudson-jaxb-ri-2.1-661 
 // See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
 // Any modifications to this file will be lost upon recompilation of the source schema. 
-// Generated on: 2009.02.21 at 02:31:02 AM PST 
+// Generated on: 2009.02.25 at 01:49:31 AM PST 
 //
 
 @javax.xml.bind.annotation.XmlSchema(namespace = "http://www.netbeams.dsp.org/message.xsd")

### 2009-02-26 versions/v2/apps/osgi-bundles/dsp/DSPPlatform/src/org/netbeams/dsp/platform/broker/MessageBroker.java
Additional trances
git-svn-id: svn://localhost/netbeams/trunk@319 559edde2-a4d4-44e0-96fc-4035e40d7396

@@ -244,14 +244,17 @@ public class MessageBroker implements MessageBrokerAccessor {
     }
 
     private String messageSummary(Message message) {
-        StringBuffer buff = new StringBuffer();
+        StringBuilder buff = new StringBuilder();
         buff.append("id=").append(message.getMessageID()).append("; ");
-        buff.append("producer=").append(message.getHeader().getProducer().getComponentType()).append("; ");
-        buff.append("content type=").append(message.getContentType()).append("; ");
+        buff.append("prod=").append(message.getHeader().getProducer().getComponentType()).append("; ");
+        buff.append("content_type=").append(message.getContentType()).append("; ");
         ComponentIdentifier consumer = message.getHeader().getConsumer();
         if (consumer != null) {
             buff.append("consumer=");
-            buff.append(consumer.getComponentType()).append(",");
+            buff.append(consumer.getComponentType()).append("; ");
+            buff.append("node=");
+            buff.append(consumer.getComponentLocator().getNodeAddress().getValue()).append("; ");
+
         }
         return buff.toString();
     }

### 2009-02-26 versions/v2/apps/osgi-bundles/dsp/DSPStockProducer/src/org/netbeams/dsp/demo/stocks/producer/StockProducer.java
Add more sample code
git-svn-id: svn://localhost/netbeams/trunk@320 559edde2-a4d4-44e0-96fc-4035e40d7396

@@ -20,10 +20,14 @@ import org.netbeams.dsp.MessageBrokerAccessor;
 import org.netbeams.dsp.MessageCategory;
 import org.netbeams.dsp.MessageFactory;
 import org.netbeams.dsp.data.property.DSProperties;
+import org.netbeams.dsp.data.property.DSProperty;
 import org.netbeams.dsp.demo.stock.StockTick;
 import org.netbeams.dsp.demo.stock.StockTicks;
+import org.netbeams.dsp.message.AcknowledgementMessage;
+import org.netbeams.dsp.message.ComponentIdentifier;
 import org.netbeams.dsp.message.MeasureMessage;
 import org.netbeams.dsp.message.Message;
+import org.netbeams.dsp.message.QueryMessage;
 import org.netbeams.dsp.message.UpdateMessage;
 import org.w3c.dom.Node;
 
@@ -181,32 +185,128 @@ public class StockProducer implements DSPComponent{
 	private void processMessage(Message message) {
 		log.debug("message class=" + message.getClass().getName());
 		
-		if(message instanceof UpdateMessage){			
-			DSProperties stockTicks = null;
-			Object content = message.getBody().getAny();
-			if(Message.isPojo(content)){
-				return;
-				
-//				if(content instanceof StockTicks){
-//					stockTicks = (StockTicks)content;
-//				}else{
-//					Log.log("StockConsumer.processMessage(): Invalid content type " + content.getClass().getName());
-//				}
-			}else{
-				try{
-					JAXBContext jc = JAXBContext.newInstance("org.netbeams.dsp.data.property",
-							org.netbeams.dsp.demo.stock.ObjectFactory.class.getClassLoader());
-					Unmarshaller unmarshaller = jc.createUnmarshaller();
-					DSProperties dspProperty= (DSProperties)unmarshaller.unmarshal((Node)content);
-				}catch(JAXBException e){
-					log.error("Error unmarhalling the message", e);
-					return;
-				}
-
-			}			
+		if(message instanceof UpdateMessage){		
+			processUpdate(message);
+		}else if(message instanceof QueryMessage){	
+			processQuery(message);			
 		}
 	}
 	
+
+	private void processQuery(Message message) {
+		DSProperties dspProperties = null;
+		try{
+			Object content = message.getBody().getAny();
+			log.debug("Content class " + content.getClass().getName());
+			
+			JAXBContext jc = JAXBContext.newInstance("org.netbeams.dsp.data.property",
[…]

### 2009-02-26 versions/v2/apps/osgi-bundles/dsp/DSPManagement/src/org/netbeams/dsp/management/Activator.java
Query and update interface
git-svn-id: svn://localhost/netbeams/trunk@321 559edde2-a4d4-44e0-96fc-4035e40d7396

@@ -4,6 +4,7 @@ import javax.servlet.ServletException;
 import javax.servlet.Servlet;
 
 import org.apache.log4j.Logger;
+import org.netbeams.dsp.management.test.TestManagement;
 import org.netbeams.dsp.platform.PlatformException;
 import org.netbeams.dsp.platform.osgi.ActivatorHelper;
 import org.osgi.framework.BundleActivator;
@@ -21,6 +22,7 @@ public class Activator implements BundleActivator {
 
 	private DSPManager dspManager;
 	private ManagementServlet mgrServlet;
+	private TestManagement tester;
 
 	/* (non-Javadoc)
 	 * @see org.osgi.framework.BundleActivator#start(org.osgi.framework.BundleContext)
@@ -30,8 +32,10 @@ public class Activator implements BundleActivator {
 
 		this.bundleContext = bundleContext;
 		dspManager = new DSPManager();
-		registerServlet();
+//		registerServlet();
 		serviceRegistration = ActivatorHelper.registerOSGIService(bundleContext, dspManager);
+		tester = new TestManagement(dspManager);
+		tester.start();
 
 	}
 
@@ -41,7 +45,11 @@ public class Activator implements BundleActivator {
 	public void stop(BundleContext context) throws Exception {
 		log.info("stop invoked.");
 		ActivatorHelper.unregisterOSGIService(bundleContext, serviceRegistration);
-		unregisterServlet();
+//		unregisterServlet();
+		if(tester != null){
+			tester.shouldStop();
+			tester.interrupt();
+		}
 		dspManager.stopComponent();
 		dspManager = null;
 		mgrServlet = null;
@@ -56,7 +64,7 @@ public class Activator implements BundleActivator {
 					HttpService httpService = (HttpService)bundleContext.getService(srvRef);
 					// Create Servlet
 					mgrServlet = new ManagementServlet(dspManager);
-					httpService.registerServlet(ManagementServlet.BASE_URI, mgrServlet, null, null);
+					httpService.registerServlet(RESTDirectory.BASE_URI, mgrServlet, null, null);
 				} catch (Exception e) {
 					log.warn("could not register management servlet", e);
 					throw new PlatformException("could not register management servlet", e);
@@ -76,7 +84,7 @@ public class Activator implements BundleActivator {
 			if(srvRef != null){
 				try {
 					HttpService httpService = (HttpService)bundleContext.getService(srvRef);
-					httpService.unregister(ManagementServlet.BASE_URI);
[…]

### 2009-02-26 versions/v2/apps/osgi-bundles/dsp/DSPManagement/src/org/netbeams/dsp/management/DSPManager.java

@@ -4,16 +4,32 @@ package org.netbeams.dsp.management;
 
 
 import java.util.Collection;
+import java.util.HashMap;
 import java.util.Map;
+import java.util.UUID;
+
+import javax.xml.bind.JAXBContext;
+import javax.xml.bind.JAXBException;
+import javax.xml.bind.Marshaller;
+import javax.xml.bind.Unmarshaller;
+import javax.xml.parsers.DocumentBuilder;
+import javax.xml.parsers.DocumentBuilderFactory;
+import javax.xml.parsers.ParserConfigurationException;
 
 import org.apache.log4j.Logger;
 import org.netbeams.dsp.ComponentDescriptor;
 import org.netbeams.dsp.data.property.DSProperties;
+import org.netbeams.dsp.message.AbstractMessage;
 import org.netbeams.dsp.message.AcknowledgementMessage;
+import org.netbeams.dsp.message.Body;
 import org.netbeams.dsp.message.ComponentIdentifier;
 import org.netbeams.dsp.message.ComponentLocator;
 import org.netbeams.dsp.message.Header;
+import org.netbeams.dsp.message.MeasureMessage;
+import org.netbeams.dsp.message.MessageContent;
+import org.netbeams.dsp.message.MessagesContainer;
 import org.netbeams.dsp.message.NodeAddress;
+import org.netbeams.dsp.message.QueryMessage;
 import org.netbeams.dsp.message.UpdateMessage;
 import org.netbeams.dsp.DSPComponent;
 import org.netbeams.dsp.DSPContext;
@@ -22,9 +38,13 @@ import org.netbeams.dsp.MessageBrokerAccessor;
 import org.netbeams.dsp.MessageFactory;
 import org.netbeams.dsp.message.Message;
 import org.netbeams.dsp.util.DSPUtils;
+import org.w3c.dom.Document;
+import org.w3c.dom.Node;
+
+import com.sun.corba.se.pept.broker.Broker;
 
 
-public class DSPManager implements DSPComponent
+public class DSPManager implements Manager, DSPComponent
 {
 	private static final Logger log = Logger.getLogger(DSPManager.class);
 			
@@ -43,10 +63,13 @@ public class DSPManager implements DSPComponent
     
     private MessageBrokerAccessor accessor;
     
-    private Map<String, Message> pendingAck;
+    private Map<String, Boolean> pendingReply;
+    private Map<String, Message> repliedMsgs;
     
     public DSPManager(){
     	isActive = false;
[…]

### 2009-02-26 versions/v2/apps/osgi-bundles/dsp/DSPManagement/src/org/netbeams/dsp/management/ManagementServlet.java

@@ -15,6 +15,8 @@ import org.apache.log4j.Logger;
 import org.netbeams.dsp.DSPException;
 import org.netbeams.dsp.data.property.DSProperties;
 import org.netbeams.dsp.data.property.DSProperty;
+import org.netbeams.dsp.management.handler.DSPGetHandler;
+import org.netbeams.dsp.management.handler.GetHandler;
 import org.netbeams.dsp.message.NodeAddress;
 
 
@@ -24,8 +26,6 @@ public class ManagementServlet
 {
 	private static final Logger log = Logger.getLogger(ManagementServlet.class);
 	
-    final public static String BASE_URI = "/dsp-management";
-
     private boolean isActive;
     private DSPManager dspManager;
     
@@ -45,10 +45,17 @@ public class ManagementServlet
         throws ServletException, IOException
     {
     	if(!dspManager.isActive()){
-    		log.warn("Management component is active ");
+    		log.warn("Management component is not active ");
     		outputNoActive(response);
     		return;
     	}
+    	
+    	GetHandler handler = selectHandler(request);
+    	if(handler != null){
+    		response.setContentType("text/html; charset=utf-8");
+    		outputFile(response, "Directory Unknown");   		
+    	}
+    	handler.doGet(request, response);
     	
     	Map params = request.getParameterMap();
     	String action = (String)params.get("ACTION");
@@ -73,7 +80,11 @@ public class ManagementServlet
 
 
 
-    private void outputNoActive(HttpServletResponse response) throws IOException {
+    private GetHandler selectHandler(HttpServletRequest request) {
+		return new DSPGetHandler();
+	}
+
+	private void outputNoActive(HttpServletResponse response) throws IOException {
     	response.sendRedirect("notActive.html");
 		
 	}
@@ -93,7 +104,7 @@ public class ManagementServlet
     	NodeAddress nodeAddress = new NodeAddress(nodeAddressStr);
     	
     	try {
-			dspManager.updateProperties(nodeAddress, componentNodeId, properties);
+			dspManager.query(componentNodeId, nodeAddressStr, properties);
 		} catch (DSPException e) {
 			log.warn("Could not update property for " + componentNodeId + ' ' + nodeAddressStr);
 		}   	

### 2009-02-26 versions/v2/apps/osgi-bundles/dsp/DSPFramework/src/org/netbeams/dsp/DSPException.java
MessageFactory enhancements
git-svn-id: svn://localhost/netbeams/trunk@322 559edde2-a4d4-44e0-96fc-4035e40d7396

@@ -8,6 +8,10 @@ public class DSPException extends Exception {
 	
 	private int error;
 	
+	public DSPException(Throwable t){
+		super(t);
+	}
+
 	public DSPException(String msg){
 		this(ErrorCode.ERROR_UNKNOWN, msg);
 	}

### 2009-02-26 versions/v2/apps/osgi-bundles/dsp/DSPFramework/src/org/netbeams/dsp/MessageFactory.java

@@ -4,8 +4,19 @@ package org.netbeams.dsp;
 import java.util.List;
 import java.util.UUID;
 
+import javax.xml.bind.JAXBContext;
+import javax.xml.bind.JAXBException;
+import javax.xml.bind.Marshaller;
+import javax.xml.parsers.DocumentBuilder;
+import javax.xml.parsers.DocumentBuilderFactory;
+import javax.xml.parsers.ParserConfigurationException;
+
+import org.apache.log4j.Logger;
 import org.netbeams.dsp.message.*;
 import org.netbeams.dsp.util.ErrorCode;
+import org.netbeams.dsp.util.NetworkUtil;
+import org.w3c.dom.Document;
+import org.w3c.dom.Node;
 
 
 /**
@@ -17,12 +28,22 @@ import org.netbeams.dsp.util.ErrorCode;
  */
 public class MessageFactory {
 	
+	private static final Logger log = Logger.getLogger(MessageFactory.class);
+	
 	private MessageFactory() {}
 	
 	public static Message newMessage(Class<? extends Message> messageType, MessageContent messageContent, BaseComponent producer)
 		throws DSPException
 	{
+		return newMessage(messageType, messageContent, producer, null, null);
+	}
 	
+	
+	public static Message newMessage(Class<? extends Message> messageType, MessageContent messageContent, BaseComponent producer,
+			String consumerNodeId, String consumerNodeAddress)
+		throws DSPException
+	{
+		
 		ComponentLocator locator = new ComponentLocator();
 		locator.setComponentNodeId(producer.getComponentNodeId());
 		locator.setNodeAddress(NodeAddressHelper.LOCAL_NODEADDRESS);
@@ -35,6 +56,60 @@ public class MessageFactory {
 		return message;
 	}
 
+	public static Message newMessage2(Class<? extends Message> messageType, MessageContent messageContent, BaseComponent producer,
+			String consumerNodeId, String componentType, String consumerNodeAddress)
+		throws DSPException
+	{
+		ComponentIdentifier ciProducer = 
+			createComponentIdentifier(producer.getComponentNodeId(), producer.getComponentType(), NetworkUtil.getCurrentEnvironmentNetworkIp());
+						
+		Message  message = createMessageInternal(messageType, ciProducer);
+		
+		ComponentIdentifier ciConsumer = 
+			createComponentIdentifier(consumerNodeId, componentType, consumerNodeAddress);		
+		setConsumer(message, ciConsumer);
+		
[…]

### 2009-02-26 versions/v2/apps/osgi-bundles/dsp/DSPFramework/src/org/netbeams/dsp/NodeAddressHelper.java

@@ -3,6 +3,8 @@ package org.netbeams.dsp;
 import org.netbeams.dsp.message.NodeAddress;
 
 public class NodeAddressHelper {
+	
+	public static String localAddressStr;
 
 	public static final NodeAddress LOCAL_NODEADDRESS = new NodeAddress();
 	static {
@@ -12,7 +14,5 @@ public class NodeAddressHelper {
 	public static final NodeAddress NO_ADDRESS = new NodeAddress();
 	static {
 		NO_ADDRESS.setValue("no-address");
-	}
-	
-	
+	}	
 }

### 2009-03-02 docs/icccn2009/paper.tex
NetBEAMS architecture
git-svn-id: svn://localhost/netbeams/trunk@318 559edde2-a4d4-44e0-96fc-4035e40d7396
Intro, Conclusion, and updated NetBEAMS diagram
git-svn-id: svn://localhost/netbeams/trunk@327 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 
-The introduction should emphasize that we interpret sensor networks in
-a different way: our focus is on off-the-shelf, highly specialized
-sensors for environmental monitoring. Our focus is end-to-end: how to
-collect data from a remote location with no stationary power or
-Internet connection and deliver the data in near-realtime to a web
-browser. Length: first page (including title and abstract).
+Environmental monitoring depends on the reliable collection of
+measurements from remotely deployed sensors and the rapid transfer of
+those measurements to data centers where processing and analysis may
+take place. Typically, sensors are deployed in out-of-the-way
+locations where physical access is limited and connections for
+telemetry are poor or nonexistent.   Maintaining the data stream from
+these sensors places demands on human resources, requiring site visits
+to service the sensors and to download data stored on internal memory.
+Consequently, there is a lag of days to months between the time the
+sensors perform the measurements and the time the measurements becomes
+available.   Advances in battery life and anti-fouling technology have
+extended service cycles for field sensors, with the unwanted result of
+further delaying access to monitoring data.  In addition, the
+dependence on field site visits to alter of sensor characteristics,
+such as sampling rate, prevents adjustment in monitoring strategy in
+response to rapidly developing events (e.g. oil spills, harmful algal
+blooms).
+
+To address these problems of data collection from remote
+sensors, we propose to develop an automated end-to-end system based on
+proven technology that provides a means to program and interrogate
+sensors at the remote field locations and to transmit collected
+measurements over long distances to data collections centers where the
+data is rapidly processed, archived, and made available to potential
+users in near real-time.
 
@@
 
+\begin{figure*}
+\centering
+\epsfig{file=netbeams, width=10cm}
+\caption{\label{FIG_NETBEAMS} NetBEAMS architecture.}
+\end{figure*}
+
 \subsection{YSI Sonde DP}
@@
 
+The California coastal region is populated by a vast number of
+remotely deployed sensors that monitor environmental conditions
+ranging from weather to water quality to ocean surface currents.
+These sensors are operated by data providers, including resource
+management groups, research scientists, municipalities, and state and
+federal agencies, who could use near real-time data streams to inform
+decision that impact public health, public safety, and the protection
+of resources.  In addition, there are efforts underway at the state
+and federal level to collate the data collected by the individual data
+providers into integrated monitoring networks that will enhance the
+distribution of monitoring data to public, private, and federal users
+groups.  Success of these efforts, whether small scale or a large
+integrated network, depends on the first step in the data management
+process: reliably obtaining data in near real-time from remotely
+located sensors.  This step can be the weak link in the data
+management pathway when telemetry options are unreliable.  The
+availability of a plug-and-play computing platform designed to
+communicate with a variety of sensor types and to provide data
+telemetry to distant brick and mortar data centers would stabilize
+data streams with poor telemetry and allowing expansion of sensor
+coverage into regions not presently served.  The low power consumption
+and low cost of these units would be particularly beneficial for
+applications with long service cycles and for data providers trying to
+maintain data streams on a limited budget.
+
 %\bibliography{../literature/lit}

### 2009-03-03 docs/icccn2009/paper.tex
Paste more text
git-svn-id: svn://localhost/netbeams/trunk@328 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 
-\section{Use Case}
+\section{Requirement Analysis}
 
@@
 
+Micro-kernel architecture.
+
 \subsection{Data Sensor Framework and Platform}
@@
 
-Should explain some specific DC and DP we have implemented for
-NetBEAMS. Length: 1 - 1.5 pages.
+In the following we present a practical application the the DSP
+presented in the previous section. We describe a deployment of the DSP
+using sensor owned and operated by the RTC. Our environmental sensor
+network is dubbed NetBEAMS (Networked Bay Environmental Assessment
+Monitoring System) and allows access to sensor equipment deployed
+throughout the San Francisco Bay Area. Figure \ref{FIG_NETBEAMS}
+provides an end-to-end overview of the NetBEAMS architecture. NetBEAMS
+is built using multiple installtion of the DSP. One DSP installation
+is collocated with each sensor and an additional DSP is running on the
+backend server.
 
@@
 \end{figure*}
+
+To describe NetBEAMS is to describe the specific DC and DP we have
+implemented. On the far left of Figure \ref{FIG_NETBEAMS} is the
+actual sensor; in our case the aforementioned Seabird sensor.
+Collocated with the Seabird is an embedded system. Here we make use of
+a platform called Gumstix.
+
+The Gumstix is a single-board computers with 256 Mb of flash and 256
+Mb of SDRAM, and running a 600 MHz TI OMAP 3503 processor. Additional
+features can be added to the motherboard with expansion cards
+connected via on-board buses. The motherboards draw less than 250 mA
+@4V at 400 MHz and less than 50 mA while idling, waiting for input.
+Gumstix run Linux 2.6 with the BusyBox utilities, and use the
+OpenEmbedded build environment to provide a complete Linux
+environment and a large range of Linux applications. We use the JamVM
+implementation of Sun Microsystem's virtual machine to execute Java
+programs.
+
+Cellular communication platform.
 

### 2009-03-04 docs/icccn2009/paper.tex
Add more material
git-svn-id: svn://localhost/netbeams/trunk@330 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 
-\author{\authorblockN{XXX}
+\author{\authorblockN{A. Puder, T. Johnson, K. Sales, M. Sales}
 \authorblockA{San Francisco State University \\
@@
 San Francisco, CA 94132 \\
-EMail: XXX@sfsu.edu}}
+EMail: \{arno$\mid$tlj$\mid$klebers$\mid$msales\}@sfsu.edu}
+\and
+\authorblockN{D. Robinson}
+\authorblockA{San Francisco State University \\
+Romberg Tiburon Center \\
+3150 Paradise Drive \\
+Tiburon, CA 94920 \\
+EMail: dhr@sfsu.edu}}
 
@@
 \section{Introduction}
+
+\cite{roemer:2004}
 
@@
 
-\section{Requirement Analysis}
+\section{Background}
 
@@
 reduces the maintenance overhead.
+
+---
+
+The YSI 6600EDS V2 Sonde is a water quality monitoring device that
+gathers water quality data, and operates in fresh, sea, or
+polluted water. It is under 22 inches (under 55 cm) in length, 3.5
+inches (8.9 cm) in diameter, and weighs approximately 7 pounds (3.18
+kg). It operates at temperatures between -5 to +50 deg C and at depths
+up to 656 feet (200 meters).  It uses 8 C-size Alkaline Batteries or
+External 12 VDC. Battery life can last up to 75 days depending on
+sensor configuration. The Sonde's bulkhead contains multiple pin
+connectors to support sensor probes for measuring the following:
+
+\begin{table}[h]
+\caption{\label{TAB_SONDE_MEASUREMENTS} YSI Sonde measurements.}
+\centering
+\begin{tabular}{|l||l|}
+\hline
+\multicolumn{1}{|c||}{\textbf{Name}} &
+\multicolumn{1}{c|}{\textbf{Description}} \\ \hline \hline
+\texttt{Date}    & Year/Month/Day. \\ \hline
+\texttt{Time}    & Hour:Minute:Second. \\ \hline
+\texttt{Temp}    & Temperature in degrees Celcius. \\ \hline
+\texttt{SpCond}  & Specific Description in microSiemens
+                   per centimeter. \\ \hline
+\texttt{Cond}    & Conductivity in microSiemens per centimeter. \\ \hline
+\texttt{Resist}  & Resistivity in Ohms $*$ centimeter. \\ \hline
+\texttt{Sal}     & Salinity in parts per thousand. \\ \hline
+\texttt{Press}   & Pressure in pounds per square inch relative. \\ \hline
+\texttt{Depth}   & Water column in meters. \\ \hline
+\texttt{pH}      & pH in standard units. \\ \hline
+\texttt{phmV}    & millivolts associated with the pH reading. \\ \hline
+\texttt{ODOSat}  & Dissolved oxygen in \% air saturation. \\ \hline
+\texttt{ODOConc} & Dissolved oxygen in mg/L. \\ \hline
+\texttt{Turbid}  & Turbidity in nephelometric turbidity units. \\ \hline
+\texttt{Battery} & Total Volts remaining in batteries.  \\ \hline
+\end{tabular}
+\end{table}
+
+Dissolved oxygen
+Conductivity/Temperature
+Fouling Resistant pH/ORP
+Chloride
+Ammonium
+Nitrate
+Turbidity
+Chlorophyll
+Rhodamine
+PC-Blue-green Algae
+PE-Blue-gren Algae
+
+Each probe may have multiple sensors.  A computer can interface with
+the Sonde via RS-232C or SDI-12. The Sonde can be configured to
+collect data samples in discrete or unnattended mode.  Discrete mode
+is normally performed while a technician manages the Sonde during the
+sampling process. In this mode the Sonde is usually connected via 650
+MDS Display/Logger or a serial cable connected to a PC.  The sampling
+frequency is usually set to a high frequency in this mode.  Unattended
+mode is performed usually when the Sonde is deployed offshore or in a
+remote body of water.  The sampling frequency is usually set for a
+longer period of time (i.e. 5 - 15 min). The Sonde is configured via a
+set of menus that can be accessed through a terminal session over
+RS232. Through these menus, the Sonde can be configured to start
+logging data to an internal file.  To view the data as it's being
+captured, the option to show live data is chosen and the data can then
+seen via a serial terminal session at each interval.
+
+---
 
@@
 
+\begin{code}
+<xs:schema ...>
+  <xs:complexType name="DSPMessage">
+    <xs:element name="Header" type="Header"/>
+  </xs:complexType>
+
+  <xs:complexType name="Header">
+    <xs:sequence>
+      <xs:attribute name="messageID"/>
+      <xs:attribute name="originatingDP"/>
+      <!-- ... -->
+    </xs:sequence>
+  </xs:complexType>
+</xs:schema>
+\end{code}
+
 Should explain the way DSP Messages are defined: XML Schema and JAXB.
@@
 \centering
-\epsfig{file=netbeams, width=10cm}
+\epsfig{file=netbeams, width=14cm}
 \caption{\label{FIG_NETBEAMS} NetBEAMS architecture.}
@@
 
-Cellular communication platform.
+The Huawei E220 is a USB modem that supports HSDPA, UMTS, EDGE packet
+data services at maximum transmission rates of 3.6Mbps, 384kbps, and
+236.8kbps respectively. HSDPA and UMTS operate at 2100MHz while GSM,
+GPRS, and EDGE operate at 900, 1800, and 1900MHz. The modem connects
+to the Verdex Pro XM4 motherboard via a mini USB interface (supporting
+USB 2.0). Linux kernels since 2.6.20 include support for E220 drivers.
+This includes the linux kernel on the gumstix which runs 2.6.21.  The
+gumstix communicates with the modem using the PPP protocol and
+accesses the cellular network using a data plan from AT\&T.
 
 \subsection{YSI Sonde DP}
+
+\begin{code}
+<xs:schema ...>
+<xs:import schemaLocation="dsp-message.xsd"/>
+
+<xs:element name="SondeDataMessage">
+  <xs:complexType>
+    <xs:complexContent>
+      <xs:extension base="DSPMessage">
+        <xs:sequence maxOccurs="unbounded"
+                     minOccurs="1">
+          <xs:element name="sondeData"
+                      type="sondeDataType"/>
+        </xs:sequence>
+      </xs:extension>
+    </xs:complexContent>
+  </xs:complexType>
+</xs:element>
+
+<xs:complexType name="sondeDataType">
+  <xs:sequence>
+    <xs:element name="Date" type="xs:date"/>
+    <xs:element name="Time" type="xs:time"/>
+    <xs:element name="Temp" type="xs:float"/>
+    <xs:element name="SpCond" type="xs:float"/>
+    <!-- ... -->
+  </xs:sequence>
+</xs:complexType>
+</xs:schema>
+\end{code}
 
@@
 
-%\bibliography{../literature/lit}
-%\bibliographystyle{plain}
+\bibliography{lit}
+\bibliographystyle{plain}
 

### 2009-03-04 docs/icccn2009/paper.tex
Some text on DSP
git-svn-id: svn://localhost/netbeams/trunk@331 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 
-\section{DSP}
+\section{Data Sensor Platform}
 
@@
 
-\subsection{Data Sensor Framework and Platform}
+\subsection{Architecture}
 
-should include a picture of a standalone DSP showing a DC and a DP
-(general description only). Explain difference between DSP Framework
-and DSP Platform. This section should give a top-level overview.
-Subsequent sub-sections should explain certain features of the DSP in
-more detail.
+Our architecture is based on a micro-kernel approach. The core
+functionality is limited to a minimum and all extra services are built
+as plug-and-play modules on top of the micro-kernel. With the help of
+this approach we can keep the basic infrastructure compact while
+allowing customization via special-purpose modules. In our terminology
+we call the basic foundation the \emph{Data Sensor Platform}
+(DSP). The scope of the DSP is one address space of an execution
+platform. A sensor network is built from several DSPs that are linked
+with each other.
+
+Figure \ref{FIG_DSP} depicts the basic architecture of the
+DSP. We refer to the plug-and-play modules as DSP Components. DSP
+Components are self-contained modules that can be added and removed to
+a DSP installation at runtime. The basic communication paradigm within
+a DSP is a message. DSP Components can exchange messages of different
+types. We distinguish between Data Producers (DP) and Data Consumers
+(DP) within the scope of a DSP. DP and DC can be seen as roles that a
+DSP Component may have. It is possible for a DSP Component to act in
+the role of DC and DP at the same time.
 
@@
 
-\subsection{DP and DC}
+The DSP is responsible for routing messages between DPs and DCs. It is
+important to note that this routing only happens withing a
+DSP. Following the paradigm of a micro-kernel approach, the DSP has no
+notion of remote communication or sensor-specific details. All this
+knowledge is embodied in DSP Components.
+
+\subsection{Components}
 
@@
 
-\subsection{DSP Message}
+\begin{code}
+interface DSPComponent
+{
+   public String getComponentType();
+   public void initComponent(...);
+   public void startComponent();
+   public void stopComponent();
+   public void deliver(DSPMessage msg);
+   public Message deliverWithReply(DSPMessage msg);
+   // ...
+}
+\end{code}
+
+\subsection{Message}
 

### 2009-03-04 docs/icccn2009/paper.tex
Some text on DSP Components
git-svn-id: svn://localhost/netbeams/trunk@332 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 
-\subsection{Components}
+\subsection{DSP Components}
+
+A DSP Component encapsulates a particular functionality. We
+distinguish between those components that produce data (DP) and those
+that consume data (DC). An example for a DSP Component acting in the
+role of a DP is a sensor module. The sensor module reads data from the
+physical sensor and converts the data into a message that is forwarded
+to the DSP. Because this module generates data with respect to the
+DSP, it is a data producer. An example for a data consumer is a
+converter module that converts the internal, DSP-specific data format,
+suitable for consumption by external applications. One such common
+format used in the domain of environmental monitoring is OpenDAP
+(REF). Since the converter module accepts messages from the DSP, it
+acts in the role of a data consumer. Finally, a DSP Component
+implementing persistency is an example of a module that acts in the
+roles of both data producer and consumer. It accepts messages from the
+DSP and stores them in a database. At a later time, it may act as a
+data producer by resending previously stored messages.
+
+In order to enrich the DSP, a programmer has to implement new DSP
+Components.
 

### 2009-03-05 versions/v2/apps/xml/jaxb/message/generated-src/org/netbeams/dsp/message/AbstractMessage.java
Deleting generated artifacts to be added to the svn ignore list in the next revision.
This is the common problem during the development/SVN updates. All the generated
java code or Jars conflict with the ones under revision control. By convention,
All generated files should be put under SVN Ignore list.

Next release will SVN ignore all the removed files/directories. The Eclipse
project will continue to require the jar files to compile locally as usual, so
run the general build to obtain the most recent jar.

git-svn-id: svn://localhost/netbeams/trunk@335 559edde2-a4d4-44e0-96fc-4035e40d7396

@@ -1,167 +1,0 @@
-//
-// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vhudson-jaxb-ri-2.1-661 
-// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
-// Any modifications to this file will be lost upon recompilation of the source schema. 
-// Generated on: 2009.02.25 at 01:49:31 AM PST 
-//
-
-
-package org.netbeams.dsp.message;
-
-import javax.xml.bind.annotation.XmlAccessType;
-import javax.xml.bind.annotation.XmlAccessorType;
-import javax.xml.bind.annotation.XmlAttribute;
-import javax.xml.bind.annotation.XmlElement;
-import javax.xml.bind.annotation.XmlSchemaType;
-import javax.xml.bind.annotation.XmlSeeAlso;
-import javax.xml.bind.annotation.XmlType;
-
-
-/**
- * <p>Java class for AbstractMessage complex type.
- * 
- * <p>The following schema fragment specifies the expected content contained within this class.
- * 
- * <pre>
- * &lt;complexType name="AbstractMessage">
- *   &lt;complexContent>
- *     &lt;restriction base="{http://www.w3.org/2001/XMLSchema}anyType">
- *       &lt;sequence>
- *         &lt;element name="Header" type="{http://www.netbeams.dsp.org/message.xsd}Header"/>
- *         &lt;element name="Body" type="{http://www.netbeams.dsp.org/message.xsd}Body"/>
- *       &lt;/sequence>
- *       &lt;attribute name="messageID" type="{http://www.w3.org/2001/XMLSchema}anySimpleType" />
- *       &lt;attribute name="ContentType" type="{http://www.w3.org/2001/XMLSchema}anySimpleType" />
- *     &lt;/restriction>
- *   &lt;/complexContent>
- * &lt;/complexType>
- * </pre>
- * 
- * 
- */
-@XmlAccessorType(XmlAccessType.FIELD)
-@XmlType(name = "AbstractMessage", propOrder = {
-    "header",
-    "body"
-})
-@XmlSeeAlso({
-    AcknowledgementMessage.class,
-    ActionMessage.class,
-    UpdateMessage.class,
-    EventMessage.class,
-    CreateMessage.class,
-    InsertMessage.class,
-    DeleteMessage.class,
-    MeasureMessage.class,
-    QueryMessage.class
-})
-public class AbstractMessage {
-
[…]

### 2009-03-05 versions/v2/apps/xml/jaxb/message/generated-src/org/netbeams/dsp/message/AbstractMessageContent.java

@@ -1,37 +1,0 @@
-//
-// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vhudson-jaxb-ri-2.1-661 
-// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
-// Any modifications to this file will be lost upon recompilation of the source schema. 
-// Generated on: 2008.11.30 at 11:56:14 AM PST 
-//
-
-
-package org.netbeams.dsp.message;
-
-import javax.xml.bind.annotation.XmlAccessType;
-import javax.xml.bind.annotation.XmlAccessorType;
-import javax.xml.bind.annotation.XmlType;
-
-
-/**
- * <p>Java class for AbstractMessageContent complex type.
- * 
- * <p>The following schema fragment specifies the expected content contained within this class.
- * 
- * <pre>
- * &lt;complexType name="AbstractMessageContent">
- *   &lt;complexContent>
- *     &lt;restriction base="{http://www.w3.org/2001/XMLSchema}anyType">
- *     &lt;/restriction>
- *   &lt;/complexContent>
- * &lt;/complexType>
- * </pre>
- * 
- * 
- */
-@XmlAccessorType(XmlAccessType.FIELD)
-@XmlType(name = "AbstractMessageContent")
-public class AbstractMessageContent {
-
-
-}

### 2009-03-05 versions/v2/apps/xml/jaxb/message/generated-src/org/netbeams/dsp/message/AbstractNodeAddress.java

@@ -1,66 +1,0 @@
-//
-// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vhudson-jaxb-ri-2.1-661 
-// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
-// Any modifications to this file will be lost upon recompilation of the source schema. 
-// Generated on: 2009.02.25 at 01:49:31 AM PST 
-//
-
-
-package org.netbeams.dsp.message;
-
-import javax.xml.bind.annotation.XmlAccessType;
-import javax.xml.bind.annotation.XmlAccessorType;
-import javax.xml.bind.annotation.XmlType;
-import javax.xml.bind.annotation.XmlValue;
-
-
-/**
- * <p>Java class for AbstractNodeAddress complex type.
- * 
- * <p>The following schema fragment specifies the expected content contained within this class.
- * 
- * <pre>
- * &lt;complexType name="AbstractNodeAddress">
- *   &lt;simpleContent>
- *     &lt;extension base="&lt;http://www.w3.org/2001/XMLSchema>string">
- *     &lt;/extension>
- *   &lt;/simpleContent>
- * &lt;/complexType>
- * </pre>
- * 
- * 
- */
-@XmlAccessorType(XmlAccessType.FIELD)
-@XmlType(name = "AbstractNodeAddress", propOrder = {
-    "value"
-})
-public class AbstractNodeAddress {
-
-    @XmlValue
-    protected String value;
-
-    /**
-     * Gets the value of the value property.
-     * 
-     * @return
-     *     possible object is
-     *     {@link String }
-     *     
-     */
-    public String getValue() {
-        return value;
-    }
-
-    /**
-     * Sets the value of the value property.
-     * 
-     * @param value
-     *     allowed object is
-     *     {@link String }
-     *     
-     */
-    public void setValue(String value) {
-        this.value = value;
-    }
-
-}

### 2009-03-05 versions/v2/apps/xml/jaxb/message/generated-src/org/netbeams/dsp/message/AcknowledgementMessage.java

@@ -1,39 +1,0 @@
-//
-// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vhudson-jaxb-ri-2.1-661 
-// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
-// Any modifications to this file will be lost upon recompilation of the source schema. 
-// Generated on: 2009.02.25 at 01:49:31 AM PST 
-//
-
-
-package org.netbeams.dsp.message;
-
-import javax.xml.bind.annotation.XmlAccessType;
-import javax.xml.bind.annotation.XmlAccessorType;
-import javax.xml.bind.annotation.XmlType;
-
-
-/**
- * <p>Java class for AcknowledgementMessage complex type.
- * 
- * <p>The following schema fragment specifies the expected content contained within this class.
- * 
- * <pre>
- * &lt;complexType name="AcknowledgementMessage">
- *   &lt;complexContent>
- *     &lt;extension base="{http://www.netbeams.dsp.org/message.xsd}AbstractMessage">
- *     &lt;/extension>
- *   &lt;/complexContent>
- * &lt;/complexType>
- * </pre>
- * 
- * 
- */
-@XmlAccessorType(XmlAccessType.FIELD)
-@XmlType(name = "AcknowledgementMessage")
-public class AcknowledgementMessage
-    extends Message
-{
-
-
-}

### 2009-03-05 versions/v2/apps/xml/jaxb/message/generated-src/org/netbeams/dsp/message/ActionMessage.java

@@ -1,39 +1,0 @@
-//
-// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vhudson-jaxb-ri-2.1-661 
-// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
-// Any modifications to this file will be lost upon recompilation of the source schema. 
-// Generated on: 2009.02.25 at 01:49:31 AM PST 
-//
-
-
-package org.netbeams.dsp.message;
-
-import javax.xml.bind.annotation.XmlAccessType;
-import javax.xml.bind.annotation.XmlAccessorType;
-import javax.xml.bind.annotation.XmlType;
-
-
-/**
- * <p>Java class for ActionMessage complex type.
- * 
- * <p>The following schema fragment specifies the expected content contained within this class.
- * 
- * <pre>
- * &lt;complexType name="ActionMessage">
- *   &lt;complexContent>
- *     &lt;extension base="{http://www.netbeams.dsp.org/message.xsd}AbstractMessage">
- *     &lt;/extension>
- *   &lt;/complexContent>
- * &lt;/complexType>
- * </pre>
- * 
- * 
- */
-@XmlAccessorType(XmlAccessType.FIELD)
-@XmlType(name = "ActionMessage")
-public class ActionMessage
-    extends Message
-{
-
-
-}

### 2009-03-05 versions/v2/apps/xml/jaxb/message/generated-src/org/netbeams/dsp/message/Body.java

@@ -1,72 +1,0 @@
-//
-// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vhudson-jaxb-ri-2.1-661 
-// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
-// Any modifications to this file will be lost upon recompilation of the source schema. 
-// Generated on: 2009.02.25 at 01:49:31 AM PST 
-//
-
-
-package org.netbeams.dsp.message;
-
-import javax.xml.bind.annotation.XmlAccessType;
-import javax.xml.bind.annotation.XmlAccessorType;
-import javax.xml.bind.annotation.XmlAnyElement;
-import javax.xml.bind.annotation.XmlType;
-import org.w3c.dom.Element;
-
-
-/**
- * <p>Java class for Body complex type.
- * 
- * <p>The following schema fragment specifies the expected content contained within this class.
- * 
- * <pre>
- * &lt;complexType name="Body">
- *   &lt;complexContent>
- *     &lt;restriction base="{http://www.w3.org/2001/XMLSchema}anyType">
- *       &lt;sequence>
- *         &lt;any/>
- *       &lt;/sequence>
- *     &lt;/restriction>
- *   &lt;/complexContent>
- * &lt;/complexType>
- * </pre>
- * 
- * 
- */
-@XmlAccessorType(XmlAccessType.FIELD)
-@XmlType(name = "Body", propOrder = {
-    "any"
-})
-public class Body {
-
-    @XmlAnyElement(lax = true)
-    protected Object any;
-
-    /**
-     * Gets the value of the any property.
-     * 
-     * @return
-     *     possible object is
-     *     {@link Object }
-     *     {@link Element }
-     *     
-     */
-    public Object getAny() {
-        return any;
-    }
-
-    /**
-     * Sets the value of the any property.
-     * 
-     * @param value
-     *     allowed object is
-     *     {@link Object }
-     *     {@link Element }
-     *     
-     */
-    public void setAny(Object value) {
-        this.any = value;
-    }
-
-}

### 2009-03-05 versions/v2/apps/xml/jaxb/message/generated-src/org/netbeams/dsp/message/ComponentIdentifier.java

@@ -1,97 +1,0 @@
-//
-// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vhudson-jaxb-ri-2.1-661 
-// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
-// Any modifications to this file will be lost upon recompilation of the source schema. 
-// Generated on: 2009.02.25 at 01:49:31 AM PST 
-//
-
-
-package org.netbeams.dsp.message;
-
-import javax.xml.bind.annotation.XmlAccessType;
-import javax.xml.bind.annotation.XmlAccessorType;
-import javax.xml.bind.annotation.XmlElement;
-import javax.xml.bind.annotation.XmlType;
-
-
-/**
- * <p>Java class for ComponentIdentifier complex type.
- * 
- * <p>The following schema fragment specifies the expected content contained within this class.
- * 
- * <pre>
- * &lt;complexType name="ComponentIdentifier">
- *   &lt;complexContent>
- *     &lt;restriction base="{http://www.w3.org/2001/XMLSchema}anyType">
- *       &lt;sequence>
- *         &lt;element name="ComponentType" type="{http://www.w3.org/2001/XMLSchema}string"/>
- *         &lt;element name="ComponentLocator" type="{http://www.netbeams.dsp.org/message.xsd}ComponentLocator"/>
- *       &lt;/sequence>
- *     &lt;/restriction>
- *   &lt;/complexContent>
- * &lt;/complexType>
- * </pre>
- * 
- * 
- */
-@XmlAccessorType(XmlAccessType.FIELD)
-@XmlType(name = "ComponentIdentifier", propOrder = {
-    "componentType",
-    "componentLocator"
-})
-public class ComponentIdentifier {
-
-    @XmlElement(name = "ComponentType", required = true)
-    protected String componentType;
-    @XmlElement(name = "ComponentLocator", required = true)
-    protected ComponentLocator componentLocator;
-
-    /**
-     * Gets the value of the componentType property.
-     * 
-     * @return
-     *     possible object is
-     *     {@link String }
-     *     
-     */
-    public String getComponentType() {
-        return componentType;
-    }
-
-    /**
-     * Sets the value of the componentType property.
-     * 
-     * @param value
-     *     allowed object is
-     *     {@link String }
-     *     
-     */
-    public void setComponentType(String value) {
-        this.componentType = value;
-    }
-
-    /**
-     * Gets the value of the componentLocator property.
-     * 
[…]

### 2009-03-05 versions/v2/apps/xml/jaxb/message/generated-src/org/netbeams/dsp/message/ComponentLocator.java

@@ -1,97 +1,0 @@
-//
-// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vhudson-jaxb-ri-2.1-661 
-// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
-// Any modifications to this file will be lost upon recompilation of the source schema. 
-// Generated on: 2009.02.25 at 01:49:31 AM PST 
-//
-
-
-package org.netbeams.dsp.message;
-
-import javax.xml.bind.annotation.XmlAccessType;
-import javax.xml.bind.annotation.XmlAccessorType;
-import javax.xml.bind.annotation.XmlElement;
-import javax.xml.bind.annotation.XmlType;
-
-
-/**
- * <p>Java class for ComponentLocator complex type.
- * 
- * <p>The following schema fragment specifies the expected content contained within this class.
- * 
- * <pre>
- * &lt;complexType name="ComponentLocator">
- *   &lt;complexContent>
- *     &lt;restriction base="{http://www.w3.org/2001/XMLSchema}anyType">
- *       &lt;sequence>
- *         &lt;element name="ComponentNodeId" type="{http://www.w3.org/2001/XMLSchema}string"/>
- *         &lt;element name="NodeAddress" type="{http://www.netbeams.dsp.org/message.xsd}AbstractNodeAddress"/>
- *       &lt;/sequence>
- *     &lt;/restriction>
- *   &lt;/complexContent>
- * &lt;/complexType>
- * </pre>
- * 
- * 
- */
-@XmlAccessorType(XmlAccessType.FIELD)
-@XmlType(name = "ComponentLocator", propOrder = {
-    "componentNodeId",
-    "nodeAddress"
-})
-public class ComponentLocator {
-
-    @XmlElement(name = "ComponentNodeId", required = true)
-    protected String componentNodeId;
-    @XmlElement(name = "NodeAddress", required = true, type = NodeAddress.class)
-    protected NodeAddress nodeAddress;
-
-    /**
-     * Gets the value of the componentNodeId property.
-     * 
-     * @return
-     *     possible object is
-     *     {@link String }
-     *     
-     */
-    public String getComponentNodeId() {
-        return componentNodeId;
-    }
-
-    /**
-     * Sets the value of the componentNodeId property.
-     * 
-     * @param value
-     *     allowed object is
-     *     {@link String }
-     *     
-     */
-    public void setComponentNodeId(String value) {
[…]

### 2009-03-05 versions/v2/apps/xml/jaxb/message/generated-src/org/netbeams/dsp/message/ContentType.java

@@ -1,57 +1,0 @@
-//
-// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vhudson-jaxb-ri-2.1-661 
-// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
-// Any modifications to this file will be lost upon recompilation of the source schema. 
-// Generated on: 2009.01.13 at 03:30:11 PM BRT 
-//
-
-
-package org.netbeams.dsp.message;
-
-import javax.xml.bind.annotation.XmlEnum;
-import javax.xml.bind.annotation.XmlType;
-
-
-/**
- * <p>Java class for ContentType.
- * 
- * <p>The following schema fragment specifies the expected content contained within this class.
- * <p>
- * <pre>
- * &lt;simpleType name="ContentType">
- *   &lt;restriction base="{http://www.w3.org/2001/XMLSchema}string">
- *     &lt;enumeration value="MEASURE"/>
- *     &lt;enumeration value="ACTION"/>
- *     &lt;enumeration value="CREATE"/>
- *     &lt;enumeration value="INSERT"/>
- *     &lt;enumeration value="QUERY"/>
- *     &lt;enumeration value="DELETE"/>
- *     &lt;enumeration value="UPDATE"/>
- *     &lt;enumeration value="EVENT"/>
- *   &lt;/restriction>
- * &lt;/simpleType>
- * </pre>
- * 
- */
-@XmlType(name = "ContentType")
-@XmlEnum
-public enum ContentType {
-
-    MEASURE,
-    ACTION,
-    CREATE,
-    INSERT,
-    QUERY,
-    DELETE,
-    UPDATE,
-    EVENT;
-
-    public String value() {
-        return name();
-    }
-
-    public static ContentType fromValue(String v) {
-        return valueOf(v);
-    }
-
-}

### 2009-03-05 versions/v2/apps/xml/jaxb/message/generated-src/org/netbeams/dsp/message/CreateMessage.java

@@ -1,39 +1,0 @@
-//
-// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vhudson-jaxb-ri-2.1-661 
-// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
-// Any modifications to this file will be lost upon recompilation of the source schema. 
-// Generated on: 2009.02.25 at 01:49:31 AM PST 
-//
-
-
-package org.netbeams.dsp.message;
-
-import javax.xml.bind.annotation.XmlAccessType;
-import javax.xml.bind.annotation.XmlAccessorType;
-import javax.xml.bind.annotation.XmlType;
-
-
-/**
- * <p>Java class for CreateMessage complex type.
- * 
- * <p>The following schema fragment specifies the expected content contained within this class.
- * 
- * <pre>
- * &lt;complexType name="CreateMessage">
- *   &lt;complexContent>
- *     &lt;extension base="{http://www.netbeams.dsp.org/message.xsd}AbstractMessage">
- *     &lt;/extension>
- *   &lt;/complexContent>
- * &lt;/complexType>
- * </pre>
- * 
- * 
- */
-@XmlAccessorType(XmlAccessType.FIELD)
-@XmlType(name = "CreateMessage")
-public class CreateMessage
-    extends Message
-{
-
-
-}

### 2009-03-05 versions/v2/apps/xml/jaxb/message/generated-src/org/netbeams/dsp/message/DeleteMessage.java

@@ -1,39 +1,0 @@
-//
-// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vhudson-jaxb-ri-2.1-661 
-// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
-// Any modifications to this file will be lost upon recompilation of the source schema. 
-// Generated on: 2009.02.25 at 01:49:31 AM PST 
-//
-
-
-package org.netbeams.dsp.message;
-
-import javax.xml.bind.annotation.XmlAccessType;
-import javax.xml.bind.annotation.XmlAccessorType;
-import javax.xml.bind.annotation.XmlType;
-
-
-/**
- * <p>Java class for DeleteMessage complex type.
- * 
- * <p>The following schema fragment specifies the expected content contained within this class.
- * 
- * <pre>
- * &lt;complexType name="DeleteMessage">
- *   &lt;complexContent>
- *     &lt;extension base="{http://www.netbeams.dsp.org/message.xsd}AbstractMessage">
- *     &lt;/extension>
- *   &lt;/complexContent>
- * &lt;/complexType>
- * </pre>
- * 
- * 
- */
-@XmlAccessorType(XmlAccessType.FIELD)
-@XmlType(name = "DeleteMessage")
-public class DeleteMessage
-    extends Message
-{
-
-
-}

### 2009-03-05 versions/v2/apps/xml/jaxb/message/generated-src/org/netbeams/dsp/message/EventMessage.java

@@ -1,39 +1,0 @@
-//
-// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vhudson-jaxb-ri-2.1-661 
-// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
-// Any modifications to this file will be lost upon recompilation of the source schema. 
-// Generated on: 2009.02.25 at 01:49:31 AM PST 
-//
-
-
-package org.netbeams.dsp.message;
-
-import javax.xml.bind.annotation.XmlAccessType;
-import javax.xml.bind.annotation.XmlAccessorType;
-import javax.xml.bind.annotation.XmlType;
-
-
-/**
- * <p>Java class for EventMessage complex type.
- * 
- * <p>The following schema fragment specifies the expected content contained within this class.
- * 
- * <pre>
- * &lt;complexType name="EventMessage">
- *   &lt;complexContent>
- *     &lt;extension base="{http://www.netbeams.dsp.org/message.xsd}AbstractMessage">
- *     &lt;/extension>
- *   &lt;/complexContent>
- * &lt;/complexType>
- * </pre>
- * 
- * 
- */
-@XmlAccessorType(XmlAccessType.FIELD)
-@XmlType(name = "EventMessage")
-public class EventMessage
-    extends Message
-{
-
-
-}

### 2009-03-05 versions/v2/apps/xml/jaxb/message/generated-src/org/netbeams/dsp/message/Header.java

@@ -1,145 +1,0 @@
-//
-// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vhudson-jaxb-ri-2.1-661 
-// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
-// Any modifications to this file will be lost upon recompilation of the source schema. 
-// Generated on: 2009.02.25 at 01:49:31 AM PST 
-//
-
-
-package org.netbeams.dsp.message;
-
-import javax.xml.bind.annotation.XmlAccessType;
-import javax.xml.bind.annotation.XmlAccessorType;
-import javax.xml.bind.annotation.XmlElement;
-import javax.xml.bind.annotation.XmlType;
-
-
-/**
- * <p>Java class for Header complex type.
- * 
- * <p>The following schema fragment specifies the expected content contained within this class.
- * 
- * <pre>
- * &lt;complexType name="Header">
- *   &lt;complexContent>
- *     &lt;restriction base="{http://www.w3.org/2001/XMLSchema}anyType">
- *       &lt;sequence>
- *         &lt;element name="CreationTime" type="{http://www.w3.org/2001/XMLSchema}long"/>
- *         &lt;element name="CorrelationID" type="{http://www.w3.org/2001/XMLSchema}anyType"/>
- *         &lt;element name="Producer" type="{http://www.netbeams.dsp.org/message.xsd}ComponentIdentifier"/>
- *         &lt;element name="Consumer" type="{http://www.netbeams.dsp.org/message.xsd}ComponentIdentifier"/>
- *       &lt;/sequence>
- *     &lt;/restriction>
- *   &lt;/complexContent>
- * &lt;/complexType>
- * </pre>
- * 
- * 
- */
-@XmlAccessorType(XmlAccessType.FIELD)
-@XmlType(name = "Header", propOrder = {
-    "creationTime",
-    "correlationID",
-    "producer",
-    "consumer"
-})
-public class Header {
-
-    @XmlElement(name = "CreationTime")
-    protected long creationTime;
-    @XmlElement(name = "CorrelationID", required = true)
-    protected Object correlationID;
-    @XmlElement(name = "Producer", required = true)
-    protected ComponentIdentifier producer;
-    @XmlElement(name = "Consumer", required = true)
-    protected ComponentIdentifier consumer;
-
-    /**
-     * Gets the value of the creationTime property.
-     * 
-     */
[…]

### 2009-03-05 versions/v2/apps/xml/jaxb/message/generated-src/org/netbeams/dsp/message/InsertMessage.java

@@ -1,39 +1,0 @@
-//
-// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vhudson-jaxb-ri-2.1-661 
-// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
-// Any modifications to this file will be lost upon recompilation of the source schema. 
-// Generated on: 2009.02.25 at 01:49:31 AM PST 
-//
-
-
-package org.netbeams.dsp.message;
-
-import javax.xml.bind.annotation.XmlAccessType;
-import javax.xml.bind.annotation.XmlAccessorType;
-import javax.xml.bind.annotation.XmlType;
-
-
-/**
- * <p>Java class for InsertMessage complex type.
- * 
- * <p>The following schema fragment specifies the expected content contained within this class.
- * 
- * <pre>
- * &lt;complexType name="InsertMessage">
- *   &lt;complexContent>
- *     &lt;extension base="{http://www.netbeams.dsp.org/message.xsd}AbstractMessage">
- *     &lt;/extension>
- *   &lt;/complexContent>
- * &lt;/complexType>
- * </pre>
- * 
- * 
- */
-@XmlAccessorType(XmlAccessType.FIELD)
-@XmlType(name = "InsertMessage")
-public class InsertMessage
-    extends Message
-{
-
-
-}

### 2009-03-05 versions/v2/apps/xml/jaxb/message/generated-src/org/netbeams/dsp/message/MeasureMessage.java

@@ -1,39 +1,0 @@
-//
-// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vhudson-jaxb-ri-2.1-661 
-// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
-// Any modifications to this file will be lost upon recompilation of the source schema. 
-// Generated on: 2009.02.25 at 01:49:31 AM PST 
-//
-
-
-package org.netbeams.dsp.message;
-
-import javax.xml.bind.annotation.XmlAccessType;
-import javax.xml.bind.annotation.XmlAccessorType;
-import javax.xml.bind.annotation.XmlType;
-
-
-/**
- * <p>Java class for MeasureMessage complex type.
- * 
- * <p>The following schema fragment specifies the expected content contained within this class.
- * 
- * <pre>
- * &lt;complexType name="MeasureMessage">
- *   &lt;complexContent>
- *     &lt;extension base="{http://www.netbeams.dsp.org/message.xsd}AbstractMessage">
- *     &lt;/extension>
- *   &lt;/complexContent>
- * &lt;/complexType>
- * </pre>
- * 
- * 
- */
-@XmlAccessorType(XmlAccessType.FIELD)
-@XmlType(name = "MeasureMessage")
-public class MeasureMessage
-    extends Message
-{
-
-
-}

### 2009-03-05 versions/v2/apps/xml/jaxb/message/generated-src/org/netbeams/dsp/message/MessagesContainer.java

@@ -1,187 +1,0 @@
-//
-// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vhudson-jaxb-ri-2.1-661 
-// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
-// Any modifications to this file will be lost upon recompilation of the source schema. 
-// Generated on: 2009.02.25 at 01:49:31 AM PST 
-//
-
-
-package org.netbeams.dsp.message;
-
-import java.util.ArrayList;
-import java.util.List;
-import javax.xml.bind.annotation.XmlAccessType;
-import javax.xml.bind.annotation.XmlAccessorType;
-import javax.xml.bind.annotation.XmlAttribute;
-import javax.xml.bind.annotation.XmlElement;
-import javax.xml.bind.annotation.XmlRootElement;
-import javax.xml.bind.annotation.XmlType;
-
-
-/**
- * <p>Java class for anonymous complex type.
- * 
- * <p>The following schema fragment specifies the expected content contained within this class.
- * 
- * <pre>
- * &lt;complexType>
- *   &lt;complexContent>
- *     &lt;restriction base="{http://www.w3.org/2001/XMLSchema}anyType">
- *       &lt;sequence>
- *         &lt;element name="message" type="{http://www.netbeams.dsp.org/message.xsd}AbstractMessage" maxOccurs="unbounded" minOccurs="0"/>
- *       &lt;/sequence>
- *       &lt;attribute name="uudi" use="required" type="{http://www.w3.org/2001/XMLSchema}string" />
- *       &lt;attribute name="creationTime" use="required" type="{http://www.w3.org/2001/XMLSchema}string" />
- *       &lt;attribute name="destinationHost" type="{http://www.w3.org/2001/XMLSchema}string" />
- *       &lt;attribute name="acknowledgeUntil" type="{http://www.w3.org/2001/XMLSchema}int" />
- *     &lt;/restriction>
- *   &lt;/complexContent>
- * &lt;/complexType>
- * </pre>
- * 
- * 
- */
-@XmlAccessorType(XmlAccessType.FIELD)
-@XmlType(name = "", propOrder = {
-    "message"
-})
-@XmlRootElement(name = "MessagesContainer")
-public class MessagesContainer {
-
-    @XmlElement(type = Message.class)
-    protected List<AbstractMessage> message;
-    @XmlAttribute(required = true)
-    protected String uudi;
[…]

### 2009-03-05 versions/v2/apps/xml/jaxb/message/generated-src/org/netbeams/dsp/message/ObjectFactory.java

@@ -1,260 +1,0 @@
-//
-// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vhudson-jaxb-ri-2.1-661 
-// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
-// Any modifications to this file will be lost upon recompilation of the source schema. 
-// Generated on: 2009.02.25 at 01:49:31 AM PST 
-//
-
-
-package org.netbeams.dsp.message;
-
-import javax.xml.bind.JAXBElement;
-import javax.xml.bind.annotation.XmlElementDecl;
-import javax.xml.bind.annotation.XmlRegistry;
-import javax.xml.namespace.QName;
-
-
-/**
- * This object contains factory methods for each 
- * Java content interface and Java element interface 
- * generated in the org.netbeams.dsp.message package. 
- * <p>An ObjectFactory allows you to programatically 
- * construct new instances of the Java representation 
- * for XML content. The Java representation of XML 
- * content can consist of schema derived interfaces 
- * and classes representing the binding of schema 
- * type definitions, element declarations and model 
- * groups.  Factory methods for each of these are 
- * provided in this class.
- * 
- */
-@XmlRegistry
-public class ObjectFactory {
-
-    private final static QName _InsertMessage_QNAME = new QName("http://www.netbeams.dsp.org/message.xsd", "InsertMessage");
-    private final static QName _DeleteMessage_QNAME = new QName("http://www.netbeams.dsp.org/message.xsd", "DeleteMessage");
-    private final static QName _CreateMessage_QNAME = new QName("http://www.netbeams.dsp.org/message.xsd", "CreateMessage");
-    private final static QName _UpdateMessage_QNAME = new QName("http://www.netbeams.dsp.org/message.xsd", "UpdateMessage");
-    private final static QName _AcknowledgementMessage_QNAME = new QName("http://www.netbeams.dsp.org/message.xsd", "AcknowledgementMessage");
-    private final static QName _ActionMessage_QNAME = new QName("http://www.netbeams.dsp.org/message.xsd", "ActionMessage");
-    private final static QName _QueryMessage_QNAME = new QName("http://www.netbeams.dsp.org/message.xsd", "QueryMessage");
-    private final static QName _MeasureMessage_QNAME = new QName("http://www.netbeams.dsp.org/message.xsd", "MeasureMessage");
[…]

### 2009-03-05 versions/v2/apps/xml/jaxb/message/generated-src/org/netbeams/dsp/message/QueryMessage.java

@@ -1,39 +1,0 @@
-//
-// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vhudson-jaxb-ri-2.1-661 
-// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
-// Any modifications to this file will be lost upon recompilation of the source schema. 
-// Generated on: 2009.02.25 at 01:49:31 AM PST 
-//
-
-
-package org.netbeams.dsp.message;
-
-import javax.xml.bind.annotation.XmlAccessType;
-import javax.xml.bind.annotation.XmlAccessorType;
-import javax.xml.bind.annotation.XmlType;
-
-
-/**
- * <p>Java class for QueryMessage complex type.
- * 
- * <p>The following schema fragment specifies the expected content contained within this class.
- * 
- * <pre>
- * &lt;complexType name="QueryMessage">
- *   &lt;complexContent>
- *     &lt;extension base="{http://www.netbeams.dsp.org/message.xsd}AbstractMessage">
- *     &lt;/extension>
- *   &lt;/complexContent>
- * &lt;/complexType>
- * </pre>
- * 
- * 
- */
-@XmlAccessorType(XmlAccessType.FIELD)
-@XmlType(name = "QueryMessage")
-public class QueryMessage
-    extends Message
-{
-
-
-}

### 2009-03-05 versions/v2/apps/xml/jaxb/message/generated-src/org/netbeams/dsp/message/UpdateMessage.java

@@ -1,39 +1,0 @@
-//
-// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vhudson-jaxb-ri-2.1-661 
-// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
-// Any modifications to this file will be lost upon recompilation of the source schema. 
-// Generated on: 2009.02.25 at 01:49:31 AM PST 
-//
-
-
-package org.netbeams.dsp.message;
-
-import javax.xml.bind.annotation.XmlAccessType;
-import javax.xml.bind.annotation.XmlAccessorType;
-import javax.xml.bind.annotation.XmlType;
-
-
-/**
- * <p>Java class for UpdateMessage complex type.
- * 
- * <p>The following schema fragment specifies the expected content contained within this class.
- * 
- * <pre>
- * &lt;complexType name="UpdateMessage">
- *   &lt;complexContent>
- *     &lt;extension base="{http://www.netbeams.dsp.org/message.xsd}AbstractMessage">
- *     &lt;/extension>
- *   &lt;/complexContent>
- * &lt;/complexType>
- * </pre>
- * 
- * 
- */
-@XmlAccessorType(XmlAccessType.FIELD)
-@XmlType(name = "UpdateMessage")
-public class UpdateMessage
-    extends Message
-{
-
-
-}

### 2009-03-05 versions/v2/apps/xml/jaxb/message/generated-src/org/netbeams/dsp/message/package-info.java

@@ -1,9 +1,0 @@
-//
-// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vhudson-jaxb-ri-2.1-661 
-// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
-// Any modifications to this file will be lost upon recompilation of the source schema. 
-// Generated on: 2009.02.25 at 01:49:31 AM PST 
-//
-
-@javax.xml.bind.annotation.XmlSchema(namespace = "http://www.netbeams.dsp.org/message.xsd")
-package org.netbeams.dsp.message;

### 2009-03-05 versions/v2/apps/xml/jaxb/message/test/org/netbeams/dsp/message/JaxbMessagesMarshallUnmarshallTest.java
Updating the test case.
git-svn-id: svn://localhost/netbeams/trunk@338 559edde2-a4d4-44e0-96fc-4035e40d7396

@@ -2,22 +2,26 @@ package org.netbeams.dsp.message;
 
 import java.io.File;
 import java.io.FileWriter;
-import java.net.InetAddress;
-import java.net.UnknownHostException;
 
 import javax.xml.bind.JAXBContext;
 import javax.xml.bind.JAXBException;
 import javax.xml.bind.Marshaller;
+import javax.xml.bind.Unmarshaller;
 import javax.xml.parsers.ParserConfigurationException;
 
 import junit.framework.TestCase;
 
+import org.netbeams.dsp.data.property.DSProperties;
+import org.netbeams.dsp.data.property.DSProperty;
+import org.netbeams.dsp.data.property.ObjectFactory;
 import org.netbeams.dsp.demo.mouseactions.ButtonName;
 import org.netbeams.dsp.demo.mouseactions.EventName;
 import org.netbeams.dsp.demo.mouseactions.MouseAction;
 import org.netbeams.dsp.demo.mouseactions.MouseActionsContainer;
 import org.netbeams.dsp.demo.stock.StockTick;
 import org.netbeams.dsp.demo.stock.StockTicks;
+import org.netbeams.dsp.util.NetworkUtil;
+import org.w3c.dom.Node;
 
 public class JaxbMessagesMarshallUnmarshallTest extends TestCase {
 
@@ -61,6 +65,48 @@ public class JaxbMessagesMarshallUnmarshallTest extends TestCase {
             fail(e.getMessage());
         }
 
+        ComponentIdentifier prod3 = this.fac.makeDSPComponentIdentifier("1234", "LOCAL",
+        "org.netbeams.dsp.platform.management.component.ComponentManager");
+        ComponentIdentifier cons3 = this.fac.makeDSPComponentIdentifier(null, "LOCAL",
+                "org.netbeams.dsp.wiretransport.client");
+        Header header3 = this.fac.makeDSPMessageHeader(null, prod3, cons3);
+        
+        /* ------------------ Creation of the Stock Ticks
+        The DSP component writer needs just to create the payload in the way it should be as the
+        body of the DSP message.
+        */  
+        
+        DSProperties properties = new ObjectFactory().createDSProperties();
+        DSProperty serverIp = new DSProperty();
+        serverIp.setName("WIRE_TRANSPORT_SERVER_IP");
+        serverIp.setValue("127.0.0.1");
+        properties.getProperty().add(serverIp);
+        
+        DSProperty serverPort = new DSProperty();
[…]

### 2009-03-05 versions/v2/apps/osgi-bundles/dsp/DSPJSwingMouseClient/src/org/netbeams/dsp/demo/mouseactions/controller/DSPMouseActionsProducer.java
Major update to the Persistence Layer (XML marshalling). This revision has refactored classes that removes the dependency with JAXB. Instead, Marshall is produced using StringBuilder. XML Unmarshalling will be introduced in the next revisions.
All DSP Message classes were refactored to have a method called 
toXml(). This method implements the marshalling of a given class.

MessageContent now implements an interface that contains the method
toXml(). This interface assures that ANYONE who creates a new MessageContent
type must implement their own XML marshalling method. Please verify the
JUnit t

@@ -3,9 +3,6 @@ package org.netbeams.dsp.demo.mouseactions.controller;
 import java.util.LinkedList;
 import java.util.List;
 
-import javax.xml.bind.JAXBException;
-import javax.xml.parsers.ParserConfigurationException;
-
 import org.apache.log4j.Logger;
 import org.netbeams.dsp.DSPContext;
 import org.netbeams.dsp.DSPException;
@@ -14,7 +11,6 @@ import org.netbeams.dsp.demo.mouseactions.ButtonName;
 import org.netbeams.dsp.demo.mouseactions.EventName;
 import org.netbeams.dsp.demo.mouseactions.MouseAction;
 import org.netbeams.dsp.demo.mouseactions.MouseActionsContainer;
-import org.netbeams.dsp.demo.mouseactions.ObjectFactory;
 import org.netbeams.dsp.demo.mouseactions.model.NetBeamsMouseInfo;
 import org.netbeams.dsp.message.ComponentIdentifier;
 import org.netbeams.dsp.message.DSPMessagesFactory;
@@ -90,7 +86,7 @@ public class DSPMouseActionsProducer implements NetBeamsMouseListener {
         Header header = DSPMessagesFactory.INSTANCE.makeDSPMessageHeader(null, producer, null);
 
         try {
-            Message message = DSPMessagesFactory.INSTANCE.makeDSPMeasureMessage(header, data, ObjectFactory.class);
+            Message message = DSPMessagesFactory.INSTANCE.makeDSPMeasureMessage(header, data);
             
             // Always check if there is a broker available
             MessageBrokerAccessor messageBroker = this.dspContext.getDataBroker();
@@ -100,10 +96,6 @@ public class DSPMouseActionsProducer implements NetBeamsMouseListener {
                     log.debug("Message broker not available");
             }
             
-        } catch (JAXBException e) {
-            log.error(e.getMessage(), e);
-        } catch (ParserConfigurationException e) {
-            log.error(e.getMessage(), e);
         } catch (DSPException e) {
             log.error(e.getMessage(), e);
         }

### 2009-03-05 versions/v2/apps/xml/jaxb/dsproperty/java-src/org/netbeams/dsp/data/property/DSProperties.java

@@ -1,0 +1,81 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vhudson-jaxb-ri-2.1-661 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2009.03.04 at 06:15:49 PM PST 
+//
+
+
+package org.netbeams.dsp.data.property;
+
+import java.util.ArrayList;
+import java.util.List;
+
+import org.netbeams.dsp.message.MessageContent;
+
+
+/**
+ * <p>Java class for anonymous complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType>
+ *   &lt;complexContent>
+ *     &lt;extension base="{http://www.netbeams.dsp.org/message.xsd}AbstractMessageContent">
+ *       &lt;sequence maxOccurs="unbounded">
+ *         &lt;element name="Property" type="{http://www.netbeams.dsp.org/dsproperty.xsd}DSProperty"/>
+ *       &lt;/sequence>
+ *     &lt;/extension>
+ *   &lt;/complexContent>
+ * &lt;/complexType>
+ * </pre>
+ * 
+ * 
+ */
+public class DSProperties
+    extends MessageContent
+{
+
+//    @XmlElement(name = "Property", required = true)
+    protected List<DSProperty> property;
+
+    /**
+     * Gets the value of the property property.
+     * 
+     * <p>
+     * This accessor method returns a reference to the live list,
+     * not a snapshot. Therefore any modification you make to the
+     * returned list will be present inside the JAXB object.
+     * This is why there is not a <CODE>set</CODE> method for the property property.
+     * 
+     * <p>
+     * For example, to add a new item, do as follows:
+     * <pre>
+     *    getProperty().add(newItem);
+     * </pre>
+     * 
+     * 
+     * <p>
+     * Objects of the following type(s) are allowed in the list
+     * {@link DSProperty }
+     * 
+     * 
+     */
+    public List<DSProperty> getProperty() {
+        if (property == null) {
+            property = new ArrayList<DSProperty>();
+        }
+        return this.property;
+    }
+
+    public String toXml() {
+        StringBuilder builder = new StringBuilder();
+        builder.append("<DSProperties>");
[…]

### 2009-03-05 versions/v2/apps/xml/jaxb/dsproperty/java-src/org/netbeams/dsp/data/property/DSProperty.java

@@ -1,0 +1,251 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vhudson-jaxb-ri-2.1-661 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2009.03.04 at 06:15:49 PM PST 
+//
+
+
+package org.netbeams.dsp.data.property;
+
+/**
+ * <p>Java class for DSProperty complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType name="DSProperty">
+ *   &lt;complexContent>
+ *     &lt;restriction base="{http://www.w3.org/2001/XMLSchema}anyType">
+ *       &lt;choice>
+ *         &lt;element name="Value" type="{http://www.w3.org/2001/XMLSchema}string"/>
+ *         &lt;element name="Values" type="{http://www.netbeams.dsp.org/dsproperty.xsd}Values"/>
+ *       &lt;/choice>
+ *       &lt;attribute name="name" type="{http://www.w3.org/2001/XMLSchema}anySimpleType" />
+ *       &lt;attribute name="type" type="{http://www.w3.org/2001/XMLSchema}anySimpleType" />
+ *       &lt;attribute name="format" type="{http://www.w3.org/2001/XMLSchema}anySimpleType" />
+ *       &lt;attribute name="unit" type="{http://www.w3.org/2001/XMLSchema}anySimpleType" />
+ *       &lt;attribute name="instruction" type="{http://www.w3.org/2001/XMLSchema}anySimpleType" />
+ *     &lt;/restriction>
+ *   &lt;/complexContent>
+ * &lt;/complexType>
+ * </pre>
+ * 
+ * 
+ */
+public class DSProperty {
+
+//    @XmlElement(name = "Value")
+    protected String value;
+//    @XmlElement(name = "Values")
+    protected Values values;
+//    @XmlAttribute
+//    @XmlSchemaType(name = "anySimpleType")
+    protected String name;
+//    @XmlAttribute
+//    @XmlSchemaType(name = "anySimpleType")
+    protected String type;
+//    @XmlAttribute
+//    @XmlSchemaType(name = "anySimpleType")
+    protected String format;
+//    @XmlAttribute
+//    @XmlSchemaType(name = "anySimpleType")
[…]

### 2009-03-05 versions/v2/apps/xml/jaxb/dsproperty/java-src/org/netbeams/dsp/data/property/Values.java

@@ -1,0 +1,77 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vhudson-jaxb-ri-2.1-661 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2009.03.04 at 06:15:49 PM PST 
+//
+
+
+package org.netbeams.dsp.data.property;
+
+import java.util.ArrayList;
+import java.util.List;
+
+/**
+ * <p>Java class for Values complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType name="Values">
+ *   &lt;complexContent>
+ *     &lt;restriction base="{http://www.w3.org/2001/XMLSchema}anyType">
+ *       &lt;sequence maxOccurs="unbounded">
+ *         &lt;element name="value" type="{http://www.w3.org/2001/XMLSchema}string"/>
+ *       &lt;/sequence>
+ *     &lt;/restriction>
+ *   &lt;/complexContent>
+ * &lt;/complexType>
+ * </pre>
+ * 
+ * 
+ */
+public class Values {
+
+//    @XmlElement(required = true)
+    protected List<String> value;
+
+    /**
+     * Gets the value of the value property.
+     * 
+     * <p>
+     * This accessor method returns a reference to the live list,
+     * not a snapshot. Therefore any modification you make to the
+     * returned list will be present inside the JAXB object.
+     * This is why there is not a <CODE>set</CODE> method for the value property.
+     * 
+     * <p>
+     * For example, to add a new item, do as follows:
+     * <pre>
+     *    getValue().add(newItem);
+     * </pre>
+     * 
+     * 
+     * <p>
+     * Objects of the following type(s) are allowed in the list
+     * {@link String }
+     * 
+     * 
+     */
+    public List<String> getValue() {
+        if (value == null) {
+            value = new ArrayList<String>();
+        }
+        return this.value;
+    }
+
+    public String toXml() {
+        StringBuilder builder = new StringBuilder();
+        builder.append("<Values>");
+        for(String value : this.value) {
+            builder.append("<value>" + value + "</value>");
+        }
+        builder.append("</Values>");
+        return builder.toString();
+    }
+
+}

### 2009-03-05 versions/v2/apps/xml/jaxb/message/java-src/org/netbeams/dsp/message/AbstractMessage.java

@@ -1,0 +1,152 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vhudson-jaxb-ri-2.1-661 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2009.03.04 at 09:49:05 PM PST 
+//
+
+
+package org.netbeams.dsp.message;
+
+/**
+ * <p>Java class for AbstractMessage complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType name="AbstractMessage">
+ *   &lt;complexContent>
+ *     &lt;restriction base="{http://www.w3.org/2001/XMLSchema}anyType">
+ *       &lt;sequence>
+ *         &lt;element name="Header" type="{http://www.netbeams.dsp.org/message.xsd}Header"/>
+ *         &lt;element name="Body" type="{http://www.netbeams.dsp.org/message.xsd}Body"/>
+ *       &lt;/sequence>
+ *       &lt;attribute name="messageID" type="{http://www.w3.org/2001/XMLSchema}anySimpleType" />
+ *       &lt;attribute name="ContentType" type="{http://www.w3.org/2001/XMLSchema}anySimpleType" />
+ *     &lt;/restriction>
+ *   &lt;/complexContent>
+ * &lt;/complexType>
+ * </pre>
+ * 
+ * 
+ */
+public abstract class AbstractMessage {
+
+    //@XmlElement(name = "Header", required = true)
+    protected Header header;
+    //@XmlElement(name = "Body", required = true)
+    protected Body body;
+    //@XmlAttribute
+    //@XmlSchemaType(name = "anySimpleType")
+    protected String messageID;
+    //@XmlAttribute(name = "ContentType")
+    //@XmlSchemaType(name = "anySimpleType")
+    protected String contentType;
+
+    /**
+     * Gets the value of the header property.
+     * 
+     * @return
+     *     possible object is
+     *     {@link Header }
+     *     
+     */
+    public Header getHeader() {
+        return header;
+    }
+
+    /**
+     * Sets the value of the header property.
+     * 
+     * @param value
+     *     allowed object is
+     *     {@link Header }
+     *     
+     */
+    public void setHeader(Header value) {
+        this.header = value;
+    }
+
[…]

### 2009-03-05 versions/v2/apps/xml/jaxb/message/java-src/org/netbeams/dsp/message/AbstractNodeAddress.java

@@ -1,0 +1,55 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vhudson-jaxb-ri-2.1-661 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2009.03.04 at 09:49:05 PM PST 
+//
+
+
+package org.netbeams.dsp.message;
+
+/**
+ * <p>Java class for AbstractNodeAddress complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType name="AbstractNodeAddress">
+ *   &lt;simpleContent>
+ *     &lt;extension base="&lt;http://www.w3.org/2001/XMLSchema>string">
+ *     &lt;/extension>
+ *   &lt;/simpleContent>
+ * &lt;/complexType>
+ * </pre>
+ * 
+ * 
+ */
+public class AbstractNodeAddress {
+
+    //@XmlValue
+    protected String value;
+
+    /**
+     * Gets the value of the value property.
+     * 
+     * @return
+     *     possible object is
+     *     {@link String }
+     *     
+     */
+    public String getValue() {
+        return value;
+    }
+
+    /**
+     * Sets the value of the value property.
+     * 
+     * @param value
+     *     allowed object is
+     *     {@link String }
+     *     
+     */
+    public void setValue(String value) {
+        this.value = value;
+    }
+}

### 2009-03-05 versions/v2/apps/xml/jaxb/message/java-src/org/netbeams/dsp/message/Body.java

@@ -1,0 +1,69 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vhudson-jaxb-ri-2.1-661 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2009.03.04 at 09:49:05 PM PST 
+//
+
+
+package org.netbeams.dsp.message;
+
+
+/**
+ * <p>Java class for Body complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType name="Body">
+ *   &lt;complexContent>
+ *     &lt;restriction base="{http://www.w3.org/2001/XMLSchema}anyType">
+ *       &lt;sequence>
+ *         &lt;any/>
+ *       &lt;/sequence>
+ *     &lt;/restriction>
+ *   &lt;/complexContent>
+ * &lt;/complexType>
+ * </pre>
+ * 
+ * 
+ */
+public class Body {
+
+    //@XmlAnyElement(lax = true)
+    protected MessageContent any;
+
+    /**
+     * Gets the value of the any property.
+     * 
+     * @return
+     *     possible object is
+     *     {@link Element }
+     *     {@link Object }
+     *     
+     */
+    public MessageContent getAny() {
+        return any;
+    }
+
+    /**
+     * Sets the value of the any property.
+     * 
+     * @param value
+     *     allowed object is
+     *     {@link Element }
+     *     {@link Object }
+     *     
+     */
+    public void setAny(MessageContent value) {
+        this.any = value;
+    }
+    
+    public String toXml() {
+        StringBuilder builder = new StringBuilder();
+        builder.append("<Body>");
+        builder.append(this.any.toXml());
+        builder.append("</Body>");
+        return builder.toString();
+    }
+}

### 2009-03-05 versions/v2/apps/xml/jaxb/message/java-src/org/netbeams/dsp/message/ComponentIdentifier.java

@@ -1,0 +1,92 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vhudson-jaxb-ri-2.1-661 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2009.03.04 at 09:49:05 PM PST 
+//
+
+
+package org.netbeams.dsp.message;
+
+/**
+ * <p>Java class for ComponentIdentifier complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType name="ComponentIdentifier">
+ *   &lt;complexContent>
+ *     &lt;restriction base="{http://www.w3.org/2001/XMLSchema}anyType">
+ *       &lt;sequence>
+ *         &lt;element name="ComponentType" type="{http://www.w3.org/2001/XMLSchema}string"/>
+ *         &lt;element name="ComponentLocator" type="{http://www.netbeams.dsp.org/message.xsd}ComponentLocator"/>
+ *       &lt;/sequence>
+ *     &lt;/restriction>
+ *   &lt;/complexContent>
+ * &lt;/complexType>
+ * </pre>
+ * 
+ * 
+ */
+public class ComponentIdentifier {
+
+    //@XmlElement(name = "ComponentType", required = true)
+    protected String componentType;
+    //@XmlElement(name = "ComponentLocator", required = true)
+    protected ComponentLocator componentLocator;
+
+    /**
+     * Gets the value of the componentType property.
+     * 
+     * @return
+     *     possible object is
+     *     {@link String }
+     *     
+     */
+    public String getComponentType() {
+        return componentType;
+    }
+
+    /**
+     * Sets the value of the componentType property.
+     * 
+     * @param value
+     *     allowed object is
+     *     {@link String }
+     *     
+     */
+    public void setComponentType(String value) {
+        this.componentType = value;
+    }
+
+    /**
+     * Gets the value of the componentLocator property.
+     * 
+     * @return
+     *     possible object is
+     *     {@link ComponentLocator }
+     *     
+     */
+    public ComponentLocator getComponentLocator() {
+        return componentLocator;
+    }
+
+    /**
+     * Sets the value of the componentLocator property.
+     * 
+     * @param value
+     *     allowed object is
+     *     {@link ComponentLocator }
+     *     
[…]

### 2009-03-05 versions/v2/apps/xml/jaxb/message/java-src/org/netbeams/dsp/message/ComponentLocator.java

@@ -1,0 +1,96 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vhudson-jaxb-ri-2.1-661 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2009.03.04 at 09:49:05 PM PST 
+//
+
+
+package org.netbeams.dsp.message;
+
+/**
+ * <p>Java class for ComponentLocator complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType name="ComponentLocator">
+ *   &lt;complexContent>
+ *     &lt;restriction base="{http://www.w3.org/2001/XMLSchema}anyType">
+ *       &lt;sequence>
+ *         &lt;element name="ComponentNodeId" type="{http://www.w3.org/2001/XMLSchema}string"/>
+ *         &lt;element name="NodeAddress" type="{http://www.netbeams.dsp.org/message.xsd}AbstractNodeAddress"/>
+ *       &lt;/sequence>
+ *     &lt;/restriction>
+ *   &lt;/complexContent>
+ * &lt;/complexType>
+ * </pre>
+ * 
+ * 
+ */
+public class ComponentLocator {
+
+//    @XmlElement(name = "ComponentNodeId", required = true)
+    protected String componentNodeId;
+//    @XmlElement(name = "NodeAddress", required = true, type = NodeAddress.class)
+    protected NodeAddress nodeAddress;
+
+    /**
+     * Gets the value of the componentNodeId property.
+     * 
+     * @return
+     *     possible object is
+     *     {@link String }
+     *     
+     */
+    public String getComponentNodeId() {
+        return componentNodeId;
+    }
+
+    /**
+     * Sets the value of the componentNodeId property.
+     * 
+     * @param value
+     *     allowed object is
+     *     {@link String }
+     *     
+     */
+    public void setComponentNodeId(String value) {
+        this.componentNodeId = value;
+    }
+
+    /**
+     * Gets the value of the nodeAddress property.
+     * 
+     * @return
+     *     possible object is
+     *     {@link AbstractNodeAddress }
+     *     
+     */
+    public AbstractNodeAddress getNodeAddress() {
+        return nodeAddress;
+    }
+
+    /**
+     * Sets the value of the nodeAddress property.
+     * 
[…]

### 2009-03-05 versions/v2/apps/xml/jaxb/message/java-src/org/netbeams/dsp/message/DSPMessagesFactory.java

@@ -4,14 +4,6 @@ import java.text.SimpleDateFormat;
 import java.util.Date;
 import java.util.UUID;
 
-import javax.xml.bind.JAXBContext;
-import javax.xml.bind.JAXBException;
-import javax.xml.bind.Marshaller;
-import javax.xml.parsers.DocumentBuilderFactory;
-import javax.xml.parsers.ParserConfigurationException;
-
-import org.w3c.dom.Element;
-
 /**
  * This is a factory method of the DSP messages
  * 
@@ -117,78 +109,62 @@ public enum DSPMessagesFactory {
      * @param header is the information of the header.
      * @param bodyNode is the instance of a JAXB generated object
      * @return an instance of the Measure Message with the given Id, header and the body with the given contentType.
-     * @throws JAXBException if there's a parsing error from JAXB
-     * @throws ParserConfigurationException if there's a problem parsing the body object
      */
-    private Message makeDSPMessage(Message m1, Header header, MessageContent bodyPayload, Class objectFactoryClass)
-            throws JAXBException, ParserConfigurationException {
+    private Message makeDSPMessage(Message m1, Header header, MessageContent bodyPayload) {
         m1.setContentType(bodyPayload.getContentContextForJAXB());
         m1.setMessageID(UUID.randomUUID().toString());
         m1.setHeader(header);
 
         Body dspMsgBody = this.factory.createBody();
-        JAXBContext context = JAXBContext.newInstance(objectFactoryClass);
-        Marshaller m = context.createMarshaller();
-        Element xmlDocumentNode = DocumentBuilderFactory.newInstance().newDocumentBuilder().newDocument()
-                .createElement("temp");
-        m.marshal(bodyPayload, xmlDocumentNode);
-        dspMsgBody.setAny(xmlDocumentNode.getFirstChild());
+        dspMsgBody.setAny(bodyPayload);
         m1.setBody(dspMsgBody);
 
         return m1;
     }
 
-    public MeasureMessage makeDSPMeasureMessage(Header header, MessageContent body, Class objectFactory)
-            throws JAXBException, ParserConfigurationException {
+    public MeasureMessage makeDSPMeasureMessage(Header header, MessageContent body) {
         MeasureMessage m1 = this.factory.createMeasureMessage();
-        return (MeasureMessage) this.makeDSPMessage(m1, header, body, objectFactory);
+        return (MeasureMessage) this.makeDSPMessage(m1, header, body);
     }
 
[…]

### 2009-03-05 versions/v2/apps/xml/jaxb/message/java-src/org/netbeams/dsp/message/DeleteMessage.java

@@ -1,0 +1,33 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vhudson-jaxb-ri-2.1-661 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2009.03.04 at 09:49:05 PM PST 
+//
+
+
+package org.netbeams.dsp.message;
+
+
+/**
+ * <p>Java class for DeleteMessage complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType name="DeleteMessage">
+ *   &lt;complexContent>
+ *     &lt;extension base="{http://www.netbeams.dsp.org/message.xsd}AbstractMessage">
+ *     &lt;/extension>
+ *   &lt;/complexContent>
+ * &lt;/complexType>
+ * </pre>
+ * 
+ * 
+ */
+public class DeleteMessage
+    extends Message
+{
+
+
+}

### 2009-03-05 versions/v2/apps/xml/jaxb/message/java-src/org/netbeams/dsp/message/Header.java

@@ -1,0 +1,150 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vhudson-jaxb-ri-2.1-661 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2009.03.04 at 09:49:05 PM PST 
+//
+
+
+package org.netbeams.dsp.message;
+
+/**
+ * <p>Java class for Header complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType name="Header">
+ *   &lt;complexContent>
+ *     &lt;restriction base="{http://www.w3.org/2001/XMLSchema}anyType">
+ *       &lt;sequence>
+ *         &lt;element name="CreationTime" type="{http://www.w3.org/2001/XMLSchema}long"/>
+ *         &lt;element name="CorrelationID" type="{http://www.w3.org/2001/XMLSchema}anyType"/>
+ *         &lt;element name="Producer" type="{http://www.netbeams.dsp.org/message.xsd}ComponentIdentifier"/>
+ *         &lt;element name="Consumer" type="{http://www.netbeams.dsp.org/message.xsd}ComponentIdentifier"/>
+ *       &lt;/sequence>
+ *     &lt;/restriction>
+ *   &lt;/complexContent>
+ * &lt;/complexType>
+ * </pre>
+ * 
+ * 
+ */
+public class Header {
+
+//    @XmlElement(name = "CreationTime")
+    protected long creationTime;
+//    @XmlElement(name = "CorrelationID", required = true)
+    protected Object correlationID;
+//    @XmlElement(name = "Producer", required = true)
+    protected ComponentIdentifier producer;
+//    @XmlElement(name = "Consumer", required = true)
+    protected ComponentIdentifier consumer;
+
+    /**
+     * Gets the value of the creationTime property.
+     * 
+     */
+    public long getCreationTime() {
+        return creationTime;
+    }
+
+    /**
+     * Sets the value of the creationTime property.
+     * 
+     */
+    public void setCreationTime(long value) {
+        this.creationTime = value;
+    }
+
+    /**
+     * Gets the value of the correlationID property.
+     * 
+     * @return
+     *     possible object is
+     *     {@link Object }
+     *     
+     */
[…]

### 2009-03-05 versions/v2/apps/xml/jaxb/message/java-src/org/netbeams/dsp/message/MessageContent.java

@@ -1,6 +1,6 @@
 package org.netbeams.dsp.message;
 
-public class MessageContent /* extends AbstractMessageContent*/ {
+public abstract class MessageContent implements XmlDecoratable /* extends AbstractMessageContent*/ {
 
     // TODO: Find way to turn this into an abstract class
     public String getContentType() {

### 2009-03-05 versions/v2/apps/xml/jaxb/message/java-src/org/netbeams/dsp/message/MessagesContainer.java

@@ -1,0 +1,193 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vhudson-jaxb-ri-2.1-661 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2009.03.04 at 09:49:05 PM PST 
+//
+
+
+package org.netbeams.dsp.message;
+
+import java.util.ArrayList;
+import java.util.List;
+
+/**
+ * <p>Java class for anonymous complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType>
+ *   &lt;complexContent>
+ *     &lt;restriction base="{http://www.w3.org/2001/XMLSchema}anyType">
+ *       &lt;sequence>
+ *         &lt;element name="message" type="{http://www.netbeams.dsp.org/message.xsd}AbstractMessage" maxOccurs="unbounded" minOccurs="0"/>
+ *       &lt;/sequence>
+ *       &lt;attribute name="uudi" use="required" type="{http://www.w3.org/2001/XMLSchema}string" />
+ *       &lt;attribute name="creationTime" use="required" type="{http://www.w3.org/2001/XMLSchema}string" />
+ *       &lt;attribute name="destinationHost" type="{http://www.w3.org/2001/XMLSchema}string" />
+ *       &lt;attribute name="acknowledgeUntil" type="{http://www.w3.org/2001/XMLSchema}int" />
+ *     &lt;/restriction>
+ *   &lt;/complexContent>
+ * &lt;/complexType>
+ * </pre>
+ * 
+ * 
+ */
+public class MessagesContainer {
+
+//    @XmlElement(type = Message.class)
+    protected List<AbstractMessage> message;
+//    @XmlAttribute(required = true)
+    protected String uudi;
+//    @XmlAttribute(required = true)
+    protected String creationTime;
+//    @XmlAttribute
+    protected String destinationHost;
+//    @XmlAttribute
+    protected Integer acknowledgeUntil;
+
+    /**
+     * Gets the value of the message property.
+     * 
+     * <p>
+     * This accessor method returns a reference to the live list,
+     * not a snapshot. Therefore any modification you make to the
+     * returned list will be present inside the JAXB object.
[…]

### 2009-03-05 versions/v2/apps/xml/jaxb/message/java-src/org/netbeams/dsp/message/NodeAddress.java

@@ -1,28 +1,32 @@
 package org.netbeams.dsp.message;
 
-
 public class NodeAddress extends AbstractNodeAddress {
 
-	public NodeAddress(){}
+    public NodeAddress() {
+    }
 
-	public NodeAddress(String nodeAddress){
-		setValue(nodeAddress);
-	}
+    public NodeAddress(String nodeAddress) {
+        setValue(nodeAddress);
+    }
 
-	public int hashCode(){
-		return value.hashCode();
-	}
+    public int hashCode() {
+        return value.hashCode();
+    }
 
-	public boolean equals(Object obj){
-		if(this == obj){
-			return true;
-		}
-		if(obj instanceof NodeAddress){
-			NodeAddress o = (NodeAddress)obj;
-			return value.equals(o.getValue());
-		}
-		return false;
-	}
+    public boolean equals(Object obj) {
+        if (this == obj) {
+            return true;
+        }
+        if (obj instanceof NodeAddress) {
+            NodeAddress o = (NodeAddress) obj;
+            return value.equals(o.getValue());
+        }
+        return false;
+    }
 
-
+    public String toXml() {
+        StringBuilder builder = new StringBuilder();
+        builder.append("<NodeAddress>" + this.value + "</NodeAddress>");
+        return builder.toString();
+    }
 }

### 2009-03-05 versions/v2/apps/xml/jaxb/message/java-src/org/netbeams/dsp/message/QueryMessage.java

@@ -1,0 +1,32 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vhudson-jaxb-ri-2.1-661 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2009.03.04 at 09:49:05 PM PST 
+//
+
+
+package org.netbeams.dsp.message;
+
+/**
+ * <p>Java class for QueryMessage complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType name="QueryMessage">
+ *   &lt;complexContent>
+ *     &lt;extension base="{http://www.netbeams.dsp.org/message.xsd}AbstractMessage">
+ *     &lt;/extension>
+ *   &lt;/complexContent>
+ * &lt;/complexType>
+ * </pre>
+ * 
+ * 
+ */
+public class QueryMessage
+    extends Message
+{
+
+
+}

### 2009-03-05 versions/v2/apps/xml/jaxb/message/java-src/org/netbeams/dsp/message/XmlDecoratable.java

@@ -1,0 +1,9 @@
+package org.netbeams.dsp.message;
+
+public interface XmlDecoratable {
+
+    /**
+     * @return the XML instance of a given class representation.
+     */
+    public String toXml();
+}

### 2009-03-05 versions/v2/apps/xml/jaxb/message/test/org/netbeams/dsp/message/JaxbMessagesMarshallUnmarshallTest.java

@@ -3,12 +3,6 @@ package org.netbeams.dsp.message;
 import java.io.File;
 import java.io.FileWriter;
 
-import javax.xml.bind.JAXBContext;
-import javax.xml.bind.JAXBException;
-import javax.xml.bind.Marshaller;
-import javax.xml.bind.Unmarshaller;
-import javax.xml.parsers.ParserConfigurationException;
-
 import junit.framework.TestCase;
 
 import org.netbeams.dsp.data.property.DSProperties;
@@ -21,12 +15,13 @@ import org.netbeams.dsp.demo.mouseactions.MouseActionsContainer;
 import org.netbeams.dsp.demo.stock.StockTick;
 import org.netbeams.dsp.demo.stock.StockTicks;
 import org.netbeams.dsp.util.NetworkUtil;
-import org.w3c.dom.Node;
+import org.netbeams.dsp.ysi.SondeDataContainer;
+import org.netbeams.dsp.ysi.SondeDataType;
 
 public class JaxbMessagesMarshallUnmarshallTest extends TestCase {
 
     private DSPMessagesFactory fac;
-    private File xmlObjectPersisted = new File("/tmp/messagesExample.xml");
+    private File xmlObjectPersisted = new File("/tmp/messagesExample2.xml");
 
     public JaxbMessagesMarshallUnmarshallTest() {
         this.fac = DSPMessagesFactory.INSTANCE;        
@@ -55,15 +50,8 @@ public class JaxbMessagesMarshallUnmarshallTest extends TestCase {
         tk1.setValue((float) 23.4);
         ticks.getStockTick().add(tk1);
 
-        MeasureMessage stocks = null;
-        try {
-            //Specifying the correct content type for the message as the package of the component
-            stocks = this.fac.makeDSPMeasureMessage(header, ticks, org.netbeams.dsp.demo.stock.ObjectFactory.class);
-        } catch (JAXBException e1) {
-            fail(e1.getMessage());
-        } catch (ParserConfigurationException e) {
-            fail(e.getMessage());
-        }
+        //Specifying the correct content type for the message as the package of the component
+        MeasureMessage stocks = this.fac.makeDSPMeasureMessage(header, ticks);
 
         ComponentIdentifier prod3 = this.fac.makeDSPComponentIdentifier("1234", "LOCAL",
         "org.netbeams.dsp.platform.management.component.ComponentManager");
@@ -98,15 +86,10 @@ public class JaxbMessagesMarshallUnmarshallTest extends TestCase {
[…]

### 2009-03-05 versions/v2/apps/xml/jaxb/sonde/java-src/org/netbeams/dsp/ysi/SondeDataContainer.java

@@ -1,0 +1,80 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vhudson-jaxb-ri-2.1-661 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2009.03.04 at 06:20:22 PM PST 
+//
+
+
+package org.netbeams.dsp.ysi;
+
+import java.util.ArrayList;
+import java.util.List;
+import org.netbeams.dsp.message.MessageContent;
+
+
+/**
+ * <p>Java class for anonymous complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType>
+ *   &lt;complexContent>
+ *     &lt;extension base="{http://www.netbeams.dsp.org/message.xsd}AbstractMessageContent">
+ *       &lt;sequence maxOccurs="unbounded">
+ *         &lt;element name="sondeData" type="{http://www.netbeams.dsp.org/sondedata.xsd}sondeDataType"/>
+ *       &lt;/sequence>
+ *     &lt;/extension>
+ *   &lt;/complexContent>
+ * &lt;/complexType>
+ * </pre>
+ * 
+ * 
+ */
+public class SondeDataContainer
+    extends MessageContent
+{
+
+//    @XmlElement(required = true)
+    protected List<SondeDataType> sondeData;
+
+    /**
+     * Gets the value of the sondeData property.
+     * 
+     * <p>
+     * This accessor method returns a reference to the live list,
+     * not a snapshot. Therefore any modification you make to the
+     * returned list will be present inside the JAXB object.
+     * This is why there is not a <CODE>set</CODE> method for the sondeData property.
+     * 
+     * <p>
+     * For example, to add a new item, do as follows:
+     * <pre>
+     *    getSondeData().add(newItem);
+     * </pre>
+     * 
+     * 
+     * <p>
+     * Objects of the following type(s) are allowed in the list
+     * {@link SondeDataType }
+     * 
+     * 
+     */
+    public List<SondeDataType> getSondeData() {
+        if (sondeData == null) {
+            sondeData = new ArrayList<SondeDataType>();
+        }
+        return this.sondeData;
+    }
+
+    public String toXml() {
+        StringBuilder builder = new StringBuilder();
[…]

### 2009-03-05 versions/v2/apps/xml/jaxb/sonde/java-src/org/netbeams/dsp/ysi/SondeDataType.java

@@ -1,0 +1,481 @@
+//
+// This file was generated by the JavaTM Architecture for XML Binding(JAXB) Reference Implementation, vhudson-jaxb-ri-2.1-661 
+// See <a href="http://java.sun.com/xml/jaxb">http://java.sun.com/xml/jaxb</a> 
+// Any modifications to this file will be lost upon recompilation of the source schema. 
+// Generated on: 2009.03.04 at 06:20:22 PM PST 
+//
+
+
+package org.netbeams.dsp.ysi;
+
+/**
+ * <p>Java class for sondeDataType complex type.
+ * 
+ * <p>The following schema fragment specifies the expected content contained within this class.
+ * 
+ * <pre>
+ * &lt;complexType name="sondeDataType">
+ *   &lt;complexContent>
+ *     &lt;restriction base="{http://www.w3.org/2001/XMLSchema}anyType">
+ *       &lt;sequence>
+ *         &lt;element name="Date" type="{http://www.w3.org/2001/XMLSchema}string"/>
+ *         &lt;element name="Time" type="{http://www.w3.org/2001/XMLSchema}string"/>
+ *         &lt;element name="Temp" type="{http://www.w3.org/2001/XMLSchema}string"/>
+ *         &lt;element name="SpCond" type="{http://www.w3.org/2001/XMLSchema}string"/>
+ *         &lt;element name="Cond" type="{http://www.w3.org/2001/XMLSchema}string"/>
+ *         &lt;element name="Resist" type="{http://www.w3.org/2001/XMLSchema}string"/>
+ *         &lt;element name="Sal" type="{http://www.w3.org/2001/XMLSchema}string"/>
+ *         &lt;element name="Press" type="{http://www.w3.org/2001/XMLSchema}string"/>
+ *         &lt;element name="Depth" type="{http://www.w3.org/2001/XMLSchema}string"/>
+ *         &lt;element name="pH" type="{http://www.w3.org/2001/XMLSchema}string"/>
+ *         &lt;element name="phmV" type="{http://www.w3.org/2001/XMLSchema}string"/>
+ *         &lt;element name="ODOSat" type="{http://www.w3.org/2001/XMLSchema}string"/>
+ *         &lt;element name="ODOConc" type="{http://www.w3.org/2001/XMLSchema}string"/>
+ *         &lt;element name="Turbid" type="{http://www.w3.org/2001/XMLSchema}string"/>
[…]

### 2009-03-05 versions/v2/apps/xml/jaxb/message/java-src/org/netbeams/dsp/message/DSPMessagesFactory.java
Updating the build scripts that were adding JAXB.
- The main build.xml does only generate Message, Sonde and DSProperties;
- Removing references from JAXB on the Wire Transport client. It still
  needs to be updated with the Unmarshalling method;
- The JUnit test JaxbMessagesMarshallUnmarshall is updated.

git-svn-id: svn://localhost/netbeams/trunk@342 559edde2-a4d4-44e0-96fc-4035e40d7396

@@ -17,27 +17,18 @@ public enum DSPMessagesFactory {
     INSTANCE;
 
     /**
-     * The DSP messages factory object generated by JAXB
-     */
-    private ObjectFactory factory;
-
-    private DSPMessagesFactory() {
-        this.factory = new ObjectFactory();
-    }
-
-    /**
      * @param compId is the identification of the component
      * @param location is the location of the component. It must be a machine name or ip address.
      * @param compType is the complete name of the
      * @return a new instance of a Component Identifier with the specified data.
      */
     public ComponentIdentifier makeDSPComponentIdentifier(String compId, String location, String compType) {
-        ComponentLocator loc = this.factory.createComponentLocator();
+        ComponentLocator loc = new ComponentLocator();
         loc.setComponentNodeId(compId);
-        AbstractNodeAddress nadd = this.factory.createAbstractNodeAddress();
+        AbstractNodeAddress nadd = new AbstractNodeAddress();
         nadd.setValue(location);
         loc.setNodeAddress(nadd);
-        ComponentIdentifier pIdent = this.factory.createComponentIdentifier();
+        ComponentIdentifier pIdent = new ComponentIdentifier();
         pIdent.setComponentLocator(loc);
         pIdent.setComponentType(compType);
         return pIdent;
@@ -51,7 +42,7 @@ public enum DSPMessagesFactory {
      * @return a new instance of the Header for the DSP Message
      */
     public Header makeDSPMessageHeader(String correlationId, ComponentIdentifier producer, ComponentIdentifier consumer) {
-        Header header = this.factory.createHeader();
+        Header header = new Header();
         header.setCorrelationID(correlationId);
         header.setCreationTime(System.currentTimeMillis());
         header.setProducer(producer);
@@ -96,7 +87,7 @@ public enum DSPMessagesFactory {
      *         time of the message.
      */
     public MessagesContainer makeDSPMessagesContainer(String destIpAddress) {
-        MessagesContainer messages = this.factory.createMessagesContainer();
+        MessagesContainer messages = new MessagesContainer();
         messages.setCreationTime(this.makeCurrentIsoDateTime());
         messages.setUudi(UUID.randomUUID().toString());
         messages.setDestinationHost(destIpAddress);
@@ -115,7 +106,7 @@ public enum DSPMessagesFactory {
[…]

### 2009-03-05 versions/v2/apps/xml/jaxb/message/java-src/org/netbeams/dsp/message/MessagesContainer.java

@@ -190,4 +190,8 @@ public class MessagesContainer {
         return builder.toString();
     }
 
+    public static MessagesContainer unmarshall(String responseStream) {
+        return null;
+    }
+
 }

### 2009-03-05 versions/v2/apps/xml/jaxb/message/test/org/netbeams/dsp/message/JaxbMessagesMarshallUnmarshallTest.java

@@ -7,7 +7,6 @@ import junit.framework.TestCase;
 
 import org.netbeams.dsp.data.property.DSProperties;
 import org.netbeams.dsp.data.property.DSProperty;
-import org.netbeams.dsp.data.property.ObjectFactory;
 import org.netbeams.dsp.demo.mouseactions.ButtonName;
 import org.netbeams.dsp.demo.mouseactions.EventName;
 import org.netbeams.dsp.demo.mouseactions.MouseAction;
@@ -64,7 +63,7 @@ public class JaxbMessagesMarshallUnmarshallTest extends TestCase {
         body of the DSP message.
         */  
         
-        DSProperties properties = new ObjectFactory().createDSProperties();
+        DSProperties properties = new DSProperties();
         DSProperty serverIp = new DSProperty();
         serverIp.setName("WIRE_TRANSPORT_SERVER_IP");
         serverIp.setValue("127.0.0.1");

### 2009-03-05 versions/v2/apps/osgi-bundles/dsp/DSPFramework/src/org/netbeams/dsp/MessageFactory.java
Missing file to be updated after the manual marshalling methods.
git-svn-id: svn://localhost/netbeams/trunk@343 559edde2-a4d4-44e0-96fc-4035e40d7396

@@ -1,227 +1,217 @@
 package org.netbeams.dsp;
 
-
 import java.util.List;
 import java.util.UUID;
 
-import javax.xml.bind.JAXBContext;
-import javax.xml.bind.JAXBException;
-import javax.xml.bind.Marshaller;
-import javax.xml.parsers.DocumentBuilder;
-import javax.xml.parsers.DocumentBuilderFactory;
-import javax.xml.parsers.ParserConfigurationException;
-
 import org.apache.log4j.Logger;
-import org.netbeams.dsp.message.*;
+import org.netbeams.dsp.message.ActionMessage;
+import org.netbeams.dsp.message.Body;
+import org.netbeams.dsp.message.ComponentIdentifier;
+import org.netbeams.dsp.message.ComponentLocator;
+import org.netbeams.dsp.message.CreateMessage;
+import org.netbeams.dsp.message.DeleteMessage;
+import org.netbeams.dsp.message.EventMessage;
+import org.netbeams.dsp.message.Header;
+import org.netbeams.dsp.message.InsertMessage;
+import org.netbeams.dsp.message.MeasureMessage;
+import org.netbeams.dsp.message.Message;
+import org.netbeams.dsp.message.MessageContent;
+import org.netbeams.dsp.message.MessagesContainer;
+import org.netbeams.dsp.message.NodeAddress;
+import org.netbeams.dsp.message.QueryMessage;
+import org.netbeams.dsp.message.UpdateMessage;
 import org.netbeams.dsp.util.ErrorCode;
 import org.netbeams.dsp.util.NetworkUtil;
-import org.w3c.dom.Document;
-import org.w3c.dom.Node;
-
 
 /**
- * Currently this factory create XML (JDOM) based message. In the future it would be possible to configure the 
- * a different encoding type.
+ * Currently this factory create XML (JDOM) based message. In the future it would be possible to configure the a
+ * different encoding type.
  * 
  * @author kleber
- *
+ * 
  */
 public class MessageFactory {
-	
-	private static final Logger log = Logger.getLogger(MessageFactory.class);
-	
-	private MessageFactory() {}
-	
-	public static Message newMessage(Class<? extends Message> messageType, MessageContent messageContent, BaseComponent producer)
-		throws DSPException
-	{
-		return newMessage(messageType, messageContent, producer, null, null);
-	}
-	
-	
[…]

### 2009-03-05 versions/v2/apps/osgi-bundles/dsp/DSPSondeProducer/src/org/netbeams/dsp/ysi/SondeTestData.java
Modified the test data class to handle the new schema types.
git-svn-id: svn://localhost/netbeams/trunk@345 559edde2-a4d4-44e0-96fc-4035e40d7396

@@ -1,6 +1,13 @@
 package org.netbeams.dsp.ysi;
 
 import java.util.ArrayList;
+import java.util.Date;
+import java.util.Calendar;
+import java.util.GregorianCalendar;
+import java.util.StringTokenizer;
+import java.text.SimpleDateFormat;
+import javax.xml.datatype.DatatypeFactory;
+import javax.xml.datatype.DatatypeConfigurationException;
 
 /**
  * 
@@ -12,10 +19,14 @@ public class SondeTestData {
 	private SondeDataContainer sdc;
 	private ArrayList<SondeDataType> list;
 	private SondeDataType sdt;
+	private static GregorianCalendar dateTime;
+	private Month m;
 		
 	public SondeTestData() {
+		m = new Month();
 		sdc = new SondeDataContainer();
 		list = new ArrayList<SondeDataType> ();
+		initDateTime();
 		initTestData();
 		sdc.sondeData = list;
 	};
@@ -23,8 +34,14 @@ public class SondeTestData {
 	public SondeDataContainer getSondeTestData() {
 		return sdc;
 	}
+		
+	private void initDateTime() {
+		dateTime = new GregorianCalendar();
+		dateTime.clear();
+		dateTime.set(2008, Calendar.DECEMBER, 2, 22, 48, 53);
+	}
 	
-	private void initTestData() {
+	private void initTestData() {;
 		setDataSet1();
 		setDataSet2();
 		setDataSet3();
@@ -32,63 +49,102 @@ public class SondeTestData {
 	
 	private void setDataSet1() {
 		sdt = new SondeDataType();
-		sdt.setDate("11/26/80");
-		sdt.setTime("05:24:51");
-		sdt.setTemp("22.69");
-		sdt.setSpCond("183.0");
-		sdt.setCond("175.0");
-		sdt.setResist("5704.66");
-		sdt.setSal("0.09");
-		sdt.setPress("-0.00");
-		sdt.setDepth("-0.00");
-		sdt.setPH("8.22");
-		sdt.setPhmV("-94.00");
-		sdt.setODOSat("111.70");
-		sdt.setODOConc("9.63");
-		sdt.setTurbid("0.30");
-		sdt.setBattery("10.10");
+		sdt.setDate(dateTime);
+		sdt.setTime(dateTime);
+		sdt.setTemp((float) 22.69);
+		sdt.setSpCond((float) 183.0);
+		sdt.setCond((float) 175.0);
[…]

### 2009-03-05 versions/v2/apps/osgi-bundles/dsp/DSPSondeProducer/src/org/netbeams/dsp/ysi/SondeProducer.java
Removed all references to Serial connections. Currently testing a serial (non-DSP) bundle which interfaces with the simulator. This hasn't been uploaded yet.  Once tests are complete, this code will merge with SondeProducer's functionality.
git-svn-id: svn://localhost/netbeams/trunk@346 559edde2-a4d4-44e0-96fc-4035e40d7396

@@ -22,17 +22,11 @@ public class SondeProducer {
 	
 private static final Logger log = Logger.getLogger(SondeProducer.class);
 	
-	public static final String SERIAL_PORT = "/dev/ttyS0";
 	private DSPContext context;
 	private SondeHandler sondeHandler;
-	private SerialHandler serialHandler;
-	private String cmdBuffer;
-	private String rcvBuffer;
 	
 	public SondeProducer (DSPContext context) {
 		this.context = context;
-		serialHandler = new SerialHandler();
-		initSerialConnection();
         log.trace("Sonde Producer initialized");
 	};
 	
@@ -48,27 +42,6 @@ private static final Logger log = Logger.getLogger(SondeProducer.class);
 			sondeHandler.stopHandler();
 			sondeHandler = null;
 		}
-	}
-	
-	private void initSerialConnection() {
-		try {
-			serialHandler.connect(SERIAL_PORT);
-		} catch (Exception e) {
-			System.err.println("Error: " + e.getMessage());
-			e.printStackTrace();
-		}
-	}
-	
-	private void configureSonde() {
-		serialHandler.sendSerial("\n");		// ----> [CR]
-		//serialHandler.recvSerial(rcvBuffer);			// <---- "#"
-		//sendSerial(cmdBuffer);		// ----> "menu"
-		//recvSerial(responseBuffer); // <---- (Main Menu)
-		//sendSerial(cmdBuffer);		// ----> "2"[CR]
-		//recvSerial(responseBuffer); // <---- (Unattended Sample Menu)
-		//sendSerial(cmdBuffer);		// ----> "1"[CR]
-		//recvSerial(responseBuffer); // <---- "00:00:00"
-		//sendSerial(cmdBuffer);		// ----> "00:15:00"
 	}
 	
 	
@@ -120,31 +93,21 @@ private static final Logger log = Logger.getLogger(SondeProducer.class);
 		private List<SondeDataType> sondeData;
 		
 		public SondeHandler () {
-			sondeTestData = new SondeTestData();  // Get the test data
+			sondeTestData = new SondeTestData();  // Get producer test data.
 			sondeDataContainer = sondeTestData.getSondeTestData();
 		};
 		
 		
 		public void run() {			
-			// Configure the sonde
-			configureSonde();
-			
-			//while (running) {
+			while (running) {
 				
-				// Communication with a consumer
-				/*
+				// Communication with the consumer
 				try {
-					
[…]

### 2009-03-05 docs/icccn2009/paper.tex
Fixing my name and some reviews to the first sections of the paper.
git-svn-id: svn://localhost/netbeams/trunk@347 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 
-\author{\authorblockN{A. Puder, T. Johnson, K. Sales, M. Sales}
+\author{\authorblockN{A. Puder, T. Johnson, K. Sales, M. de Sales}
 \authorblockA{San Francisco State University \\
@@
 take place. Typically, sensors are deployed in out-of-the-way
-locations where physical access is limited and connections for
+locations, where physical access is limited and connections for
 telemetry are poor or nonexistent. Maintaining the data stream from
@@
 Consequently, there is a lag of days to months between the time the
-sensors perform the measurements and the time the measurements becomes
+sensors perform the measurements and the time the measurements become
 available. Advances in battery life and anti-fouling technology have
@@
 sensors, we propose to develop an automated end-to-end system based on
-proven technology that provides a means to program and interrogate
+proven technology. It provides means to program and interrogate
 sensors at the remote field locations and to transmit collected
-measurements over long distances to data collections centers where the
+measurements over long distances to data collections centers, where the
 data is rapidly processed, archived, and made available to potential
@@
 maintenance difficult, but also access to their measurements.
-Typically the sensors are autonomous in the sense that they can run
+Typically, the sensors are autonomous in the sense that they can run
 for a certain time on battery power and store the measurement data in
 internal memory. Only when the sensors are serviced, the measurement
-data they have collected since the last servicing are uploaded to a
-portable storage device. From there the data can be uploaded to
-server. This manual procedure to retrieve sensor data is not only
-error prone, but also results in a serious time lag between the time
+data, which was collected since the last servicing, is uploaded to a 
+portable storage device. From there, the data can be uploaded to
+server. Notice that this manual procedure to retrieve sensor data is not only
+error prone, but also results are in a serious time lag between the time
 the sensor performs the measurement and when it becomes available for
-further processing. For many applications it would be beneficial to
-have near-realtime access to the sensor measurement. Being able to
-access data when they become available on the sensor device also
-reduces the maintenance overhead.
+further processing. For many applications, however, it would be beneficial to
+have near-realtime access to the sensor measurement. In this way, being able 
+to access data when they become available on the sensor device would also
+reduce the maintenance overhead.
 

### 2009-03-06 docs/icccn2009/paper.tex
More text
git-svn-id: svn://localhost/netbeams/trunk@349 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 
-\title{An OSGi-based Sensor Network for \\ Environmental Monitoring}
+\title{A Component-Based Sensor Network for \\ Environmental Monitoring}
 
@@
 
+Mention the nature of our sensor network: highly specialized sensors,
+sparse topology, remote locations, long range communication.
 \cite{roemer:2004}
+
+---
 
@@
 \section{Background}
+\label{SEC_BACKGROUND}
 
@@
 The YSI 6600EDS V2 Sonde is a water quality monitoring device that
-gathers water quality data, and operates in fresh, sea, or
-polluted water. It is under 22 inches (under 55 cm) in length, 3.5
-inches (8.9 cm) in diameter, and weighs approximately 7 pounds (3.18
-kg). It operates at temperatures between -5 to +50 deg C and at depths
-up to 656 feet (200 meters).  It uses 8 C-size Alkaline Batteries or
-External 12 VDC. Battery life can last up to 75 days depending on
-sensor configuration. The Sonde's bulkhead contains multiple pin
-connectors to support sensor probes for measuring the following:
+gathers water quality data, and operates in fresh, sea, or polluted
+water. Some of the measurements that the YSI is capable of are
+conductivity, temperature, chloride, ammonium, nitrate, turbidity, and
+chlorophyll.  It is under 55 cm in length, 8.9 cm in diameter, and
+weighs approximately 3.18 kg. It operates at temperatures between -5
+to +50 deg C and at depths up to 200 meters.  It uses 8 C-size
+Alkaline Batteries or External 12 VDC. Battery life can last up to 75
+days depending on sensor configuration. The Sonde's bulkhead contains
+multiple pin connectors to support sensor probes for measuring the
+following:
 
@@
 
-Dissolved oxygen
-Conductivity/Temperature
-Fouling Resistant pH/ORP
-Chloride
-Ammonium
-Nitrate
-Turbidity
-Chlorophyll
-Rhodamine
-PC-Blue-green Algae
-PE-Blue-gren Algae
-
 Each probe may have multiple sensors.  A computer can interface with
@@
 capabilities, time-delayed communication, etc. In this section should
-be no mentioned of DSP or OSGi. Length: 1 page.
+be no mentioning of DSP. Length: 1 page.
 
@@
 
-Should describe the OSGi-based NetBEAMS architecture. Total length of
-this section: 2 pages.
-
-Micro-kernel architecture.
+Should describe DSP architecture. Only mention OSGi in last subsection
+and explain that components can be mapped to bundles. Currently no
+distinction between Platform and Framework. Necessary?
 
@@
 
-In order to enrich the DSP, a programmer has to implement new DSP
-Components.
-
-Should give some API sniplets for DP and DC.
+All DSP Components need to implement the following Java interface:
 
@@
 
+By implementing the above Java interface, the resuling code
+effectively becomes a DSP Component that can be deployed at runtime. A
+component is identified by its type. Only one component of a certain
+type per DSP is permissible. A string serves as a type identifier that
+much be unique across all DSP Components. Upon deploying a DSP
+Component, an initialization function is called (line 4). Once a
+component is initialized, it may be started (line 5) and stopped (line
+6) multiple time (TODO: SHOULDN'T THERE BE ALSO A WAY TO UNDEPLOY A
+COMPONENT?) Whenever the DSP wishes to pass a message (that was
+generated by some DP), it invokes the \texttt{deliver()} method (ine
+7). The message structure will be explained in the following section.
+
 \subsection{Message}
+
+Message passing is the main paradigm for exchanging information within
+a DSP. A data producer sends a message with data consumers accept
+messages. We make use of XML technologies to define the format of a
+message. Specifically, the common header of all messages are defined
+via the following XML Schema:
 
@@
 
-Should explain the way DSP Messages are defined: XML Schema and JAXB.
+The above XML Schema DSPMessage serves as a base schema for all
+messages and itself defines information typically contained in the
+header of a protocol data unit. Amongst others, the header contains a
+unique message ID (line 8) and the originating DP (line 9). Using the
+XML Schema, we use JAXB to generate a type safe API for handling
+instances of DSPMessage. Different messages with different payloads
+can be defined by creating schemas that derive from DSPMessage. While
+inside the DSP all messages are of type DSPMessage, specific DP and DC
+can distinguish between different message types based on their runtime
+type information.
 
@@
 
+can only route messages within one DSP. No notion of remote DSP
+components.
+
+\subsection{OSGi}
+
+OSGi lends itself naturally to implement the DSP.
+
+DSP can be implemented in different ways, but OSGi is a natural match.
+DSP Components are mapped to bundles. DSP itself is a bundle.
+
 \section{NetBEAMS}
 
-In the following we present a practical application the the DSP
+In the following we present a practical application of the DSP
 presented in the previous section. We describe a deployment of the DSP
@@
 Monitoring System) and allows access to sensor equipment deployed
-throughout the San Francisco Bay Area. Figure \ref{FIG_NETBEAMS}
-provides an end-to-end overview of the NetBEAMS architecture. NetBEAMS
-is built using multiple installtion of the DSP. One DSP installation
-is collocated with each sensor and an additional DSP is running on the
-backend server.
+throughout the San Francisco Bay Area.  Our goal is to address the
+problem of data collection from remote sensors by using the DSP to
+automatically interrogate sensors at the remote field locations and
+transmit the collected measurements over long distances to a backend
+server where the data is made available to potential users in near
+real-time. This end-to-end system will be a ready-to-use turnkey
+solution that is adaptable to a wide variety of sensor types and
+requires no programming expertise to operate. Configuration of the
+system will be possible when the sensor is deployed in the field via a
+web-based interface, eliminating the need for extra site visits by
+technicians between service cycles to change systems functionalities
+(e.g. sampling rate). We present the NetBEAMS infrastructure by
+discussing its hardware and software components.
+
+\subsection{Hardware}
+
+Figure \ref{FIG_NETBEAMS} provides an end-to-end overview of the
+NetBEAMS architecture. NetBEAMS is built using multiple installation
+of the DSP and by providing special purpose components for the various
+tasks. One DSP installation is collocated with each sensor and an
+additional DSP is running on the backend server. Because of the remote
+locations where each individual sensor is deployed in the field and
+the sparse nature of the network, the resulting topology is a tree
+structure of depth 1.
 
@@
 
-To describe NetBEAMS is to describe the specific DC and DP we have
-implemented. On the far left of Figure \ref{FIG_NETBEAMS} is the
-actual sensor; in our case the aforementioned Seabird sensor.
-Collocated with the Seabird is an embedded system. Here we make use of
-a platform called Gumstix.
+NetBEAMS supports environmental sensor such as the YSI Sonde
+introduced in Section \ref{SEC_BACKGROUND}. In order to access its
+sensor readings, we collocate it with an off-the-shelf embedded
+computing platform called Gumstix (REF) (far left in Figure
+\ref{FIG_NETBEAMS}). The Gumstix is a single-board computers with 256
+Mb of flash and 256 Mb of SDRAM, and running a 600 MHz TI OMAP 3503
+processor. Additional features can be added to the motherboard with
+expansion cards connected via on-board buses. The motherboards draw
+less than 250 mA @4V at 400 MHz and less than 50 mA while idling,
+waiting for input.  Gumstix runs Linux 2.6 with the BusyBox utilities,
+and use the OpenEmbedded build environment to provide a complete Linux
+environment and a large range of Linux applications. We use the JamVM
+(REF) implementation of Sun Microsystem's virtual machine in order to
+run the DSP
 
-The Gumstix is a single-board computers with 256 Mb of flash and 256
-Mb of SDRAM, and running a 600 MHz TI OMAP 3503 processor. Additional
-features can be added to the motherboard with expansion cards
-connected via on-board buses. The motherboards draw less than 250 mA
-@4V at 400 MHz and less than 50 mA while idling, waiting for input.
-Gumstix run Linux 2.6 with the BusyBox utilities, and use the
-OpenEmbedded build environment to provide a complete Linux
-environment and a large range of Linux applications. We use the JamVM
-implementation of Sun Microsystem's virtual machine to execute Java
-programs.
+We make use of the cellular phone network to establish a communication
+link with the backend server. The cellular phone network has the
+advantages of reaching well beyond telemetry options such as wireless
+networks and being a significantly lower cost solution than satellite
+communications. We make use of the Huawei E220. The E220 is a USB
+modem that supports HSDPA, UMTS, EDGE packet data services at maximum
+transmission rates of 3.6Mbps, 384kbps, and 236.8kbps respectively.
+HSDPA and UMTS operate at 2100MHz while GSM, GPRS, and EDGE operate at
+900, 1800, and 1900MHz. The modem connects to the Verdex Pro XM4
+motherboard via a mini USB interface (supporting USB 2.0). Linux
+kernels since 2.6.20 include support for E220 drivers.  This includes
+the linux kernel on the Gumstix which runs 2.6.21.  The Gumstix
+communicates with the modem using the PPP protocol and accesses the
+cellular network using a data plan from AT\&T.
 
-The Huawei E220 is a USB modem that supports HSDPA, UMTS, EDGE packet
-data services at maximum transmission rates of 3.6Mbps, 384kbps, and
-236.8kbps respectively. HSDPA and UMTS operate at 2100MHz while GSM,
-GPRS, and EDGE operate at 900, 1800, and 1900MHz. The modem connects
-to the Verdex Pro XM4 motherboard via a mini USB interface (supporting
-USB 2.0). Linux kernels since 2.6.20 include support for E220 drivers.
-This includes the linux kernel on the gumstix which runs 2.6.21.  The
-gumstix communicates with the modem using the PPP protocol and
-accesses the cellular network using a data plan from AT\&T.
+\subsection{YSI Sonde Component}
 
-\subsection{YSI Sonde DP}
+Our implementation of the DSP is based on the Open Source
+implementation of OSGi called Knopflerfish (REF). Since Sun
+Microsystems does not support its JDK for the Gumstix, we use JamVM
+(REF) as the Java virtual machine implementation and GNU Classpath
+(REF) for the runtime libraries. We have implemented various DSP
+Components that implement the functionality needed for NetBEAMS. On
+the far left of Figure \ref{FIG_NETBEAMS} is the actual sensor; in our
+case the aforementioned YSI Sonde. A YSI Sonde data procuder (1) is a DSP
+component that uses the RS232 to read the physical measurements from
+the device. This component converts the physical measurement into a
+DSP-specific message. The following XML Schema defines the YSI Sonde
+specific message that captures all measurement points mentioned in
+Table \ref{TAB_SONDE_MEASUREMENTS}:
 
@@
 
-\subsection{Wire transport DC/DP}
+The schema is derived from DSPMessage. As a payload it adds an
+unbounded sequence of sondeDataType. Each instance of this type
+captures one timestamped measure point. The YSI Sonde component can be
+configured via special DSP messages that define properties such as the
+sampling frequency. These property messages originate from the web
+frontend that we will explain in a subsequent section.
 
-\subsection{Web Managements}
+\subsection{Remote Communication}
+
+The YSI Sonde component will pass any messages it generates to the
+local DSP. Via the matcher configuration these messages are forwarded
+to the remote communication component (2). It is the responsibility of
+this component to upload the message to its remote peer component
+(3). Since we use a data channel of a cellular phone network, the
+remote communication component only uploads the buffered data in
+certain configurable time intervals.
+
+\subsection{Web Management}
+
+Web Interface. Management messages are regular
+DSPMessages. Asynchronous UI: UI may only build up after next
+communication cycle.
 
 \section{Conclusions and Outlook}
-
-Length (including bibliography): 0.5 pages.
 

### 2009-03-06 versions/v2/apps/osgi-bundles/test/SerialBundle/src/org/netbeams/test/serial/SerialBundle.java
Submitted test bundle which confirms successful asynchronous communication between the simulator and a bundle.
Tests the interval rate change from 5 seconds to 1 second: Pauses for the first 30 seconds while data is sent every 5 seconds. Then data is sent every 1 second.

git-svn-id: svn://localhost/netbeams/trunk@350 559edde2-a4d4-44e0-96fc-4035e40d7396

@@ -1,0 +1,27 @@
+package org.netbeams.test.serial;
+
+/**
+ * 
+ * @author Teresa L. Johnson <gamma.particle@gmail.com>
+ *
+ */
+public class SerialBundle extends Thread {
+
+	private static final String SERIAL_PORT = "/dev/ttyS0";
+	private SerialHandler sh;
+	
+	public SerialBundle() {
+		sh = new SerialHandler();
+	};
+	
+	public void run() {
+		try {
+			System.out.println("Serial connection established from bundle...");
+			sh.connect(SERIAL_PORT);
+		} catch (Exception e) {
+			System.out.println("ERROR: " + e.getMessage());
+			e.printStackTrace();
+		}
+	}
+	
+}

### 2009-03-06 versions/v2/apps/osgi-bundles/test/SerialBundle/src/org/netbeams/test/serial/SerialHandler.java

@@ -1,0 +1,58 @@
+package org.netbeams.test.serial;
+
+import gnu.io.CommPort;
+import gnu.io.CommPortIdentifier;
+import gnu.io.SerialPort;
+
+import java.io.InputStream;
+import java.io.OutputStream;
+
+/**
+ * 
+ * @author Teresa L. Johnson <gamma.particle@gmail.com>
+ *
+ */
+public class SerialHandler {
+
+	private static final int BAUD_RATE = 9600;
+	private InputStream in;
+	private OutputStream out;
+	private SerialReader sr;
+	private SerialWriter sw;
+		
+	public SerialHandler() {
+		super();
+	};
+	
+	public void connect (String portName) throws Exception {
+
+		CommPortIdentifier portIdentifier = CommPortIdentifier.getPortIdentifier(portName);
+		if (portIdentifier.isCurrentlyOwned()) {
+			System.err.println("Error: Port is currently in use");
+		} else {
+			CommPort commPort = portIdentifier.open(this.getClass().getName(),2000);
+			if (commPort instanceof SerialPort) {
+				
+				SerialPort serialPort = (SerialPort) commPort;
+	            serialPort.setSerialPortParams(BAUD_RATE,SerialPort.DATABITS_8,SerialPort.STOPBITS_1,SerialPort.PARITY_NONE);
+
+	            in = serialPort.getInputStream();
+	            out = serialPort.getOutputStream();
+
+	            sr = new SerialReader(in);
+	    		sw = new SerialWriter(out);
+	    		
+	    		Thread srThread = new Thread(sr);
+	    		Thread swThread = new Thread(sw);
+	    		
+	    		srThread.start();
+	    		swThread.start();
+	            
+	            System.out.println("Success: Connection established");
+	        } else {
+	            System.out.println("Error: Only serial ports are handled.");
+	        }
+		}		        
+	}
+	
+}

### 2009-03-06 versions/v2/apps/simulator/YSI/Sonde/src/org/netbeams/sim/ysi/SondeSerialReader.java
Made interval publicly accessible.
git-svn-id: svn://localhost/netbeams/trunk@353 559edde2-a4d4-44e0-96fc-4035e40d7396

@@ -11,8 +11,10 @@ import java.io.IOException;
 public class SondeSerialReader implements Runnable {
 	
 	private InputStream in;
+	private static String cmd;
 	
 	public SondeSerialReader (InputStream in) {
+		cmd = new String("");
 		this.in = in;
 	}
 
@@ -22,7 +24,23 @@ public class SondeSerialReader implements Runnable {
         int len = -1;
         try {
             while ((len = this.in.read(buffer)) > -1) {
-                System.out.print(new String(buffer,0,len));
+            	cmd = new String(buffer,0,len);
+            	char option = cmd.charAt(0);
+            	System.out.print(option);
+            	switch (option) {
+            		case '1':
+            			// set interval to 1s
+            			System.out.print(option);
+            			SondeSerialWriter.interval = 1000;
+            			break;
+            		case '2':
+            			// set interval to 2s
+            			System.out.print(option);
+            			SondeSerialWriter.interval = 2000;
+            			break;
+            		default :
+            			break;
+            	};
             }
         } catch (IOException e) {
             e.printStackTrace();

### 2009-03-06 versions/v2/apps/simulator/YSI/Sonde/src/org/netbeams/sim/ysi/SondeSerialWriter.java

@@ -10,10 +10,10 @@ import java.io.IOException;
  */
 public class SondeSerialWriter implements Runnable {
 
-	private long interval = 5000;
+	public static long interval = 5000;
 	private OutputStream out;
 	private SondeTestData sdt;
-	private StringBuffer strBuffer; 
+	private StringBuffer strBuffer;
 
 	
 	public SondeSerialWriter(OutputStream out) {

### 2009-03-11 docs/icccn2009/paper.tex
This is an intermediate commit: Added more citations to lit.bib and the package "url" to paper.tex. Also added citation references in the document itself.
git-svn-id: svn://localhost/netbeams/trunk@362 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 \usepackage{fancyvrb}
+\usepackage{url}
 
@@
 
-The YSI 6600EDS V2 Sonde is a water quality monitoring device that
+The YSI 6600EDS V2 Sonde \cite{Sonde01} is a water quality monitoring device that
 gathers water quality data, and operates in fresh, sea, or polluted
@@
 format used in the domain of environmental monitoring is OpenDAP
-(REF). Since the converter module accepts messages from the DSP, it
+\cite{opendap01}. Since the converter module accepts messages from the DSP, it
 acts in the role of a data consumer. Finally, a DSP Component
@@
 Figure \ref{FIG_NETBEAMS} provides an end-to-end overview of the
-NetBEAMS architecture. NetBEAMS is built using multiple installation
+NetBEAMS architecture. NetBEAMS is built using multiple installations
 of the DSP and by providing special purpose components for the various
@@
 
-NetBEAMS supports environmental sensor such as the YSI Sonde
+NetBEAMS supports an environmental sensor such as the YSI Sonde,
 introduced in Section \ref{SEC_BACKGROUND}. In order to access its
 sensor readings, we collocate it with an off-the-shelf embedded
-computing platform called Gumstix (REF) (far left in Figure
+computing platform called Gumstix \cite{gumstix01} (far left in Figure
 \ref{FIG_NETBEAMS}). The Gumstix is a single-board computers with 256
@@
 waiting for input.  Gumstix runs Linux 2.6 with the BusyBox utilities,
-and use the OpenEmbedded build environment to provide a complete Linux
+and uses the OpenEmbedded build environment to provide a complete Linux
 environment and a large range of Linux applications. We use the JamVM
-(REF) implementation of Sun Microsystem's virtual machine in order to
-run the DSP
+\cite{jamvm01} implementation of Sun Microsystem's virtual machine in order to
+run the DSP.
+
+The Gumstix connects to the RS232 serial port of the YSI Sonde via via a
+null-modem cable.
 
@@
 communications. We make use of the Huawei E220. The E220 is a USB
-modem that supports HSDPA, UMTS, EDGE packet data services at maximum
+modem that supports HSDPA, UMTS, and EDGE packet data services at maximum
 transmission rates of 3.6Mbps, 384kbps, and 236.8kbps respectively.
@@
 Our implementation of the DSP is based on the Open Source
-implementation of OSGi called Knopflerfish (REF). Since Sun
+implementation of OSGi called Knopflerfish \cite{knopflerfish01}. Since Sun
 Microsystems does not support its JDK for the Gumstix, we use JamVM
-(REF) as the Java virtual machine implementation and GNU Classpath
-(REF) for the runtime libraries. We have implemented various DSP
+\cite{jamvm01} as the Java virtual machine implementation and GNU Classpath
+\cite{classpath01} for the runtime libraries. We have implemented various DSP
 Components that implement the functionality needed for NetBEAMS. On
 the far left of Figure \ref{FIG_NETBEAMS} is the actual sensor; in our
-case the aforementioned YSI Sonde. A YSI Sonde data procuder (1) is a DSP
-component that uses the RS232 to read the physical measurements from
-the device. This component converts the physical measurement into a
-DSP-specific message. The following XML Schema defines the YSI Sonde
+case the aforementioned YSI Sonde. A YSI Sonde data producer (1) is a DSP
+component that uses the RS232 standard to read the physical measurements from
+the device over a serial port. This component converts the physical measurements 
+into a DSP-specific message. The following XML Schema defines the YSI Sonde
 specific message that captures all measurement points mentioned in

### 2009-03-12 docs/icccn2009/paper.tex
Updated the Hardware subsection.
git-svn-id: svn://localhost/netbeams/trunk@363 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 computing platform called Gumstix \cite{gumstix01} (far left in Figure
-\ref{FIG_NETBEAMS}). The Gumstix is a single-board computers with 256
-Mb of flash and 256 Mb of SDRAM, and running a 600 MHz TI OMAP 3503
-processor. Additional features can be added to the motherboard with
-expansion cards connected via on-board buses. The motherboards draw
-less than 250 mA @4V at 400 MHz and less than 50 mA while idling,
-waiting for input.  Gumstix runs Linux 2.6 with the BusyBox utilities,
-and uses the OpenEmbedded build environment to provide a complete Linux
-environment and a large range of Linux applications. We use the JamVM
-\cite{jamvm01} implementation of Sun Microsystem's virtual machine in order to
-run the DSP.
+\ref{FIG_NETBEAMS}).  The Gumstix is an ARM architecture and comes in various 
+hardware configurations.  A typical Gumstix configuration consists 
+of a motherboard and one or more expansion boards which connect to the 
+motherboard via on-board buses.  The motherboards draw less than 250 mA @4V 
+at 400 MHz and less than 50 mA while idling, waiting for input.
 
-The Gumstix connects to the RS232 serial port of the YSI Sonde via via a
-null-modem cable.
+In our project, we use the Verdex Pro XM4 motherboard with 64MB of SDRAM, 16MB of Flash, and 
+a 400MHz Marvell PXA270 with XScale processor.  We also use two expansion boards:
+the Console-VX which comes with three RS232 serial ports and the Netpro-VX which
+features one 10/100baseT ethernet port.  The Gumstix runs Linux 2.6 with the 
+BusyBox utilities, and uses the OpenEmbedded cross-compile build environment to 
+provide a complete Linux environment and a large range of Linux applications for
+the ARM architecture.  We use the JamVM \cite{jamvm01} implementation of Sun 
+Microsystem's virtual machine in order to run the DSP. The Gumstix communicates 
+with the YSI Sonde via an RS232 serial port using a null-modem cable since both 
+the Sonde and the Gumstix are DTE devices. The serial port configuration is:
+
+\begin{table}
+\caption{\label{tab:RS232_Config}RS232 Serial Port Configuration}
+\centering
+\begin{tabular}{l l}
+\hline
+Baud Rate&9600\\
+Data Bit&8\\
+Parity&None\\
+Handshake&None\\
+\hline
+\end{tabular}
+\end{table}
+
+\begin{table}
+\caption{\label{TAB_RS232_Pinout}RS232 Serial Pinout}
+\centering
+\begin{tabular} {|l|l|l|}
+\hline
+Wire Color	&Pin Description        &DB-9\\ 
+\hline
+Yellow 		&RS232 TX    &2\\
+Orange		&RS232 RX    &3\\  
+Green  		&Alarm       &----\\ 
+Grey			&RTS         &----\\
+Blue  		&CTS         &----\\
+Red			&+ 12V DC    &9\\
+Black			&GND         &5\\  
+Purple		&SDI-12      &----\\ 
+Bare			&Shield      &----\\
+\hline
+\end{tabular}
+\end{table}
 
@@
 unbounded sequence of sondeDataType. Each instance of this type
-captures one timestamped measure point. The YSI Sonde component can be
+captures one timestamped measure point.
+
+The YSI Sonde component can be
 configured via special DSP messages that define properties such as the

### 2009-03-12 docs/icccn2009/paper.tex
Updated the YSI Sonde Component subsection.
git-svn-id: svn://localhost/netbeams/trunk@364 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 the device over a serial port. This component converts the physical measurements 
-into a DSP-specific message. The following XML Schema defines the YSI Sonde
-specific message that captures all measurement points mentioned in
-Table \ref{TAB_SONDE_MEASUREMENTS}:
+into a DSP-specific message. The following XML Schema defines the YSI Sonde's
+specific message format which includes each measurement data point, described in
+Table \ref{TAB_SONDE_MEASUREMENTS}, and its type:
 
 \begin{code}
-<xs:schema ...>
-<xs:import schemaLocation="dsp-message.xsd"/>
+<?xml version="1.0" encoding="UTF-8"?>
+<xs:schema targetNamespace="http://www.netbeams.dsp.org/sondedata.xsd" xmlns:jaxb="http://java.sun.com/xml/ns/jaxb" jaxb:version="2.0" xmlns:mc="http://www.netbeams.dsp.org/message.xsd" xmlns:sndata="http://www.netbeams.dsp.org/sondedata.xsd" xmlns:xs="http://www.w3.org/2001/XMLSchema">
+
+    <xs:import namespace="http://www.netbeams.dsp.org/message.xsd" schemaLocation="message-content.xsd"/>
+
+    <xs:annotation><xs:appinfo>
+    <jaxb:globalBindings>
+      <jaxb:javaType name="java.util.Calendar" xmlType="xs:date"
+        parseMethod="javax.xml.bind.DatatypeConverter.parseDate"
+        printMethod="javax.xml.bind.DatatypeConverter.printDate"
+        />
+      <jaxb:javaType name="java.util.Calendar" xmlType="xs:time"
+        parseMethod="javax.xml.bind.DatatypeConverter.parseDate"
+        printMethod="javax.xml.bind.DatatypeConverter.printDate"
+        />
+    </jaxb:globalBindings>
+    </xs:appinfo></xs:annotation>
 
@@
         <xs:complexContent>
-      <xs:extension base="DSPMessage">
-        <xs:sequence maxOccurs="unbounded"
-                     minOccurs="1">
+            <xs:extension base="mc:AbstractMessageContent">
+                <xs:sequence maxOccurs="unbounded" minOccurs="1">
                     <xs:element name="sondeData"
-                      type="sondeDataType"/>
+                        type="sndata:sondeDataType" />
                 </xs:sequence>
+                <xs:attribute name="date" type="xs:date"></xs:attribute>
+                <xs:attribute name="time" type="xs:time"></xs:attribute>
             </xs:extension>
@@
         <xs:sequence>
-    <xs:element name="Date" type="xs:date"/>
-    <xs:element name="Time" type="xs:time"/>
                 <xs:element name="Temp" type="xs:float"/>
                 <xs:element name="SpCond" type="xs:float"/>
-    <!-- ... -->
+                <xs:element name="Cond" type="xs:float"/>
+                <xs:element name="Resist" type="xs:float"/>
+                <xs:element name="Sal" type="xs:float"/>
+                <xs:element name="Press" type="xs:float"/>
+                <xs:element name="Depth" type="xs:float"/>
+                <xs:element name="pH" type="xs:float"/>
+                <xs:element name="phmV" type="xs:float"/>
+                <xs:element name="ODOSat" type="xs:float"/>
+                <xs:element name="ODOConc" type="xs:float"/>
+                <xs:element name="Turbid" type="xs:float"/>
+                <xs:element name="Battery" type="xs:float"/>
         </xs:sequence>
@@
 
-The schema is derived from DSPMessage. As a payload it adds an
-unbounded sequence of sondeDataType. Each instance of this type
-captures one timestamped measure point.
+The schema describes two Java classes, sondeDataType and SondeDataMessage, 
+which are used to package the Sonde's incoming data into a DSP message for delivery 
+to a remote consumer.  The SondeDataMessage class extends the AbstractMessageContent 
+class by adding a new type to the message payload, SondeDataType.  Each time the Sonde
+transmits its payload of measurement data during a sampling interval, the data is converted 
+into a SondeDataType which is then encapsulated within SondeDataMessage and then sent 
+to a remote consumer.
 
-The YSI Sonde component can be
-configured via special DSP messages that define properties such as the
-sampling frequency. These property messages originate from the web
-frontend that we will explain in a subsequent section.
+The YSI Sonde component can also be configured via special DSP messages, to direct the 
+Sonde to change its sampling frequency.  These property messages originate from the web 
+frontend that we will explained in a subsequent section.
 

### 2009-03-12 docs/icccn2009/paper.tex
ICCCN 2009 Paper update, sections Message, Matcher, Broker, YSI sonde
- Section Message has updates regarding the types of 
  messages, and others;
  
- Adding images of XML schema to clean up a bit the
  presentation of the paper. This is just an experimental
  version, and can be replaced back; YSI XML Schema
  code was replaced by the XML Schema type image;
  
- Creating the match_config.xsd, the XML Schema that
  validates the Matcher rules. Instead of using code,
  adding the XML Schema type image for Matching Rule.
 
- Adding a DSP Broker section, since it's one critical
  component for

--- 2009-03-12 docs/icccn2009/paper.tex
@@
 \centering
-\epsfig{file=dsp, width=7cm}
+\epsfig{file=dsp-platform, width=7cm}
 \caption{\label{FIG_DSP} Data Sensor Platform (DSP).}
@@
 Message passing is the main paradigm for exchanging information within
-a DSP. A data producer sends a message with data consumers accept
-messages. We make use of XML technologies to define the format of a
-message. Specifically, the common header of all messages are defined
-via the following XML Schema:
+a DSP. After producing data, DP component wraps that data in the 
+body of a message, so that DC components can receive the message and consume 
+the data on the message's body. When it comes to passing messages to a remote
+DSP instane, messages are serialized in XML \cite{xml2000} and automatically 
+deserialized for DC located in the remote DSP. In order to serialize and
+deserialize DSP messages, the DSP infrastructure uses the XML Schema 
+standard \cite{xml-schema2004} to generate and validate the DSP message. Figure
+\ref{FIG_DSP_MESSAGE} defines the structure of a general DSP Message.
 
-\begin{code}
-<xs:schema ...>
-  <xs:complexType name="DSPMessage">
-    <xs:element name="Header" type="Header"/>
-  </xs:complexType>
+\begin{figure}[!htb]
+ \centering
+ \epsfig{file=dsp-message, width=9cm, height=4cm}
+ \caption{\label{FIG_DSP_MESSAGE} Structure of a DSP Message.}
+\end{figure}
 
-  <xs:complexType name="Header">
-    <xs:sequence>
-      <xs:attribute name="messageID"/>
-      <xs:attribute name="originatingDP"/>
-      <!-- ... -->
-    </xs:sequence>
-  </xs:complexType>
-</xs:schema>
-\end{code}
+The above XML Schema is defined for an ``Abstract Message'' type, whose
+structure serves as a base schema for all different types of messages. Any
+message will contain two basic attributes: ``messageID'' and ``ContentType''. 
+The former uniquely identifies a message and the latter the data in the body of
+the message. Complementary identification information such as the message 
+producer and consumer are located in the header of the message, which are
+defined by the schema type ``ComponenIdentifier''. A ``ComponentIdentifier''
+carries out information about the component name and its physical host address.
+Finally, the body of the message may contain any data type defined by the component 
+developer, and it can be interpreted as the DSP payload.
 
-The above XML Schema DSPMessage serves as a base schema for all
-messages and itself defines information typically contained in the
-header of a protocol data unit. Amongst others, the header contains a
-unique message ID (line 8) and the originating DP (line 9). Using the
-XML Schema, we use JAXB to generate a type safe API for handling
-instances of DSPMessage. Different messages with different payloads
-can be defined by creating schemas that derive from DSPMessage. While
-inside the DSP all messages are of type DSPMessage, specific DP and DC
-can distinguish between different message types based on their runtime
-type information.
+Although XML schema is the de-facto standard to bind and validate xml data to
+its type, the architectural infrastructure of DSP uses an XML Data Bindining
+Technology \cite{xml-dbind} called JAXB \cite{xml-jaxb}. It is used to generate
+a type safe API for handling in-memory object graphs from the XML instances of 
+DSP Messages and the content of its data payload. Moreover, taking advantage 
+of the data type polimorphirsm of XML schema data types - concept borrowed from 
+Object-Oriented Design \cite{Rumbaugh1991} - we defined different types of 
+messages in order to allow specific DP and DC interpret the payload of any 
+given message solely based on their runtime type, as follows:
+
+\begin{itemize}
+  \item Action Message;
+  \item Event Message;
+  \item Create Message;
+  \item Insert Message;
+  \item Update Message;
+  \item Delete Message;
+  \item Query Message;
+  \item Measure Message.
+\end{itemize}
+
+Not only the message structure is defined by XML schemas, but also the body
+content is and may be of any user-defined type. Therefore, the data
+defined to be sent from any given DP to any DC can be also in XML format or
+in-memory. An example illustrating a DSP message is given in the YSI section.
+
+\subsection{Broker}
+
+As soon as a DP produces data and wraps it up in a DSP Message, the DP only
+interact with the DSP platform via the DSP Broker by making the call
+``Broker.deliver(DSP Message)''. In this way, DSP broker is responsible for
+delivering the produced message to DSP consumers DCs, as well as verifying
+whether the delivery of the message is going to be done locally or remotely.
+
+\begin{itemize}
+ \item Local delivery of messages: it uses local object reference to deliver the
+DSP message;
+ \item Remote delivery of messages: it uses any DSP component that is
+specialized in remote delivery defined as ``Gateway DSP Component''.
+\end{itemize}
+
+The selection of the DCs for a given message to where the Broker will deliver
+messages or use a gateway component is determined by an object called Matcher.
 
@@
 
-can only route messages within one DSP. No notion of remote DSP
-components.
+The Matcher is an object responsible to find the possible set of DCs for a given
+DSP message, and its ``deterministic intelligence'' relies on matching rules
+defined by the DSP administrator. In this way, a set of matching rules is define
+in XML and stored at the DSP run-time directory, where each matching rule is
+uniquely identified by a rule ID and determines the following (see
+figure \ref{FIG_MATCHER_RULE}):
+
+\begin{figure}[!htb]
+ \epsfig{file=dsp-match-rule, width=9cm, height=4cm}
+ \caption{\label{FIG_MATCHER_RULE} Definition of Matcher Rule.}
+\end{figure}
+
+\begin{itemize}
+ \item Match Criteria: given the DSP message delivered to the Broker, which
+component(s) produced that message and in which node? It defines the name of
+the component type of the producer and its physical location (IP address);
+ \item Match Target: given the DSP message, which component(s) are going
+to consume the message, and in which node? It defines the name of the component
+type of the consumer, and optionally defines a gateway component.
+\end{itemize}
+
+\subsection{Matcher replies to Broker}
+
+When analyzing the matching rules, the Matcher only selects the rules that
+satisfy the Matching Criteria and forwards them back to the Broker. Once the
+broker has the matching rules, it determines whether the delivery is done
+directly to a component defined on the Match Target, or if the message is
+delivered to the optional gateway. We have defined special gateway DSP
+components that is used to transport the DSP Messages, serialized in
+XML, to any remote DSP instance.
 
@@
 the device over a serial port. This component converts the physical measurements 
-into a DSP-specific message. The following XML Schema defines the YSI Sonde
-specific message that captures all measurement points mentioned in
-Table \ref{TAB_SONDE_MEASUREMENTS}:
+into a DSP-specific message. The following XML Schema on figure \ref{FIG_DSP_YSI_DATA} 
+defines the YSI Sonde's specific message format which includes each measurement data point, described in
+Table \ref{TAB_SONDE_MEASUREMENTS}, and its type:
 
-\begin{code}
-<xs:schema ...>
-<xs:import schemaLocation="dsp-message.xsd"/>
+\begin{figure}[!htb]
+\centering
+\epsfig{file=dsp-message-body-ysi, width=8cm, height=7cm}
+\caption{\label{FIG_DSP_YSI_DATA} Structure of the YSI Component Data.}
+\end{figure}
 
-<xs:element name="SondeDataMessage">
-  <xs:complexType>
-    <xs:complexContent>
-      <xs:extension base="DSPMessage">
-        <xs:sequence maxOccurs="unbounded"
-                     minOccurs="1">
-          <xs:element name="sondeData"
-                      type="sondeDataType"/>
-        </xs:sequence>
-      </xs:extension>
-    </xs:complexContent>
-  </xs:complexType>
-</xs:element>
+The schema describes two Java classes, sondeDataType and SondeDataMessage, 
+which are used to package the Sonde's incoming data into a DSP message for delivery 
+to a remote consumer.  The SondeDataMessage class extends the AbstractMessageContent 
+class by adding a new type to the message payload, SondeDataType.  Each time the Sonde
+transmits its payload of measurement data during a sampling interval, the data is converted 
+into a SondeDataType which is then encapsulated within SondeDataMessage and then sent 
+to a remote consumer.
 
-<xs:complexType name="sondeDataType">
-  <xs:sequence>
-    <xs:element name="Date" type="xs:date"/>
-    <xs:element name="Time" type="xs:time"/>
-    <xs:element name="Temp" type="xs:float"/>
-    <xs:element name="SpCond" type="xs:float"/>
-    <!-- ... -->
-  </xs:sequence>
-</xs:complexType>
-</xs:schema>
-\end{code}
-
-The schema is derived from DSPMessage. As a payload it adds an
-unbounded sequence of sondeDataType. Each instance of this type
-captures one timestamped measure point.
-
-The YSI Sonde component can be
-configured via special DSP messages that define properties such as the
-sampling frequency. These property messages originate from the web
-frontend that we will explain in a subsequent section.
+The YSI Sonde component can also be configured via special DSP messages, to direct the 
+Sonde to change its sampling frequency.  These property messages originate from the web 
+frontend that we will explained in a subsequent section.
 

### 2009-03-12 docs/icccn2009/paper.tex
ICCCN 2009 Paper Update;
Integrating Kleber's review. All changes to this
revision were submitted by Kleber via email.

git-svn-id: svn://localhost/netbeams/trunk@367 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 
-Should describe DSP architecture. Only mention OSGi in last subsection
-and explain that components can be mapped to bundles. Currently no
-distinction between Platform and Framework. Necessary?
+Sensor networks, especially of environmental sensors, are normally deployed
+over distributed locations where communication infrastructure are not always 
+reliable or, at least, not countinuously available. Additionally, different
+sensors and communication protocols might be available in throughout the network.
+
+As the network grows in size and complexity scalability, reliabilty and maintainability
+become critical issues. Additional components 
+might be added to the network, the network configuration might be altered or 
+new versions of components might be deployed. The sensor network must be able to 
+support such adjustments without significant overhead for network administrators and
+the components in the network.
+
+From system development and integration point of view, interfaces among components and
+between components and the network should be simple. Complex interaction can be 
+constructed on top of these simple interfaces ( a good example of this strategy is the 
+world-wide-web).
+
+In the following we discuss how our proposed architechure addresses those issues.
+
+\subsection{Architecture}
 
@@
 a DSP is a message. DSP Components can exchange messages of different
-types. We distinguish between Data Producers (DP) and Data Consumers
-(DP) within the scope of a DSP. DP and DC can be seen as roles that a
+types. Actually, all interactions between DSP Components are implemented
+through messages yielding simple but powerful communication instructure.
+We distinguish between Data Producers (DP) and Data Consumers
+(DC) within the scope of a DSP. DP and DC can be seen as roles that a
 DSP Component may have. It is possible for a DSP Component to act in
-the role of DC and DP at the same time.
+the role of DC and DP at the same time. Messages in DSPs are routed based 
+on rules. Rules can be statically or dynamically configured. Rules can use 
+different source information when routing a message: producers or consumers of the message, 
+the message itself or specific configuration provided by the administrators.
 
@@
 effectively becomes a DSP Component that can be deployed at runtime. A
-component is identified by its type. Only one component of a certain
-type per DSP is permissible. A string serves as a type identifier that
-much be unique across all DSP Components. Upon deploying a DSP
-Component, an initialization function is called (line 4). Once a
+component is identified by its type. Only one component of a certain type per DSP 
+is permissible. A type identifier much be unique across all DSP Components in DSP. 
+Upon deploying a DSP Component, an initialization function is called (line 4). Once a
 component is initialized, it may be started (line 5) and stopped (line
-6) multiple time (TODO: SHOULDN'T THERE BE ALSO A WAY TO UNDEPLOY A
-COMPONENT?) Whenever the DSP wishes to pass a message (that was
+6) multiple time. Whenever the DSP wishes to pass a message (that was
 generated by some DP), it invokes the \texttt{deliver()} method (ine
 7). The message structure will be explained in the following section.
+One important feature of the DSP is the ability to redeploy (maybe with a new version) at 
+runtime. This allows DSPs to be updated remotely. This process is managed through
+messages sent to a special DSP Component which represents the DSP itself.
 

### 2009-03-12 docs/icccn2009/paper.tex
Adding the Remote Communication section. It must be finished up...
git-svn-id: svn://localhost/netbeams/trunk@368 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 
+\begin{figure*} 
+\centering
+\epsfig{file=netbeams, width=14cm}
+\caption{\label{FIG_NETBEAMS} NetBEAMS architecture.}
+\end{figure*}
+
 In the following we present a practical application of the DSP
@@
 
-\begin{figure*}
-\centering
-\epsfig{file=netbeams, width=14cm}
-\caption{\label{FIG_NETBEAMS} NetBEAMS architecture.}
-\end{figure*}
-
 NetBEAMS supports an environmental sensor such as the YSI Sonde,
@@
 
-\begin{table}
+\begin{table}[!htb]
 \caption{\label{tab:RS232_Config}RS232 Serial Port Configuration}
@@
 
-The YSI Sonde component will pass any messages it generates to the
-local DSP. Via the matcher configuration these messages are forwarded
-to the remote communication component (2). It is the responsibility of
-this component to upload the message to its remote peer component
-(3). Since we use a data channel of a cellular phone network, the
-remote communication component only uploads the buffered data in
-certain configurable time intervals.
+After producing the SondeDataContainer data, the YSI Sonde component will wrap
+it up in a DSP Message Message and pass it to the DSP Broker for its delivery.
+However, our use case scenario is that the message must go from the local
+sensor device to the server. For this reason, we have developed a pair of
+client and server DSP components specialized in transferring and receiving 
+serialized objects throught the communication channel using the HTTP protocol
+\cite{RFC2068}.
+
+\begin{itemize}
+  \item DSP Wire Transport Client: takes a collection of DSP messages and
+  transmit them using an HTTP client API written in JAVA \cite{http-client}. As
+  a non-functional requirement, a DSP administrator can set the interval in
+  which this component will transmit messages to other components;
+  \item DSP Wire Transport Server: is implemented using the Java Servlets
+  technology \cite{java-servlets}, and is used to receive the collection of DSP
+  messages transmitted, and potentially send DSP messages back to the sensor
+  station. 
+\end{itemize}
+
+One of the non-functional requirements for the DSP is to maximize the
+utilization of the communication channel. For this specific reason, we opted to
+add the type Messages Container, which contains a collection of DSP messages,
+and specific identifiers of the hosting sensor. Figure
+\ref{FIG_DSP_MSGS_CONTAINER} contains the structure of the DSP Messages
+Container.
+
+\begin{figure}[!htb]
+\centering
+\epsfig{file=dsp-messages-container, width=8cm}
+\caption{\label{FIG_DSP_MSGS_CONTAINER} The structure of Messages Container.}
+\end{figure}
+
+Some remarks regarding the Messages Container and it's implementation are as
+follows:
+
+\begin{itemize}
+  \item Container with or without Messages Container: an instance of
+  MessagesContainer may or may not have DSP messages;
+  \item Container may be used to acknowledge messages using a similar strategy
+  as the Slide Windowing protocol \cite{slide-window}. 
+\end{itemize}
+
+\subsection{Sending Messages Container}
+
+It's important to note that, at this point, the DSP Broker had already checked
+with the Matcher which are the DSP components that will be receive a copy of
+the message in question. The matcher had specific instructions to use the DSP
+Wire Transport Client component as the default ``Proxy'' or default gateway, if
+you prefer.
+
+By building a MessagesContainer instance, the Wire Transport Client adds all
+the messages addressed to the ``destinationHost'', with specific UUID of
+the component (2).
+
+\subsection{Receiving Messages Container}
+
+After the DSP server receives the message, it directly delivers to the
+local DSP Broker all the messages contained in the Messages Container server.
+In addition to that, the Messages Container might also include an attribute
+called acknowledgeUntil, which determines the next actions that will occur in
+the system.
 

### 2009-03-12 docs/icccn2009/paper.tex
Adding
git-svn-id: svn://localhost/netbeams/trunk@369 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 
-After the DSP server receives the message, it directly delivers to the
-local DSP Broker all the messages contained in the Messages Container server.
-In addition to that, the Messages Container might also include an attribute
-called acknowledgeUntil, which determines the next actions that will occur in
-the system.
+After the DSP Wire Transport Server receives the Messages Container (3), it
+unmarshalls the messages and forward them to the local DSP Broker.
+
+The reception of those messages will ALWAYS depend on the local DSP Broker and
+its combination with the Matcher. The steps (4), (5) and (6) are the
+actual reception of DSP messages. Container might also include an attribute
+called acknowledgeUntil, which determines the next actions that will occur in the system.
 

### 2009-03-12 docs/icccn2009/paper.tex
Abstract/Introduction
git-svn-id: svn://localhost/netbeams/trunk@370 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 \begin{abstract}
-  Abstract.
+  In this paper we describe a sensor network for environmental
+  monitoring based on highly specialized, off-the-shelf sensing
+  devices.  Environmental monitoring depends on the reliable
+  collection of measurements from remotely deployed sensors and the
+  rapid transfer of those measurements to data centers. Typically,
+  sensors are deployed in out-of-the-way locations, where physical
+  access is limited and connections for telemetry are poor or
+  nonexistent. Maintaining the data stream from these sensors places
+  demands on human resources, requiring site visits to service the
+  sensors and to download data stored on internal memory.  The work
+  described in this paper is a collaborative effort between the
+  Computer Science Department at San Francisco State University and
+  the Romberg Tiburon Center. The result of this collaboration is an
+  environmental sensor network deployed in the San Francisco Bay. It
+  provides means to program and interrogate sensors at the remote
+  field locations and to transmit collected measurements to a remote
+  database.
 \end{abstract}
@@
 
-Mention the nature of our sensor network: highly specialized sensors,
-sparse topology, remote locations, long range communication.
-\cite{roemer:2004}
-
----
+A wireless sensor network consists of spatially distributed autonomous
+devices using sensors to cooperatively monitor physical or
+environmental conditions at different locations \cite{roemer:2004}.
+Much research effort is focused on building wireless ad-hoc networks
+where each sensor node participates in a multi-hop routing algorithm.
+One the earliest examples of a wireless sensor network is the
+SmartDust project \cite{smartdust:2001} where one objective was to
+create autonomous sensing and communication in a cubic
+millimeter-sized device.  While most wireless sensor networks focus on
+small sensing devices such as the SmartDust mote, there exist a wide
+variety of off-the-shelf commercially available sensors that are used
+by environmental researchers. In contrast to the small sensing
+devices, sensors for environmental monitoring are typically on the
+order of one to two feet in length weighing several pounds. Our
+focus is on building an end-to-end infrastructure for environmental
+monitoring using specialized sensing devices.
 
@@
 measurements from remotely deployed sensors and the rapid transfer of
-those measurements to data centers where processing and analysis may
-take place. Typically, sensors are deployed in out-of-the-way
-locations, where physical access is limited and connections for
-telemetry are poor or nonexistent. Maintaining the data stream from
-these sensors places demands on human resources, requiring site visits
-to service the sensors and to download data stored on internal memory.
-Consequently, there is a lag of days to months between the time the
-sensors perform the measurements and the time the measurements become
-available. Advances in battery life and anti-fouling technology have
-extended service cycles for field sensors, with the unwanted result of
-further delaying access to monitoring data.  In addition, the
-dependence on field site visits to alter of sensor characteristics, 
-such as sampling rate, prevents adjustment in monitoring strategy in
-response to rapidly developing events (e.g. oil spills, harmful algal
-blooms).
+those measurements to data centers. Typically, sensors are deployed in
+out-of-the-way locations, where physical access is limited and
+connections for telemetry are poor or nonexistent. Maintaining the
+data stream from these sensors places demands on human resources,
+requiring site visits to service the sensors and to download data
+stored on internal memory.  Consequently, there is a lag of days to
+months between the time the sensors perform the measurements and the
+time the measurements become available. Advances in battery life and
+anti-fouling technology have extended service cycles for field
+sensors, with the unwanted result of further delaying access to
+monitoring data.  In addition, the dependence on field site visits to
+alter of sensor characteristics, such as sampling rate, prevents
+adjustment in monitoring strategy in response to rapidly developing
+events (e.g. oil spills, harmful algal blooms).
 
-To address these problems of data collection from remote
-sensors, we propose to develop an automated end-to-end system based on
-proven technology. It provides means to program and interrogate
-sensors at the remote field locations and to transmit collected
-measurements over long distances to data collections centers, where the
-data is rapidly processed, archived, and made available to potential
-users in near real-time.
+The work described in this paper is a collaborative effort between the
+Computer Science Department at San Francisco State University and the
+Romberg Tiburon Center (RTC), a research institute focusing on the
+understanding of complex marine and estuarine environments. The result
+of this collaboration is NetBEAMS (Networked Bay Environmental
+Assessment Monitoring System), an environmental sensor network
+deployed in the San Francisco Bay. NetBEAMS provides means to program
+and interrogate sensors at the remote field locations and to transmit
+collected measurements to a database, where the data is rapidly
+processed, archived, and made available to potential users in near
+real-time. In the following we give an overview of the NetBEAMS
+architecture and implementation.  Section \ref{SEC_BACKGROUND}
+describes a typical environmental sensor used by the RTC. Section
+\ref{SEC_DSP} describes our Data Sensor Platform (DSP) that allows us
+to efficiently build up an end-to-end environmental sensor network.
+Section \ref{SEC_NETBEAMS} describes NetBEAMS; a practical application
+of the DSP. In Section \ref{SEC_CONCLUSION} we provide a conclusion
+and outlook.
 
@@
 \section{Data Sensor Platform}
+\label{SEC_DSP}
 
@@
 In the following we discuss how our proposed architechure addresses those issues.
-
-\subsection{Architecture}
 
@@
 \section{NetBEAMS}
+\label{SEC_NETBEAMS}
 
@@
 \centering
-\epsfig{file=dsp-messages-container, width=8cm}
+%\epsfig{file=dsp-messages-container, width=8cm}
 \caption{\label{FIG_DSP_MSGS_CONTAINER} The structure of Messages Container.}
@@
 \section{Conclusions and Outlook}
+\label{SEC_CONCLUSION}
 

### 2009-03-13 docs/icccn2009/paper.tex
Revised background section
git-svn-id: svn://localhost/netbeams/trunk@372 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 
-Researchers in the geosciences rely on sensor data collected by highly
-specialized sensors such as the Seabird or YSI. These sensors are
-usually deployed in remote locations that not only make their
-maintenance difficult, but also access to their measurements.
-Typically, the sensors are autonomous in the sense that they can run
-for a certain time on battery power and store the measurement data in
-internal memory. Only when the sensors are serviced, the measurement
-data, which was collected since the last servicing, is uploaded to a 
-portable storage device. From there, the data can be uploaded to
-server. Notice that this manual procedure to retrieve sensor data is not only
-error prone, but also results are in a serious time lag between the time
-the sensor performs the measurement and when it becomes available for
-further processing. For many applications, however, it would be beneficial to
-have near-realtime access to the sensor measurement. In this way, being able 
-to access data when they become available on the sensor device would also
-reduce the maintenance overhead.
-
----
-
-The YSI 6600EDS V2 Sonde \cite{Sonde01} is a water quality monitoring device that
-gathers water quality data, and operates in fresh, sea, or polluted
-water. Some of the measurements that the YSI is capable of are
-conductivity, temperature, chloride, ammonium, nitrate, turbidity, and
-chlorophyll.  It is under 55 cm in length, 8.9 cm in diameter, and
-weighs approximately 3.18 kg. It operates at temperatures between -5
-to +50 deg C and at depths up to 200 meters.  It uses 8 C-size
-Alkaline Batteries or External 12 VDC. Battery life can last up to 75
-days depending on sensor configuration. The Sonde's bulkhead contains
-multiple pin connectors to support sensor probes for measuring the
-following:
+Environmental researchers typically use standard, off-the-shelf
+sensing devices for their purposes. These sensing devices are at the
+opposite spectrum of miniature sensors such as SmartDust. One such
+device is the YSI 6600EDS V2 Sonde \cite{Sonde01}. It is a water
+quality monitoring device that gathers water quality data, and
+operates in fresh, sea, or polluted water. Some of the measurements
+that the YSI is capable of are conductivity, temperature, chloride,
+ammonium, nitrate, turbidity, and chlorophyll.  It is under 55 cm in
+length, 8.9 cm in diameter, and weighs approximately 3.18 kg. It
+operates at temperatures between -5 to +50 deg C and at depths up to
+200 meters.  It uses 8 C-size Alkaline Batteries or External 12 VDC.
+Battery life can last up to 75 days depending on sensor configuration.
+The Sonde's bulkhead contains multiple pin connectors to support
+sensor probes for measuring parameters as shown in Table
+\ref{TAB_SONDE_MEASUREMENTS}.
 
@@
 
----
+We provide these details to show that sensing devices such as the YSI
+are highly specialized devices with a unit price of several thousand
+dollars. These sensors are usually sparsely deployed in remote
+locations that not only make their maintenance difficult, but also
+access to their measurements.  Typically, the sensors are autonomous
+in the sense that they can run for a certain time on battery power and
+store the measurement data in internal memory. Only when the sensors
+are serviced, the measurement data, which was collected since the last
+servicing, is uploaded to a portable storage device. From there, the
+data can be uploaded to server. This manual procedure to retrieve
+sensor data is not only error prone, but also results are in a serious
+time lag between the time the sensor performs the measurement and when
+it becomes available for further processing. For many applications,
+however, it would be beneficial to have near-realtime access to the
+sensor measurement. We identify the following requirements that an
+end-to-end infrastructure supporting devices such as the YSI should
+offer:
 
-Start with some related work. Who else has done an OSGi-based
-infrastructure for sensor networks? Kleber: I believe you have a few
-references. This section should be a combination of a use case (using
-the RTC and their YSI sonde as an example) and a requirements
-analysis: requirements such as remote management and monitoring
-capabilities, time-delayed communication, etc. In this section should
-be no mentioning of DSP. Length: 1 page.
+\begin{itemize}
+\item Allow near-realtime access to sensor measurements.
+\item Allow remote servicing that include changing of configuration
+  parameters as well as software upgrades.
+\item Support cost-effective long distance communication link.
+\item Can be deployed in remote locations in unprotected outdoor
+  conditions.
+\end{itemize}
+
+In the next section we describe a software architecture that provides
+a foundation to fulfill these requirements.
 
@@
 \centering
-%\epsfig{file=dsp-messages-container, width=8cm}
+\epsfig{file=dsp-messages-container, width=8cm}
 \caption{\label{FIG_DSP_MSGS_CONTAINER} The structure of Messages Container.}

### 2009-03-13 docs/icccn2009/paper.tex
Edited section on DSP
git-svn-id: svn://localhost/netbeams/trunk@373 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 
-Sensor networks, especially of environmental sensors, are normally deployed
-over distributed locations where communication infrastructure are not always 
-reliable or, at least, not countinuously available. Additionally, different
-sensors and communication protocols might be available in throughout the network.
-
-As the network grows in size and complexity scalability, reliabilty and maintainability
-become critical issues. Additional components 
-might be added to the network, the network configuration might be altered or 
-new versions of components might be deployed. The sensor network must be able to 
-support such adjustments without significant overhead for network administrators and
-the components in the network.
-
-From system development and integration point of view, interfaces among components and
-between components and the network should be simple. Complex interaction can be 
-constructed on top of these simple interfaces ( a good example of this strategy is the 
-world-wide-web).
-
-In the following we discuss how our proposed architechure addresses those issues.
+Sensor networks, especially of environmental sensors, are normally
+deployed over a geographically wide area where the communication
+infrastructure is not always reliable or, at least, not countinuously
+available. As the sensor network grows in size and complexity
+scalability, reliabilty and maintainability become critical issues.
+Additional components might be added to the network, sensor
+configuration might be altered, or new versions of components might be
+deployed. The sensor network must be able to support such adjustments
+without significant overhead for maintainers.  From system
+development and integration point of view, interfaces among components
+and between components and the network should be simple. Complex
+interaction can be constructed on top of these simple interfaces.  In
+the following we discuss how our proposed architechure addresses those
+issues.
 
@@
 
-Figure \ref{FIG_DSP} depicts the basic architecture of the
-DSP. We refer to the plug-and-play modules as DSP Components. DSP
-Components are self-contained modules that can be added and removed to
-a DSP installation at runtime. The basic communication paradigm within
-a DSP is a message. DSP Components can exchange messages of different
-types. Actually, all interactions between DSP Components are implemented
-through messages yielding simple but powerful communication instructure.
-We distinguish between Data Producers (DP) and Data Consumers
-(DC) within the scope of a DSP. DP and DC can be seen as roles that a
-DSP Component may have. It is possible for a DSP Component to act in
-the role of DC and DP at the same time. Messages in DSPs are routed based 
-on rules. Rules can be statically or dynamically configured. Rules can use 
-different source information when routing a message: producers or consumers of the message, 
-the message itself or specific configuration provided by the administrators.
+Figure \ref{FIG_DSP} depicts the basic architecture of the DSP. We
+refer to the plug-and-play modules as DSP Components. DSP Components
+are self-contained modules that can be added and removed to a DSP
+installation at runtime. The basic communication paradigm within a DSP
+is a message. DSP Components can exchange messages of different types.
+Actually, all interactions between DSP Components are implemented
+through messages yielding simple but powerful communication
+instructure.  We distinguish between \emph{Data Producers} (DP) and
+\emph{Data Consumers} (DC) within the scope of a DSP. DP and DC can be
+seen as roles that a DSP Component may have. It is possible for a DSP
+Component to act in the role of DC and DP at the same time. Messages
+in DSPs are routed based on rules. Rules can be configured statically
+or dynamically. Rules for routing messages can be based on different
+information: producers or consumers of the message, the message
+itself, or specific configuration parameters provided externally.
 
@@
 \centering
-\epsfig{file=dsp-platform, width=7cm}
+\epsfig{file=dsp, width=7cm}
 \caption{\label{FIG_DSP} Data Sensor Platform (DSP).}
@@
 The DSP is responsible for routing messages between DPs and DCs. It is
-important to note that this routing only happens withing a
+important to note that this routing only happens within a
 DSP. Following the paradigm of a micro-kernel approach, the DSP has no
 notion of remote communication or sensor-specific details. All this
-knowledge is embodied in DSP Components.
+knowledge is embodied in special purpose DSP Components.
 
@@
 that consume data (DC). An example for a DSP Component acting in the
-role of a DP is a sensor module. The sensor module reads data from the
-physical sensor and converts the data into a message that is forwarded
-to the DSP. Because this module generates data with respect to the
-DSP, it is a data producer. An example for a data consumer is a
-converter module that converts the internal, DSP-specific data format,
-suitable for consumption by external applications. One such common
-format used in the domain of environmental monitoring is OpenDAP
-\cite{opendap01}. Since the converter module accepts messages from the DSP, it
-acts in the role of a data consumer. Finally, a DSP Component
-implementing persistency is an example of a module that acts in the
-roles of both data producer and consumer. It accepts messages from the
-DSP and stores them in a database. At a later time, it may act as a
-data producer by resending previously stored messages.
+role of a DP is a module that interfaces with a physical sensor. Such
+a module reads data from the sensor and converts the data into a
+message that is forwarded to the DSP. Because this module generates
+data with respect to the DSP, it is a data producer. An example for a
+data consumer is a converter module that converts the internal,
+DSP-specific data format to a format suitable for consumption by
+external applications. One such common format used in the domain of
+environmental monitoring is OpenDAP \cite{opendap01}.  Since the
+converter module accepts messages from the DSP, it acts in the role of
+a data consumer. Finally, a DSP Component implementing persistency is
+an example of a module that acts in the roles of both data producer
+and consumer. It accepts messages from the DSP and stores them in a
+database. At a later time, it may act as a data producer by resending
+previously stored messages.
 
@@
    public void deliver(DSPMessage msg);
-   public Message deliverWithReply(DSPMessage msg);
    // ...
@@
 By implementing the above Java interface, the resuling code
-effectively becomes a DSP Component that can be deployed at runtime. A
-component is identified by its type. Only one component of a certain type per DSP 
-is permissible. A type identifier much be unique across all DSP Components in DSP. 
-Upon deploying a DSP Component, an initialization function is called (line 4). Once a
+effectively becomes a DSP Component that can be deployed on a DSP at
+runtime. A component is identified by its type (line 3). Only one
+component of a certain type per DSP is permissible. A type identifier
+must be unique across all DSP Components in a DSP.  Upon deploying a
+component, an initialization function is called (line 4). Once a
 component is initialized, it may be started (line 5) and stopped (line
 6) multiple time. Whenever the DSP wishes to pass a message (that was
-generated by some DP), it invokes the \texttt{deliver()} method (ine
+generated by some DP), it invokes the \texttt{deliver()} method (line
 7). The message structure will be explained in the following section.
-One important feature of the DSP is the ability to redeploy (maybe with a new version) at 
-runtime. This allows DSPs to be updated remotely. This process is managed through
-messages sent to a special DSP Component which represents the DSP itself.
+One important feature of the DSP is the ability to redeploy (maybe
+with a new version) at runtime. This allows DSPs to be updated
+remotely. This process is managed through messages sent to a special
+DSP Component that is able to manage its local DSP.
 
-\subsection{Message}
+\subsection{Message Structure}
 
 Message passing is the main paradigm for exchanging information within
-a DSP. After producing data, DP component wraps that data in the 
-body of a message, so that DC components can receive the message and consume 
-the data on the message's body. When it comes to passing messages to a remote
-DSP instane, messages are serialized in XML \cite{xml2000} and automatically 
-deserialized for DC located in the remote DSP. In order to serialize and
-deserialize DSP messages, the DSP infrastructure uses the XML Schema 
-standard \cite{xml-schema2004} to generate and validate the DSP message. Figure
-\ref{FIG_DSP_MESSAGE} defines the structure of a general DSP Message.
+a DSP. After producing data, a DP wraps that data in the body
+of a message, so that DC components can receive the message and
+consume the data contained in the message's body. When it comes to passing
+messages to a remote DSP instance, messages are serialized in XML
+\cite{xml2000} and automatically deserialized for DC located in the
+remote DSP. In order to serialize and deserialize DSP messages, the
+DSP infrastructure uses the XML Schema standard \cite{xml-schema2004}
+to generate and validate the DSP message. Figure \ref{FIG_DSP_MESSAGE}
+defines the structure of a general DSP message.
 
@@
  \epsfig{file=dsp-message, width=9cm, height=4cm}
- \caption{\label{FIG_DSP_MESSAGE} Structure of a DSP Message.}
+ \caption{\label{FIG_DSP_MESSAGE} Structure of a DSP message.}
 \end{figure}
 
-The above XML Schema is defined for an ``Abstract Message'' type, whose
-structure serves as a base schema for all different types of messages. Any
-message will contain two basic attributes: ``messageID'' and ``ContentType''. 
-The former uniquely identifies a message and the latter the data in the body of
-the message. Complementary identification information such as the message 
-producer and consumer are located in the header of the message, which are
-defined by the schema type ``ComponenIdentifier''. A ``ComponentIdentifier''
-carries out information about the component name and its physical host address.
-Finally, the body of the message may contain any data type defined by the component 
-developer, and it can be interpreted as the DSP payload.
+The basic message structure is defined by an \texttt{AbstractMessage}
+type, whose structure serves as a base schema for all different types
+of messages. Any message will contain two basic attributes:
+\texttt{messageID} and \texttt{ContentType}.  The former uniquely
+identifies a message and the latter the data in the body of the
+message. Complementary identification information such as the message
+producer and consumer are located in the header of the message, which
+are defined by the schema type \texttt{ComponenIdentifier}. A
+\texttt{ComponentIdentifier} carries out information about the
+component name and its physical host address.  Finally, the body of
+the message represents the payload and may contain any data type
+defined by the component developer.
 
-Although XML schema is the de-facto standard to bind and validate xml data to
-its type, the architectural infrastructure of DSP uses an XML Data Bindining
-Technology \cite{xml-dbind} called JAXB \cite{xml-jaxb}. It is used to generate
-a type safe API for handling in-memory object graphs from the XML instances of 
-DSP Messages and the content of its data payload. Moreover, taking advantage 
-of the data type polimorphirsm of XML schema data types - concept borrowed from 
-Object-Oriented Design \cite{Rumbaugh1991} - we defined different types of 
-messages in order to allow specific DP and DC interpret the payload of any 
-given message solely based on their runtime type, as follows:
+Although XML schema is the de-facto standard to bind and validate XML
+data to its type, the architectural infrastructure of DSP uses an XML
+data binding technology \cite{xml-dbind} called JAXB \cite{xml-jaxb}.
+It is used to generate a type safe API for handling in-memory object
+graphs from the XML instances of DSP Messages and the content of its
+data payload. Moreover, taking advantage of the data type polymorphism
+of XML schema data types -- a concept borrowed from object oriented
+principles -- we defined different categories of messages in order to
+reflect the nature of the message content. Among those categories are
+query and update messages for configuring various properties supported
+by DSP Components and measurement messages that actually contain
+sensor readings.
+
+The payload of a message is itself described via an XML schema in
+order to integrate it seemingly with the general message structure. A
+developer of a DSP Component can define a schema that represents the
+type of payload processed by that component. An example illustrating a
+DSP message is given when we discuss the YSI Sonde Component.
+
+\subsection{DSP Broker}
+
+In order to keep the API symmetrical, the DSP is itself implemented as
+a DSP Component, the so-called DSP Broker. While technically the DSP
+Broker has the same API as any component, it serves a special role and
+cannot be removed. When a DP sends a message, it actually
+invokes the \texttt{deliver()} method of the DSP Broker. The broker is
+responsible for routing the message to one or several DCs. It
+distinguishes between local and remote destinations:
 
 \begin{itemize}
-  \item Action Message;
-  \item Event Message;
-  \item Create Message;
-  \item Insert Message;
-  \item Update Message;
-  \item Delete Message;
-  \item Query Message;
-  \item Measure Message.
+\item Local delivery of messages: it uses local object reference to
+  deliver the DSP message.
+\item Remote delivery of messages: it uses a DSP Component marked as a
+  default gateway to handle remote message delivery.
 \end{itemize}
 
-Not only the message structure is defined by XML schemas, but also the body
-content is and may be of any user-defined type. Therefore, the data
-defined to be sent from any given DP to any DC can be also in XML format or
-in-memory. An example illustrating a DSP message is given in the YSI section.
-
-\subsection{Broker}
-
-As soon as a DP produces data and wraps it up in a DSP Message, the DP only
-interact with the DSP platform via the DSP Broker by making the call
-``Broker.deliver(DSP Message)''. In this way, DSP broker is responsible for
-delivering the produced message to DSP consumers DCs, as well as verifying
-whether the delivery of the message is going to be done locally or remotely.
-
-\begin{itemize}
- \item Local delivery of messages: it uses local object reference to deliver the
-DSP message;
- \item Remote delivery of messages: it uses any DSP component that is
-specialized in remote delivery defined as ``Gateway DSP Component''.
-\end{itemize}
-
-The selection of the DCs for a given message to where the Broker will deliver
-messages or use a gateway component is determined by an object called Matcher.
-
-\subsection{Matcher}
-
-The Matcher is an object responsible to find the possible set of DCs for a given
-DSP message, and its ``deterministic intelligence'' relies on matching rules
-defined by the DSP administrator. In this way, a set of matching rules is define
-in XML and stored at the DSP run-time directory, where each matching rule is
-uniquely identified by a rule ID and determines the following (see
-figure \ref{FIG_MATCHER_RULE}):
+The selection of the DCs for a given message to where the broker will
+deliver messages to is determined by a matcher.  The matcher is
+responsible to find the possible set of DCs for a given DSP message,
+and it relies on matching rules defined statically or at runtim. In
+this way, a set of matching rules is defined in XML and stored at the
+DSP run-time directory, where each matching rule is uniquely
+identified by a rule ID and determines the following (see Figure
+\ref{FIG_MATCHER_RULE}):
 
@@
 \begin{itemize}
- \item Match Criteria: given the DSP message delivered to the Broker, which
-component(s) produced that message and in which node? It defines the name of
-the component type of the producer and its physical location (IP address);
- \item Match Target: given the DSP message, which component(s) are going
-to consume the message, and in which node? It defines the name of the component
-type of the consumer, and optionally defines a gateway component.
+\item The \emph{match criteria} defines the name of the component type
+  of the producer and its physical location (IP address);
+\item The \emph{match target} defines the name of the component type
+  of the consumer, and optionally defines a gateway component.
 \end{itemize}
 
-\subsection{Matcher replies to Broker}
-
-When analyzing the matching rules, the Matcher only selects the rules that
-satisfy the Matching Criteria and forwards them back to the Broker. Once the
-broker has the matching rules, it determines whether the delivery is done
-directly to a component defined on the Match Target, or if the message is
-delivered to the optional gateway. We have defined special gateway DSP
-components that is used to transport the DSP Messages, serialized in
-XML, to any remote DSP instance.
-
-\subsection{OSGi}
-
-OSGi lends itself naturally to implement the DSP.
-
-DSP can be implemented in different ways, but OSGi is a natural match.
-DSP Components are mapped to bundles. DSP itself is a bundle.
+When analyzing the matching rules, the matcher only selects the rules
+that satisfy the matching criteria and forwards them back to the
+broker. Once the broker has the matching rules, it determines whether
+the delivery is done directly to a component defined on the match
+target, or if the message is delivered to the optional gateway. We
+have defined special gateway DSP components that is used to transport
+the DSP Messages, serialized in XML, to any remote DSP instance.
 
@@
 
-In our project, we use the Verdex Pro XM4 motherboard with 64MB of SDRAM, 16MB of Flash, and 
-a 400MHz Marvell PXA270 with XScale processor.  We also use two expansion boards:
-the Console-VX which comes with three RS232 serial ports and the Netpro-VX which
-features one 10/100baseT ethernet port.  The Gumstix runs Linux 2.6 with the 
-BusyBox utilities, and uses the OpenEmbedded cross-compile build environment to 
-provide a complete Linux environment and a large range of Linux applications for
-the ARM architecture.  We use the JamVM \cite{jamvm01} implementation of Sun 
-Microsystem's virtual machine in order to run the DSP. The Gumstix communicates 
-with the YSI Sonde via an RS232 serial port using a null-modem cable since both 
-the Sonde and the Gumstix are DTE devices. The serial port configuration is:
+In our project, we use the Verdex Pro XM4 motherboard with 64MB of
+SDRAM, 16MB of Flash, and a 400MHz Marvell PXA270 with XScale
+processor.  We also use two expansion boards: the Console-VX which
+comes with three RS232 serial ports and the Netpro-VX which features
+one 10/100baseT ethernet port.  The Gumstix runs Linux 2.6 with the
+BusyBox utilities, and uses the OpenEmbedded cross-compile build
+environment to provide a complete Linux environment and a large range
+of Linux applications for the ARM architecture.  We use the JamVM
+\cite{jamvm01} implementation of Sun Microsystem's virtual machine in
+order to run the DSP. The Gumstix communicates with the YSI Sonde via
+an RS232 serial port using a null-modem cable since both the Sonde and
+the Gumstix are DTE devices. The serial port configuration is:
 
@@
 Our implementation of the DSP is based on the Open Source
-implementation of OSGi called Knopflerfish \cite{knopflerfish01}. Since Sun
-Microsystems does not support its JDK for the Gumstix, we use JamVM
-\cite{jamvm01} as the Java virtual machine implementation and GNU Classpath
-\cite{classpath01} for the runtime libraries. We have implemented various DSP
-Components that implement the functionality needed for NetBEAMS. On
-the far left of Figure \ref{FIG_NETBEAMS} is the actual sensor; in our
-case the aforementioned YSI Sonde. A YSI Sonde data producer (1) is a DSP
-component that uses the RS232 standard to read the physical measurements from
-the device over a serial port. This component converts the physical measurements 
-into a DSP-specific message. The following XML Schema on figure \ref{FIG_DSP_YSI_DATA} 
-defines the YSI Sonde's specific message format which includes each measurement data point, described in
-Table \ref{TAB_SONDE_MEASUREMENTS}, and its type:
+implementation of OSGi called Knopflerfish \cite{knopflerfish01}.
+Since Sun Microsystems does not support its JDK for the Gumstix, we
+use JamVM \cite{jamvm01} as the Java virtual machine implementation
+and GNU Classpath \cite{classpath01} for the runtime libraries. We
+have implemented various DSP Components that implement the
+functionality needed for NetBEAMS. On the far left of Figure
+\ref{FIG_NETBEAMS} is the actual sensor; in our case the
+aforementioned YSI Sonde. A YSI Sonde data producer (1) is a DSP
+component that uses the RS232 standard to read the physical
+measurements from the device over a serial port. This component
+converts the physical measurements into a DSP-specific message. The
+following XML Schema on figure \ref{FIG_DSP_YSI_DATA} defines the YSI
+Sonde's specific message format which includes each measurement data
+point, described in Table \ref{TAB_SONDE_MEASUREMENTS}, and its type:
 
@@
 
-The schema describes two Java classes, sondeDataType and SondeDataMessage, 
-which are used to package the Sonde's incoming data into a DSP message for delivery 
-to a remote consumer.  The SondeDataMessage class extends the AbstractMessageContent 
-class by adding a new type to the message payload, SondeDataType.  Each time the Sonde
-transmits its payload of measurement data during a sampling interval, the data is converted 
-into a SondeDataType which is then encapsulated within SondeDataMessage and then sent 
-to a remote consumer.
+The schema describes two Java classes, sondeDataType and
+SondeDataMessage, which are used to package the Sonde's incoming data
+into a DSP message for delivery to a remote consumer.  The
+SondeDataMessage class extends the AbstractMessageContent class by
+adding a new type to the message payload, SondeDataType.  Each time
+the Sonde transmits its payload of measurement data during a sampling
+interval, the data is converted into a SondeDataType which is then
+encapsulated within SondeDataMessage and then sent to a remote
+consumer.
 
-The YSI Sonde component can also be configured via special DSP messages, to direct the 
-Sonde to change its sampling frequency.  These property messages originate from the web 
-frontend that we will explained in a subsequent section.
+The YSI Sonde component can also be configured via special DSP
+messages, to direct the Sonde to change its sampling frequency.  These
+property messages originate from the web frontend that we will
+explained in a subsequent section.
 
@@
 
-After producing the SondeDataContainer data, the YSI Sonde component will wrap
-it up in a DSP Message Message and pass it to the DSP Broker for its delivery.
-However, our use case scenario is that the message must go from the local
-sensor device to the server. For this reason, we have developed a pair of
-client and server DSP components specialized in transferring and receiving 
-serialized objects throught the communication channel using the HTTP protocol
-\cite{RFC2068}.
+After producing the SondeDataContainer data, the YSI Sonde component
+will wrap it up in a DSP Message Message and pass it to the DSP Broker
+for its delivery.  However, our use case scenario is that the message
+must go from the local sensor device to the server. For this reason,
+we have developed a pair of client and server DSP components
+specialized in transferring and receiving serialized objects throught
+the communication channel using the HTTP protocol \cite{RFC2068}.
 
 \begin{itemize}
-  \item DSP Wire Transport Client: takes a collection of DSP messages and
-  transmit them using an HTTP client API written in JAVA \cite{http-client}. As
-  a non-functional requirement, a DSP administrator can set the interval in
-  which this component will transmit messages to other components;
-  \item DSP Wire Transport Server: is implemented using the Java Servlets
-  technology \cite{java-servlets}, and is used to receive the collection of DSP
-  messages transmitted, and potentially send DSP messages back to the sensor
-  station. 
+\item DSP Wire Transport Client: takes a collection of DSP messages
+  and transmit them using an HTTP client API written in JAVA
+  \cite{http-client}. As a non-functional requirement, a DSP
+  administrator can set the interval in which this component will
+  transmit messages to other components;
+\item DSP Wire Transport Server: is implemented using the Java
+  Servlets technology \cite{java-servlets}, and is used to receive the
+  collection of DSP messages transmitted, and potentially send DSP
+  messages back to the sensor station.
 \end{itemize}
@@
 One of the non-functional requirements for the DSP is to maximize the
-utilization of the communication channel. For this specific reason, we opted to
-add the type Messages Container, which contains a collection of DSP messages,
-and specific identifiers of the hosting sensor. Figure
-\ref{FIG_DSP_MSGS_CONTAINER} contains the structure of the DSP Messages
-Container.
+utilization of the communication channel. For this specific reason, we
+opted to add the type Messages Container, which contains a collection
+of DSP messages, and specific identifiers of the hosting sensor.
+Figure \ref{FIG_DSP_MSGS_CONTAINER} contains the structure of the DSP
+Messages Container.
 
@@
 
-It's important to note that, at this point, the DSP Broker had already checked
-with the Matcher which are the DSP components that will be receive a copy of
-the message in question. The matcher had specific instructions to use the DSP
-Wire Transport Client component as the default ``Proxy'' or default gateway, if
-you prefer.
+It's important to note that, at this point, the DSP Broker had already
+checked with the Matcher which are the DSP components that will be
+receive a copy of the message in question. The matcher had specific
+instructions to use the DSP Wire Transport Client component as the
+default ``Proxy'' or default gateway, if you prefer.
 
-By building a MessagesContainer instance, the Wire Transport Client adds all
-the messages addressed to the ``destinationHost'', with specific UUID of
-the component (2).
+By building a MessagesContainer instance, the Wire Transport Client
+adds all the messages addressed to the ``destinationHost'', with
+specific UUID of the component (2).
 
@@
 
-After the DSP Wire Transport Server receives the Messages Container (3), it
-unmarshalls the messages and forward them to the local DSP Broker.
+After the DSP Wire Transport Server receives the Messages Container
+(3), it unmarshalls the messages and forward them to the local DSP
+Broker.
 

### 2009-03-13 docs/icccn2009/paper.tex
Edited section on DSP
git-svn-id: svn://localhost/netbeams/trunk@374 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 In the following we present a practical application of the DSP
-presented in the previous section. We describe a deployment of the DSP
-using sensor owned and operated by the RTC. Our environmental sensor
-network is dubbed NetBEAMS (Networked Bay Environmental Assessment
-Monitoring System) and allows access to sensor equipment deployed
-throughout the San Francisco Bay Area.  Our goal is to address the
-problem of data collection from remote sensors by using the DSP to
-automatically interrogate sensors at the remote field locations and
-transmit the collected measurements over long distances to a backend
-server where the data is made available to potential users in near
-real-time. This end-to-end system will be a ready-to-use turnkey
-solution that is adaptable to a wide variety of sensor types and
-requires no programming expertise to operate. Configuration of the
+introduced in the previous section. Our environmental sensor network
+is dubbed NetBEAMS (Networked Bay Environmental Assessment Monitoring
+System) and allows access to sensor equipment deployed throughout the
+San Francisco Bay Area. This end-to-end system is a ready-to-use
+turnkey solution that is adaptable to a wide variety of sensor types
+and requires no programming expertise to operate. Configuration of the
 system will be possible when the sensor is deployed in the field via a
@@
 locations where each individual sensor is deployed in the field and
-the sparse nature of the network, the resulting topology is a tree
+the sparse density of the network, the resulting topology is a tree
 structure of depth 1.
@@
 computing platform called Gumstix \cite{gumstix01} (far left in Figure
-\ref{FIG_NETBEAMS}).  The Gumstix is an ARM architecture and comes in various 
-hardware configurations.  A typical Gumstix configuration consists 
-of a motherboard and one or more expansion boards which connect to the 
-motherboard via on-board buses.  The motherboards draw less than 250 mA @4V 
-at 400 MHz and less than 50 mA while idling, waiting for input.
+\ref{FIG_NETBEAMS}).  The Gumstix is an ARM architecture and comes in
+various hardware configurations.  A typical Gumstix configuration
+consists of a motherboard and one or more expansion boards which
+connect to the motherboard via on-board buses.  The motherboards draw
+less than 250 mA @4V at 400 MHz and less than 50 mA while idling,
+waiting for input.
 
-In our project, we use the Verdex Pro XM4 motherboard with 64MB of
-SDRAM, 16MB of Flash, and a 400MHz Marvell PXA270 with XScale
-processor.  We also use two expansion boards: the Console-VX which
-comes with three RS232 serial ports and the Netpro-VX which features
-one 10/100baseT ethernet port.  The Gumstix runs Linux 2.6 with the
-BusyBox utilities, and uses the OpenEmbedded cross-compile build
-environment to provide a complete Linux environment and a large range
-of Linux applications for the ARM architecture.  We use the JamVM
-\cite{jamvm01} implementation of Sun Microsystem's virtual machine in
-order to run the DSP. The Gumstix communicates with the YSI Sonde via
-an RS232 serial port using a null-modem cable since both the Sonde and
-the Gumstix are DTE devices. The serial port configuration is:
+The Gumstix runs Linux 2.6 with the BusyBox utilities, and uses the
+OpenEmbedded cross-compile build environment to provide a complete
+Linux environment and a large range of Linux applications for the ARM
+architecture.  We use the JamVM \cite{jamvm01} implementation of Sun
+Microsystem's virtual machine in order to run the DSP. The Gumstix
+communicates with the YSI Sonde via an RS232 serial port using a
+null-modem cable since both the Sonde and the Gumstix are DTE devices.
+The serial port configuration is:
 

### 2009-03-13 docs/icccn2009/paper.tex
Edited NetBEAMS section
git-svn-id: svn://localhost/netbeams/trunk@375 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 null-modem cable since both the Sonde and the Gumstix are DTE devices.
-The serial port configuration is:
+%The serial port configuration is:
 
-\begin{table}[!htb]
-\caption{\label{tab:RS232_Config}RS232 Serial Port Configuration}
-\centering
-\begin{tabular}{l l}
-\hline
-Baud Rate&9600\\
-Data Bit&8\\
-Parity&None\\
-Handshake&None\\
-\hline
-\end{tabular}
-\end{table}
+%\begin{table}[!htb]
+%\caption{\label{tab:RS232_Config}RS232 Serial Port Configuration}
+%\centering
+%\begin{tabular}{l l}
+%\hline
+%Baud Rate&9600\\
+%Data Bit&8\\
+%Parity&None\\
+%Handshake&None\\
+%\hline
+%\end{tabular}
+%\end{table}
 
-\begin{table}
-\caption{\label{TAB_RS232_Pinout}RS232 Serial Pinout}
-\centering
-\begin{tabular} {|l|l|l|}
-\hline
-Wire Color	&Pin Description        &DB-9\\ 
-\hline
-Yellow 		&RS232 TX    &2\\
-Orange		&RS232 RX    &3\\  
-Green  		&Alarm       &----\\ 
-Grey			&RTS         &----\\
-Blue  		&CTS         &----\\
-Red			&+ 12V DC    &9\\
-Black			&GND         &5\\  
-Purple		&SDI-12      &----\\ 
-Bare			&Shield      &----\\
-\hline
-\end{tabular}
-\end{table}
+%\begin{table}
+%\caption{\label{TAB_RS232_Pinout}RS232 Serial Pinout}
+%\centering
+%\begin{tabular} {|l|l|l|}
+%\hline
+%Wire Color	&Pin Description        &DB-9\\ 
+%\hline
+%Yellow 		&RS232 TX    &2\\
+%Orange		&RS232 RX    &3\\  
+%Green  		&Alarm       &----\\ 
+%Grey			&RTS         &----\\
+%Blue  		&CTS         &----\\
+%Red			&+ 12V DC    &9\\
+%Black			&GND         &5\\  
+%Purple		&SDI-12      &----\\ 
+%Bare			&Shield      &----\\
+%\hline
+%\end{tabular}
+%\end{table}
 
@@
 communications. We make use of the Huawei E220. The E220 is a USB
-modem that supports HSDPA, UMTS, and EDGE packet data services at maximum
-transmission rates of 3.6Mbps, 384kbps, and 236.8kbps respectively.
-HSDPA and UMTS operate at 2100MHz while GSM, GPRS, and EDGE operate at
-900, 1800, and 1900MHz. The modem connects to the Verdex Pro XM4
-motherboard via a mini USB interface (supporting USB 2.0). Linux
-kernels since 2.6.20 include support for E220 drivers.  This includes
-the linux kernel on the Gumstix which runs 2.6.21.  The Gumstix
-communicates with the modem using the PPP protocol and accesses the
-cellular network using a data plan from AT\&T.
+modem that supports HSDPA, UMTS, and EDGE packet data services at
+maximum transmission rates of 3.6Mbps, 384kbps, and 236.8kbps
+respectively.  HSDPA and UMTS operate at 2100MHz while GSM, GPRS, and
+EDGE operate at 900, 1800, and 1900MHz. The modem connects to the
+Gumstix motherboard via a mini USB interface (supporting USB 2.0).The
+Gumstix communicates with the modem using the PPP protocol and
+accesses the cellular network using a data plan from AT\&T.
 
@@
 \ref{FIG_NETBEAMS} is the actual sensor; in our case the
-aforementioned YSI Sonde. A YSI Sonde data producer (1) is a DSP
-component that uses the RS232 standard to read the physical
-measurements from the device over a serial port. This component
-converts the physical measurements into a DSP-specific message. The
-following XML Schema on figure \ref{FIG_DSP_YSI_DATA} defines the YSI
-Sonde's specific message format which includes each measurement data
-point, described in Table \ref{TAB_SONDE_MEASUREMENTS}, and its type:
+aforementioned YSI Sonde. The YSI Sonde DP (1) uses the RS232 standard
+to read the physical measurements from the device over a serial port.
+This component converts the physical measurements into a DSP-specific
+message. Figure \ref{FIG_DSP_YSI_DATA} depicts the YSI Sonde's
+specific message format which includes each measurement data point,
+described in Table \ref{TAB_SONDE_MEASUREMENTS}.
 
@@
 \epsfig{file=dsp-message-body-ysi, width=8cm, height=7cm}
-\caption{\label{FIG_DSP_YSI_DATA} Structure of the YSI Component Data.}
+\caption{\label{FIG_DSP_YSI_DATA} YSI message format.}
 \end{figure}
 
-The schema describes two Java classes, sondeDataType and
-SondeDataMessage, which are used to package the Sonde's incoming data
-into a DSP message for delivery to a remote consumer.  The
-SondeDataMessage class extends the AbstractMessageContent class by
-adding a new type to the message payload, SondeDataType.  Each time
-the Sonde transmits its payload of measurement data during a sampling
-interval, the data is converted into a SondeDataType which is then
-encapsulated within SondeDataMessage and then sent to a remote
-consumer.
-
-The YSI Sonde component can also be configured via special DSP
-messages, to direct the Sonde to change its sampling frequency.  These
-property messages originate from the web frontend that we will
-explained in a subsequent section.
+The SondeDataContainerType defines the payload of messages generated
+by the YSI Sonde DP. As all DSP messages, it is derived from the
+AbstractMessageContent. The payload mainly consists of the
+SondeDataType that defines all the sensor data that the YSI Sonde is
+capable of producing. The YSI Sonde DP uses the RS232 to query the
+physical device for its data in regular intervals. The measurements
+are then converted into a DSP message and forwarded to the DSP Broker
+from there they are forwarded to the remote communication component
+described in the next section.  The YSI Sonde DP can also be
+configured via special DSP messages, to direct the Sonde to change its
+sampling frequency.  These property messages originate from a web
+management interface (component (6) in Figure \ref{FIG_NETBEAMS}).
 
@@
 
-After producing the SondeDataContainer data, the YSI Sonde component
-will wrap it up in a DSP Message Message and pass it to the DSP Broker
-for its delivery.  However, our use case scenario is that the message
-must go from the local sensor device to the server. For this reason,
-we have developed a pair of client and server DSP components
-specialized in transferring and receiving serialized objects throught
-the communication channel using the HTTP protocol \cite{RFC2068}.
+All messages sent by the YSI Sonde DP need to be uploaded to a remote
+database. Since the DSP itself can only handle communication within
+its local DPs and DCs, we have created a matching pair of DC and DP
+that can handle remote communication (components (2) and (3) in Figure
+\ref{FIG_NETBEAMS}). As explained earlier, we make use of the cellular
+phone network to upload sensor data to a remote backend server. While
+the cellular phone network has sufficient coverage for more
+environmental sensor networks, it is not cost effective to keep a data
+link permanently open. The remote communiction components can be
+configured to upload buffered sensor data only in certain time
+intervals.Messages are transported via HTTP \cite{RFC2068}. The
+matching DP/DC pair for remote communication has the following
+properties:
 
@@
 \item DSP Wire Transport Client: takes a collection of DSP messages
-  and transmit them using an HTTP client API written in JAVA
-  \cite{http-client}. As a non-functional requirement, a DSP
-  administrator can set the interval in which this component will
-  transmit messages to other components;
-\item DSP Wire Transport Server: is implemented using the Java
-  Servlets technology \cite{java-servlets}, and is used to receive the
-  collection of DSP messages transmitted, and potentially send DSP
-  messages back to the sensor station.
+  and transmit them via HTTP. As a non-functional requirement, the
+  interval in which this component will transmit messages to other
+  components can be configured via the management interface.
+\item DSP Wire Transport Server: is a HTTP server that receives the
+  collection of DSP messages sent by the client, and piggy backs any
+  messages destined for the remote client that may have queued up in
+  the meantime.
 \end{itemize}
 
-One of the non-functional requirements for the DSP is to maximize the
-utilization of the communication channel. For this specific reason, we
-opted to add the type Messages Container, which contains a collection
-of DSP messages, and specific identifiers of the hosting sensor.
-Figure \ref{FIG_DSP_MSGS_CONTAINER} contains the structure of the DSP
-Messages Container.
-
-\begin{figure}[!htb]
-\centering
-\epsfig{file=dsp-messages-container, width=8cm}
-\caption{\label{FIG_DSP_MSGS_CONTAINER} The structure of Messages Container.}
-\end{figure}
-
-Some remarks regarding the Messages Container and it's implementation are as
-follows:
-
-\begin{itemize}
-  \item Container with or without Messages Container: an instance of
-  MessagesContainer may or may not have DSP messages;
-  \item Container may be used to acknowledge messages using a similar strategy
-  as the Slide Windowing protocol \cite{slide-window}. 
-\end{itemize}
-
-\subsection{Sending Messages Container}
-
-It's important to note that, at this point, the DSP Broker had already
-checked with the Matcher which are the DSP components that will be
-receive a copy of the message in question. The matcher had specific
-instructions to use the DSP Wire Transport Client component as the
-default ``Proxy'' or default gateway, if you prefer.
-
-By building a MessagesContainer instance, the Wire Transport Client
-adds all the messages addressed to the ``destinationHost'', with
-specific UUID of the component (2).
-
-\subsection{Receiving Messages Container}
-
-After the DSP Wire Transport Server receives the Messages Container
-(3), it unmarshalls the messages and forward them to the local DSP
-Broker.
-
-The reception of those messages will ALWAYS depend on the local DSP Broker and
-its combination with the Matcher. The steps (4), (5) and (6) are the
-actual reception of DSP messages. Container might also include an attribute
-called acknowledgeUntil, which determines the next actions that will occur in the system.
+The HTTP request and response are used to upload and download any
+messages that have been queued at the respective end of the
+communication link. The client and server component are capable of
+buffering messages until they are sent to the remote destination
+during the next communication cycle. Once messages have been
+transported to their final destinations, they are unmarshalled and
+passed on to the local DSP Broker for further dispatching. In this
+regard the client and server are completely symmetrical. Messages are
+dispatched with the help of the matcher to their final destinations.
 

### 2009-03-13 docs/icccn2009/paper.tex
Edited conclusions
git-svn-id: svn://localhost/netbeams/trunk@376 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 Section \ref{SEC_NETBEAMS} describes NetBEAMS; a practical application
-of the DSP. In Section \ref{SEC_CONCLUSION} we provide a conclusion
-and outlook.
+of the DSP. Section \ref{SEC_CONCLUSION} concludes this paper.
 
@@
 
-\section{Conclusions and Outlook}
+\section{Conclusions}
 \label{SEC_CONCLUSION}
@@
 decision that impact public health, public safety, and the protection
-of resources.  In addition, there are efforts underway at the state
-and federal level to collate the data collected by the individual data
-providers into integrated monitoring networks that will enhance the
-distribution of monitoring data to public, private, and federal users
-groups.  Success of these efforts, whether small scale or a large
+of resources. Success of these efforts, whether small scale or a large
 integrated network, depends on the first step in the data management
@@
 availability of a plug-and-play computing platform designed to
-communicate with a variety of sensor types and to provide data
-telemetry to distant brick and mortar data centers would stabilize
-data streams with poor telemetry and allowing expansion of sensor
-coverage into regions not presently served.  The low power consumption
-and low cost of these units would be particularly beneficial for
-applications with long service cycles and for data providers trying to
-maintain data streams on a limited budget.
+communicate with a variety of sensor types would stabilize data
+streams with poor telemetry and allowing expansion of sensor coverage
+into regions not presently served.
+
+In this paper we introduce a practical application of a sensor network
+in the domain of environmental monitoring. We have built an
+environmental sensor network in the San Francisco Bay using highly
+specialized, off-the-shelf water quality sensors. The resulting
+infrastructure enables near-realtime access to sensor data collected
+from remote locations. Our end-to-end solution furthermore allows
+configuration and management of remotely deployed sensors. The key to
+our infrastructure is the DSP; a micro-kernel plug-and-play
+architecure that can be configured flexibly to specific needs through
+special purpose components. The NetBEAMS infrastructure as described
+in the paper is available as Open Source and can be viewed in
+operation at \texttt{www.netbeams.org}.
 

### 2009-03-13 docs/icccn2009/paper.tex
Adding better quality EPS images and adjusting their sizes.
Adding 2 the header image in case we wanna add more details
about it.

git-svn-id: svn://localhost/netbeams/trunk@377 559edde2-a4d4-44e0-96fc-4035e40d7396
Incorporated Dale's feedback
git-svn-id: svn://localhost/netbeams/trunk@379 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 processed, archived, and made available to potential users in near
-real-time. In the following we give an overview of the NetBEAMS
+real-time. In the following sections we give an overview of the NetBEAMS
 architecture and implementation.  Section \ref{SEC_BACKGROUND}
@@
 sensing devices for their purposes. These sensing devices are at the
-opposite spectrum of miniature sensors such as SmartDust. One such
-device is the YSI 6600EDS V2 Sonde \cite{Sonde01}. It is a water
+opposite end of the size spectrum of miniature sensors such as SmartDust. One such
+device is the YSI 6600EDS V2 Sonde \cite{Sonde01}, a water
 quality monitoring device that gathers water quality data, and
@@
 \centering
-\epsfig{file=dsp-message-body-ysi, width=8cm, height=7cm}
+\epsfig{file=dsp-message-body-ysi, width=9cm}
 \caption{\label{FIG_DSP_YSI_DATA} YSI message format.}
@@
 configured to upload buffered sensor data only in certain time
-intervals.Messages are transported via HTTP \cite{RFC2068}. The
+intervals. Messages are transported via HTTP \cite{RFC2068}. The
 matching DP/DC pair for remote communication has the following

### 2009-03-13 docs/icccn2009/paper.tex
Add OSGi section
git-svn-id: svn://localhost/netbeams/trunk@381 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 
+\subsection{OSGi}
+
+OSGi \cite{osgi} is a generic platform for "dynamic module system for Java\texttrademark".
+\cite{Mueller07} used OSGi as platform their sensor network.
+In \cite{Baude07} OSGi is used to support large scale deployment. OSGi is flexible
+enough to be used in mobile devices and in application servers. One aspect not addressed
+by OSGi is distributed communication (although some extensions have
+been proposed,e.g., \cite{Rellermeyer}). DSP was designed to handle remote communication and,
+therefore, can complement OSGi capability. DSP Components are OSGi bundles (in fact, DSP itself
+is a bundle) and can be remotely and dynamically maintained. OSGi capabilities along with DSP
+design make well suitable for our project.
+
 \section{NetBEAMS}

### 2009-03-13 docs/icccn2009/paper.tex
Added ding-symbols
git-svn-id: svn://localhost/netbeams/trunk@382 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 \usepackage{url}
+\usepackage{pifont}
 
@@
 \ref{FIG_NETBEAMS} is the actual sensor; in our case the
-aforementioned YSI Sonde. The YSI Sonde DP (1) uses the RS232 standard
+aforementioned YSI Sonde. The YSI Sonde DP \ding{192} uses the RS232 standard
 to read the physical measurements from the device over a serial port.
@@
 sampling frequency.  These property messages originate from a web
-management interface (component (6) in Figure \ref{FIG_NETBEAMS}).
+management interface (component \ding{197} in Figure \ref{FIG_NETBEAMS}).
 
@@
 its local DPs and DCs, we have created a matching pair of DC and DP
-that can handle remote communication (components (2) and (3) in Figure
+that can handle remote communication (components \ding{193} and \ding{194} in Figure
 \ref{FIG_NETBEAMS}). As explained earlier, we make use of the cellular

### 2009-03-13 docs/icccn2009/paper.tex
Added section on DSP Implementation
git-svn-id: svn://localhost/netbeams/trunk@383 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 \subsection{DSP Broker}
+\label{SEC_DSP_BROKER}
 
@@
 
-\subsection{YSI Sonde Component}
+\subsection{DSP Implementation}
+
+The DSP is implemented in Java using the OSGi platform. OSGi
+\cite{osgi} is a Java-based component system. So-called bundles, which
+are self-contained Java JAR files, can be deployed and undeployed at
+runtime via special purpose class loaders. In fact, the API of the DSP
+is inspired by OSGi. All DSP Components depicted in Figure
+\ref{FIG_NETBEAMS} are implemented as OSGi bundles. The DSP Broker
+discussed in Section \ref{DSP_BROKER} is mapped to an OSGi bundle as well.
 
@@
 use JamVM \cite{jamvm01} as the Java virtual machine implementation
-and GNU Classpath \cite{classpath01} for the runtime libraries. We
-have implemented various DSP Components that implement the
-functionality needed for NetBEAMS. On the far left of Figure
-\ref{FIG_NETBEAMS} is the actual sensor; in our case the
-aforementioned YSI Sonde. The YSI Sonde DP \ding{192} uses the RS232 standard
-to read the physical measurements from the device over a serial port.
-This component converts the physical measurements into a DSP-specific
-message. Figure \ref{FIG_DSP_YSI_DATA} depicts the YSI Sonde's
-specific message format which includes each measurement data point,
-described in Table \ref{TAB_SONDE_MEASUREMENTS}.
+and GNU Classpath \cite{classpath01} for the runtime libraries. In the
+following sections we provide further details of some of the DSP
+Components that we have implemented as part of the NetBEAMS project.
+
+\subsection{YSI Sonde Component}
+
+On the far left of Figure \ref{FIG_NETBEAMS} is the actual sensor; in
+our case the aforementioned YSI Sonde. The YSI Sonde DP \ding{192}
+uses the RS232 standard to read the physical measurements from the
+device over a serial port.  This component converts the physical
+measurements into a DSP-specific message. Figure
+\ref{FIG_DSP_YSI_DATA} depicts the YSI Sonde's specific message format
+which includes each measurement data point, described in Table
+\ref{TAB_SONDE_MEASUREMENTS}.
 
@@
 
-The SondeDataContainerType defines the payload of messages generated
+The \texttt{SondeDataContainerType} defines the payload of messages generated
 by the YSI Sonde DP. As all DSP messages, it is derived from the
-AbstractMessageContent. The payload mainly consists of the
-SondeDataType that defines all the sensor data that the YSI Sonde is
+\texttt{AbstractMessageContent}. The payload mainly consists of the
+\texttt{SondeDataType} that defines all the sensor data that the YSI Sonde is
 capable of producing. The YSI Sonde DP uses the RS232 to query the

### 2009-03-13 docs/icccn2009/paper.tex
Added section on web interface
git-svn-id: svn://localhost/netbeams/trunk@384 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 
-Web Interface. Management messages are regular
-DSPMessages. Asynchronous UI: UI may only build up after next
-communication cycle.
+One of NetBEAMS goals is the ability to easily configure the sensor
+network without programming skills. Technicians should be able to
+change configuration parameters of various DSP Components via an
+easy-to-use web interface. With regards to the NetBEAMS specific
+components discussed in earlier sections, these configuration
+parameters are for example the sampling frequency of the YSI Sonde DP
+and the upload frequency of the the wire transport DC.
+
+The way this configuration is done by a special DSP Component that
+offers a web interface through which a technician can define the
+configuration parameters (component \ding{197} in Figure
+\ref{FIG_NETBEAMS}). The web management component first queries the remote
+component to get a list of configuration parameters. Once the remote
+component sends a reply, a user interface is automatically built up
+to allow the entry of user defined values. When those values are
+submitted, an update message is sent back to the remote component that
+will then change the appropriate configuration parameters.
+
+The user interface offered by the web management component is
+asynchronous in the sense that there might be a time lag between a
+request and a response. This has to do with the fact that the remote
+DSP will only establish a communication link in certain time intervals
+during which the query and update messages will be forwarded. Apart
+from component configuration, the web management component is also
+capable to remotely start, stop, and re-deploy new versions of DSP
+Components.
 

### 2009-03-13 docs/icccn2009/paper.tex
This is my paper revision in all sections. Please, diff this revision with the previous one for better capturing the details.
git-svn-id: svn://localhost/netbeams/trunk@385 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 environmental conditions at different locations \cite{roemer:2004}.
-Much research effort is focused on building wireless ad-hoc networks
+Much research effort is focused on building wireless ad-hoc networks, 
 where each sensor node participates in a multi-hop routing algorithm.
-One the earliest examples of a wireless sensor network is the
-SmartDust project \cite{smartdust:2001} where one objective was to
+One of the earliest examples of a wireless sensor network is the
+SmartDust project \cite{smartdust:2001}, where one objectives was to
 create autonomous sensing and communication in a cubic
@@
 devices, sensors for environmental monitoring are typically on the
-order of one to two feet in length weighing several pounds. Our
+order of one to two feet in length, weighing several pounds. Our
 focus is on building an end-to-end infrastructure for environmental
@@
 adjustment in monitoring strategy in response to rapidly developing
-events (e.g. oil spills, harmful algal blooms).
+events (e.g. oil spills, harmful algal blooms, among others).
 
@@
 \ref{SEC_DSP} describes our Data Sensor Platform (DSP) that allows us
-to efficiently build up an end-to-end environmental sensor network.
-Section \ref{SEC_NETBEAMS} describes NetBEAMS; a practical application
-of the DSP. Section \ref{SEC_CONCLUSION} concludes this paper.
+to efficiently build up an end-to-end environmental sensor network while
+Section \ref{SEC_NETBEAMS} describes NetBEAMS. Finally, Section 
+\ref{SEC_CONCLUSION} presents our conclusions and future works.
 
@@
 is normally performed while a technician manages the Sonde during the
-sampling process. In this mode the Sonde is usually connected via 650
-MDS Display/Logger or a serial cable connected to a PC.  The sampling
-frequency is usually set to a high frequency in this mode.  Unattended
+sampling process. In this mode, the Sonde is usually connected via 650
+MDS Display/Logger or a serial cable connected to a PC, where the sampling
+frequency is usually set to a high frequency. In contrast, Unattended
 mode is performed usually when the Sonde is deployed offshore or in a
-remote body of water.  The sampling frequency is usually set for a
-longer period of time (i.e. 5 - 15 min). The Sonde is configured via a
-set of menus that can be accessed through a terminal session over
-RS232. Through these menus, the Sonde can be configured to start
-logging data to an internal file.  To view the data as it's being
-captured, the option to show live data is chosen and the data can then
+remote body of water, where the sampling frequency is usually set for a
+longer period of time (i.e. 5 - 15 min). In order to be configured,
+the Sonde offers a set of menus that can be accessed through a terminal session
+over RS232. In this way, the Sonde can be configured to
+start logging data to an internal file. In order to view the data as it's
+being captured, the option to show live data is chosen and the data can then be 
 seen via a serial terminal session at each interval.
@@
 locations that not only make their maintenance difficult, but also
-access to their measurements.  Typically, the sensors are autonomous
-in the sense that they can run for a certain time on battery power and
+the access to their measurements.  Typically, the sensors are autonomous
+in the sense that they can run for a certain amount of time on battery power and
 store the measurement data in internal memory. Only when the sensors
@@
 \begin{itemize}
-\item Allow near-realtime access to sensor measurements.
+\item Allow near-realtime access to sensor measurements;
 \item Allow remote servicing that include changing of configuration
-  parameters as well as software upgrades.
-\item Support cost-effective long distance communication link.
+  parameters as well as software upgrades;
+\item Support cost-effective long distance communication link;
 \item Can be deployed in remote locations in unprotected outdoor
@@
 
-In the next section we describe a software architecture that provides
+In the next section, we describe a software architecture that provides
 a foundation to fulfill these requirements.
@@
 Sensor networks, especially of environmental sensors, are normally
-deployed over a geographically wide area where the communication
-infrastructure is not always reliable or, at least, not countinuously
-available. As the sensor network grows in size and complexity
-scalability, reliabilty and maintainability become critical issues.
-Additional components might be added to the network, sensor
-configuration might be altered, or new versions of components might be
-deployed. The sensor network must be able to support such adjustments
-without significant overhead for maintainers.  From system
+deployed over a geographically wide area, where the communication
+infrastructure is not always reliable or, at least, not coutinuously
+available. As the sensor network grows in size and complexity, 
+scalability, reliability and maintainability become critical issues.
+For instance, additional components might be added to the network, sensor
+configuration might be altered, or new versions of components might have to be
+deployed. As a consequence, the sensor network must be able to support such
+adjustments without significant overhead for maintainers.  From system
 development and integration point of view, interfaces among components
-and between components and the network should be simple. Complex
+and between components and the network should be simple. As a result, complex
 interaction can be constructed on top of these simple interfaces.  In
-the following we discuss how our proposed architechure addresses those
-issues.
+the following subsections, we discuss how our proposed architecture addresses
+those issues.
 
@@
 
-Our architecture is based on a micro-kernel approach. The core
+NetBEAMS's architecture is based on a micro-kernel approach. The core
 functionality is limited to a minimum and all extra services are built
 as plug-and-play modules on top of the micro-kernel. With the help of
-this approach we can keep the basic infrastructure compact while
-allowing customization via special-purpose modules. In our terminology
+this approach, we can keep the basic infrastructure compact while
+allowing customization via special-purpose modules. In our terminology,
 we call the basic foundation the \emph{Data Sensor Platform}
 (DSP). The scope of the DSP is one address space of an execution
-platform. A sensor network is built from several DSPs that are linked
+platform and a sensor network is built from several DSPs that are linked
 with each other.
@@
 are self-contained modules that can be added and removed to a DSP
-installation at runtime. The basic communication paradigm within a DSP
+installation at runtime. The basic communication unit within a DSP
 is a message. DSP Components can exchange messages of different types.
@@
 through messages yielding simple but powerful communication
-instructure.  We distinguish between \emph{Data Producers} (DP) and
+structure. Moreover, we distinguish between \emph{Data Producers} (DP) and
 \emph{Data Consumers} (DC) within the scope of a DSP. DP and DC can be
-seen as roles that a DSP Component may have. It is possible for a DSP
-Component to act in the role of DC and DP at the same time. Messages
-in DSPs are routed based on rules. Rules can be configured statically
+seen as roles that a DSP Component may have, as it may also 
+act in the role of DC and DP at the same time. Messages
+in DSPs are routed based on rules, which can be configured statically
 or dynamically. Rules for routing messages can be based on different
-information: producers or consumers of the message, the message
+information such as producers or consumers of the message, the message
 itself, or specific configuration parameters provided externally.
@@
 converter module accepts messages from the DSP, it acts in the role of
-a data consumer. Finally, a DSP Component implementing persistency is
+a data consumer. Finally, a DSP Component implementing persistence is
 an example of a module that acts in the roles of both data producer
-and consumer. It accepts messages from the DSP and stores them in a
-database. At a later time, it may act as a data producer by resending
-previously stored messages.
+and consumer, since it accepts messages from the DSP and stores them in a
+database. Furthermore, it may act as a data producer by resending previously
+stored messages at a later time.
 
@@
 
-By implementing the above Java interface, the resuling code
+By implementing the above Java interface, the resulting code
 effectively becomes a DSP Component that can be deployed on a DSP at
@@
 component is initialized, it may be started (line 5) and stopped (line
-6) multiple time. Whenever the DSP wishes to pass a message (that was
+6) multiple times. Whenever the DSP wishes to pass a message (that was
 generated by some DP), it invokes the \texttt{deliver()} method (line
-7). The message structure will be explained in the following section.
-One important feature of the DSP is the ability to redeploy (maybe
-with a new version) at runtime. This allows DSPs to be updated
+7). The message structure will be explained in the next section.
+One important feature of the DSP is the ability to redeploy itself (maybe
+with a new version with updates) at runtime. This allows DSPs to be updated
 remotely. This process is managed through messages sent to a special
@@
 messages to a remote DSP instance, messages are serialized in XML
-\cite{xml2000} and automatically deserialized for DC located in the
+\cite{xml2000} and automatically deserialized for DCs located in the
 remote DSP. In order to serialize and deserialize DSP messages, the
@@
 identifies a message and the latter the data in the body of the
-message. Complementary identification information such as the message
-producer and consumer are located in the header of the message, which
-are defined by the schema type \texttt{ComponenIdentifier}. A
+message. Complementary identification information, such as the message
+producer and consumer, is located in the header of the message and 
+is defined by the schema type \texttt{ComponenIdentifier}. A
 \texttt{ComponentIdentifier} carries out information about the
@@
 reflect the nature of the message content. Among those categories are
-query and update messages for configuring various properties supported
-by DSP Components and measurement messages that actually contain
-sensor readings.
+query and update messages, where the former is used for configuring various
+properties supported by DSP Components, and the latter is used for measurement
+messages that actually contain sensor readings.
 
 The payload of a message is itself described via an XML schema in
-order to integrate it seemingly with the general message structure. A
-developer of a DSP Component can define a schema that represents the
-type of payload processed by that component. An example illustrating a
-DSP message is given when we discuss the YSI Sonde Component.
+order to integrate it seemingly with the general message structure. By this
+means, a developer of a DSP Component can define a schema that represents the
+type of payload processed by that component. An example illustrating a DSP
+message is given when we discuss the YSI Sonde Component.
 
@@
 invokes the \texttt{deliver()} method of the DSP Broker. The broker is
-responsible for routing the message to one or several DCs. It
+responsible for routing the message to one or several DCs. Furthermore, it
 distinguishes between local and remote destinations:
@@
 \item Local delivery of messages: it uses local object reference to
-  deliver the DSP message.
+  deliver the DSP message;
 \item Remote delivery of messages: it uses a DSP Component marked as a
@@
 responsible to find the possible set of DCs for a given DSP message,
-and it relies on matching rules defined statically or at runtim. In
+and it relies on matching rules defined statically or at runtime. In
 this way, a set of matching rules is defined in XML and stored at the
@@
 OSGi \cite{osgi} is a generic platform for "dynamic module system for Java\texttrademark".
-\cite{Mueller07} used OSGi as platform their sensor network.
-In \cite{Baude07} OSGi is used to support large scale deployment. OSGi is flexible
-enough to be used in mobile devices and in application servers. One aspect not addressed
-by OSGi is distributed communication (although some extensions have
-been proposed,e.g., \cite{Rellermeyer}). DSP was designed to handle remote communication and,
-therefore, can complement OSGi capability. DSP Components are OSGi bundles (in fact, DSP itself
-is a bundle) and can be remotely and dynamically maintained. OSGi capabilities along with DSP
-design make well suitable for our project.
+Many researchers have been using OSGi to support different requirements similar
+to the DSP. For instance, \cite{Mueller07} used OSGi as a platform to their
+sensor network, while \cite{Baude07} used OSGi to support large scale
+deployment. As many have researchers have noticed, OSGi is flexible enough to be
+used in memory-limited mobile devices or in larger application servers.
+However, as being primarily focused on providing an infrastructure for
+loosely-coupled module interaction, OSGi does not address distributed
+communication although some extensions have been proposed, e.g.,
+\cite{Rellermeyer}). As one of the contributions of this project, DSP was
+designed to handle remote communication and, therefore, can complement OSGi
+capability. Therefore, DSP Components are OSGi bundles (in fact, DSP itself is a
+bundle) and they can be remotely and dynamically maintained, so that OSGi
+capabilities along with DSP design make well suitable for our project.
 
@@
 
-In the following we present a practical application of the DSP
+In the following subsection, we present a practical application of the DSP
 introduced in the previous section. Our environmental sensor network
@@
 technicians between service cycles to change systems functionalities
-(e.g. sampling rate).  We present the NetBEAMS infrastructure by
+such as sampling rate.  We present the NetBEAMS infrastructure by
 discussing its hardware and software components.
@@
 computing platform called Gumstix \cite{gumstix01} (far left in Figure
-\ref{FIG_NETBEAMS}).  The Gumstix is an ARM architecture and comes in
-various hardware configurations.  A typical Gumstix configuration
-consists of a motherboard and one or more expansion boards which
+\ref{FIG_NETBEAMS}).  The Gumstix is an ARM architecture hardware and comes in
+various configurations.  A typical Gumstix configuration
+consists of a motherboard and one or more expansion boards that
 connect to the motherboard via on-board buses.  The motherboards draw
@@
 communicates with the YSI Sonde via an RS232 serial port using a
-null-modem cable since both the Sonde and the Gumstix are DTE devices.
+null-modem cable, since both the Sonde and the Gumstix are DTE devices.
 %The serial port configuration is:
@@
 networks and being a significantly lower cost solution than satellite
-communications. We make use of the Huawei E220. The E220 is a USB
+communications. We make use of the Huawei E220, which is a USB
 modem that supports HSDPA, UMTS, and EDGE packet data services at
@@
 EDGE operate at 900, 1800, and 1900MHz. The modem connects to the
-Gumstix motherboard via a mini USB interface (supporting USB 2.0).The
-Gumstix communicates with the modem using the PPP protocol and
+Gumstix motherboard via a mini USB interface (supporting USB 2.0). In this
+way, the Gumstix communicates with the modem using the PPP protocol and
 accesses the cellular network using a data plan from AT\&T.
@@
 
-The DSP is implemented in Java using the OSGi platform. OSGi
-\cite{osgi} is a Java-based component system. So-called bundles, which
-are self-contained Java JAR files, can be deployed and undeployed at
-runtime via special purpose class loaders. In fact, the API of the DSP
-is inspired by OSGi. All DSP Components depicted in Figure
-\ref{FIG_NETBEAMS} are implemented as OSGi bundles. The DSP Broker
-discussed in Section \ref{DSP_BROKER} is mapped to an OSGi bundle as well.
+As mentioned earlier, the DSP is implemented in Java using the OSGi
+platform. The so-called bundles, which are self-contained Java JAR files, can
+be deployed and undeployed at runtime via special purpose class loaders. In 
+fact, the API of the DSP is inspired by OSGi. All DSP Components depicted in Figure
+\ref{FIG_NETBEAMS} are implemented as OSGi bundles. The DSP Broker discussed in
+Section \ref{SEC_DSP_BROKER} is mapped to an OSGi bundle as well.
 
-Our implementation of the DSP is based on the Open Source
+Our implementation of the DSP is based on the Open-Source
 implementation of OSGi called Knopflerfish \cite{knopflerfish01}.
@@
 use JamVM \cite{jamvm01} as the Java virtual machine implementation
-and GNU Classpath \cite{classpath01} for the runtime libraries. In the
-following sections we provide further details of some of the DSP
+and the GNU Classpath \cite{classpath01} project for the runtime libraries. In
+the following sections, we provide further details of some of the DSP
 Components that we have implemented as part of the NetBEAMS project.
@@
 On the far left of Figure \ref{FIG_NETBEAMS} is the actual sensor; in
-our case the aforementioned YSI Sonde. The YSI Sonde DP \ding{192}
+our case, the aforementioned YSI Sonde. The YSI Sonde DP \ding{192}
 uses the RS232 standard to read the physical measurements from the
@@
 measurements into a DSP-specific message. Figure
-\ref{FIG_DSP_YSI_DATA} depicts the YSI Sonde's specific message format
+\ref{FIG_DSP_YSI_DATA} depicts the YSI Sonde's specific message format, 
 which includes each measurement data point, described in Table
@@
 physical device for its data in regular intervals. The measurements
-are then converted into a DSP message and forwarded to the DSP Broker
-from there they are forwarded to the remote communication component
+are then converted into a DSP message and forwarded to the DSP Broker,
+from where they are forwarded to the remote communication component
 described in the next section.  The YSI Sonde DP can also be
-configured via special DSP messages, to direct the Sonde to change its
+configured via special Update DSP messages, to direct the Sonde to change its
 sampling frequency.  These property messages originate from a web
@@
 environmental sensor networks, it is not cost effective to keep a data
-link permanently open. The remote communiction components can be
+link permanently open. The remote communication components can be
 configured to upload buffered sensor data only in certain time
-intervals. Messages are transported via HTTP \cite{RFC2068}. The
-matching DP/DC pair for remote communication has the following
-properties:
+intervals. Having this in mind, DSP Messages are transported via HTTP
+\cite{RFC2068}. In this way, the matching DP/DC pair for remote communication
+has the following properties:
 
@@
   interval in which this component will transmit messages to other
-  components can be configured via the management interface.
-\item DSP Wire Transport Server: is a HTTP server that receives the
+  components can be configured via the management interface;
+\item DSP Wire Transport Server: is an HTTP server that receives the
   collection of DSP messages sent by the client, and piggy backs any
@@
 messages that have been queued at the respective end of the
-communication link. The client and server component are capable of
+communication link. The client and server components are capable of
 buffering messages until they are sent to the remote destination
@@
 passed on to the local DSP Broker for further dispatching. In this
-regard the client and server are completely symmetrical. Messages are
-dispatched with the help of the matcher to their final destinations.
+regard, the client and server components are completely symmetrical. Messages
+are dispatched with the help of the matcher to their final destinations.
 
@@
 These sensors are operated by data providers, including resource
-management groups, research scientists, municipalities, and state and
+management groups, research scientists, municipalities, state and
 federal agencies, who could use near real-time data streams to inform
-decision that impact public health, public safety, and the protection
+decisions that impact public health, public safety, and the protection
 of resources. Success of these efforts, whether small scale or a large
@@
 
-In this paper we introduce a practical application of a sensor network
+In this paper, we introduce a practical application of a sensor network
 in the domain of environmental monitoring. We have built an
@@
 infrastructure enables near-realtime access to sensor data collected
-from remote locations. Our end-to-end solution furthermore allows
-configuration and management of remotely deployed sensors. The key to
-our infrastructure is the DSP; a micro-kernel plug-and-play
-architecure that can be configured flexibly to specific needs through
-special purpose components. The NetBEAMS infrastructure as described
-in the paper is available as Open Source and can be viewed in
-operation at \texttt{www.netbeams.org}.
+from remote locations. Our end-to-end solution, furthermore, allows
+configuration and management of remotely deployed sensors. In this sense, the
+key to our infrastructure is the DSP, a micro-kernel plug-and-play
+architecture that can flexibly be configured to specific needs through
+special purpose components. Furthermore, DSP is a contribution to the OSGi
+project, providing an optional way of unified loosely-coupled bundle
+interaction and distributed communication. The NetBEAMS infrastructure, as
+described in the paper, is available as an Open-Source project and can be 
+viewed in operation at \texttt{www.netbeams.org}.
 

### 2009-03-14 versions/v2/apps/osgi-bundles/dsp/DSPKPIndicators/src/org/netbeams/dsp/demo/kpi/DSPKPIndicators.java
DSPKPIndicators component
git-svn-id: svn://localhost/netbeams/trunk@386 559edde2-a4d4-44e0-96fc-4035e40d7396

@@ -1,0 +1,169 @@
+package org.netbeams.dsp.demo.kpi;
+
+
+import javax.xml.bind.JAXBContext;
+import javax.xml.bind.JAXBException;
+import javax.xml.bind.Unmarshaller;
+
+import org.apache.log4j.Logger;
+import org.netbeams.dsp.ComponentDescriptor;
+import org.netbeams.dsp.message.ComponentLocator;
+import org.netbeams.dsp.DSPComponent;
+import org.netbeams.dsp.DSPContext;
+import org.netbeams.dsp.DSPException;
+
+import org.netbeams.dsp.message.MeasureMessage;
+import org.netbeams.dsp.message.Message;
+
+import org.netbeams.dsp.weblogger.Buffer;
+
+
+import org.w3c.dom.Node;
+import org.netbeams.dsp.demo.stock.StockTick;
+import org.netbeams.dsp.demo.stock.StockTicks;
+import org.netbeams.dsp.demo.mouseactions.MouseAction;
+import org.netbeams.dsp.demo.mouseactions.MouseActionsContainer;
+
+
+public class DSPKPIndicators implements DSPComponent{
+	
+	
+	private static final Logger log = Logger.getLogger(DSPKPIndicators.class);
+	
+	// Component Type
+	public final static String COMPONENT_TYPE = "org.netbeams.dsp.demo.kpi";
+	
+	private static ComponentDescriptor componentDescriptor;
+
+	
+	private String componentNodeId;
+	
+	private DSPContext context;
+	private ComponentLocator locator;
+	private ComponentDescriptor descriptor;
+
+	/////////////////////////////////////////////
+	////////// DSP Component Interface //////////
+	/////////////////////////////////////////////
+
+
+	/**
+	 * @Override
+	 */
+	public String getComponentNodeId() {
+		return componentNodeId;
+	}
+
+	/**
+	 * @Override
+	 */
+	public String getComponentType() {
+		return COMPONENT_TYPE;
+	}
+
+	/**
+	 * @Override
+	 */
+	public void initComponent(String componentNodeId, DSPContext context) throws DSPException {
+		log.info("Initializing...");
+		
+		this.context = context;
+		this.componentNodeId = componentNodeId;
+	}
+
+	/**
+	 * @Override
+	 */
+	public ComponentDescriptor getComponentDescriptor() {
+		return componentDescriptor;
+	}
+
+
+	/**
+	 * @Override
+	 */
[…]

### 2009-03-14 docs/icccn2009/paper.tex
Revise DSP Implementation, Broker, matcher and OSGi
git-svn-id: svn://localhost/netbeams/trunk@387 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
    public String getComponentType();
-   public void initComponent(...);
+   public void initComponent(String     id, 
+                             DSPContext context);
    public void startComponent();
@@
 component, an initialization function is called (line 4). Once a
-component is initialized, it may be started (line 5) and stopped (line
-6) multiple times. Whenever the DSP wishes to pass a message (that was
+component is initialized, it may be started (line 4) and stopped (line
+7) multiple times. Whenever the DSP wishes to pass a message (that was
 generated by some DP), it invokes the \texttt{deliver()} method (line
-7). The message structure will be explained in the next section.
+8). DP sends messages through \texttt{Context} implementation (provided 
+by the DSP during DSP initialization)The message structure will be explained in the next section.
 One important feature of the DSP is the ability to redeploy itself (maybe
@@
 \subsection{DSP Broker}
-\label{SEC_DSP_BROKER}
-
-In order to keep the API symmetrical, the DSP is itself implemented as
-a DSP Component, the so-called DSP Broker. While technically the DSP
-Broker has the same API as any component, it serves a special role and
-cannot be removed. When a DP sends a message, it actually
-invokes the \texttt{deliver()} method of the DSP Broker. The broker is
-responsible for routing the message to one or several DCs. Furthermore, it
+Broker is responsible for routing the message to one or several DCs. It
 distinguishes between local and remote destinations:
@@
 \item Local delivery of messages: it uses local object reference to
-  deliver the DSP message;
+  deliver the DSP message.
 \item Remote delivery of messages: it uses a DSP Component marked as a
-  default gateway to handle remote message delivery.
+  gateway (configured in matching rules) to handle remote message delivery.
 \end{itemize}
 
-The selection of the DCs for a given message to where the broker will
-deliver messages to is determined by a matcher.  The matcher is
-responsible to find the possible set of DCs for a given DSP message,
-and it relies on matching rules defined statically or at runtime. In
-this way, a set of matching rules is defined in XML and stored at the
-DSP run-time directory, where each matching rule is uniquely
-identified by a rule ID and determines the following (see Figure
-\ref{FIG_MATCHER_RULE}):
+Broker provides the infrastructure necessary to deliver messages concurrently and
+efficiently. Itis  also responsible for setting any additional message header 
+information necessary for proper routing. Broker does not dictates which DCs should
+receive a message. It delegates this responsibility to the Matcher.
+
+\subsection{Matcher}
+
+Matcher is the component in DSP responsible for matching messages from a DP to one or more
+DCs. The decision process is based on evaluation of rules (see Figure
+\ref{FIG_MATCHER_RULE}). Rules can be statically or dynamically
+configured.
 
@@
 
-When analyzing the matching rules, the matcher only selects the rules
-that satisfy the matching criteria and forwards them back to the
-broker. Once the broker has the matching rules, it determines whether
-the delivery is done directly to a component defined on the match
-target, or if the message is delivered to the optional gateway. We
-have defined special gateway DSP components that is used to transport
-the DSP Messages, serialized in XML, to any remote DSP instance.
+Matcher's main responsibility is to shield DPs from the responsibility
+for specifying consumers of their messages. DPs may create messages without any pre-determinted
+DCs. Once the message is sent through DSP, the matcher will determine which DCs should
+receive the message. Note that, a DP can still specify a DC for its message. In this case, Matcher
+includes this DC in list of DCs to receive the message. Matcher also plays an important role on
+the dynamic aspect of the DSP. Suppose the communication between two DSPs becomes unreliable and
+it necessary to persist messages before they are sent through the communication channel. It is possible to
+achieve this goal by deploying a DSP component which implements persistence and modifying the matching
+rules to divert messages to the new DSP Component and from the component to the original destination.
 
@@
 
-As mentioned earlier, the DSP is implemented in Java using the OSGi
-platform. The so-called bundles, which are self-contained Java JAR files, can
-be deployed and undeployed at runtime via special purpose class loaders. In 
-fact, the API of the DSP is inspired by OSGi. All DSP Components depicted in Figure
-\ref{FIG_NETBEAMS} are implemented as OSGi bundles. The DSP Broker discussed in
-Section \ref{SEC_DSP_BROKER} is mapped to an OSGi bundle as well.
+The DSP is implemented in java and serves as a container environment where 
+DSP Components execute. Due to its micro-kernel, modular and dynamic characteristics 
+the current DSP implementation uses OSGi \cite{osgi} as a deployment and execution 
+infrastucture. One important DSP requirement is to impose on DSP components minimal 
+dependency on DSP itself and no depency on OSGi. To this extent, all DSP resources
+accessible to DSP components are specified by \texttt{DSPContext} interface.
+DSP also uses message to communicate with DSP Components. It has a special "adapter"
+that acts like a DSPComponent. Therefore, configuring DSP is no different from configuring
+a DSP component.
+
+DSP compoments are deployed as OSGi bundles. OSGi defines bundles life-cycle and
+runtime upgrade mechanism. DSP uses these capabilities to allow remote deployment 
+(bundles with DSP Components are trasmitted as message to the DSP Component Adaptor) 
+and activation/deactivation of DSP Components.
 

### 2009-03-14 docs/icccn2009/paper.tex
Remove OSGi writeup in architecture section
git-svn-id: svn://localhost/netbeams/trunk@388 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 
-\subsection{OSGi}
-
-OSGi \cite{osgi} is a generic platform for "dynamic module system for Java\texttrademark".
-Many researchers have been using OSGi to support different requirements similar
-to the DSP. For instance, \cite{Mueller07} used OSGi as a platform to their
-sensor network, while \cite{Baude07} used OSGi to support large scale
-deployment. As many have researchers have noticed, OSGi is flexible enough to be
-used in memory-limited mobile devices or in larger application servers.
-However, as being primarily focused on providing an infrastructure for
-loosely-coupled module interaction, OSGi does not address distributed
-communication although some extensions have been proposed, e.g.,
-\cite{Rellermeyer}). As one of the contributions of this project, DSP was
-designed to handle remote communication and, therefore, can complement OSGi
-capability. Therefore, DSP Components are OSGi bundles (in fact, DSP itself is a
-bundle) and they can be remotely and dynamically maintained, so that OSGi
-capabilities along with DSP design make well suitable for our project.
-
 \section{NetBEAMS}

### 2009-03-14 docs/icccn2009/paper.tex
Cleanup architecture section
git-svn-id: svn://localhost/netbeams/trunk@389 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 component, an initialization function is called (line 4). Once a
-component is initialized, it may be started (line 4) and stopped (line
+component is initialized, it may be started (line 6) and stopped (line
 7) multiple times. Whenever the DSP wishes to pass a message (that was
 generated by some DP), it invokes the \texttt{deliver()} method (line
-8). DP sends messages through \texttt{Context} implementation (provided 
-by the DSP during DSP initialization)The message structure will be explained in the next section.
-One important feature of the DSP is the ability to redeploy itself (maybe
-with a new version with updates) at runtime. This allows DSPs to be updated
-remotely. This process is managed through messages sent to a special
-DSP Component that is able to manage its local DSP.
+8). A DP can send messages via a \texttt{DSPContext} object that was
+passed to the DP during initialization.
 
@@
 producer and consumer, is located in the header of the message and 
-is defined by the schema type \texttt{ComponenIdentifier}. A
+is defined by the schema type \texttt{ComponentIdentifier}. A
 \texttt{ComponentIdentifier} carries out information about the
@@
 
-Although XML schema is the de-facto standard to bind and validate XML
-data to its type, the architectural infrastructure of DSP uses an XML
+XML schema is the de-facto standard to bind and validate XML
+data to its type. Within the DSP we use an XML
 data binding technology \cite{xml-dbind} called JAXB \cite{xml-jaxb}.
@@
 data payload. Moreover, taking advantage of the data type polymorphism
-of XML schema data types -- a concept borrowed from object oriented
-principles -- we defined different categories of messages in order to
+of XML schema data types, we defined different categories of messages in order to
 reflect the nature of the message content. Among those categories are
@@
 
-\subsection{DSP Broker}
-Broker is responsible for routing the message to one or several DCs. It
-distinguishes between local and remote destinations:
+\subsection{DSP Broker and Matcher}
+
+The \emph{broker} is responsible for routing the message to one or
+several DCs. It distinguishes between local and remote destinations:
 
@@
 \item Remote delivery of messages: it uses a DSP Component marked as a
-  gateway (configured in matching rules) to handle remote message delivery.
+  gateway to handle remote message delivery.
 \end{itemize}
 
-Broker provides the infrastructure necessary to deliver messages concurrently and
-efficiently. Itis  also responsible for setting any additional message header 
-information necessary for proper routing. Broker does not dictates which DCs should
-receive a message. It delegates this responsibility to the Matcher.
-
-\subsection{Matcher}
-
-Matcher is the component in DSP responsible for matching messages from a DP to one or more
-DCs. The decision process is based on evaluation of rules (see Figure
-\ref{FIG_MATCHER_RULE}). Rules can be statically or dynamically
+The broker is can deliver messages concurrently and asynchronously. It
+sets any additional message header information necessary for proper
+routing. The broker does not make and routing decisions. It delegates
+this responsibility to the \emph{matcher}. The matcher is responsible
+for matching messages sent by a DP and route it to one or more DCs.
+This decision is based on a set of routing rules (see Figure
+\ref{FIG_MATCHER_RULE}) which can be statically or dynamically
 configured.
@@
 \item The \emph{match criteria} defines the name of the component type
-  of the producer and its physical location (IP address);
+  of the producer and its physical location (IP address).
 \item The \emph{match target} defines the name of the component type
@@
 
-Matcher's main responsibility is to shield DPs from the responsibility
-for specifying consumers of their messages. DPs may create messages without any pre-determinted
-DCs. Once the message is sent through DSP, the matcher will determine which DCs should
-receive the message. Note that, a DP can still specify a DC for its message. In this case, Matcher
-includes this DC in list of DCs to receive the message. Matcher also plays an important role on
-the dynamic aspect of the DSP. Suppose the communication between two DSPs becomes unreliable and
-it necessary to persist messages before they are sent through the communication channel. It is possible to
-achieve this goal by deploying a DSP component which implements persistence and modifying the matching
-rules to divert messages to the new DSP Component and from the component to the original destination.
+The matcher's main purpose is to achieve a loosely coupling between
+DCs and DPs. DPs may create messages without specifying a specific
+destination DC. Once the message is sent through DSP, the matcher will
+determine which DCs should receive the message. Note that, a DP can
+still specify a DC for its message. In this case, the matcher includes
+this DC in list of DCs to receive the message.
 

### 2009-03-14 versions/v2/apps/osgi-bundles/dsp/DSPPlatform/src/org/netbeams/dsp/platform/broker/MessageBroker.java
Adding changes to make the mouse actions to compile;
Also changing the Matcher and Broker to deliver messages locally.

Removing the JAXB code dependency from the RandomNumberGEnerator.

Marcello did these changes...

git-svn-id: svn://localhost/netbeams/trunk@392 559edde2-a4d4-44e0-96fc-4035e40d7396

@@ -230,7 +230,8 @@ public class MessageBroker implements MessageBrokerAccessor {
         if (consumerIdentifier != null) {
             String comonentNodeId = consumerIdentifier.getComponentLocator().getComponentNodeId();
             if (comonentNodeId != null) {
-                return components.get(comonentNodeId);
+                DSPComponent a = components.get(comonentNodeId);  
+                return a != null ? a : componentsByType.get(consumerIdentifier.getComponentType());
             } else {
                 return componentsByType.get(consumerIdentifier.getComponentType());
             }

### 2009-03-14 versions/v2/apps/osgi-bundles/dsp/DSPPlatform/src/org/netbeams/dsp/platform/matcher/MatchTarget.java

@@ -47,11 +47,10 @@ public class MatchTarget {
     }
 
     public int hashCode() {
-        if (this.gatewayComponentType != null)
-            return 10 * this.consumerComponentType.hashCode() + 20 * this.locator.getNodeAddress().getValue().hashCode() + 30
-                    * this.gatewayComponentType.hashCode();
-        else
-            return 10 * this.consumerComponentType.hashCode() + 20 * this.locator.getNodeAddress().getValue().hashCode();
+    	int value = (this.gatewayComponentType != null) ? 10 * this.consumerComponentType.hashCode() : 0;
+    	value += this.gatewayComponentType != null ? 30 * this.gatewayComponentType.hashCode() : 0;
+    	value += this.locator != null ? 20 * this.locator.getNodeAddress().getValue().hashCode() : 0;
+        return value;
     }
 
     public String getConsumerComponentType() {

### 2009-03-14 versions/v2/apps/osgi-bundles/dsp/DSPPlatform/src/org/netbeams/dsp/platform/matcher/Matcher.java

@@ -96,20 +96,27 @@ public class Matcher implements BaseComponent {
         ComponentIdentifier producer = message.getHeader().getProducer();
         ComponentIdentifier consumer = message.getHeader().getConsumer();
         
-        MatchCriteria crit = new MatchCriteria(producer.getComponentType(), producer.getComponentLocator());
-        MatchTarget targ = new MatchTarget(consumer.getComponentType(), consumer.getComponentLocator(), null);        
-        MatchRule localRule = new MatchRule("defaultRule", true, crit, targ);
-        consumers.add(localRule);
+        ComponentLocator locator = new ComponentLocator();
+        locator.setComponentNodeId("local_id");
+        NodeAddress nodeAddr = new NodeAddress();
+        nodeAddr.setValue("LOCAL");
+        locator.setNodeAddress(nodeAddr);
         
+        if (consumer != null) {
+	        MatchCriteria crit = new MatchCriteria(producer.getComponentType(), locator);
+	        MatchTarget targ = new MatchTarget(consumer.getComponentType(), locator, null);        
+	        MatchRule localRule = new MatchRule("defaultRule", true, crit, targ);
+	        consumers.add(localRule);
+        }
         for (MatchRule mr : rules) {
             log.debug("Verifying matcher for rule: " + mr.getRuleID() + " Is Default? " + mr.isDefault());
             log.debug("Message's producer: " + producer.getComponentType());
             log.debug("Message's consumer: " + consumer.getComponentType());
-            log.debug("component type matches rule criteria?");
+            log.debug("Component type matches rule criteria?");
             log.debug("Rule Criteria: Producer=" + mr.getCriteria().getProducerComponentType());
             log.debug("Rule Criteria: Consumer=" + mr.getTarget().getConsumerComponentType());
             if (isMatchForComponentType(producer, consumer, mr)) {
-                log.debug("Producer matches for producer and consumer types...");
+                log.debug("Component type matches for producer and consumer types...");
                 consumers.add(mr);
             } else {
                 log.debug("Message doesn't match on Component Type");

### 2009-03-14 versions/v2/apps/osgi-bundles/dsp/DSPTestRandomNumGen/src/org/netbeams/dsp/test/randomnumber/RandomNumberProducer.java

@@ -5,15 +5,12 @@ import java.util.ArrayList;
 import java.io.ByteArrayInputStream;
 import java.io.DataInputStream;
 import java.security.SecureRandom;
-import javax.xml.bind.JAXBException;
-import javax.xml.parsers.ParserConfigurationException;
 import org.apache.log4j.Logger;
 import org.netbeams.dsp.DSPContext;
 import org.netbeams.dsp.DSPException;
 import org.netbeams.dsp.MessageBrokerAccessor;
 import org.netbeams.dsp.data.property.DSProperties;
 import org.netbeams.dsp.data.property.DSProperty;
-import org.netbeams.dsp.data.property.ObjectFactory;
 import org.netbeams.dsp.message.ComponentIdentifier;
 import org.netbeams.dsp.message.DSPMessagesFactory;
 import org.netbeams.dsp.message.Header;
@@ -76,7 +73,7 @@ public class RandomNumberProducer {
         prop.setValue(data.getRandomNumbers().get(0).toString());
         
         try {
-        	Message message = DSPMessagesFactory.INSTANCE.makeDSPMeasureMessage(header, props, ObjectFactory.class);
+        	Message message = DSPMessagesFactory.INSTANCE.makeDSPMeasureMessage(header, props);
 		
         	// Always check if there is a broker available
         	MessageBrokerAccessor messageBroker = context.getDataBroker();
@@ -85,10 +82,7 @@ public class RandomNumberProducer {
         	}else{
         		log.debug("Message broker not available");
         	}
-        } catch (JAXBException e) {
-        	log.error(e.getMessage(), e);
-        } catch (ParserConfigurationException e) {
-        	log.error(e.getMessage(), e);        	
+        	
         } catch (DSPException e) {
         	log.error(e.getMessage(), e);
         }		

### 2009-03-14 versions/v2/apps/osgi-bundles/dsp/DSPTestRandomNumGen/src/org/netbeams/dsp/test/randomnumber/RandomNumbers.java

@@ -16,4 +16,9 @@ public class RandomNumbers extends MessageContent {
 		return rNumbers;
 	}
 
+	public String toXml() {
+		// TODO Auto-generated method stub
+		return null;
+	}
+
 }

### 2009-03-14 docs/icccn2009/paper.tex
Modified the Conclusion by removing the first paragraph and adding another that sums up related work.  Supporting bibliography was added.
git-svn-id: svn://localhost/netbeams/trunk@393 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 
-The California coastal region is populated by a vast number of
-remotely deployed sensors that monitor environmental conditions
-ranging from weather to water quality to ocean surface currents.
-These sensors are operated by data providers, including resource
-management groups, research scientists, municipalities, state and
-federal agencies, who could use near real-time data streams to inform
-decisions that impact public health, public safety, and the protection
-of resources. Success of these efforts, whether small scale or a large
-integrated network, depends on the first step in the data management
-process: reliably obtaining data in near real-time from remotely
-located sensors.  This step can be the weak link in the data
-management pathway when telemetry options are unreliable.  The
-availability of a plug-and-play computing platform designed to
-communicate with a variety of sensor types would stabilize data
-streams with poor telemetry and allowing expansion of sensor coverage
-into regions not presently served.
+%The California coastal region is populated by a vast number of
+%remotely deployed sensors that monitor environmental conditions
+%ranging from weather to water quality to ocean surface currents.
+%These sensors are operated by data providers, including resource
+%management groups, research scientists, municipalities, state and
+%federal agencies, who could use near real-time data streams to inform
+%decisions that impact public health, public safety, and the protection
+%of resources. Success of these efforts, whether small scale or a large
+%integrated network, depends on the first step in the data management
+%process: reliably obtaining data in near real-time from remotely
+%located sensors.  This step can be the weak link in the data
+%management pathway when telemetry options are unreliable.  The
+%availability of a plug-and-play computing platform designed to
+%communicate with a variety of sensor types would stabilize data
+%streams with poor telemetry and allowing expansion of sensor coverage
+%into regions not presently served.
 
@@
 architecture that can flexibly be configured to specific needs through
-special purpose components. Furthermore, DSP is a contribution to the OSGi
+special purpose components.
+
+Furthermore, our work improves upon existing technologies in earlier sensor network implementations.
+Many sensor networks use PIC microcontrollers \cite{Pic01} and specialized operating systems
+such as TinyOS \cite{tinyos01}. However PICs are usually too buggy,
+requiring non-standard operating systems \cite{Martinez04}. Likewise, TinyOS exhibits limited 
+constraints on storage, restricting any complex processing of sensor data
+that can be performed \cite{Martinez04}. OSGi has been used in other sensor network
+project \cite{Mueller07, Baude07}. Although some OSGi sensor networks also include remote 
+communication capabilities \cite{Rellermeyer}, our DSP is designed to support asynchronous 
+communication in loosely coupled networks. It is a contribution to the OSGi
 project, providing an optional way of unified loosely-coupled bundle

### 2009-03-14 versions/v2/apps/osgi-bundles/dsp/DSPPlatform/src/org/netbeams/dsp/platform/matcher/MatchRule.java
Testing the matcher updates
git-svn-id: svn://localhost/netbeams/trunk@394 559edde2-a4d4-44e0-96fc-4035e40d7396

@@ -24,8 +24,8 @@ public class MatchRule {
     }
 
     public int hashCode() {
-        return 10 * this.ruleID.hashCode() + 20 * this.criteria.hashCode() + 30 * this.target.hashCode() + 
-               (this.isDefault ? -1 : 1);
+        return 10 * this.ruleID.hashCode() + 20 * this.criteria.hashCode()
+                + (this.target != null ? 30 * this.target.hashCode() : 0) + (this.isDefault ? -1 : 1);
     }
 
     public String getRuleID() {

### 2009-03-14 versions/v2/apps/osgi-bundles/dsp/DSPTestRandomNumGen/src/org/netbeams/dsp/test/randomnumber/RandomNumberProducer.java

@@ -1,10 +1,11 @@
 package org.netbeams.dsp.test.randomnumber;
 
-import java.util.List;
-import java.util.ArrayList;
 import java.io.ByteArrayInputStream;
 import java.io.DataInputStream;
 import java.security.SecureRandom;
+import java.util.ArrayList;
+import java.util.List;
+
 import org.apache.log4j.Logger;
 import org.netbeams.dsp.DSPContext;
 import org.netbeams.dsp.DSPException;
@@ -20,121 +21,116 @@ import org.netbeams.dsp.util.NetworkUtil;
 /**
  * 
  * @author Teresa L. Johnson <gamma.particle@gmail.com>
- *
+ * 
  */
 public class RandomNumberProducer {
 
-	private static final Logger log = Logger.getLogger(RandomNumberProducer.class);
-	
-	private List<RandomNumber> numberList;
-	private DSPContext context;
-	private RandomNumberGenerator rng;
-	
-	
-	public RandomNumberProducer (DSPContext context) {
-		this.context = context;
-	    numberList = new ArrayList<RandomNumber>();
+    private static final Logger log = Logger.getLogger(RandomNumberProducer.class);
+
+    private List<RandomNumber> numberList;
+    private DSPContext context;
+    private RandomNumberGenerator rng;
+
+    public RandomNumberProducer(DSPContext context) {
+        this.context = context;
+        numberList = new ArrayList<RandomNumber>();
         log.trace("The DSPRandomNumberProducer initialized: sends every 5 seconds");
-	};
-	
-	
-	public void startProducer() {
-		rng = new RandomNumberGenerator();
-		rng.start();
-	}
-	
-	
-	public void stopProducer() {
-		if (rng != null) {
-			rng.stopGenerator();
-			rng = null;
-		}
-	}
-	
-	private void send(RandomNumbers data) throws DSPException{
-		
-		String localIPAddress = NetworkUtil.getCurrentEnvironmentNetworkIp();
-		
-		log.debug("Random Numbers to be sent from " + localIPAddress);
+    };
+
+    public void startProducer() {
+        rng = new RandomNumberGenerator();
+        rng.start();
+    }
+
+    public void stopProducer() {
+        if (rng != null) {
+            rng.stopGenerator();
+            rng = null;
+        }
+    }
+
+    private void send(RandomNumbers data) throws DSPException {
+
+        String localIPAddress = NetworkUtil.getCurrentEnvironmentNetworkIp();
+
[…]

### 2009-03-14 versions/v2/apps/osgi-bundles/dsp/DSPTestRandomNumGen/src/org/netbeams/dsp/test/randomnumber/RandomNumbers.java

@@ -4,17 +4,16 @@ import java.util.ArrayList;
 import org.netbeams.dsp.message.MessageContent;
 
 public class RandomNumbers extends MessageContent {
-	
-	private ArrayList<RandomNumber> rNumbers;
-	
-	
-	public RandomNumbers(ArrayList<RandomNumber> rNumbers) {
-		this.rNumbers = rNumbers;
-	};
-	
-	public ArrayList<RandomNumber> getRandomNumbers() {
-		return rNumbers;
-	}
+
+    private ArrayList<RandomNumber> rNumbers;
+
+    public RandomNumbers(ArrayList<RandomNumber> rNumbers) {
+        this.rNumbers = rNumbers;
+    };
+
+    public ArrayList<RandomNumber> getRandomNumbers() {
+        return rNumbers;
+    }
 
 	public String toXml() {
 		// TODO Auto-generated method stub

### 2009-03-14 docs/icccn2009/paper.tex
Uploaded more polishing changes to the paper.
git-svn-id: svn://localhost/netbeams/trunk@395 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 One of the earliest examples of a wireless sensor network is the
-SmartDust project \cite{smartdust:2001}, where one objectives was to
+SmartDust project \cite{smartdust:2001}, where one objective was to
 create autonomous sensing and communication in a cubic
@@
 monitoring data.  In addition, the dependence on field site visits to
-alter of sensor characteristics, such as sampling rate, prevents
-adjustment in monitoring strategy in response to rapidly developing
+alter sensor characteristics, such as sampling rate, prevents
+adjustments in monitoring strategy in response to rapidly developing
 events (e.g. oil spills, harmful algal blooms, among others).
@@
 Assessment Monitoring System), an environmental sensor network
-deployed in the San Francisco Bay. NetBEAMS provides means to program
-and interrogate sensors at the remote field locations and to transmit
+deployed in the San Francisco Bay. NetBEAMS provides the means to program
+and interrogate sensors at remote field locations and to transmit
 collected measurements to a database, where the data is rapidly
@@
 is normally performed while a technician manages the Sonde during the
-sampling process. In this mode, the Sonde is usually connected via 650
+sampling process. In this mode, the Sonde is connected via 650
 MDS Display/Logger or a serial cable connected to a PC, where the sampling
 frequency is usually set to a high frequency. In contrast, Unattended
-mode is performed usually when the Sonde is deployed offshore or in a
+mode is performed when the Sonde is deployed offshore or in a
 remote body of water, where the sampling frequency is usually set for a
@@
 locations that not only make their maintenance difficult, but also
-the access to their measurements.  Typically, the sensors are autonomous
+access to their measurements difficult.  Typically, the sensors are autonomous
 in the sense that they can run for a certain amount of time on battery power and
@@
 data can be uploaded to server. This manual procedure to retrieve
-sensor data is not only error prone, but also results are in a serious
-time lag between the time the sensor performs the measurement and when
+sensor data is not only error prone, but the results greatly differ
+between the time the sensor performs the measurement and when
 it becomes available for further processing. For many applications,
 however, it would be beneficial to have near-realtime access to the
-sensor measurement. We identify the following requirements that an
+sensor measurements. We identify the following requirements that an
 end-to-end infrastructure supporting devices such as the YSI should
@@
 
-Sensor networks, especially of environmental sensors, are normally
+Sensor networks, especially environmental sensors, are normally
 deployed over a geographically wide area, where the communication
@@
 deployed. As a consequence, the sensor network must be able to support such
-adjustments without significant overhead for maintainers.  From system
+adjustments without significant overhead for maintainers.  From a system
 development and integration point of view, interfaces among components
@@
 Actually, all interactions between DSP Components are implemented
-through messages yielding simple but powerful communication
+through messages yielding a simple but powerful communication
 structure. Moreover, we distinguish between \emph{Data Producers} (DP) and
@@
 The payload of a message is itself described via an XML schema in
-order to integrate it seemingly with the general message structure. By this
+order to integrate it seemingly with the general message structure. By these
 means, a developer of a DSP Component can define a schema that represents the
@@
 
-The broker is can deliver messages concurrently and asynchronously. It
+The broker can deliver messages concurrently and asynchronously. It
 sets any additional message header information necessary for proper
-routing. The broker does not make and routing decisions. It delegates
+routing. The broker does not make any routing decisions. It delegates
 this responsibility to the \emph{matcher}. The matcher is responsible
-for matching messages sent by a DP and route it to one or more DCs.
+for matching messages sent by a DP and routes it to one or more DCs.
 This decision is based on a set of routing rules (see Figure
@@
 
-The matcher's main purpose is to achieve a loosely coupling between
+The matcher's main purpose is to achieve a loose coupling between
 DCs and DPs. DPs may create messages without specifying a specific
-destination DC. Once the message is sent through DSP, the matcher will
+destination DC. Once the message is sent through the DSP, the matcher will
 determine which DCs should receive the message. Note that, a DP can
 still specify a DC for its message. In this case, the matcher includes
-this DC in list of DCs to receive the message.
+this DC in a list of DCs to receive the message.
 
@@
 web-based interface, eliminating the need for extra site visits by
-technicians between service cycles to change systems functionalities
-such as sampling rate.  We present the NetBEAMS infrastructure by
+technicians between service cycles to change system characteristics
+such as the sampling rate.  We present the NetBEAMS infrastructure by
 discussing its hardware and software components.
@@
 computing platform called Gumstix \cite{gumstix01} (far left in Figure
-\ref{FIG_NETBEAMS}).  The Gumstix is an ARM architecture hardware and comes in
+\ref{FIG_NETBEAMS}).  The Gumstix is an ARM-based single board computer and comes in
 various configurations.  A typical Gumstix configuration
@@
 
-The DSP is implemented in java and serves as a container environment where 
-DSP Components execute. Due to its micro-kernel, modular and dynamic characteristics 
+The DSP is implemented in Java and serves as a container environment where 
+DSP Components execute. Due to its micro-kernel, modular and dynamic characteristics, 
 the current DSP implementation uses OSGi \cite{osgi} as a deployment and execution 
 infrastucture. One important DSP requirement is to impose on DSP components minimal 
-dependency on DSP itself and no depency on OSGi. To this extent, all DSP resources
+dependency on the DSP itself and no depency on OSGi. To this extent, all DSP resources
 accessible to DSP components are specified by \texttt{DSPContext} interface.
-DSP also uses message to communicate with DSP Components. It has a special "adapter"
-that acts like a DSPComponent. Therefore, configuring DSP is no different from configuring
+The DSP also uses messages to communicate with DSP Components. It has a special "adapter"
+that acts like a DSPComponent. Therefore, configuring the DSP is no different from configuring
 a DSP component.
@@
 runtime upgrade mechanism. DSP uses these capabilities to allow remote deployment 
-(bundles with DSP Components are trasmitted as message to the DSP Component Adaptor) 
+(bundles with DSP Components are transmitted as message to the DSP Component Adaptor) 
 and activation/deactivation of DSP Components.
@@
 are then converted into a DSP message and forwarded to the DSP Broker,
-from where they are forwarded to the remote communication component
+from which they are forwarded to the remote communication component
 described in the next section.  The YSI Sonde DP can also be
@@
 configured to upload buffered sensor data only in certain time
-intervals. Having this in mind, DSP Messages are transported via HTTP
+intervals. With this in mind, DSP Messages are transported via HTTP
 \cite{RFC2068}. In this way, the matching DP/DC pair for remote communication
@@
 
-One of NetBEAMS goals is the ability to easily configure the sensor
+One of the NetBEAMS goals is to easily configure the sensor
 network without programming skills. Technicians should be able to
@@
 components discussed in earlier sections, these configuration
-parameters are for example the sampling frequency of the YSI Sonde DP
+parameters are, for example, the sampling frequency of the YSI Sonde DP
 and the upload frequency of the the wire transport DC.
 
-The way this configuration is done by a special DSP Component that
+This configuration is performed by a special DSP Component that
 offers a web interface through which a technician can define the
@@
 request and a response. This has to do with the fact that the remote
-DSP will only establish a communication link in certain time intervals
+DSP will only establish a communication link at certain time intervals
 during which the query and update messages will be forwarded. Apart
@@
 Many sensor networks use PIC microcontrollers \cite{Pic01} and specialized operating systems
-such as TinyOS \cite{tinyos01}. However PICs are usually too buggy,
+such as TinyOS \cite{tinyos01}. However, PICs are usually too buggy,
 requiring non-standard operating systems \cite{Martinez04}. Likewise, TinyOS exhibits limited 
@@
 that can be performed \cite{Martinez04}. OSGi has been used in other sensor network
-project \cite{Mueller07, Baude07}. Although some OSGi sensor networks also include remote 
+projects \cite{Mueller07, Baude07}. Although some OSGi sensor networks also include remote 
 communication capabilities \cite{Rellermeyer}, our DSP is designed to support asynchronous 

### 2009-03-20 docs/icccn2009/paper.tex
Kleber's feedback
git-svn-id: svn://localhost/netbeams/trunk@419 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 San Francisco, CA 94132 \\
-EMail: \{arno$\mid$tlj$\mid$klebers$\mid$msales\}@sfsu.edu}
+EMail: \{arno$\mid$tlj$\mid$kleber$\mid$msales\}@sfsu.edu}
 \and
@@
 communication capabilities \cite{Rellermeyer}, our DSP is designed to support asynchronous 
-communication in loosely coupled networks. It is a contribution to the OSGi
-project, providing an optional way of unified loosely-coupled bundle
-interaction and distributed communication. The NetBEAMS infrastructure, as
+communication in loosely coupled networks. The NetBEAMS infrastructure, as
 described in the paper, is available as an Open-Source project and can be 

### 2009-03-20 docs/icccn2009/paper.tex
Minor edits
git-svn-id: svn://localhost/netbeams/trunk@420 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 remote body of water, where the sampling frequency is usually set for a
-longer period of time (i.e. 5 - 15 min). In order to be configured,
+longer period of time (i.e. 5 - 15 min). Assuming a 5 min sampling
+frequency, the Sonde produces 267 KB of data per day.
+In order to be configured,
 the Sonde offers a set of menus that can be accessed through a terminal session
@@
 the current DSP implementation uses OSGi \cite{osgi} as a deployment and execution 
-infrastucture. One important DSP requirement is to impose on DSP components minimal 
+infrastucture. One important DSP requirement is to impose on DSP Components minimal 
 dependency on the DSP itself and no depency on OSGi. To this extent, all DSP resources
-accessible to DSP components are specified by \texttt{DSPContext} interface.
-The DSP also uses messages to communicate with DSP Components. It has a special "adapter"
-that acts like a DSPComponent. Therefore, configuring the DSP is no different from configuring
-a DSP component.
+accessible to DSP Components are specified by \texttt{DSPContext} interface.
+The DSP also uses messages to communicate with DSP Components. It has a special adapter
+that acts like a \texttt{DSPComponent}. Therefore, configuring the DSP is no different from configuring
+a DSP Component.
 
-DSP compoments are deployed as OSGi bundles. OSGi defines bundles life-cycle and
+DSP Compoments are deployed as OSGi bundles. OSGi defines bundles life-cycle and
 runtime upgrade mechanism. DSP uses these capabilities to allow remote deployment 
@@
 architecture that can flexibly be configured to specific needs through
-special purpose components.
+special purpose components. We plan to use the Gumstix's processing
+power to pre-process data intensive sensors such as the HF Radar to
+reduce the data volume.
 

### 2009-09-21 docs/icccn2009/paper.tex
Replaced former DSP EPS images with simpler ones.
git-svn-id: svn://localhost/netbeams/trunk@421 559edde2-a4d4-44e0-96fc-4035e40d7396
Resubmitting paper with DSP diagrams using Open Office. Both the new EPS and original ODG files have been uploaded.
git-svn-id: svn://localhost/netbeams/trunk@422 559edde2-a4d4-44e0-96fc-4035e40d7396
A very tiny edit.
git-svn-id: svn://localhost/netbeams/trunk@423 559edde2-a4d4-44e0-96fc-4035e40d7396
REVIEWS DONE BY KLEBER... I'M JUST COMMITTING HIS CHANGES DESCRIBED AT REVISION 77 OF THE PAPER REVISION GOOGLE DOCUMENT.
http://docs.google.com/Revs?id=0Af5ISdu-g4

@@
 millimeter-sized device.  While most wireless sensor networks focus on
-small sensing devices such as the SmartDust mote, there exist a wide
+small sensing devices such as the SmartDust mote, there exists a wide
 variety of off-the-shelf commercially available sensors that are used
@@
 we call the basic foundation the \emph{Data Sensor Platform}
-(DSP). The scope of the DSP is one address space of an execution
-platform and a sensor network is built from several DSPs that are linked
+(DSP). DSP is collection of local built-in and plug-in components; 
+a sensor network is built from several DSPs that are linked
 with each other.
@@
 generated by some DP), it invokes the \texttt{deliver()} method (line
-8). A DP can send messages via a \texttt{DSPContext} object that was
+8). A DP can send messages via a \texttt{DSP Broker} (a DSP Component itself) 
+obtained from \texttt{DSPContext} object that was
 passed to the DP during initialization.
@@
  \centering
- \epsfig{file=dsp-message, width=9cm, height=4cm}
+ \epsfig{file=AbstractMessage2, width=9cm, height=4cm}
  \caption{\label{FIG_DSP_MESSAGE} Structure of a DSP message.}
@@
 of messages. Any message will contain two basic attributes:
-\texttt{messageID} and \texttt{ContentType}.  The former uniquely
+\texttt{messageID} and \texttt{contentType}.  The former uniquely
 identifies a message and the latter the data in the body of the
@@
 \begin{figure}[!htb]
- \epsfig{file=dsp-match-rule, width=9cm, height=4cm}
+ \centering
+ \epsfig{file=MatcherRule2, width=8cm, height=5cm}
  \caption{\label{FIG_MATCHER_RULE} Definition of Matcher Rule.}
@@
 accessible to DSP Components are specified by \texttt{DSPContext} interface.
-The DSP also uses messages to communicate with DSP Components. It has a special adapter
-that acts like a \texttt{DSPComponent}. Therefore, configuring the DSP is no different from configuring
-a DSP Component.
+The DSP communicates with DSP Components through the same message mechanism DSP components
+used among themselves. Therefore, configuring the DSP is no different from configuring
+a DSP Component, which simplifies administration and monitoring implementations.
 
@@
 \centering
-\epsfig{file=dsp-message-body-ysi, width=9cm}
+\epsfig{file=YSI_Message_Format2, width=9cm}
 \caption{\label{FIG_DSP_YSI_DATA} YSI message format.}

### 2009-09-21 versions/v2/apps/osgi-bundles/dsp/DSPDataPersistence/src/org/netbeams/dsp/persistence/controller/DSPDataPersistence.java
This revision includes the last changes needed to run the DSP Data Persistence with mongoDB. Other changes include the Eclipse code formatter, addition of the separate experiment Java artifact, and description on the experiment script.
* .classpath: adding needed jars to the Eclipse classpath;

* eclipse.netbeans.codestyle.xml: java code formatter for Eclipse, updating the 
  conventions for Javadocs;

* DSPDataPersistence/build.xml: updating the build system to include the mongoDB
  Java drivers into the MANIFEST.MF;

* DSPDataPersistence/META-INF/MANIFEST.MF: adding the mongoDB Java driver i

@@ -1,6 +1,5 @@
 package org.netbeams.dsp.persistence.controller;
 
-import java.net.UnknownHostException;
 import java.util.List;
 import java.util.concurrent.Executors;
 import java.util.concurrent.ScheduledExecutorService;
@@ -22,14 +21,7 @@ import org.netbeams.dsp.message.QueryMessage;
 import org.netbeams.dsp.message.UpdateMessage;
 import org.netbeams.dsp.persistence.model.TransientPersistenceLayer;
 import org.netbeams.dsp.persistence.model.component.data.PersistentMessageUnit;
-import org.netbeams.dsp.persistence.model.component.data.PropertyKey;
-import org.netbeams.dsp.persistence.util.DSPJSONUtil;
 import org.netbeams.dsp.util.NetworkUtil;
-
-import com.mongodb.DBCollection;
-import com.mongodb.Mongo;
-import com.mongodb.MongoAdmin;
-import com.mongodb.MongoException;
 
 public class DSPDataPersistence implements DSPComponent {
 
@@ -94,30 +86,8 @@ public class DSPDataPersistence implements DSPComponent {
                     thLog.debug("Preparing to transfer messages " + tranMsgs.size() + " to database...");
 
                     for (PersistentMessageUnit tranMsg : tranMsgs) {
-
-                        DSPMongoCRUDService.insertMessageUnit(tranMsg);
-
+                        DSPMongoCRUDService.insertPersistentUnitMessageContents(tranMsg);
                         tranMsg.setStateToFlushed();
-
-                        // TODO: save the key_value pair into the persistent data store.
-                        // TODO: SEPARATE ALL THE PROPERTIES FROM EACH COLLECTION OF ITEMS
-                        /*
-                         * Imagine from JSON values... Write a parser or the driver that returns all the properties to
-                         * the application. { "sondeData": [ { "samplingTimeStamp": { "time": 1242009066438, "timezone":
-                         * "America/Los_Angeles" }, "temp": 87.0, "cond": 35.0, "resist": 45.0, "sal": 67.0, "press":
-                         * 443.0, "depth": 565.0, "ph": 12.0, "phmV": 65.0, "odoSat": 39.0, "odoConc": 88.0, "turbid":
-                         * 40.0, "battery": 5.0 }, { "temp": 87.0, "cond": 35.0, "resist": 45.0, "sal": 67.0, "press":
[…]

### 2009-09-21 versions/v2/apps/osgi-bundles/dsp/DSPDataPersistence/src/org/netbeams/dsp/persistence/controller/DSPMessageToMongoDBExperiment.java

@@ -1,0 +1,202 @@
+package org.netbeams.dsp.persistence.controller;
+
+import java.net.UnknownHostException;
+import java.text.SimpleDateFormat;
+import java.util.Date;
+import java.util.UUID;
+
+import org.netbeams.dsp.message.ComponentIdentifier;
+import org.netbeams.dsp.message.DSPMessagesFactory;
+import org.netbeams.dsp.message.Header;
+import org.netbeams.dsp.message.Message;
+import org.netbeams.dsp.message.MessageContent;
+import org.netbeams.dsp.persistence.model.component.data.PersistentMessageUnit;
+import org.netbeams.dsp.persistence.model.location.SensorLocation;
+import org.netbeams.dsp.ysi.SondeDataContainer;
+import org.netbeams.dsp.ysi.SondeDataType;
+import org.netbeams.dsp.ysi.SondeTestData;
+
+import com.commsen.stopwatch.Report;
+import com.commsen.stopwatch.Stopwatch;
+import com.mongodb.BasicDBObject;
+import com.mongodb.DBCollection;
+import com.mongodb.MongoException;
+
+/**
+ * The experiment to transfer DSP Messages to mongoDB. It uses random data of DSP Message Contents.
+ * 
+ * @author Marcello de Sales (marcello.sales@gmail.com)
+ *
+ */
+public class DSPMessageToMongoDBExperiment {
+
+    private static final SimpleDateFormat DATE_FORMATTER = new SimpleDateFormat("MM/dd/yyyy HH:mm:ss:SS");
+    /**
+     * Sleep interval for garbage collection
+     */
+    private static long fSLEEP_INTERVAL = 100;
+
+    /**
+     * Inserts the Message Content's instance into the mongoDB.
+     * 
+     * @param tranMsg is a persistent message unit from a DSP message transmitted by a sensor.
+     * @throws UnknownHostException
+     * @throws MongoException
+     */
+    public static void insertMessageUnit(PersistentMessageUnit tranMsg) throws UnknownHostException, MongoException {
+        long start = System.currentTimeMillis();
+        long startMemoryUse = getMemoryUse();
+        System.out.println("Started saving netbeams samples as mongodb objects at "
+                + DATE_FORMATTER.format(new Date(start)));
+
+        long watchTransaction = Stopwatch.start("Transaction time for inserting", "sonde-2-mongodb");
+        insertPersistentUnitMessageContents(tranMsg);
[…]

### 2009-09-21 versions/v2/apps/osgi-bundles/dsp/DSPDataPersistence/src/org/netbeams/dsp/persistence/controller/DSPMongoCRUDService.java

@@ -1,22 +1,13 @@
 package org.netbeams.dsp.persistence.controller;
 
 import java.net.UnknownHostException;
-import java.text.SimpleDateFormat;
-import java.util.Date;
 import java.util.HashMap;
 import java.util.Map;
-import java.util.UUID;
 
-import org.netbeams.dsp.message.ComponentIdentifier;
-import org.netbeams.dsp.message.DSPMessagesFactory;
-import org.netbeams.dsp.message.Header;
-import org.netbeams.dsp.message.Message;
 import org.netbeams.dsp.message.MessageContent;
 import org.netbeams.dsp.persistence.model.component.data.PersistentMessageUnit;
-import org.netbeams.dsp.persistence.model.location.SensorLocation;
 import org.netbeams.dsp.ysi.SondeDataContainer;
 import org.netbeams.dsp.ysi.SondeDataType;
-import org.netbeams.dsp.ysi.SondeTestData;
 
 import com.mongodb.BasicDBObject;
 import com.mongodb.DBCollection;
@@ -24,25 +15,34 @@ import com.mongodb.Mongo;
 import com.mongodb.MongoAdmin;
 import com.mongodb.MongoException;
 
+/**
+ * This class is responsible for the CRUD services with the mongoDB. Although a regular CRUD service includes the Update
+ * (U) of the CRUD, it will NOT be implemented as data in sensor networks DOES NOT CHANGE.
+ * 
+ * C - Create: Any document created will be created <li>sensor_ip_address - the ip of the sensor <li>message_id - the
+ * identification of the DSP message that transferred the message <li>transaction_time - the time when the data was
+ * received by the server. <li>fact_time - the time when the data was transferred from the sensor. <li>data - the actual
+ * sample data
+ * 
+ * R - Retrieve: Retrieve any created documents from the mongoDB given the message type U - Update: NOT Implemented as
+ * data can't be updated. D - Delete: Deletes an item from a collection.
+ * 
+ * @author Marcello de Sales (marcello.sales@gmail.com)
+ * 
+ */
 public class DSPMongoCRUDService {
 
-    private static final SimpleDateFormat DATE_FORMATTER = new SimpleDateFormat("MM/dd/yyyy HH:mm:ss:SS");
     /**
      * The default mongoDB database name for the project.
      */
     private static final String NETBEAMS_DATASTORE_NAME = "netbeams";
-    /**
-     * Sleep interval for garbage collection
-     */
-    private static long fSLEEP_INTERVAL = 100;
     /**
[…]

### 2009-09-21 versions/v2/apps/osgi-bundles/dsp/DSPSondeProducer/src/org/netbeams/dsp/ysi/SondeTestData.java

@@ -1,10 +1,10 @@
 package org.netbeams.dsp.ysi;
 
+import java.text.DecimalFormat;
+import java.text.SimpleDateFormat;
 import java.util.ArrayList;
 import java.util.Calendar;
 import java.util.List;
-import java.text.DecimalFormat;
-import java.text.SimpleDateFormat;
 
 /**
  * Generates random test data. The number of data must be specified during the creation of an instance of this class.

### 2009-09-29 docs/icccn2009/paper.tex
These are reviews and updates to the text before the camera-ready.
I've updated paragraphs that repeats the same contents in order
to be back to 6 pages.

git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@606 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 
-\author{\authorblockN{A. Puder, T. Johnson, K. Sales, M. de Sales}
+\author{\authorblockN{A. Puder, K. Sales, M. de Sales, T. Johnson}
 \authorblockA{San Francisco State University \\
@@
 San Francisco, CA 94132 \\
-EMail: \{arno$\mid$tlj$\mid$kleber$\mid$msales\}@sfsu.edu}
+EMail: \{arno$\mid$kleber$\mid$msales$\mid$tlj\}@sfsu.edu}
 \and
@@
 \begin{abstract}
-  In this paper we describe a sensor network for environmental
+ In this paper, we describe a sensor network for environmental
  monitoring based on highly specialized, off-the-shelf sensing
@@
 processed, archived, and made available to potential users in near
-real-time. In the following sections we give an overview of the NetBEAMS
+realtime. In the following sections we give an overview of the NetBEAMS
 architecture and implementation. Section \ref{SEC_BACKGROUND}
@@
 that the YSI is capable of are conductivity, temperature, chloride,
-ammonium, nitrate, turbidity, and chlorophyll.  It is under 55 cm in
-length, 8.9 cm in diameter, and weighs approximately 3.18 kg. It
-operates at temperatures between -5 to +50 deg C and at depths up to
-200 meters.  It uses 8 C-size Alkaline Batteries or External 12 VDC.
-Battery life can last up to 75 days depending on sensor configuration.
+among others. It is under 55 cm in length, 8.9 cm in diameter, and weighs
+approximately 3.18 kg, and operates at temperatures between -5 to +50 deg C and
+at depths up to 200 meters. It uses 8 C-size Alkaline Batteries or External 12 VDC,
+making the device last up to 75 days depending on the sensor configuration.
 The Sonde's bulkhead contains multiple pin connectors to support
@@
 MDS Display/Logger or a serial cable connected to a PC, where the sampling
-frequency is usually set to a high frequency. In contrast, Unattended
+frequency is usually set to a high frequency. In contrast, unattended
 mode is performed when the Sonde is deployed offshore or in a
@@
 over RS232. In this way, the Sonde can be configured to
-start logging data to an internal file. In order to view the data as it's
+start logging data to an internal file. In order to view the data as it is
 being captured, the option to show live data is chosen and the data can then be
@@
 however, it would be beneficial to have near-realtime access to the
-sensor measurements. We identify the following requirements that an
-end-to-end infrastructure supporting devices such as the YSI should
+sensor measurements. In this way, we identify the following requirements
+that an end-to-end infrastructure supporting devices such as the YSI should
 offer:
@@
 deployed over a geographically wide area, where the communication
-infrastructure is not always reliable or, at least, not coutinuously
+infrastructure is not always reliable or, at least, not continuously
 available. As the sensor network grows in size and complexity,
@@
 allowing customization via special-purpose modules. In our terminology,
-we call the basic foundation the \emph{Data Sensor Platform}
-(DSP). DSP is collection of local built-in and plug-in components; 
+we call this basic foundation the \emph{Data Sensor Platform}
+(DSP). DSP is a collection of local built-in and plug-in components;
 a sensor network is built from several DSPs that are linked
@@
 are self-contained modules that can be added and removed to a DSP
-installation at runtime. The basic communication unit within a DSP
+installation at run-time. The basic communication unit within a DSP
 is a message. DSP Components can exchange messages of different types.
 Actually, all interactions between DSP Components are implemented
-through messages yielding a simple but powerful communication
+through messages yielding a simple, yet powerful communication
 structure. Moreover, we distinguish between \emph{Data Producers} (DP) and
@@
 
-The DSP is responsible for routing messages between DPs and DCs. It is
-important to note that this routing only happens within a
-DSP. Following the paradigm of a micro-kernel approach, the DSP has no
-notion of remote communication or sensor-specific details. All this
-knowledge is embodied in special purpose DSP Components.
-
 \subsection{DSP Components}
@@
 external applications. One such common format used in the domain of
-environmental monitoring is OpenDAP \cite{opendap01}.  Since the
+environmental monitoring is OPEnDAP \cite{opendap01}. Since the
 converter module accepts messages from the DSP, it acts in the role of
@@
 effectively becomes a DSP Component that can be deployed on a DSP at
-runtime. A component is identified by its type (line 3). Only one
+run-time. A component is identified by its type (line 3). Only one
 component of a certain type per DSP is permissible. A type identifier
@@
 generated by some DP), it invokes the \texttt{deliver()} method (line
-8). A DP can send messages via a \texttt{DSP Broker} (a DSP Component itself) 
-obtained from \texttt{DSPContext} object that was
-passed to the DP during initialization.
+8). A DP can send messages via a \texttt{DSP Broker} through the 
+\texttt{DSPContext}, which is passed to the DP during initialization.
 
@@
 This decision is based on a set of routing rules (see Figure
-\ref{FIG_MATCHER_RULE}) which can be statically or dynamically
+\ref{FIG_MATCHER_RULE}), which can be statically or dynamically
 configured.
@@
 destination DC. Once the message is sent through the DSP, the matcher will
-determine which DCs should receive the message. Note that, a DP can
+determine which DCs should receive the message. Note that a DP can
 still specify a DC for its message. In this case, the matcher includes
@@
 
-The Gumstix runs Linux 2.6 with the BusyBox utilities, and uses the
-OpenEmbedded cross-compile build environment to provide a complete
+The Gumstix runs Linux 2.6 with the BusyBox utilities \cite{busybox01}, and uses
+the OpenEmbedded cross-compile build environment to provide a complete
 Linux environment and a large range of Linux applications for the ARM
 architecture. We use the JamVM \cite{jamvm01} implementation of Sun
-Microsystem's virtual machine in order to run the DSP. The Gumstix
-communicates with the YSI Sonde via an RS232 serial port using a
+Microsystem's virtual machine in order to run the DSP, together with
+the GNU Classpath \cite{classpath01} as the runtime libraries. 
+The Gumstix communicates with the YSI Sonde via an RS232 serial port using a
 null-modem cable, since both the Sonde and the Gumstix are DTE devices.
@@
 DSP Compoments are deployed as OSGi bundles. OSGi defines bundles life-cycle and
-runtime upgrade mechanism. DSP uses these capabilities to allow remote deployment 
+run-time upgrade mechanism. In this way, DSP uses these capabilities to allow remote deployment 
 (bundles with DSP Components are transmitted as message to the DSP Component Adaptor) 
-and activation/deactivation of DSP Components.
+and activation/deactivation of DSP Components. Our solution is implemented using 
+Knopflerfish \cite{knopflerfish01}, an Open-Source OSGi implementation.
 
-Our implementation of the DSP is based on the Open-Source
-implementation of OSGi called Knopflerfish \cite{knopflerfish01}.
-Since Sun Microsystems does not support its JDK for the Gumstix, we
-use JamVM \cite{jamvm01} as the Java virtual machine implementation
-and the GNU Classpath \cite{classpath01} project for the runtime libraries. In
-the following sections, we provide further details of some of the DSP
+In the following sections, we provide further details of some of the DSP
 Components that we have implemented as part of the NetBEAMS project.
@@
 change configuration parameters of various DSP Components via an
-easy-to-use web interface. With regards to the NetBEAMS specific
-components discussed in earlier sections, these configuration
-parameters are, for example, the sampling frequency of the YSI Sonde DP
-and the upload frequency of the the wire transport DC.
+easy-to-use web interface. As discussed in earlier sections, these 
+configuration parameters are, for example, the sampling frequency of the YSI
+Sonde DP or the upload frequency of the the wire transport DC.
 
 This configuration is performed by a special DSP Component that
-offers a web interface through which a technician can define the
+offers a web interface through which a technician can define
 configuration parameters (component \ding{197} in Figure
@@
 to allow the entry of user defined values. When those values are
-submitted, an update message is sent back to the remote component that
-will then change the appropriate configuration parameters.
+submitted, an update message is sent back to the remote component to be 
+adjusted accordingly.
 
@@
 %management groups, research scientists, municipalities, state and
-%federal agencies, who could use near real-time data streams to inform
+%federal agencies, who could use near realtime data streams to inform
 %decisions that impact public health, public safety, and the protection
@@
 %integrated network, depends on the first step in the data management
-%process: reliably obtaining data in near real-time from remotely
+%process: reliably obtaining data in near realtime from remotely
 %located sensors. This step can be the weak link in the data
@@
 key to our infrastructure is the DSP, a micro-kernel plug-and-play
-architecture that can flexibly be configured to specific needs through
+architecture that can be configured to specific needs through
 special purpose components. We plan to use the Gumstix's processing
@@
 communication capabilities \cite{Rellermeyer}, our DSP is designed to support asynchronous 
-communication in loosely coupled networks. The NetBEAMS infrastructure, as
-described in the paper, is available as an Open-Source project and can be 
-viewed in operation at \texttt{www.netbeams.org}.
+communication in loosely coupled networks. The NetBEAMS infrastructure is available as an Open-Source
+project and can be viewed in operation at \texttt{www.netbeams.org}.
 

### 2009-11-11 docs/marcello-dissertation/thesis/chapter1.tex
Drafts of the dissertation in latex format. The latex template was downloaded from the Math department. Next revision will update the format as mentioned by Niki.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@634 559edde2-a4d4-44e0-96fc-4035e40d7396

% main.tex, to be used with thesis.tex
% This contains the main work of your thesis.

%\bibliography{thesis}  % uses the references stored in Chapter1Radar.bib

\chapter{Introduction}

Sensor networks are commonly used in different areas for different purposes
such as scientific measurements and general population use
\cite{snintro01} \cite{snintro02}. In the scientific community, they serve as
tools to monitor the state of the environment by storing samples of data in specific 
periods of time. For instance, the NASA Jet Propulsion Laboratory uses the 
Volcano SensorWeb \cite{snex02}, a sensor network which aims to collect data
from specific volcanoes in Alaska and South Pacific to be used with heuristics 
regarding dangerous activities. Similarly, the National Data Buoy Center, a
department of the National Oceanic and Atmospheric Administration (NOAA)
\cite{snex03}, provides online data from different buoys in different coast
shores around the world regarding water quality, temperature, tides, etc, what 
characterizes sensor network for environmental monitoring \cite{snex01}.
Therefore, sensor networks can be seen as a very important scientific research and 
commercial field with direct application and use of its produced data.

In order to provide access to the collected sensor data, different approaches
can be taken to interrogate sensor devices. In general, the query process can
take place on the network nodes themselves, as it is called in-network, or in 
a centralized data sink. The former is generally used when the data is
collected at the sensor node on its local storage, being shown by \cite{sn-storage01}
\cite{sn-storage03} \cite{sn-storage04} that this strategy helps mitigating energy
consumption. However, this strategy is used when data archival is not the
primary reason, given the limited resources on the sensor devices. For this
reason, the the latter strategy has been used when the collected data is
primarily archived in a centralized database system for later reuse
\cite{sn-storage02}.

When it comes to data model implementation, the relational data model
\cite{db01} is one of the most used approaches to represent data in sensor 
network databases in any of the query processing strategies, and in most cases 
with a modified version of the SQL language \cite{sn-db-newop01}
\cite{sn-db-newop02}. However, the design of the database system must be done
with prior knowledge of the sensor types when  choosing the relational data
model, as the database design must be normalized \cite{db02} for the entities
already identified for the system. For instance, the introduction of a new
sensor device might represent a potential change on the database structure by
seeking a new normalized version of the database structure. In addition to
challenges of maintaining the database schema normalized, the use of Data
Provenance approaches in sensor network data \cite{sn-provenance01} usually
addresses the problem related with the lineage of data  and its full history.
However, once the data model is in place to receive the collected data, one of
the most important non-functional requirements in sensor networks is
scalability of the persistence storage. Since an operational sensor network
can produce huge amounts of data, it must be able to cope with the increase
and data load during data gathering, and yet provide users the same
performance during data retrieval by using different strategies such a
data-centric storage \cite{sn-storage03}.

In summary, this dissertation discusses the solution of a centralized
data-centric persistence storage component for the NetBEAMS \cite{nb01}, Data
Sensor Platform (DSP), an open-source \cite{se01} implementation of a sensor
network that primarily focused on environmental data produced by NetBEAMS. 
Different challenges can be related to the DSP in the context of sensor 
networks: data model, data persistence organization and strategy, and user data
access and distribution. First, NetBEAMS needs to provide means to safely 
represent sensor devices' properties without the need of restructuring any 
existing data model, taking into account the use cases from the main users. 
For example, a marine biologist may query the the persistence storage regarding
the temperature of a specific geographic region at the San Francisco Bay at a 
given time frame. Second, since a sensor network may experiment grown in terms 
of number of sensors, and consequently the number of data produced, the
persistence  storage must be able to reasonably scale and be well structured
in terms of storage space with the least service interruption. For instance,
if the persistence storage is close to reach its storage capacity, what
should the network administrator do? Last, but not least, the means to access 
the data for reuse is the most important features of the system, taking into 
account the different data formats used by the main users of the system. How 
to export specific collected data to a spreadsheet or archive files to be 
published or shared with other users? For these, and many other reasons, the 
persistence storage problem addressed by this work can be seen as a very 
challenging experience to solve.

In view of the fact that the persistence storage of a sensor network imposes
many different challenges, and taking into account the case study NetBEAMS, this
work's main strategy focused analyzing the different specifications of data
persistence in the literature in order to find new types of mechanisms to
provide data persistence for sensor network data. In this way, the kernel of
this dissertation is focused on providing solutions to the real-world problem
face by the Romberg Tiburon Center and its use of NetBEAMS. Furthermore,
the evaluation of different technologies is discussed in order to provide a
loosely-coupled component-based solution based on the infrastructure of the 
DSP Platform. For the reason of providing a 'lighter' data model, a subtype of
the Key-Value Pair \cite{kvp01} data model was chosen to easily capture the
description of any sensor device's properties, along with the use of time
series data models \cite{sn-provenance01} for time snapshots. In that way, data
modeling process does not represent a new refactoring each time a new sensor 
type is added or removed from the persistence system, giving users means to 
retrieve data accordingly. Similarly, in contemplation of scalability, this 
work proposes the use of an increasingly popular infrastructure technique called
Database Sharding \cite{dbshard01} \cite{dbshard02}, which takes the approach of
Data-Centric storage \cite{sn-storage03} on in-network query processing in a
centralized database system. In order to provide such mechanism, this work
evaluates the use of shard keys based on the geographic origin of the collected
data as cited to be one of the options when working with Provenance data in 
sensor networks \cite{sn-provenance01}. As a result, the data persistence system
component was conceived using the plug-and-play capability from the DSP 
platform without doing  any changes to the current DSP Platform, providing an
scalable solution that can adjust according to any data load increase. Finally,
the selected technology provides different data access mechanisms, which
directly addressed problems related to data access and Information Fusion 
\cite{sndp06}, providing export-import capabilities and visualization through
the availability of APIs written in different popular programming languages.

This dissertation report is organized as follows: the next chapter is an 
overview of the state of art on topics that supports the implementation of the
solution of this work, providing the necessary background foundation for the 
following chapters. Chapter 3 presents the design and architecture of the 
conceived persistence module, describing the requirements needed to be
addressed,  while Chapter 4 describes the implementation based on the design. 
The analysis and investigation of the implemented solution is shown through 
the analyzes of the experiment results on Chapter 5. Last, but not least, 
Chapter 6 presents the conclusions of this work, followed by the possible 
future works and recommendations presented on Chapter 7.

### 2009-11-11 docs/marcello-dissertation/thesis/chapter2.tex
Drafts of the dissertation in latex format. The latex template was downloaded from the Math department. Next revision will update the format as mentioned by Niki.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@634 559edde2-a4d4-44e0-96fc-4035e40d7396

% main.tex, to be used with thesis.tex
% This contains the main work of your thesis.

%\bibliography{thesis}  % uses the references stored in Chapter1Radar.bib

\chapter{State of the Art}

Different aspects related to two major fields of Computer Science have
motivated to the inception of this work. In the perspective of Database
Systems and Distributed Systems, this work seeks a different approach for a
persisting collected data from sensor networks into a storage system that can
be seen as a cluster of distributed databases. Likewise, in the perspective of
Software Engineering, this work adds a major capability to an existing sensor 
network framework, meeting the requirements and interfaces. In general, this 
chapter gives the foundation based on the motivations for the inception of a 
different approach to persisting data using a different data model, covering 
topics that will be referred throughout the following other chapters.

This chapter will first give an overview on different related problems
described  in the literature that solved problems related to data modeling and 
representation in sensor networks. Then, it presents the design, architecture 
and implementation details of the DSP Platform, which is the bases for this 
dissertation. Finally, this chapter will briefly give a tutorial on advanced 
topics such as the differences between the KVP data model and the relational 
data model, as well as discussing about the approaches used on Database
Sharding.

\section{Sensor Networks}

Sensor Networks are specialized types of network systems comprised of network 
nodes with specific goals to observe and collect data. The following
subsections describe classes of characteristis of these types of sensor
networks, with applications on research, industrial or militery.

\subsection{Deployment and Mobility}

Each of the nodes can be deployed in a random way such as being dropped
from an airplane, or in a manual fashion at deliberately chosen geographical 
points \cite{snintro01}. Moreover, the deployment can be done as a one-time
event or an iterative way.

However, nodes can be able to move around the deployed area or be static in
this place. In the case of the degree of the former behaviour, a sensor node
can move occasionally or continuously, being a result of an incidental 
side-effect or a desired property of the device, where they are an automotive
device or attached to another moving object. For this reason, \cite{snintro01} 
describes those types of mobility as active or passive.

\subsection{Device Size, Resources, Cost, Energy, Heterogeneity}

Such sensor nodes can be of different sizes, ranging from microscopically small
particles, small devices to bigger ones with a few feet cubic. In addition to
being limited in size, they are also limited in resources and computational
power, forcing their conception limited to have its energy stored in either
batteries or scavenged from solar cells \cite{snintro01}.

Since sensor devices range on size and capabilities, the cost for each of the
devices very from hundreds to thousands of dollars, depending on its
functionality, energy utilization, etc. For thie reason, \cite{snintro01}
explains that a sensor network may potentially contain similar sensor devices,
as they can contain dissimilar ones from each other, which characterizes the heterogeneity 
of a sensor network by the hardware capabilities such as GPS enabled or having
GSM  network connectivity.

\subsection{Communication Modality, Network Infrastructure, Coverage and
Connectivity}
\label{sec:sn-infrastructure}

Sensor nodes can communicate in different modalities such as radio, diffuse
light, inductive or sound, which may influence on the type of protocols used to
communicate among the nodes, where they can be said to be in an organized
infrastructure or as an ad hoc way \cite{snintro01}.

When it comes to the sensor network different organizational infrastructure,
they can differ in types of topology: single-hop, star, networked stars, tree
and graph, where each of them may affect the characteristics of the network as it 
may also influence on data gathering and routing mechanisms. While in a
single-hop sensor network there are only two levels of sensor connectivity, the
tree, graph or networked stars networks uses communications among the network
hosts for data communication.

Last, but not least, sensor networks may be influenced its coverage designed in
its purpose, its size in terms of number of nodes and its relating lifetime. As
sensors are deployed in different areas, they can be distributed in a sparse or
dense way, yet providing necessary coverage for the environment. However, 
considering that sensor devices are very likely to constant failures, projects 
have provided redundant nodes \cite{snintro01}. This degree of coverage also 
influences on the connectivity of the sensor nodes, where they can be
considered connected when a connection is available at all times. On the other
hand, nodes can communicate with other nodes in an intermittent way or sporadic
way. In the former, the sensor node is partitioned, while in the latter the
nodes are isolated most of the time and communicates at certain moments in
time.

\subsection{Network Size, Lifetime and Quality-Of-Service}

Similar to the degree of coverage of the sensor network is the number of
sensor nodes deployed in the network, where they can be in the order of
dozens, hundreds or even thousand of nodes \cite{snintro01}.

The application of the sensor network determines how long the sensor
network may be exist. For example, the lifetime of an environmental sensor
networks \cite{snex01} may be not determined since its service serves in a
regular basis, whereas the lifetime of a sensor network to determine the number
of people in an event is limited to the event.

Finally, the design of sensor networks can be related to the constraints
related to quality-of-service (QoS) such as real-time delivery within certain
period of time or robustness such as maintaining data alive even though there
a communication link is not available \cite{snintro02}.

\section{Data Persistence in Sensor Networks}

The most important purpose of a sensor network is related to its main product:
the use of the raw data. Data collection in sensor networks is directly
dependent on the specifications of the sensor network, its routing mechanisms
and the final destination of the produced data. In a nutshell,
\cite{sn-storage03} describes data as follows:

\begin{itemize}
  \item \textbf{low-level property} as the raw data collected from the
  sensor device; For example, the sea level as a number;
  \item \textbf{Observation or Event} as the processed data in format of the final
  observation. For instance, if the ocean is in its high tight or low tight.
\end{itemize}

\subsection{The Purpose of the Collected Data}
\label{sec:sn-data-purpose}

At first glance, the raw data serves as the primary product of the sensor
network. However, the first consideration that must take place is regarding the
purpose of the collected data, as \cite{sn-provenance01} explains
differentiates between two main use of the sensor data:

\begin{itemize}
  \item \textbf{Real-time Data Stream}, as the generated data from the sensor is
  actually not transmitted, but extracted from the device as it is requested.
  The data is still in its natural format specified by the sensor manufacturer;
  \item \textbf{Archival Data}, as the collected data have historical value or will
  be use by other applications at a future time. 
\end{itemize}

While real-time data stream are used and can potentially be discarded, the
archival data needs to be stored in a secondary memory. Nonetheless, given the
nature of the sensor device specification, the low-level data produced
by a given sensor device might not contain necessary information regarding the
identity of the data itself. The use of Data Provenance techniques is described
by \cite{sn-provenance01} as an important step to provide the collected data 
specific metadata regarding its location and lineage.

Under these circumstances, the collected data needs first to face its jorney
starting from its transmission from its producing device to the specified data 
sink as specified by \cite{sn-storage01} \cite{sn-storage02} \cite{sn-storage03}.

\subsubsection{Providing Identity to the Collected Sensor Data}

Most of sensor networks implementations use the collected data from the sensor
devices after they observe the environment. \cite{sn-provenance01} showed that
collected data used for archival purposes are usually missing important
features. First, the device's sampling are usually in raw format or without
obvious names. In order to provide search over the collected sensor data, the
data needs first to be described and indexed, and this process can be related
to Data Provenance \cite{db-provenance}.

An example of Data Provenance application is the use of aggregation of data
over time to estimate the changing effects of the weather in a given region.
For this reason, the collected data needs descriptions or annotations that can
give enough information to track changes, that is, the history of how and when
a given data has changed. The following should be the foundation of a
provenance-aware collected data according to \cite{sn-provenance01}:

\begin{itemize}
  \item each low-level observation as it describes the actual data;
  \item the location where the collected data was generated;
  \item the time when the data was collected.
\end{itemize}

\subsection{Storage Location for Collected Data}
\label{sec:sn-storage-locations}

Given that collected data can be used for different purposes as described in
section \ref{sec:sn-data-purpose}, the sensor node that produces the data is
responsible to send the raw data to a specified data storage destination. Upon
producing data, a sensor network node is specified to deal with the produced
data, where in most cases, \cite{sn-storage03} described that depending on the
sensor network infrastructure (see section \ref{sec:sn-infrastructure}),
different storage mechanisms are provided:

\begin{itemize}
  \item \textbf{External Storage}: the sensor device sends the produced data to
  another sensor node that contains an external storage. 
  \item \textbf{Local Storage}: as it is the case of real-time data stream, where data
  is allocated on the sensor device's main read-only memory.
\end{itemize}

It is clear that the data flows from the sensor node to another depending on
the specification of the network. On the whole, the sensor network
infrastructure dictates where the data will flow to, as described in section
\ref{sec:sn-infrastructure}. \cite{sn-storage02} sees a wireless sensor
networks as a many-to-one data gathering pattern, when its organization is
based on single-hop or star. On the other hand, when the data organization is
structured as a graph, tree, etc, the data flows from one intermediate sensor
node to another until it reaches its final storage node as described by
\cite{sn-storage01} \cite{sn-storage03}.

Last, but not least, yet a different storage strategy can be used: Data-Centric
Storage. This strategy aims at clustering the collected data based on a given nature of the
data such as location. Regardless the data storage mechanism used, the query
processing will depend on the sensor network infrastructure and data location.

\subsection{Query Processing for Storage Locations}

Given the different locations where the collected data might be as described in
section \ref{sec:sn-storage-locations}, different query processing for the
collected sensor data can be used as described by \cite{sn-storage03}:

\begin{itemize}
  \item {In-Network Query Processing}: sensor nodes that provides real-time
  data stream usually replies to requests to properties. Likewise, when the
  data is used for archival purpose, the data can be located at a network node
  specialized in storing the neighbors collected data. In this way, when a
  query is issued to the sensor network, the query processor will
  collectively verify every storage node for the collected data.
  \item {Centralized Query Processing}: a main network sink receives the
  collected data from the peer nodes. In this case, the query processor queries
  the solo centralized node for the overall values of the participating sensor
  devices.
\end{itemize}

There are pros and cons related to both of the approaches of query processing,
which are also related to the infrastructure, sensor network nature and 
requirements. The main goal of the in-network query processing is to decrease
the amount of data transmitted from the sensor devices to the final data sink, as
sensor devices are usually limited by constraints such as energy utilization 
\cite{sn-storage03}. Along with this strategy, the literature describes
different proposals that aims at upgrading the Structured Query Language (SQL)
adding location-sensitive clauses to the language \cite{sn-db-newop01}
\cite{sn-db-newop02}. However, this main problem related to this approach is
that it may potentially flood all nodes with query messages \cite{sn-storage04}.

In order to mitigate the problems related flooded nodes in in-network query
processing, the use of a Data-Centric storage strategy has been shown to be a
proven alternative for the sake of energy consumption \cite{sn-storage03}
\cite{sn-storage01}. Its main goal is to classify the collected data based in a
specific property, say location or event, in order to be sent to specific
storage nodes that relates to the observation. As the main outcome of this
strategy is to create network partitioning strategy, the query processor would
hit specific sensor nodes with the needed data, and therefore, would
decrease the load and maximize the capacity on the actual sensor devices.

A centralized network query processing processes data in a network node
specified as the network sink. The network sink usually is a network
node with a database server with the collection of data delivered by
all the sensor devices from the network. As a consequence of this
many-to-one communication strategy \cite{sn-storage02}, the query process does
not affect any of the sensor nodes in the network and, thus, maximizes the
sensor device utilization. Moreover, the use of the data for historical
purposes is easier to manage and give access to the primary end users of the
data, following the process of data processing, compression, aggregation, etc
\cite{sn-db02}.

Although this approach seems to be convenient to the point of view of the
sensor nodes and used as the tradeoff between achieving capacity and energy
consumption, the centralized data collector suffers from the annavoiable
creation of a point of traffic concentration \cite{sn-storage02}, also referred
to as the phenomenon of congestion, or funneling effect \cite{sn-storage04}.
For this reason, simple distributed system techniques such as the use of
master-slave or data replication can be used to decrease the overhead.

\section{SF-BEAMS: a Marine Sensor Network for Water Quality Monitoring}

The San Francisco Bay Environmental Assessment and Monitoring Station,
or SF-BEAMS, is a marine sensor network operated by the Romberg Tiburon Center
(RTC). This section details the SF-BEAMS sensor network located at
the Tiburon Island, California, and then presents the details of NetBEAMS, as
it is part of the case study of this work.

\subsection{The SF-BEAMS Infrastructure}

The RTC's main focuses are in the study of complex marine and estuarine
environments using the SF-BEAMS sensors deployed offshore of its pier located
on the banks of the San Francisco Bay at the Tiburon Island, California. In
this way, the project is responsible for providing data for water quality
monitoring, as well as weather and surface conditions.

In general, the SF-BEAMS network infrastructure contains heterogeneous wired and
wireless devices standing on pilons docked in the pier, where each of them is
responsible for observing different conditions of the area, as well as having
its own mechanisms for internal storage for the collected data. In this way,
data can be directly transferred to the labs via ethernet cables or collected
manually download by a staff's laptop computer. Figure \ref{fig:sf-beams}
shows a picture taken from the SF-BEAMS web-cam placed in-site.

\begin{figure}
  \centering
    \includegraphics{../diagrams/cam_image-oct15}
  \caption{Picture of the SF-BEAMS Network, at the RTC pier. Tiburon, CA.}
  \label{fig:sf-beams}
\end{figure}

\subsection{SF-BEAMS produced data and collection process}
\label{sec:sfbeams}

As described in section \ref{sec:sn-infrastructure}, sensor devices generates
its observed data based on properties of measurements defined by its
manufacturer. SF-BEAMS is a single-hop star sensor network containing
different instances of sensors, including the "YSI 6600 ESD V2", as seen on
figure \ref{fig:ysi-device}, which is a powerful water quality monitoring
device that produces around 52 bytes (13x4 Bytes) on a single real-time data
stream as shown on table \ref{tab:ysi-data-stream}.

\begin{figure}
  \centering
  \includegraphics{../diagrams/ysi-device}
  \caption{Picture of the case study sonde device: YSI 6600 ESD V2}
  \label{fig:ysi-device}
\end{figure}

\begin{table}
    \caption{YSI Data Stream}
    \begin{center}
        \begin{tabular}{lr}
  12.20    192    179 5588.40   0.09   0.084   0.059  7.98   -79.6   99.5  8.83 
  0.4     8.7
        \end{tabular}
    \end{center}
    \label{tab:ysi-data-stream}
\end{table}

In order to have access to the observed data, the network staff downloads
the data using one of the device's connection such as the RS-232 serial
connector \cite{rs232}. Then, the data is transferred to the RTC labs for
archival, where it is described, indexed and distributed to its main users.
According to its staff members, in May 2009 the infrastructure of the SF-BEAMS
included the following was organized as follows:

\begin{itemize}
  \item 5 YSI sondes in operation at the RTC site Pier in Tiburon Island, San
  Francisco Bay;
  \item Each YSI sonde produces 52 Bytes for each its observations;
  \item The sampling frequency rate could be configured in ranges of 1, 6 or
  15 minutes, depending on the staff's speficiation. 
\end{itemize}

Considering the infrastructure previously described, the size of the data
produced by the YSI sonde sensors at the RTC pier can estimated by its
measurement frequency as follows:

\begin{table}
    \label{tab:ysi-data-distribution}
    \caption{Amount of data produced by the RTC's YSI sondes}
        \begin{center}
        \begin{tabular}{|c|c|c|c|c|c|c|}\hline 
        \textbf{YSIs} & \textbf{Rate} & \textbf{Hourly} & \textbf{Daily} &
        \textbf{Weeky} & \textbf{Monthly} & \textbf{Yearly}\\\hline 
        1 & 1 min & 3.04 Kb & 73.12 Kb & 511.87 Kb & 1.99 Mb & 23.99 Mb\\\hline 
        5 & 6 min & 15.23 Kb & 365.62 Kb & 2.5 Mb & 9.99 Mb & 119.97 Mb\\\hline 
        1 & 15 min & 0.5 Kb & 12.18 Kb & 85.31 Kb & 341.25 Kb & 3.99 Mb\\\hline 
        5 & 1 min & 2.54 Kb & 60.93 Kb & 426.56 Kb & 1.67 Mb & 19.99 Mb\\\hline
        1 & 6 min & 0.2 Kb & 4.87 Kb & 34.12 Kb & 136.5 Kb & 1.6 Mb\\\hline 
        5 & 15 min & 1.0 Kb & 24.37 Kb & 170.62 Kb & 682.5 Kb & 7.99 Mb\\\hline
        \end{tabular}
        \end{center}
\end{table}

As it is shown, the data produced by this type of sensor runs in the order of
at most hundreds of Megabytes a year, what can be considered a very low data
storage requirement. Upon collecting the data from sensors, the RTC lab staff
use automation scripts written in Matlab to process, index and distribute the
raw data in different formats, including the OPeNDAP \cite{opendap}, a format
widely used in the research institutions that promotes easier data exchange
among them.

Although the execution of the RTC SF-BEAM sensor network produces its
execution, many operational challenges were faced by this type of
infrastructure, when it comes to the data utilization of wireless sensors that
are deployed in offshore conditions, \cite{netbeams2009} described the
operational challenges faced by the RTC staff and proposed NetBEAMS, a
component-based infrastructure that uses common-off-the-shelfs (COTS) embedded
systems and software to automate the process of data collection. Next section
describes how NetBEAMS is used to extract data from the SF-BEAMS sensor network.

\section{NetBEAMS: a Real-World Case Study on Environmental Sensor Network}

The Networked Bay Environmental Assessment and Monitoring System, or NetBEAMS, 
is a work-in-progress joint project proposed by the department of Computer
Science and the RTC at San Francisco State University. As described by
\cite{netbeams2009}, the solution can be used to automated the data gathering
process of the SF-BEAMS as described in section \ref{sec:sfbeams}. This
section fully details how NetBEAMS is executed in order to collect data from
SF-BEAMS, describing pertinent design and architectural details.

One of the motivations for the inception of NetBEAMS was the automation
of the data gathering and processing from the SF-BEAMS sondes without requiring
human intervention. Most importantly, the decrease with the costs of constant 
offshore site visits by boat, where biologists need manually collect the data
from the sensor devices in order to take them for further processing at the
lab. For this reason, the NetBEAMS research group proposed the use of the Data
Sensor Platform (DSP) \cite{netbeams2009}, as the second attempt to automate
the operation of SF-BEAMS.

This following sections details the DSP, its architecture and how data is
transported from the network nodes to the main network sink, the RTC labs.

\subsection{The Data Sensor Platform (DSP)}

The Data Sensor Platform, in short DSP Platform, is based on a micro-kernel
architecture developed on top of the Java modular framework called OSGi
\cite{osgi}. The platform represents the foundation of the execution of
NetBEAMS, since it is executed from an embedded hardware called Gumstix
\cite{gumstix} using a Linux operating system and on top of the Java Virtual
Machine (JVM). In order to transport data from sensors hosts to the network
sink, the use of a cellular connection is used to transport the collected data
from the senso devices using the HTTP Protocol, as described by
\cite{netbeams09}. Figure \ref{fig:sf-netbeams-node} shows the
architecture of the inhanced node designed by the NetBEAMS research group.

\begin{figure}
  \centering
  \includegraphics{../diagrams/DSP-Gateway-Node}
  \caption{The NetBEAMS Node enhancement with the Gumstix and the GSM Modem}
  \label{fig:sf-netbeams-node}
\end{figure}

The following list describes each of the devices in the NetBEAMS Node.

\begin{itemize}
  \item \textbf{YSI 6600EDS V2}: the YSI device is the regular sonde
  responsible for observing the environment. It is connected to the Gumstix
  embedded system, where the DSP is executed;
  \item \textbf{Gumstix console-vx}: it's a COTS ARM-based hardware that
  provides a computer-on-module environment for the development of small
  embedded systems. This is the main environment of data extraction and remote
  processing is accomplished. It runs a cross-compiled version of Linux Kernel
  2.6;
  \item \textbf{Huawei E220 USB Modem}: it is used to transfer the data from the
  Gumstix to the RTC Labs Data Center.
\end{itemize}

The DSP runs inside of the embedded system Gumstix. The following sections
details its design and process for extracting data.

\subsection{The DSP System Design}

This section focus on the development of a Software Platform for the NetBEAMS Gateway Embedded System, or on the system developed in the Gumstix device. As shown on Image 2.4.1.1, the NetBEAMS Gateway Node is an embedded system that contains a set of layer found in any computer. However, each of the resources described from the system were customized for the NetBEAMS system. The main components of such system can be summarized as follows:

    * Operating System: it uses a cross-compiled Gentoo Linux, which supports a wide variety of development tools, including the Java Virtual Machine (JVM), the underlying platform system;
    * Java Virtual Machine: it uses the JamVM, a ~200 KB version of the Sun Microsystems Java Virtual Machine, version 2.0;
    * OSGi Framework: it uses the Knopflerfish implementation of the OSGi 4.1 specification. More details in the following sections.
    * DSP Platform and other Bundles: The plug-and-play DSP Components are based on the OSGi bundles capabilities, which can reuse services registered in the OSGi framework.

### 2009-11-11 docs/marcello-dissertation/thesis/main.tex
Drafts of the dissertation in latex format. The latex template was downloaded from the Math department. Next revision will update the format as mentioned by Niki.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@634 559edde2-a4d4-44e0-96fc-4035e40d7396

% main.tex, to be used with thesis.tex
% This contains the main work of your thesis.

%\chapter*{Appendix A: Triangulations of Polytopes}
%\addcontentsline{toc}{chapter}{Appendix A: Triangulations of Polytopes}

\appendix{Appendix A: Triangulations of Polytopes}

This creates an appendix, which is not numbered (and therefore has to be added to the table of contents semi-manually).

For code, the {\tt verbatim} environment (maybe in conjunction with $\backslash${\tt include}) is helpful:

\begin{verbatim}
It reproduces text
  exactly
    as
      it
        is
typed.
\end{verbatim}

Good luck and enjoy writing!

### 2009-11-11 docs/marcello-dissertation/thesis/preamble.tex
Drafts of the dissertation in latex format. The latex template was downloaded from the Math department. Next revision will update the format as mentioned by Niki.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@634 559edde2-a4d4-44e0-96fc-4035e40d7396

% preamble.tex, to be used with thesis.tex
% This contains the TeX definitions for layout, style, etc., as well as the first few pages of your thesis: title page, copyright page, approval page, abstract, acknowledgments, tables of contents, tables, and figures.
% The layout commands should give the correct margins according to the graduate division's guidelines

%%%%% TeX class and packages

\documentclass[12pt,oneside]{sfsuthesis}

\usepackage{amsthm,amsmath,amssymb,amsfonts,latexsym,graphicx,enumerate,setspace,verbatim,tocloft}

%\usepackage{color}                    % For creating colored text and background
%\usepackage{hyperref}                 % For creating hyperlinks in cross references
% other possibly useful packages: textcomp,mathrsfs,amscd,epsfig,euscript,cancel

%%%%% Layout
% These numbers might depend on your printer. Check the margins and compare them to the Graduate Division's
% guidelines. If there's something off, try playing with the numbers...
%
% For chapters:
%     Must have a minimum of 1.5in margin on left and 1in on all other sides.  Where there are page numbers
%     (whether on top or bottom), must have one additional inch between the page number and the text, for a
%     total of 2in between the edge of the paper and the text.
% For frontmatter pages:
%     The same margin numbers generally work, except for the Title Page, so you will notice that we use
%     some numbers for \textheight and \footskip right here, and then change them below, right after
%     generating the Title Page.

\hoffset=.5in 
\oddsidemargin=0in   % = 1in because LaTeX adds 1in
\evensidemargin=0in  % = 1in because LaTeX adds 1in
\topmargin=0in       % = 1in because LaTeX adds 1in
\headheight=0in
\headsep=1in         % Distance from top of pagenum (for page numbers at top-right corner of page) to text
\footskip=1.2in      % Distance from bottom of text to the page number (for page number at bottom of page)
\textwidth=5.9in     % Should be 6in, but use 5.9in to be conservative
\textheight=8.0in    % Best for Title Page (will change after the Title Page)

\pagestyle{plain}

\doublespacing

%%%%% Style of theorems, definitions, examples, equations, etc.

\theoremstyle{plain} % Heading is bold, text italic.
\newtheorem{theorem}{Theorem}[chapter]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{conjecture}{Conjecture}[chapter]

\theoremstyle{definition}  % Heading is bold, text is roman
\newtheorem{definition}{Definition}[chapter]
\newtheorem{example}{Example}[chapter]

\theoremstyle{remark}  % Heading is italic, text is roman
\newtheorem*{remark}{Remark}
\newtheorem*{note}{Note}
\newtheorem{claim}{Claim}[chapter]

%%%%% Appendix style

\renewcommand\appendix[1]{
\chapter*{#1}
\addcontentsline{toc}{chapter}{#1}
}

%%%%% Source-Code listings

\usepackage{courier}
\usepackage{color}
\usepackage{xcolor}
\usepackage{listings}

\lstset{commentstyle=\tiny,captionpos=t,tabsize=2,frame=lines,keywordstyle=\color{blue}\tiny,commentstyle=\color{gray}\tiny,stringstyle=\color{red}\tiny,numbers=left,numberstyle=\tiny,breaklines=true,showstringspaces=false,basicstyle=\tiny,emph={label}}

\usepackage{caption}
\DeclareCaptionFont{white}{\color{white}}
\DeclareCaptionFormat{listing}{\colorbox{gray}{\parbox{\textwidth}{#1#2#3}}}
\captionsetup[lstlisting]{format=listing,labelfont=white,textfont=white}

%% Use fancy chapter headers, with Jos Dingjan's modifications,
%% plus my own tweaks. This style is not part of teTeX,
%% so we are using a local (and renamed) copy.
\usepackage[Lenny]{fncychap}

%Hifens

\usepackage[latin1]{inputenc}
\usepackage[USenglish]{babel}
\usepackage[dvips]{hyperref}

\begin{document}

\pagenumbering{roman}
\thispagestyle{empty}

\[ \]
\vspace{-1.9in}

\begin{center}
{\mytitle}

\vspace{1.4in}

\singlespace{A thesis presented to the faculty of\\
San Francisco State University\\
In partial fulfilment of\\
The requirements for\\ The degree}

\vspace{.5in}

\singlespace{Master of Science\\ In\\ Computer Science}

\vspace{3.1in}

{by \\[12pt] 
\myname \\[12pt]
San Francisco, California\\[12pt]
\thismonth
\thisyear}
\end{center}

\newpage
\textheight=7.1in    % For all pages after the Title Page -- try making this number smaller (7.0 or 6.9) if the bottom margins are too small
\footskip=1.1in      % Distance from bottom of text to the page number (for page number at bottom of page)
\thispagestyle{empty}

$\mbox{}$
\vspace{3in}
\begin{center}
\singlespace{
    Copyright \copyright{} \thisyear by \myname

    \bigskip

    All rights reserved. No part of the material protected by this
    copyright notice may be reproduced or utilized in any form or by any
    means, electronic or mechanical, including photocopying, recording or
    by any information storage and retrieval system, without the prior
    permission of the author.
}
\end{center}

\newpage
\thispagestyle{empty}
\[ \]
\vspace{-1.8in}
\begin{center}
{CERTIFICATION OF APPROVAL}
\end{center}
\vspace{.6in}
\begin{quote}
I certify that I have read {\it \mytitle} by \myname and that in my opinion this work meets the criteria for approving a thesis submitted in partial fulfillment of the requirements for the degree: Master of Science in Computer Science at San Francisco State University.
\end{quote}

\vspace{0.75in}

\hspace*{\fill}\parbox{3.5in}{
\singlespace{

\hrule{\hspace{3.5in}} \\ 
Prof. Arno Puder, Ph.D.\\
Professor of Computer Science\\

\vspace{1in}

\hrule{\hspace{3.5in}} \\
Prof. Marguerite C. Murphy, Ph.D.\\
Professor of Computer Science

}
}

\newpage
\thispagestyle{empty}
\[ \]
\vspace{-1.8in}
\begin{center}
{\mytitle} \\

\vspace{.5in}

\singlespace{
\myname \\
San Francisco State University \\ 
\thisyear \\
}

\end{center}

\vspace{.3in}

\doublespacing{\noindent
Your abstract comes here. % maximum one page
}

\vspace*{\fill}

\hspace*{\fill}

\noindent
I certify that the Abstract is a correct representation of the content of this thesis.

\vspace{.6in}

\hrule{\hspace{3.75in}} \\[-10pt]
Chair, Thesis Committee 
\hspace{2.5in}
Date

\newpage
\[ \]
\vspace{-1.8in}
\begin{center}{ACKNOWLEDGMENTS}\end{center}

\vspace{.3in}
\begin{quote}
\noindent
I thank everybody and their mothers\dots (optional).
\end{quote}

\renewcommand{\contentsname}{\vspace{-1.8in} \begin{center} \normalsize \rm TABLE OF CONTENTS \end{center}}
\renewcommand{\listfigurename}{\vspace{-1.8in} \begin{center} \normalsize \rm LIST OF FIGURES \end{center}}
\renewcommand{\listtablename}{\vspace{-1.8in} \begin{center} \normalsize \rm LIST OF TABLES \end{center}}
\renewcommand{\cftchapfont}{\normalfont}
\renewcommand{\cftchappagefont}{\normalfont}
\renewcommand{\cftchapleader}{\cftdotfill{\cftdotsep}} % formatting commands for table of contents
\renewcommand{\cftsecfont}{\normalfont}
\renewcommand{\cftsecpagefont}{\normalfont}
\renewcommand{\cftsecleader}{\cftdotfill{\cftdotsep}}

\newpage \tableofcontents 
\newpage \listoftables % comment out if you don't use tables
\newpage \listoffigures % comment out if you don't use figures

\newpage
\pagestyle{myheadings}
\pagenumbering{arabic} 
\setcounter{page}{1}

### 2009-11-11 docs/marcello-dissertation/thesis/source-code.tex
Drafts of the dissertation in latex format. The latex template was downloaded from the Math department. Next revision will update the format as mentioned by Niki.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@634 559edde2-a4d4-44e0-96fc-4035e40d7396

% main.tex, to be used with thesis.tex
% This contains the main work of your thesis.

%\chapter*{Appendix A: Triangulations of Polytopes}
%\addcontentsline{toc}{chapter}{Appendix A: Triangulations of Polytopes}

\appendix{Appendix A: DSP Data Persistence OSGi Bundle}

This section includes the source-code for the DSP Data Persistence. Each
section details one single main artifact used on the development of the
component. However, only the main ones will be added into this document, while
the remainder can be downloaded directly from the Subversion repository.

The source code is located in the Subversion repository for the NetBEAMS
project. This is how it is structured:

%    DSP_HOME - http://code.google.com/p/netbeams/source/browse/ main
%    repository for evaluation.
%    - $PERSISTENCE = DSP_HOME/browse/branches/marcello/persistence/ = defines
%    the project home for the persistence component.
%    
%    - PERSISTENCE/versions/v2/apps/osgi-bundles/dsp/DSPDataPersistence =
%    defines the directory for the persistence bundle.

\section{OSGi Bundle Build}

This artifact defines the ANT build file.

\lstinputlisting[language=Ant,label=persistence-builder,caption=Build
system using Apache
ANT]{../../../../netbeams/versions/v2/apps/osgi-bundles/dsp/DSPDataPersistence/build.xml}

\section{DSP Data Persistence Service}

\lstinputlisting[language=Java,label=persistence-component,caption=DSP
Data Persistence
Component]{../../../../netbeams/versions/v2/apps/osgi-bundles/dsp/DSPDataPersistence/src/org/netbeams/dsp/persistence/controller/DSPDataPersistence.java}

### 2009-11-11 docs/marcello-dissertation/thesis/chapter1.tex
* This revision fixes the problem related to the table of contents that was not printed.   - The package for hiper-link is responsible for the bug in the preamble.tex.
* References to the misc should be done by using \ref instead of \cite.

* Renaming a few bibliographic entries.

git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@635 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 such as scientific measurements and general population use
-\cite{snintro01} \cite{snintro02}. In the scientific community, they serve as
+\cite{sn-intro01} \cite{sn-intro02}. In the scientific community, they serve as
 tools to monitor the state of the environment by storing samples of data in specific 
 periods of time. For instance, the NASA Jet Propulsion Laboratory uses the 
-Volcano SensorWeb \cite{snex02}, a sensor network which aims to collect data
+Volcano SensorWeb \cite{sn-ex02}, a sensor network which aims to collect data
 from specific volcanoes in Alaska and South Pacific to be used with heuristics 
@@
 department of the National Oceanic and Atmospheric Administration (NOAA)
-\cite{snex03}, provides online data from different buoys in different coast
+\cite{sn-ex03}, provides online data from different buoys in different coast
 shores around the world regarding water quality, temperature, tides, etc, what 
-characterizes sensor network for environmental monitoring \cite{snex01}.
+characterizes sensor network for environmental monitoring \cite{sn-ex01}.
 Therefore, sensor networks can be seen as a very important scientific research and 
@@
 When it comes to data model implementation, the relational data model
-\cite{db01} is one of the most used approaches to represent data in sensor 
-network databases in any of the query processing strategies, and in most cases 
+\cite{relational-model} is one of the most used approaches to represent data in
+sensor network databases in any of the query processing strategies, and in most cases 
 with a modified version of the SQL language \cite{sn-db-newop01}
@@
 with prior knowledge of the sensor types when  choosing the relational data
-model, as the database design must be normalized \cite{db02} for the entities
+model, as the database design must be normalized \cite{db-normalization} for the entities
 already identified for the system. For instance, the introduction of a new
@@
 In summary, this dissertation discusses the solution of a centralized
-data-centric persistence storage component for the NetBEAMS \cite{nb01}, Data
-Sensor Platform (DSP), an open-source \cite{se01} implementation of a sensor
-network that primarily focused on environmental data produced by NetBEAMS. 
+data-centric persistence storage component for the NetBEAMS \cite{netbeams09},
+Data Sensor Platform (DSP), an open-source \cite{open-source} implementation of
+a sensor network that primarily focused on environmental data produced by NetBEAMS. 
 Different challenges can be related to the DSP in the context of sensor 
@@
 DSP Platform. For the reason of providing a 'lighter' data model, a subtype of
-the Key-Value Pair \cite{kvp01} data model was chosen to easily capture the
+the Key-Value Pair \cite{db-kvp01} data model was chosen to easily capture the
 description of any sensor device's properties, along with the use of time
@@
 work proposes the use of an increasingly popular infrastructure technique called
-Database Sharding \cite{dbshard01} \cite{dbshard02}, which takes the approach of
-Data-Centric storage \cite{sn-storage03} on in-network query processing in a
+Database Sharding \cite{db-shard01} \cite{db-shard02}, which takes the approach
+of Data-Centric storage \cite{sn-storage03} on in-network query processing in a
 centralized database system. In order to provide such mechanism, this work
@@
 directly addressed problems related to data access and Information Fusion 
-\cite{sndp06}, providing export-import capabilities and visualization through
+\cite{sn-info-fusion}, providing export-import capabilities and visualization through
 the availability of APIs written in different popular programming languages.

### 2009-11-11 docs/marcello-dissertation/thesis/chapter2.tex
* This revision fixes the problem related to the table of contents that was not printed.   - The package for hiper-link is responsible for the bug in the preamble.tex.
* References to the misc should be done by using \ref instead of \cite.

* Renaming a few bibliographic entries.

git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@635 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 from an airplane, or in a manual fashion at deliberately chosen geographical 
-points \cite{snintro01}. Moreover, the deployment can be done as a one-time
+points \cite{sn-intro01}. Moreover, the deployment can be done as a one-time
 event or an iterative way.
@@
 side-effect or a desired property of the device, where they are an automotive
-device or attached to another moving object. For this reason, \cite{snintro01} 
+device or attached to another moving object. For this reason, \cite{sn-intro01} 
 describes those types of mobility as active or passive.
@@
 power, forcing their conception limited to have its energy stored in either
-batteries or scavenged from solar cells \cite{snintro01}.
+batteries or scavenged from solar cells \cite{sn-intro01}.
 
@@
 devices very from hundreds to thousands of dollars, depending on its
-functionality, energy utilization, etc. For thie reason, \cite{snintro01}
+functionality, energy utilization, etc. For thie reason, \cite{sn-intro01}
 explains that a sensor network may potentially contain similar sensor devices,
@@
 communicate among the nodes, where they can be said to be in an organized
-infrastructure or as an ad hoc way \cite{snintro01}.
+infrastructure or as an ad hoc way \cite{sn-intro01}.
 
@@
 considering that sensor devices are very likely to constant failures, projects 
-have provided redundant nodes \cite{snintro01}. This degree of coverage also 
+have provided redundant nodes \cite{sn-intro01}. This degree of coverage also 
 influences on the connectivity of the sensor nodes, where they can be
@@
 sensor nodes deployed in the network, where they can be in the order of
-dozens, hundreds or even thousand of nodes \cite{snintro01}.
+dozens, hundreds or even thousand of nodes \cite{sn-intro01}.
 
@@
 network may be exist. For example, the lifetime of an environmental sensor
-networks \cite{snex01} may be not determined since its service serves in a
+networks \cite{sn-ex01} may be not determined since its service serves in a
 regular basis, whereas the lifetime of a sensor network to determine the number
@@
 period of time or robustness such as maintaining data alive even though there
-a communication link is not available \cite{snintro02}.
+a communication link is not available \cite{sn-intro02}.
 
@@
 use automation scripts written in Matlab to process, index and distribute the
-raw data in different formats, including the OPeNDAP \cite{opendap}, a format
+raw data in different formats, including the OPeNDAP \ref{opendap}, a format
 widely used in the research institutions that promotes easier data exchange
@@
 architecture developed on top of the Java modular framework called OSGi
-\cite{osgi}. The platform represents the foundation of the execution of
+\ref{osgi}. The platform represents the foundation of the execution of
 NetBEAMS, since it is executed from an embedded hardware called Gumstix
-\cite{gumstix} using a Linux operating system and on top of the Java Virtual
+\ref{gumstix} using a Linux operating system and on top of the Java Virtual
 Machine (JVM). In order to transport data from sensors hosts to the network

### 2009-11-11 docs/marcello-dissertation/thesis/chapter2.tex
Bugs in the bibtex file were preventing the references to be printed.
This revision also carries the format submitted by Niki
 - preamble and sfsuthesis format

git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@636 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 
-%\bibliography{thesis}  % uses the references stored in Chapter1Radar.bib
+\bibliography{thesis}  % uses the references stored in Chapter1Radar.bib
 
@@
 use automation scripts written in Matlab to process, index and distribute the
-raw data in different formats, including the OPeNDAP \ref{opendap}, a format
+raw data in different formats, including the OPeNDAP \cite{opendap}, a format
 widely used in the research institutions that promotes easier data exchange
@@
 architecture developed on top of the Java modular framework called OSGi
-\ref{osgi}. The platform represents the foundation of the execution of
+\cite{osgi}. The platform represents the foundation of the execution of
 NetBEAMS, since it is executed from an embedded hardware called Gumstix
-\ref{gumstix} using a Linux operating system and on top of the Java Virtual
+\cite{gumstix} using a Linux operating system and on top of the Java Virtual
 Machine (JVM). In order to transport data from sensors hosts to the network
@@
 from the senso devices using the HTTP Protocol, as described by
-\cite{netbeams09}. Figure \ref{fig:sf-netbeams-node} shows the
+\cite{netbeams2009}. Figure \ref{fig:sf-netbeams-node} shows the
 architecture of the inhanced node designed by the NetBEAMS research group.

### 2009-11-11 docs/marcello-dissertation/thesis/preamble.tex
* This revision fixes the problem related to the table of contents that was not printed.   - The package for hiper-link is responsible for the bug in the preamble.tex.
* References to the misc should be done by using \ref instead of \cite.

* Renaming a few bibliographic entries.

git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@635 559edde2-a4d4-44e0-96fc-4035e40d7396
Bugs in the bibtex file were preventing the references to be printed.
This revision also carries the format submitted by Niki
 - preamble and sfsuthesis format

git-svn-id: svn://localhost/netbeams/branches/mar

@@
 \documentclass[12pt,oneside]{sfsuthesis}
-
-\usepackage{amsthm,amsmath,amssymb,amsfonts,latexsym,graphicx,enumerate,setspace,verbatim,tocloft}
-
-%\usepackage{color}                    % For creating colored text and background
-%\usepackage{hyperref}                 % For creating hyperlinks in cross references
+\usepackage{amsthm,amsmath,amssymb,amsfonts,latexsym,graphicx,enumerate,setspace,verbatim,makeidx,subfigure, multirow, url, soul}
+\usepackage[subfigure]{tocloft}
+\usepackage{color}                       % For creating colored text and background
+%\usepackage{hyperref}                 % For creating hyperlinks in cross
+%references
 % other possibly useful packages: textcomp,mathrsfs,amscd,epsfig,euscript,cancel
@@
 \hoffset=.5in 
-\oddsidemargin=0in   % = 1in because LaTeX adds 1in
-\evensidemargin=0in  % = 1in because LaTeX adds 1in
-\topmargin=0in       % = 1in because LaTeX adds 1in
+\oddsidemargin=0in
+\evensidemargin=0in
+\topmargin=0.1in 
 \headheight=0in
-\headsep=1in         % Distance from top of pagenum (for page numbers at top-right corner of page) to text
-\footskip=1.2in      % Distance from bottom of text to the page number (for page number at bottom of page)
-\textwidth=5.9in     % Should be 6in, but use 5.9in to be conservative
-\textheight=8.0in    % Best for Title Page (will change after the Title Page)
+\headsep=1in
+\footskip=.8in
+\textwidth=5.9in
+\textheight=7in
+
+\setcounter{secnumdepth}{3}
+\setcounter{tocdepth}{3}
 
@@
 \usepackage[USenglish]{babel}
-\usepackage[dvips]{hyperref}
 
@@
 \[ \]
-\vspace{-1.9in}
+\vspace{-1.8in}
 
@@
 San Francisco State University\\
-In partial fulfilment of\\
+In partial fulfillment of\\
 The requirements for\\ The degree}
@@
 
-\vspace{3.1in}
+\vspace*{\fill}
 
@@
 \newpage
-\textheight=7.1in    % For all pages after the Title Page -- try making this number smaller (7.0 or 6.9) if the bottom margins are too small
-\footskip=1.1in      % Distance from bottom of text to the page number (for page number at bottom of page)
 \thispagestyle{empty}
@@
 \end{center}
-\vspace{.6in}
+\vspace{.5in}
 \begin{quote}
-I certify that I have read {\it \mytitle} by \myname and that in my opinion this work meets the criteria for approving a thesis submitted in partial fulfillment of the requirements for the degree: Master of Science in Computer Science at San Francisco State University.
+I certify that I have read {\it \mytitle} by \myname, and that in my opinion this work meets the criteria for approving a thesis submitted in partial fulfillment of the requirements for the degree: Master of Science in Computer Science at San Francisco State University.
 \end{quote}
 
-\vspace{0.75in}
+\vspace{1.5in}
 
@@
 Professor of Computer Science\\
-
 \vspace{1in}
-
 \hrule{\hspace{3.5in}} \\
@@
 \myname \\
-San Francisco State University \\ 
+San Francisco, California \\  
 \thisyear \\
@@
 
-\vspace{.3in}
+\vspace{.5in}
 
-\doublespacing{\noindent
-Your abstract comes here. % maximum one page
+%\doublespacing{\noindent
+\onehalfspacing{\noindent      
+Abstract goes here
 }
@@
 
-\vspace{.6in}
+\vspace{.5in}
 
@@
 Date
-
+\doublespacing
 \newpage
@@
 
-\vspace{.3in}
+\vspace{.5in}
 \begin{quote}
 \noindent
-I thank everybody and their mothers\dots (optional).
+Acknowledgments
 \end{quote}
 
-\renewcommand{\contentsname}{\vspace{-1.8in} \begin{center} \normalsize \rm TABLE OF CONTENTS \end{center}}
-\renewcommand{\listfigurename}{\vspace{-1.8in} \begin{center} \normalsize \rm LIST OF FIGURES \end{center}}
-\renewcommand{\listtablename}{\vspace{-1.8in} \begin{center} \normalsize \rm LIST OF TABLES \end{center}}
+\renewcommand{\contentsname}{\vspace{-1.7in} \begin{center} \normalsize \rm TABLE OF CONTENTS \end{center}}
+\renewcommand{\listfigurename}{\vspace{-1.7in} \begin{center} \normalsize \rm LIST OF FIGURES \end{center}}
+\renewcommand{\listtablename}{\vspace{-1.7in} \begin{center} \normalsize \rm LIST OF TABLES \end{center}}
 \renewcommand{\cftchapfont}{\normalfont}

### 2009-11-14 docs/marcello-dissertation/thesis/chapter1-introduction.tex
Updating the thesis and added chapters 2-5. Still revisions are needed in section 4 and 5, adjusting images and tables. Images are in the eps format.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@642 559edde2-a4d4-44e0-96fc-4035e40d7396

--- 2009-11-11 docs/marcello-dissertation/thesis/chapter1.tex
@@
 Sensor networks are commonly used in different areas for different purposes
-such as scientific measurements and general population use
-\cite{sn-intro01} \cite{sn-intro02}. In the scientific community, they serve as
-tools to monitor the state of the environment by storing samples of data in specific 
-periods of time. For instance, the NASA Jet Propulsion Laboratory uses the 
-Volcano SensorWeb \cite{sn-ex02}, a sensor network which aims to collect data
-from specific volcanoes in Alaska and South Pacific to be used with heuristics 
+such as scientific measurements and general population use \cite{sn-intro01}
+\cite{sn-intro02}. In the scientific community, they serve as tools to monitor
+the state of the environment by storing samples of data in specific periods of
+time. For instance, the NASA Jet Propulsion Laboratory uses the Volcano
+SensorWeb \cite{sn-ex02}, a sensor network which aims to collect data from
+specific volcanoes in Alaska and South Pacific to be used with heuristics
 regarding dangerous activities. Similarly, the National Data Buoy Center, a
@@
 characterizes sensor network for environmental monitoring \cite{sn-ex01}.
-Therefore, sensor networks can be seen as a very important scientific research and 
-commercial field with direct application and use of its produced data.
+Therefore, sensor networks can be seen as a very important scientific research
+and commercial field with direct application and use of its produced data.
 
 In order to provide access to the collected sensor data, different approaches
-can be taken to interrogate sensor devices. In general, the query process can
-take place on the network nodes themselves, as it is called in-network, or in 
-a centralized data sink. The former is generally used when the data is
-collected at the sensor node on its local storage, being shown by \cite{sn-storage01}
-\cite{sn-storage03} \cite{sn-storage04} that this strategy helps mitigating energy
-consumption. However, this strategy is used when data archival is not the
-primary reason, given the limited resources on the sensor devices. For this
-reason, the the latter strategy has been used when the collected data is
-primarily archived in a centralized database system for later reuse
-\cite{sn-storage02}.
+can be taken to interrogate sensor devices indirectly by accessing its data in
+places called network sink. These data nodes can be located in the network or 
+in a centralized host. In this way, the query process can take place on the
+network nodes themselves or on the centralized data sink. The former is
+generally used when the data is collected at the sensor node on its local
+storage, being shown by \cite{sn-storage01} \cite{sn-storage03}
+\cite{sn-storage04} that this strategy helps mitigating energy consumption.
+However, this strategy is used when data archival is not the primary reason,
+given the limited resources on the sensor devices. For this reason, the the
+latter strategy has been used when the collected data is primarily archived in
+a centralized database system for later reuse \cite{sn-storage02}.
 
@@
 sensor network databases in any of the query processing strategies, and in most cases 
-with a modified version of the SQL language \cite{sn-db-newop01}
-\cite{sn-db-newop02}. However, the design of the database system must be done
-with prior knowledge of the sensor types when  choosing the relational data
-model, as the database design must be normalized \cite{db-normalization} for the entities
-already identified for the system. For instance, the introduction of a new
-sensor device might represent a potential change on the database structure by
-seeking a new normalized version of the database structure. In addition to
-challenges of maintaining the database schema normalized, the use of Data
-Provenance approaches in sensor network data \cite{sn-provenance01} usually
-addresses the problem related with the lineage of data  and its full history.
-However, once the data model is in place to receive the collected data, one of
-the most important non-functional requirements in sensor networks is
-scalability of the persistence storage. Since an operational sensor network
-can produce huge amounts of data, it must be able to cope with the increase
-and data load during data gathering, and yet provide users the same
-performance during data retrieval by using different strategies such a
-data-centric storage \cite{sn-storage03}.
+with a modified version of the SQL language \cite{sn-db-newop}. However, the
+design of the database system must be done with prior knowledge of the sensor
+types when choosing the relational data model, as the database design must be
+normalized \cite{db-normalization} for the entities already identified for the
+system. For instance, the introduction of a new sensor device might represent
+a potential change on the database structure by seeking a new normalized
+version of the database structure. In addition to challenges of maintaining
+the database schema normalized, the use of Data Provenance approaches in
+sensor network data \cite{sn-provenance} usually addresses the problem related
+with the lineage of data  and its full history. However, once the data model
+is chosen and implemented, one of the most important non-functional
+requirements in sensor networks is scalability of the persistence storage.
+Since an operational sensor network can produce huge amounts of data, it must
+be able to cope with the increase and data load during data gathering, and yet
+provide users the same performance during data retrieval by using different
+strategies such a data-centric storage \cite{sn-storage03}.
 
-In summary, this dissertation discusses the solution of a centralized
-data-centric persistence storage component for the NetBEAMS \cite{netbeams09},
-Data Sensor Platform (DSP), an open-source \cite{open-source} implementation of
-a sensor network that primarily focused on environmental data produced by NetBEAMS. 
-Different challenges can be related to the DSP in the context of sensor 
-networks: data model, data persistence organization and strategy, and user data
-access and distribution. First, NetBEAMS needs to provide means to safely 
-represent sensor devices' properties without the need of restructuring any 
-existing data model, taking into account the use cases from the main users. 
-For example, a marine biologist may query the the persistence storage regarding
-the temperature of a specific geographic region at the San Francisco Bay at a 
-given time frame. Second, since a sensor network may experiment grown in terms 
-of number of sensors, and consequently the number of data produced, the
-persistence  storage must be able to reasonably scale and be well structured
-in terms of storage space with the least service interruption. For instance,
-if the persistence storage is close to reach its storage capacity, what
-should the network administrator do? Last, but not least, the means to access 
-the data for reuse is the most important features of the system, taking into 
-account the different data formats used by the main users of the system. How 
-to export specific collected data to a spreadsheet or archive files to be 
-published or shared with other users? For these, and many other reasons, the 
-persistence storage problem addressed by this work can be seen as a very 
-challenging experience to solve.
+In summary, this dissertation discusses the solution of a
+centralized data-centric persistence storage component for the NetBEAMS
+\cite{netbeams2009}, an implementation of a sensor network that primarily
+focused on environmental data. Different challenges can be related to the
+selection of a technology that handles data persistence in the context of sensor
+networks: the network infrastructure, the nature and location of the collected
+data, the chosen data model, among others. First, NetBEAMS needs to provide
+means to safely represent sensor devices' properties without the need of
+restructuring any existing data model, taking into account the use cases from
+the main users. For example, a marine biologist may query the the persistence
+storage regarding the temperature of a specific geographic region at the San
+Francisco Bay at a given time frame. Second, since a sensor network may
+experiment grown in terms of number of sensors, and consequently the number of
+data produced, the persistence  storage must be able to reasonably scale and
+be well structured in terms of storage space with the least service
+interruption. For instance, if the persistence storage is close to reach its
+storage capacity, what should the network administrator do? Last, but not
+least, the means to access the data for reuse is the most important features
+of the system, taking into account the different data formats used by the main
+users of the system. How to export specific collected data to a spreadsheet or
+archive files to be published or shared with other users? For these, and many
+other reasons, the persistence storage problem addressed by this work can be
+seen as a very challenging experience to solve.
 
@@
 DSP Platform. For the reason of providing a 'lighter' data model, a subtype of
-the Key-Value Pair \cite{db-kvp01} data model was chosen to easily capture the
+the Key-Value Pair \cite{db-kvp} data model was chosen to easily capture the
 description of any sensor device's properties, along with the use of time
-series data models \cite{sn-provenance01} for time snapshots. In that way, data
+series data models \cite{sn-provenance} for time snapshots. In that way, data
 modeling process does not represent a new refactoring each time a new sensor 
 type is added or removed from the persistence system, giving users means to 
-retrieve data accordingly. Similarly, in contemplation of scalability, this 
-work proposes the use of an increasingly popular infrastructure technique called
-Database Sharding \cite{db-shard01} \cite{db-shard02}, which takes the approach
-of Data-Centric storage \cite{sn-storage03} on in-network query processing in a
-centralized database system. In order to provide such mechanism, this work
-evaluates the use of shard keys based on the geographic origin of the collected
-data as cited to be one of the options when working with Provenance data in 
-sensor networks \cite{sn-provenance01}. As a result, the data persistence system
-component was conceived using the plug-and-play capability from the DSP 
-platform without doing  any changes to the current DSP Platform, providing an
-scalable solution that can adjust according to any data load increase. Finally,
-the selected technology provides different data access mechanisms, which
-directly addressed problems related to data access and Information Fusion 
-\cite{sn-info-fusion}, providing export-import capabilities and visualization through
-the availability of APIs written in different popular programming languages.
+retrieve data accordingly. Similarly, in contemplation of scalability and the
+use of Data-Centric approach \cite{sn-storage03}, this work proposes the use of
+Database Sharding \cite{db-shard01} \cite{db-shard02}, on a centralized
+database system.
+
+In order to provide such mechanism, this work first takes a look at the related
+literature to study the properties of sensor networks, in the context of its
+produced data and infrastructure for storage and query processing. Then, a
+summarized list of taxonomies is proposed to be the basis of an empirical  the
+use of shard keys based on the geographic origin of the collected data as cited to be one of the options when working with Provenance data in sensor networks \cite{sn-provenance}.
+As a result, the data persistence system component was conceived using the plug-and-play capability from the DSP platform without doing  any changes to the current DSP Platform, providing an scalable solution that can adjust according to any data load increase.
+Finally, the selected technology provides different data access
+mechanisms, which directly addressed problems related to data access and
+Information Fusion \cite{sn-info-fusion}, providing export-import capabilities
+and visualization through the availability of APIs written in different
+popular programming languages.
 

### 2009-11-14 docs/marcello-dissertation/thesis/chapter2-state-of-the-art.tex
Updating the thesis and added chapters 2-5. Still revisions are needed in section 4 and 5, adjusting images and tables. Images are in the eps format.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@642 559edde2-a4d4-44e0-96fc-4035e40d7396

--- 2009-11-11 docs/marcello-dissertation/thesis/chapter2.tex
@@
 
-\section{Sensor Networks}
+\section{Sensor Networks Survey}
 
@@
 
-\section{Data Persistence in Sensor Networks}
+\section{Persistence Storage for Sensor Networks}
 
@@
 network. However, the first consideration that must take place is regarding the
-purpose of the collected data, as \cite{sn-provenance01} explains
+purpose of the collected data, as \cite{sn-provenance} explains
 differentiates between two main use of the sensor data:
@@
 identity of the data itself. The use of Data Provenance techniques is described
-by \cite{sn-provenance01} as an important step to provide the collected data 
+by \cite{sn-provenance} as an important step to provide the collected data 
 specific metadata regarding its location and lineage.
@@
 sink as specified by \cite{sn-storage01} \cite{sn-storage02} \cite{sn-storage03}.
-
-\subsubsection{Providing Identity to the Collected Sensor Data}
-
-Most of sensor networks implementations use the collected data from the sensor
-devices after they observe the environment. \cite{sn-provenance01} showed that
-collected data used for archival purposes are usually missing important
-features. First, the device's sampling are usually in raw format or without
-obvious names. In order to provide search over the collected sensor data, the
-data needs first to be described and indexed, and this process can be related
-to Data Provenance \cite{db-provenance}.
-
-An example of Data Provenance application is the use of aggregation of data
-over time to estimate the changing effects of the weather in a given region.
-For this reason, the collected data needs descriptions or annotations that can
-give enough information to track changes, that is, the history of how and when
-a given data has changed. The following should be the foundation of a
-provenance-aware collected data according to \cite{sn-provenance01}:
-
-\begin{itemize}
-  \item each low-level observation as it describes the actual data;
-  \item the location where the collected data was generated;
-  \item the time when the data was collected.
-\end{itemize}
 
@@
 \subsection{Query Processing for Storage Locations}
+\label{sec:query-process}
 
@@
 \begin{itemize}
-  \item {In-Network Query Processing}: sensor nodes that provides real-time
-  data stream usually replies to requests to properties. Likewise, when the
+  \item \textbf{In-Network Query Processing}: sensor nodes that provides
+  real-time data stream usually replies to requests to properties. Likewise, when the
   data is used for archival purpose, the data can be located at a network node
@@
   collectively verify every storage node for the collected data.
-  \item {Centralized Query Processing}: a main network sink receives the
+  \item \textbf{Centralized Query Processing}: a main network sink receives the
   collected data from the peer nodes. In this case, the query processor queries
@@
 different proposals that aims at upgrading the Structured Query Language (SQL)
-adding location-sensitive clauses to the language \cite{sn-db-newop01}
-\cite{sn-db-newop02}. However, this main problem related to this approach is
-that it may potentially flood all nodes with query messages \cite{sn-storage04}.
+adding location-sensitive clauses to the language \cite{sn-db-newop}. However,
+this main problem related to this approach is that it may potentially flood
+all nodes with query messages \cite{sn-storage04}.
 
@@
 data, following the process of data processing, compression, aggregation, etc
-\cite{sn-db02}.
+\cite{sn-db-modeling02}.
 
@@
 
-\section{SF-BEAMS: a Marine Sensor Network for Water Quality Monitoring}
+\section{Data Models and Query Engines for Sensor Networks}
+\label{sec:data-models}
 
-The San Francisco Bay Environmental Assessment and Monitoring Station,
-or SF-BEAMS, is a marine sensor network operated by the Romberg Tiburon Center
-(RTC). This section details the SF-BEAMS sensor network located at
-the Tiburon Island, California, and then presents the details of NetBEAMS, as
-it is part of the case study of this work.
-
-\subsection{The SF-BEAMS Infrastructure}
-
-The RTC's main focuses are in the study of complex marine and estuarine
-environments using the SF-BEAMS sensors deployed offshore of its pier located
-on the banks of the San Francisco Bay at the Tiburon Island, California. In
-this way, the project is responsible for providing data for water quality
-monitoring, as well as weather and surface conditions.
-
-In general, the SF-BEAMS network infrastructure contains heterogeneous wired and
-wireless devices standing on pilons docked in the pier, where each of them is
-responsible for observing different conditions of the area, as well as having
-its own mechanisms for internal storage for the collected data. In this way,
-data can be directly transferred to the labs via ethernet cables or collected
-manually download by a staff's laptop computer. Figure \ref{fig:sf-beams}
-shows a picture taken from the SF-BEAMS web-cam placed in-site.
-
-\begin{figure}
-  \centering
-    \includegraphics{../diagrams/cam_image-oct15}
-  \caption{Picture of the SF-BEAMS Network, at the RTC pier. Tiburon, CA.}
-  \label{fig:sf-beams}
-\end{figure}
-
-\subsection{SF-BEAMS produced data and collection process}
-\label{sec:sfbeams}
-
-As described in section \ref{sec:sn-infrastructure}, sensor devices generates
-its observed data based on properties of measurements defined by its
-manufacturer. SF-BEAMS is a single-hop star sensor network containing
-different instances of sensors, including the "YSI 6600 ESD V2", as seen on
-figure \ref{fig:ysi-device}, which is a powerful water quality monitoring
-device that produces around 52 bytes (13x4 Bytes) on a single real-time data
-stream as shown on table \ref{tab:ysi-data-stream}.
-
-\begin{figure}
-  \centering
-  \includegraphics{../diagrams/ysi-device}
-  \caption{Picture of the case study sonde device: YSI 6600 ESD V2}
-  \label{fig:ysi-device}
-\end{figure}
-
-\begin{table}
-    \caption{YSI Data Stream}
-    \begin{center}
-        \begin{tabular}{lr}
-  12.20    192    179 5588.40   0.09   0.084   0.059  7.98   -79.6   99.5  8.83 
-  0.4     8.7
-        \end{tabular}
-    \end{center}
-    \label{tab:ysi-data-stream}
-\end{table}
-
-In order to have access to the observed data, the network staff downloads
-the data using one of the device's connection such as the RS-232 serial
-connector \cite{rs232}. Then, the data is transferred to the RTC labs for
-archival, where it is described, indexed and distributed to its main users.
-According to its staff members, in May 2009 the infrastructure of the SF-BEAMS
-included the following was organized as follows:
+Different approaches are used to represent the collected data from sensors. The
+use of simple data descriptors have been used in research groups which
+expertise is not database systems, while others use such technology and the way
+to represent data using more robust approaches. \cite{sn-data-model-survey}
+have shown different data models and the advantages and desadvantages. The
+following is a summary of these approaches used by the sensor network
+community.
 
 \begin{itemize}
-  \item 5 YSI sondes in operation at the RTC site Pier in Tiburon Island, San
-  Francisco Bay;
-  \item Each YSI sonde produces 52 Bytes for each its observations;
-  \item The sampling frequency rate could be configured in ranges of 1, 6 or
-  15 minutes, depending on the staff's speficiation. 
+  \item \textbf{Tabular Data Model}: comma-separeted values are very popular
+  due to its simplicity. Although data is not index, it is an easy approach for data
+ exchange among researchers.
+  \item \textbf{Relational Data Model}: it is used in the majority of the
+  projects. Most of them due to the easy access to database systems, with
+  refactored versions that support the implementation of sensor networks and
+  its requirements \cite{sn-db-tinydb};
+  \item \textbf{Structured Data Model}: simiar to the relational model,
+  structured data models are becoming popular due to the Internet. The use of
+  XML data \cite{xml} is not yet popular, but there are a few projects using it
+  \cite{sn-xml-usage01}\cite{sn-xml-usage01}, implementing its query by using a
+  middleware \cite{sn-xml-middleware} \cite{sn-xml-query-engines}.
+  \item \textbf{Key-Value Pair and Document-Oriented Data Model}: as
+  an emerging data model, the Key-Value Pair (KVP) \cite{db-kvp} data model is
+  becoming popular with the up-coming trend of cloud computing. Although there are no
+  records of the use of this model data model in sensor networks, this work
+  will investigate it.
 \end{itemize}
 
-Considering the infrastructure previously described, the size of the data
-produced by the YSI sonde sensors at the RTC pier can estimated by its
-measurement frequency as follows:
+Most importantly, the use of Data Provenance approaches to describe the
+collected data is shown to be important \cite{sn-provenance}, given the nature
+of the collected data. Therefore, this section describes these approaches for
+modeling data collected from sensor devices, taking into account the need of
+metadata for the data.
 
-\begin{table}
-    \label{tab:ysi-data-distribution}
-    \caption{Amount of data produced by the RTC's YSI sondes}
-        \begin{center}
-        \begin{tabular}{|c|c|c|c|c|c|c|}\hline 
-        \textbf{YSIs} & \textbf{Rate} & \textbf{Hourly} & \textbf{Daily} &
-        \textbf{Weeky} & \textbf{Monthly} & \textbf{Yearly}\\\hline 
-        1 & 1 min & 3.04 Kb & 73.12 Kb & 511.87 Kb & 1.99 Mb & 23.99 Mb\\\hline 
-        5 & 6 min & 15.23 Kb & 365.62 Kb & 2.5 Mb & 9.99 Mb & 119.97 Mb\\\hline 
-        1 & 15 min & 0.5 Kb & 12.18 Kb & 85.31 Kb & 341.25 Kb & 3.99 Mb\\\hline 
-        5 & 1 min & 2.54 Kb & 60.93 Kb & 426.56 Kb & 1.67 Mb & 19.99 Mb\\\hline
-        1 & 6 min & 0.2 Kb & 4.87 Kb & 34.12 Kb & 136.5 Kb & 1.6 Mb\\\hline 
-        5 & 15 min & 1.0 Kb & 24.37 Kb & 170.62 Kb & 682.5 Kb & 7.99 Mb\\\hline
-        \end{tabular}
-        \end{center}
-\end{table}
+\subsection{Tabular Data Model and Query}
 
-As it is shown, the data produced by this type of sensor runs in the order of
-at most hundreds of Megabytes a year, what can be considered a very low data
-storage requirement. Upon collecting the data from sensors, the RTC lab staff
-use automation scripts written in Matlab to process, index and distribute the
-raw data in different formats, including the OPeNDAP \ref{opendap}, a format
-widely used in the research institutions that promotes easier data exchange
-among them.
-
-Although the execution of the RTC SF-BEAM sensor network produces its
-execution, many operational challenges were faced by this type of
-infrastructure, when it comes to the data utilization of wireless sensors that
-are deployed in offshore conditions, \cite{netbeams2009} described the
-operational challenges faced by the RTC staff and proposed NetBEAMS, a
-component-based infrastructure that uses common-off-the-shelfs (COTS) embedded
-systems and software to automate the process of data collection. Next section
-describes how NetBEAMS is used to extract data from the SF-BEAMS sensor network.
-
-\section{NetBEAMS: a Real-World Case Study on Environmental Sensor Network}
-
-The Networked Bay Environmental Assessment and Monitoring System, or NetBEAMS, 
-is a work-in-progress joint project proposed by the department of Computer
-Science and the RTC at San Francisco State University. As described by
-\cite{netbeams2009}, the solution can be used to automated the data gathering
-process of the SF-BEAMS as described in section \ref{sec:sfbeams}. This
-section fully details how NetBEAMS is executed in order to collect data from
-SF-BEAMS, describing pertinent design and architectural details.
-
-One of the motivations for the inception of NetBEAMS was the automation
-of the data gathering and processing from the SF-BEAMS sondes without requiring
-human intervention. Most importantly, the decrease with the costs of constant 
-offshore site visits by boat, where biologists need manually collect the data
-from the sensor devices in order to take them for further processing at the
-lab. For this reason, the NetBEAMS research group proposed the use of the Data
-Sensor Platform (DSP) \cite{netbeams2009}, as the second attempt to automate
-the operation of SF-BEAMS.
-
-This following sections details the DSP, its architecture and how data is
-transported from the network nodes to the main network sink, the RTC labs.
-
-\subsection{The Data Sensor Platform (DSP)}
-
-The Data Sensor Platform, in short DSP Platform, is based on a micro-kernel
-architecture developed on top of the Java modular framework called OSGi
-\ref{osgi}. The platform represents the foundation of the execution of
-NetBEAMS, since it is executed from an embedded hardware called Gumstix
-\ref{gumstix} using a Linux operating system and on top of the Java Virtual
-Machine (JVM). In order to transport data from sensors hosts to the network
-sink, the use of a cellular connection is used to transport the collected data
-from the senso devices using the HTTP Protocol, as described by
-\cite{netbeams09}. Figure \ref{fig:sf-netbeams-node} shows the
-architecture of the inhanced node designed by the NetBEAMS research group.
-
-\begin{figure}
-  \centering
-  \includegraphics{../diagrams/DSP-Gateway-Node}
-  \caption{The NetBEAMS Node enhancement with the Gumstix and the GSM Modem}
-  \label{fig:sf-netbeams-node}
-\end{figure}
-
-The following list describes each of the devices in the NetBEAMS Node.
+Simple use of Comma-Separated Values is an example os sensor network data. Data
+is structured by the top header, and the values are in each line of the file.
+The implementation of the SF-BEAMS \cite{sfbeams2006} uses this approach. In
+summary, in a tabular data model
 
 \begin{itemize}
-  \item \textbf{YSI 6600EDS V2}: the YSI device is the regular sonde
-  responsible for observing the environment. It is connected to the Gumstix
-  embedded system, where the DSP is executed;
-  \item \textbf{Gumstix console-vx}: it's a COTS ARM-based hardware that
-  provides a computer-on-module environment for the development of small
-  embedded systems. This is the main environment of data extraction and remote
-  processing is accomplished. It runs a cross-compiled version of Linux Kernel
-  2.6;
-  \item \textbf{Huawei E220 USB Modem}: it is used to transfer the data from the
-  Gumstix to the RTC Labs Data Center.
+  \item data is represented by columns in a flat file separated by a character
+  that separates the values. One example of such data in sensor networks is the
+  OPenDAP \cite{opendap} format, used by the SF-BEAMS project
+  \cite{sfbeams2006}.
+  \item data is queried using a text editor file, without any indexing
+  associated with it. 
 \end{itemize}
 
-The DSP runs inside of the embedded system Gumstix. The following sections
-details its design and process for extracting data.
+\subsection{Relational Data Model and Query}
 
-\subsection{The DSP System Design}
+Many different projects have used the relational model \cite{relational-model}
+as the representation of data for sensor networks. In general, the following
+properties apply to sensor networks:
 
-This section focus on the development of a Software Platform for the NetBEAMS Gateway Embedded System, or on the system developed in the Gumstix device.
-As shown on Image 2.4.1.1, the NetBEAMS Gateway Node is an embedded system that contains a set of layer found in any computer.
-However, each of the resources described from the system were customized for the NetBEAMS system.
-The main components of such system can be summarized as follows:
+\begin{itemize}
+  \item data is structured in tables, and sensor properties are represented 
+ by the columns of these tables. Data instances are represented by tuples or
+ rows of these tables. Finally, table relationships are implemented based
+ on the relational algebra abstraction, which is the mechanism used to 
+describe the relationship among different data types.
+  \item data is queried using the Structured Query Languange (SQL), and uses
+  the definitions of the tables and columns to extract data.
+\end{itemize}
 
-    * Operating System: it uses a cross-compiled Gentoo Linux, which supports a wide variety of development tools, including the Java Virtual Machine (JVM), the underlying platform system;
-    * Java Virtual Machine: it uses the JamVM, a ~200 KB version of the Sun Microsystems Java Virtual Machine, version 2.0;
-    * OSGi Framework: it uses the Knopflerfish implementation of the OSGi 4.1 specification. More details in the following sections.
-    * DSP Platform and other Bundles: The plug-and-play DSP Components are based on the OSGi bundles capabilities, which can reuse services registered in the OSGi framework.
+Before the conception of such model, a database normalization process is
+necessary to describe the data in an appropriate way. An example of such use
+of the relational model is the TinyDB database systems \cite{sn-db-tinydb},
+which uses the relational model to represent the types of sensors. Due to the
+popularity of relational models, this is a popular choice since of the vast
+availability of research on relational database models.
+
+Although the implementation of a relational database seems to cover the needs
+of data representation, \cite{sn-data-model-survey} described the problems
+related to this approach. One such problem is that the relational model is not
+suited for modeling time series data and metadata related to identify, time and
+location. For the problems related to it, came the implementation of different
+approaches that adds semantics to data types and different query clauses to the
+Structured Query Language (SQL) to solve join \cite{sn-db-newop}.
+
+\subsection{Structured Data Model and Query}
+
+The use of XML is used to represent data in sensor networks has become popular
+due the popularity of this data format. In general, the use of XML for
+persistence can be summarized as follows:
+
+\begin{itemize}
+  \item data is structured in the collected data properties can be represented through the
+XML \cite{xml} instances by using tags, having its model validated by the use
+of XML Schema \cite{xml-schema}.
+  \item in order to extract data from XML-based databases, the use of XML XPath
+  is used to query XML documents.
+\end{itemize}
+
+Some examples of the application of XML in sensor networks in 
+\cite{sn-xml-usage02} some have investigated the effecient use of XML in
+sensor networks to represent the collected data \cite{sn-xml-usage01}. An
+example of querying xml data in sensor networks described by
+\cite{sn-xml-query-engines}, while others use XML as a means of data exchange
+by implementing of middlewares to query and extract query data
+\cite{sn-xml-middleware}. NetBEAMS is one of such sensor network that contains
+an XML-based communication middleware for data exchange among the sensor nodes
+\cite{netbeams2009}.
+
+\subsection{Key-Value Pair and Document-Oriented Data Models}
+
+The advance of distributed systems and the Internet have enabled the
+development of powerful database systems that can be organized in the context
+of database systems and how to represent models. An important up-coming
+approach is the use of database systems that does not use structured language
+for quering its data. One such example is called Key-Value Pair (KVP) data model
+\cite{db-kvp}. Such model has the following properties:
+
+\begin{itemize}
+  \item data is structured in collections of key-value pairs, as it is done in
+  hash data structure. The definition of the key is a given property with its
+  associating value.
+  \item the query process is using mechanics similar to programming or
+  scripting language, that is, the use of APIs in a given programming language.
+\end{itemize}
+
+Different variants of such data model is the document-oriented data model,
+where data is modelled using structured documents. However, these documents
+have a dynamic structure that can be freely described without the use of an
+schema validation mechanism as used in XML Schema. One example of such
+approach is the use of the JSON data format \cite{json} in the implementation
+of database systems which are becoming popular with the new trend called Cloud
+Computing \cite{cloud-comp-architectures}. Databases implementing this strategy
+is the mongoDB \cite{mongodb} and CouchDB \cite{mongodb}.
+
+\subsection{Data Provenance for Sensor Networks}
+
+Most of sensor networks implementations use the collected data from the sensor
+devices after they observe the environment. As shown by
+\cite{sn-data-model-survey}, different data models contains its pros and
+cons to represent data. However, one important aspect in sensor networks is the
+description of it collected data, as \cite{sn-provenance} discussed the
+properties and nature of sensor networks generated data. First, the device's
+sampling are usually in raw format or without obvious names. In order to
+provide search over the collected sensor data, the data needs first to be
+described and indexed, and this process can be related to Data Provenance
+\cite{db-provenance}.
+
+An example of Data Provenance application is the use of aggregation of data
+over time to estimate the changing effects of the weather in a given region.
+For this reason, the collected data needs descriptions or annotations that can
+give enough information to track changes, that is, the history of how and when
+a given data has changed. The following should be the foundation of a
+provenance-aware collected data according to \cite{sn-provenance}:
+
+\begin{itemize}
+  \item each low-level observation as it describes the actual data;
+  \item the location where the collected data was generated;
+  \item the time when the data was collected.
+\end{itemize}
+
+While Data Provenance provides necessary mechanisms to describe sensor networks
+data in general, others see the importance of the time space of sensor
+networks. \cite{sn-time-series} describes strategies to used to cluster
+distributed sensor data as time series of data.

### 2009-11-14 docs/marcello-dissertation/thesis/chapter3-taxonomic-classification-of-persistence-storage-for-sensor-networks.tex
Updating the thesis and added chapters 2-5. Still revisions are needed in section 4 and 5, adjusting images and tables. Images are in the eps format.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@642 559edde2-a4d4-44e0-96fc-4035e40d7396

% main.tex, to be used with thesis.tex
% This contains the main work of your thesis.

%\bibliography{thesis}  % uses the references stored in Chapter1Radar.bib

\chapter{Taxonomic Classification of Persistence Storage Systems for Sensor
Networks}
\label{chap:taxonomies}

Different mechanisms are used for persisting data collected from sensor
networks, as described by the section regarding the state of the art in sensor
networks and different approaches for persistence. This section analyzes these
approaches and presents the different approaches through taxonomy
\footnote{Taxonomy is the practice and science of classification, by using
taxonomic units known as taxa (singular taxon)}, which will be used throughout
this work to evaluate different existing database systems implementations,
taking into account different requirements and specifications of different
sensor networks.

\section{Nature and Location of Sensor Data Taxonomy}

No matter the type of resource management used, the first taxonomy presented is
related to the purpose of the use of the collected sensor data. That is, when
and how the collected data is going to be used as described in section
\ref{sec:sn-data-purpose}. In addition to it, the mechanism for data storage
described in section \ref{sec:sn-infrastructure} can be related to each of
them. The related taxon is depicted by figure \ref{fig:taxonomy-data-purpose}.

\begin{figure}[h]
  \centering
  \includegraphics{../diagrams/taxonomy-data-purpose}
  \caption{The Nature and Location of Sensor Data Taxonomy}
  \label{fig:taxonomy-data-purpose}
\end{figure}

\begin{itemize}
  \item \textbf{Real-time Data Stream}: data is used as real-time data stream
  during data query \cite{sn-intro01}.
    \begin{itemize}[label=\textbullet]
        \item \textbf{Local Storage}: The data can be temporarily persisted
        in-memory, or a secondary memory device in a local storage device.
    \end{itemize}
  \item \textbf{Archival Data}: data is used for historical purposes such as
  data analysis or data fusion \cite{sn-intro01, sn-intro02}.  
    \begin{itemize}[label=\textbullet]
      \item \textbf{External Storage}: The data is usually located in an
       external storage device with a data management system such as a database
       system.
      \item \textbf{Data-Centric Storage}: the collected data can arranged in a
      categorized fashion as described by \cite{sn-storage03}. This taxon of
      system allows the data to be partitioned by a specific property. Nowadays,
      the strategy is often related to database partition \cite{db-partition}
      and database sharding approaches \cite{db-shard01} \cite{db-shard02}.
    \end{itemize}
\end{itemize}

\section{Data Model Taxonomy}

One of the major challenges in the development of a persistence storage for
sensor network is related to the persistence model chosen. The addition of new
sensor devices may trigger refactoring of the existing model or not, depending
on how the data model accommodates the data. Different data models are reported
to be used in the literature, as described in section \ref{sec:data-models}.
For this reason, taxonomy related to the data model depicts two different
classes of data models, as shown in figure \ref{fig:taxonomy-data-model}.

\begin{figure}[h]
  \centering
  \includegraphics{../diagrams/taxonomy-data-model}
  \caption{Data Model Taxonomy}
  \label{fig:taxonomy-data-model}
\end{figure}

\begin{itemize}
  \item \textbf{Schema-Dependent Models}: this type of data model require
  the definition of a master schema that describes the data through a regorous
  data modeling process prior to the use of the data. In addition, an addition
 to a new sensor device, and consequently a new data entity, requires a
 re-evaluation of the existing data schema. For instance, relational model is
 governed by the relational algebra model, while the structured models such as
 XML needs a definition of an XML Schema.
  \item \textbf{Schema-less Models}: contrary to the previous taxon, 
  schema-less data models does not require the definition of a schema or master
  table definition and relationships. Instances of such data model is the KVP
  and the document-oriented models, whose usage in wireless sensor networks
  have not been reported.
\end{itemize}

\section{Query Processing Mechanism Taxonomy}

The query processing mechanism in sensor networks is directly related to how
and where the collected data is used and stored, as shown in the previous
section. As a consequence, the following taxonomy taxon is related to the
different types of query processing mechanisms described in section
\ref{sec:query-process}, and can be seen in figure
\ref{fig:taxonomy-query-mechanism}.

\begin{figure}[h]
  \centering
  \includegraphics{../diagrams/taxonomy-query-mechanism}
  \caption{Query Processing Mechanism Taxonomy}
  \label{fig:taxonomy-query-mechanism}
\end{figure}

\begin{itemize}
  \item \textbf{In-Network Query Processing}: the sensor node receives a
  request to the properties of the sensor, and returns the current values
  related to them \cite{sn-intro01}.
  \item \textbf{Centralized Query Processing}: as it is done in regular single
  server relational databases, the data is collected to a centralized data
  sink in order to be used.
\end{itemize}

\section{Database System Organization Taxonomy}

Different approaches of data storage can also be related to the performance in
the database system used, and how its architecture provides the data storage
solution. For example, one important problem related to the centralized
network sink in sensor networks is the presence of an unavoidable creation of a
point of traffic concentration at the data collector or network nodes as
described by \cite{sn-storage02}. In this way, in order to mitigate the
problems associated with the so-called ``Funneling Effect", figure
\ref{fig:taxonomy-database-architecture} summarizes the different capabilities
of different types of database systems.

\begin{figure}[h]
  \centering
  \includegraphics{../diagrams/taxonomy-database-architecture}
  \caption{Database Architecture Taxonomy}
  \label{fig:taxonomy-database-architecture}
\end{figure}

\begin{itemize}
  \item \textbf{Centralized System}: this is a database system that runs in
  one single centralized host, being the main point of data traffic as read
  and writes \cite{sn-intro01}.
  \item \textbf{Distributed System}: a database system that can be composed by
  more than one node. Usually, the use of the paradigm of Master-Slave
  mechanisms can be used by implementing ``Replication". In addition to it,
  ``Database Sharding" is used to portion the data related to a given shard
  key.
\end{itemize}

### 2009-11-14 docs/marcello-dissertation/thesis/chapter4-netbeams-in-depth-requirements-analysis-and-specification.tex
Updating the thesis and added chapters 2-5. Still revisions are needed in section 4 and 5, adjusting images and tables. Images are in the eps format.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@642 559edde2-a4d4-44e0-96fc-4035e40d7396

% main.tex, to be used with thesis.tex
% This contains the main work of your thesis.

%\bibliography{thesis}  % uses the references stored in Chapter1Radar.bib

\chapter{Netbeams in Depth: Requirements Analysis and Specification}

This chapter presents NetBEAMS and its implementation details, showing its
main application of the the environmental sensor network deployed at the
Romberg Tiburon Center (RTC) as it is the main case study.

\section{SF-BEAMS: a Marine Sensor Network for Water Quality Monitoring}

The San Francisco Bay Environmental Assessment and Monitoring Station,
or SF-BEAMS, is a marine sensor network operated by the Romberg Tiburon Center
(RTC). This section details the SF-BEAMS sensor network located at
the Tiburon Island, California, and then presents the details of NetBEAMS, as
it is part of the case study of this work.

\subsection{The SF-BEAMS Infrastructure}

The RTC's main focuses are in the study of complex marine and estuarine
environments using the SF-BEAMS sensors deployed offshore of its pier located
on the banks of the San Francisco Bay at the Tiburon Island, California. In
this way, the project is responsible for providing data for water quality
monitoring, as well as weather and surface conditions. Figure 
\ref{fig:sf-beams} shows a picture taken from the SF-BEAMS web-cam placed
in-site.

\begin{figure}[!b]
  \centering
    \includegraphics[scale=0.7]{../diagrams/cam_image-oct15}
  \caption{Picture of the SF-BEAMS Network, at the RTC pier. Tiburon, CA.}
  \label{fig:sf-beams}
\end{figure}

In general, the SF-BEAMS network infrastructure contains heterogeneous wired and
wireless devices standing on pilons docked in the pier, where each of them is
responsible for observing different conditions of the area, as well as having
its own mechanisms for internal storage for the collected data. In this way,
data can be directly transferred to the labs via ethernet cables or collected
manually download by a staff's laptop computer.

\subsection{SF-BEAMS produced data and collection process}
\label{sec:sfbeams}

As described in section \ref{sec:sn-infrastructure}, sensor devices generates
its observed data based on properties of measurements defined by its
manufacturer. SF-BEAMS is a single-hop star sensor network containing
different instances of sensors, including the "YSI 6600 ESD V2", as seen on
figure \ref{fig:ysi-device}, which is a powerful water quality monitoring
device that produces around 52 bytes (13x4 Bytes) on a single real-time data
stream as shown on table \ref{tab:ysi-data-stream}.

\begin{figure}[!b]
  \centering
  \includegraphics[scale=0.7]{../diagrams/ysi-device}
  \caption{Picture of the case study sonde device: YSI 6600 ESD V2}
  \label{fig:ysi-device}
\end{figure}

\begin{table}
    \caption{YSI Data Stream}
    \begin{center}
        \begin{tabular}{lr}
  12.20    192    179 5588.40   0.09   0.084   0.059  7.98   -79.6   99.5  8.83 
  0.4     8.7
        \end{tabular}
    \end{center}
    \label{tab:ysi-data-stream}
\end{table}

In order to have access to the observed data, the network staff downloads
the data using one of the device's connection such as the RS-232 serial
connector \cite{rs232}. Then, the data is transferred to the RTC labs for
archival, where it is described, indexed and distributed to its main users.
According to its staff members, in May 2009 the infrastructure of the SF-BEAMS
included the following was organized as follows:

\begin{itemize}
  \item 5 YSI sondes in operation at the RTC site Pier in Tiburon Island, San
  Francisco Bay;
  \item Each YSI sonde produces 52 Bytes for each its observations;
  \item The sampling frequency rate could be configured in ranges of 1, 6 or
  15 minutes, depending on the staff's speficiation. 
\end{itemize}

Considering the infrastructure previously described, the size of the data
produced by the YSI sonde sensors at the RTC pier can estimated by its
measurement frequency as follows:

\begin{table}
    \label{tab:ysi-data-distribution}
    \caption{Amount of data produced by the RTC's YSI sondes}
        \begin{center}
        \begin{tabular}{|c|c|c|c|c|c|c|}\hline 
        \textbf{YSIs} & \textbf{Rate} & \textbf{Hourly} & \textbf{Daily} &
        \textbf{Weeky} & \textbf{Monthly} & \textbf{Yearly}\\\hline 
        1 & 1 min & 3.04 Kb & 73.12 Kb & 511.87 Kb & 1.99 Mb & 23.99 Mb\\\hline 
        5 & 6 min & 15.23 Kb & 365.62 Kb & 2.5 Mb & 9.99 Mb & 119.97 Mb\\\hline 
        1 & 15 min & 0.5 Kb & 12.18 Kb & 85.31 Kb & 341.25 Kb & 3.99 Mb\\\hline 
        5 & 1 min & 2.54 Kb & 60.93 Kb & 426.56 Kb & 1.67 Mb & 19.99 Mb\\\hline
        1 & 6 min & 0.2 Kb & 4.87 Kb & 34.12 Kb & 136.5 Kb & 1.6 Mb\\\hline 
        5 & 15 min & 1.0 Kb & 24.37 Kb & 170.62 Kb & 682.5 Kb & 7.99 Mb\\\hline
        \end{tabular}
        \end{center}
\end{table}

As it is shown, the data produced by this type of sensor runs in the order of
at most hundreds of Megabytes a year, what can be considered a very low data
storage requirement as compared to other systems such as web applications. Upon
collecting the data from sensors, the RTC lab staff use automation scripts
written in Matlab to process, index and distribute the raw data in different
formats, including the OPeNDAP \cite{opendap}, a format widely used in the
research institutions that promotes easier data exchange among them. The
access to the archived data is done by using the OPeNDAP server is located at
http://sfbeams.sfsu.edu:8080/opendap.

Although the execution of the RTC SF-BEAM sensor network produces its
execution, many operational challenges were faced by this type of
infrastructure, when it comes to the data utilization of wireless sensors that
are deployed in offshore conditions, \cite{netbeams2009} described the
operational challenges faced by the RTC staff and proposed NetBEAMS, a
component-based infrastructure that uses COTS \footnote{Common-Off-The-Shelf
designates a product that is produced and sold in bulk} embedded systems and
software to automate the process of data collection. Next section describes
how NetBEAMS is used to extract data from the SF-BEAMS sensor network.

\section{RTC Requirements for Data Use}
\label{sec:problem-requirements}

Main users are biologists, students and researchers with some knowledge in
programming and scripting languages.

\subsection{Functional Requirements}

\begin{itemize}
  \item Reuse the NetBEAMS infrastructure and develop a component responsible
  for data persistence in a database system;
\end{itemize}

\subsection{Non-Functional Requirements}

\begin{itemize}
  \item Main users are biologists, no expertise in Database Systems;
  \item Data Representation in human-readable format;
  \item Scalable system with low degree of maintanance;
  \item Data searchable in near-real-time
\end{itemize}

\section{Understanding NetBEAMS Architecture}
\label{netbeams-architecture}

The Networked Bay Environmental Assessment and Monitoring System, or NetBEAMS, 
is a work-in-progress joint project proposed by the department of Computer
Science and the RTC at San Francisco State University. As described by
\cite{netbeams2009}, the solution can be used to automated the data gathering
process of the SF-BEAMS as described in section \ref{sec:sfbeams}. This
section fully details how NetBEAMS is executed in order to collect data from
SF-BEAMS, describing pertinent design and architectural details.

One of the motivations for the inception of NetBEAMS was the automation
of the data gathering and processing from the SF-BEAMS sondes without requiring
human intervention. Most importantly, the decrease with the costs of constant 
offshore site visits by boat, where biologists need manually collect the data
from the sensor devices in order to take them for further processing at the
lab. For this reason, the NetBEAMS research group proposed the use of the Data
Sensor Platform (DSP) \cite{netbeams2009}, as the second attempt to automate
the operation of SF-BEAMS.

This following sections details the DSP, its architecture and how data is
transported from the network nodes to the main network sink, the RTC labs.

\subsection{The Data Sensor Platform (DSP)}

The Data Sensor Platform, in short DSP Platform, is based on a micro-kernel
architecture developed on top of the Java modular framework called OSGi
\cite{osgi}. The platform represents the foundation of the execution of
NetBEAMS, since it is executed from an embedded hardware called Gumstix
\cite{gumstix} using a Linux operating system and on top of the Java Virtual
Machine (JVM). In order to transport data from sensors hosts to the network
sink, the use of a cellular connection is used to transport the collected data
from the senso devices using the HTTP Protocol, as described by
\citet{netbeams2009}. Figure \ref{fig:sf-netbeams-node} shows the
architecture of the inhanced node designed by the NetBEAMS research group.

\begin{figure}[h]
  \centering
  \includegraphics[scale=0.5]{../diagrams/DSP-Gateway-Node}
  \caption{The NetBEAMS Node enhancement with the Gumstix and the GSM Modem}
  \label{fig:sf-netbeams-node}
\end{figure}

The following list describes each of the devices in the NetBEAMS Node.

\begin{itemize}
  \item \textbf{YSI 6600EDS V2}: the YSI device is the regular sonde
  responsible for observing the environment. It is connected to the Gumstix
  embedded system, where the DSP is executed;
  \item \textbf{Gumstix console-vx}: it's a COTS ARM-based hardware that
  provides a computer-on-module environment for the development of small
  embedded systems. This is the main environment of data extraction and remote
  processing is accomplished. It runs a cross-compiled version of Linux Kernel
  2.6;
  \item \textbf{Huawei E220 USB Modem}: it is used to transfer the data from the
  Gumstix to the RTC Labs Data Center.
\end{itemize}

The DSP runs inside of the embedded system Gumstix. The following sections
details its design and process for extracting data.

\subsection{The DSP System Design}

This section focus on the development of a Software Platform for the NetBEAMS
Gateway Embedded System, or on the system developed in the Gumstix device. As
shown in figure \ref{fig:netbeams-software-stack}, the NetBEAMS Gateway Node is
an embedded system that contains a set of layer found in any computer.
However, each of the resources described from the system were customized for
the NetBEAMS system. The main components of such system can be summarized as
follows:

\begin{itemize}
  \item \textbf{Operating System}: it uses a cross-compiled Gentoo Linux, which
  supports a wide variety of development tools, including the Java Virtual Machine
  (JVM), the underlying platform system;
  \item \textbf{Java Virtual Machine}: it uses the JamVM, a ~200 KB version of
  the Sun Microsystems Java Virtual Machine, version 2.0;
  \item \textbf{OSGi Framework}: it uses the Knopflerfish implementation of the
  OSGi 4.1 specification. More details in the following sections.
  \item \textbf{DSP Platform and other Bundles}: The plug-and-play DSP
  Components are based on the OSGi bundles capabilities, which can reuse services registered in the OSGi framework.
\end{itemize}

\begin{figure}[h]
  \centering
  \includegraphics[scale=0.6]{../diagrams/NetBEAMS-Client-Node-Components}
  \caption{The NetBEAMS's Gumstix Software Stack}
  \label{fig:netbeams-software-stack}
\end{figure}

As depicted in figure \ref{fig:netbeams-software-stack}, the DSP Sonde Data
Producer component opens a serial connection with the device through the Java
Native Interface (JNI) driver and produces the data flow 1. The data stream
format at this stage is as described on table \ref{tab:ysi-data-stream}.
Later, by parsing and transforming the data to DSP internal representation, it
sends the data to the DSP Platform as depicted on flow 2. Finally, the DSP
Platform decides to send the data to the DSP Wire Transport Client, as shown
on flow 3, by following rules defined by the system administrator. At this
point, the data is ready to be transmitted to the server, which is covered on
the following sections.

\subsection{The DSP Platform Architecture}

The architecture of the DSP, aligned with the OSGi capabilities, reuses the
Service-Oriented Architecture that can easily enable and disable plug-and-play
components. By following a publish-subscriber design-pattern \cite{gof}, the
service producer registers into the service broker while the consumer looks
for a given registered service in the Service Broker. For this reason, this
approach allows decoupled interoperability among modules as it is shown in
figure \ref{fig:publish-subscriber-pattern}.

\begin{figure}[h]
  \centering
  \includegraphics[scale=0.6]{../diagrams/publish-subscriber-pattern}
  \caption{The publish-subscriber design-pattern for services}
  \label{fig:publish-subscriber-pattern}
\end{figure}

The DSP infrastructure is based on a loosely-coupled architecture hosted in a
given host computer with an OSGi container, where the entire system is
composed by a set of plug-n-play OSGi modules. Each DSP component, as an
instance of an OSGi Module, can perform specialized tasks according to its
specification. In general, each DSP component can be categorized as follows:

\begin{itemize}
  \item \textbf{Data Producer (DP)}, whose main function is to produce data for
  the system;
  \item \textbf{Data Consumer (DC)}, whose main function is to consume the
  received data.
\end{itemize}

\begin{figure}[!b]
  \centering
  \includegraphics[scale=0.7]{../diagrams/DSP-Producers-Consumers-Components-Interactions}
  \caption{The basic architecture of a Data Sensor Platform (DSP)}
  \label{fig:DSP-Producers-Consumers-Components-Interactions}
\end{figure}

The Data Producers and Data Consumers exchange data through the DSP Message
Broker using a message unit called DSP Message. As it is depicted in figure
\ref{fig:DSP-Producers-Consumers-Components-Interactions}, a Data Producer
creates a set of one or more DSP Messages wrapped up in a DSP
MessagesContainer in order to be sent to the DSP Message Broker. When the
broker receives the set of DSP Messages, it requests a list of Data Consumers
to each of the DSP Messages based on rules managed by the DSP Matcher.
Whenever the broker acquires the list of Data Consumers, it delivers a copy of
the DSP Message to each of the Data Consumers in the receiving list. Upon the
DSP Message arrival, each Data Consumer can extract the actual data from the
DSPMessage. Finally, although the components can be distinguished between Data
Producers and Consumers, any of them are entitled to send and receive
administrative DSP Messages, which is covered in the following sections.

When the DSP Platform is installed into the OSGi framework, it follows the
OSGi specifications without any changes to remain it in the installed state.
In this way, it is important to note that the DSP Platform and its
plug-and-play components are directly mapped with the OSGi Framework as
follows:

\begin{itemize}
  \item \textbf{OSGi Bundle} = DSP Component Artifact in Jar format
  \begin{itemize}
    \item SondeDSPProducer.jar is the component developed for the YSI sonde
  \end{itemize}  
  \item \textbf{OSGi Activator} = DSP Component Activator
  \begin{itemize}
    \item Class specified in the MANIFEST.MF artifact (next sections)
  \end{itemize} 
  \item \textbf{OSGi Service Reference} = DSP Component Class
  \begin{itemize}
    \item  Class specified by the component author.
  \end{itemize}
\end{itemize}

The DSP Platform, along with the DSP Framework, are the only DSP Components
needed to be installed into the OSGi container. For this reason,the DSP
Platform is responsible to install and activate the additional DSP components
that makes part of the platform during the start up process, which leverages
its bootstrap process. Nevertheless, in order to automate the process of
automatic install and removal of the plug-and-play components, the DSP Platform
provides a deployment configuration file that describes which DSP components
must be installed with a given priority number. The name of this configuration
artifact is called config.xml (see listing \ref{file:dsp-config.xml}), and is
depicted by the XML Schema \cite{xml-schema} in figure
\ref{fig:dsp-config-schema}.

\begin{figure}[h]
  \centering
  \includegraphics[scale=0.8]{../diagrams/config-schema}
  \caption{XML Schema for the main DSP deployment configuration artifact}
  \label{fig:dsp-config-schema}
\end{figure}

In order to show an example, consider the DSP Platform on the embedded system
Gumstix in figure \ref{fig:sf-netbeams-node}. It is started by providing a
deployment configuration artifact config.xml included in the appendix. It lists
the components that the DSP Platform needs to install and load for its
operation. As it is shown, the DSPWireTransportClient and the DSPSondeProducer
are the identification of the components, followed by the name of the physical
file and an assigned loading priority. While the DSPWireTransportClient is
responsible for the serialization and deserialization of the DSP Messages
exchanged by the DSPSondeProducer and any remote DSP component. In this case,
the DSPSondeProducer is the DSP component responsible to collect measurement
data from the YSI Sonde and temporarily store them before being transmitted.

Therefore, in order to orchestrate the DSP Components together, the DSP
Platform component is responsible for installing and uninstalling during its
state "Starting" and "Stopping", respectively, as it is shown in figure
\ref{fig:DSPPlatform-Install-Usage-State-Diagram}. When the system
administrator starts the DSP Platform, it performs the regular OSGi procedures
to start the bundle. Up until that state, it will get the list of DSP
Components to be installed as described in the config.xml artifact, trying to
install each of them in an optimistic way. After that, it will continue on the
state "Active" until the system administrator decides to stop it.

\begin{figure}[!b]
  \centering
  \includegraphics[scale=0.6]{../diagrams/DSPPlatform-Install-Usage-State-Diagram}
  \caption{The DSP Platform State Diagram}
  \label{fig:DSPPlatform-Install-Usage-State-Diagram}
\end{figure}

When the DSP Platform is on the Active state, it is ready to manage the DSP
Messages exchanges as described through figure
\ref{fig:DSP-Producers-Consumers-Components-Interactions}. The following
sections describe the participating classes in starting the DSP Platform and
its execution.

In a nutshell, the DSP Platform is an extension of the OSGi Platform, as it can
be seen in the UML Class Diagram \cite{uml} in figure
\ref{fig:DSP-Platform-Class-Diagram-Simple}. The class Activator contains a
unique instance of the Platform class, which contains instances of the classes
ComponentManager, Matcher, the MessageBroker and DSPContextImpl. The
DSPBundleController is the main instance that manages the plug-and-play DSP
components, and therefore, is tightly-coupled with the OSGi Framework classes.
Additionally, this class is part of the publish-subscriber design-pattern, as
it it implements the interfaces Bundle Listener and Service Listener, whose
responsibility is to track listens to the changes of the state of
ServicesReferences on the DSP Platform.

\begin{figure}[!t]
  \centering
  \includegraphics[scale=0.7]{../diagrams/DSP-Platform-Class-Diagram-Simple}
  \caption{The DSP Platform UML Class Diagram}
  \label{fig:DSP-Platform-Class-Diagram-Simple}
\end{figure}

\section{DSP Message and DSP Messages Container}

As shown in the previous section, a DSP Message is the main unit of
communication among DSP Components. In a nutshell, a DSP Message can be seen
as an abstraction of an envelop containing the sections: a header and a body.
The former is used for identification and routing purposes, while the latter is
used to carry the payload of the message, that is, the collected data from
a sensor. Figure \ref{fig:DSP-Message-Representation} depicts the abstraction
of a DSP Message and its main components.

\begin{figure}[!b]
  \centering
  \includegraphics[scale=0.8]{../diagrams/DSP-Message-Representation}
  \caption{An Abstraction of a DSP Message and its components}
  \label{fig:DSP-Message-Representation}
\end{figure}

\begin{itemize}
  \item \textbf{Header} is composed by 2 blocks that identify the data
  producer and consumer of the message, as well as the time of creation, UUID,
  etc. It is an important piece of information for the data delivery and
  routing purposes.
  \item \textbf{Body} contains the payload of the message. This section carries
  the collected data from a sensor.
\end{itemize}

In general, the DSP Platform offers a veriety of types of DSP Messages for
different purposes. For example, any measurement data must be
wrapped up in a Measurement Message, while a Query Message is used to exchange
messages among the components for the purpose of management. In this way, the
main DSP Messages can be summarized as follows:

\begin{itemize}
  \item \textbf{Measurement Message} is used to transport any sensor collected
  data;
  \item \textbf{Query Message} is used to query a DSP component about its
  configuration properties;
  \item \textbf{Update Message} used to update a DSP component's configuration
  properties;
  \item \textbf{Acknowledgement Message} is used for the transport communication
  protocol, by acknowledging the reception of a DSP Message. More details in the
  Remote Data delivery section.
\end{itemize}

These DSP Classes and other can be seen in figure the  UML Class
Diagram \cite{uml} of figure \ref{fig:DSP-Messages-Classes}. Each DSP Message
contains an instance of a Header and Body. Whenever a DSP component is ready
to transmit messages, it wraps up the set of DSP Messages into the instance of
the Messages Container, which contains information about the collection of
messages being transmitted with its own identification. In this fashion, the
Messages Container is the main communication unit between 2 different DSP
Components.

\begin{figure}[!t]
  \centering
  \includegraphics[scale=0.6]{../diagrams/DSP-Messages-Classes}
  \caption{The DSP Messages UML Class Diagram}
  \label{fig:DSP-Messages-Classes}
\end{figure}

Another use of the DSP Messages is during the DSP Platform activation process,
where bootstrap messages are used to configure each DSP component. In order to
send configuration parameters, the DSP Platform uses Update Messages using a
unique type of Message Content called DSP Properties, which is described in the
next section.

\section{Collected Sensor Data as the DSP Message Content}

As mentioned in the previsou section, a DSP Message can carry any data on its
body, also called the payload. In such a way, any data representation must
just extend from the Message Content abstract class as it is shown
\ref{fig:Sonde-MessageContainer-Class-Diagram}, as defines the data
representation of the YSI Sonde data.

\begin{figure}[!h]
  \centering
  \includegraphics[scale=0.6]{../diagrams/Sonde-MessageContainer-Class-Diagram}
  \caption{The YSI Payload - Message Content UML Class Diagram}
  \label{fig:Sonde-MessageContainer-Class-Diagram}
\end{figure}

The properties measured from the YSI Sonde can be captured by the class
SondeDataType as a direct render of the measurements data depicted in table
\ref{tab:ysi-data-stream}. In addition to the regular data types, the class
also provides additional properties such as the attribute "samplingTimeStamp"
that carries the time when the data was collected from the YSI Sonde. However,
in order to capture a set of measurements at once, the SondeDataContainer
class used to carry a set of SondeDataType classes, as it is composed by one
or more instances of SondeDataType. In order to illustrate other examples,
consider the UML Class Diagram \cite{uml} of image
\ref{fig:Mouse-Actions--MessageContainer-Class-Diagram} as a payload of
observations made from a mouse over a screen, which captures the name of the
event, the button name and the x/y coordinates.

\begin{figure}[!h]
  \centering
  \includegraphics[scale=0.5]{../diagrams/Mouse-Actions--MessageContainer-Class-Diagram}
  \caption{The Mouse Actions Payload - Message Content UML Class Diagram}
  \label{fig:Mouse-Actions--MessageContainer-Class-Diagram}
\end{figure}

As described in the previous section, the DSP Platform uses DSP Update
Messages to configure internal components' initial configuration values. In
order to do so, the DSP Platform uses an instance of the DSP Properties, which
is composed of a list of initial DSP Property and their relating Values
instance depicted on the UML class diagram \cite{uml} of figure
\ref{fig:DSP-Property-Class-Diagram}.

\begin{itemize}
  \item \textbf{Property Instance}: initial-rate
  \item \textbf{Value Instance}: 40
\end{itemize}

\begin{figure}[!b]
  \centering
  \includegraphics[scale=0.6]{../diagrams/DSP-Property-Class-Diagram}
  \caption{The DSP Property Payload - Message Content UML Class Diagram}
  \label{fig:DSP-Property-Class-Diagram}
\end{figure}

\subsection{DSP Platform Initialization Process}

All in all, the DSP Platform requires the knowledge of the aforementioned DSP
classes in order to initialize. The UML Sequence Diagram \cite{uml} in figure
\ref{fig:From-OSGi-DSP-Platform-Start-to-All-Components-Sequence-Diagram}
depicts the exact moments when the additional the plug-and-play DSP components
are installed and bootstrapped by the DSP Platform, as the UML State Diagram
\cite{uml} in figure \ref{fig:DSPPlatform-Install-Usage-State-Diagram}
described.

\begin{itemize}
  \item \textbf{Start DSP Activator Gate}: after installing the Activator
  indicated by the DSP Bundle's MANIFEST.MF artifact, the OSGi Framework makes
  a call to the method Activator.start();
  \item \textbf{Bootstrap DSP Component Gate}: after the component has been
  initialized (by the call the method init() from the class ComponentManager),
  the DSP Component is ready to be bootstrapped, and the optional bootstrap
  message is delivered.
\end{itemize}

\begin{figure}[!b]
  \centering
  \includegraphics[scale=0.5]{../diagrams/From-OSGi-DSP-Platform-Start-to-All-Components-Sequence-Diagram}
  \caption{The DSP Platform Initialization UML Sequence Diagram}
  \label{fig:From-OSGi-DSP-Platform-Start-to-All-Components-Sequence-Diagram}
\end{figure}

\subsection{Data Delivery Mechanism}

In general, when a DSP Component finishes preparing the DSP Messages Container,
it contacts an entity responsible for sending and proxying message to other
components. The DSP Broker acts like a mailman, receiving messages and
forwarding copies of it to other entities in the system, abstraction in which
is shown in figure \ref{fig:DSP-Message-Broker-Abstraction}. In this way, the
DSP Broker receives a DSP Message and makes necessary deliveries to DSP
components that it has knowledge of.

\begin{figure}[!t]
  \centering
  \includegraphics[scale=0.5]{../diagrams/DSP-Message-Broker-Abstraction}
  \caption{A DSP Message Broker Abstraction}
  \label{fig:DSP-Message-Broker-Abstraction}
\end{figure}

The use of messaging exchange paradigm is used in NetBEAMS because it decouples
the system by using messages take into account their structure to filter the
messages. In this way, the DSP Broker uses a filter component called DSP
Matcher, which provides a list of DSP Components that needs to receive a copy
of the DSP Message. It turns out that the DSP Matcher bases its decisions on
structured rules provided by the system administrator for an specific DSP
Platform instance. The UML Class diagram for the DSP Broker, Matcher and
particiaping classes is depicted in figure
\ref{fig:DSPBroker-Matcher-Class-Diagram}, and describes how the DSP Broker
implementation is tightly coupled with the DSP Matcher, as well as to the
lists of DSP Components sorted by name and type.

\begin{figure}[!t]
  \centering
  \includegraphics[scale=0.5]{../diagrams/DSPBroker-Matcher-Class-Diagram}
  \caption{The DSP Broker-Matcher UML Class Diagram}
  \label{fig:DSPBroker-Matcher-Class-Diagram}
\end{figure}

In general, the DSP Matcher is composed by a list of MatchRule, which contains
a reference to a MatchCriteria and a MatchTarget. The former identifies the
rule for matching a given DSP Message based on the producer and consumer's
message type and location, while the latter describes the DSP Component that
will receive the DSP Message, and if its delivery is done by means of a gateway
component. In this way, the DSP Matcher can be seen as a function that takes a
DSP Message as an input and returns a list of DSP consumers, and the
matching configuration artifact provided by the administrator. See listing
\ref{file:dsp-matcher-config.xml}) for details.

\begin{table}
    \caption{DSP Matcher Rule Algorithm to filter Data Consumer components to
    receive a given Message}
    \begin{center}
        \begin{tabular}{lr}
          \textit{Data Consumers ( DSP Message ) := Verify the DSP
          Message's Header + Verify the matcher rules (which contains the list
          of consumers)}
        \end{tabular}
    \end{center}
    \label{tab:ysi-data-stream}
\end{table}

One example of such matching rule used by the DSP installed at the Gumstix can
be seen in listing \ref{file:dsp-matcher-config-gumstix.xml}. the Gumstix
system to configure the data delivery of the messages produced by the DSP
Component DSP Sonde Producer. It describes a list of maching rules that
reseambles a filtering system. This configuration contains 2 match rules: one
related to internal DSP Configuration and another for a remote destination.
The match criteria describes what the DSP Message Header properties must be,
while the match target is the destination of the message. One important
definition is the gateway Component Type, which defines the name of a remote
DSP Component that is capable of receiving a remote DSP Message.

Upon receiving all the matching rules and analyzing the DSP Message against
the rules, the DSP Broker selects a set of the running DSP components to receive
a copy of the DSP Message object in two different ways:

\begin{itemize}
  \item \textbf{In-memory local message delivery}: if the receiving DSP
  Component is located in the current local host, a deep copy of the instance
  of the DSP message is delivered;
    \begin{itemize}[label=\textbullet]
        \item IP Addresses are correctly resolved by using the Ethernet card in
        the device: localhost, 127.0.0.1, or the same IP Address for Producer
        and Receiver resolves into a local device. An example for such type of
        delivery can be seen in the match ruleID
        all messages sent to dsp manager of listing 
        \ref{file:dsp-matcher-config-gumstix.xml}.
    \end{itemize}
  \item \textbf{Serialized remote message delivery}: if the receiving DSP
  Component is located in a foreign/remote host, the message is serialized in a
  format defined by the transport protocol chosen the DSP Component responsible
  for the transport. The following section describes the existing DSP Data
  Transport component;     
  \begin{itemize}[label=\textbullet]
        \item IP Address from the Producer and Consumer are different, and are
        not resolved to be in the same host. An instance of such match rule can
        be seen in the match rule ID send-message-to-remote url dsp of the code
        snippet \ref{file:dsp-matcher-config-gumstix.xml}.
    \end{itemize}
\end{itemize}

\subsection{Sending DSP Messages through the Wire}

In order to send messages through the wire, the DSP Platform uses two symmetric
components that serialize and deserialize DSP Messages to XML and back to Java
POJO, respectively, and vice-versa. Named DSP Wire Transport Client and DSP
Wire Transport Server \cite{netbeams2009}, these components use the HTTP
\cite{http} protocol to transport the serialized version of the DSP Messages.
Figure \ref{fig:DSP-to-DSP-Remote-Communication} depicts the scenario where a
DSP Sensor Node transmits a DSP Message through an HTTP POST request.

\begin{figure}[!t]
  \centering
  \includegraphics[scale=0.6]{../diagrams/DSP-to-DSP-Remote-Communication}
  \caption{The remote communication between two DSP Components using the HTTP
  Protocol}
  \label{fig:DSP-to-DSP-Remote-Communication}
\end{figure}

\begin{itemize}
  \item \textbf{DSP Wire Transport Client}: responsible for making HTTP POST
  Requests to the service provided by the DSP Wire Transport Server component.
  The body of the HTTP POST request contains a serialized version of an
  instance of a DSP MessagesContainer in XML format;
  \item \textbf{DSP Wire Transport Server}: exposes a Web Server listening to
  port 8080 that receives the HTTP POST request from the Client. Upon receiving
  the request, the server deserializes the DSP MessagesContainer back to the
  in-memory POJO and proceeds with the delivery to the DSP Broker. At this
  situation, the component packages acknowledgement messages and any queued
  message to the client and piggybacks \cite{xml-piggybacking} it to the HTTP
  response body.
\end{itemize}

As described by \cite{netbeams2009}, NetBEAMS is a single-hop sensor network
whose sensors transmits data from the sensors to the server. Upon receiving the
data stream, the DSP Wire Transport Server component desirializes the DSP
MessagesContainer and its enclosing DSP Messages back to a Java class instance
and delivers each of the DSP Messages to the DSP Broker, which delivers a
copy of the message to each DSP Component listed by the maching rule. 
%This
%process is depicted by the UML Sequence Diagram \cite{uml} in figure \ref{fig:}

Whenever a DSP bundle producer is ready to transmit messages, they are added
into a DSP Messages Container (see UML Class Diagram in figure
\ref{fig:DSP-Messages-Classes}), which is serialized in XML to be transmitted.
Listing \ref{file:dsp-message-serialized-ysi} shows an instance of a DSP
MessagesContainer in serialized in XML, ready for transmission. In fact, the
listing is an example of the transmitted data collected from a sensor located
at the 192.168.0.103 to be transmitted to the host 192.168.0.106, and the
process of transmission in the server-side can be followed in figure
\ref{fig:From-DSPSondeComponent-to-DSPWireTransportClient-Sequence}.

\begin{figure}[h]
  \centering
  \includegraphics[scale=0.5]{../diagrams/From-DSPSondeComponent-to-DSPWireTransportClient-Sequence}
  \caption{The UML Sequence Diagram for the production of a YSI Message}
  \label{fig:From-DSPSondeComponent-to-DSPWireTransportClient-Sequence}
\end{figure}

The Sonde DSP Component is responsible for starting a worker thread called
SondeProducer, which is responsible for constantly read from the RS-232 serial
port by using an instance of the SerialReader class. This process depicts the
production of the data flow "flow 1" of figure
\ref{fig:netbeams-software-stack}, which is a collection of SondeDataType for
each of the measurement data read on a SondeDataContainer. As depicted in
figure \ref{fig:From-DSPSondeComponent-to-DSPWireTransportClient-Sequence}, it
first read the serial connection using the class SerialReader, which parses 
the data stream. After parsing the data, the reader creates an instance of the
class SondeDataType, the POJO responsible for carrying out the measurement
data. As described earlier, the data is added into the SondeDataContainer to
dispatched to the DSP Broker, whose responsibility is to get the list of DSP
Components that is setup to receive a copy of the DSP message (see previous
code snippet). As for the setup done for the Gumstix, the measurement message
is supposed to be sent to a remote destination of host 192.168.0.7. At this
point, the DSP Broker makes a decision to send the DSP Message to the default
gateway component DSPWireTransportClient.

\begin{figure}[!t]
  \centering
  \includegraphics[scale=0.5]{../diagrams/From-DSPWireTransport-Client-to-DSPWireTransportServer-Sequence}
  \caption{The UML Sequence Diagram describing the Wire Transport Server
  receiving a DSP Message}
  \label{fig:From-DSPWireTransport-Client-to-DSPWireTransportServer-Sequence}
\end{figure}

After the message arrives in the DSPWireTransportClient, the message is
maintained in a temporary repository of outbound DSP Messages called
MessagesQueues after the message is hashed by the destination IP address and
the DSP Client sender using an instance of QueueMessageData. As shown in figure
\ref{fig:From-DSPWireTransport-Client-to-DSPWireTransportServer-Sequence}, the
DSP TransportSender is a worker thread that is started by the DSP Wire
Transport Client and is constantly transmits the data in the outbound queues
after a configurable rate. It makes a transport of the message and receives
the response, updates the state of the messages to transmitted, to finally
send the response messages to the DSP Message Broker.

While the DSPWireTransportClient makes an HTTP Request with the DSP Message,
the DSPWireTransportServer's servlet receives the request. It first
deserializes the MessagesContainer and delivers each of its messages to the DSP
Broker. At this point, the server verifies if there are messages in the
outbound queue that are addressed for the host that requested the messages
delivery. By using a technique called piggybacking, the server adds to the HTTP
Response a serialized version of the MessagesContainer with any DSP Messages.
In addition to any DSP Message, the server prepares and sends acknowledgement
messages back to the requesting host to the host.

\begin{figure}[!t]
  \centering
  \includegraphics[scale=0.5]{../diagrams/From-DSPWireTransport-Server-To-DSP-Broker}
  \caption{The UML Sequence diagram describing the DSP
Message delivery on the server-side and the response}
  \label{fig:From-DSPWireTransport-Server-To-DSP-Broker}
\end{figure}

The gate "From DSP WireTransport Server to Broker Gate" defines the actual
delivery of the DSP Message to the DSP Broker. As this section is a tutorial
about the DSP Platform, the remaining steps will be explained on Section 3,
while showing the design and implementation of the component DSP Data
Persistence.

### 2009-11-14 docs/marcello-dissertation/thesis/chapter5-taxonomic-persistence-storage-selection-for-netbeams.tex
Updating the thesis and added chapters 2-5. Still revisions are needed in section 4 and 5, adjusting images and tables. Images are in the eps format.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@642 559edde2-a4d4-44e0-96fc-4035e40d7396

% main.tex, to be used with thesis.tex
% This contains the main work of your thesis.

%\bibliography{thesis}  % uses the references stored in Chapter1Radar.bib

\chapter{Data Persistence Technology for Sensor Networks: An Empirical-based
Choice for NetBEAMS}

This chapter analyzes each of the taxonomies defined in chapter
\ref{chap:taxonomies} against the requirements defined by our case study,
NetBEAMS, in chapter \ref{sec:problem-requirements}. Moreover, it evaluates
different database technologies against on those taxonomies, and tries to select
one of them which can be a candidate for an experimental analysis.

First, the contender technologies are as follows:

\begin{itemize}
  \item MySQL
  \item Microsoft SQL 
  \item Berkeley TinyDB
  \item MongoDB
  \item CouchDB
  \item IBM DB2
\end{itemize}

\section{Analysis of the Purpose of Sensor Data Taxonomy}

Considering the infrastructure of NetBEAMS, as described in the previous
chapter, and the characteristics of the sensor network defined at the SF-BEAMS
\cite{sfbeams2006} site, the data characteristics are as follows:

\begin{itemize}
  \item Data is generated by sensor devices, such as the YSI sonde
  \cite{YSI-Sonde}, and manually collected by using a laptop to the network
  sink at the RTC laboratories;
  \item Upon data reception, the RTC staff index and archive the data for
  for distribution using the OPEnDAP format;
  \item Another approach for data collection of data generated by the SF-BEAMS
  is the use of NetBEAMS \cite{netbeams2009}, as described in section
  \ref{netbeams-architecture}.
\end{itemize}

Given the facts collected from the observations of the execution of NetBEAMS
through the SF-BEAMS sensor network, we can say that the purpose of collected
data is \textbf{Data Archival}.

\subsection{Technology Analysis}

Any of the contenders are options to cover for Data Archival.

\begin{itemize}
  \item \textbf{MySQL}: Supports
  \item \textbf{Microsoft SQL}: Supports
  \item \textbf{Berkeley TinyDB}: Supports
  \item \textbf{MongoDB}: Supports
  \item \textbf{CouchDB}: Supports
  \item \textbf{IBM DB2}: Supports
\end{itemize}

\section{Analysis of the Location of Sensor Data Taxonomy}

Some observations regarding the location of the collected sensor data for
NetBEAMS were described in the previous section and summarized as follows:

\begin{itemize}
  \item NetBEAMS's architecture is based on a single-hop star infrastructure 
   with a direct network sink being the RTC server \ref{sec:sn-infrastructure}.
\end{itemize}

Based in the observations described, the type used for storing
data can be either the \textbf{External Storage} or the
\textbf{Data-Centric Storage}. The former is the current approach used by the
SF-BEAMS sensor network, while the latter adopted as an improvement of the
persistence infrastructure for NetBEAMS.

\subsection{Advantages of the External Storage Approach}

The External Storage approach is the most common way to implement persistence
for any type of systems, including sensor networks, due to its the simplistic
data management. Furthermore, this type of setup is easier to use of
distributed systems approaches such as Replication to help preventing scaling
the system.

\subsection{Disadvantages of the External Storage Approach}

The most common problem related to a single External Data Storage is that
resources may run into data management problems such as full disk space or disk
failures.

\subsection{Advantages of the Data-Centric Approach}

The advantage of the Data-Centric approach lies on the way that data is
distributed into different locations. By using techniques of distributed
database systems, the approach improves overall performance of the use of
massive data sets since the operations of the data are directed to specific
locations. Besides, this approach helps the network managers scale the data
storage just by adding new data nodes as required.

\subsection{Disadvantages of the Data-Centric Approach}

The data partitioning is a very restricted technology, mostly available to
advanced users of distributed database systems. The main users of wireless
sensor networks do not have such expertise.

\subsection{Technology Analysis}

\begin{itemize}
  \item \textbf{MySQL}: Supports
  \item \textbf{Microsoft SQL}: Supports
  \item \textbf{Berkeley TinyDB}: Supports
  \item \textbf{MongoDB}: Supports
  \item \textbf{CouchDB}: Supports
  \item \textbf{IBM DB2}: Supports
\end{itemize}

\section{Analysis of the Query Processing Mechanism Taxonomy}

As a direct result from the previous section, the use of \textbf{Centralized Query
Processing} is indicated for NetBEAMS, given the fact that NetBEAMS contains
a single network sink, and therefore, a centralized location of the data.

\subsection{Advantages of the Centralized Query Processing}

Centralized data management and query processing is simpler than in in-network.
The centralized query processing can be

\subsection{Disadvantages of the Centralized Query Processing}

The creation of the so-called Funneling Affect, since the point-of-traffic is
concentrated in the centralized system. Any data exchange will be using the
same channel of data exchange.

\subsection{Technology Analysis}

\begin{itemize}
  \item \textbf{MySQL}: Supports
  \item \textbf{Microsoft SQL}: Supports
  \item \textbf{Berkeley TinyDB}: Supports
  \item \textbf{MongoDB}: Supports
  \item \textbf{CouchDB}: Supports
  \item \textbf{IBM DB2}: Supports
\end{itemize}

\section{Analysis of the Data Model Taxonomy}

One of the most common practices in the area of database system is to use the
relational model to persist data, although the application of the system may
not fit to solve the problem. In fact, the main users of sensor networks may
not hold any expertise in database systems or data modeling, given they come
from different science areas. Taking NetBEAMS as an example, we see a Sensor
Network managed by Marine Biologists without expertise in Data Management,
Modeling systems, and for this reason, one of the requirements of the system is
the use of schema-less approaches. The \textbf{Key-Value-Pair Data Model} or the
\textbf{Document-Oriented Data Model} seems to be the simplest choices of the
models.

\subsection{Analysis of the Schema-Dependent Models}

Considering the inception of a relational data model \cite{relational-model} and
the use of the YSI Sonde data as the main entity in the system, let figure 
\ref{fig:Relational-Model-Original} represent a prototype of the relational
model, after passing through the process of normalization
\cite{db-normalization}.

\begin{figure}
  \centering
  \includegraphics[scale=0.65]{../diagrams/Relational-Model-Original}
  \caption{Relational Data Model for NetBEAMS - A first prototype}
  \label{fig:Relational-Model-Original}
\end{figure}

Supposing a new type is added into the system, let the refactored
version of the relational model be depicted in figure
\ref{fig:Relational-Model-Addition-Modified}.

\begin{figure}
  \centering
  \includegraphics[scale=0.65]{../diagrams/Relational-Model-Addition-Modified}
  \caption{Relational Data Model for NetBEAMS - Modified version with new
  entity}
  \label{fig:Relational-Model-Addition-Modified}
\end{figure}

\begin{itemize}
  \item Constant data schema changes require constant database normalization
  process, changes to structure, database management, etc;
  \item Some research have shown that the Relational Model does not fit the
  nature of collected data from wired or wireless sensor networks. They usually
  does not support time-series data nor provenance.
  \item Even project aiming at updating the relational sensor database systems
  and SQL clauses have been proposed. 
\end{itemize}

Other projects have been using the XML data models. It falls into the same
category as the Relational Model, since the XML documents, when in the context
of a database, must be complaint to an XML Schema. As shown in section 3, this
model uses XPath technology for querying documents, although some hibrid
technologies may still use SQL for that matter.

\subsection{Analysis of the Schema-less Models}

The advance of Cloud Computing, and dynamic web applications have motivated the
database community to implement different mechanisms for persisting data
without the heavy-process of data modeling. Given the fact that sensor networks
may be updated at any time by the addition of new sensor types, the use of
moding approaches such as Key-Value Pairs have been proposed []. One could be
tempted to implement a KVP over relational \ref{fig:KVP-on-Relational-Model}:

\begin{figure}
  \centering
  \includegraphics[scale=0.65]{../diagrams/KVP-on-Relational-Model}
  \caption{KVP Data Model implementation using Relational Model}
  \label{fig:KVP-on-Relational-Model}
\end{figure}

\begin{itemize}
  \item tules are not complaint to the normalization rules;
  \item difficult access to key and values
\end{itemize}

The real use of KVP and Document-Oriented data models mitigates the problem of
data modeling process.

\subsection{Technology Analysis}

\begin{itemize}
  \item MySQL: NO SUPPORT
  \item Microsoft SQL: NO SUPPORT
  \item Berkeley TinyDB: NO SUPPORT
  \item \textbf{MongoDB}: Supports Both
  \item \textbf{CouchDB}: Supports Both
  \item IBM DB2: NO SUPPORT
\end{itemize}

\section{Analysis of the Database System Organization Taxonomy}

Sensor Network data can be saved in either Centralized or Distributed Systems.
While Centralized Database Systems are easier to manage, it may face challenges
regarding its data. For this reason, the use of partitioning or sharding would
be an addition to persist sensor data.

\subsection{Advantages of Database Partitioning and Sharding}

\begin{itemize}
  \item Data-Centric queries and data use;
  \item Solves bottleneck problems related to reads/writes;
  \item Decrease the funneling effect by directing queries to given data
  partition;
\end{itemize}

\subsection{Disadvantages of Database Partitioning and Sharding}

\begin{itemize}
  \item Very advanced topics in Database System;
  \item May be easy of use, depending on the application.
\end{itemize}

\subsection{Technology Analysis}

\begin{itemize}
  \item \textbf{MySQL}: Supports
  \item \textbf{Microsoft SQL}: Supports
  \item \textbf{Berkeley TinyDB}: Supports
  \item \textbf{MongoDB}: Supports
  \item \textbf{CouchDB}: Supports
  \item \textbf{IBM DB2}: Supports
\end{itemize}

\section{Other Non-Functional Analysis}

\begin{itemize}
  \item Given the fact of the nature of the project, the technology to be used
  must be \textbf{open-source} \cite{open-source}, that is, no costs involved
  in the adoption of such technology;
  \item A solution for persisting collected sensor data must be not only limited
  to a technology that provides data access, such as SQL, but also by other
  \textbf{data access mechanisms} such as through \textbf{native APIs}. The
  reason is that part of the sensor network data does not have skills in
  database models, but may have skills with programming languages;
  \item Supports hot backup without service interruption.
\end{itemize}

\subsection{Technology Analysis}

Most of the technologies listed provides access to the data sets through the
use of drivers in different programming and scripting languages such as Java,
Python, Perl, etc.

\begin{itemize}
  \item \textbf{MySQL}: Supports
  \item \textbf{Microsoft SQL}: Supports
  \item \textbf{Berkeley TinyDB}: Supports
  \item \textbf{MongoDB}: Supports
  \item \textbf{CouchDB}: Supports
  \item \textbf{IBM DB2}: Supports
\end{itemize}

\section{Global Analysis Results and Technology Selection}

\begin{table}
    \label{tab:ysi-data-distribution}
    \caption{Amount of data produced by the RTC's YSI sondes}
        \begin{center}
        \begin{tabular}{|c|c|c|c|c|c|c|}\hline 
        \textbf{Database} & \textbf{MySQL} & \textbf{MicrosoftSQL} & \textbf{TinyDB} &
        \textbf{MongoDB} & \textbf{CouchDB} & \textbf{IBM DB2}\\\hline 
        Centralized Query & + & + & + & + & + & + \\\hline 
        Distributed System & ++ & ++ & - & ++ & ++ & +\\\hline 
        Schema-less & - & - & - & + & + & +\\\hline 
        Provenance Support & + & + & + & + & + & +\\\hline 
        NO-SQL Query & - & - & - & + & + & +\\\hline 
        Data Partition & + & + & - & + & + & +\\\hline 
        Hot Backup & + & + & - & - & - & +\\\hline 
        Export Capability & + & + & - & + & + & -\\\hline 
        Programming Lang. & + & + & + & + & + & +\\\hline
        Script Lang. & + & + & - & + & + & +\\\hline
        Open-Source & + & - & + & + & + & -\\\hline
        \end{tabular}
        \end{center}
\end{table}

### 2009-11-14 docs/marcello-dissertation/thesis/chapter9-appendix.tex
Updating the thesis and added chapters 2-5. Still revisions are needed in section 4 and 5, adjusting images and tables. Images are in the eps format.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@642 559edde2-a4d4-44e0-96fc-4035e40d7396

\chapter{Appendix - Source Code / Artifacts}

This section includes the source-code for the DSP Data Persistence. Each
section details one single main artifact used on the development of the
component. However, only the main ones will be added into this document, while
the remainder can be downloaded directly from the Subversion repository.

The source code is located in the Subversion repository for the NetBEAMS
project. This is how it is structured:

\section{DSP Data Persistence Component}

This artifact defines the ANT build file.

\lstinputlisting[language=Ant,label=file:dsp-build.xml,caption=Build
system using Apache ANT]{../../../../netbeams/versions/v2/apps/osgi-bundles/dsp/DSPDataPersistence/build.xml}

%\lstinputlisting[language=bash,label=file:osgi-manifest,caption=OSGi
%Manifest
%Descriptor]{../../../../netbeams/versions/v2/apps/osgi-bundles/dsp/DSPDataPersistence/META-INF/MANIFEST.MF}

\lstinputlisting[language=XML,label=file:dsp-config.xml,caption=DSP
Deployment
Configuration]{../../../../netbeams/versions/v2/config/deployment/config.xml}

\lstinputlisting[language=XML,label=file:dsp-matcher-config.xml,caption=DSP
Matching Rules]{../../../../netbeams/versions/v2/config/deployment/matcher_config.xml}

\lstinputlisting[language=XML,label=file:dsp-matcher-config-gumstix.xml,caption=DSP
Matching Rules
for
Gumstix]{../../../../netbeams/versions/v2/config/deployment/matcher_config-gumstix.xml}

%TODO: add bootstrap message in case it is implemented

\lstinputlisting[language=Java,label=file:dsp-activator,caption=OSGi
Activator]{../../../../netbeams/versions/v2/apps/osgi-bundles/dsp/DSPDataPersistence/src/org/netbeams/dsp/persistence/osgi/DataPersistenceActivator.java}

\lstinputlisting[language=Java,label=file:dsp-component-service,caption=OSGi
Service - Component]{../../../../netbeams/versions/v2/apps/osgi-bundles/dsp/DSPDataPersistence/src/org/netbeams/dsp/persistence/controller/DSPDataPersistence.java}

\lstinputlisting[language=Java,label=file:dsp-mongo-service,caption=DSP Data to
Mongo DB
Service]{../../../../netbeams/versions/v2/apps/osgi-bundles/dsp/DSPDataPersistence/src/org/netbeams/dsp/persistence/controller/DSPMongoCRUDService.java}

\lstset{language=XML,morecomment=[s]{!--}{--},label=file:dsp-message-serialized-ysi,caption=DSP
Message with a YSI sonde data payload}
\begin{lstlisting}
<?xml version="1.0" encoding="UTF-8" standalone="yes"?>
<MessagesContainer uudi="24929c29-60ee-4d17-af08-64d9446277ef"
        creationTime="2009-03-06T15:17:18-0800" destinationHost="192.168.0.106">
   <MeasureMessage ContentType="org.netbeams.dsp.ysi"
         messageID="435a61f6-370f-458d-aeb7-6e92270a79cb">
      <Header>
         <CreationTime>1236381438480</CreationTime>

         <Producer>   
            <ComponentType>
                 org.netbeams.dsp.platform.management.component.ComponentManager
            </ComponentType>

            <ComponentLocator>
               <ComponentNodeId>1234</ComponentNodeId>
               <NodeAddress>192.168.0.103</NodeAddress>
            </ComponentLocator>
         </Producer>

         <Consumer>
            <ComponentType>
                 org.netbeams.dsp.wiretransport.client
            </ComponentType>

            <ComponentLocator>
               <NodeAddress>LOCAL</NodeAddress>
            </ComponentLocator>
         </Consumer>
      </Header>
      <Body>
         <SondeDataContainer>
            <soundeData date="15:17:18" time="03-06-2009">  
               <Temp>21.20</Temp>
               <SpCond>193</SpCond>
               <Cond>179</Cond>
               <Resist>5588.40</Resist>
               <Sal>0.09</Sal>
               <Press>0.084</Press>
               <Depth>0.059</Depth>
               <pH>7.98</pH>
               <phmV>-79.6</phmV>
               <ODOSat>99.5</ODOSat>
               <ODOConc>8.83</ODOConc>
               <Turbid>0.4</Turbid>
               <Battery>8.7</Battery>
            </soundeData>
         </SondeDataContainer>
      </Body>
   </MeasureMessage>
</MessagesContainer>
\end{lstlisting}

\section{Experiment Implementation}

\lstinputlisting[language=Bash,label=file:experiment-setup-executor,caption=Experiment
Implementation]{../../../../netbeams/versions/v2/persistence/run-persistence-experiment}

\lstinputlisting[language=Java,label=file:experiment-dsp-implementation,caption=Experiment
Implementation]{../../../../netbeams/versions/v2/apps/osgi-bundles/dsp/DSPDataPersistence/src/org/netbeams/dsp/persistence/controller/DSPMessageToMongoDBExperiment.java}

### 2009-11-15 docs/marcello-dissertation/thesis/chapter10-appendix.tex
Updating the previous chapters 4, 5, and adding 6 and 7 unfinished.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@645 559edde2-a4d4-44e0-96fc-4035e40d7396

--- 2009-11-14 docs/marcello-dissertation/thesis/chapter9-appendix.tex
@@
 Implementation]{../../../../netbeams/versions/v2/apps/osgi-bundles/dsp/DSPDataPersistence/src/org/netbeams/dsp/persistence/controller/DSPMessageToMongoDBExperiment.java}
+
+\lstset{label=file:mongodb-ysi-data-format,caption=JSON representation of the
+data produced by a YSI sensor on mongoDB}
+\begin{lstlisting}
+            {
+             "_id" :  ObjectId( "d36f4007b7e7ac4a03c60000")  , 
+             "sensor_ip_address" : "192.168.0.136", 
+             "message_id" : "7b6624d6-0ca1-4cba-a343-f166e88da73b",
+             "transaction_time" : 1252845473412 , 
+             "fact_time" : 1252845346000,
+             "latitude" : 12.3450, 
+             "longitude" : -123.4456 , 
+             "data" : {
+                       "temperature" : 45.01, 
+                       "sp_condition" : 37.6, 
+                       "condition" : 145.8, 
+                       "resistence" : 159.77, 
+                       "salinitude" : 0.0, 
+                       "pressure" : 0.391, 
+                       "depth" : 0.46, 
+                       "ph" : 5.64, 
+                       "pH_mv" : -62.1, 
+                       "odo_sat" : 89.7, 
+                       "odo_condition" : 59.34, 
+                       "turbidity" : 0.0,
+                       "battery" : 9.4
+                      }
+            }
+\end{lstlisting}
+
+\lstset{label=file:rtc-ysi-opendap,caption=HTTP GET Request x Response to the OPeNDAP server at SF-BEAMS}
+\begin{lstlisting}
+http://sfbeams.sfsu.edu:8080/opendap/sfbeams/data_ctd/rtc_ctd5-ysimoor/real-time/sfb_CTD5_PUF.dat.ascii?
+GET /opendap/sfbeams/data_ctd/rtc_ctd5-ysimoor/real-time/sfb_CTD5_PUF.dat.dods?&Month=11&Day=12 HTTP/1.1
+
+HTTP/1.x 200 OK
+Server: Apache-Coyote/1.1
+Last-Modified: Thu, 12 Nov 2009 17:55:06 GMT
+XDODS-Server: Server-Version-Unknown
+XOPeNDAP-Server: bes/3.5.1 freeform_handler/3.7.5, netcdf_handler/3.7.6
+XDAP: 3.2
+Content-Description: dods_data
+Content-Type: text/txt
+Date: Sun, 15 Nov 2009 03:08:45 GMT
+
+Dataset: sfb_CTD5_PUF.dat
+YSI_REALTIME_CSV.Month, YSI_REALTIME_CSV.Day, YSI_REALTIME_CSV.Year, YSI_REALTIME_CSV.Hour, YSI_REALTIME_CSV.Min, YSI_REALTIME_CSV.Sec, YSI_REALTIME_CSV.WaterTemperature, YSI_REALTIME_CSV.SpecificConductivity, YSI_REALTIME_CSV.Conductivity, YSI_REALTIME_CSV.Resistivity, YSI_REALTIME_CSV.TDS, YSI_REALTIME_CSV.Salinity, YSI_REALTIME_CSV.Pressure, YSI_REALTIME_CSV.Depth, YSI_REALTIME_CSV.pH, YSI_REALTIME_CSV.pHmV, YSI_REALTIME_CSV.Turbidity, YSI_REALTIME_CSV.ODOSaturation, YSI_REALTIME_CSV.ODO, YSI_REALTIME_CSV.Chlor, YSI_REALTIME_CSV.Chlor_RFU, YSI_REALTIME_CSV.Battery, YSI_REALTIME_CSV.InstSN
+11, 12, 2009, 16, 30, 51, 13.78, 4.3945, 3.4532, 0, 28.56, 28.35, 2.65, 2.651, 7.9, -54.7, 9, -99999, -99999, 3.7, 1, 9.4, -99999
+11, 12, 2009, 16, 36, 50, 13.78, 4.4095, 3.4642, 0, 28.66, 28.46, 2.628, 2.63, 7.91, -54.9, 11.9, -99999, -99999, 4, 1.1, 9.4, -99999
+11, 12, 2009, 16, 42, 51, 13.77, 4.3943, 3.4515, 0, 28.56, 28.35, 2.631, 2.632, 7.9, -54.9, 9.8, -99999, -99999, 3.6, 1, 9.4, -99999
+11, 12, 2009, 16, 48, 50, 13.76, 4.3893, 3.447, 0, 28.53, 28.32, 2.628, 2.63, 7.9, -54.8, 10.6, -99999, -99999, 3.8, 1, 9.4, -99999
+\end{lstlisting}

### 2009-11-15 docs/marcello-dissertation/thesis/chapter4-netbeams-in-depth-requirements-analysis-and-specification.tex
Updating the previous chapters 4, 5, and adding 6 and 7 unfinished.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@645 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 access to the archived data is done by using the OPeNDAP server is located at
-http://sfbeams.sfsu.edu:8080/opendap.
+http://sfbeams.sfsu.edu:8080/opendap, as a request and the related output for
+one of the YSI devices can be see in listing \ref{file:rtc-ysi-opendap}.
 
@@
 
-\begin{figure}[!b]
+\begin{figure}[!t]
   \centering
@@
 
+\newpage
+
 \subsection{Data Delivery Mechanism}
@@
 
-\begin{figure}[!t]
+\begin{figure}[!h]
   \centering
-  \includegraphics[scale=0.5]{../diagrams/DSP-Message-Broker-Abstraction}
+  \includegraphics[scale=0.8]{../diagrams/DSP-Message-Broker-Abstraction}
   \caption{A DSP Message Broker Abstraction}
@@
 
-\begin{figure}[!t]
+\begin{figure}[!b]
   \centering
-  \includegraphics[scale=0.5]{../diagrams/DSPBroker-Matcher-Class-Diagram}
+  \includegraphics[scale=0.6]{../diagrams/DSPBroker-Matcher-Class-Diagram}
   \caption{The DSP Broker-Matcher UML Class Diagram}
@@
 
-\begin{figure}[!t]
+\begin{figure}[!b]
   \centering
@@
 
-\begin{figure}[h]
+\begin{figure}[!h]
   \centering
@@
 
-\begin{figure}[!t]
-  \centering
-  \includegraphics[scale=0.5]{../diagrams/From-DSPWireTransport-Client-to-DSPWireTransportServer-Sequence}
-  \caption{The UML Sequence Diagram describing the Wire Transport Server
-  receiving a DSP Message}
-  \label{fig:From-DSPWireTransport-Client-to-DSPWireTransportServer-Sequence}
-\end{figure}
-
 After the message arrives in the DSPWireTransportClient, the message is
@@
 send the response messages to the DSP Message Broker.
+
+\begin{figure}[!t]
+  \centering
+  \includegraphics[scale=0.5]{../diagrams/From-DSPWireTransport-Client-to-DSPWireTransportServer-Sequence}
+  \caption{The UML Sequence Diagram describing the Wire Transport Server
+  receiving a DSP Message}
+  \label{fig:From-DSPWireTransport-Client-to-DSPWireTransportServer-Sequence}
+\end{figure}
 

### 2009-11-15 docs/marcello-dissertation/thesis/chapter5-taxonomic-persistence-storage-selection-for-netbeams.tex
Updating the previous chapters 4, 5, and adding 6 and 7 unfinished.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@645 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 model uses XPath technology for querying documents, although some hibrid
-technologies may still use SQL for that matter.
+technologies may still use SQL for that matter. In this way, an instance of
+such schema dependent types can be seen in figure
+\ref{fig:persistence-example-relational}.
+
+\begin{figure}[!h]
+  \centering
+  \includegraphics[scale=0.65]{../diagrams/persistence-example-relational}
+  \caption{Instance of a Relational Model Database Prototype}
+  \label{fig:persistence-example-relational}
+\end{figure}
+
+Although this data model have been used in different types of applications,
+developers have tried to use the Relational Model to model an application that
+could prevent any database change through the use of a technique that relates a
+key to a value, called Key-Value data model. With the advent of Internet
+applications, the need of such a model that does not need constant refactoring
+led developers to propose such approach using the relational model as
+documented in technical articles such as \cite{db-kvp-in-relational01} and
+\cite{db-kvp-in-relational02}. Based in these articles, consider the relational
+model depicted in figure \ref{fig:KVP-on-Relational-Model} as model for the
+persistence data for our case study using the key-value strategy. The problem
+with it is that key repetition occurs, since the nature of time-series data
+requires the use of a timestamp key for each property of the sensor. Therefore,
+this strategy is not well-suited to provide persistence data from collected
+sensor networks.
+
+\begin{figure}[!h]
+  \centering
+  \includegraphics[scale=0.5]{../diagrams/KVP-on-Relational-Model}
+  \caption{KVP Data Model implementation using Relational Model}
+  \label{fig:KVP-on-Relational-Model}
+\end{figure}
+
+\begin{figure}[!h]
+  \centering
+  \includegraphics[scale=0.5]{../diagrams/persistence-example-relational-kvp}
+  \caption{KVP over Relational Model Database Prototype}
+  \label{fig:persistence-example-relational-kvp}
+\end{figure}
 
@@
 may be updated at any time by the addition of new sensor types, the use of
-moding approaches such as Key-Value Pairs have been proposed []. One could be
-tempted to implement a KVP over relational \ref{fig:KVP-on-Relational-Model}:
+moding approaches such as Key-Value Database have been proposed \cite{db-kvp}.
 
-\begin{figure}
+\begin{itemize}
+  \item entities from the same domains are placed into a bucket;
+  \item entities have a set of attributes and relating values
+  \item entities with different set of attributes may be contained in the same
+  bucket, since there is no schema to govern the bucket items restriction.
+\end{itemize}
+
+For instance, all data needed for the YSI sonde data, as well as all necessary
+provenance data that describes the data, are represented by means of key-value
+pairs \ref{fig:persistence-example-kvp}.
+
+\begin{figure}[!h]
   \centering
-  \includegraphics[scale=0.65]{../diagrams/KVP-on-Relational-Model}
-  \caption{KVP Data Model implementation using Relational Model}
-  \label{fig:KVP-on-Relational-Model}
+  \includegraphics[scale=0.5]{../diagrams/persistence-example-kvp}
+  \caption{KVP Database instance Prototype}
+  \label{fig:persistence-example-kvp}
 \end{figure}
 
-\begin{itemize}
-  \item tules are not complaint to the normalization rules;
-  \item difficult access to key and values
-\end{itemize}
+\subsection{Technology Analysis}
 
-The real use of KVP and Document-Oriented data models mitigates the problem of
-data modeling process.
+The data models were compared by \cite{db-is-rdbs-dommed} and can be summarized
+as follows:
 
-\subsection{Technology Analysis}
+\begin{table}
+    \label{tab:ysi-data-distribution}
+    \caption{Schema-Dependent X Schema-less Databases Compared: Properties}
+    \begin{center}
+    \begin{tabular}{|p{210pt}|p{210pt}|}\hline
+    Schema-Dependent Databases & Schema-less Databases\\\hline
+    \begin{enumerate}
+      \item Real-world model by entities, classified in tables;
+      \item Tables composed by columns and rows. Rows are comprised by column
+      values, which have the same schema;
+      \item Data Model defined in advance with a schema, which contains
+      relationships and constraints to enforce data integrity;
+      \item Data represents ``natural'' entities, not application specific;
+      \item Data Normalization is the data structuring process to remove data
+      duplication, as well as establishing data associations through table
+      relationships.
+    \end{enumerate} 
+    & 
+    \begin{enumerate}
+      \item Real-world model by entities, classified in Domains;
+      \item Domains are similar to tables, but like buckets that contains items
+      without a pre-defined schema, what enable them to have different schemas;
+      \item Items are identified by keys, as well as have a dynamic set of
+      attributes attached to it, however with no schema defined;
+      \item Items may represent not only the natural representation of data, but
+      also include application-specific data;
+      \item Since data may be repeated, no data normalization is done, so that
+      integrity is done in the application layer.
+    \end{enumerate}
+    \\\hline
+    \end{tabular}
+    \end{center}
+\end{table}
+
+\begin{table}
+    \label{tab:ysi-data-distribution}
+    \caption{Schema-Dependent X Schema-less Databases: Data Access}
+    \begin{center}
+    \begin{tabular}{|p{210pt}|p{210pt}|}\hline
+    Schema-Dependent Databases & Schema-less Databases\\\hline
+    \begin{enumerate}
+      \item The basic operations CRUD \footnote{Create-Retrieve-Update-Update
+      database operations} data are performed using a structured language such
+      as the SQL or XPath;
+      \item Query languages can access data from different tables through
+      joins, contains functions for aggregation and complex filter.
+    \end{enumerate} 
+    & 
+    \begin{enumerate}
+      \item The CRUD operations are performed via API \footnote{Application
+      Programming Interface} through programming languages;
+      \item Some technologies provide basic SQL-like syntax for filter criteria
+      with some predicates like =, !=, <, > that ca be applied;
+      \item The data and application integrity logic is placed in the
+      application layer.
+    \end{enumerate}
+    \\\hline
+    \end{tabular}
+    \end{center}
+\end{table}
+
+\begin{table}
+    \label{tab:ysi-data-distribution}
+    \caption{Schema-Dependent X Schema-less Databases: Application Interface}
+    \begin{center}
+    \begin{tabular}{|p{210pt}|p{210pt}|}\hline
+    Schema-Dependent Databases & Schema-less Databases\\\hline
+    \begin{enumerate}
+      \item Have their own specific API or through ODBC\footnote{Open Database
+      Connectivity};
+      \item Data is stored in a format that represents its natural structure,
+      and for this reason, in a single or distributed fashion.
+    \end{enumerate} 
+    & 
+    \begin{enumerate}
+      \item Systems tend to provide SOAP/REST \cite{http-rest};
+      \item \cite{db-is-rdbs-dommed} claims that data is stored in a more
+      effective way, requiring only code plumbing for the relational code;
+    \end{enumerate}
+    \\\hline
+    \end{tabular}
+    \end{center}
+\end{table}
 
@@
   \item \textbf{Microsoft SQL}: Supports
-  \item \textbf{Berkeley TinyDB}: Supports
+  \item \textbf{Berkeley TinyDB}: NO Supports
   \item \textbf{MongoDB}: Supports

### 2009-11-15 docs/marcello-dissertation/thesis/chapter6-data-persistence-component-design-and-architecture.tex
Updating the previous chapters 4, 5, and adding 6 and 7 unfinished.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@645 559edde2-a4d4-44e0-96fc-4035e40d7396

% main.tex, to be used with thesis.tex
% This contains the main work of your thesis.

%\bibliography{thesis}  % uses the references stored in Chapter1Radar.bib

\chapter{DSP Data Persistence: Component Design and Architecture}

In order to provide data to users, the RTC scientists use OPeNDAP \cite{opendap}
server to distribute collected data over the Internet. However, this work
proposes the use a KVP database system as described in the previous chapter,
giving options to advanced techniques to scale on data structure and
organization.

This section shows the proposed DSP Data Persistence component for the DSP
Platform as reviewed in chapter 4, as it is responsible for gathering
the DSP Messages sent from any remote DSP instance. First, section 3.1 briefly
describes main events, processes, input and outputs generated by analyzing a
scenario of remote communications between sensors and the DSP Client and
Server nodes. Then, section 3.2 shows the requirements specification based on
the analysis of section 3.1 and the documentation of the DSP Platform of
section 2.4.

\section{Business Process Analysis}

As described in section 2.4, the Data Sensor Platform was designed to support
an environmental sensor network based in San Francisco Bay, composed by
different types of sensors. The process of extracting the measurements from the
sensors to be reused using the DSP Platform can be summarized following the UML
Business Process diagram depicted in figure
\ref{fig:fig:dsp-persistence-business-process}. It is important to note that
this diagrams details remote communication between sensor devices and a
centralized server, as part of the main scenario to solve the problem of data
persistence.

\begin{figure}[!b]
  \centering
  \includegraphics[scale=0.5]{../diagrams/DSP-DataPersistence-Business-Diagram}
  \caption{UML Business Process Model diagram - Adding Persistence for NetBEAMS}
  \label{fig:dsp-persistence-business-process}
\end{figure}

At the event of data collection, the DSP Client receives measurements collected
from each of the sensor devices. Each of the co-located DSP Client processes
the collected measurements and transforms them into the so-called DSP Message
(section 2.4), which embodies the actual data values. After proper data
processing, transformation and categorization, the DSP Message is added into an
in-memory data storage identified (section 2.4). Then, as an asynchronous data
transmission event occurs with an specified rate, transferring all the queued
messages to be processed by a DSP Server host. As it is detailed in image
3.1.1, the server is centralized and is able to concurrently process messages.

This work proposes the addition of a database system that is responsible for
storing the gathered measurement data from sensor hosts on a given network. As
it is shown in image 3.1.1,  the processed measurement data from sensors 1
through N must be collected in the same fashion and categorized as necessary.
In this way, end users such as Biologists and/or Programmers can access the
collected measurements directly from the Database. Furthermore, the same
measurement data is supposed to be exported to different formats as the users
find necessary.

As a matter of fact, the measurements data collected from the sensor devices
are simple  properties with and associated values as described on section 2.1.
According to the descriptions of the DSP Platform and the DSP Messages
structure on section 2.4, the data is represented in programming language
formats such as Java or in a serialized version in XML. Similarly, the
collected data contains measurements that uniquely identifies them, together
with other properties such as time stamps of data creation and collected, as
well as the the IP address of the DSP Client that produced the data. Therefore,
these are important properties for the proposed solution of this work.

\section{Requirements Specification}

By analyzing the main scenario of data process on section 3.1, this section
lists the main use cases that the system and the primary users require. One of
the primary requirements of the system is to enable the use a persistence
system that does not promote the use of specialized professionals to deal with
data modeling, as it is required in most projects using the Structured Query
Language (SQL) [db01], used to extract data from a Relational Database [db01].

\subsection{Functional Requirements}
As part of the user stories related to data access, figure
\ref{fig:DSP-Data-Persistence-UseCases-Diagram-Users} describes the use cases
that the end users can perform with the system.

\begin{figure}[!b]
  \centering
  \includegraphics[scale=0.5]{../diagrams/DSP-Data-Persistence-UseCases-Diagram-Users}
  \caption{UML Use Case diagram for Persistence Functions}
  \label{fig:DSP-Data-Persistence-UseCases-Diagram-Users}
\end{figure}

The functional requirements listed in this section are in the format of UML
User Cases [se05]. Figure
\ref{fig:DSP-Data-Persistence-UseCases-Diagram-System} summarizes the use
cases that the system, as the main actor of the system, must perform:

\begin{figure}[!b]
  \centering
  \includegraphics[scale=0.5]{../diagrams/DSP-Data-Persistence-UseCases-Diagram-System}
  \caption{UML Use Case diagram for System Persistence Functions}
  \label{fig:DSP-Data-Persistence-UseCases-Diagram-System}
\end{figure}

\subsection{Non-Functional Requirements}

\begin{itemize}
  \item The DSP Data Persistence component must maintain the received DSP
  Measurement Messages in-memory for a specific rate in order to decrease the
  write load on the database;
  \item The Persistence storage system must be able to cope with dozens of DSP
  Messages;
  \item The Persistence storage system must be able to save measurements quickly;
  \item The Persistence storage system must be able to scale without too much
  architectural changes;
  \item The Persistence storage system must not impose the knowledge of
  complicated database system languages such as SQL [db01] or XML XPath [db05],
  but use access through API calls, since it is closer to researchers. 
\end{itemize}

\section{Data Model Design}

After analyzing section 2.1 along with the functional and non-functional
requirements described on section 3.2.2, this work considered the
Key-Value-Pair Data model as described on Section 2.5. Taking into account the
mongoDB architecture and the properties of a DSP Message (see section
"Acquiring the properties of a DSP Message Content" at DSPDataPersistence),
here are the conventions followed on revision r585. The following is the list
of properties that composes the Key of a document:

\begin{itemize}
  \item \textbf{sensor\underline{ }ip\underline{ }address}: it's extracted from
  the DSP Message Produce and identifies which sensor generated the sampling;
  \item \textbf{message\underline{ }id}: it's extracted from each of the messages contained
  in the DSP Message Container;
  \item \textbf{transaction\underline{ }time}: it's extracted from the DSP message container
  creation time and it is used to identify when the transaction occurred (see Section 2.2);
  \item \textbf{fact\underline{ }time}: it's extracted from the SondeDataContainer's date
  and time, and identifies the time in which the collected data occurred (see Section 2.2);
  \item \textbf{latitude}: the latitude variable of the position;
  \item \textbf{longitude}: the longitude variable of the position.
\end{itemize}

The definition of the Value of a document is as follows:

\begin{itemize}
  \item data: this key defines the values of the document, and will have every
  different property of the sensor (see Section 2.4):
   \begin{itemize}
     \item battery
     \item temperature
     \item salinity
     \item \ldots
   \end{itemize}
\end{itemize}

Some remarks about the creation of the items contained with the collections. In
case a message container contains multiple readings in a message container,
each item will be counted as individual snapshots of the data. Also, it is
important to note that the description of the data follows no pattern other
than the description of the properties of the sensor device. However, some
studies have shown that when disk space maximization is required, the name of
database columns are important.

\section{High-Level System Architecture}

Considering the business analysis described in section 3.1, the addition of a
data persistence layer to the DSP Platform can be achieved by adding Data
Consumer (DC) as described in section 2.4. Based on the Requirements
specification described in section 3.2, this component must be responsible for
providing the processing of the DSP Measure Messages into Database-ready
representation, as well as providing the necessary tools for the transmission
of the data to a database system. Figure
\ref{fig:NetBEAMS-Persistence-Server-Node-Components} provides an
architectural view of the DSP Server node with the loosely-coupled components,
highlighting the inclusion of the DSP Data Persistence.

\begin{figure}[!b]
  \centering
  \includegraphics[scale=0.5]{../diagrams/NetBEAMS-Persistence-Server-Node-Components}
  \caption{UML Components diagram of the DSP Data Collector}
  \label{fig:NetBEAMS-Persistence-Server-Node-Components}
\end{figure}

As it is depicted on image 3.4.1, the DSP Data Persistence component can be
added into the system as an independent plug-and-play component. In addition,
the database system and the necessary drivers are also added without any
changes to the existing DSP infrastructure. First, as DSP Messages are
delivered to the server through the DSP Wire Transport Server, as shown by flow
1, the DSP Platform is responsible to forward a copy of the Measurement
messages directly to the DSP Data Persistence as shown by flow 2. Then, the
sole responsibility of the DSP Data Persistence is to transfer the received
measurement messages to the database system by using the connection drivers, as
depicted by flow 3.

The following sections focus on the DSP Data Persistence component design, and
in this way, any references to the DSP Platform can be seen on section 2.4.

\section{DSP Data Persistence: OSGi-DSP Bundle Design}

The new proposed Data Persistence component must follow the DSP requirements
for a new DSP Component. As described on section 2.4, any DSP component follows
the design-pattern of Data Producer/Consumer. Similarly, section 3.2 described
the requirements that the DSP Data Persistence must handle, taking into account
the data model and database connection.

After reviewing the requirements of section 3.2, the resulting model maintains
the dependency among 4 different major Java packages. The addition of a DSP
Data Persistence component depends on 2 existing packages, namely the OSGi
Framework and the DSP Platform as in any DSP Component, and an additional
package for the database access driver API. The UML Package dependency diagram
of image 3.5.x shows these dependencies, which are used to better track changes
among the packages.

\begin{figure}[!h]
  \centering
  \includegraphics[scale=0.5]{../diagrams/DSP-Data-Persistence-Packages-Dependency}
  \caption{UML Components diagram of the DSP Data Collector}
  \label{fig:DSP-Data-Persistence-Packages-Dependency}
\end{figure}

The following sections details each step of the DSP Data Persistence activation
and parallel execution of the contained classes.

\subsection{DSP Data Persistence Activation and Message Delivery}

As shown on section 2.4.3, the DSP Platform is composed of loosely-coupled
components. In this way, this section describes the addition of the DSP Data
Persistence, as shown on the UML Class diagram of image 3.5.1.1.

\begin{figure}[!h]
  \centering
  \includegraphics[scale=0.5]{../diagrams/DSP-DataPersistence-Activator-Class-Diagram}
  \caption{UML Class diagram for the DSP Data Persistence Activator and Component}
  \label{fig:DSP-DataPersistence-Activator-Class-Diagram}
\end{figure}

The DSP Data Persistence Activator is the main class responsible for the
activation of the DSP Data Persistence component, as it extends the OSGi class
BundleActivator and maintains a reference of the DSP Data Persistence class.
Furthermore, the activator maintains references to the classes BundleContext
and ServiceReference both from the OSGi Framework. The former is responsible
for maintaining the link with the OSGi Framework, while the latter is used to
register the DSP Component as a service. Therefore, it is clear that the class
DSP Data Persistence Activator is the main communication interface between the
DSP Platform and the OSGi Framework.

While the DSP Data Persistence Activator is responsible for the orchestration
of the classes during the execution on the OSGi Framework, the class DSP Data
Persistence extends from the DSP Platform class DSPComponent. For this reason,
this class inherits all the behavior described on section 2.4 and can be in the
roles of Data Consumer and a Data Producer.

\begin{itemize}
  \item \textbf{DSP Data Persistence as Data Consumer}: receives any measurement
  message from remote sensors;
  \item \textbf{DSP Data Persistence as Data Producer}: transforms any
  measurement message received into a format that must be ready to be saved on the
  database, based on the description of the data model of section 3.3.
\end{itemize}

The DSP Data Persistence Activator follows the states described on the state
diagram of image 2.4.3.4, as it is an extension of the OSGI class
BundleActivator, as well as being responsible for stopping or starting the the
DSP Data Persistence component instance. As described on section 2.4, the
following steps must take place in order to activate and start the DSP Data
Persistence component, described partially on image 2.4.6.3:

\begin{enumerate}
  \item During the DSP Platform activation, it will get the name of the
  persistence component from the configuration artifact config.xml;
  \item After being selected based on the configuration priority, the DSP
  Bundle artifact is identified and installed, by creating an instance of the
  class DSP DataPersistence Activator and making a call to the method start(),
  as shown on image 3.5.1.2;
  \item During the activation, an instance of the DSP Data Persistence
  component is created and registered as an OSGi Service;
  \item Upon registering the DSP Data Persistence component, the DSP Platform
  "listens" the event serviceUpdate() and, and triggers the last operations to
  bootstrap the DSP Component:
   \begin{enumerate}
      \item Initializes the component by calling the method initComponent();
      \item Starts the component by calling the method startComponent();
      \item Bootstrap messages are delivered to the component.
   \end{enumerate}
\end{enumerate}

As the requirements defined on section 3.2 specifies the update of the DSP
Component during its initialization, an optional DSP Update Message will be
sent to the component during its initialization process, as described on
section 2.4. The initialization of the DSP Component Activator and the DSP
Component registration are shown on image 3.5.1.2.

\begin{figure}[!b]
  \centering
  \includegraphics[scale=0.5]{../diagrams/From-OSGi-Framework-to-DSP-Data-PersistenceActivator-Sequence-Diagram}
  \caption{UML Sequence diagram describing the DSP Data Persistence bundle activation}
  \label{fig:From-OSGi-Framework-to-DSP-Data-PersistenceActivator-Sequence-Diagram}
\end{figure}

The the gate "Start DSP Activator Gate" reaches the DSP Data Persistence, it
first creates the unique instance of the class DSP Data Persisence and uses
that instance to register it as the OSGi service through the call to the DSP
Platform helper class ActivatorHelper. At this point, the DSP Data Persistence
has been installed and is in the state "Active".

\subsection{Delivering Bootstrap and Measurement Messages}

Although the DSP Data Persistence component has been instantiated as of image
3.5.1.2, the process of delivering messages to the DSP Data Persistence has yet
to be completed. As defined by the requirements, the DSP Component must execute
its function in a concurrent fashion, using configuration parameters during its
bootstrap process. The class DSP Data Flusher is responsible for running as a
"worker thread", whose function is to verify if there are messages to be
flushed into the Database with the class Transient Persistence Layer. Image
3.5.2.1 describes the remaining UML Class diagram for the main classes of this
functionality.

\begin{figure}[!b]
  \centering
  \includegraphics[scale=0.5]{../diagrams/DSP-DataPersistence-Flusher-Classes}
  \caption{UML Class diagram with the DSP Data Flusher worker thread}
  \label{fig:DSP-DataPersistence-Flusher-Classes}
\end{figure}

The DSP Data Flusher is a thread that is designed to be in two different
states: RUNNABLE and TIMED\underline{ }WAITING. In the former, the DSP Data Flusher will be
contacting the Transient Persistence Layer to verify if there are any message
waiting to be sent to the Database service, while in the latter the DSP Data
flusher will be waiting for the next cycle defined by the bootstrap parameter
TRANSIENT\underline{ }DATA\underline{ }FLUSHER\underline{ }DELAY.

\begin{figure}[!b]
  \centering
  \includegraphics[scale=0.5]{../diagrams/DSP-DataPersistence-Flusher-State-Diagram}
  \caption{UML State diagram for the DSP Data Flusher Worker Thread}
  \label{fig:DSP-DataPersistence-Flusher-State-Diagram}
\end{figure}

Upon receiving a new DSP measurement message, the DSP Data Persistence
component must send its reference to the temporary memory space, as described
by the non-functional requirement on section 3.2. In this way, the DSP Data
Persistence component maintains a dependency on the Singleton class
TransientPersistenceLayer, which is responsible for keeping track of messages
that are received by working component. However, in order to add the DSP
message into its aggregated map, it first transforms the message into an
instance of the class PersistentMessageUnit, which has references to the class
regarding the location of the message,  SensorLocation.

The last paragraph of section 2.4.7 described the last steps of the Sequence
Diagram regarding DSP Message delivery to the DSP Broker on the DSP Server
host. For this reason, Image 3.4.1.2 continues the steps of the DSP Broker from
the gate "From DSP WireTransport Server to Broker Gate", showing the steps
taken to transfer a DSP Message to the Singleton class
TransientPersistenceLayer.

According to the DSP Broker model described on section 2.4.6, the DSP Matcher
needs to be updated with the addition of a matching rule that filters a copy of
any Measurement Message (section 2.4.4) to the DSP Data Persistence component
designed. As it is shown in figure
\ref{fig:From-DSP-Broker-To-DSPDataPersistence-General-Sequence}, the worker
Thread DSP Data Flusher is the link between the Transient and Persistent
layers, as it depends on the references from the classes
TransientPersistenceLayer and the DSPMongoCRUDService. The former maintains
DSP Messages in memory wrapped by an instance of PersistenceMessageUnit and
the latter is responsible for sending the MessageContent from the DSPMessage
body to the Database.

\begin{figure}[!b]
  \centering
  \includegraphics[scale=0.5]{../diagrams/From-DSP-Broker-To-DSPDataPersistence-General-Sequence}
  \caption{UML Sequence Diagram - Flow 1: From the DSP Broker to the Data
  Persistence Component}
  \label{fig:From-DSP-Broker-To-DSPDataPersistence-General-Sequence}
\end{figure}

The DSP Broker retrieves the list of matching rules by contacting the Matcher.
As the DSP Platform have already loaded all the Matching rules, a copy of a
given DSP message is delivered to the DSP Data Persistence component. However,
the sequence diagram shows two different stages of reception:

\begin{itemize}
  \item A bootstrap DSP Update Message is delivered to the DSP Data Persistence
  component when the DSP Platform is initializing the component. In this way,
  the DSP Data Persistence can initialize the DspDataFlusher worker thread to
  flush temporary messages into the database at the rate of
  TRANSIENT\underline{ }DATA\underline{ }FLUSHER\underline{ }DELAY in seconds;
  \item Any DSP Measure Message can be persisted by the DSP Data Persistence
  component. When the DSP Broker delivers a DSP MeasureMessage to this
  component, it will add the message to a temporary memory location to decrease
  I/O on the database, as depicted by the gate "create PersistentMessageUnit
  Gate". The simple procedure to add the message into the transient persistent
  layer is shown on image 3.4.3.
\end{itemize}

The class PersistenceMessageUnit is the major transient persistence unit that
carries references regarding the originating DSP Message that was collected by
the DSP Data Persistence component. It is composed by a SensorLocation and
contains a reference to the PersistentMessageState enumaration. The former
identifies which sensor produced the collected data from the DSP Message by the
IP address, while the latter identifies whether the PersistenceMessageUnit has
been saved into the Database or not as depicted by the UML State diagram in
figure \ref{fig:PersistentMessageState-Diagram}.

\begin{figure}[!b]
  \centering
  \includegraphics[scale=0.5]{../diagrams/PersistentMessageState-Diagram}
  \caption{UML State Diagram for the instance of the class
  PersistentMessageState}
  \label{fig:PersistentMessageState-Diagram}
\end{figure}

\begin{figure}[!b]
  \centering
  \includegraphics[scale=0.5]{../diagrams/From-Create-PersistentMessageUnit-to-TransientPersistence-Layer-Sequence}
  \caption{UML Sequence diagram - Adding a DSP Message into the Transient Persistence layer}
  \label{fig:From-Create-PersistentMessageUnit-to-TransientPersistence-Layer-Sequence}
\end{figure}

According to section 3.4.1, the persistence model must carry information
regarding which sensor device produced the collected data. For this reason, an
instance of the class SensorLocation carries the information about the IP
address from the component that originally produced the DSP Message. Then, an
instance of the class PersistenceMessageUnit is created, having its state on
TRANSIENT, as detailed on UML State diagram on Image 3.4.2.

\subsection{Flushing data into the Database}

The last stage sequence described on the Sequence diagram of Image 3.4.3
depicts the DspDataFlusher worker thread continuously waking up at the rate of
TRANSIENT\underline{ }DATA\underline{ }FLUSHER\underline{ }DELAY. Its purpose is as simple as to iterate over the
list of PersitenceMessageUnit on the state of TRANSIENT, and send each of them
to be saved on the persistence storage. Figure
\ref{fig:DSP-Data-Persistence-Classes} shows the participating classes of the
component.

\begin{figure}[!b]
  \centering
  \includegraphics[scale=0.5]{../diagrams/DSP-Data-Persistence-Classes}
  \caption{UML Class diagram for the DSP Data Persistence component.}
  \label{fig:DSP-Data-Persistence-Classes}
\end{figure}

Finally, this sequence of events is continued by the "insert
PersistenceMessageUnit Gate", which triggers the persistence service from the
chosen database system as shown on Image 3.4.5.

\begin{figure}[!b]
  \centering
  \includegraphics[scale=0.5]{../diagrams/From-Insert-PersistentMessageUnit-to-mongoDB}
  \caption{UML Sequence Diagram - MessageContent instance saved by the class
  mongoDB service}
  \label{fig:From-Insert-PersistentMessageUnit-to-mongoDB}
\end{figure}

In order to build the document key and value segments as described on section
3.2. First, the value is built by extracting the DSP Message Content from the
body of the DSP Message.

\begin{figure}[!b]
  \centering
  \includegraphics[scale=0.5]{../diagrams/From-Creating-Value-Segment-Sequence}
  \caption{UML Sequence diagram showing the creation of the value document segment}
  \label{fig:From-Creating-Value-Segment-Sequence}
\end{figure}

Finally, the document key segment is prepared as follows on image 3.4.7.

\begin{figure}[!b]
  \centering
  \includegraphics[scale=0.5]{../diagrams/From-Creating-Key-Segment-Sequence}
  \caption{UML Sequence diagram showing the creation of the key document segment}
  \label{fig:From-Creating-Key-Segment-Sequence}
\end{figure}

\subsection{DSP Data Persistence Implementation}

This section shows the implementation of the DSP Data Persistence, which
follows the specifications of the DSP Components described on section 2.4,
along with the setup of the mongoDB, the chosen database system that handles
document-oriented instances. The implementation of the DSP Data Persistence
component was developed using the DSP Subversion repository at
http://code.google.com/p/netbeams.

The Deployment diagram with a single host, single database;
Deployment diagram with single host, multiple shards;
Deployment diagram with dedicated database host, single database;
Deployment diagram with dedicated cluster host, multiple shards/replicas;

Each of the deployment schema have two different deployment components: the DSP
component deployment and the database instance installation and configuration.
The following section shows each of the components artifacts and configuration
necessary.

\subsubsection{DSP Deployment on OSGi Framework}

    * Description of the OSGi MANIFEST.MF
    * Description of the component on the config.xml
    * Adding a matching rule on the matcher\underline{ }config.xml
    * Adding a bootstrap message the component and database configuration

    * Adding a new entry into the deployment configuration artifact config.xml;
    * Adding a new entry into the matcher configuration artifact matcher\underline{ }config.xml;
    * Adding an optional configuration message that sets up the new DSP Component;
    * Adding new Java drivers that communicates with the Database system;
    * Installing and configuring the proposed database system.

### 2009-11-15 docs/marcello-dissertation/thesis/chapter7-data-persistence-component-implementation.tex
Updating the previous chapters 4, 5, and adding 6 and 7 unfinished.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@645 559edde2-a4d4-44e0-96fc-4035e40d7396

% main.tex, to be used with thesis.tex
% This contains the main work of your thesis.

%\bibliography{thesis}  % uses the references stored in Chapter1Radar.bib

\chapter{DSP Data Persistence: Component Implementation}

This section shows the implementation of the DSP Data Persistence, which
follows the specifications of the DSP Components described on section 2.4,
along with the setup of the mongoDB, the chosen database system that handles
document-oriented instances. The implementation of the DSP Data Persistence
component was developed using the DSP Subversion repository at
http://code.google.com/p/netbeams.

\section{DSP Deployment on OSGi Framework}

    * Description of the OSGi MANIFEST.MF
    * Description of the component on the config.xml
    * Adding a matching rule on the matcher\underline{ }config.xml
    * Adding a bootstrap message the component and database configuration

    * Adding a new entry into the deployment configuration artifact config.xml;
    * Adding a new entry into the matcher configuration artifact matcher\underline{ }config.xml;
    * Adding an optional configuration message that sets up the new DSP Component;
    * Adding new Java drivers that communicates with the Database system;
    * Installing and configuring the proposed database system.

\subsection{Starting the DSP Platform}

    * command-line (Knopflerfish Framework)
    * user interface (Knopflerfish Framework)

\subsection{Stopping the the DSP Platform}

    * command-line (Knopflerfish Framework)
    * user interface (Knopflerfish Framework)

\subsection{Execution Logs}

\section{mongoDB Deployment}

Given that Document-Oriented Model makes a good candidate persist sensors'
properties and the recently enumerated list technologies in the previous
article DSPDataPersistence, the open-source project called mongoDB was chosen
for the evaluation on our case study, the netBEAMS DSP Platform.

mongoDB supports storage based on collections of data, stored using BSON, a
binary representation the JSON data representation format, including dynamic
queries and indexing support. As it's stated in their web site, mongoDB
"bridges the gap between key/value stores (which are fast and highly scalable)
and traditional RDBMS systems (which are deep in functionality)".

\begin{itemize}
  \item mongoDB implements a document-oriented structure, which is similar to
  KVP;
  \item mongoDB is written in C++, and therefore, can is available in any major
  platform, as well as offers a broad range of API drivers written in different
  languages such as Java, Python, Perl and Ruby; 
  \item mongoDB is open-source, with good community support and availability
  through mailing lists, freenode IRC channel, and commercial support through
  10gen company; 
  \item mongoDB has support to distributed systems properties such as
  Master-Slave replication, and features like Database Shards with
  auto-sharding based on shard keys.
\end{itemize}

nstallation of mongoDB
- Explanation about mongoDB processes: mongod, mongos, mongo
- Specification of the shards, shard key, etc

    * The database instance is called "netbeams";
    * The database "netbeams" may contain different collections, categorized by
    the Sensor Content Type, that is, depending on how the DSP Component was
    described;

\subsection{Document-Oriented Data Model}

An example of the data is as folows (using the JSON syntax). Listing
\ref{file:mongodb-ysi-data-format} follows the strategy described on Section X,
where the entity representation is denormalized, and each of the properties of
the sensor is repeated in each instance of the data. The fact and transaction
times are expressed in milliseconds. The data element contains the complete
structure of the originating sensor such as the IP address. The collection of
instances of this document represents the sensor data type.

\subsection{Starting the Database}

\subsection{Stopping the Database}

\subsection{Execution Logs}

- Java API implementation
- Python API implementation

\section{Data Access Using Database Shell}

The mongoDB client can be started by using the following command. Make sure you
have started the mongoDB server before executing the mongoDB client.

mongo netbeams | tee output\underline{ }number\underline{ }date.log

Here, the iterative mongo client shell offers users to verify and navigate on a
given database and its collections. This first section shows the connection of
the mongo client to the database netbeams. It also highlights the query for the
collections available. During the experiment, the SondeDataContainer?
collection was created as related to the type from the DSP Messages for the YSI
Sonde.

The shell references to the mongoDB system can be found at
http://www.mongodb.org/display/DOCS/dbshell+Reference

### 2009-11-16 docs/marcello-dissertation/thesis/chapter4-netbeams-in-depth-requirements-analysis-and-specification.tex
Major update in the thesis write up, now containing all the needed chapters moved from Google Docs. Updates to chapters 4 to 7 done, adding structure of 8 and 9.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@646 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 \chapter{Netbeams in Depth: Requirements Analysis and Specification}
-
+\label{chap:netbeams-overview}
 This chapter presents NetBEAMS and its implementation details, showing its
@@
 \section{DSP Message and DSP Messages Container}
+\label{sec:dsp-message}
 

### 2009-11-16 docs/marcello-dissertation/thesis/chapter5-taxonomic-persistence-storage-selection-for-netbeams.tex
Major update in the thesis write up, now containing all the needed chapters moved from Google Docs. Updates to chapters 4 to 7 done, adding structure of 8 and 9.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@646 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 \section{Analysis of the Location of Sensor Data Taxonomy}
+\label{sec:sn-data-location}
 
@@
 
+In order to support this strategy, a technique called Database Sharding
+\cite{db-shard02} is a technique used in distributed database systems that
+uses denormalized data models to improve scalability in terms of data storage
+and data location. It is a type of database system patitioning where
+de-normalized data is spread across different other hosts, or shard \cite{db-shard01}.
+Different models of database sharding can be developed \cite{db-shard-schemas}.
+
+\begin{itemize}
+  \item Vertical Partitioning: each feature of the system, usually represented
+  by a table in the relational model, is placed in a different server host. For
+  example, each sensor in a given network can be divided by its device type;
+  \item Range Based Partitioning: after partitioning each feature on a
+  different server host, different multiple shards, or similar hosts, can be
+  added by using the range of a given key of the entity of a function. For
+  example, all sensor types in a geographic location may contain its own shard
+  based on its IP address;
+  \item Key or Hash Based Partitioning: a given value for a given property of
+  an instance of data that is chosen to be used in a hash function, whose
+  output determines which shard to use;
+  \item Directory Based Partitioning: a simple lookup service is used to
+  identify which database shard to use, where the lookup table can be stored
+  in a secondary area. Each approach for database sharding used be used
+  together on a given function depending on the needs for performance.
+\end{itemize}
+
+For example, figure \ref{fig:database-sharding-by-region} is an example of
+data storage with shards \cite{db-shard01} related the location of data, where
+each shard is assumed to be placed in different server host. Since the data
+model is denormalized, the organization of data is entirely placed in each of
+these shards. As it is shown, the shards by location can be defined by a shard
+key, as described before. The horizontal scalability is given related to
+positions, while the vertical by the shards added.
+
+\begin{figure}
+  \centering
+  \includegraphics[scale=0.65]{../diagrams/database-sharding-by-region}
+  \caption{Example of Data-Centric Storage for Sensor Networks using Database
+  Sharding}
+  \label{fig:database-sharding-by-region}
+\end{figure}
+
+%This data organization architecture is currently being used to manage
+%structured data that scales to petabytes or more data. One example of such use
+%is on BigTable \cite{bigtable}, which provides support to different Google
+%products with requirements on high availability and dynamic data models and
+%uses automatic sharding to its data. The BigTable project uses MapReduce
+%[ds02], a programming model designed to search huge amounts of data in
+%paralell.
+
 \subsection{Disadvantages of the Data-Centric Approach}
@@
 sensor networks do not have such expertise.
+
+\begin{itemize}
+  \item Referential integrity: any cross-shard queries can represent problems 
+  since the application layer is responsible to enforce data integrity, like
+  verifying constraints of foreign keys;
+  \item Rebalancing Shards: If the schema model changes for the shards, the
+  strategy of rebalancing must be done. For instance, when one of the shards 
+  outgrows more than the others. As a consequence, the database shards must be
+  rebalanced by transferring data to new locations whenever necessary. This
+  technique is starting to appear as COTS implemetations such as
+  \cite{mongodb}.
+\end{itemize}
 
@@
 Other projects have been using the XML data models. It falls into the same
-category as the Relational Model, since the XML documents, when in the context
-of a database, must be complaint to an XML Schema. As shown in section 3, this
-model uses XPath technology for querying documents, although some hibrid
-technologies may still use SQL for that matter. In this way, an instance of
-such schema dependent types can be seen in figure
-\ref{fig:persistence-example-relational}.
+category as the Relational Model, since the XML documents \cite{xml}, when in
+the context of a database, must be complaint to an XML Schema \cite{xml-schema}. As shown in
+section 3, this model uses XPath \cite{xml-xpath} technology for querying
+documents, although some hibrid technologies may still use SQL \cite{sql} for
+that matter. In this way, an instance of such schema dependent types can be
+seen in figure \ref{fig:persistence-example-relational}.
 
@@
   \centering
-  \includegraphics[scale=0.65]{../diagrams/persistence-example-relational}
+  \includegraphics[scale=0.7]{../diagrams/persistence-example-relational}
   \caption{Instance of a Relational Model Database Prototype}
@@
   \centering
-  \includegraphics[scale=0.5]{../diagrams/KVP-on-Relational-Model}
+  \includegraphics[scale=0.7]{../diagrams/KVP-on-Relational-Model}
   \caption{KVP Data Model implementation using Relational Model}
@@
   \centering
-  \includegraphics[scale=0.5]{../diagrams/persistence-example-relational-kvp}
+  \includegraphics[scale=0.7]{../diagrams/persistence-example-relational-kvp}
   \caption{KVP over Relational Model Database Prototype}
@@
 Sensor Network data can be saved in either Centralized or Distributed Systems.
-While Centralized Database Systems are easier to manage, it may face challenges
-regarding its data. For this reason, the use of partitioning or sharding would
-be an addition to persist sensor data.
+As described in  While Centralized Database Systems are easier to manage, it may
+face challenges regarding its data. For this reason, the use of partitioning or
+sharding would be an addition to persist sensor data.
 

### 2009-11-16 docs/marcello-dissertation/thesis/chapter6-data-persistence-component-design-and-architecture.tex
Major update in the thesis write up, now containing all the needed chapters moved from Google Docs. Updates to chapters 4 to 7 done, adding structure of 8 and 9.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@646 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 In order to provide data to users, the RTC scientists use OPeNDAP \cite{opendap}
-server to distribute collected data over the Internet. However, this work
-proposes the use a KVP database system as described in the previous chapter,
-giving options to advanced techniques to scale on data structure and
-organization.
+server to distribute collected data over the Internet. As discussed in the
+previous chapter, this work proposes the use a Key-Value database system
+in order to provide an alternative way for data persistence in sensor
+networks, taking into account the taxonomies discussed earlier in chapter 
+\ref{chap:taxonomies}.
 
-This section shows the proposed DSP Data Persistence component for the DSP
-Platform as reviewed in chapter 4, as it is responsible for gathering
-the DSP Messages sent from any remote DSP instance. First, section 3.1 briefly
-describes main events, processes, input and outputs generated by analyzing a
-scenario of remote communications between sensors and the DSP Client and
-Server nodes. Then, section 3.2 shows the requirements specification based on
-the analysis of section 3.1 and the documentation of the DSP Platform of
-section 2.4.
+Considering our case study being NetBEAMS, and its infrastructure, this chapter
+describes the design and architecture component designed specifically for the
+Data Sensor Platform, DSP, in order to provide a persistence layer for
+NetBEAMS. It is important to note that the technology chosen in the previous
+chapter can be adapted by any alternative that uses the data model and
+infrastructure options presented.
+
+The proposed DSP Data Persistence component for the DSP Platform was
+implemented by using the same approaches described in chapter
+\ref{chap:netbeams-overview}. The following sections describe the scenarios
+for persistence, analyzes the requirements from RTC, and describes the design
+and architecture of the solution.
 
 \section{Business Process Analysis}
-
-As described in section 2.4, the Data Sensor Platform was designed to support
-an environmental sensor network based in San Francisco Bay, composed by
-different types of sensors. The process of extracting the measurements from the
-sensors to be reused using the DSP Platform can be summarized following the UML
-Business Process diagram depicted in figure
-\ref{fig:fig:dsp-persistence-business-process}. It is important to note that
-this diagrams details remote communication between sensor devices and a
+\label{sec:business-process-analysis}
+As described in section \ref{chap:netbeams-overview}, the DSP Platform was
+designed to support an environmental sensor network based in San Francisco
+Bay, composed by different types of sensors. The process of extracting the
+measurements from the sensors to be reused using the DSP Platform can be
+summarized following the UML Business Process Diagram \cite{uml} depicted in
+figure \ref{fig:fig:dsp-persistence-business-process}. It is important to note
+that this diagrams details remote communication between sensor devices and a
 centralized server, as part of the main scenario to solve the problem of data
@@
 the collected measurements and transforms them into the so-called DSP Message
-(section 2.4), which embodies the actual data values. After proper data
-processing, transformation and categorization, the DSP Message is added into an
-in-memory data storage identified (section 2.4). Then, as an asynchronous data
+(section \ref{sec:dsp-message}), which embodies the actual data values. After
+proper data processing, transformation and categorization, the DSP Message is
+added into temporary outbound queue. Then, as an asynchronous data
 transmission event occurs with an specified rate, transferring all the queued
 messages to be processed by a DSP Server host. As it is detailed in image
-3.1.1, the server is centralized and is able to concurrently process messages.
+\ref{fig:dsp-persistence-business-process}, the server is centralized and is
+able to concurrently process messages.
 
 This work proposes the addition of a database system that is responsible for
-storing the gathered measurement data from sensor hosts on a given network. As
-it is shown in image 3.1.1,  the processed measurement data from sensors 1
-through N must be collected in the same fashion and categorized as necessary.
-In this way, end users such as Biologists and/or Programmers can access the
-collected measurements directly from the Database. Furthermore, the same
-measurement data is supposed to be exported to different formats as the users
-find necessary.
+storing the gathered measurement data from sensor nodes of the SF-BEAMS through
+the use of NetBEAMS. Also in figure \ref{fig:dsp-persistence-business-process}, 
+the processed measurement data from sensors 1 through N must be collected in
+the same fashion and categorized as necessary. In this way, end users such as
+Biologists and/or Programmers can access the collected measurements directly
+from the Database. Furthermore, the same measurement data can be exported to
+different formats as the users find necessary.
 
-As a matter of fact, the measurements data collected from the sensor devices
-are simple  properties with and associated values as described on section 2.1.
-According to the descriptions of the DSP Platform and the DSP Messages
-structure on section 2.4, the data is represented in programming language
-formats such as Java or in a serialized version in XML. Similarly, the
-collected data contains measurements that uniquely identifies them, together
-with other properties such as time stamps of data creation and collected, as
-well as the the IP address of the DSP Client that produced the data. Therefore,
-these are important properties for the proposed solution of this work.
+As a matter of fact, the data measurements collected from the sensor devices
+are simple properties with keys and associated values as described in section
+\ref{sec:sfbeams}, after data has been described. According to the descriptions
+of the DSP Platform and the DSP Messages structure, the data is represented in
+programming language formats such as Java \cite{java} or in a serialized
+version in XML \cite{xml}. Similarly, the collected data contains measurements
+that uniquely identifies them, together with other properties such as time
+stamps of data creation and collected, as well as the the IP address
+\footnote{Internet Protocol Address} of the DSP Client that produced the data.
+Therefore, these are important properties for the proposed solution of this work.
 
@@
 
-By analyzing the main scenario of data process on section 3.1, this section
-lists the main use cases that the system and the primary users require. One of
-the primary requirements of the system is to enable the use a persistence
-system that does not promote the use of specialized professionals to deal with
-data modeling, as it is required in most projects using the Structured Query
-Language (SQL) [db01], used to extract data from a Relational Database [db01].
+By analyzing the main scenario of collected data in the previous section, as
+well as by collected requirements from the RTC staff, this section lists the
+main use cases that the system and the primary users require. One of the
+primary requirements of the system is to enable the use a persistence system
+that does not promote the use of specialized professionals to deal with data
+modeling, as it is required in most projects using the Structured Query
+Language (SQL) \cite{sql}, used to extract data from a Relational Databases.
 
@@
 The functional requirements listed in this section are in the format of UML
-User Cases [se05]. Figure
+User Cases Diagrams \cite{uml}. Figure
 \ref{fig:DSP-Data-Persistence-UseCases-Diagram-System} summarizes the use
@@
   \item The Persistence storage system must not impose the knowledge of
-  complicated database system languages such as SQL [db01] or XML XPath [db05],
-  but use access through API calls, since it is closer to researchers. 
+  complicated database system languages such as SQL \cite{sql} or XML XPath
+  \cite{xml-xpath}, but use access through API calls, since it is closer to
+  researchers.
 \end{itemize}
@@
 \section{Data Model Design}
+\label{sec:dsp-persistence-data-model}
+As described in the previous chapter, the data model that might give better
+support for collected data from sensor networks is the Key-Value-Pair Data
+Model. Beyond the data collected from the sensor devices, Data Provenance 
+techniques must also be considered to better describe the collected data. In
+general, the data model must include the following metadata besed on Data
+Provenance:
 
-After analyzing section 2.1 along with the functional and non-functional
-requirements described on section 3.2.2, this work considered the
-Key-Value-Pair Data model as described on Section 2.5. Taking into account the
-mongoDB architecture and the properties of a DSP Message (see section
-"Acquiring the properties of a DSP Message Content" at DSPDataPersistence),
-here are the conventions followed on revision r585. The following is the list
-of properties that composes the Key of a document:
+\begin{itemize}
+  \item \textbf{Locality}: properties that describe the location of the data.
+  For example, the host machine name or IP address, the GPS coordinate
+  \cite{gps};
+  \item \textbf{Identity}: what uniquely identities the data among the others
+  produced by other sensor devices in the network, as for instance a
+  UUID\footnote{Universally Unique Identifier};
+  \item \textbf{Time-Space}: when the data was ``sensed'' \cite{db-provenance}
+  and when the collection transaction happened \cite{sn-time-series}.
+\end{itemize}
+
+After analyzing the DSP Platform and the structure of the DSP Messages in
+chapter \ref{netbeams-architecture}, the following set of keys are expected to
+be part of the design of the data model:
 
@@
   \item \textbf{transaction\underline{ }time}: it's extracted from the DSP message container
-  creation time and it is used to identify when the transaction occurred (see Section 2.2);
+  creation time and it is used to identify when the transaction occurred;
   \item \textbf{fact\underline{ }time}: it's extracted from the SondeDataContainer's date
-  and time, and identifies the time in which the collected data occurred (see Section 2.2);
-  \item \textbf{latitude}: the latitude variable of the position;
-  \item \textbf{longitude}: the longitude variable of the position.
+  and time, and identifies the time in which the collected data occurred;
+  \item \textbf{latitude}: the latitude coordinate of the position of
+  the sensor device, captured from references of the device (not supported by
+  NetBEAMS);
+  \item \textbf{longitude}: the longitude coordinate of the position of the
+  sensor device, captured from references of the device (not supported by
+  NetBEAMS).
 \end{itemize}
 
-The definition of the Value of a document is as follows:
+It is important to note that this set of metadata must occur in every
+collected data from any sensor device. However, on the contrary of the
+metadata, a different set of properties will occur to describe the sensed
+properties of a given sensor, and for this reason, the following structure is
+defined to accommodate such data:
 
@@
   \item data: this key defines the values of the document, and will have every
-  different property of the sensor (see Section 2.4):
-   \begin{itemize}
-     \item battery
-     \item temperature
-     \item salinity
-     \item \ldots
+  different property of the sensor. For instance, the data key for the
+  properties of the YSI device will contain the set of properties defined for
+  the device, such as battery, temperature, salinity, etc.
 \end{itemize}
-\end{itemize}
-
-Some remarks about the creation of the items contained with the collections. In
-case a message container contains multiple readings in a message container,
-each item will be counted as individual snapshots of the data. Also, it is
-important to note that the description of the data follows no pattern other
-than the description of the properties of the sensor device. However, some
-studies have shown that when disk space maximization is required, the name of
-database columns are important.
 
@@
 
-Considering the business analysis described in section 3.1, the addition of a
-data persistence layer to the DSP Platform can be achieved by adding Data
-Consumer (DC) as described in section 2.4. Based on the Requirements
-specification described in section 3.2, this component must be responsible for
-providing the processing of the DSP Measure Messages into Database-ready
-representation, as well as providing the necessary tools for the transmission
-of the data to a database system. Figure
-\ref{fig:NetBEAMS-Persistence-Server-Node-Components} provides an
-architectural view of the DSP Server node with the loosely-coupled components,
-highlighting the inclusion of the DSP Data Persistence.
+Considering the business analysis described in section
+\ref{sec:business-process-analysis}, the addition of a data persistence layer
+to the DSP Platform can be achieved by adding Data Consumer (DC) as described
+in section \ref{netbeams-architecture}. Based on the Requirements
+specification described earlier, this component must be responsible for
+providing the processing of the DSP Measure Messages, extracting the payload of
+the body of the message and converting it to the data model defined in the
+previous section. Finally, the data can be transmitted to the chosen database
+system. Figure \ref{fig:NetBEAMS-Persistence-Server-Node-Components} provides
+an architectural view of the DSP Server node with the loosely-coupled
+components, highlighting the inclusion of the DSP Data Persistence.
 
@@
 
-As it is depicted on image 3.4.1, the DSP Data Persistence component can be
-added into the system as an independent plug-and-play component. In addition,
-the database system and the necessary drivers are also added without any
-changes to the existing DSP infrastructure. First, as DSP Messages are
-delivered to the server through the DSP Wire Transport Server, as shown by flow
-1, the DSP Platform is responsible to forward a copy of the Measurement
-messages directly to the DSP Data Persistence as shown by flow 2. Then, the
-sole responsibility of the DSP Data Persistence is to transfer the received
-measurement messages to the database system by using the connection drivers, as
-depicted by flow 3.
+As it is depicted in figure
+\ref{fig:NetBEAMS-Persistence-Server-Node-Components}, the DSP Data
+Persistence component can be added into the system as an independent
+plug-and-play component. In addition, the database system and the necessary
+drivers are also added without any changes to the existing DSP infrastructure.
+First, as DSP Messages are delivered to the server through the DSP Wire
+Transport Server, as shown by flow 1, the DSP Platform is responsible to
+forward a copy of the Measurement messages directly to the DSP Data
+Persistence as shown by flow 2. Then, the sole responsibility of the DSP Data
+Persistence is to transfer the received measurement messages to the database
+system by using the connection drivers, as depicted by flow 3.
 
-The following sections focus on the DSP Data Persistence component design, and
-in this way, any references to the DSP Platform can be seen on section 2.4.
+The following sections focus on the DSP Data Persistence component design in
+the context of an OSGi bundle \cite{osgi}.
 
@@
 The new proposed Data Persistence component must follow the DSP requirements
-for a new DSP Component. As described on section 2.4, any DSP component follows
-the design-pattern of Data Producer/Consumer. Similarly, section 3.2 described
-the requirements that the DSP Data Persistence must handle, taking into account
-the data model and database connection.
-
-After reviewing the requirements of section 3.2, the resulting model maintains
-the dependency among 4 different major Java packages. The addition of a DSP
-Data Persistence component depends on 2 existing packages, namely the OSGi
-Framework and the DSP Platform as in any DSP Component, and an additional
-package for the database access driver API. The UML Package dependency diagram
-of image 3.5.x shows these dependencies, which are used to better track changes
-among the packages.
+for a new DSP Component following the the design-pattern of Data
+Producer/Consumer. In a nutshell, the design of a DSP Component resulting
+maintains the dependency among 3 different major Java packages, as the addition
+of such a DSP Data Persistence component depends on 2 existing packages, namely
+the OSGi Framework and the DSP Platform, as well as and an additional package
+for the database access driver API, as shown in the UML Package Dependency
+Diagram \cite{uml} of figure
+\ref{fig:DSP-Data-Persistence-Packages-Dependency}.
 
@@
 
-As shown on section 2.4.3, the DSP Platform is composed of loosely-coupled
-components. In this way, this section describes the addition of the DSP Data
-Persistence, as shown on the UML Class diagram of image 3.5.1.1.
+The DSP Data Persistence component starts with the design of an OSGi Activator
+class, and the implementation of the DSP Component, resposible to provide
+services to the platform. Based on these specifications, the following set of
+classes in figure \ref{fig:DSP-DataPersistence-Activator-Class-Diagram} shows
+the UML Class Diagram \cite{uml} implemented for the component.
 
@@
 Persistence extends from the DSP Platform class DSPComponent. For this reason,
-this class inherits all the behavior described on section 2.4 and can be in the
-roles of Data Consumer and a Data Producer.
+this class inherits all the behavior described on \ref{netbeams-architecture}
+and can be in the roles of Data Consumer and a Data Producer.
 
@@
   measurement message received into a format that must be ready to be saved on the
-  database, based on the description of the data model of section 3.3.
+  database, based on the description of the data model in section
+  \ref{sec:dsp-persistence-data-model}.
 \end{itemize}
 
-The DSP Data Persistence Activator follows the states described on the state
-diagram of image 2.4.3.4, as it is an extension of the OSGI class
-BundleActivator, as well as being responsible for stopping or starting the the
-DSP Data Persistence component instance. As described on section 2.4, the
-following steps must take place in order to activate and start the DSP Data
-Persistence component, described partially on image 2.4.6.3:
+The behavior of the activator class follows the specifications of the
+OSGi Framework, and for this reason, instances of the DSP Data Persistence
+Activator can be in the states described on the UML State Diagram depicted in
+figure \ref{fig:DSPPlatform-Install-Usage-State-Diagram}, as it is an extension
+of the OSGI class BundleActivator, as well as being responsible for stopping
+or starting the the DSP Data Persistence component instance. Following the
+specifications of section \ref{netbeams-architecture}, the DSP Data Persistence
+Platform can be initialized as described by the UML Sequence Diagram of figure
+\ref{fig:From-OSGi-Framework-to-DSP-Data-PersistenceActivator-Sequence-Diagram}:
 
@@
   \item During the DSP Platform activation, it will get the name of the
-  persistence component from the configuration artifact config.xml;
+  persistence component from the configuration artifact config.xml as seen in
+  listing \ref{file:dsp-config.xml};
   \item After being selected based on the configuration priority, the DSP
   Bundle artifact is identified and installed, by creating an instance of the
-  class DSP DataPersistence Activator and making a call to the method start(),
-  as shown on image 3.5.1.2;
+  class DSP DataPersistence Activator and making a call to the method start();
   \item During the activation, an instance of the DSP Data Persistence
@@
 
-As the requirements defined on section 3.2 specifies the update of the DSP
-Component during its initialization, an optional DSP Update Message will be
-sent to the component during its initialization process, as described on
-section 2.4. The initialization of the DSP Component Activator and the DSP
-Component registration are shown on image 3.5.1.2.
+Any configuration parameter required by a DSP Component is sent
+during its bootstrap process through the use of an optional DSP Update Message,
+as described in section \ref{sec:dsp-message}. The initialization of the DSP
+Component Activator and the DSP Component registration are also shown in figure
+\ref{fig:From-OSGi-Framework-to-DSP-Data-PersistenceActivator-Sequence-Diagram}.
 
@@
 Although the DSP Data Persistence component has been instantiated as of image
-3.5.1.2, the process of delivering messages to the DSP Data Persistence has yet
-to be completed. As defined by the requirements, the DSP Component must execute
-its function in a concurrent fashion, using configuration parameters during its
+\ref{fig:From-OSGi-Framework-to-DSP-Data-PersistenceActivator-Sequence-Diagram}, 
+the process of delivering messages to the DSP Data Persistence has yet to be
+completed. As defined by the requirements, the DSP Component must execute its
+function in a concurrent fashion, using configuration parameters during its
 bootstrap process. The class DSP Data Flusher is responsible for running as a
 "worker thread", whose function is to verify if there are messages to be
-flushed into the Database with the class Transient Persistence Layer. Image
-3.5.2.1 describes the remaining UML Class diagram for the main classes of this
-functionality.
+flushed into the Database with the class Transient Persistence Layer. Figure 
+\ref{fig:DSP-DataPersistence-Flusher-Classes} describes the UML Class
+Diagram \cite{uml} for the main classes of this functionality.
 
@@
 The DSP Data Flusher is a thread that is designed to be in two different
-states: RUNNABLE and TIMED\underline{ }WAITING. In the former, the DSP Data Flusher will be
+states as depicted in figure
+\ref{fig:DSP-DataPersistence-Flusher-State-Diagram}: RUNNABLE and
+TIMED\underline{ }WAITING. In the former, the DSP Data Flusher will be
 contacting the Transient Persistence Layer to verify if there are any message
@@
 component must send its reference to the temporary memory space, as described
-by the non-functional requirement on section 3.2. In this way, the DSP Data
-Persistence component maintains a dependency on the Singleton class
-TransientPersistenceLayer, which is responsible for keeping track of messages
-that are received by working component. However, in order to add the DSP
-message into its aggregated map, it first transforms the message into an
-instance of the class PersistentMessageUnit, which has references to the class
-regarding the location of the message,  SensorLocation.
+by the non-functional requirement in the beginning of this chapter. In this
+way, the DSP Data Persistence component maintains a dependency on the
+Singleton class TransientPersistenceLayer, which is responsible for keeping
+track of messages that are received by working component. However, in order to
+add the DSP message into its aggregated map, it first transforms the message
+into an instance of the class PersistentMessageUnit, which has references to
+the class regarding the location of the message, SensorLocation.
 
-The last paragraph of section 2.4.7 described the last steps of the Sequence
-Diagram regarding DSP Message delivery to the DSP Broker on the DSP Server
-host. For this reason, Image 3.4.1.2 continues the steps of the DSP Broker from
-the gate "From DSP WireTransport Server to Broker Gate", showing the steps
-taken to transfer a DSP Message to the Singleton class
-TransientPersistenceLayer.
+As depicted by the the UML Sequence Diagram \cite{uml} of figure
+\ref{fig:From-DSPWireTransport-Server-To-DSP-Broker}, a DSP Message
+is delivery to the DSP Broker for its delivery to Data Consumers configured to
+receive that specific message. For this reason, figre
+\ref{fig:From-DSP-Broker-To-DSPDataPersistence-General-Sequence} continues the
+steps of the DSP Broker from the gate "From DSP WireTransport Server to Broker
+Gate", showing the steps taken to transfer a DSP Message to the Singleton
+class TransientPersistenceLayer, responsible for maintaining a queue of
+messages waiting to be sent to the database system.
 
-According to the DSP Broker model described on section 2.4.6, the DSP Matcher
-needs to be updated with the addition of a matching rule that filters a copy of
-any Measurement Message (section 2.4.4) to the DSP Data Persistence component
+According to the DSP Broker model described in section \ref{sec:dsp-message}, 
+the DSP Matcher needs to be updated with the addition of a matching rule that
+filters a copy of any Measurement Message to the DSP Data Persistence component
 designed. As it is shown in figure
@@
   Gate". The simple procedure to add the message into the transient persistent
-  layer is shown on image 3.4.3.
+  layer.
 \end{itemize}
@@
 IP address, while the latter identifies whether the PersistenceMessageUnit has
-been saved into the Database or not as depicted by the UML State diagram in
-figure \ref{fig:PersistentMessageState-Diagram}.
+been saved into the Database or not as depicted by the UML State Diagram
+\cite{uml} in figure \ref{fig:PersistentMessageState-Diagram}.
 
@@
 
-According to section 3.4.1, the persistence model must carry information
-regarding which sensor device produced the collected data. For this reason, an
-instance of the class SensorLocation carries the information about the IP
-address from the component that originally produced the DSP Message. Then, an
-instance of the class PersistenceMessageUnit is created, having its state on
-TRANSIENT, as detailed on UML State diagram on Image 3.4.2.
+According to the data model specifications defined in section
+\ref{sec:dsp-persistence-data-model}, the persistence model must carry
+information regarding which sensor device produced the collected data. For
+this reason, an instance of the class SensorLocation carries the information
+about the IP address from the component that originally produced the DSP
+Message. Then, an instance of the class PersistenceMessageUnit is created,
+having its state on TRANSIENT, as detailed on UML State Diagram \cite{uml} in
+figure \ref{fig:PersistentMessageState-Diagram}.
 
@@
 
-The last stage sequence described on the Sequence diagram of Image 3.4.3
-depicts the DspDataFlusher worker thread continuously waking up at the rate of
-TRANSIENT\underline{ }DATA\underline{ }FLUSHER\underline{ }DELAY. Its purpose is as simple as to iterate over the
-list of PersitenceMessageUnit on the state of TRANSIENT, and send each of them
-to be saved on the persistence storage. Figure
-\ref{fig:DSP-Data-Persistence-Classes} shows the participating classes of the
-component.
+The last stage sequence described in the UML Sequence diagram of figure
+\ref{fig:From-DSP-Broker-To-DSPDataPersistence-General-Sequence} depicts the
+DspDataFlusher worker thread continuously waking up at the rate of
+TRANSIENT\underline{ }DATA\underline{ }FLUSHER\underline{ }DELAY. Its purpose
+is as simple as to iterate over the list of PersitenceMessageUnit on the state
+of TRANSIENT, and send each of them to be saved on the persistence storage.
+Figure \ref{fig:DSP-Data-Persistence-Classes} shows the participating classes
+of the component.
 
@@
 PersistenceMessageUnit Gate", which triggers the persistence service from the
-chosen database system as shown on Image 3.4.5.
+chosen database system as shown in figure
+\ref{fig:From-Insert-PersistentMessageUnit-to-mongoDB}.
 
@@
 In order to build the document key and value segments as described on section
-3.2. First, the value is built by extracting the DSP Message Content from the
-body of the DSP Message.
+\ref{sec:dsp-persistence-data-model}. First, the value is built by extracting
+the DSP Message Content from the body of the DSP Message.
 
@@
 
-Finally, the document key segment is prepared as follows on image 3.4.7.
+Finally, the document key segment is prepared as follows in figure
+\ref{fig:From-Creating-Key-Segment-Sequence}.
 
@@
 
-\subsection{DSP Data Persistence Implementation}
-
-This section shows the implementation of the DSP Data Persistence, which
-follows the specifications of the DSP Components described on section 2.4,
-along with the setup of the mongoDB, the chosen database system that handles
-document-oriented instances. The implementation of the DSP Data Persistence
-component was developed using the DSP Subversion repository at
-http://code.google.com/p/netbeams.
-
-The Deployment diagram with a single host, single database;
-Deployment diagram with single host, multiple shards;
-Deployment diagram with dedicated database host, single database;
-Deployment diagram with dedicated cluster host, multiple shards/replicas;
-
-Each of the deployment schema have two different deployment components: the DSP
-component deployment and the database instance installation and configuration.
-The following section shows each of the components artifacts and configuration
-necessary.
-
-\subsubsection{DSP Deployment on OSGi Framework}
-
-    * Description of the OSGi MANIFEST.MF
-    * Description of the component on the config.xml
-    * Adding a matching rule on the matcher\underline{ }config.xml
-    * Adding a bootstrap message the component and database configuration
-
-    * Adding a new entry into the deployment configuration artifact config.xml;
-    * Adding a new entry into the matcher configuration artifact matcher\underline{ }config.xml;
-    * Adding an optional configuration message that sets up the new DSP Component;
-    * Adding new Java drivers that communicates with the Database system;
-    * Installing and configuring the proposed database system.
+Next section details the implementation and deployment of the DSP Data
+Persistence component, as well as the database technology chosen in the previous
+section to provide the persistence for the collected data by the DSP Data
+Persistence component.

### 2009-11-16 docs/marcello-dissertation/thesis/chapter7-data-persistence-component-implementation.tex
Major update in the thesis write up, now containing all the needed chapters moved from Google Docs. Updates to chapters 4 to 7 done, adding structure of 8 and 9.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@646 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 http://www.mongodb.org/display/DOCS/dbshell+Reference
+
+\lstset{label=cmd:mongo,caption=Execution of mongo client}
+\begin{lstlisting}
+MongoDB shell version: 1.1.0-
+url: netbeams
+connecting to: netbeams
+type "help" for help
+> show collections
+SondeDataContainer
+system.indexes
+>
+> db.SondeDataContainer.count()
+1000000
+\end{lstlisting}
+
+Then, the first verification of the data integrity is regarding the number of
+elements created. Here, the first count() function on the collection returned
+1000000.
+
+An example about retrieving the first element of the collection can be done
+using the findOne() function. It will return an element instance on the
+JSONnotation.
+
+\lstset{label=cmd:mongo-findone,caption=Querying the database: one item}
+\begin{lstlisting}
+> db.SondeDataContainer.findOne()
+{"_id" :  ObjectId( "d36f4007b7e7ac4a03c60000")  , "sensor_ip_address" : "192.168.0.136" , "message_id" : "7b6624d6-0ca1-4cba-a343-f166e88da73b" ,
+"transaction_time" : 1252845473412 , "fact_time" : 1252845346000 , "data" : {"temperature" : "45.01" , "sp_condition" : "37.6"
+, "condition" : "145.8" , "resistence" : "159.77" , "salinitude" : "0.0" , "pressure" : "0.391" , "depth" : "0.46" , "ph" : "5.64" ,
+"pH_mv" : "-62.1" , "odo_sat" : "89.7" , "odo_condition" : "59.34" , "turbidity" : "0.0" , "battery" : "9.4"}}
+\end{lstlisting}
+
+The query based on attributes can be done using the "dot" notation, as you
+navigate through the JSON documents. Additionally, you can use the functions as
+aggregated on the result of others. The example in listing
+\ref{cmd:mongo-find-list} counts the number of documents with the key
+"data.ph" equals to "5.64".
+
+\lstset{label=cmd:mongo-find-list,caption=Execution of mongo client}
+\begin{lstlisting}
+> db.SondeDataContainer.find({"data.ph":5.64)}).count()
+1226
+\end{lstlisting}
+
+The following example is the output of the first 2 documents from the same
+previous query using the limit() function, as shown in figure
+\ref{cmd:mongo-find-limit}.
+
+\lstset{label=cmd:mongo-find-limit,caption=Query Element with specific
+projection limiting the result set size}
+\begin{lstlisting}
+> db.SondeDataContainer.find({"data.ph":5.64}).limit(2)
+{"_id" :  ObjectId( "d36f4007b7e7ac4a03c60000")  , "sensor_ip_address" : "192.168.0.136" , "message_id" : "7b6624d6-0ca1-4cba-a343-f166e88da73b"
+, "transaction_time" : 1252845473412 , "fact_time" : 1252845346000 , "data" : {"temperature" : "45.01" , "sp_condition" : "37.6" ,
+"condition" : "145.8" , "resistence" : "159.77" , "salinitude" : "0.0" , "pressure" : "0.391" , "depth" : "0.46" , "ph" : "5.64" ,
+"pH_mv" : "-62.1" , "odo_sat" : "89.7" , "odo_condition" : "59.34" , "turbidity" : "0.0" , "battery" : "9.4"}}
+{"_id" :  ObjectId( "d36f4007b7e7ac4a1fc80000")  , "sensor_ip_address" : "192.168.0.136" , "message_id" : "7b6624d6-0ca1-4cba-a343-f166e88da73b" ,
+"transaction_time" : 1252845473412 , "fact_time" : 1252845346000 , "data" : {"temperature" : "46.71" , "sp_condition" : "60.8" ,
+"condition" : "160.6" , "resistence" : "1399.4" , "salinitude" : "0.01" , "pressure" : "1.057" , "depth" : "2.485" , "ph" : "5.64" ,
+"pH_mv" : "-16.3" , "odo_sat" : "58.8" , "odo_condition" : "19.29" , "turbidity" : "0.2" , "battery" : "9.2"}}
+>
+\end{lstlisting}
+
+Other logs are located at
+http://code.google.com/p/netbeams/downloads/list
+
+\section{Data Access Using APIs}
+
+The mongoDB server offers different drivers to access the data, as well as the
+Web Services.
+
+\begin{itemize}
+  \item The Java tutorial on the drivers is located at
+    http://www.mongodb.org/display/DOCS/Java+Tutorial. The driver is located on
+    the NETBEAMS/versions/v2/thirdparty/mongodb directory;  
+  \item The REST API can be called from any HTTP client. At the time of the
+  editing, this feature is still under alpha version. Check the documentation
+  athttp://www.mongodb.org/display/DOCS/Http+Interface for details 
+\end{itemize}
+
+The following HTTP GET Request method returns the first 5 documents in the collection:
+http://127.0.0.1:28017/netbeams/SondeDataContainer/?limit=-5, as shown in
+listing \ref{cmd:mongo-rest-request}.
+
+\lstset{label=cmd:mongo-rest-request,caption=REST HTTP GET Request Example}
+\begin{lstlisting}
+HTTP/1.0 200 OK
+x-action:
+x-ns: netbeams.SondeDataContainer
+Content-Type: text/plain;charset=utf-8
+
+{
+  "offset" : 0,
+  "rows": [
+    { "_id" : "156f4007e4c3b74a36ed3100", "sensor_ip_address" : "192.168.0.117", "message_id" : "08b02c08-9290-4517-9a28-c6ee7e16509a",
+"transaction_time" : 1253557219486, "fact_time" : 1253557217000, "data" : { "temperature" : "31.44", "sp_condition" : "99.8", "condition" : "53.5",
+"resistence" : "1157.08", "salinity" : "0.0", "pressure" : "1.066", "depth" : "0.161", "ph" : "1.08", "pH_mv" : "-82.0", "odo_sat" : "40.3",
+"odo_condition" : "56.85", "turbidity" : "0.2", "battery" : "8.2" } } ,
+    { "_id" : "156f4007e5c3b74a37ed3100", "sensor_ip_address" : "192.168.0.117", "message_id" : "08b02c08-9290-4517-9a28-c6ee7e16509a",
+"transaction_time" : 1253557219486, "fact_time" : 1253557217000, "data" : { "temperature" : "37.83", "sp_condition" : "176.3", "condition" : "2.6",
+"resistence" : "1324.97", "salinity" : "0.01", "pressure" : "1.36", "depth" : "1.564", "ph" : "0.12", "pH_mv" : "-23.5", "odo_sat" : "104.5",
+"odo_condition" : "19.44", "turbidity" : "0.1", "battery" : "5.0" } } ,
+    { "_id" : "156f4007e5c3b74a38ed3100", "sensor_ip_address" : "192.168.0.117", "message_id" : "08b02c08-9290-4517-9a28-c6ee7e16509a",
+"transaction_time" : 1253557219486, "fact_time" : 1253557217000, "data" : { "temperature" : "74.3", "sp_condition" : "84.0", "condition" : "104.7",
+"resistence" : "4089.13", "salinity" : "0.01", "pressure" : "1.222", "depth" : "2.788", "ph" : "6.56", "pH_mv" : "-78.1", "odo_sat" : "40.0",
+"odo_condition" : "6.02", "turbidity" : "0.3", "battery" : "3.2" } },
+  "total_rows" : 3 ,
+  "query" : {} ,
+  "millis" : 0
+}
+\end{lstlisting}
+
+Data visualisation tools for mongoDB is slowly being developed by open-source
+developers. The next picture shows the database "netbeams" and the collection
+"SondeDataContainer  ?" being rendered by futon4mongodb, one of the
+open-source tools developed to visualise mongoDB data.
+
+\section{Visualizing Data on The Browser}
+
+Futon4Mongo is an open-source software adapted from CouchDB \ref{couchdb}.
+First, the collections list is shown. Figure
+\ref{fig:view-collections-instance-browser-futondb} shows one single collection
+for the YSI data containing 1 million objects.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=0.3]{../diagrams/view-collections-instance-browser-futondb}
+  \caption{Viewing a partial list of data using the Futon for CouchDB/MongoDB}
+  \label{fig:view-collections-instance-browser-futondb}
+\end{figure}
+
+The collection is composed by instances of documents of representing the YSI
+Sonde type, being indexed by a document ID and the values being the keys
+defined in the previous chapter, as shown in figure
+\ref{fig:view-collected-data-list-browser-futondb}.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=0.3]{../diagrams/view-collected-data-list-browser-futondb}
+  \caption{Viewing a partial list of data using the Futon for CouchDB/MongoDB}
+  \label{fig:view-collected-data-list-browser-futondb}
+\end{figure}
+
+By clicking in one of the items, one can view the instance of the document, as
+it is shown in figure \ref{fig:view-collections-instance-browser-futondb}.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=0.3]{../diagrams/view-collected-data-instance-browser-futondb}
+  \caption{Viewing an instance of collected data using the Futon for
+  CouchDB/MongoDB}
+  \label{fig:view-collected-data-instance-browser-futondb}
+\end{figure}
+
+\section{Exporting Data to Spreadsheets}
+
+mongoDB has an export facility shell called mongoexport. It can export the data
+in JSON format or CSV. One may also write its own export tool in any of the
+languages such as Java, PHP, Python, Perl, Ruby, among others. A list of the
+existing drivers in different languages is provided
+athttp://www.mongodb.org/display/DOCS/Drivers. The following command can be
+executed to have the exported version of the data in CSV (read the help output
+of the command for details).
+
+An example to export the data in CSV format can be seen in listing
+\ref{cmd:mongoexport}, and an example containing 1 million objects can be
+downloaded at
+http://netbeams.googlecode.com/files/experiment-1000000-data-exported-20090913-053538.csv.tar.gz.
+
+\lstset{label=cmd:mongoexport,caption=Command to export data in CSV format}
+\begin{lstlisting}
+mongoexport -d netbeams -c SondeDataContainer --dbpath ./data/ --csv -f "_id,sensor_ip_address,transaction_time,fact_time,
+data.temperature,data.sp_condition,data.condition,data.resistence,data.salinitude,data.pressure,data.depth,data.ph,data.pH_mv,data.odo_sat,
+data.odo_condition,data.turbidity,data.battery" -o sonde-data-exported.csv
+\end{lstlisting}
+
+\section{Exporting Data to OPeNDAP Format}
+
+By using any programming language driver, one can export data in the OPeNDAP
+format.

### 2009-11-16 docs/marcello-dissertation/thesis/chapter8-experimental-results-correct-behavior-and-performance.tex
Major update in the thesis write up, now containing all the needed chapters moved from Google Docs. Updates to chapters 4 to 7 done, adding structure of 8 and 9.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@646 559edde2-a4d4-44e0-96fc-4035e40d7396

--- 2009-11-14 docs/marcello-dissertation/thesis/chapter2-state-of-the-art.tex
@@
 
-\chapter{State of the Art}
+\chapter{Experimental Results: Correct Behavior and Performance}
 
-Different aspects related to two major fields of Computer Science have
-motivated to the inception of this work. In the perspective of Database
-Systems and Distributed Systems, this work seeks a different approach for a
-persisting collected data from sensor networks into a storage system that can
-be seen as a cluster of distributed databases. Likewise, in the perspective of
-Software Engineering, this work adds a major capability to an existing sensor 
-network framework, meeting the requirements and interfaces. In general, this 
-chapter gives the foundation based on the motivations for the inception of a 
-different approach to persisting data using a different data model, covering 
-topics that will be referred throughout the following other chapters.
-
-This chapter will first give an overview on different related problems
-described  in the literature that solved problems related to data modeling and 
-representation in sensor networks. Then, it presents the design, architecture 
-and implementation details of the DSP Platform, which is the bases for this 
-dissertation. Finally, this chapter will briefly give a tutorial on advanced 
-topics such as the differences between the KVP data model and the relational 
-data model, as well as discussing about the approaches used on Database
-Sharding.
-
-\section{Sensor Networks Survey}
+\section{Design Model For Sensor Networks}
 
@@
 
-\subsection{Deployment and Mobility}
+\subsection{Strengths as compared to literature}
 
-Each of the nodes can be deployed in a random way such as being dropped
-from an airplane, or in a manual fashion at deliberately chosen geographical 
-points \cite{sn-intro01}. Moreover, the deployment can be done as a one-time
-event or an iterative way.
+\subsection{Weaknesses as compared to literature}
 
-However, nodes can be able to move around the deployed area or be static in
-this place. In the case of the degree of the former behaviour, a sensor node
-can move occasionally or continuously, being a result of an incidental 
-side-effect or a desired property of the device, where they are an automotive
-device or attached to another moving object. For this reason, \cite{sn-intro01} 
-describes those types of mobility as active or passive.
+\section{Data-Centric Storage and Database Shards}
 
-\subsection{Device Size, Resources, Cost, Energy, Heterogeneity}
+Horizontal scalability for data based on specific keys.
 
-Such sensor nodes can be of different sizes, ranging from microscopically small
-particles, small devices to bigger ones with a few feet cubic. In addition to
-being limited in size, they are also limited in resources and computational
-power, forcing their conception limited to have its energy stored in either
-batteries or scavenged from solar cells \cite{sn-intro01}.
+\subsection{Scalability facing High Volume of Collected Data}
 
-Since sensor devices range on size and capabilities, the cost for each of the
-devices very from hundreds to thousands of dollars, depending on its
-functionality, energy utilization, etc. For thie reason, \cite{sn-intro01}
-explains that a sensor network may potentially contain similar sensor devices,
-as they can contain dissimilar ones from each other, which characterizes the heterogeneity 
-of a sensor network by the hardware capabilities such as GPS enabled or having
-GSM  network connectivity.
+Adding new shards as new hardware into the cluster, facing increasing data load
+produced by the sensor network.
 
-\subsection{Communication Modality, Network Infrastructure, Coverage and
-Connectivity}
-\label{sec:sn-infrastructure}
+\subsection{Weakness}
 
-Sensor nodes can communicate in different modalities such as radio, diffuse
-light, inductive or sound, which may influence on the type of protocols used to
-communicate among the nodes, where they can be said to be in an organized
-infrastructure or as an ad hoc way \cite{sn-intro01}.
-
-When it comes to the sensor network different organizational infrastructure,
-they can differ in types of topology: single-hop, star, networked stars, tree
-and graph, where each of them may affect the characteristics of the network as it 
-may also influence on data gathering and routing mechanisms. While in a
-single-hop sensor network there are only two levels of sensor connectivity, the
-tree, graph or networked stars networks uses communications among the network
-hosts for data communication.
-
-Last, but not least, sensor networks may be influenced its coverage designed in
-its purpose, its size in terms of number of nodes and its relating lifetime. As
-sensors are deployed in different areas, they can be distributed in a sparse or
-dense way, yet providing necessary coverage for the environment. However, 
-considering that sensor devices are very likely to constant failures, projects 
-have provided redundant nodes \cite{sn-intro01}. This degree of coverage also 
-influences on the connectivity of the sensor nodes, where they can be
-considered connected when a connection is available at all times. On the other
-hand, nodes can communicate with other nodes in an intermittent way or sporadic
-way. In the former, the sensor node is partitioned, while in the latter the
-nodes are isolated most of the time and communicates at certain moments in
-time.
-
-\subsection{Network Size, Lifetime and Quality-Of-Service}
-
-Similar to the degree of coverage of the sensor network is the number of
-sensor nodes deployed in the network, where they can be in the order of
-dozens, hundreds or even thousand of nodes \cite{sn-intro01}.
-
-The application of the sensor network determines how long the sensor
-network may be exist. For example, the lifetime of an environmental sensor
-networks \cite{sn-ex01} may be not determined since its service serves in a
-regular basis, whereas the lifetime of a sensor network to determine the number
-of people in an event is limited to the event.
-
-Finally, the design of sensor networks can be related to the constraints
-related to quality-of-service (QoS) such as real-time delivery within certain
-period of time or robustness such as maintaining data alive even though there
-a communication link is not available \cite{sn-intro02}.
-
-\section{Persistence Storage for Sensor Networks}
-
-The most important purpose of a sensor network is related to its main product:
-the use of the raw data. Data collection in sensor networks is directly
-dependent on the specifications of the sensor network, its routing mechanisms
-and the final destination of the produced data. In a nutshell,
-\cite{sn-storage03} describes data as follows:
-
-\begin{itemize}
-  \item \textbf{low-level property} as the raw data collected from the
-  sensor device; For example, the sea level as a number;
-  \item \textbf{Observation or Event} as the processed data in format of the final
-  observation. For instance, if the ocean is in its high tight or low tight.
-\end{itemize}
-
-\subsection{The Purpose of the Collected Data}
-\label{sec:sn-data-purpose}
-
-At first glance, the raw data serves as the primary product of the sensor
-network. However, the first consideration that must take place is regarding the
-purpose of the collected data, as \cite{sn-provenance} explains
-differentiates between two main use of the sensor data:
-
-\begin{itemize}
-  \item \textbf{Real-time Data Stream}, as the generated data from the sensor is
-  actually not transmitted, but extracted from the device as it is requested.
-  The data is still in its natural format specified by the sensor manufacturer;
-  \item \textbf{Archival Data}, as the collected data have historical value or will
-  be use by other applications at a future time. 
-\end{itemize}
-
-While real-time data stream are used and can potentially be discarded, the
-archival data needs to be stored in a secondary memory. Nonetheless, given the
-nature of the sensor device specification, the low-level data produced
-by a given sensor device might not contain necessary information regarding the
-identity of the data itself. The use of Data Provenance techniques is described
-by \cite{sn-provenance} as an important step to provide the collected data 
-specific metadata regarding its location and lineage.
-
-Under these circumstances, the collected data needs first to face its jorney
-starting from its transmission from its producing device to the specified data 
-sink as specified by \cite{sn-storage01} \cite{sn-storage02} \cite{sn-storage03}.
-
-\subsection{Storage Location for Collected Data}
-\label{sec:sn-storage-locations}
-
-Given that collected data can be used for different purposes as described in
-section \ref{sec:sn-data-purpose}, the sensor node that produces the data is
-responsible to send the raw data to a specified data storage destination. Upon
-producing data, a sensor network node is specified to deal with the produced
-data, where in most cases, \cite{sn-storage03} described that depending on the
-sensor network infrastructure (see section \ref{sec:sn-infrastructure}),
-different storage mechanisms are provided:
-
-\begin{itemize}
-  \item \textbf{External Storage}: the sensor device sends the produced data to
-  another sensor node that contains an external storage. 
-  \item \textbf{Local Storage}: as it is the case of real-time data stream, where data
-  is allocated on the sensor device's main read-only memory.
-\end{itemize}
-
-It is clear that the data flows from the sensor node to another depending on
-the specification of the network. On the whole, the sensor network
-infrastructure dictates where the data will flow to, as described in section
-\ref{sec:sn-infrastructure}. \cite{sn-storage02} sees a wireless sensor
-networks as a many-to-one data gathering pattern, when its organization is
-based on single-hop or star. On the other hand, when the data organization is
-structured as a graph, tree, etc, the data flows from one intermediate sensor
-node to another until it reaches its final storage node as described by
-\cite{sn-storage01} \cite{sn-storage03}.
-
-Last, but not least, yet a different storage strategy can be used: Data-Centric
-Storage. This strategy aims at clustering the collected data based on a given nature of the
-data such as location. Regardless the data storage mechanism used, the query
-processing will depend on the sensor network infrastructure and data location.
-
-\subsection{Query Processing for Storage Locations}
-\label{sec:query-process}
-
-Given the different locations where the collected data might be as described in
-section \ref{sec:sn-storage-locations}, different query processing for the
-collected sensor data can be used as described by \cite{sn-storage03}:
-
-\begin{itemize}
-  \item \textbf{In-Network Query Processing}: sensor nodes that provides
-  real-time data stream usually replies to requests to properties. Likewise, when the
-  data is used for archival purpose, the data can be located at a network node
-  specialized in storing the neighbors collected data. In this way, when a
-  query is issued to the sensor network, the query processor will
-  collectively verify every storage node for the collected data.
-  \item \textbf{Centralized Query Processing}: a main network sink receives the
-  collected data from the peer nodes. In this case, the query processor queries
-  the solo centralized node for the overall values of the participating sensor
-  devices.
-\end{itemize}
-
-There are pros and cons related to both of the approaches of query processing,
-which are also related to the infrastructure, sensor network nature and 
-requirements. The main goal of the in-network query processing is to decrease
-the amount of data transmitted from the sensor devices to the final data sink, as
-sensor devices are usually limited by constraints such as energy utilization 
-\cite{sn-storage03}. Along with this strategy, the literature describes
-different proposals that aims at upgrading the Structured Query Language (SQL)
-adding location-sensitive clauses to the language \cite{sn-db-newop}. However,
-this main problem related to this approach is that it may potentially flood
-all nodes with query messages \cite{sn-storage04}.
-
-In order to mitigate the problems related flooded nodes in in-network query
-processing, the use of a Data-Centric storage strategy has been shown to be a
-proven alternative for the sake of energy consumption \cite{sn-storage03}
-\cite{sn-storage01}. Its main goal is to classify the collected data based in a
-specific property, say location or event, in order to be sent to specific
-storage nodes that relates to the observation. As the main outcome of this
-strategy is to create network partitioning strategy, the query processor would
-hit specific sensor nodes with the needed data, and therefore, would
-decrease the load and maximize the capacity on the actual sensor devices.
-
-A centralized network query processing processes data in a network node
-specified as the network sink. The network sink usually is a network
-node with a database server with the collection of data delivered by
-all the sensor devices from the network. As a consequence of this
-many-to-one communication strategy \cite{sn-storage02}, the query process does
-not affect any of the sensor nodes in the network and, thus, maximizes the
-sensor device utilization. Moreover, the use of the data for historical
-purposes is easier to manage and give access to the primary end users of the
-data, following the process of data processing, compression, aggregation, etc
-\cite{sn-db-modeling02}.
-
-Although this approach seems to be convenient to the point of view of the
-sensor nodes and used as the tradeoff between achieving capacity and energy
-consumption, the centralized data collector suffers from the annavoiable
-creation of a point of traffic concentration \cite{sn-storage02}, also referred
-to as the phenomenon of congestion, or funneling effect \cite{sn-storage04}.
-For this reason, simple distributed system techniques such as the use of
-master-slave or data replication can be used to decrease the overhead.
-
-\section{Data Models and Query Engines for Sensor Networks}
-\label{sec:data-models}
-
-Different approaches are used to represent the collected data from sensors. The
-use of simple data descriptors have been used in research groups which
-expertise is not database systems, while others use such technology and the way
-to represent data using more robust approaches. \cite{sn-data-model-survey}
-have shown different data models and the advantages and desadvantages. The
-following is a summary of these approaches used by the sensor network
-community.
-
-\begin{itemize}
-  \item \textbf{Tabular Data Model}: comma-separeted values are very popular
-  due to its simplicity. Although data is not index, it is an easy approach for data
- exchange among researchers.
-  \item \textbf{Relational Data Model}: it is used in the majority of the
-  projects. Most of them due to the easy access to database systems, with
-  refactored versions that support the implementation of sensor networks and
-  its requirements \cite{sn-db-tinydb};
-  \item \textbf{Structured Data Model}: simiar to the relational model,
-  structured data models are becoming popular due to the Internet. The use of
-  XML data \cite{xml} is not yet popular, but there are a few projects using it
-  \cite{sn-xml-usage01}\cite{sn-xml-usage01}, implementing its query by using a
-  middleware \cite{sn-xml-middleware} \cite{sn-xml-query-engines}.
-  \item \textbf{Key-Value Pair and Document-Oriented Data Model}: as
-  an emerging data model, the Key-Value Pair (KVP) \cite{db-kvp} data model is
-  becoming popular with the up-coming trend of cloud computing. Although there are no
-  records of the use of this model data model in sensor networks, this work
-  will investigate it.
-\end{itemize}
-
-Most importantly, the use of Data Provenance approaches to describe the
-collected data is shown to be important \cite{sn-provenance}, given the nature
-of the collected data. Therefore, this section describes these approaches for
-modeling data collected from sensor devices, taking into account the need of
-metadata for the data.
-
-\subsection{Tabular Data Model and Query}
-
-Simple use of Comma-Separated Values is an example os sensor network data. Data
-is structured by the top header, and the values are in each line of the file.
-The implementation of the SF-BEAMS \cite{sfbeams2006} uses this approach. In
-summary, in a tabular data model
-
-\begin{itemize}
-  \item data is represented by columns in a flat file separated by a character
-  that separates the values. One example of such data in sensor networks is the
-  OPenDAP \cite{opendap} format, used by the SF-BEAMS project
-  \cite{sfbeams2006}.
-  \item data is queried using a text editor file, without any indexing
-  associated with it. 
-\end{itemize}
-
-\subsection{Relational Data Model and Query}
-
-Many different projects have used the relational model \cite{relational-model}
-as the representation of data for sensor networks. In general, the following
-properties apply to sensor networks:
-
-\begin{itemize}
-  \item data is structured in tables, and sensor properties are represented 
- by the columns of these tables. Data instances are represented by tuples or
- rows of these tables. Finally, table relationships are implemented based
- on the relational algebra abstraction, which is the mechanism used to 
-describe the relationship among different data types.
-  \item data is queried using the Structured Query Languange (SQL), and uses
-  the definitions of the tables and columns to extract data.
-\end{itemize}
-
-Before the conception of such model, a database normalization process is
-necessary to describe the data in an appropriate way. An example of such use
-of the relational model is the TinyDB database systems \cite{sn-db-tinydb},
-which uses the relational model to represent the types of sensors. Due to the
-popularity of relational models, this is a popular choice since of the vast
-availability of research on relational database models.
-
-Although the implementation of a relational database seems to cover the needs
-of data representation, \cite{sn-data-model-survey} described the problems
-related to this approach. One such problem is that the relational model is not
-suited for modeling time series data and metadata related to identify, time and
-location. For the problems related to it, came the implementation of different
-approaches that adds semantics to data types and different query clauses to the
-Structured Query Language (SQL) to solve join \cite{sn-db-newop}.
-
-\subsection{Structured Data Model and Query}
-
-The use of XML is used to represent data in sensor networks has become popular
-due the popularity of this data format. In general, the use of XML for
-persistence can be summarized as follows:
-
-\begin{itemize}
-  \item data is structured in the collected data properties can be represented through the
-XML \cite{xml} instances by using tags, having its model validated by the use
-of XML Schema \cite{xml-schema}.
-  \item in order to extract data from XML-based databases, the use of XML XPath
-  is used to query XML documents.
-\end{itemize}
-
-Some examples of the application of XML in sensor networks in 
-\cite{sn-xml-usage02} some have investigated the effecient use of XML in
-sensor networks to represent the collected data \cite{sn-xml-usage01}. An
-example of querying xml data in sensor networks described by
-\cite{sn-xml-query-engines}, while others use XML as a means of data exchange
-by implementing of middlewares to query and extract query data
-\cite{sn-xml-middleware}. NetBEAMS is one of such sensor network that contains
-an XML-based communication middleware for data exchange among the sensor nodes
-\cite{netbeams2009}.
-
-\subsection{Key-Value Pair and Document-Oriented Data Models}
-
-The advance of distributed systems and the Internet have enabled the
-development of powerful database systems that can be organized in the context
-of database systems and how to represent models. An important up-coming
-approach is the use of database systems that does not use structured language
-for quering its data. One such example is called Key-Value Pair (KVP) data model
-\cite{db-kvp}. Such model has the following properties:
-
-\begin{itemize}
-  \item data is structured in collections of key-value pairs, as it is done in
-  hash data structure. The definition of the key is a given property with its
-  associating value.
-  \item the query process is using mechanics similar to programming or
-  scripting language, that is, the use of APIs in a given programming language.
-\end{itemize}
-
-Different variants of such data model is the document-oriented data model,
-where data is modelled using structured documents. However, these documents
-have a dynamic structure that can be freely described without the use of an
-schema validation mechanism as used in XML Schema. One example of such
-approach is the use of the JSON data format \cite{json} in the implementation
-of database systems which are becoming popular with the new trend called Cloud
-Computing \cite{cloud-comp-architectures}. Databases implementing this strategy
-is the mongoDB \cite{mongodb} and CouchDB \cite{mongodb}.
-
-\subsection{Data Provenance for Sensor Networks}
-
-Most of sensor networks implementations use the collected data from the sensor
-devices after they observe the environment. As shown by
-\cite{sn-data-model-survey}, different data models contains its pros and
-cons to represent data. However, one important aspect in sensor networks is the
-description of it collected data, as \cite{sn-provenance} discussed the
-properties and nature of sensor networks generated data. First, the device's
-sampling are usually in raw format or without obvious names. In order to
-provide search over the collected sensor data, the data needs first to be
-described and indexed, and this process can be related to Data Provenance
-\cite{db-provenance}.
-
-An example of Data Provenance application is the use of aggregation of data
-over time to estimate the changing effects of the weather in a given region.
-For this reason, the collected data needs descriptions or annotations that can
-give enough information to track changes, that is, the history of how and when
-a given data has changed. The following should be the foundation of a
-provenance-aware collected data according to \cite{sn-provenance}:
-
-\begin{itemize}
-  \item each low-level observation as it describes the actual data;
-  \item the location where the collected data was generated;
-  \item the time when the data was collected.
-\end{itemize}
-
-While Data Provenance provides necessary mechanisms to describe sensor networks
-data in general, others see the importance of the time space of sensor
-networks. \cite{sn-time-series} describes strategies to used to cluster
-distributed sensor data as time series of data.
+Technological, and skills-based.

### 2009-11-16 docs/marcello-dissertation/thesis/chapter9-conclusions-and-future-works.tex
Major update in the thesis write up, now containing all the needed chapters moved from Google Docs. Updates to chapters 4 to 7 done, adding structure of 8 and 9.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@646 559edde2-a4d4-44e0-96fc-4035e40d7396

% main.tex, to be used with thesis.tex
% This contains the main work of your thesis.

%\bibliography{thesis}  % uses the references stored in Chapter1Radar.bib

\chapter{Conclusions and Future Works}

This work presented a solution to store data from sensor networks.....
...the main result of this work was that the persistence is

There are many important features covered in this work, including...
  - Use of open-source UI for KVP Sensor Data

Other important features that COULD HAVE BEEN covered, includes\ldots
  - Different Dababase system, use of cloud computing, etc
  - Use of MapReduce technology and implementations, distributed Processing that scales the proposed KVP storage

... The main results from this dissertation were important because of the contributions on new technologies...
  - An alternative Data Model for Data Storage in Sensor Networks
  - An alternative apprach for Data-Centric Storage in Sensor Networks

\section{TODO list}

### 2009-11-20 docs/marcello-dissertation/thesis/chapter2-state-of-the-art.tex
Revision of the chapters 2-6 regarding Data Provenance in sensor networks:
* Chapter 2
  - Adding section regarding provenance

* Chapter 3
  - Splitting section 1 into 2 different ones; TODO: Draw another diagram. 
  - Adding provenance information to the taxonomy of data model
  
* Chapter 4
  - Adding references regarding to the missing location in the NetBEAMS data;

* Chapter 5
  - References to the provenance sections to the keys definition;
  
* Chapter 6
  - References to the provenance sections to the keys definition;

git-svn-id: svn://localhost/netbeams/branches/marcello/persistence

@@
   due to its simplicity. Although data is not index, it is an easy approach for data
- exchange among researchers.
+ exchange among researchers \cite{sn-provenance}.
   \item \textbf{Relational Data Model}: it is used in the majority of the
@@
   middleware \cite{sn-xml-middleware} \cite{sn-xml-query-engines}.
-  \item \textbf{Key-Value Pair and Document-Oriented Data Model}: as
-  an emerging data model, the Key-Value Pair (KVP) \cite{db-kvp} data model is
-  becoming popular with the up-coming trend of cloud computing. Although there are no
-  records of the use of this model data model in sensor networks, this work
-  will investigate it.
 \end{itemize}
 
-Most importantly, the use of Data Provenance approaches to describe the
-collected data is shown to be important \cite{sn-provenance}, given the nature
-of the collected data. Therefore, this section describes these approaches for
-modeling data collected from sensor devices, taking into account the need of
-metadata for the data.
+Although these approaches have been used in Sensor Networks, this work will
+also take a look at others still not used, as follows:
+
+\begin{itemize}
+  \item \textbf{Key-Value Pair Model}: as an emerging data model, the Key-Value
+  Pair (KVP) \cite{db-kvp} data model is becoming popular with the up-coming
+  trend of cloud computing. Although there are no records of the use of this
+  model data model in sensor networks, this work will investigate it.
+\end{itemize}
 
@@
 
-\subsection{Key-Value Pair and Document-Oriented Data Models}
+\subsection{Key-Value Pair Data Models}
 
@@
 for quering its data. One such example is called Key-Value Pair (KVP) data model
-\cite{db-kvp}. Such model has the following properties:
+\cite{db-kvp}, which is also known as Document-Oriented, Distributed Hash Table.
+Such model has the following properties:
 
@@
 
-\subsection{Data Provenance for Sensor Networks}
+\section{How to describe collected sensor data using Data Provenance}
+\label{sec:sn-provenance}
 
@@
 described and indexed, and this process can be related to Data Provenance
-\cite{db-provenance}.
+\cite{db-provenance}. Needless to say, Data Provenance is the 
+Metadata\footnote{Data about data} about the collected sensor data.
 
@@
 \begin{itemize}
-  \item each low-level observation as it describes the actual data;
-  \item the location where the collected data was generated;
-  \item the time when the data was collected.
+  \item \textbf{what was collected?}: it can be simply one or more low-level
+  observations, which is a set of values with or without a description;
+  \item \textbf{When was the data collected?}: is the the time when the data was
+  collected, as most sensor data must be used depending on its existance;
+  \item \textbf{Where was the data collected?}: it is primarily the location
+  where the collected data was observed. However, It is important to note that
+  the use of this metadata depends on the applicability of the data.
 \end{itemize}
@@
 networks. \cite{sn-time-series} describes strategies to used to cluster
-distributed sensor data as time series of data.
+distributed sensor data as time series of data, as well as to decrease the size
+of data streams being transmitted in real-time sensor networks when these
+collected data does not need to be transferred \cite{sn-data-reduction}.
+
+Finally, an important application of collected sensor data is the so-called
+blackbox of an aircraft, which carries the collection of raw data collected
+from a group of sensors placed in different places of the aircraft, as
+described by \cite{sn-exemple-blackbox}. The author describes the data is
+used in a process of data recovery and analysis in order to identify the
+possible reason of an aircraft accident, by using the time, and location of
+the aircraft, along with the different variables from the sensors.
+
+\subsection{What was collected?}
+
+Different sensor devices will provide different attribute sets regarding a
+specific observation. In general, these data sets does not have obvious names,
+which leads to difficulties when indicies are necessary in order to make the
+data searcheable. Specially for the purpose of data archival, as described
+earlier in this chapter, the collected data needs to receive and identity,
+which identifies the data by itself in a regular database
+\cite{relational-model}. Equally, \cite{sn-provenance} described that
+traditional sensor networks uses the collected data for analysis long after
+the attribute set has been collected, usually before it is categorized and
+indexed. Annotations and tags are also names given by \cite{sn-provenance} as
+means to describe collected sensor data.
+
+In a nutshell, data is described by the name of the observation, when this is
+not provided. For example,
+
+\begin{itemize}
+  \item \textbf{temperature} = 73.45: this is the value of the collected
+  observation, which in general, the attribute name ``temperature'' is not
+  provided in the data stream;
+  \item \textbf{temperature-scale} = ``Fahrenheit'': the metadata regarding the
+  temperature attribute, which describes the scale of the temperature.
+  \item \textbf{device-version} = 5: the metadata that describes the version of
+  the sensor device, since sensors can be replaced, and as a consequence, data
+  accuracy may be as well be changed.
+\end{itemize}
+
+\subsection{When was the data collected?}
+
+In order to use historical sensor data, the use of a point in time is the most
+important characteristic of data, as it is considered an important metadata
+for different types of sensor networks such as an aircraft the blackbox
+\cite{sn-exemple-blackbox}. Other projects which classifies data in a
+time-series \cite{sn-time-series-example}, while others provide approaches to
+decrease the creation of data based on clustering data before they are sent to
+the sink \cite{sn-time-series}.
+
+In different areas, temporal databases or time-series are the ones that takes
+the real-world time into consideration. In general, data on a temporal
+database can be classified by its dimention, as surveyed by \cite{db-temporal}:
+
+\begin{itemize}
+  \item \textbf{Valid time}: it is the point in time during which a fact
+  happened with respect to the real world. This value is only when data is captured, an
+  cannot be changed because of its nature, as well as placing a restriction over
+  Update operations;
+  \item \textbf{Transaction time}: is the point in time at which a fact is
+  collected, transferred or processed in the database. This value is used
+  during Insertion operations, as well as in Update ones.
+\end{itemize}
+
+When both temporal dimensions are applied, the data is consided to be 
+Bitemporal when \cite{db-temporal}. Usually, the data type of the valid and
+transaction time columns are simple timestamps, with any supported accuracy.
+For example, the following set of attribute and values shows a temperature of
+70 degrees that was observed on November 15, 2009, but collected only 3 days
+later.
+
+\begin{itemize}
+  \item \textbf{temperature} = 70;
+  \item \textbf{fact-time} = Nov 15, 2009 10:45 PM PST
+  \item \textbf{transaction-time} = Nov 18, 2009 11:11 PM PST
+\end{itemize}
+
+When it comes to Data Provenance in sensor networks, \cite{sn-provenance}
+shows that the collected data for historical purpose is usually saved in
+self-described flat-files by file-names, containing complicated naming
+conventions that includes the timestamp information about the data. For this
+reason, time is an important metadata needed in any data model used, since it 
+helps identifying data uniquely for a given data set.
+
+\subsection{Where was the data collected from?}
+
+Depending on the application of the collected sensor data, the location in
+which the data was collected may be needed. For instance, the geographic
+location of the collected data is important to identify the specific
+GPS\footnote{Global Positioning System} coordinates of the sensor device
+\cite{sn-ex02}. Similarly, the position of any event captured in a black-box of
+an aircraft carries is also used as a metadata for the event. It will either
+depend on the sensor device provide such sampling attribute, or entered
+manually at the database.
+
+A correlation between time and location is presented by \cite{sn-time-series}
+as an important feature to reduce the traffic generated in a sensor network by
+clustering similar data when the sensor devices are placed in concentrated
+environment.

### 2009-11-20 docs/marcello-dissertation/thesis/chapter3-taxonomic-classification-of-persistence-storage-for-sensor-networks.tex
Revision of the chapters 2-6 regarding Data Provenance in sensor networks:
* Chapter 2
  - Adding section regarding provenance

* Chapter 3
  - Splitting section 1 into 2 different ones; TODO: Draw another diagram. 
  - Adding provenance information to the taxonomy of data model
  
* Chapter 4
  - Adding references regarding to the missing location in the NetBEAMS data;

* Chapter 5
  - References to the provenance sections to the keys definition;
  
* Chapter 6
  - References to the provenance sections to the keys definition;

git-svn-id: svn://localhost/netbeams/branches/marcello/persistence

@@
 
-\section{Nature and Location of Sensor Data Taxonomy}
+\section{Nature of Sensor Data Taxonomy}
 
@@
   \item \textbf{Real-time Data Stream}: data is used as real-time data stream
-  during data query \cite{sn-intro01}.
-    \begin{itemize}[label=\textbullet]
-        \item \textbf{Local Storage}: The data can be temporarily persisted
-        in-memory, or a secondary memory device in a local storage device.
-    \end{itemize}
+  during data query \cite{sn-intro01};
   \item \textbf{Archival Data}: data is used for historical purposes such as
   data analysis or data fusion \cite{sn-intro01, sn-intro02}.
-    \begin{itemize}[label=\textbullet]
+\end{itemize}
+
+\section{Location of Sensor Data Taxonomy}
+
+No matter the type of resource management used, the first taxonomy presented is
+related to the purpose of the use of the collected sensor data. That is, when
+and how the collected data is going to be used as described in section
+\ref{sec:sn-data-purpose}. In addition to it, the mechanism for data storage
+described in section \ref{sec:sn-infrastructure} can be related to each of
+them. The related taxon is depicted by figure \ref{fig:taxonomy-data-purpose}.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics{../diagrams/taxonomy-data-purpose}
+  \caption{The Nature and Location of Sensor Data Taxonomy}
+  \label{fig:taxonomy-data-purpose}
+\end{figure}
+
+\begin{itemize}
+  \item \textbf{Local Storage}: The data can be temporarily persisted
+   in-memory, or a secondary memory device in a local storage device. It is
+  directly related to the real-time data stream taxon, since it occurs at
+  the place of the sensor device;
   \item \textbf{External Storage}: The data is usually located in an
@@
   and database sharding approaches \cite{db-shard01} \cite{db-shard02}.
-    \end{itemize}
 \end{itemize}
@@
 \end{itemize}
+
+No matter the type of data model used, the support for Data Provenance as shown
+in section \ref{sec:sn-provenance} is as important as the data being
+collected. For this reason, the data model taxonomy must have support for the
+data types that support the use of provenance-related attributes in the
+collected sensor data.
 

### 2009-11-20 docs/marcello-dissertation/thesis/chapter4-netbeams-in-depth-requirements-analysis-and-specification.tex
Revision of the chapters 2-6 regarding Data Provenance in sensor networks:
* Chapter 2
  - Adding section regarding provenance

* Chapter 3
  - Splitting section 1 into 2 different ones; TODO: Draw another diagram. 
  - Adding provenance information to the taxonomy of data model
  
* Chapter 4
  - Adding references regarding to the missing location in the NetBEAMS data;

* Chapter 5
  - References to the provenance sections to the keys definition;
  
* Chapter 6
  - References to the provenance sections to the keys definition;

git-svn-id: svn://localhost/netbeams/branches/marcello/persistence

--- 2009-11-15 docs/marcello-dissertation/thesis/chapter4-netbeams-in-depth-requirements-analysis-and-specification.tex
@@
 \chapter{Netbeams in Depth: Requirements Analysis and Specification}
+\label{chap:netbeams-overview}
 
@@
   \item The sampling frequency rate could be configured in ranges of 1, 6 or
-  15 minutes, depending on the staff's speficiation. 
+  15 minutes, depending on the staff's speficiation;
+  \item Data for final users are published with Provenance Data regarding time,
+  but not regarding location.
 \end{itemize}
@@
   \item Scalable system with low degree of maintanance;
-  \item Data searchable in near-real-time
+  \item Data searchable in near-real-time.
 \end{itemize}
@@
         \begin{tabular}{lr}
-          \textit{Data Consumers ( DSP Message ) := Verify the DSP
-          Message's Header + Verify the matcher rules (which contains the list
-          of consumers)}
+          \textit{Data Consumers ( DSP Message )} := \\ 
+          Verify the DSP Message's Header + \\
+          Verify the matcher rules, which contains the list of consumers for
+          the given message
         \end{tabular}

### 2009-11-20 docs/marcello-dissertation/thesis/chapter5-taxonomic-persistence-storage-selection-for-netbeams.tex
Revision of the chapters 2-6 regarding Data Provenance in sensor networks:
* Chapter 2
  - Adding section regarding provenance

* Chapter 3
  - Splitting section 1 into 2 different ones; TODO: Draw another diagram. 
  - Adding provenance information to the taxonomy of data model
  
* Chapter 4
  - Adding references regarding to the missing location in the NetBEAMS data;

* Chapter 5
  - References to the provenance sections to the keys definition;
  
* Chapter 6
  - References to the provenance sections to the keys definition;

git-svn-id: svn://localhost/netbeams/branches/marcello/persistence

@@
 For instance, all data needed for the YSI sonde data, as well as all necessary
-provenance data that describes the data, are represented by means of key-value
-pairs \ref{fig:persistence-example-kvp}.
+provenance data that describes the data, depicted in section
+\ref{sec:sn-provenance}, are represented by means of key-value pairs
+\ref{fig:persistence-example-kvp}.
 
@@
       duplication, as well as establishing data associations through table
-      relationships.
+      relationships;
+      \item Data Provenance can be provided through data types such as
+      ``Timestamp'' for time or float for the GPS float-based coordinates; 
     \end{enumerate} 
@@
       \item Since data may be repeated, no data normalization is done, so that
-      integrity is done in the application layer.
+      integrity is done in the application layer;
+      \item Data provenance can be added through the use of keys and relating
+      values for the time information and location of the data.
     \end{enumerate}

### 2009-11-20 docs/marcello-dissertation/thesis/chapter6-data-persistence-component-design-and-architecture.tex
Revision of the chapters 2-6 regarding Data Provenance in sensor networks:
* Chapter 2
  - Adding section regarding provenance

* Chapter 3
  - Splitting section 1 into 2 different ones; TODO: Draw another diagram. 
  - Adding provenance information to the taxonomy of data model
  
* Chapter 4
  - Adding references regarding to the missing location in the NetBEAMS data;

* Chapter 5
  - References to the provenance sections to the keys definition;
  
* Chapter 6
  - References to the provenance sections to the keys definition;

git-svn-id: svn://localhost/netbeams/branches/marcello/persistence

@@
 chapter \ref{netbeams-architecture}, the following set of keys are expected to
-be part of the design of the data model:
+be part of the design of the data model. Any key represented by time or
+location are related to the provenance support discussed in section
+\ref{sec:sn-provenance}:
 

### 2009-11-22 docs/marcello-dissertation/thesis/chapter1-introduction.tex
Review of chapter 1 after Prof. Puder's feedback.
 - More abstract introduction, without mentioning keywords
   used in the work such as NetBEAMS, DSP, KVP, etc;

 - Reduced structure

git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@655 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 \cite{sn-ex03}, provides online data from different buoys in different coast
-shores around the world regarding water quality, temperature, tides, etc, what
+shores around the world regarding water quality, tide information, etc, what
 characterizes sensor network for environmental monitoring \cite{sn-ex01}.
@@
 can be taken to interrogate sensor devices indirectly by accessing its data in
-places called network sink. These data nodes can be located in the network or 
-in a centralized host. In this way, the query process can take place on the
-network nodes themselves or on the centralized data sink. The former is
-generally used when the data is collected at the sensor node on its local
-storage, being shown by \cite{sn-storage01} \cite{sn-storage03}
-\cite{sn-storage04} that this strategy helps mitigating energy consumption.
-However, this strategy is used when data archival is not the primary reason,
-given the limited resources on the sensor devices. For this reason, the the
-latter strategy has been used when the collected data is primarily archived in
-a centralized database system for later reuse \cite{sn-storage02}.
-
-When it comes to data model implementation, the relational data model
-\cite{relational-model} is one of the most used approaches to represent data in
-sensor network databases in any of the query processing strategies, and in most cases 
-with a modified version of the SQL language \cite{sn-db-newop}. However, the
-design of the database system must be done with prior knowledge of the sensor
-types when choosing the relational data model, as the database design must be
-normalized \cite{db-normalization} for the entities already identified for the
-system. For instance, the introduction of a new sensor device might represent
-a potential change on the database structure by seeking a new normalized
-version of the database structure. In addition to challenges of maintaining
-the database schema normalized, the use of Data Provenance approaches in
-sensor network data \cite{sn-provenance} usually addresses the problem related
-with the lineage of data  and its full history. However, once the data model
-is chosen and implemented, one of the most important non-functional
-requirements in sensor networks is scalability of the persistence storage.
-Since an operational sensor network can produce huge amounts of data, it must
-be able to cope with the increase and data load during data gathering, and yet
-provide users the same performance during data retrieval by using different
-strategies such a data-centric storage \cite{sn-storage03}.
-
-In summary, this dissertation discusses the solution of a
-centralized data-centric persistence storage component for the NetBEAMS
-\cite{netbeams2009}, an implementation of a sensor network that primarily
-focused on environmental data. Different challenges can be related to the
-selection of a technology that handles data persistence in the context of sensor
-networks: the network infrastructure, the nature and location of the collected
-data, the chosen data model, among others. First, NetBEAMS needs to provide
-means to safely represent sensor devices' properties without the need of
-restructuring any existing data model, taking into account the use cases from
-the main users. For example, a marine biologist may query the the persistence
-storage regarding the temperature of a specific geographic region at the San
-Francisco Bay at a given time frame. Second, since a sensor network may
-experiment grown in terms of number of sensors, and consequently the number of
-data produced, the persistence  storage must be able to reasonably scale and
-be well structured in terms of storage space with the least service
-interruption. For instance, if the persistence storage is close to reach its
-storage capacity, what should the network administrator do? Last, but not
-least, the means to access the data for reuse is the most important features
-of the system, taking into account the different data formats used by the main
-users of the system. How to export specific collected data to a spreadsheet or
-archive files to be published or shared with other users? For these, and many
-other reasons, the persistence storage problem addressed by this work can be
-seen as a very challenging experience to solve.
+places called network sink. Similarly, the data model and database system
+location may also influence in the process of reusing the collected data. In
+this way, the lack of a persistence layer for existing sensor networks requires
+a deep understanding about its infrastructure, the nature and location of the
+collected data, the chosen data model representation, among others. For
+instance, how can a data persistence layer safely represent the collected sensor
+devices' properties without the need of restructuring any existing data model?
+Should the end users' skills in database system be taken into account or the
+use of programming languages to access the data \cite{sn-programming-language}?
+This problem is currently faced by the departent of Biology at San
+Francisco State University with a marine sensor network managed by
+researchers without Computer Science skills. While the data access may impose
+the difficulties, technical issues may also rise while the execution of a
+sensor network such as growth in terms of number of sensor devices and,
+consequently, the number of data produced. Last, but not least, the means to
+access the data for reuse is the most important features of the sensor
+network, taking into account the different data formats used by the main users
+of the system. How can specific collected sensor data be exported to other
+formats such as spreadsheets or archived files meant to be shared among
+research peers? For these, and many other reasons, the inception of a
+persistence storage can be a very challenging problem to solve.
 
 In view of the fact that the persistence storage of a sensor network imposes
-many different challenges, and taking into account the case study NetBEAMS, this
-work's main strategy focused analyzing the different specifications of data
-persistence in the literature in order to find new types of mechanisms to
-provide data persistence for sensor network data. In this way, the kernel of
-this dissertation is focused on providing solutions to the real-world problem
-face by the Romberg Tiburon Center and its use of NetBEAMS. Furthermore,
-the evaluation of different technologies is discussed in order to provide a
-loosely-coupled component-based solution based on the infrastructure of the 
-DSP Platform. For the reason of providing a 'lighter' data model, a subtype of
-the Key-Value Pair \cite{db-kvp} data model was chosen to easily capture the
-description of any sensor device's properties, along with the use of time
-series data models \cite{sn-provenance} for time snapshots. In that way, data
-modeling process does not represent a new refactoring each time a new sensor 
-type is added or removed from the persistence system, giving users means to 
-retrieve data accordingly. Similarly, in contemplation of scalability and the
-use of Data-Centric approach \cite{sn-storage03}, this work proposes the use of
-Database Sharding \cite{db-shard01} \cite{db-shard02}, on a centralized
-database system.
+many different challenges, this work proposes the review of the current state of
+the art for sensor networks in order to provide a data persistence layer for an
+existing sensor network. As a matter of fact, this work's first contribution is
+the inception of a set of basic taxonomies for data persistence in sensor
+networks, proposed to drive an empirical evaluation of different technologies
+used in the sensor network community. Furthermore, it also takes a look at
+different alternatives not yet explored through experiments for a selected
+sensor network as a case study. Similarly, this work makes another novice
+contribution by suggesting a ``lighter" data model and database infrastructure
+that easily captures sensor network device's observations, taking into account
+the non-predictable scalability of its infrastructure, as well as the aspects of
+the data's nature such as identity, time and location of the collected data.
 
-In order to provide such mechanism, this work first takes a look at the related
-literature to study the properties of sensor networks, in the context of its
-produced data and infrastructure for storage and query processing. Then, a
-summarized list of taxonomies is proposed to be the basis of an empirical  the
-use of shard keys based on the geographic origin of the collected data as cited to be one of the options when working with Provenance data in sensor networks \cite{sn-provenance}.
-As a result, the data persistence system component was conceived using the plug-and-play capability from the DSP platform without doing  any changes to the current DSP Platform, providing an scalable solution that can adjust according to any data load increase.
-Finally, the selected technology provides different data access
-mechanisms, which directly addressed problems related to data access and
-Information Fusion \cite{sn-info-fusion}, providing export-import capabilities
-and visualization through the availability of APIs written in different
-popular programming languages.
-
-This dissertation report is organized as follows: the next chapter is an 
-overview of the state of art on topics that supports the implementation of the
-solution of this work, providing the necessary background foundation for the 
-following chapters. Chapter 3 presents the design and architecture of the 
-conceived persistence module, describing the requirements needed to be
-addressed,  while Chapter 4 describes the implementation based on the design. 
-The analysis and investigation of the implemented solution is shown through 
-the analyzes of the experiment results on Chapter 5. Last, but not least, 
-Chapter 6 presents the conclusions of this work, followed by the possible 
-future works and recommendations presented on Chapter 7.
+This dissertation is organized as follows: the next chapter describes the
+literature review as the state of art on sensor networks and data persistence, 
+providing the needed foundation for the design of taxonomies in data
+persistence for sensor networks in chapter 3. Then, chapter 4 examines the
+selected case study and its requirements for a persistence layer, whose
+technology properties are evaluated in chapter 5 against the developed
+taxonomies. The culmination of this work is chapter 6, which details the design
+and implementation of a data persistence solution for the selected case
+study. Likewise, chapter 7 discusses the solution by presenting the
+measurements and conducted experiments, and describing the findings
+and contributions of this work. Last, but not least, chapter 8 presents the
+conclusions of this work, together with the possible future works not addressed
+here.

### 2009-11-23 docs/marcello-dissertation/thesis/chapter2-state-of-the-art.tex
Revisions for chapters 2 and 3 after prof. Puder's reviews.
- Removing blurried images;
- Adding the data provenance taxonomy

git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@656 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 described in the literature that solved problems related to data modeling and
-representation in sensor networks. Then, it presents the design, architecture 
-and implementation details of the DSP Platform, which is the bases for this 
-dissertation. Finally, this chapter will briefly give a tutorial on advanced 
-topics such as the differences between the KVP data model and the relational 
-data model, as well as discussing about the approaches used on Database
-Sharding.
+representation in sensor networks.
 
@@
 Sensor Networks are specialized types of network systems comprised of network 
-nodes with specific goals to observe and collect data. The following
-subsections describe classes of characteristis of these types of sensor
-networks, with applications on research, industrial or militery.
+nodes with specific goals to observe and collect data. Data Persistence in
+Sensor Networks, however, depends on the approaches used to store the data,
+since data nodes can be located close to the devices in the network or in a
+centralized node. In this way, the query process can take place on the network
+nodes themselves or on the centralized data sink. The former approach is
+generally used when the data is collected at the sensor node on its local
+storage \cite{sn-storage01}\cite{sn-storage03}, strategy mainly used to
+maximize the sensor's resources while minimizing energy consumption
+\cite{sn-storage04}. However, this strategy is used when data archival is not
+the primary reason, given the limited resources on the sensor devices. For
+this reason, the the latter strategy has been used when the collected data is
+primarily archived in a centralized database system for later reuse
+\cite{sn-storage02}.
+
+When it comes to data representation, the relational data model
+\cite{relational-model} is one of the most used approaches to represent data
+in sensor network databases in any of the query processing strategies and, in
+most cases, with a modified version of the SQL language \cite{sn-db-newop}.
+However, the design of the database system must be done with prior knowledge
+of the sensor types when choosing the relational data model, as the database
+design must be normalized \cite{db-normalization} for the entities already
+identified for the system. For instance, the introduction of a new sensor
+device might represent a potential change on the database structure by seeking
+a new normalized version of the database structure. In addition to challenges
+of maintaining the database schema normalized, the use of Data Provenance
+approaches in sensor network data \cite{sn-provenance} usually addresses the
+problem related with the lineage of data  and its full history. However, once
+the data model is chosen and implemented, one of the most important
+non-functional requirements in sensor networks is scalability of the
+persistence storage. Since an operational sensor network can produce huge
+amounts of data, it must be able to cope with the increase and data load
+during data gathering, and yet provide users the same performance during data
+retrieval by using different strategies such a data-centric storage
+\cite{sn-storage03}.
+
+The following subsections describe classes of characteristis of these types of
+sensor networks, with applications on research, industrial or militery.
 
@@
 
-\section{How to describe collected sensor data using Data Provenance}
+\section{Data Provenance: How to describe collected data?}
 \label{sec:sn-provenance}

### 2009-11-23 docs/marcello-dissertation/thesis/chapter3-taxonomic-classification-of-persistence-storage-for-sensor-networks.tex
Revisions for chapters 2 and 3 after prof. Puder's reviews.
- Removing blurried images;
- Adding the data provenance taxonomy

git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@656 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 networks and different approaches for persistence. This section analyzes these
-approaches and presents the different approaches through taxonomy
-\footnote{Taxonomy is the practice and science of classification, by using
-taxonomic units known as taxa (singular taxon)}, which will be used throughout
-this work to evaluate different existing database systems implementations,
-taking into account different requirements and specifications of different
-sensor networks.
+approaches and presents the different approaches through 
+taxonomy\footnote{Taxonomy is the practice and science of classification, by
+using taxonomic units known as taxa (singular taxon)}, which will be used
+throughout this work to evaluate different existing database systems
+implementations, taking into account different requirements and specifications
+of different sensor networks.
 
@@
 described in section \ref{sec:sn-infrastructure} can be related to each of
-them. The related taxon is depicted by figure \ref{fig:taxonomy-data-purpose}.
+them. The related taxon is depicted by figure \ref{fig:taxonomy-data-location}.
 
@@
   \centering
-  \includegraphics{../diagrams/taxonomy-data-purpose}
+  \includegraphics{../diagrams/taxonomy-data-location}
   \caption{The Nature and Location of Sensor Data Taxonomy}
-  \label{fig:taxonomy-data-purpose}
+  \label{fig:taxonomy-data-location}
 \end{figure}
@@
 
+\section{Data Provenance Taxonomy}
+
 No matter the type of data model used, the support for Data Provenance as shown
-in section \ref{sec:sn-provenance} is as important as the data being
-collected. For this reason, the data model taxonomy must have support for the
-data types that support the use of provenance-related attributes in the
-collected sensor data.
+in section \ref{sec:sn-provenance} is as important as the data being collected.
+For this reason, the data model taxonomy must have support for the data types
+that support the use of provenance-related attributes in the collected sensor
+data. In this way, Figure \ref{fig:taxonomy-data-provenance} describes the data types
+for the metadata support.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics{../diagrams/taxonomy-data-provenance}
+  \caption{Data Provenance Taxonomy}
+  \label{fig:taxonomy-data-provenance}
+\end{figure}
+
+\begin{itemize}
+  \item \textbf{What: Data Identity}: the data type that uniquely identifies
+  the collected sensor dat;
+  \item \textbf{When: Time Dimension}: the data type that identifies the
+  collected data;
+  \item \textbf{Where: Data Location}: an optional data type that identifies
+  where the data came from.
+\end{itemize}
 

### 2009-11-23 docs/marcello-dissertation/thesis/chapter1-introduction.tex
Word Fixes in chapter one.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@657 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 Sensor networks are commonly used in different areas for different purposes
-such as scientific measurements and general population use \cite{sn-intro01}
-\cite{sn-intro02}. In the scientific community, they serve as tools to monitor
-the state of the environment by storing samples of data in specific periods of
-time. For instance, the NASA Jet Propulsion Laboratory uses the Volcano
-SensorWeb \cite{sn-ex02}, a sensor network which aims to collect data from
-specific volcanoes in Alaska and South Pacific to be used with heuristics
-regarding dangerous activities. Similarly, the National Data Buoy Center, a
-department of the National Oceanic and Atmospheric Administration (NOAA)
-\cite{sn-ex03}, provides online data from different buoys in different coast
-shores around the world regarding water quality, tide information, etc, what
-characterizes sensor network for environmental monitoring \cite{sn-ex01}.
-Therefore, sensor networks can be seen as a very important scientific research
-and commercial field with direct application and use of its produced data.
+such as accurate measurements and the general public use \cite{sn-intro01}
+\cite{sn-intro02}. In the scientific community, they serve as tools to watch
+the state of the environment by storing samples of data at particular periods
+of time. For example, the NASA Jet Propulsion Laboratory uses the Volcano
+SensorWeb \cite{sn-ex02}, a sensor network which aims at collecting data from
+individual volcanoes in Alaska and South Pacific to be used with heuristics
+regarding hazardous activities. Similarly, the National Data Buoy Center, a
+division of the National Oceanic and Atmospheric Administration (NOAA)
+\cite{sn-ex03}, provides online information from different buoys in different
+shores around the world regarding water quality, current information, etc,
+what characterizes sensor network for environmental monitoring \cite{sn-ex01}.
+Therefore, sensor networks can be seen as an extremely valuable scientific
+research and commercial sector with direct application and use of its produced
+data.
 
@@
 collected data, the chosen data model representation, among others. For
-instance, how can a data persistence layer safely represent the collected sensor
-devices' properties without the need of restructuring any existing data model?
-Should the end users' skills in database system be taken into account or the
-use of programming languages to access the data \cite{sn-programming-language}?
-This problem is currently faced by the departent of Biology at San
-Francisco State University with a marine sensor network managed by
-researchers without Computer Science skills. While the data access may impose
-the difficulties, technical issues may also rise while the execution of a
-sensor network such as growth in terms of number of sensor devices and,
-consequently, the number of data produced. Last, but not least, the means to
-access the data for reuse is the most important features of the sensor
-network, taking into account the different data formats used by the main users
-of the system. How can specific collected sensor data be exported to other
-formats such as spreadsheets or archived files meant to be shared among
-research peers? For these, and many other reasons, the inception of a
-persistence storage can be a very challenging problem to solve.
+instance, how can a data persistence layer safely represent the collected
+sensor devices’ properties without the necessity of restructuring any existing
+data model? Should the end users’ skills in database system be taken into
+account or the use of programming languages to access the data
+\cite{sn-programming-language}? This problem is currently faced by the
+department of Biology at San Francisco State University with a marine sensor
+network managed by researchers without Computer Science skills. While the data
+access may enforce the difficulties, technical issues may also increase while
+the implementation of a sensor network such as growth in terms of the number of
+sensor devices and, consequently, the number of data produced. Last, but not
+least, the way to read the data for reuse is the most striking features of the
+sensor network, taking into account the different data formats used by the
+primary users of the system. How can individual collected sensor data be
+exported to other formats such as spreadsheets or archived files meant to be
+shared among research peers? For these reasons, the inception of a persistence
+storage can be an exceptionally challenging problem to solve.
 
 In view of the fact that the persistence storage of a sensor network imposes
-many different challenges, this work proposes the review of the current state of
-the art for sensor networks in order to provide a data persistence layer for an
-existing sensor network. As a matter of fact, this work's first contribution is
-the inception of a set of basic taxonomies for data persistence in sensor
-networks, proposed to drive an empirical evaluation of different technologies
-used in the sensor network community. Furthermore, it also takes a look at
-different alternatives not yet explored through experiments for a selected
-sensor network as a case study. Similarly, this work makes another novice
-contribution by suggesting a ``lighter" data model and database infrastructure
-that easily captures sensor network device's observations, taking into account
-the non-predictable scalability of its infrastructure, as well as the aspects of
-the data's nature such as identity, time and location of the collected data.
+many different challenges, this work proposes the assessment of the current
+state of the art for sensor networks in order to provide a data persistence
+layer for an existing sensor network. As a matter of fact, this work’s first
+contribution is the creation of a set of basic taxonomies for data persistence
+in sensor networks, proposed to drive an empirical analysis of different
+technologies used in the sensor network community. Furthermore, it also takes a
+look at different alternatives not yet explored through experiments for a
+selected sensor network as a case study. Similarly, this work makes another
+novice contribution by suggesting a “lighter” data model and database
+infrastructure that simply captures sensor network devices' observations,
+taking into account the non-predictable scalability of its infrastructure, as
+well as aspects of the nature of the data such as identity, time and location
+of the collected data.
 
-This dissertation is organized as follows: the next chapter describes the
-literature review as the state of art on sensor networks and data persistence, 
-providing the needed foundation for the design of taxonomies in data
-persistence for sensor networks in chapter 3. Then, chapter 4 examines the
-selected case study and its requirements for a persistence layer, whose
-technology properties are evaluated in chapter 5 against the developed
-taxonomies. The culmination of this work is chapter 6, which details the design
-and implementation of a data persistence solution for the selected case
-study. Likewise, chapter 7 discusses the solution by presenting the
-measurements and conducted experiments, and describing the findings
-and contributions of this work. Last, but not least, chapter 8 presents the
-conclusions of this work, together with the possible future works not addressed
-here.
+This paper is organized as follows: the next chapter describes the literature
+review as the state of art on sensor networks and data persistence, providing
+the underlying foundation for the design of taxonomies in data persistence for
+sensor networks in chapter 3. Then, chapter 4 examines the selected case study
+and its requirements for a persistence layer, whose technology properties are
+evaluated in chapter 5 against the developed taxonomies. The culmination of
+this work is chapter 6, which details the design and implementation of a data
+persistence solution for the selected case study. Likewise, chapter 7 discusses
+the solution by presenting the measurements and conducted experiments, and
+describes the findings and contributions of this work. Last, but not least,
+chapter 8 presents the conclusions of this work, together with the possible
+future works not addressed here.

### 2009-11-24 docs/marcello-dissertation/thesis/chapter1-introduction.tex
Fixing sentences, references, and misspelled words in chapters 1 and 2.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@658 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 instance, how can a data persistence layer safely represent the collected
-sensor devices’ properties without the necessity of restructuring any existing
-data model? Should the end users’ skills in database system be taken into
+sensor devices' properties without the necessity of restructuring any existing
+data model? Should the end users' skills in database system be taken into
 account or the use of programming languages to access the data
@@
 state of the art for sensor networks in order to provide a data persistence
-layer for an existing sensor network. As a matter of fact, this work’s first
+layer for an existing sensor network. As a matter of fact, this work's first
 contribution is the creation of a set of basic taxonomies for data persistence
@@
 selected sensor network as a case study. Similarly, this work makes another
-novice contribution by suggesting a “lighter” data model and database
+novice contribution by suggesting a ``lighter" data model and database
 infrastructure that simply captures sensor network devices' observations,

### 2009-11-24 docs/marcello-dissertation/thesis/chapter2-state-of-the-art.tex
Fixing sentences, references, and misspelled words in chapters 1 and 2.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@658 559edde2-a4d4-44e0-96fc-4035e40d7396

--- 2009-11-14 docs/marcello-dissertation/thesis/chapter2-state-of-the-art.tex
@@
 
-Different aspects related to two major fields of Computer Science have
-motivated to the inception of this work. In the perspective of Database
-Systems and Distributed Systems, this work seeks a different approach for a
-persisting collected data from sensor networks into a storage system that can
-be seen as a cluster of distributed databases. Likewise, in the perspective of
-Software Engineering, this work adds a major capability to an existing sensor 
-network framework, meeting the requirements and interfaces. In general, this 
-chapter gives the foundation based on the motivations for the inception of a 
+Different aspects related to two fields of Computer Science have motivated to
+the commencement of this work. In the prospect of Database Systems and
+Distributed Systems, this work seeks a different approach for a persisting
+collected data from sensor networks into a storage system that can be seen as a
+cluster of distributed databases. Likewise, in the perspective of Software
+Engineering, this work adds a crucial capability to an existing sensor network
+framework, meeting the requirements and interfaces. In general, this chapter
+gives the foundation based on the motivations for the establishment of a
 different approach to persisting data using a different data model, covering
@@
 described in the literature that solved problems related to data modeling and
-representation in sensor networks. Then, it presents the design, architecture 
-and implementation details of the DSP Platform, which is the bases for this 
-dissertation. Finally, this chapter will briefly give a tutorial on advanced 
-topics such as the differences between the KVP data model and the relational 
-data model, as well as discussing about the approaches used on Database
-Sharding.
+representation in sensor networks.
 
-\section{Sensor Networks Survey}
+\section{Survey in Sensor Networks}
 
-Sensor Networks are specialized types of network systems comprised of network 
-nodes with specific goals to observe and collect data. The following
-subsections describe classes of characteristis of these types of sensor
-networks, with applications on research, industrial or militery.
+Sensor Networks are specialized types of network systems comprised of nodes
+with specific goals to monitor and collect data. Data Persistence in Sensor
+Networks, however, depends on the approaches used to store the data, since data
+nodes can be located close to the devices in the network or in a centralized
+node. In this way, the query process can take place on the network nodes
+themselves or in the centralized data sink. The former approach is used when
+the data is collected at the sensor node on its local storage
+\cite{sn-storage01}\cite{sn-storage03}, strategy typically used to maximize
+the sensor's resources while minimizing energy consumption
+\cite{sn-storage04}. However, this strategy is used when data archival is not
+the main purpose, given the limited resources on the sensor devices. For this
+reason, the latter strategy has been used when the collected data is generally
+archived in a centralized database system for later reuse \cite{sn-storage02}.
+
+When it comes to data representation, the relational data model
+\cite{relational-model} is one of the most used approaches to represent data in
+sensor network databases in any of the query processing strategies and, in most
+cases, with a modified version of the SQL language \cite{sn-db-newop}. However,
+the design of the database system must be done with prior knowledge of the
+sensor types when choosing the relational data model, as the database design
+must be normalized \cite{db-normalization} for the entities already identified
+for the system. For example, the introduction of a new sensor device might
+represent a potential change on the database structure by seeking a new
+normalized version of the database design. In addition to challenges of
+maintaining the database schema normalized, the use of Data Provenance
+approaches in sensor network data \cite{sn-provenance} usually addresses the
+problem related with the origin of data and its entire history. However, once
+the data model is chosen and implemented, one of the most prominent
+nonfunctional requirements in sensor networks is scalability of the persistence
+storage. Since an operational sensor network can generate enormous amounts of
+data, it must be able to cope with the increase and data load during data
+gathering, and yet provide users the same performance during data retrieval by
+using different strategies such a data-centric storage \cite{sn-storage03}.
+
+The following subsections describe classes of characteristis of these types of
+sensor networks, with applications on research, industrial or militery.
+
+The following subsections describe classes of characteristics of these types of
+sensor networks, with applications on research, commercial or military.
 
@@
 
-Each of the nodes can be deployed in a random way such as being dropped
-from an airplane, or in a manual fashion at deliberately chosen geographical 
-points \cite{sn-intro01}. Moreover, the deployment can be done as a one-time
-event or an iterative way.
+Each of the nodes can be deployed in a arbitrary method
+such as being dropped from an airplane, or in a manual mode at carefully chosen
+geographical points \cite{sn-intro01}. Moreover, the deployment can be done like a
+one-time event or an iterative process.
 
-However, nodes can be able to move around the deployed area or be static in
-this place. In the case of the degree of the former behaviour, a sensor node
-can move occasionally or continuously, being a result of an incidental 
+However, nodes can be able to move around the deployed locality or remain fixed
+in this place. In the case of the amount of the former behaviour, a sensor node
+can move occasionally or continuously, being a result of an accidental
 side-effect or a desired property of the device, where they are an automotive
@@
 being limited in size, they are also limited in resources and computational
-power, forcing their conception limited to have its energy stored in either
-batteries or scavenged from solar cells \cite{sn-intro01}.
+power, forcing their  conception limited to maintain its energy stored  either
+on batteries or scavenged from solar cells \cite{sn-intro01}.
 
-Since sensor devices range on size and capabilities, the cost for each of the
+Since sensor devices range in size and capabilities, the cost for each of the
 devices very from hundreds to thousands of dollars, depending on its
-functionality, energy utilization, etc. For thie reason, \cite{sn-intro01}
+functionality, energy utilization, etc. For this reason, \cite{sn-intro01}
 explains that a sensor network may potentially contain similar sensor devices,
-as they can contain dissimilar ones from each other, which characterizes the heterogeneity 
-of a sensor network by the hardware capabilities such as GPS enabled or having
-GSM  network connectivity.
+as they can contain dissimilar ones from each other, which characterizes the
+heterogeneity of a sensor network by the hardware capabilities such as GPS
+enabled or having GSM network connectivity.
 
@@
 they can differ in types of topology: single-hop, star, networked stars, tree
-and graph, where each of them may affect the characteristics of the network as it 
-may also influence on data gathering and routing mechanisms. While in a
+and graph, where each of them may affect the characteristics of the network as
+it may also influence on data collection and routing mechanisms. While in a
 single-hop sensor network there are only two levels of sensor connectivity, the
@@
 Last, but not least, sensor networks may be influenced its coverage designed in
-its purpose, its size in terms of number of nodes and its relating lifetime. As
-sensors are deployed in different areas, they can be distributed in a sparse or
-dense way, yet providing necessary coverage for the environment. However, 
-considering that sensor devices are very likely to constant failures, projects 
-have provided redundant nodes \cite{sn-intro01}. This degree of coverage also 
-influences on the connectivity of the sensor nodes, where they can be
-considered connected when a connection is available at all times. On the other
-hand, nodes can communicate with other nodes in an intermittent way or sporadic
-way. In the former, the sensor node is partitioned, while in the latter the
-nodes are isolated most of the time and communicates at certain moments in
-time.
+its purpose, its size in terms of the number of nodes and its relating
+lifetime. As sensors are deployed in different areas, they can be distributed
+in a sparse or dense nature, yet providing essential coverage for the
+environment. However, considering that sensor devices are very likely to
+regular failures, projects have provided redundant nodes \cite{sn-intro01}.
+This level of coverage also influences on the connectivity of the sensor
+nodes, where they can be considered connected when a connection is available
+at all times. On the other hand, nodes can communicate with other nodes in a
+sporadic way. In the former, the sensor node is partitioned, while in the
+latter the nodes are isolated most of the time and communicates at certain
+moments in time.
 
@@
 Similar to the degree of coverage of the sensor network is the number of
-sensor nodes deployed in the network, where they can be in the order of
-dozens, hundreds or even thousand of nodes \cite{sn-intro01}.
+devices deployed in the network, where they can be in the order of dozens,
+hundreds or even thousand of nodes \cite{sn-intro01}.
 
-The application of the sensor network determines how long the sensor
-network may be exist. For example, the lifetime of an environmental sensor
-networks \cite{sn-ex01} may be not determined since its service serves in a
-regular basis, whereas the lifetime of a sensor network to determine the number
-of people in an event is limited to the event.
+The application of the sensor network determines how long the sensor network
+may be exist. For example, the existence of an environmental sensor networks
+\cite{sn-ex01} may be not determined since its use serves in a regular basis,
+whereas the lifetime of a sensor network to determine the number of people in
+an event is limited to the event.
 
-Finally, the design of sensor networks can be related to the constraints
-related to quality-of-service (QoS) such as real-time delivery within certain
-period of time or robustness such as maintaining data alive even though there
-a communication link is not available \cite{sn-intro02}.
+Finally, the design of sensor networks can be
+related to the constraints related to quality-of-service (QoS) such as
+real-time delivery within a certain period of time or robustness such as
+maintaining data alive even though there a communication link is not available
+\cite{sn-intro02}.
 
@@
 
-The most important purpose of a sensor network is related to its main product:
-the use of the raw data. Data collection in sensor networks is directly
-dependent on the specifications of the sensor network, its routing mechanisms
-and the final destination of the produced data. In a nutshell,
+The most valuable objective of a sensor network is related to its first
+product: the use of the raw data. Data collection in sensor networks is
+directly dependent on the specifications of the sensor network, its routing
+mechanisms and the final destination of the produced data. In a nutshell,
 \cite{sn-storage03} describes data as follows:
@@
 \begin{itemize}
-  \item \textbf{low-level property} as the raw data collected from the
-  sensor device; For example, the sea level as a number;
-  \item \textbf{Observation or Event} as the processed data in format of the final
-  observation. For instance, if the ocean is in its high tight or low tight.
+  \item \textbf{low-level property} as the raw data collected from the sensor
+  device; For example, the sea level as a number;
+  \item \textbf{Observation or Event} as the processed data in the format of
+  the final observation. For instance, if the ocean is in its highlight or low
+  tight.
 \end{itemize}
@@
 
-At first glance, the raw data serves as the primary product of the sensor
-network. However, the first consideration that must take place is regarding the
-purpose of the collected data, as \cite{sn-provenance} explains
-differentiates between two main use of the sensor data:
+At first glance, the raw data serves as the main outcome of the sensor network.
+However, the first consideration that must be considered is regarding the
+purpose of the collected data, as \cite{sn-provenance} explains differentiates
+between two main benefit of the sensor data:
 
 \begin{itemize}
-  \item \textbf{Real-time Data Stream}, as the generated data from the sensor is
-  actually not transmitted, but extracted from the device as it is requested.
-  The data is still in its natural format specified by the sensor manufacturer;
-  \item \textbf{Archival Data}, as the collected data have historical value or will
-  be use by other applications at a future time. 
+  \item \textbf{Real-time Data Stream}, as the generated data from the sensor
+  is not transmitted, but extracted from the device as it is requested. The
+  data is still in its standard form specified by the sensor manufacturer;
+  \item \textbf{Archival Data}, as the collected data include historical
+  significance or will be use by other applications at a later time.
 \end{itemize}
@@
 archival data needs to be stored in a secondary memory. Nonetheless, given the
-nature of the sensor device specification, the low-level data produced
-by a given sensor device might not contain necessary information regarding the
-identity of the data itself. The use of Data Provenance techniques is described
-by \cite{sn-provenance} as an important step to provide the collected data 
+nature of the sensor device specification, the low-level data produced by a
+given sensor device might not include vital information regarding the identity
+of the data itself. The use of Data Provenance techniques is described by
+\cite{sn-provenance} as an essential step to provide the collected data
 specific metadata regarding its location and lineage.
 
-Under these circumstances, the collected data needs first to face its jorney
+Under these circumstances, the collected data needs first to face its journey
 starting from its transmission from its producing device to the specified data
-sink as specified by \cite{sn-storage01} \cite{sn-storage02} \cite{sn-storage03}.
+sink as specified by \cite{sn-storage01}\cite{sn-storage02}\cite{sn-storage03}.
 
@@
 Given that collected data can be used for different purposes as described in
-section \ref{sec:sn-data-purpose}, the sensor node that produces the data is
-responsible to send the raw data to a specified data storage destination. Upon
-producing data, a sensor network node is specified to deal with the produced
-data, where in most cases, \cite{sn-storage03} described that depending on the
-sensor network infrastructure (see section \ref{sec:sn-infrastructure}),
-different storage mechanisms are provided:
+section \ref{sec:sn-data-purpose}, the sensor node that produces the data is responsible to send
+the raw data to a specified data storage target. Upon producing data, a sensor
+network node is specified to deal with the produced data where, in most cases,
+depending on the sensor network infrastructure used \cite{sn-storage03} (see
+section \ref{sec:sn-infrastructure}). In such a way, different storage
+mechanisms are provided:
 
@@
   another sensor node that contains an external storage.
-  \item \textbf{Local Storage}: as it is the case of real-time data stream, where data
-  is allocated on the sensor device's main read-only memory.
+  \item \textbf{Local Storage}: as it is the case of real-time data stream,
+  where data is allocated on the sensor device's main read-only memory.
 \end{itemize}
@@
 It is clear that the data flows from the sensor node to another depending on
-the specification of the network. On the whole, the sensor network
-infrastructure dictates where the data will flow to, as described in section
+the definition of the network. On the whole, the sensor network infrastructure
+dictates where the data will flow to, as described in section
 \ref{sec:sn-infrastructure}. \cite{sn-storage02} sees a wireless sensor
-networks as a many-to-one data gathering pattern, when its organization is
+networks as a many-to-one data collection pattern, when its organization is
 based on single-hop or star. On the other hand, when the data organization is
@@
 node to another until it reaches its final storage node as described by
-\cite{sn-storage01} \cite{sn-storage03}.
+\cite{sn-storage01}\cite{sn-storage03}.
 
 Last, but not least, yet a different storage strategy can be used: Data-Centric
-Storage. This strategy aims at clustering the collected data based on a given nature of the
-data such as location. Regardless the data storage mechanism used, the query
-processing will depend on the sensor network infrastructure and data location.
+Storage. This strategy aims at clustering the collected data based on a given
+nature of the data such as location. Regardless the data storage mechanism
+used, the query processing will depend on the sensor network infrastructure and
+data location.
 
@@
 \begin{itemize}
-  \item \textbf{In-Network Query Processing}: sensor nodes that provides
-  real-time data stream usually replies to requests to properties. Likewise, when the
-  data is used for archival purpose, the data can be located at a network node
-  specialized in storing the neighbors collected data. In this way, when a
-  query is issued to the sensor network, the query processor will
-  collectively verify every storage node for the collected data.
-  \item \textbf{Centralized Query Processing}: a main network sink receives the
-  collected data from the peer nodes. In this case, the query processor queries
-  the solo centralized node for the overall values of the participating sensor
-  devices.
+  \item \textbf{In-Network Query Processing}: sensor nodes that provide
+  real-time data stream usually replies to requests to properties.
+  Likewise, when the data is used for archival purpose, the data can be
+  located at a network node specialized in storing the neighbors
+  collected data. In this way, when a query is issued to the sensor
+  network, the query processor will collectively verify every storage
+  node for the collected data.
+  \item \textbf{Centralized Query Processing}: a leading network sink receives
+  the collected data from the peer nodes. In this case, the query processor
+  queries the solo centralized node for the global values of the
+  participating sensor devices.
 \end{itemize}
@@
 which are also related to the infrastructure, sensor network nature and
-requirements. The main goal of the in-network query processing is to decrease
-the amount of data transmitted from the sensor devices to the final data sink, as
-sensor devices are usually limited by constraints such as energy utilization 
-\cite{sn-storage03}. Along with this strategy, the literature describes
-different proposals that aims at upgrading the Structured Query Language (SQL)
-adding location-sensitive clauses to the language \cite{sn-db-newop}. However,
-this main problem related to this approach is that it may potentially flood
-all nodes with query messages \cite{sn-storage04}.
+requirements. The primary purpose of the in-network query processing is to
+minimize the number of data transmitted from the sensor devices to the final
+data sink, as sensor devices are generally limited by constraints such as
+energy utilization \cite{sn-storage03}. Along with this strategy, the literature
+describes different proposals that aim at upgrading the Structured Query
+Language (SQL) adding location-sensitive clauses to the language \cite{sn-db-newop}.
+However, this crucial problem related to this method is that it may potentially
+flood all nodes with query messages \cite{sn-storage04}.
 
@@
 processing, the use of a Data-Centric storage strategy has been shown to be a
-proven alternative for the sake of energy consumption \cite{sn-storage03}
-\cite{sn-storage01}. Its main goal is to classify the collected data based in a
-specific property, say location or event, in order to be sent to specific
-storage nodes that relates to the observation. As the main outcome of this
-strategy is to create network partitioning strategy, the query processor would
-hit specific sensor nodes with the needed data, and therefore, would
-decrease the load and maximize the capacity on the actual sensor devices.
+proven alternative for the purpose of energy consumption
+\cite{sn-storage03}\cite{sn-storage01}.Its primary goal is to organize the
+collected data based on a particular property, say location or event, in order
+to be sent to individual storage nodes that relate to the observation. As the
+main outcome of this strategy is to create network partitioning strategy, the
+query processor would hit specific sensor nodes with the needed data and,
+therefore, would decrease the load and maximize the capacity on the actual
+sensor devices.
 
 A centralized network query processing processes data in a network node
-specified as the network sink. The network sink usually is a network
-node with a database server with the collection of data delivered by
-all the sensor devices from the network. As a consequence of this
-many-to-one communication strategy \cite{sn-storage02}, the query process does
-not affect any of the sensor nodes in the network and, thus, maximizes the
-sensor device utilization. Moreover, the use of the data for historical
-purposes is easier to manage and give access to the primary end users of the
-data, following the process of data processing, compression, aggregation, etc
+specified as the network sink. The network sink usually is a network node with
+a database server with the collection of data delivered by all the sensor
+devices from the network. As a consequence of this many-to-one communication
+strategy \cite{sn-storage02}, the query process does not affect any of the
+sensor nodes in the network and, thus, maximizes the sensor device
+utilization. Moreover, the use of the data for historical purposes is easier
+to manage and provide access to the primary users of the data, following the
+process of data processing, compression, aggregation, etc
 \cite{sn-db-modeling02}.
 
-Although this approach seems to be convenient to the point of view of the
+Although this approach seemed to be appropriate to the point of view of the
 sensor nodes and used as the tradeoff between achieving capacity and energy
-consumption, the centralized data collector suffers from the annavoiable
-creation of a point of traffic concentration \cite{sn-storage02}, also referred
-to as the phenomenon of congestion, or funneling effect \cite{sn-storage04}.
-For this reason, simple distributed system techniques such as the use of
-master-slave or data replication can be used to decrease the overhead.
+consumption, the centralized data collector may suffer from the unavoidable
+development of a point of traffic concentration \cite{sn-storage02}, also
+referred to as the phenomenon of congestion, or funneling effect
+\cite{sn-storage04}. For this reason, simple distributed system techniques
+such as the use of master-slave or data replication can be used to reduce the
+overhead.
 
@@
 Different approaches are used to represent the collected data from sensors. The
-use of simple data descriptors have been used in research groups which
-expertise is not database systems, while others use such technology and the way
-to represent data using more robust approaches. \cite{sn-data-model-survey}
-have shown different data models and the advantages and desadvantages. The
-following is a summary of these approaches used by the sensor network
+use of basic data descriptors has been used in research groups which expertise
+is not database systems, while others use such technology and the way to
+represent data using more pragmatic approaches. \cite{sn-data-model-survey}
+have shown different data models and the advantages and disadvantages. The 
+following is a overview of these approaches used by the sensor network
 community.
@@
 \begin{itemize}
-  \item \textbf{Tabular Data Model}: comma-separeted values are very popular
-  due to its simplicity. Although data is not index, it is an easy approach for data
- exchange among researchers.
+  \item \textbf{Tabular Data Model}: comma-separated values are immensely
+  popular due to its simplicity. Although data is not index , it is an easy
+  approach for data exchange among researchers \cite{sn-provenance};
   \item \textbf{Relational Data Model}: it is used in the majority of the
@@
   its requirements \cite{sn-db-tinydb};
-  \item \textbf{Structured Data Model}: simiar to the relational model,
+  \item \textbf{Structured Data Model}: similar to the relational model,
   structured data models are becoming popular due to the Internet. The use of
-  XML data \cite{xml} is not yet popular, but there are a few projects using it
-  \cite{sn-xml-usage01}\cite{sn-xml-usage01}, implementing its query by using a
-  middleware \cite{sn-xml-middleware} \cite{sn-xml-query-engines}.
-  \item \textbf{Key-Value Pair and Document-Oriented Data Model}: as
-  an emerging data model, the Key-Value Pair (KVP) \cite{db-kvp} data model is
-  becoming popular with the up-coming trend of cloud computing. Although there are no
-  records of the use of this model data model in sensor networks, this work
-  will investigate it.
+  XML data \cite{xml} is not yet popular, but there are a few projects using
+  it \cite{sn-xml-usage01}\cite{sn-xml-usage02}, implementing its query by
+  using a middleware \cite{sn-xml-middleware}\cite{sn-xml-query-engines}.
 \end{itemize}
 
-Most importantly, the use of Data Provenance approaches to describe the
-collected data is shown to be important \cite{sn-provenance}, given the nature
-of the collected data. Therefore, this section describes these approaches for
-modeling data collected from sensor devices, taking into account the need of
-metadata for the data.
+Although these approaches have been used in Sensor Networks, this work will
+also take a look at others still not used, as follows:
+
+\begin{itemize}
+  \item \textbf{Key-Value Pair Model}: as an emerging data model, the Key-Value
+  Pair (KVP) \cite{db-kvp} data model is becoming popular with the up-coming
+  movement of cloud computing. Although there are no records of the use of this
+  model data model in sensor networks, this work will consider it.
+\end{itemize}
 
@@
 
-Simple use of Comma-Separated Values is an example os sensor network data. Data
+Simple use of Comma-Separated Values is an example of sensor network data. Data
 is structured by the top header, and the values are in each line of the file.
@@
   OPenDAP \cite{opendap} format, used by the SF-BEAMS project
-  \cite{sfbeams2006}.
+  \cite{sfbeams2006};
   \item data is queried using a text editor file, without any indexing
@@
 \begin{itemize}
-  \item data is structured in tables, and sensor properties are represented 
- by the columns of these tables. Data instances are represented by tuples or
- rows of these tables. Finally, table relationships are implemented based
- on the relational algebra abstraction, which is the mechanism used to 
-describe the relationship among different data types.
-  \item data is queried using the Structured Query Languange (SQL), and uses
-  the definitions of the tables and columns to extract data.
+  \item data is structured in tables, and sensor properties are represented by
+  the columns of these tables. Data instances are represented by tuples or rows
+  of these tables. Finally, table relationships are implemented based on the
+  relational algebra theory, which is the mechanism used to describe the
+  relationship among different data types.
+  \item data is queried using the Structured Query Languange (SQL) and uses the
+  definitions of the tables and columns to extract data.
 \end{itemize}
 
-Before the conception of such model, a database normalization process is
-necessary to describe the data in an appropriate way. An example of such use
-of the relational model is the TinyDB database systems \cite{sn-db-tinydb},
-which uses the relational model to represent the types of sensors. Due to the
-popularity of relational models, this is a popular choice since of the vast
-availability of research on relational database models.
+Before the creation of such a model, a database normalization process is
+necessary to describe the data in a convenient way. An example of such use of
+the relational model is the TinyDB database systems \cite{sn-db-tinydb}, which
+uses the relational model to represent the types of sensors. This is a popular
+choice because of the vast availability of research on relational database
+models.
 
@@
 related to this approach. One such problem is that the relational model is not
-suited for modeling time series data and metadata related to identify, time and
-location. For the problems related to it, came the implementation of different
-approaches that adds semantics to data types and different query clauses to the
-Structured Query Language (SQL) to solve join \cite{sn-db-newop}.
+suited for modeling time series data and metadata related to identify, time
+and location. For problems related to it, came the implementation of different
+approaches that add semantics to data types and different query clauses to the
+Structured Query Language (SQL) to solve join operations \cite{sn-db-newop}.
 
@@
 \begin{itemize}
-  \item data is structured in the collected data properties can be represented through the
-XML \cite{xml} instances by using tags, having its model validated by the use
-of XML Schema \cite{xml-schema}.
+  \item data is structured in the collected data properties can be represented
+  through the XML \cite{xml} instances by using tags, having its model validated
+  by the use of XML Schema \cite{xml-schema}.
   \item in order to extract data from XML-based databases, the use of XML XPath
-  is used to query XML documents.
+  \cite{xml-xpath} is used to query XML documents.
 \end{itemize}
@@
 Some examples of the application of XML in sensor networks in
-\cite{sn-xml-usage02} some have investigated the effecient use of XML in
+\cite{sn-xml-usage02} some have investigated the effective use of XML in
 sensor networks to represent the collected data \cite{sn-xml-usage01}. An
-example of querying xml data in sensor networks described by
+example of querying XML data in sensor networks described by
 \cite{sn-xml-query-engines}, while others use XML as a means of data exchange
-by implementing of middlewares to query and extract query data
+by implementing of middleware to query and extract query data
 \cite{sn-xml-middleware}. NetBEAMS is one of such sensor network that contains
@@
 
-\subsection{Key-Value Pair and Document-Oriented Data Models}
+\subsection{Key-Value Pair Data Models}
 
-The advance of distributed systems and the Internet have enabled the
-development of powerful database systems that can be organized in the context
-of database systems and how to represent models. An important up-coming
-approach is the use of database systems that does not use structured language
-for quering its data. One such example is called Key-Value Pair (KVP) data model
-\cite{db-kvp}. Such model has the following properties:
+The growth of distributed systems and the Internet have enabled the development
+of powerful database servers that can be organized in the context of
+infrastructure and how to represent models. An powerful up-coming approach is
+the use of database systems that does not use structured language for querying
+its data. One such example is called Key-Value Pair (KVP) data model
+\cite{db-kvp}, which is also known as Document-Oriented, Distributed Hash
+Table. Such a model has the following properties:
 
@@
   hash data structure. The definition of the key is a given property with its
-  associating value.
+  associating value;
   \item the query process is using mechanics similar to programming or
@@
 where data is modelled using structured documents. However, these documents
-have a dynamic structure that can be freely described without the use of an
-schema validation mechanism as used in XML Schema. One example of such
-approach is the use of the JSON data format \cite{json} in the implementation
-of database systems which are becoming popular with the new trend called Cloud
-Computing \cite{cloud-comp-architectures}. Databases implementing this strategy
-is the mongoDB \cite{mongodb} and CouchDB \cite{mongodb}.
+have a dynamic structure that can be freely described without the use of a
+schema validation mechanism as used in XML Schema. One example of such approach
+is the use of the JSON data format \cite{json} in the implementation of
+database systems which are becoming popular with the new trend called Cloud
+Computing \cite{cloud-comp-architectures}. Databases implementing this
+strategy is the mongoDB \cite{mongodb} and CouchDB \cite{couchdb}.
 
-\subsection{Data Provenance for Sensor Networks}
+\section{Data Provenance: How to describe collected data?}
+\label{sec:sn-provenance}
 
 Most of sensor networks implementations use the collected data from the sensor
-devices after they observe the environment. As shown by
-\cite{sn-data-model-survey}, different data models contains its pros and
-cons to represent data. However, one important aspect in sensor networks is the
+devices after they perceive the environment. As shown by
+\cite{sn-data-model-survey}, different data models contains their pros and
+cons to represent data. However, one crucial feature in sensor networks is the
 description of it collected data, as \cite{sn-provenance} discussed the
@@
 described and indexed, and this process can be related to Data Provenance
-\cite{db-provenance}.
+\cite{db-provenance}. Needless to say, Data Provenance is the
+Metadata\footnote{Data about data} about the collected sensor data.
 
@@
 For this reason, the collected data needs descriptions or annotations that can
-give enough information to track changes, that is, the history of how and when
-a given data has changed. The following should be the foundation of a
-provenance-aware collected data according to \cite{sn-provenance}:
+provide enough information to track changes, that is, the history of how and
+when a given data has changed. The following should be the essence of a
+provenance-aware collected data according to \cite{sn-provenance}.
 
 \begin{itemize}
-  \item each low-level observation as it describes the actual data;
-  \item the location where the collected data was generated;
-  \item the time when the data was collected.
+  \item \textbf{what was collected?}: it can be merely one or more low-level
+  observations,  which is a set of values with or without a description;
+  \item \textbf{When was the data collected?}: is the time when the data was
+  collected,  as most sensor data must be used depending on its existence;
+  \item \textbf{Where was the data collected?}: it is typically the value of 
+  the of the location from where the data was collected. However, it is vital to
+  note that the use of this metadata depends on the applicability of the data.
 \end{itemize}
 
-While Data Provenance provides necessary mechanisms to describe sensor networks
+While Data Provenance provides basic mechanisms to describe sensor networks
 data in general, others see the importance of the time space of sensor
 networks. \cite{sn-time-series} describes strategies to used to cluster
-distributed sensor data as time series of data.
+distributed sensor data as time series of data, as well as to decrease the
+size of data streams being transmitted in real-time sensor networks when these
+collected data does not need to be transferred \cite{sn-data-reduction}.
+
+Finally, an indispensable application of collected sensor data is the so-called
+blackbox of an aircraft, which carries the collection of raw data collected
+from a group of sensors placed in different places of the aircraft, as
+described by \cite{sn-exemple-blackbox}. The author describes the data is used
+in a process of data recovery and analysis in order to identify the probable
+reason of an aircraft accident, by using the time, and location of the
+aircraft, along with the independent variables from the sensors.
+
+\subsection{What was collected?}
+
+Different sensor devices will produce different attribute sets regarding a
+particular observation. In general, these data sets does not have distinct
+names, which leads to difficulties when indices are required in order to make
+the data searchable. Specially for the end of data archival, as described
+earlier in this chapter, the collected data needs to receive and identity,
+which identifies the data by itself in a regular database
+\cite{relational-model}. Equally, \cite{sn-provenance} described that
+traditional sensor networks uses the collected data for analysis long after
+the attribute set has been collected, usually before it is categorized and
+indexed. Annotations and tags are also names given by \cite{sn-provenance} as
+means to describe collected sensor data.
+
+In a nutshell, data is described by the name of the observation, when that is
+not provided. For example,
+
+\begin{itemize}
+  \item \textbf{temperature} = 73.45: this is the value of the collected
+  observation, which in general, the attribute name ``temperature" is not
+  provided in the data stream;
+  \item \textbf{temperature-scale} = ``Fahrenheit": the metadata regarding the
+  temperature attribute, which describes the scale of the temperature;
+  \item \textbf{device-version} = 5: the metadata that describes the version of
+  the sensor device, since sensors can be replaced, and as a consequence, data
+  accuracy may be as well be changed.
+\end{itemize}
+
+\subsection{When was the data collected?}
+
+In order to use historical sensor data, the use of a point in time is the most
+prominent feature of data, as it is considered a pivotal metadata for different
+types of sensor networks such as an aircraft the blackbox
+\cite{sn-exemple-blackbox}. Other projects classify data in time-series
+\cite{sn-time-series-example}, while others provide approaches to decrease
+the creation of data based on clustering data before they are sent to the sink
+\cite{sn-time-series}.
+
+In different areas, temporal databases or time-series are the ones that take
+the real-world time into consideration. In general, data in a temporal database
+can be classified by its dimension, as surveyed by \cite{db-temporal}:
+
+\begin{itemize}
+  \item \textbf{Valid time}: it is the point in time during which a fact
+  happened with respect to the real world. This value is only used when data is
+  captured, and cannot be changed because of its nature, as well as placing a
+  restriction over Update operations;
+  \item \textbf{Transaction time}: is the point in time at which a fact is
+  collected, transferred or processed in the database. This value is used
+  during Insertion operations, as well as in Update ones.
+\end{itemize}
+
+When both temporal dimensions are applied, the data is considered to be
+Bitemporal when \cite{db-temporal}. Usually, the data type of the valid and
+transaction time columns are straightforward timestamps, with any supported
+precision. For example, the following set of attribute and values shows a
+temperature of 70 degrees that was observed on ``November 15, 2009'', but
+collected only 3 days later.
+
+\begin{itemize}
+  \item \textbf{temperature} = 70;
+  \item \textbf{fact-time} = ``Nov 15, 2009 10:45 PM PST''
+  \item \textbf{transaction-time} = ``Nov 18, 2009 11:11 PM PST''
+\end{itemize}
+
+When it comes to Data Provenance in sensor networks, \cite{sn-provenance}
+shows that the collected data for archival purpose is usually saved in 
+self-described flat-files by file-names, containing complex naming conventions
+that contain the timestamp information about the data. For this reason, time
+is a central metadata needed in any data model used, since it helps
+identifying data uniquely for a given data set.
+
+\subsection{Where was the data collected from?}
+
+Depending on the application of the collected sensor data, the location in
+which the data was collected may be needed. For example, the geographic
+location of the collected data is critical to identify the specific
+GPS\footnote{Global Positioning System} coordinates of the sensor device
+\cite{sn-ex02}. Similarly, the position of any  event captured in a black-box
+of an aircraft carries is also used as metadata for the event. It will either
+depend on the sensor device provide such sampling attribute, or entered
+manually on the database.
+
+A correlation between time and location is presented by \cite{sn-time-series}
+as a noteworthy feature to reduce the traffic generated in a sensor network by
+clustering similar data when the sensor devices are placed in concentrated environment.

### 2009-11-24 docs/marcello-dissertation/thesis/chapter10-appendix.tex
Updates to chapter 2 through 7.
Chapter 2 lost the KVP section, since this approach had not been identified in
the literature for sensor networks. In this way, this makes the primary finding
and contribution of the dissertation. It will only appear during the technology
selection, what is the direct representation of what had happened;

chapter 3 adds details to the taxonomies;

Chapter 4 was cut short to describe SF-BEAMS, NetBEAMS in separate, as well as
keeps the entire implementation to the last chapter (appendix). 

Chapter 5 is the selection of the technology, but still misses highlights

@@
-\chapter{Appendix - Source Code / Artifacts}
+\chapter{Appendix}
 
@@
 
-The source code is located in the Subversion repository for the NetBEAMS
-project. This is how it is structured:
+\section{The Data Sensor Platform (DSP)}
+\label{sec:dsp-details}
 
-\section{DSP Data Persistence Component}
+The Data Sensor Platform, in short DSP Platform, is based on a micro-kernel
+architecture developed on top of the Java modular framework called OSGi
+\cite{osgi}. The platform represents the foundation of the execution of
+NetBEAMS, since it is executed from an embedded hardware called Gumstix
+\cite{gumstix} using a Linux operating system and on top of the Java Virtual
+Machine (JVM). In order to transport data from sensors hosts to the network
+sink, the use of a cellular connection is used to transport the collected data
+from the senso devices using the HTTP Protocol, as described by
+\citet{netbeams2009}. Figure \ref{fig:sf-netbeams-node} shows the
+architecture of the inhanced node designed by the NetBEAMS research group.
 
-This artifact defines the ANT build file.
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=0.5]{../diagrams/DSP-Gateway-Node}
+  \caption{The NetBEAMS Node enhancement with the Gumstix and the GSM Modem}
+  \label{fig:sf-netbeams-node}
+\end{figure}
+
+The following list describes each of the devices in the NetBEAMS Node.
+
+\begin{itemize}
+  \item \textbf{YSI 6600EDS V2}: the YSI device is the regular sonde
+  responsible for observing the environment. It is connected to the Gumstix
+  embedded system, where the DSP is executed;
+  \item \textbf{Gumstix console-vx}: it's a COTS ARM-based hardware that
+  provides a computer-on-module environment for the development of small
+  embedded systems. This is the main environment of data extraction and remote
+  processing is accomplished. It runs a cross-compiled version of Linux Kernel
+  2.6;
+  \item \textbf{Huawei E220 USB Modem}: it is used to transfer the data from the
+  Gumstix to the RTC Labs Data Center.
+\end{itemize}
+
+The DSP runs inside of the embedded system Gumstix. The following sections
+details its design and process for extracting data.
+
+\subsection{The DSP System Design}
+
+This section focus on the development of a Software Platform for the NetBEAMS
+Gateway Embedded System, or on the system developed in the Gumstix device. As
+shown in figure \ref{fig:netbeams-software-stack}, the NetBEAMS Gateway Node is
+an embedded system that contains a set of layer found in any computer.
+However, each of the resources described from the system were customized for
+the NetBEAMS system. The main components of such system can be summarized as
+follows:
+
+\begin{itemize}
+  \item \textbf{Operating System}: it uses a cross-compiled Gentoo Linux, which
+  supports a wide variety of development tools, including the Java Virtual Machine
+  (JVM), the underlying platform system;
+  \item \textbf{Java Virtual Machine}: it uses the JamVM, a ~200 KB version of
+  the Sun Microsystems Java Virtual Machine, version 2.0;
+  \item \textbf{OSGi Framework}: it uses the Knopflerfish implementation of the
+  OSGi 4.1 specification. More details in the following sections.
+  \item \textbf{DSP Platform and other Bundles}: The plug-and-play DSP
+  Components are based on the OSGi bundles capabilities, which can reuse services registered in the OSGi framework.
+\end{itemize}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=0.6]{../diagrams/NetBEAMS-Client-Node-Components}
+  \caption{The NetBEAMS's Gumstix Software Stack}
+  \label{fig:netbeams-software-stack}
+\end{figure}
+
+As depicted in figure \ref{fig:netbeams-software-stack}, the DSP Sonde Data
+Producer component opens a serial connection with the device through the Java
+Native Interface (JNI) driver and produces the data flow 1. The data stream
+format at this stage is as described on table \ref{tab:ysi-data-stream}.
+Later, by parsing and transforming the data to DSP internal representation, it
+sends the data to the DSP Platform as depicted on flow 2. Finally, the DSP
+Platform decides to send the data to the DSP Wire Transport Client, as shown
+on flow 3, by following rules defined by the system administrator. At this
+point, the data is ready to be transmitted to the server, which is covered on
+the following sections.
+
+\subsection{The DSP Platform Architecture}
+\label{netbeams-architecture}
+
+The architecture of the DSP, aligned with the OSGi capabilities, reuses the
+Service-Oriented Architecture that can easily enable and disable plug-and-play
+components. By following a publish-subscriber design-pattern \cite{gof}, the
+service producer registers into the service broker while the consumer looks
+for a given registered service in the Service Broker. For this reason, this
+approach allows decoupled interoperability among modules as it is shown in
+figure \ref{fig:publish-subscriber-pattern}.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=0.6]{../diagrams/publish-subscriber-pattern}
+  \caption{The publish-subscriber design-pattern for services}
+  \label{fig:publish-subscriber-pattern}
+\end{figure}
+
+The DSP infrastructure is based on a loosely-coupled architecture hosted in a
+given host computer with an OSGi container, where the entire system is
+composed by a set of plug-n-play OSGi modules. Each DSP component, as an
+instance of an OSGi Module, can perform specialized tasks according to its
+specification. In general, each DSP component can be categorized as follows:
+
+\begin{itemize}
+  \item \textbf{Data Producer (DP)}, whose main function is to produce data for
+  the system;
+  \item \textbf{Data Consumer (DC)}, whose main function is to consume the
+  received data.
+\end{itemize}
+
+\begin{figure}[!b]
+  \centering
+  \includegraphics[scale=0.7]{../diagrams/DSP-Producers-Consumers-Components-Interactions}
+  \caption{The basic architecture of a Data Sensor Platform (DSP)}
+  \label{fig:DSP-Producers-Consumers-Components-Interactions}
+\end{figure}
+
+The Data Producers and Data Consumers exchange data through the DSP Message
+Broker using a message unit called DSP Message. As it is depicted in figure
+\ref{fig:DSP-Producers-Consumers-Components-Interactions}, a Data Producer
+creates a set of one or more DSP Messages wrapped up in a DSP
+MessagesContainer in order to be sent to the DSP Message Broker. When the
+broker receives the set of DSP Messages, it requests a list of Data Consumers
+to each of the DSP Messages based on rules managed by the DSP Matcher.
+Whenever the broker acquires the list of Data Consumers, it delivers a copy of
+the DSP Message to each of the Data Consumers in the receiving list. Upon the
+DSP Message arrival, each Data Consumer can extract the actual data from the
+DSPMessage. Finally, although the components can be distinguished between Data
+Producers and Consumers, any of them are entitled to send and receive
+administrative DSP Messages, which is covered in the following sections.
+
+When the DSP Platform is installed into the OSGi framework, it follows the
+OSGi specifications without any changes to remain it in the installed state.
+In this way, it is important to note that the DSP Platform and its
+plug-and-play components are directly mapped with the OSGi Framework as
+follows:
+
+\begin{itemize}
+  \item \textbf{OSGi Bundle} = DSP Component Artifact in Jar format
+  \begin{itemize}
+    \item SondeDSPProducer.jar is the component developed for the YSI sonde
+  \end{itemize}  
+  \item \textbf{OSGi Activator} = DSP Component Activator
+  \begin{itemize}
+    \item Class specified in the MANIFEST.MF artifact (next sections)
+  \end{itemize} 
+  \item \textbf{OSGi Service Reference} = DSP Component Class
+  \begin{itemize}
+    \item  Class specified by the component author.
+  \end{itemize}
+\end{itemize}
+
+The DSP Platform, along with the DSP Framework, are the only DSP Components
+needed to be installed into the OSGi container. For this reason,the DSP
+Platform is responsible to install and activate the additional DSP components
+that makes part of the platform during the start up process, which leverages
+its bootstrap process. Nevertheless, in order to automate the process of
+automatic install and removal of the plug-and-play components, the DSP Platform
+provides a deployment configuration file that describes which DSP components
+must be installed with a given priority number. The name of this configuration
+artifact is called config.xml (see listing \ref{file:dsp-config.xml}), and is
+depicted by the XML Schema \cite{xml-schema} in figure
+\ref{fig:dsp-config-schema}.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=0.8]{../diagrams/config-schema}
+  \caption{XML Schema for the main DSP deployment configuration artifact}
+  \label{fig:dsp-config-schema}
+\end{figure}
+
+In order to show an example, consider the DSP Platform on the embedded system
+Gumstix in figure \ref{fig:sf-netbeams-node}. It is started by providing a
+deployment configuration artifact config.xml included in the appendix. It lists
+the components that the DSP Platform needs to install and load for its
+operation. As it is shown, the DSPWireTransportClient and the DSPSondeProducer
+are the identification of the components, followed by the name of the physical
+file and an assigned loading priority. While the DSPWireTransportClient is
+responsible for the serialization and deserialization of the DSP Messages
+exchanged by the DSPSondeProducer and any remote DSP component. In this case,
+the DSPSondeProducer is the DSP component responsible to collect measurement
+data from the YSI Sonde and temporarily store them before being transmitted.
+
+Therefore, in order to orchestrate the DSP Components together, the DSP
+Platform component is responsible for installing and uninstalling during its
+state "Starting" and "Stopping", respectively, as it is shown in figure
+\ref{fig:DSPPlatform-Install-Usage-State-Diagram}. When the system
+administrator starts the DSP Platform, it performs the regular OSGi procedures
+to start the bundle. Up until that state, it will get the list of DSP
+Components to be installed as described in the config.xml artifact, trying to
+install each of them in an optimistic way. After that, it will continue on the
+state "Active" until the system administrator decides to stop it.
+
+\begin{figure}[!b]
+  \centering
+  \includegraphics[scale=0.6]{../diagrams/DSPPlatform-Install-Usage-State-Diagram}
+  \caption{The DSP Platform State Diagram}
+  \label{fig:DSPPlatform-Install-Usage-State-Diagram}
+\end{figure}
+
+When the DSP Platform is on the Active state, it is ready to manage the DSP
+Messages exchanges as described through figure
+\ref{fig:DSP-Producers-Consumers-Components-Interactions}. The following
+sections describe the participating classes in starting the DSP Platform and
+its execution.
+
+In a nutshell, the DSP Platform is an extension of the OSGi Platform, as it can
+be seen in the UML Class Diagram \cite{uml} in figure
+\ref{fig:DSP-Platform-Class-Diagram-Simple}. The class Activator contains a
+unique instance of the Platform class, which contains instances of the classes
+ComponentManager, Matcher, the MessageBroker and DSPContextImpl. The
+DSPBundleController is the main instance that manages the plug-and-play DSP
+components, and therefore, is tightly-coupled with the OSGi Framework classes.
+Additionally, this class is part of the publish-subscriber design-pattern, as
+it it implements the interfaces Bundle Listener and Service Listener, whose
+responsibility is to track listens to the changes of the state of
+ServicesReferences on the DSP Platform.
+
+\begin{figure}[!t]
+  \centering
+  \includegraphics[scale=0.7]{../diagrams/DSP-Platform-Class-Diagram-Simple}
+  \caption{The DSP Platform UML Class Diagram}
+  \label{fig:DSP-Platform-Class-Diagram-Simple}
+\end{figure}
+
+\section{DSP Message and DSP Messages Container}
+\label{sec:dsp-message}
+
+As shown in the previous section, a DSP Message is the main unit of
+communication among DSP Components. In a nutshell, a DSP Message can be seen
+as an abstraction of an envelop containing the sections: a header and a body.
+The former is used for identification and routing purposes, while the latter is
+used to carry the payload of the message, that is, the collected data from
+a sensor. Figure \ref{fig:DSP-Message-Representation} depicts the abstraction
+of a DSP Message and its main components.
+
+\begin{figure}[!b]
+  \centering
+  \includegraphics[scale=0.8]{../diagrams/DSP-Message-Representation}
+  \caption{An Abstraction of a DSP Message and its components}
+  \label{fig:DSP-Message-Representation}
+\end{figure}
+
+\begin{itemize}
+  \item \textbf{Header} is composed by 2 blocks that identify the data
+  producer and consumer of the message, as well as the time of creation, UUID,
+  etc. It is an important piece of information for the data delivery and
+  routing purposes.
+  \item \textbf{Body} contains the payload of the message. This section carries
+  the collected data from a sensor.
+\end{itemize}
+
+In general, the DSP Platform offers a veriety of types of DSP Messages for
+different purposes. For example, any measurement data must be
+wrapped up in a Measurement Message, while a Query Message is used to exchange
+messages among the components for the purpose of management. In this way, the
+main DSP Messages can be summarized as follows:
+
+\begin{itemize}
+  \item \textbf{Measurement Message} is used to transport any sensor collected
+  data;
+  \item \textbf{Query Message} is used to query a DSP component about its
+  configuration properties;
+  \item \textbf{Update Message} used to update a DSP component's configuration
+  properties;
+  \item \textbf{Acknowledgement Message} is used for the transport communication
+  protocol, by acknowledging the reception of a DSP Message. More details in the
+  Remote Data delivery section.
+\end{itemize}
+
+These DSP Classes and other can be seen in figure the  UML Class
+Diagram \cite{uml} of figure \ref{fig:DSP-Messages-Classes}. Each DSP Message
+contains an instance of a Header and Body. Whenever a DSP component is ready
+to transmit messages, it wraps up the set of DSP Messages into the instance of
+the Messages Container, which contains information about the collection of
+messages being transmitted with its own identification. In this fashion, the
+Messages Container is the main communication unit between 2 different DSP
+Components.
+
+\begin{figure}[!t]
+  \centering
+  \includegraphics[scale=0.6]{../diagrams/DSP-Messages-Classes}
+  \caption{The DSP Messages UML Class Diagram}
+  \label{fig:DSP-Messages-Classes}
+\end{figure}
+
+Another use of the DSP Messages is during the DSP Platform activation process,
+where bootstrap messages are used to configure each DSP component. In order to
+send configuration parameters, the DSP Platform uses Update Messages using a
+unique type of Message Content called DSP Properties, which is described in the
+next section.
+
+\section{Collected Sensor Data as the DSP Message Content}
+
+As mentioned in the previsou section, a DSP Message can carry any data on its
+body, also called the payload. In such a way, any data representation must
+just extend from the Message Content abstract class as it is shown
+\ref{fig:Sonde-MessageContainer-Class-Diagram}, as defines the data
+representation of the YSI Sonde data.
+
+\begin{figure}[!h]
+  \centering
+  \includegraphics[scale=0.6]{../diagrams/Sonde-MessageContainer-Class-Diagram}
+  \caption{The YSI Payload - Message Content UML Class Diagram}
+  \label{fig:Sonde-MessageContainer-Class-Diagram}
+\end{figure}
+
+The properties measured from the YSI Sonde can be captured by the class
+SondeDataType as a direct render of the measurements data depicted in table
+\ref{tab:ysi-data-stream}. In addition to the regular data types, the class
+also provides additional properties such as the attribute "samplingTimeStamp"
+that carries the time when the data was collected from the YSI Sonde. However,
+in order to capture a set of measurements at once, the SondeDataContainer
+class used to carry a set of SondeDataType classes, as it is composed by one
+or more instances of SondeDataType. In order to illustrate other examples,
+consider the UML Class Diagram \cite{uml} of image
+\ref{fig:Mouse-Actions--MessageContainer-Class-Diagram} as a payload of
+observations made from a mouse over a screen, which captures the name of the
+event, the button name and the x/y coordinates.
+
+\begin{figure}[!h]
+  \centering
+  \includegraphics[scale=0.5]{../diagrams/Mouse-Actions--MessageContainer-Class-Diagram}
+  \caption{The Mouse Actions Payload - Message Content UML Class Diagram}
+  \label{fig:Mouse-Actions--MessageContainer-Class-Diagram}
+\end{figure}
+
+As described in the previous section, the DSP Platform uses DSP Update
+Messages to configure internal components' initial configuration values. In
+order to do so, the DSP Platform uses an instance of the DSP Properties, which
+is composed of a list of initial DSP Property and their relating Values
+instance depicted on the UML class diagram \cite{uml} of figure
+\ref{fig:DSP-Property-Class-Diagram}.
+
+\begin{itemize}
+  \item \textbf{Property Instance}: initial-rate
+  \item \textbf{Value Instance}: 40
+\end{itemize}
+
+\begin{figure}[!t]
+  \centering
+  \includegraphics[scale=0.6]{../diagrams/DSP-Property-Class-Diagram}
+  \caption{The DSP Property Payload - Message Content UML Class Diagram}
+  \label{fig:DSP-Property-Class-Diagram}
+\end{figure}
+
+\subsection{DSP Platform Initialization Process}
+
+All in all, the DSP Platform requires the knowledge of the aforementioned DSP
+classes in order to initialize. The UML Sequence Diagram \cite{uml} in figure
+\ref{fig:From-OSGi-DSP-Platform-Start-to-All-Components-Sequence-Diagram}
+depicts the exact moments when the additional the plug-and-play DSP components
+are installed and bootstrapped by the DSP Platform, as the UML State Diagram
+\cite{uml} in figure \ref{fig:DSPPlatform-Install-Usage-State-Diagram}
+described.
+
+\begin{itemize}
+  \item \textbf{Start DSP Activator Gate}: after installing the Activator
+  indicated by the DSP Bundle's MANIFEST.MF artifact, the OSGi Framework makes
+  a call to the method Activator.start();
+  \item \textbf{Bootstrap DSP Component Gate}: after the component has been
+  initialized (by the call the method init() from the class ComponentManager),
+  the DSP Component is ready to be bootstrapped, and the optional bootstrap
+  message is delivered.
+\end{itemize}
+
+\begin{figure}[!b]
+  \centering
+  \includegraphics[scale=0.5]{../diagrams/From-OSGi-DSP-Platform-Start-to-All-Components-Sequence-Diagram}
+  \caption{The DSP Platform Initialization UML Sequence Diagram}
+  \label{fig:From-OSGi-DSP-Platform-Start-to-All-Components-Sequence-Diagram}
+\end{figure}
+
+\newpage
+
+\subsection{Data Delivery Mechanism}
+
+In general, when a DSP Component finishes preparing the DSP Messages Container,
+it contacts an entity responsible for sending and proxying message to other
+components. The DSP Broker acts like a mailman, receiving messages and
+forwarding copies of it to other entities in the system, abstraction in which
+is shown in figure \ref{fig:DSP-Message-Broker-Abstraction}. In this way, the
+DSP Broker receives a DSP Message and makes necessary deliveries to DSP
+components that it has knowledge of.
+
+\begin{figure}[!h]
+  \centering
+  \includegraphics[scale=0.8]{../diagrams/DSP-Message-Broker-Abstraction}
+  \caption{A DSP Message Broker Abstraction}
+  \label{fig:DSP-Message-Broker-Abstraction}
+\end{figure}
+
+The use of messaging exchange paradigm is used in NetBEAMS because it decouples
+the system by using messages take into account their structure to filter the
+messages. In this way, the DSP Broker uses a filter component called DSP
+Matcher, which provides a list of DSP Components that needs to receive a copy
+of the DSP Message. It turns out that the DSP Matcher bases its decisions on
+structured rules provided by the system administrator for an specific DSP
+Platform instance. The UML Class diagram for the DSP Broker, Matcher and
+particiaping classes is depicted in figure
+\ref{fig:DSPBroker-Matcher-Class-Diagram}, and describes how the DSP Broker
+implementation is tightly coupled with the DSP Matcher, as well as to the
+lists of DSP Components sorted by name and type.
+
+\begin{figure}[!b]
+  \centering
+  \includegraphics[scale=0.6]{../diagrams/DSPBroker-Matcher-Class-Diagram}
+  \caption{The DSP Broker-Matcher UML Class Diagram}
+  \label{fig:DSPBroker-Matcher-Class-Diagram}
+\end{figure}
+
+In general, the DSP Matcher is composed by a list of MatchRule, which contains
+a reference to a MatchCriteria and a MatchTarget. The former identifies the
+rule for matching a given DSP Message based on the producer and consumer's
+message type and location, while the latter describes the DSP Component that
+will receive the DSP Message, and if its delivery is done by means of a gateway
+component. In this way, the DSP Matcher can be seen as a function that takes a
+DSP Message as an input and returns a list of DSP consumers, and the
+matching configuration artifact provided by the administrator. See listing
+\ref{file:dsp-matcher-config.xml}) for details.
+
+\begin{table}
+    \caption{DSP Matcher Rule Algorithm to filter Data Consumer components to
+    receive a given Message}
+    \begin{center}
+        \begin{tabular}{lr}
+          \textit{Data Consumers ( DSP Message )} := \\ 
+          Verify the DSP Message's Header + \\
+          Verify the matcher rules, which contains the list of consumers for
+          the given message
+        \end{tabular}
+    \end{center}
+    \label{tab:ysi-data-stream}
+\end{table}
+
+One example of such matching rule used by the DSP installed at the Gumstix can
+be seen in listing \ref{file:dsp-matcher-config-gumstix.xml}. the Gumstix
+system to configure the data delivery of the messages produced by the DSP
+Component DSP Sonde Producer. It describes a list of maching rules that
+reseambles a filtering system. This configuration contains 2 match rules: one
+related to internal DSP Configuration and another for a remote destination.
+The match criteria describes what the DSP Message Header properties must be,
+while the match target is the destination of the message. One important
+definition is the gateway Component Type, which defines the name of a remote
+DSP Component that is capable of receiving a remote DSP Message.
+
+Upon receiving all the matching rules and analyzing the DSP Message against
+the rules, the DSP Broker selects a set of the running DSP components to receive
+a copy of the DSP Message object in two different ways:
+
+\begin{itemize}
+  \item \textbf{In-memory local message delivery}: if the receiving DSP
+  Component is located in the current local host, a deep copy of the instance
+  of the DSP message is delivered;
+    \begin{itemize}[label=\textbullet]
+        \item IP Addresses are correctly resolved by using the Ethernet card in
+        the device: localhost, 127.0.0.1, or the same IP Address for Producer
+        and Receiver resolves into a local device. An example for such type of
+        delivery can be seen in the match ruleID
+        all messages sent to dsp manager of listing 
+        \ref{file:dsp-matcher-config-gumstix.xml}.
+    \end{itemize}
+  \item \textbf{Serialized remote message delivery}: if the receiving DSP
+  Component is located in a foreign/remote host, the message is serialized in a
+  format defined by the transport protocol chosen the DSP Component responsible
+  for the transport. The following section describes the existing DSP Data
+  Transport component;     
+  \begin{itemize}[label=\textbullet]
+        \item IP Address from the Producer and Consumer are different, and are
+        not resolved to be in the same host. An instance of such match rule can
+        be seen in the match rule ID send-message-to-remote url dsp of the code
+        snippet \ref{file:dsp-matcher-config-gumstix.xml}.
+    \end{itemize}
+\end{itemize}
+
+\subsection{Sending DSP Messages through the Wire}
+
+In order to send messages through the wire, the DSP Platform uses two symmetric
+components that serialize and deserialize DSP Messages to XML and back to Java
+POJO, respectively, and vice-versa. Named DSP Wire Transport Client and DSP
+Wire Transport Server \cite{netbeams2009}, these components use the HTTP
+\cite{http} protocol to transport the serialized version of the DSP Messages.
+Figure \ref{fig:DSP-to-DSP-Remote-Communication} depicts the scenario where a
+DSP Sensor Node transmits a DSP Message through an HTTP POST request.
+
+\begin{figure}[!b]
+  \centering
+  \includegraphics[scale=0.6]{../diagrams/DSP-to-DSP-Remote-Communication}
+  \caption{The remote communication between two DSP Components using the HTTP
+  Protocol}
+  \label{fig:DSP-to-DSP-Remote-Communication}
+\end{figure}
+
+\begin{itemize}
+  \item \textbf{DSP Wire Transport Client}: responsible for making HTTP POST
+  Requests to the service provided by the DSP Wire Transport Server component.
+  The body of the HTTP POST request contains a serialized version of an
+  instance of a DSP MessagesContainer in XML format;
+  \item \textbf{DSP Wire Transport Server}: exposes a Web Server listening to
+  port 8080 that receives the HTTP POST request from the Client. Upon receiving
+  the request, the server deserializes the DSP MessagesContainer back to the
+  in-memory POJO and proceeds with the delivery to the DSP Broker. At this
+  situation, the component packages acknowledgement messages and any queued
+  message to the client and piggybacks \cite{xml-piggybacking} it to the HTTP
+  response body.
+\end{itemize}
+
+As described by \cite{netbeams2009}, NetBEAMS is a single-hop sensor network
+whose sensors transmits data from the sensors to the server. Upon receiving the
+data stream, the DSP Wire Transport Server component desirializes the DSP
+MessagesContainer and its enclosing DSP Messages back to a Java class instance
+and delivers each of the DSP Messages to the DSP Broker, which delivers a
+copy of the message to each DSP Component listed by the maching rule. 
+%This
+%process is depicted by the UML Sequence Diagram \cite{uml} in figure \ref{fig:}
+
+Whenever a DSP bundle producer is ready to transmit messages, they are added
+into a DSP Messages Container (see UML Class Diagram in figure
+\ref{fig:DSP-Messages-Classes}), which is serialized in XML to be transmitted.
+Listing \ref{file:dsp-message-serialized-ysi} shows an instance of a DSP
+MessagesContainer in serialized in XML, ready for transmission. In fact, the
+listing is an example of the transmitted data collected from a sensor located
+at the 192.168.0.103 to be transmitted to the host 192.168.0.106, and the
+process of transmission in the server-side can be followed in figure
+\ref{fig:From-DSPSondeComponent-to-DSPWireTransportClient-Sequence}.
+
+\begin{figure}[!h]
+  \centering
+  \includegraphics[scale=0.5]{../diagrams/From-DSPSondeComponent-to-DSPWireTransportClient-Sequence}
+  \caption{The UML Sequence Diagram for the production of a YSI Message}
+  \label{fig:From-DSPSondeComponent-to-DSPWireTransportClient-Sequence}
+\end{figure}
+
+The Sonde DSP Component is responsible for starting a worker thread called
+SondeProducer, which is responsible for constantly read from the RS-232 serial
+port by using an instance of the SerialReader class. This process depicts the
+production of the data flow "flow 1" of figure
+\ref{fig:netbeams-software-stack}, which is a collection of SondeDataType for
+each of the measurement data read on a SondeDataContainer. As depicted in
+figure \ref{fig:From-DSPSondeComponent-to-DSPWireTransportClient-Sequence}, it
+first read the serial connection using the class SerialReader, which parses 
+the data stream. After parsing the data, the reader creates an instance of the
+class SondeDataType, the POJO responsible for carrying out the measurement
+data. As described earlier, the data is added into the SondeDataContainer to
+dispatched to the DSP Broker, whose responsibility is to get the list of DSP
+Components that is setup to receive a copy of the DSP message (see previous
+code snippet). As for the setup done for the Gumstix, the measurement message
+is supposed to be sent to a remote destination of host 192.168.0.7. At this
+point, the DSP Broker makes a decision to send the DSP Message to the default
+gateway component DSPWireTransportClient.
+
+After the message arrives in the DSPWireTransportClient, the message is
+maintained in a temporary repository of outbound DSP Messages called
+MessagesQueues after the message is hashed by the destination IP address and
+the DSP Client sender using an instance of QueueMessageData. As shown in figure
+\ref{fig:From-DSPWireTransport-Client-to-DSPWireTransportServer-Sequence}, the
+DSP TransportSender is a worker thread that is started by the DSP Wire
+Transport Client and is constantly transmits the data in the outbound queues
+after a configurable rate. It makes a transport of the message and receives
+the response, updates the state of the messages to transmitted, to finally
+send the response messages to the DSP Message Broker.
+
+\begin{figure}[!t]
+  \centering
+  \includegraphics[scale=0.5]{../diagrams/From-DSPWireTransport-Client-to-DSPWireTransportServer-Sequence}
+  \caption{The UML Sequence Diagram describing the Wire Transport Server
+  receiving a DSP Message}
+  \label{fig:From-DSPWireTransport-Client-to-DSPWireTransportServer-Sequence}
+\end{figure}
+
+While the DSPWireTransportClient makes an HTTP Request with the DSP Message,
+the DSPWireTransportServer's servlet receives the request. It first
+deserializes the MessagesContainer and delivers each of its messages to the DSP
+Broker. At this point, the server verifies if there are messages in the
+outbound queue that are addressed for the host that requested the messages
+delivery. By using a technique called piggybacking, the server adds to the HTTP
+Response a serialized version of the MessagesContainer with any DSP Messages.
+In addition to any DSP Message, the server prepares and sends acknowledgement
+messages back to the requesting host to the host.
+
+\begin{figure}[!t]
+  \centering
+  \includegraphics[scale=0.5]{../diagrams/From-DSPWireTransport-Server-To-DSP-Broker}
+  \caption{The UML Sequence diagram describing the DSP
+Message delivery on the server-side and the response}
+  \label{fig:From-DSPWireTransport-Server-To-DSP-Broker}
+\end{figure}
+
+The gate "From DSP WireTransport Server to Broker Gate" defines the actual
+delivery of the DSP Message to the DSP Broker. As this section is a tutorial
+about the DSP Platform, the remaining steps will be explained on Section 3,
+while showing the design and implementation of the component DSP Data
+Persistence.
+
+\section{Source-Code and Implemented Artifacts}
 
@@
 
-%\lstinputlisting[language=bash,label=file:osgi-manifest,caption=OSGi
-%Manifest
-%Descriptor]{../../../../netbeams/versions/v2/apps/osgi-bundles/dsp/DSPDataPersistence/META-INF/MANIFEST.MF}
+\lstinputlisting[label=file:osgi-manifest,caption=OSGi Manifest
+Descriptor]{../../../../netbeams/versions/v2/apps/osgi-bundles/dsp/DSPDataPersistence/META-INF/MANIFEST.MF}
 
@@
 Gumstix]{../../../../netbeams/versions/v2/config/deployment/matcher_config-gumstix.xml}
+
+\lstinputlisting[language=XML,label=file:dsp-data-persistence-bootstrap.xml,caption=DSP
+Message used to bootstrap the DSP Data Persistence Component]
+{../../../../netbeams/versions/v2/apps/xml/samples/DSPDataPersistence_001.xml}
 

### 2009-11-24 docs/marcello-dissertation/thesis/chapter2-state-of-the-art.tex
Updates to chapter 2 through 7.
Chapter 2 lost the KVP section, since this approach had not been identified in
the literature for sensor networks. In this way, this makes the primary finding
and contribution of the dissertation. It will only appear during the technology
selection, what is the direct representation of what had happened;

chapter 3 adds details to the taxonomies;

Chapter 4 was cut short to describe SF-BEAMS, NetBEAMS in separate, as well as
keeps the entire implementation to the last chapter (appendix). 

Chapter 5 is the selection of the technology, but still misses highlights

@@
 
-Although these approaches have been used in Sensor Networks, this work will
-also take a look at others still not used, as follows:
-
-\begin{itemize}
-  \item \textbf{Key-Value Pair Model}: as an emerging data model, the Key-Value
-  Pair (KVP) \cite{db-kvp} data model is becoming popular with the up-coming
-  movement of cloud computing. Although there are no records of the use of this
-  model data model in sensor networks, this work will consider it.
-\end{itemize}
-
 \subsection{Tabular Data Model and Query}
@@
 \cite{netbeams2009}.
-
-\subsection{Key-Value Pair Data Models}
-
-The growth of distributed systems and the Internet have enabled the development
-of powerful database servers that can be organized in the context of
-infrastructure and how to represent models. An powerful up-coming approach is
-the use of database systems that does not use structured language for querying
-its data. One such example is called Key-Value Pair (KVP) data model
-\cite{db-kvp}, which is also known as Document-Oriented, Distributed Hash
-Table. Such a model has the following properties:
-
-\begin{itemize}
-  \item data is structured in collections of key-value pairs, as it is done in
-  hash data structure. The definition of the key is a given property with its
-  associating value;
-  \item the query process is using mechanics similar to programming or
-  scripting language, that is, the use of APIs in a given programming language.
-\end{itemize}
-
-Different variants of such data model is the document-oriented data model,
-where data is modelled using structured documents. However, these documents
-have a dynamic structure that can be freely described without the use of a
-schema validation mechanism as used in XML Schema. One example of such approach
-is the use of the JSON data format \cite{json} in the implementation of
-database systems which are becoming popular with the new trend called Cloud
-Computing \cite{cloud-comp-architectures}. Databases implementing this
-strategy is the mongoDB \cite{mongodb} and CouchDB \cite{couchdb}.
 

### 2009-11-24 docs/marcello-dissertation/thesis/chapter3-taxonomic-classification-of-persistence-storage-for-sensor-networks.tex
Updates to chapter 2 through 7.
Chapter 2 lost the KVP section, since this approach had not been identified in
the literature for sensor networks. In this way, this makes the primary finding
and contribution of the dissertation. It will only appear during the technology
selection, what is the direct representation of what had happened;

chapter 3 adds details to the taxonomies;

Chapter 4 was cut short to describe SF-BEAMS, NetBEAMS in separate, as well as
keeps the entire implementation to the last chapter (appendix). 

Chapter 5 is the selection of the technology, but still misses highlights

@@
 
-\chapter{Taxonomic Classification of Persistence Storage Systems for Sensor
-Networks}
+\chapter{Taxonomic Classification of Data Persistence in Sensor Networks}
 \label{chap:taxonomies}
@@
 
-\section{Nature of Sensor Data Taxonomy}
+\section{Purpose of the Sensor Data Taxonomy}
 
@@
 and how the collected data is going to be used as described in section
-\ref{sec:sn-data-purpose}. In addition to it, the mechanism for data storage
-described in section \ref{sec:sn-infrastructure} can be related to each of
-them. The related taxon is depicted by figure \ref{fig:taxonomy-data-purpose}.
+\ref{sec:sn-data-purpose}.
 
@@
   \includegraphics{../diagrams/taxonomy-data-purpose}
-  \caption{The Nature and Location of Sensor Data Taxonomy}
+  \caption{The Purpose of Sensor Data Taxonomy}
   \label{fig:taxonomy-data-purpose}
@@
 
-\section{Location of Sensor Data Taxonomy}
+\section{The Location of the Sensor Data Taxonomy}
 
-No matter the type of resource management used, the first taxonomy presented is
-related to the purpose of the use of the collected sensor data. That is, when
-and how the collected data is going to be used as described in section
-\ref{sec:sn-data-purpose}. In addition to it, the mechanism for data storage
-described in section \ref{sec:sn-infrastructure} can be related to each of
-them. The related taxon is depicted by figure \ref{fig:taxonomy-data-location}.
+Directly related to the purpose of data, the mechanism for data storage
+described in section \ref{sec:sn-infrastructure} depends on the purpose of
+the use of the collected data. The related taxon is depicted by figure
+\ref{fig:taxonomy-data-location}, and described as follows:
 
@@
   \includegraphics{../diagrams/taxonomy-data-location}
-  \caption{The Nature and Location of Sensor Data Taxonomy}
+  \caption{The Location of Sensor Data Taxonomy}
   \label{fig:taxonomy-data-location}
@@
 One of the major challenges in the development of a persistence storage for
-sensor network is related to the persistence model chosen. The addition of new
-sensor devices may trigger refactoring of the existing model or not, depending
-on how the data model accommodates the data. Different data models are reported
-to be used in the literature, as described in section \ref{sec:data-models}.
-For this reason, taxonomy related to the data model depicts two different
-classes of data models, as shown in figure \ref{fig:taxonomy-data-model}.
+sensor network is related to the data model chosen to represent the collected
+data. The addition of new sensor devices may trigger refactoring of the
+existing model or not, depending on how the data model accommodates the data.
+Different data models are reported to be used in the literature, as described
+in section \ref{sec:data-models}. For this reason, taxonomy related to the
+data model depicts two different classes of data models, as shown in figure
+\ref{fig:taxonomy-data-model}.
 

### 2009-11-24 docs/marcello-dissertation/thesis/chapter4-netbeams-in-depth-requirements-analysis-and-specification.tex
Updates to chapter 2 through 7.
Chapter 2 lost the KVP section, since this approach had not been identified in
the literature for sensor networks. In this way, this makes the primary finding
and contribution of the dissertation. It will only appear during the technology
selection, what is the direct representation of what had happened;

chapter 3 adds details to the taxonomies;

Chapter 4 was cut short to describe SF-BEAMS, NetBEAMS in separate, as well as
keeps the entire implementation to the last chapter (appendix). 

Chapter 5 is the selection of the technology, but still misses highlights

--- 2009-11-15 docs/marcello-dissertation/thesis/chapter4-netbeams-in-depth-requirements-analysis-and-specification.tex
@@
 
-\chapter{Netbeams in Depth: Requirements Analysis and Specification}
+\chapter{Data Persistence for Sensor Networks: A Case Study for NetBEAMS}
+\label{chap:netbeams-overview}
 
-This chapter presents NetBEAMS and its implementation details, showing its
-main application of the the environmental sensor network deployed at the
-Romberg Tiburon Center (RTC) as it is the main case study.
+This chapter presents the main motivation and scope of of this dissertation.
+The infrastructure of SF-BEAMS is presented as a real-world environmental
+sensor network, deployed by the department of Biology at the San Francisco State
+University. It serves as the foundation of the execution of NetBEAMS, a
+joint project between the department of Biology and the Computer Science.
+As an alternative approach for data collection, NetBEAMS is used to
+automate the data management and collection process of SF-BEAMS. However, it
+offered different open problems, including Data Persistence. Therefore, the
+scope of the solution presented in the following sections and chapters are
+based in the case study presented in this current chapter.
 
@@
 
-The San Francisco Bay Environmental Assessment and Monitoring Station,
-or SF-BEAMS, is a marine sensor network operated by the Romberg Tiburon Center
-(RTC). This section details the SF-BEAMS sensor network located at
-the Tiburon Island, California, and then presents the details of NetBEAMS, as
-it is part of the case study of this work.
+The San Francisco Bay Environmental Assessment and Monitoring Station, or
+SF-BEAMS \cite{sfbeams2006}, is a marine sensor network operated by the Romberg
+Tiburon Center (RTC), whose main focus is the study of complex marine and
+estuarine environments using the SF-BEAMS sensors. The network is deployed
+offshore of its pier located on the banks of the San Francisco Bay at the
+Tiburon Island, California, being operated by the department of Biology from
+San Francisco State University.
 
@@
 
-The RTC's main focuses are in the study of complex marine and estuarine
-environments using the SF-BEAMS sensors deployed offshore of its pier located
-on the banks of the San Francisco Bay at the Tiburon Island, California. In
-this way, the project is responsible for providing data for water quality
+The SF-BEAMS sensor network is responsible for providing data for water quality
 monitoring, as well as weather and surface conditions. Figure
@@
 data can be directly transferred to the labs via ethernet cables or collected
-manually download by a staff's laptop computer.
+manually by a laptop computer.
 
@@
 
-As described in section \ref{sec:sn-infrastructure}, sensor devices generates
+As described in section \ref{sec:sn-infrastructure}, sensor devices generate
 its observed data based on properties of measurements defined by its
-manufacturer. SF-BEAMS is a single-hop star sensor network containing
+manufacturer. SF-BEAMS can be seen a single-hop star sensor network containing
 different instances of sensors, including the "YSI 6600 ESD V2", as seen on
@@
 archival, where it is described, indexed and distributed to its main users.
-According to its staff members, in May 2009 the infrastructure of the SF-BEAMS
-included the following was organized as follows:
+According to one of its staff members, the infrastructure of the SF-BEAMS in
+May 2009 was comprised of the following:
 
 \begin{itemize}
-  \item 5 YSI sondes in operation at the RTC site Pier in Tiburon Island, San
-  Francisco Bay;
+  \item Among other devices, 5 YSI sondes in operation at the RTC pier;
   \item Each YSI sonde produces 52 Bytes for each its observations;
   \item The sampling frequency rate could be configured in ranges of 1, 6 or
-  15 minutes, depending on the staff's speficiation. 
+  15 minutes, depending on the specification;
+  \item Data for final users are published with Provenance Data regarding time,
+  but no metadata regarding the origin of data.
 \end{itemize}
@@
 As it is shown, the data produced by this type of sensor runs in the order of
-at most hundreds of Megabytes a year, what can be considered a very low data
-storage requirement as compared to other systems such as web applications. Upon
-collecting the data from sensors, the RTC lab staff use automation scripts
-written in Matlab to process, index and distribute the raw data in different
-formats, including the OPeNDAP \cite{opendap}, a format widely used in the
-research institutions that promotes easier data exchange among them. The
-access to the archived data is done by using the OPeNDAP server is located at
-http://sfbeams.sfsu.edu:8080/opendap, as a request and the related output for
-one of the YSI devices can be see in listing \ref{file:rtc-ysi-opendap}.
+at most hundreds of Megabytes a year for a maximun of \textbf{483,840} samples,
+what can be considered a very low data storage requirement as compared to
+other networks. Upon collecting the data from sensors, the RTC lab staff use
+automation scripts written in Matlab to process, index and distribute the raw
+data in different formats, including the OPeNDAP \cite{opendap}, a format
+widely used in the research institutions that promotes easier data exchange
+among them. The access to the archived data is done by using the OPeNDAP
+server is located at http://sfbeams.sfsu.edu:8080/opendap, as a request and
+the related output for one of the YSI devices can be see in Listing
+\ref{file:rtc-ysi-opendap}.
 
-Although the execution of the RTC SF-BEAM sensor network produces its
-execution, many operational challenges were faced by this type of
-infrastructure, when it comes to the data utilization of wireless sensors that
-are deployed in offshore conditions, \cite{netbeams2009} described the
-operational challenges faced by the RTC staff and proposed NetBEAMS, a
-component-based infrastructure that uses COTS \footnote{Common-Off-The-Shelf
-designates a product that is produced and sold in bulk} embedded systems and
-software to automate the process of data collection. Next section describes
-how NetBEAMS is used to extract data from the SF-BEAMS sensor network.
+\subsection{Taxonomic Classification of SF-BEAMS}
 
-\section{RTC Requirements for Data Use}
+In order to distinct each of the properties of the SF-BEAMS sensor
+networks, consider the following list as a classification of the data
+persistence mechanisms from the sensor network used by the project:
+
+\begin{itemize}
+  \item \textbf{The Purpose of Sensor Data}: the data from SF-BEAMS is purely
+  used for \textbf{Data Archival} purposes, since they are stored in the RTC's
+  lab as part of the collaboration research project that monitors each part of
+  the California's coast. The data sets are distributed by specific date in a
+  yearly calendar;
+  \item \textbf{The Location of the Sensor Data}: it is clear that the purpose
+  of data is directly related to its location and, as consequence, the strategy
+  of \textbf{External Data Storage} is used to store its collected data at the
+  network sink at the RTC's labs;
+  \item \textbf{Data Model}: no data model is used by the RTC's staff to model
+  the data. A simple transformation of the data collected from the sensors
+  directly to spreadsheets, using comma-delimited values, is used. That is, the
+  data mode used is \textbf{Schema-less};
+  \item \textbf{Data Provenance}: data is stored in files with conventions that
+  takes the time when the data was collected. The contents of the files include
+  the \textbf{Data Identity} and \textbf{Time Dimensions}, as shown in
+  the samples collected and shown in Listing \ref{file:rtc-ysi-opendap};
+  \item \textbf{Query Processing Mechanism}: data is queried in a
+  \textbf{Centralized} fashion, due to the nature of its purpose and locality;
+  \item \textbf{Database System Organization}: there are no database in place.
+  The SF-BEAMS labs use the OPeNDAP Hyrax server as a middleware used to read
+  and interpret the collected data. 
+\end{itemize}
+
+\section{NetBEAMS: a component-based approach for SF-BEAMS}
 \label{sec:problem-requirements}
 
-Main users are biologists, students and researchers with some knowledge in
-programming and scripting languages.
+Although the execution of the RTC's sensor network can be operated as it
+described in the previous sections, \cite{netbeams2009} described some
+operational challenges faced by the RTC staff during regular activities during
+the data collection process. It was clear to the research group that SF-BEAMS
+could have not only its data gathering process automated, but also its data
+management and distribution. By the use of COTS\footnote{Common-Off-The-Shelf
+designates a product that is produced and sold in bulk} embedded
+devices and open-source \cite{open-source} software, the research group 
+developed a second version of NetBEAMS, a component-based approach suggested
+to improve the operational activities from SF-BEAMS by using systems
+automation techniques. In brief, one of the open problems of NetBEAMS was
+regarding Data Persistence, the main motivation of this dissertation.
+
+The Networked Bay Environmental Assessment and Monitoring System, or NetBEAMS, 
+is a work-in-progress joint project proposed by the department of Computer
+Science and the RTC at San Francisco State University. The main reason for the
+inception of NetBEAMS is the potential to decrease operational costs and speed
+the data gathering process. In the current state of SF-BEAMS, Biologists need
+to manually collect the data from the sensor devices in order to take them to
+the network sink (lab) for further processing. For this reason, the Data
+Sensor Platform (DSP) \cite{netbeams2009} was proposed as the system
+architecture that can address the operation of SF-BEAMS. Therefore, an in-depth
+detail of the the DSP Platform, its architecture and data gathering process
+can are provided in the appendix section \ref{sec:dsp-details}, and may be
+referred at any time.
+
+The scope of this work then is defined as to provide a persistence layer for
+NetBEAMS, making a selection of a persistence layer that is complaint to the
+taxonomical classification of SF-BEAMS. Furthermore, the technology to be
+selected for the persistence layer must take into account the requirements of
+NetBEAMS primary users such as Computer Science researchers, as well as end
+users such as Biologists from the RTC laboratories.
+
+\subsection{Requirements for NetBEAMS Data Use}
+
+The main reason for the operation of SF-BEAMS using NetBEAMS's automated
+approach is that it can potentially reduce operational costs and improve the
+data collection process. In review of this, the scope of a persistence layer
+can be summarized as a group of functional and non-functional requirements in
+order to select a database system for the sensor network supported by NetBEAMS.
 
@@
   for data persistence in a database system;
+  \item The Persistence System must obey to the characteristics of the
+  categorization of SF-BEAMS used by the taxonomies. 
 \end{itemize}
@@
 \begin{itemize}
-  \item Main users are biologists, no expertise in Database Systems;
-  \item Data Representation in human-readable format;
-  \item Scalable system with low degree of maintanance;
-  \item Data searchable in near-real-time
+  \item The data model used to describe the data must not impose restrictions
+  to the users of the system: Biologists and students without expertise in
+  Database Systems;
+  \item Data Representation must be similar to the ones used by RTC,
+  with a more human-readable format;
+  \item Scalable System with low degree of maintanance, in a way that data
+  gathering service should be able to manage the system with less service
+  interruption;
+  \item Data must be searchable in near-real-time with good performance, as
+  compared to the time when data was collected to the time when they are available;
+  \item Easy data integration and export capabilities must be also taken into
+  account;
+  \item The database system must be free of charge, following the
+  implementation specifications of NetBEAMS.
 \end{itemize}
-
-\section{Understanding NetBEAMS Architecture}
-\label{netbeams-architecture}
-
-The Networked Bay Environmental Assessment and Monitoring System, or NetBEAMS, 
-is a work-in-progress joint project proposed by the department of Computer
-Science and the RTC at San Francisco State University. As described by
-\cite{netbeams2009}, the solution can be used to automated the data gathering
-process of the SF-BEAMS as described in section \ref{sec:sfbeams}. This
-section fully details how NetBEAMS is executed in order to collect data from
-SF-BEAMS, describing pertinent design and architectural details.
-
-One of the motivations for the inception of NetBEAMS was the automation
-of the data gathering and processing from the SF-BEAMS sondes without requiring
-human intervention. Most importantly, the decrease with the costs of constant 
-offshore site visits by boat, where biologists need manually collect the data
-from the sensor devices in order to take them for further processing at the
-lab. For this reason, the NetBEAMS research group proposed the use of the Data
-Sensor Platform (DSP) \cite{netbeams2009}, as the second attempt to automate
-the operation of SF-BEAMS.
-
-This following sections details the DSP, its architecture and how data is
-transported from the network nodes to the main network sink, the RTC labs.
-
-\subsection{The Data Sensor Platform (DSP)}
-
-The Data Sensor Platform, in short DSP Platform, is based on a micro-kernel
-architecture developed on top of the Java modular framework called OSGi
-\cite{osgi}. The platform represents the foundation of the execution of
-NetBEAMS, since it is executed from an embedded hardware called Gumstix
-\cite{gumstix} using a Linux operating system and on top of the Java Virtual
-Machine (JVM). In order to transport data from sensors hosts to the network
-sink, the use of a cellular connection is used to transport the collected data
-from the senso devices using the HTTP Protocol, as described by
-\citet{netbeams2009}. Figure \ref{fig:sf-netbeams-node} shows the
-architecture of the inhanced node designed by the NetBEAMS research group.
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[scale=0.5]{../diagrams/DSP-Gateway-Node}
-  \caption{The NetBEAMS Node enhancement with the Gumstix and the GSM Modem}
-  \label{fig:sf-netbeams-node}
-\end{figure}
-
-The following list describes each of the devices in the NetBEAMS Node.
-
-\begin{itemize}
-  \item \textbf{YSI 6600EDS V2}: the YSI device is the regular sonde
-  responsible for observing the environment. It is connected to the Gumstix
-  embedded system, where the DSP is executed;
-  \item \textbf{Gumstix console-vx}: it's a COTS ARM-based hardware that
-  provides a computer-on-module environment for the development of small
-  embedded systems. This is the main environment of data extraction and remote
-  processing is accomplished. It runs a cross-compiled version of Linux Kernel
-  2.6;
-  \item \textbf{Huawei E220 USB Modem}: it is used to transfer the data from the
-  Gumstix to the RTC Labs Data Center.
-\end{itemize}
-
-The DSP runs inside of the embedded system Gumstix. The following sections
-details its design and process for extracting data.
-
-\subsection{The DSP System Design}
-
-This section focus on the development of a Software Platform for the NetBEAMS
-Gateway Embedded System, or on the system developed in the Gumstix device. As
-shown in figure \ref{fig:netbeams-software-stack}, the NetBEAMS Gateway Node is
-an embedded system that contains a set of layer found in any computer.
-However, each of the resources described from the system were customized for
-the NetBEAMS system. The main components of such system can be summarized as
-follows:
-
-\begin{itemize}
-  \item \textbf{Operating System}: it uses a cross-compiled Gentoo Linux, which
-  supports a wide variety of development tools, including the Java Virtual Machine
-  (JVM), the underlying platform system;
-  \item \textbf{Java Virtual Machine}: it uses the JamVM, a ~200 KB version of
-  the Sun Microsystems Java Virtual Machine, version 2.0;
-  \item \textbf{OSGi Framework}: it uses the Knopflerfish implementation of the
-  OSGi 4.1 specification. More details in the following sections.
-  \item \textbf{DSP Platform and other Bundles}: The plug-and-play DSP
-  Components are based on the OSGi bundles capabilities, which can reuse services registered in the OSGi framework.
-\end{itemize}
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[scale=0.6]{../diagrams/NetBEAMS-Client-Node-Components}
-  \caption{The NetBEAMS's Gumstix Software Stack}
-  \label{fig:netbeams-software-stack}
-\end{figure}
-
-As depicted in figure \ref{fig:netbeams-software-stack}, the DSP Sonde Data
-Producer component opens a serial connection with the device through the Java
-Native Interface (JNI) driver and produces the data flow 1. The data stream
-format at this stage is as described on table \ref{tab:ysi-data-stream}.
-Later, by parsing and transforming the data to DSP internal representation, it
-sends the data to the DSP Platform as depicted on flow 2. Finally, the DSP
-Platform decides to send the data to the DSP Wire Transport Client, as shown
-on flow 3, by following rules defined by the system administrator. At this
-point, the data is ready to be transmitted to the server, which is covered on
-the following sections.
-
-\subsection{The DSP Platform Architecture}
-
-The architecture of the DSP, aligned with the OSGi capabilities, reuses the
-Service-Oriented Architecture that can easily enable and disable plug-and-play
-components. By following a publish-subscriber design-pattern \cite{gof}, the
-service producer registers into the service broker while the consumer looks
-for a given registered service in the Service Broker. For this reason, this
-approach allows decoupled interoperability among modules as it is shown in
-figure \ref{fig:publish-subscriber-pattern}.
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[scale=0.6]{../diagrams/publish-subscriber-pattern}
-  \caption{The publish-subscriber design-pattern for services}
-  \label{fig:publish-subscriber-pattern}
-\end{figure}
-
-The DSP infrastructure is based on a loosely-coupled architecture hosted in a
-given host computer with an OSGi container, where the entire system is
-composed by a set of plug-n-play OSGi modules. Each DSP component, as an
-instance of an OSGi Module, can perform specialized tasks according to its
-specification. In general, each DSP component can be categorized as follows:
-
-\begin{itemize}
-  \item \textbf{Data Producer (DP)}, whose main function is to produce data for
-  the system;
-  \item \textbf{Data Consumer (DC)}, whose main function is to consume the
-  received data.
-\end{itemize}
-
-\begin{figure}[!b]
-  \centering
-  \includegraphics[scale=0.7]{../diagrams/DSP-Producers-Consumers-Components-Interactions}
-  \caption{The basic architecture of a Data Sensor Platform (DSP)}
-  \label{fig:DSP-Producers-Consumers-Components-Interactions}
-\end{figure}
-
-The Data Producers and Data Consumers exchange data through the DSP Message
-Broker using a message unit called DSP Message. As it is depicted in figure
-\ref{fig:DSP-Producers-Consumers-Components-Interactions}, a Data Producer
-creates a set of one or more DSP Messages wrapped up in a DSP
-MessagesContainer in order to be sent to the DSP Message Broker. When the
-broker receives the set of DSP Messages, it requests a list of Data Consumers
-to each of the DSP Messages based on rules managed by the DSP Matcher.
-Whenever the broker acquires the list of Data Consumers, it delivers a copy of
-the DSP Message to each of the Data Consumers in the receiving list. Upon the
-DSP Message arrival, each Data Consumer can extract the actual data from the
-DSPMessage. Finally, although the components can be distinguished between Data
-Producers and Consumers, any of them are entitled to send and receive
-administrative DSP Messages, which is covered in the following sections.
-
-When the DSP Platform is installed into the OSGi framework, it follows the
-OSGi specifications without any changes to remain it in the installed state.
-In this way, it is important to note that the DSP Platform and its
-plug-and-play components are directly mapped with the OSGi Framework as
-follows:
-
-\begin{itemize}
-  \item \textbf{OSGi Bundle} = DSP Component Artifact in Jar format
-  \begin{itemize}
-    \item SondeDSPProducer.jar is the component developed for the YSI sonde
-  \end{itemize}  
-  \item \textbf{OSGi Activator} = DSP Component Activator
-  \begin{itemize}
-    \item Class specified in the MANIFEST.MF artifact (next sections)
-  \end{itemize} 
-  \item \textbf{OSGi Service Reference} = DSP Component Class
-  \begin{itemize}
-    \item  Class specified by the component author.
-  \end{itemize}
-\end{itemize}
-
-The DSP Platform, along with the DSP Framework, are the only DSP Components
-needed to be installed into the OSGi container. For this reason,the DSP
-Platform is responsible to install and activate the additional DSP components
-that makes part of the platform during the start up process, which leverages
-its bootstrap process. Nevertheless, in order to automate the process of
-automatic install and removal of the plug-and-play components, the DSP Platform
-provides a deployment configuration file that describes which DSP components
-must be installed with a given priority number. The name of this configuration
-artifact is called config.xml (see listing \ref{file:dsp-config.xml}), and is
-depicted by the XML Schema \cite{xml-schema} in figure
-\ref{fig:dsp-config-schema}.
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[scale=0.8]{../diagrams/config-schema}
-  \caption{XML Schema for the main DSP deployment configuration artifact}
-  \label{fig:dsp-config-schema}
-\end{figure}
-
-In order to show an example, consider the DSP Platform on the embedded system
-Gumstix in figure \ref{fig:sf-netbeams-node}. It is started by providing a
-deployment configuration artifact config.xml included in the appendix. It lists
-the components that the DSP Platform needs to install and load for its
-operation. As it is shown, the DSPWireTransportClient and the DSPSondeProducer
-are the identification of the components, followed by the name of the physical
-file and an assigned loading priority. While the DSPWireTransportClient is
-responsible for the serialization and deserialization of the DSP Messages
-exchanged by the DSPSondeProducer and any remote DSP component. In this case,
-the DSPSondeProducer is the DSP component responsible to collect measurement
-data from the YSI Sonde and temporarily store them before being transmitted.
-
-Therefore, in order to orchestrate the DSP Components together, the DSP
-Platform component is responsible for installing and uninstalling during its
-state "Starting" and "Stopping", respectively, as it is shown in figure
-\ref{fig:DSPPlatform-Install-Usage-State-Diagram}. When the system
-administrator starts the DSP Platform, it performs the regular OSGi procedures
-to start the bundle. Up until that state, it will get the list of DSP
-Components to be installed as described in the config.xml artifact, trying to
-install each of them in an optimistic way. After that, it will continue on the
-state "Active" until the system administrator decides to stop it.
-
-\begin{figure}[!b]
-  \centering
-  \includegraphics[scale=0.6]{../diagrams/DSPPlatform-Install-Usage-State-Diagram}
-  \caption{The DSP Platform State Diagram}
-  \label{fig:DSPPlatform-Install-Usage-State-Diagram}
-\end{figure}
-
-When the DSP Platform is on the Active state, it is ready to manage the DSP
-Messages exchanges as described through figure
-\ref{fig:DSP-Producers-Consumers-Components-Interactions}. The following
-sections describe the participating classes in starting the DSP Platform and
-its execution.
-
-In a nutshell, the DSP Platform is an extension of the OSGi Platform, as it can
-be seen in the UML Class Diagram \cite{uml} in figure
-\ref{fig:DSP-Platform-Class-Diagram-Simple}. The class Activator contains a
-unique instance of the Platform class, which contains instances of the classes
-ComponentManager, Matcher, the MessageBroker and DSPContextImpl. The
-DSPBundleController is the main instance that manages the plug-and-play DSP
-components, and therefore, is tightly-coupled with the OSGi Framework classes.
-Additionally, this class is part of the publish-subscriber design-pattern, as
-it it implements the interfaces Bundle Listener and Service Listener, whose
-responsibility is to track listens to the changes of the state of
-ServicesReferences on the DSP Platform.
-
-\begin{figure}[!t]
-  \centering
-  \includegraphics[scale=0.7]{../diagrams/DSP-Platform-Class-Diagram-Simple}
-  \caption{The DSP Platform UML Class Diagram}
-  \label{fig:DSP-Platform-Class-Diagram-Simple}
-\end{figure}
-
-\section{DSP Message and DSP Messages Container}
-
-As shown in the previous section, a DSP Message is the main unit of
-communication among DSP Components. In a nutshell, a DSP Message can be seen
-as an abstraction of an envelop containing the sections: a header and a body.
-The former is used for identification and routing purposes, while the latter is
-used to carry the payload of the message, that is, the collected data from
-a sensor. Figure \ref{fig:DSP-Message-Representation} depicts the abstraction
-of a DSP Message and its main components.
-
-\begin{figure}[!b]
-  \centering
-  \includegraphics[scale=0.8]{../diagrams/DSP-Message-Representation}
-  \caption{An Abstraction of a DSP Message and its components}
-  \label{fig:DSP-Message-Representation}
-\end{figure}
-
-\begin{itemize}
-  \item \textbf{Header} is composed by 2 blocks that identify the data
-  producer and consumer of the message, as well as the time of creation, UUID,
-  etc. It is an important piece of information for the data delivery and
-  routing purposes.
-  \item \textbf{Body} contains the payload of the message. This section carries
-  the collected data from a sensor.
-\end{itemize}
-
-In general, the DSP Platform offers a veriety of types of DSP Messages for
-different purposes. For example, any measurement data must be
-wrapped up in a Measurement Message, while a Query Message is used to exchange
-messages among the components for the purpose of management. In this way, the
-main DSP Messages can be summarized as follows:
-
-\begin{itemize}
-  \item \textbf{Measurement Message} is used to transport any sensor collected
-  data;
-  \item \textbf{Query Message} is used to query a DSP component about its
-  configuration properties;
-  \item \textbf{Update Message} used to update a DSP component's configuration
-  properties;
-  \item \textbf{Acknowledgement Message} is used for the transport communication
-  protocol, by acknowledging the reception of a DSP Message. More details in the
-  Remote Data delivery section.
-\end{itemize}
-
-These DSP Classes and other can be seen in figure the  UML Class
-Diagram \cite{uml} of figure \ref{fig:DSP-Messages-Classes}. Each DSP Message
-contains an instance of a Header and Body. Whenever a DSP component is ready
-to transmit messages, it wraps up the set of DSP Messages into the instance of
-the Messages Container, which contains information about the collection of
-messages being transmitted with its own identification. In this fashion, the
-Messages Container is the main communication unit between 2 different DSP
-Components.
-
-\begin{figure}[!t]
-  \centering
-  \includegraphics[scale=0.6]{../diagrams/DSP-Messages-Classes}
-  \caption{The DSP Messages UML Class Diagram}
-  \label{fig:DSP-Messages-Classes}
-\end{figure}
-
-Another use of the DSP Messages is during the DSP Platform activation process,
-where bootstrap messages are used to configure each DSP component. In order to
-send configuration parameters, the DSP Platform uses Update Messages using a
-unique type of Message Content called DSP Properties, which is described in the
-next section.
-
-\section{Collected Sensor Data as the DSP Message Content}
-
-As mentioned in the previsou section, a DSP Message can carry any data on its
-body, also called the payload. In such a way, any data representation must
-just extend from the Message Content abstract class as it is shown
-\ref{fig:Sonde-MessageContainer-Class-Diagram}, as defines the data
-representation of the YSI Sonde data.
-
-\begin{figure}[!h]
-  \centering
-  \includegraphics[scale=0.6]{../diagrams/Sonde-MessageContainer-Class-Diagram}
-  \caption{The YSI Payload - Message Content UML Class Diagram}
-  \label{fig:Sonde-MessageContainer-Class-Diagram}
-\end{figure}
-
-The properties measured from the YSI Sonde can be captured by the class
-SondeDataType as a direct render of the measurements data depicted in table
-\ref{tab:ysi-data-stream}. In addition to the regular data types, the class
-also provides additional properties such as the attribute "samplingTimeStamp"
-that carries the time when the data was collected from the YSI Sonde. However,
-in order to capture a set of measurements at once, the SondeDataContainer
-class used to carry a set of SondeDataType classes, as it is composed by one
-or more instances of SondeDataType. In order to illustrate other examples,
-consider the UML Class Diagram \cite{uml} of image
-\ref{fig:Mouse-Actions--MessageContainer-Class-Diagram} as a payload of
-observations made from a mouse over a screen, which captures the name of the
-event, the button name and the x/y coordinates.
-
-\begin{figure}[!h]
-  \centering
-  \includegraphics[scale=0.5]{../diagrams/Mouse-Actions--MessageContainer-Class-Diagram}
-  \caption{The Mouse Actions Payload - Message Content UML Class Diagram}
-  \label{fig:Mouse-Actions--MessageContainer-Class-Diagram}
-\end{figure}
-
-As described in the previous section, the DSP Platform uses DSP Update
-Messages to configure internal components' initial configuration values. In
-order to do so, the DSP Platform uses an instance of the DSP Properties, which
-is composed of a list of initial DSP Property and their relating Values
-instance depicted on the UML class diagram \cite{uml} of figure
-\ref{fig:DSP-Property-Class-Diagram}.
-
-\begin{itemize}
-  \item \textbf{Property Instance}: initial-rate
-  \item \textbf{Value Instance}: 40
-\end{itemize}
-
-\begin{figure}[!t]
-  \centering
-  \includegraphics[scale=0.6]{../diagrams/DSP-Property-Class-Diagram}
-  \caption{The DSP Property Payload - Message Content UML Class Diagram}
-  \label{fig:DSP-Property-Class-Diagram}
-\end{figure}
-
-\subsection{DSP Platform Initialization Process}
-
-All in all, the DSP Platform requires the knowledge of the aforementioned DSP
-classes in order to initialize. The UML Sequence Diagram \cite{uml} in figure
-\ref{fig:From-OSGi-DSP-Platform-Start-to-All-Components-Sequence-Diagram}
-depicts the exact moments when the additional the plug-and-play DSP components
-are installed and bootstrapped by the DSP Platform, as the UML State Diagram
-\cite{uml} in figure \ref{fig:DSPPlatform-Install-Usage-State-Diagram}
-described.
-
-\begin{itemize}
-  \item \textbf{Start DSP Activator Gate}: after installing the Activator
-  indicated by the DSP Bundle's MANIFEST.MF artifact, the OSGi Framework makes
-  a call to the method Activator.start();
-  \item \textbf{Bootstrap DSP Component Gate}: after the component has been
-  initialized (by the call the method init() from the class ComponentManager),
-  the DSP Component is ready to be bootstrapped, and the optional bootstrap
-  message is delivered.
-\end{itemize}
-
-\begin{figure}[!b]
-  \centering
-  \includegraphics[scale=0.5]{../diagrams/From-OSGi-DSP-Platform-Start-to-All-Components-Sequence-Diagram}
-  \caption{The DSP Platform Initialization UML Sequence Diagram}
-  \label{fig:From-OSGi-DSP-Platform-Start-to-All-Components-Sequence-Diagram}
-\end{figure}
-
-\newpage
-
-\subsection{Data Delivery Mechanism}
-
-In general, when a DSP Component finishes preparing the DSP Messages Container,
-it contacts an entity responsible for sending and proxying message to other
-components. The DSP Broker acts like a mailman, receiving messages and
-forwarding copies of it to other entities in the system, abstraction in which
-is shown in figure \ref{fig:DSP-Message-Broker-Abstraction}. In this way, the
-DSP Broker receives a DSP Message and makes necessary deliveries to DSP
-components that it has knowledge of.
-
-\begin{figure}[!h]
-  \centering
-  \includegraphics[scale=0.8]{../diagrams/DSP-Message-Broker-Abstraction}
-  \caption{A DSP Message Broker Abstraction}
-  \label{fig:DSP-Message-Broker-Abstraction}
-\end{figure}
-
-The use of messaging exchange paradigm is used in NetBEAMS because it decouples
-the system by using messages take into account their structure to filter the
-messages. In this way, the DSP Broker uses a filter component called DSP
-Matcher, which provides a list of DSP Components that needs to receive a copy
-of the DSP Message. It turns out that the DSP Matcher bases its decisions on
-structured rules provided by the system administrator for an specific DSP
-Platform instance. The UML Class diagram for the DSP Broker, Matcher and
-particiaping classes is depicted in figure
-\ref{fig:DSPBroker-Matcher-Class-Diagram}, and describes how the DSP Broker
-implementation is tightly coupled with the DSP Matcher, as well as to the
-lists of DSP Components sorted by name and type.
-
-\begin{figure}[!b]
-  \centering
-  \includegraphics[scale=0.6]{../diagrams/DSPBroker-Matcher-Class-Diagram}
-  \caption{The DSP Broker-Matcher UML Class Diagram}
-  \label{fig:DSPBroker-Matcher-Class-Diagram}
-\end{figure}
-
-In general, the DSP Matcher is composed by a list of MatchRule, which contains
-a reference to a MatchCriteria and a MatchTarget. The former identifies the
-rule for matching a given DSP Message based on the producer and consumer's
-message type and location, while the latter describes the DSP Component that
-will receive the DSP Message, and if its delivery is done by means of a gateway
-component. In this way, the DSP Matcher can be seen as a function that takes a
-DSP Message as an input and returns a list of DSP consumers, and the
-matching configuration artifact provided by the administrator. See listing
-\ref{file:dsp-matcher-config.xml}) for details.
-
-\begin{table}
-    \caption{DSP Matcher Rule Algorithm to filter Data Consumer components to
-    receive a given Message}
-    \begin{center}
-        \begin{tabular}{lr}
-          \textit{Data Consumers ( DSP Message ) := Verify the DSP
-          Message's Header + Verify the matcher rules (which contains the list
-          of consumers)}
-        \end{tabular}
-    \end{center}
-    \label{tab:ysi-data-stream}
-\end{table}
-
-One example of such matching rule used by the DSP installed at the Gumstix can
-be seen in listing \ref{file:dsp-matcher-config-gumstix.xml}. the Gumstix
-system to configure the data delivery of the messages produced by the DSP
-Component DSP Sonde Producer. It describes a list of maching rules that
-reseambles a filtering system. This configuration contains 2 match rules: one
-related to internal DSP Configuration and another for a remote destination.
-The match criteria describes what the DSP Message Header properties must be,
-while the match target is the destination of the message. One important
-definition is the gateway Component Type, which defines the name of a remote
-DSP Component that is capable of receiving a remote DSP Message.
-
-Upon receiving all the matching rules and analyzing the DSP Message against
-the rules, the DSP Broker selects a set of the running DSP components to receive
-a copy of the DSP Message object in two different ways:
-
-\begin{itemize}
-  \item \textbf{In-memory local message delivery}: if the receiving DSP
-  Component is located in the current local host, a deep copy of the instance
-  of the DSP message is delivered;
-    \begin{itemize}[label=\textbullet]
-        \item IP Addresses are correctly resolved by using the Ethernet card in
-        the device: localhost, 127.0.0.1, or the same IP Address for Producer
-        and Receiver resolves into a local device. An example for such type of
-        delivery can be seen in the match ruleID
-        all messages sent to dsp manager of listing 
-        \ref{file:dsp-matcher-config-gumstix.xml}.
-    \end{itemize}
-  \item \textbf{Serialized remote message delivery}: if the receiving DSP
-  Component is located in a foreign/remote host, the message is serialized in a
-  format defined by the transport protocol chosen the DSP Component responsible
-  for the transport. The following section describes the existing DSP Data
-  Transport component;     
-  \begin{itemize}[label=\textbullet]
-        \item IP Address from the Producer and Consumer are different, and are
-        not resolved to be in the same host. An instance of such match rule can
-        be seen in the match rule ID send-message-to-remote url dsp of the code
-        snippet \ref{file:dsp-matcher-config-gumstix.xml}.
-    \end{itemize}
-\end{itemize}
-
-\subsection{Sending DSP Messages through the Wire}
-
-In order to send messages through the wire, the DSP Platform uses two symmetric
-components that serialize and deserialize DSP Messages to XML and back to Java
-POJO, respectively, and vice-versa. Named DSP Wire Transport Client and DSP
-Wire Transport Server \cite{netbeams2009}, these components use the HTTP
-\cite{http} protocol to transport the serialized version of the DSP Messages.
-Figure \ref{fig:DSP-to-DSP-Remote-Communication} depicts the scenario where a
-DSP Sensor Node transmits a DSP Message through an HTTP POST request.
-
-\begin{figure}[!b]
-  \centering
-  \includegraphics[scale=0.6]{../diagrams/DSP-to-DSP-Remote-Communication}
-  \caption{The remote communication between two DSP Components using the HTTP
-  Protocol}
-  \label{fig:DSP-to-DSP-Remote-Communication}
-\end{figure}
-
-\begin{itemize}
-  \item \textbf{DSP Wire Transport Client}: responsible for making HTTP POST
-  Requests to the service provided by the DSP Wire Transport Server component.
-  The body of the HTTP POST request contains a serialized version of an
-  instance of a DSP MessagesContainer in XML format;
-  \item \textbf{DSP Wire Transport Server}: exposes a Web Server listening to
-  port 8080 that receives the HTTP POST request from the Client. Upon receiving
-  the request, the server deserializes the DSP MessagesContainer back to the
-  in-memory POJO and proceeds with the delivery to the DSP Broker. At this
-  situation, the component packages acknowledgement messages and any queued
-  message to the client and piggybacks \cite{xml-piggybacking} it to the HTTP
-  response body.
-\end{itemize}
-
-As described by \cite{netbeams2009}, NetBEAMS is a single-hop sensor network
-whose sensors transmits data from the sensors to the server. Upon receiving the
-data stream, the DSP Wire Transport Server component desirializes the DSP
-MessagesContainer and its enclosing DSP Messages back to a Java class instance
-and delivers each of the DSP Messages to the DSP Broker, which delivers a
-copy of the message to each DSP Component listed by the maching rule. 
-%This
-%process is depicted by the UML Sequence Diagram \cite{uml} in figure \ref{fig:}
-
-Whenever a DSP bundle producer is ready to transmit messages, they are added
-into a DSP Messages Container (see UML Class Diagram in figure
-\ref{fig:DSP-Messages-Classes}), which is serialized in XML to be transmitted.
-Listing \ref{file:dsp-message-serialized-ysi} shows an instance of a DSP
-MessagesContainer in serialized in XML, ready for transmission. In fact, the
-listing is an example of the transmitted data collected from a sensor located
-at the 192.168.0.103 to be transmitted to the host 192.168.0.106, and the
-process of transmission in the server-side can be followed in figure
-\ref{fig:From-DSPSondeComponent-to-DSPWireTransportClient-Sequence}.
-
-\begin{figure}[!h]
-  \centering
-  \includegraphics[scale=0.5]{../diagrams/From-DSPSondeComponent-to-DSPWireTransportClient-Sequence}
-  \caption{The UML Sequence Diagram for the production of a YSI Message}
-  \label{fig:From-DSPSondeComponent-to-DSPWireTransportClient-Sequence}
-\end{figure}
-
-The Sonde DSP Component is responsible for starting a worker thread called
-SondeProducer, which is responsible for constantly read from the RS-232 serial
-port by using an instance of the SerialReader class. This process depicts the
-production of the data flow "flow 1" of figure
-\ref{fig:netbeams-software-stack}, which is a collection of SondeDataType for
-each of the measurement data read on a SondeDataContainer. As depicted in
-figure \ref{fig:From-DSPSondeComponent-to-DSPWireTransportClient-Sequence}, it
-first read the serial connection using the class SerialReader, which parses 
-the data stream. After parsing the data, the reader creates an instance of the
-class SondeDataType, the POJO responsible for carrying out the measurement
-data. As described earlier, the data is added into the SondeDataContainer to
-dispatched to the DSP Broker, whose responsibility is to get the list of DSP
-Components that is setup to receive a copy of the DSP message (see previous
-code snippet). As for the setup done for the Gumstix, the measurement message
-is supposed to be sent to a remote destination of host 192.168.0.7. At this
-point, the DSP Broker makes a decision to send the DSP Message to the default
-gateway component DSPWireTransportClient.
-
-After the message arrives in the DSPWireTransportClient, the message is
-maintained in a temporary repository of outbound DSP Messages called
-MessagesQueues after the message is hashed by the destination IP address and
-the DSP Client sender using an instance of QueueMessageData. As shown in figure
-\ref{fig:From-DSPWireTransport-Client-to-DSPWireTransportServer-Sequence}, the
-DSP TransportSender is a worker thread that is started by the DSP Wire
-Transport Client and is constantly transmits the data in the outbound queues
-after a configurable rate. It makes a transport of the message and receives
-the response, updates the state of the messages to transmitted, to finally
-send the response messages to the DSP Message Broker.
-
-\begin{figure}[!t]
-  \centering
-  \includegraphics[scale=0.5]{../diagrams/From-DSPWireTransport-Client-to-DSPWireTransportServer-Sequence}
-  \caption{The UML Sequence Diagram describing the Wire Transport Server
-  receiving a DSP Message}
-  \label{fig:From-DSPWireTransport-Client-to-DSPWireTransportServer-Sequence}
-\end{figure}
-
-While the DSPWireTransportClient makes an HTTP Request with the DSP Message,
-the DSPWireTransportServer's servlet receives the request. It first
-deserializes the MessagesContainer and delivers each of its messages to the DSP
-Broker. At this point, the server verifies if there are messages in the
-outbound queue that are addressed for the host that requested the messages
-delivery. By using a technique called piggybacking, the server adds to the HTTP
-Response a serialized version of the MessagesContainer with any DSP Messages.
-In addition to any DSP Message, the server prepares and sends acknowledgement
-messages back to the requesting host to the host.
-
-\begin{figure}[!t]
-  \centering
-  \includegraphics[scale=0.5]{../diagrams/From-DSPWireTransport-Server-To-DSP-Broker}
-  \caption{The UML Sequence diagram describing the DSP
-Message delivery on the server-side and the response}
-  \label{fig:From-DSPWireTransport-Server-To-DSP-Broker}
-\end{figure}
-
-The gate "From DSP WireTransport Server to Broker Gate" defines the actual
-delivery of the DSP Message to the DSP Broker. As this section is a tutorial
-about the DSP Platform, the remaining steps will be explained on Section 3,
-while showing the design and implementation of the component DSP Data
-Persistence.

### 2009-11-24 docs/marcello-dissertation/thesis/chapter5-taxonomic-persistence-storage-selection-for-netbeams.tex
Updates to chapter 2 through 7.
Chapter 2 lost the KVP section, since this approach had not been identified in
the literature for sensor networks. In this way, this makes the primary finding
and contribution of the dissertation. It will only appear during the technology
selection, what is the direct representation of what had happened;

chapter 3 adds details to the taxonomies;

Chapter 4 was cut short to describe SF-BEAMS, NetBEAMS in separate, as well as
keeps the entire implementation to the last chapter (appendix). 

Chapter 5 is the selection of the technology, but still misses highlights

@@
 This chapter analyzes each of the taxonomies defined in chapter
-\ref{chap:taxonomies} against the requirements defined by our case study,
-NetBEAMS, in chapter \ref{sec:problem-requirements}. Moreover, it evaluates
-different database technologies against on those taxonomies, and tries to select
-one of them which can be a candidate for an experimental analysis.
+\ref{chap:taxonomies} against the requirements defined by NetBEAMS as our case
+study in chapter \ref{sec:problem-requirements}, following the scope of a given
+solution also defined in that chapter. The contenders were drawn from a mixed
+list of technologies either related in the literature review of chapter 2 or
+from the Internet.
 
-First, the contender technologies are as follows:
+\section{Database Systems Contenders}
+
+This section describes the database systems contenders considered to be used
+for the implementation of Data Persistence layer for NetBEAMS. As described in
+the previous chapter, the selection of the technologies must be based on the
+list of functional and non-functional requirements specified within the scope
+of implementation defined.
+
+One of the primary questions related to data persistence is regarding the data
+model and type of the database, such as the use of relational model or
+structured models. In addition to that, and given the current time of academic
+research and new trends distributed computing such as Cloud Computing and the
+Internet \cite{cloud-comp-survey}, the selection of a data persistence for
+sensor networks should maintain a wider spectrum of the types of database
+systems. Among other intriguing questions, \cite{db-is-rdbs-dommed} reflects 
+about the use of alternatives to the relational model for dynamic
+environments such as data persistence for the Web. Similarly, these
+trends are also related to the advances of Cloud Computing and its persistence
+layer as a solution to dynamic systems \cite{cloud-comp-architectures}. For
+this reason, this work also included the overview of alternative data models
+for the development of an experiment that brings data persistence to NetBEAMS.
 
 \begin{itemize}
-  \item MySQL
-  \item Microsoft SQL 
-  \item Berkeley TinyDB
-  \item MongoDB
-  \item CouchDB
-  \item IBM DB2
+  \item \textbf{MySQL}: an open-source relational database system that is the
+  primary choice of the development of projects of different sizes. It was selected
+  because of its popularity in the current community;
+  \item \textbf{Oracle}: is the most popular commercial database system, having
+  a campus wide academic licenses availabe;
+  \item \textbf{TinyDB}\cite{tinydb}: as referred in the literature review,
+  TinyDB is a database system used in many different sensor networks implementation
+  \cite{sn-db-tinydb};
+  \item \textbf{MongoDB}\cite{mongodb}: an open-source key-value database system
+  described in the article that questions the use of the relational model
+  \cite{db-is-rdbs-dommed}, as well as the Cloud Computing trends survey
+  \cite{cloud-comp-architectures}.
+  \item \textbf{DB2}\cite{db2}: in the search of an XML database system, DB2 is
+  listed as a hybrid database system that offers both XML and relational model
+  \cite{db-xml-enabled}. It was selected because NetBEAMS can represent the
+  collected data in the XML format \ref{sec:dsp-message}.
 \end{itemize}
 
-\section{Analysis of the Purpose of Sensor Data Taxonomy}
+The following sections analyzes the taxonomies defined in chapter 3, along with
+the requirements for the Data Persistence Layer for NetBEAMS, in section 4.
+
+\section{Analysis of the Purpose of Sensor Data}
 
@@
 chapter, and the characteristics of the sensor network defined at the SF-BEAMS
-\cite{sfbeams2006} site, the data characteristics are as follows:
+site \cite{sfbeams2006}, the data characteristics are as follows:
 
@@
 
-Given the facts collected from the observations of the execution of NetBEAMS
-through the SF-BEAMS sensor network, we can say that the purpose of collected
-data is \textbf{Data Archival}.
+As summarized in the previous section, the nature of data from NetBEAMS through
+SF-BEAMS is used for the purpose of \textbf{Data Archival}.
 
@@
 
-Any of the contenders are options to cover for Data Archival.
+As it is supposed to be, all the contenders are developed to support data
+persistence and management, what is the requirement for the Data Archival
+purpose of the sensor network data.
 
@@
   \item \textbf{MySQL}: Supports
-  \item \textbf{Microsoft SQL}: Supports
-  \item \textbf{Berkeley TinyDB}: Supports
+  \item \textbf{Oracle}: Supports
+  \item \textbf{TinyDB}: Supports
   \item \textbf{MongoDB}: Supports
-  \item \textbf{CouchDB}: Supports
-  \item \textbf{IBM DB2}: Supports
+  \item \textbf{DB2}: Supports
 \end{itemize}
@@
 
-Based in the observations described, the type used for storing
-data can be either the \textbf{External Storage} or the
-\textbf{Data-Centric Storage}. The former is the current approach used by the
-SF-BEAMS sensor network, while the latter adopted as an improvement of the
-persistence infrastructure for NetBEAMS.
+Based in the observations also described in the previous chapter, the type used
+for storing data can be either the \textbf{External Storage}. However, the use
+of a \textbf{Data-Centric Storage} approach can be considered. The former is the
+current approach used by the SF-BEAMS sensor network, while the latter can be
+explored in the experiments.
 
@@
 
-In order to support this strategy, a technique called Database Sharding
-\cite{db-shard02} is a technique used in distributed database systems that
-uses denormalized data models to improve scalability in terms of data storage
-and data location. It is a type of database system patitioning where
-de-normalized data is spread across different other hosts, or shard \cite{db-shard01}.
-Different models of database sharding can be developed \cite{db-shard-schemas}.
+In order to support this strategy, approaches such as data partitioning based
+on the entities called Database Sharding
+\cite{db-shard02}, or the partitioning based on tables
+\cite{db-table-partition}. The former is a technique used in distributed
+database systems that uses denormalized data models to improve scalability in
+terms of data storage and data location, while the latter defines partitioning
+techniques in relational databases. It is a type of database system patitioning
+where de-normalized data is spread across different other hosts, or shard
+\cite{db-shard01}. Different models of database sharding can be developed
+\cite{db-shard-schemas}.
 
@@
   \item \textbf{MySQL}: Supports
-  \item \textbf{Microsoft SQL}: Supports
+  \item \textbf{Oracle}: Supports
   \item \textbf{Berkeley TinyDB}: Supports
@@
   \item \textbf{CouchDB}: Supports
-  \item \textbf{IBM DB2}: Supports
+  \item \textbf{DB2}: Supports
 \end{itemize}
@@
   \centering
-  \includegraphics[scale=0.7]{../diagrams/KVP-on-Relational-Model}
+  \includegraphics[scale=0.6]{../diagrams/KVP-on-Relational-Model}
   \caption{KVP Data Model implementation using Relational Model}
@@
   \centering
-  \includegraphics[scale=0.7]{../diagrams/persistence-example-relational-kvp}
+  \includegraphics[scale=0.75]{../diagrams/persistence-example-relational-kvp}
   \caption{KVP over Relational Model Database Prototype}
@@
 
-The advance of Cloud Computing, and dynamic web applications have motivated the
-database community to implement different mechanisms for persisting data
-without the heavy-process of data modeling. Given the fact that sensor networks
-may be updated at any time by the addition of new sensor types, the use of
-moding approaches such as Key-Value Database have been proposed \cite{db-kvp}.
+The growth of distributed systems and the Internet have enabled the development
+of powerful database servers that can be organized in the context of
+infrastructure and how to represent models. An powerful up-coming approach is
+the use of database systems that does not use structured language for querying
+its data. One such example is called Key-Value Pair (KVP) data model
+\cite{db-kvp}, which is also known as Document-Oriented, Distributed Hash
+Table. Such a model has the following properties:
+
+\begin{itemize}
+  \item data is structured in collections of key-value pairs, as it is done in
+  hash data structure. The definition of the key is a given property with its
+  associating value;
+  \item the query process is using mechanics similar to programming or
+  scripting language, that is, the use of APIs in a given programming language.
+\end{itemize}
+
+There are no records of the use of this data model with Sensor Networks.
+Different variants of such data model is the document-oriented data model,
+where data is modeled using structured documents. However, these documents
+have a dynamic structure that can be freely described without the use of a
+schema validation mechanism as used in XML Schema. One example of such
+approach is the use of the JSON data format \cite{json} in the implementation
+of database systems which are becoming popular with the new trend called Cloud
+Computing \cite{cloud-comp-architectures}. Databases implementing this
+strategy is the mongoDB \cite{mongodb}.
 
@@
   \centering
-  \includegraphics[scale=0.5]{../diagrams/persistence-example-kvp}
+  \includegraphics[scale=0.75]{../diagrams/persistence-example-kvp}
   \caption{KVP Database instance Prototype}
@@
 
-The data models were compared by \cite{db-is-rdbs-dommed} and can be summarized
-as follows:
+The relational and key-value data models were compared by
+\cite{db-is-rdbs-dommed} and can be summarized as follows:
 
@@
     \begin{enumerate}
-      \item The basic operations CRUD \footnote{Create-Retrieve-Update-Update
+      \item The basic operations CRUD\footnote{Create-Retrieve-Update-Update
       database operations} data are performed using a structured language such
@@
     \begin{enumerate}
-      \item The CRUD operations are performed via API \footnote{Application
+      \item The CRUD operations are performed via API\footnote{Application
       Programming Interface} through programming languages;
@@
   \item MySQL: NO SUPPORT
-  \item Microsoft SQL: NO SUPPORT
+  \item Oracle: NO SUPPORT
   \item Berkeley TinyDB: NO SUPPORT
   \item \textbf{MongoDB}: Supports Both
-  \item \textbf{CouchDB}: Supports Both
   \item IBM DB2: NO SUPPORT
@@
 Sensor Network data can be saved in either Centralized or Distributed Systems.
-As described in  While Centralized Database Systems are easier to manage, it may
-face challenges regarding its data. For this reason, the use of partitioning or
-sharding would be an addition to persist sensor data.
+While Centralized Database Systems are easier to manage, it may
+face challenges regarding its data. In order to implement a Data-Centric
+solution, a distribute database system must be in place in order to manage
+the different data by a single node.
 
@@
   \item \textbf{MySQL}: Supports
-  \item \textbf{Microsoft SQL}: Supports
+  \item \textbf{Oracle}: Supports
   \item \textbf{Berkeley TinyDB}: NO Supports
@@
   \item \textbf{MySQL}: Supports
-  \item \textbf{Microsoft SQL}: Supports
-  \item \textbf{Berkeley TinyDB}: Supports
+  \item \textbf{Oracle}: Supports, Not Open-Source
+  \item \textbf{Berkeley TinyDB}: Supports, Open-Source
   \item \textbf{MongoDB}: Supports
   \item \textbf{CouchDB}: Supports
-  \item \textbf{IBM DB2}: Supports
+  \item \textbf{IBM DB2}: Supports, Open-Source
 \end{itemize}
@@
 \section{Global Analysis Results and Technology Selection}
+
+Each of the databases were scored according to its support to the different
+types of taxonomies, and randomly selected by using the literature review's
+list, as well as the Internet.
 
@@
         \begin{tabular}{|c|c|c|c|c|c|c|}\hline 
-        \textbf{Database} & \textbf{MySQL} & \textbf{MicrosoftSQL} & \textbf{TinyDB} &
-        \textbf{MongoDB} & \textbf{CouchDB} & \textbf{IBM DB2}\\\hline 
-        Centralized Query & + & + & + & + & + & + \\\hline 
-        Distributed System & ++ & ++ & - & ++ & ++ & +\\\hline 
-        Schema-less & - & - & - & + & + & +\\\hline 
-        Provenance Support & + & + & + & + & + & +\\\hline 
-        NO-SQL Query & - & - & - & + & + & +\\\hline 
-        Data Partition & + & + & - & + & + & +\\\hline 
-        Hot Backup & + & + & - & - & - & +\\\hline 
-        Export Capability & + & + & - & + & + & -\\\hline 
-        Programming Lang. & + & + & + & + & + & +\\\hline
-        Script Lang. & + & + & - & + & + & +\\\hline
-        Open-Source & + & - & + & + & + & -\\\hline
+        \textbf{Database} & \textbf{MySQL} & \textbf{Oracle} &
+        \textbf{TinyDB} & \textbf{MongoDB} & \textbf{IBM DB2}\\\hline 
+        Centralized Query & + & + & + & + & + \\\hline 
+        Distributed System & ++ & ++ & - & ++ & +\\\hline 
+        Schema-less & - & - & - & + & +\\\hline 
+        Provenance Support & + & + & + & + & +\\\hline 
+        NO-SQL Query & - & - & - & + & +\\\hline 
+        Data Partition & + & + & - & + & +\\\hline 
+        Hot Backup & + & + & - & - & +\\\hline 
+        Export Capability & + & + & - & + & -\\\hline 
+        Programming Lang. & + & + & + & + & +\\\hline
+        Script Lang. & + & + & - & + & +\\\hline
+        Open-Source & + & - & + & + & -\\\hline
         \end{tabular}

### 2009-11-24 docs/marcello-dissertation/thesis/chapter6-data-persistence-component-design-and-architecture.tex
Updates to chapter 2 through 7.
Chapter 2 lost the KVP section, since this approach had not been identified in
the literature for sensor networks. In this way, this makes the primary finding
and contribution of the dissertation. It will only appear during the technology
selection, what is the direct representation of what had happened;

chapter 3 adds details to the taxonomies;

Chapter 4 was cut short to describe SF-BEAMS, NetBEAMS in separate, as well as
keeps the entire implementation to the last chapter (appendix). 

Chapter 5 is the selection of the technology, but still misses highlights

@@
 summarized following the UML Business Process Diagram \cite{uml} depicted in
-figure \ref{fig:fig:dsp-persistence-business-process}. It is important to note
+figure \ref{fig:dsp-persistence-business-process}. It is important to note
 that this diagrams details remote communication between sensor devices and a
@@
 that uniquely identifies them, together with other properties such as time
-stamps of data creation and collected, as well as the the IP address
-\footnote{Internet Protocol Address} of the DSP Client that produced the data.
-Therefore, these are important properties for the proposed solution of this work.
+stamps of data creation and collected, as well as the the IP
+address\footnote{Internet Protocol Address} of the DSP Client that produced
+the data. Therefore, these are important properties for the proposed solution
+of this work.
 
@@
 
-Any configuration parameter required by a DSP Component is sent
-during its bootstrap process through the use of an optional DSP Update Message,
-as described in section \ref{sec:dsp-message}. The initialization of the DSP
-Component Activator and the DSP Component registration are also shown in figure
-\ref{fig:From-OSGi-Framework-to-DSP-Data-PersistenceActivator-Sequence-Diagram}.
+Any configuration parameter required by a DSP Component is sent during its
+bootstrap process through the use of an optional DSP Update Message, as
+described in section \ref{sec:dsp-message}. The initialization of the DSP
+Component Activator and the DSP Component registration are also shown in
+figure \ref{fig:From-OSGi-Framework-to-DSP-Data-PersistenceActivator-Sequence-Diagram}.
 
@@
 
+\begin{figure}[!b]
+  \centering
+  \includegraphics[scale=0.5]{../diagrams/DSP-Data-Persistence-Mongo-Classes}
+  \caption{UML Class diagram with Persistence Service for MongoDB}
+  \label{fig:DSP-Data-Persistence-Mongo-Classes}
+\end{figure}
+
 The DSP Data Flusher is a thread that is designed to be in two different
@@
 is delivery to the DSP Broker for its delivery to Data Consumers configured to
-receive that specific message. For this reason, figre
+receive that specific message. For this reason, Figure
 \ref{fig:From-DSP-Broker-To-DSPDataPersistence-General-Sequence} continues the
@@
 
-Next section details the implementation and deployment of the DSP Data
-Persistence component, as well as the database technology chosen in the previous
-section to provide the persistence for the collected data by the DSP Data
-Persistence component.
+\section{DSP Data Persistence Deployment}
+
+The deployment of the persistence layer is comprised of two main
+components: the DSP Data Persistence and the MongoDB Database Server. Two
+different approaches can be taken for the deployment: a single server of a
+cluster of mongoDB instances for a sharded environment.
+
+The single server approach maintains all data located in a simple space.
+Distributed Systems applications such as Master-Slave, Data Replication, etc
+can be used in this approach for scalability
+\ref{fig:DSP-Data-Persistence-Deployment-Single}.
+
+\begin{figure}[!b]
+  \centering
+  \includegraphics[scale=0.7]{../diagrams/DSP-Data-Persistence-Deployment-Single}
+  \caption{UML Deployment diagram with a single server}
+  \label{fig:DSP-Data-Persistence-Deployment-Single}
+\end{figure}
+
+However, given that mongoDB provides support to Shards, then the implementation
+of a Data-Centric approach can be done by deploying different mongoDB instances
+in a distributed cluster, where each node can hold a given type of data. The
+specification of what types of data should be stored in each of the cluster
+nodes is defined by a shard-key.
+
+\begin{figure}[!b]
+  \centering
+  \includegraphics[scale=0.7]{../diagrams/DSP-Data-Persistence-Deployment-Sharded}
+  \caption{UML Deployment Diagram for a sharded server}
+  \label{fig:DSP-Data-Persistence-Deployment-Sharded}
+\end{figure}
+
+More details about the implementation of the DSP Data Persistence can be see in
+the next chapter.

### 2009-11-24 docs/marcello-dissertation/thesis/chapter7-data-persistence-component-implementation.tex
Updates to chapter 2 through 7.
Chapter 2 lost the KVP section, since this approach had not been identified in
the literature for sensor networks. In this way, this makes the primary finding
and contribution of the dissertation. It will only appear during the technology
selection, what is the direct representation of what had happened;

chapter 3 adds details to the taxonomies;

Chapter 4 was cut short to describe SF-BEAMS, NetBEAMS in separate, as well as
keeps the entire implementation to the last chapter (appendix). 

Chapter 5 is the selection of the technology, but still misses highlights

@@
 This section shows the implementation of the DSP Data Persistence, which
-follows the specifications of the DSP Components described on section 2.4,
-along with the setup of the mongoDB, the chosen database system that handles
-document-oriented instances. The implementation of the DSP Data Persistence
-component was developed using the DSP Subversion repository at
-http://code.google.com/p/netbeams.
+follows the specifications of the DSP Components in the previous chapter,
+along with the setup of the mongoDB, the chosen database system that models
+data in a type of key-value pair model. Moreover, this chapter also describes
+the experiments conducted for the persistence.
 
-\section{DSP Deployment on OSGi Framework}
+The implementation of the DSP Data Persistence component, as well as the
+implementation of the experiments, were developed and managed using the DSP a
+reserved development branch at the Subversion \cite{subversion} repository
+located at the Google Code website. In this way, anyone can checkout the entire
+source-code from the branch
+\textit{https://netbeams.googlecode.com/svn/branches/marcello/persistence},
+as shown in Figure \ref{fig:dsp-data-persistence-dir}. In this way, the
+recommended checkout branch directory will give the remaining directory
+structure as summarized:
 
-    * Description of the OSGi MANIFEST.MF
-    * Description of the component on the config.xml
-    * Adding a matching rule on the matcher\underline{ }config.xml
-    * Adding a bootstrap message the component and database configuration
+\begin{itemize}
+  \item \textbf{NEBEAMS-DIR/}: the directories under the subdirectory
+  ``persistence'';
+  \item \textbf{DEV-VERSION}: the directory of the current development
+  version for the DSP; That is, ``\textbf{NEBEAMS-DIR}/versions/v2/''
+  \item \textbf{PERSISTENCE-DIR}: the directory of the DSP component, that is,
+  ``\textbf{DEV-VERSION}/apps/osgi-intro-bundles/dsp/DSPDataPersistence''.
+\end{itemize}
 
-    * Adding a new entry into the deployment configuration artifact config.xml;
-    * Adding a new entry into the matcher configuration artifact matcher\underline{ }config.xml;
-    * Adding an optional configuration message that sets up the new DSP Component;
-    * Adding new Java drivers that communicates with the Database system;
-    * Installing and configuring the proposed database system.
+\begin{figure}[!b]
+  \centering
+  \includegraphics[scale=0.6]{../diagrams/dsp-data-persistence-dir}
+  \caption{The DSP Data Persistence Directory Structure}
+  \label{fig:dsp-data-persistence-dir}
+\end{figure}
 
+The DSP Platform defines its own run-time directory, which organizes the
+artifacts used by the osgi-intro Framework, as well as the its own configuration
+descriptor and matcher, as described in the previous section.
+
+\begin{itemize}
+  \item \textbf{RUNTIME-DIR}: the main directory for the execution of the DSP
+Platform, with all the necessary osgi-intro-related artifacts are placed, as shown in
+Figure \ref{fig:dsp-runtime-dir}.
+\end{itemize}
+
+\begin{figure}[!t]
+  \centering
+  \includegraphics[scale=0.5]{../diagrams/dsp-runtime-dir}
+  \caption{The DSP Data Persistence Directory Structure}
+  \label{fig:dsp-runtime-dir}
+\end{figure}
+
+\section{DSP Platform Deployment}
+
+As described in the previous chapter, the DSP Data Persistence Component was
+implemented using the Java Programming language, on top of the osgi-intro Framework
+API. This section describes the setup process of the implementation, as well as
+the artifacts use during the implementation, which will be listed in the
+appendix section.
+
+The structure of the artifacts implemented for the DSP Data Persistence
+Component follow the conventions of the NetBEAMS implementation, where the main
+directory \textbf{PERSISTENCE-DIR}, depicted by Figure
+\ref{fig:dsp-data-persistence-dir-checkedout}, holds each of the development
+artifacts.
+
+\begin{itemize}
+  \item \textbf{DSPDataPersistence}: Main directory with the build.xml (Listing
+  \ref{file:dsp-build.xml}) and other Eclipse-related artifacts. The other
+  major directories are also listed:
+  \item \textbf{META-INF}: this directory contains the descriptor file for
+  the osgi-intro Framework, this this case the MANIFEST.MF;
+  \item \textbf{src}: the main directory structure for the source-code
+  implemented. Note that it follows the Java specification for packaging, and
+  therefore, lists the package org.netbeams.dsp.persistence as the main
+  package, including the layers controller, model and osgi-intro.
+\end{itemize}
+
+\begin{figure}[!h]
+  \centering
+  \includegraphics[scale=0.5]{../diagrams/dsp-data-persistence-dir-checkedout}
+  \caption{The DSP Data Persistence Directory Structure}
+  \label{fig:dsp-data-persistence-dir-checkedout}
+\end{figure}
+
+\subsection{osgi-intro Deployment Process}
+
+Since each NetBEAMS component is managed by an osgi-intro component and its
+infrastructure, this section describes the basic functionality of the osgi-intro
+platform \cite{osgi-intro}, a framework was conceived to support modularity in
+terms resources-limited environments such mobile devices and vehicles, but it
+was first widely deployed on Eclipe IDE\footnote{Integrated Development
+Environment}, because it promotes the use of reusable loosely-coupled modules
+using the Java Programming Language \cite{java}. The most basic layer of the
+osgi-intro Framework depicted in Figure \ref{fig:layering-osgi-intro} are as follows:
+
+\begin{itemize}
+  \item \textbf{Module Layer}: manages the osgi-intro bundles deployed on the osgi-intro
+  Platform, providing the necessary "wiring" of the components. In other words,
+  the modules, called osgi-intro Bundles, can export and/or import packages in the
+  level of a Java Class managed by the osgi-intro Framework;
+  \item \textbf{Service Layer}: resposible for the interoperability between 2
+  or more bundles, enabling the bundles to register services offered by
+  the its specification;
+  \item \textbf{Execution Layer}: executes the bundles, managing and changing
+  their life-cycle through the bundle execution.
+\end{itemize}
+
+\begin{figure}[!h]
+  \centering
+  \includegraphics[scale=0.5]{../diagrams/layering-osgi}
+  \caption{The osgi-intro Framework Layers}
+  \label{fig:layering-osgi-intro}
+\end{figure}
+
+NetBEAMS uses the concept of the modularization through the use of the the
+Producer-Consumer paradigm described in the previous section. Since the DSP
+Components are essentially osgi-intro bundles, the interoperability between them are
+given in the implementation and reuse existing bundles in the osgi-intro Framework
+and the DSP Platform, described by an artifact called MANIFEST.MF as seen in
+listing \cite{file:osgi-intro-manifest}. In general, an osgi-intro bundle must
+provide specifications that describes the module to be published into the osgi-intro
+Framework. In this way, the main properties of the osgi-intro MANIFEST.MF artifact can
+used by the DSP Data Persistence can be summarized as follows:
+
+\begin{itemize}
+  \item \textbf{Bundle-Activator}: The name of the instance of an osgi-intro Activator
+  class, responsible to manage the bundle. The implemented class
+  ``DataPersistenceActivator'' provides the activation mechanisms for the
+  services provided by the DSP Data Persistence Component;
+  \item \textbf{Bundle-ClassPath}: the necessary Java Jars list needed to run
+  the bundle. It lists the mongoDB Java driver as one of the required
+  dependencies deployed in the same package;
+  \item \textbf{Import-Package}: the Java Packages needed by the DSP Data
+  Persistence Component Bundle. These Java Packages must be provided by other
+  DSP components, forcing the execution to be dependent on those package.
+  They are provided through a ``Export-Package'' section in other osgi-intro
+  bundles and, as it is shown in listing \cite{file:osgi-intro-manifest}, they
+  are from different packages.
+\end{itemize}
+
+Once the osgi-intro bundle is installed into the osgi-intro Platform, it will be managed by
+the osgi-intro Execution layer and change the bundle state according to its
+lify-cycle, as described in the previous chapter. Therefore, the packaging of
+the osgi-intro source-code is done using a JAR\footnote{Java Archival Repository}
+artifact specification \cite{java-tutorial}. In this way, the DSP Data
+Platform osgi-intro bundle JAR is created by the Apache ANT build script
+\cite{apache-ant} in Listing \ref{file:dsp-build.xml}. The task
+``dsp-data-persistence.all'' compiles all the source-code created from the
+design of the previous chapter and packages everything into the artifact
+``DSPDataPersistence-x.x.jar'', where x.x is the version of the bundle. Figure
+\ref{fig:dsp-runtime-dir} shows the DSP Data Persistence file under the
+dirctory ``\textbf{RUNTIME-DIR}/deployment/''.
+
+\subsection{Adding DSP Data Persistence into DSP Platform }
+
+Once the new component was developed, the DSP Data Persistence component was
+deployed by editing two main descriptors, as well as the optional bootstrap
+message, as described in chapter 5:
+
+\begin{itemize}
+  \item \textbf{config.xml}: the DSP Data Persistence is added to the DSP
+  Platform adding an entry for the component;
+  \item \textbf{matcher\underline{ }config.xml}: the addition of the rules that
+  filters the messages to the DSP Data Persistence;
+  \item \textbf{DSPDataPersistence\underline{ }001.xml}: depicts the bootstrap
+  message for the DSP Data Persistence Component, as shown in Listing 
+  \ref{file:dsp-data-pers-bootstrap.xml}.
+\end{itemize}
+
+The first configuration artifact enables NetBEAMS to initialize contain the
+descriptor of the component, as shown in Listing \ref{file:dsp-config.xml}.
+The component is added with the highest priority, since it contains
+dependencies to other DSP Components. 
 \subsection{Starting the DSP Platform}
 
-    * command-line (Knopflerfish Framework)
-    * user interface (Knopflerfish Framework)
+Once the DSP Platform automatically installs all the components described by
+the configuration descriptor, the container verifies the dependencies and
+starts each of the DSP components of Figure \ref{fig:knopflerfish-execution},
+the osgi-intro Framework container Knopflerfish \cite{knopflerfish}.
 
-\subsection{Stopping the the DSP Platform}
-
-    * command-line (Knopflerfish Framework)
-    * user interface (Knopflerfish Framework)
+\begin{figure}[!t]
+  \centering
+  \includegraphics[scale=0.65]{../diagrams/knopflerfish-execution}
+  \caption{The execution of the Knopflerfish Container}
+  \label{fig:knopflerfish-execution}
+\end{figure}
 
@@
 
-nstallation of mongoDB
+The artifacts of the mongoDB are located in the thirdparty directory of the
+NetBEAMS resources. However, the build script from the DSP Data Persistence
+component can be executed to produce the persistence directory
+structure for NetBEAMS, as described in figure
+\ref{fig:dsp-persistence-system-dir}.
+
+\begin{figure}[!h]
+  \centering
+  \includegraphics[scale=0.5]{../diagrams/dsp-persistence-system-dir}
+  \caption{The osgi-intro Framework Layers}
+  \label{fig:dsp-persistence-system-dir}
+\end{figure}
+
+Installation of mongoDB
 - Explanation about mongoDB processes: mongod, mongos, mongo
@@
 > db.SondeDataContainer.findOne()
-{"_id" :  ObjectId( "d36f4007b7e7ac4a03c60000")  , "sensor_ip_address" : "192.168.0.136" , "message_id" : "7b6624d6-0ca1-4cba-a343-f166e88da73b" ,
-"transaction_time" : 1252845473412 , "fact_time" : 1252845346000 , "data" : {"temperature" : "45.01" , "sp_condition" : "37.6"
-, "condition" : "145.8" , "resistence" : "159.77" , "salinitude" : "0.0" , "pressure" : "0.391" , "depth" : "0.46" , "ph" : "5.64" ,
-"pH_mv" : "-62.1" , "odo_sat" : "89.7" , "odo_condition" : "59.34" , "turbidity" : "0.0" , "battery" : "9.4"}}
+{
+    "_id" : ObjectId("5d6f40078ec8074bba980900"),
+    "message_id" : "020d82e1-18b2-4fe2-800c-a0f68e22ea86",
+    "sensor" : {
+        "ip_address" : "192.168.0.91",
+        "location" : {
+            "latitude" : 37.89155,
+            "longitude" : -122.4464
+        }
+    },
+    "time" : {
+        "valid" : "Wed Nov 04 2009 02:30:12 GMT-0800 (PST)",
+        "transaction" : "Sat Nov 21 2009 03:01:33 GMT-0800 (PST)"
+    },
+    "observation" : {
+        "WaterTemperature" : 88.58,
+        "SpecificConductivity" : 180.5,
+        "Conductivity" : 167.3,
+        "Resistivity" : 491.89,
+        "Salinity" : 0.06,
+        "Pressure" : 1.109,
+        "Depth" : 2.642,
+        "pH" : 7.11,
+        "pHmV" : -42.9,
+        "Turbidity" : 0.2,
+        "ODOSaturation" : 72.4,
+        "ODO" : 10.85,
+        "Battery" : 2.7
+    }
+}
 \end{lstlisting}

### 2009-11-27 docs/marcello-dissertation/thesis/chapter10-appendix.tex
Major revision of the thesis structure, and review of chapters, moving chapter 7 to the appendix, and adding the first revision of the abstract in the preamble.
Chap 3. Reviewed the structure, misspelled words, sentences, etc;

Chap 4. Changes to the structure of the case study. The documentation regarding netbeams was pushed to the last chapter: appendix;

Chap 5. The selection of the technology follows the review with Prof. Puder. One section describes each of the technologies briefly, and details their characteristics during the analysis of each of the taxonomies;

Chap 6. Describes the imp

@@
 \section{Collected Sensor Data as the DSP Message Content}
+\label{sec:dsp-payload-implementation}
 
@@
 
+\chapter{DSP Data Persistence: Component Implementation}
+\label{sec:dsp-data-persistence-implementation}
+
+This section shows the implementation of the DSP Data Persistence, which
+follows the specifications of the DSP Components in the previous chapter,
+along with the setup of the mongoDB, the chosen database system that models
+data in a type of key-value pair model. Moreover, this chapter also describes
+the experiments conducted for the persistence.
+
+The implementation of the DSP Data Persistence component, as well as the
+implementation of the experiments, were developed and managed using the DSP a
+reserved development branch at the Subversion \cite{subversion} repository
+located at the Google Code website. In this way, anyone can checkout the entire
+source-code from the branch
+\textit{https://netbeams.googlecode.com/svn/branches/marcello/persistence},
+as shown in Figure \ref{fig:dsp-data-persistence-dir}. In this way, the
+recommended checkout branch directory will give the remaining directory
+structure as summarized:
+
+\begin{itemize}
+  \item \textbf{NEBEAMS-DIR/}: the directories under the subdirectory
+  ``persistence'';
+  \item \textbf{DEV-VERSION}: the directory of the current development
+  version for the DSP; That is, ``\textbf{NEBEAMS-DIR}/versions/v2/''
+  \item \textbf{PERSISTENCE-DIR}: the directory of the DSP component, that is,
+  ``\textbf{DEV-VERSION}/apps/osgi-intro-bundles/dsp/DSPDataPersistence''.
+\end{itemize}
+
+\begin{figure}[!b]
+  \centering
+  \includegraphics[scale=0.6]{../diagrams/dsp-data-persistence-dir}
+  \caption{The DSP Data Persistence Directory Structure}
+  \label{fig:dsp-data-persistence-dir}
+\end{figure}
+
+The DSP Platform defines its own run-time directory, which organizes the
+artifacts used by the osgi-intro Framework, as well as the its own configuration
+descriptor and matcher, as described in the previous section.
+
+\begin{itemize}
+  \item \textbf{RUNTIME-DIR}: the main directory for the execution of the DSP
+Platform, with all the necessary osgi-intro-related artifacts are placed, as shown in
+Figure \ref{fig:dsp-runtime-dir}.
+\end{itemize}
+
+\begin{figure}[!t]
+  \centering
+  \includegraphics[scale=0.5]{../diagrams/dsp-runtime-dir}
+  \caption{The DSP Data Persistence Directory Structure}
+  \label{fig:dsp-runtime-dir}
+\end{figure}
+
+\section{DSP Platform Deployment}
+
+As described in the previous chapter, the DSP Data Persistence Component was
+implemented using the Java Programming language, on top of the osgi-intro Framework
+API. This section describes the setup process of the implementation, as well as
+the artifacts use during the implementation, which will be listed in the
+appendix section.
+
+The structure of the artifacts implemented for the DSP Data Persistence
+Component follow the conventions of the NetBEAMS implementation, where the main
+directory \textbf{PERSISTENCE-DIR}, depicted by Figure
+\ref{fig:dsp-data-persistence-dir-checkedout}, holds each of the development
+artifacts.
+
+\begin{itemize}
+  \item \textbf{DSPDataPersistence}: Main directory with the build.xml (Listing
+  \ref{file:dsp-build.xml}) and other Eclipse-related artifacts. The other
+  major directories are also listed:
+  \item \textbf{META-INF}: this directory contains the descriptor file for
+  the osgi-intro Framework, this this case the MANIFEST.MF;
+  \item \textbf{src}: the main directory structure for the source-code
+  implemented. Note that it follows the Java specification for packaging, and
+  therefore, lists the package org.netbeams.dsp.persistence as the main
+  package, including the layers controller, model and osgi-intro.
+\end{itemize}
+
+\begin{figure}[!h]
+  \centering
+  \includegraphics[scale=0.5]{../diagrams/dsp-data-persistence-dir-checkedout}
+  \caption{The DSP Data Persistence Directory Structure}
+  \label{fig:dsp-data-persistence-dir-checkedout}
+\end{figure}
+
+\subsection{osgi-intro Deployment Process}
+
+Since each NetBEAMS component is managed by an osgi-intro component and its
+infrastructure, this section describes the basic functionality of the osgi-intro
+platform \cite{osgi-intro}, a framework was conceived to support modularity in
+terms resources-limited environments such mobile devices and vehicles, but it
+was first widely deployed on Eclipe IDE\footnote{Integrated Development
+Environment}, because it promotes the use of reusable loosely-coupled modules
+using the Java Programming Language \cite{java}. The most basic layer of the
+osgi-intro Framework depicted in Figure \ref{fig:layering-osgi-intro} are as follows:
+
+\begin{itemize}
+  \item \textbf{Module Layer}: manages the osgi-intro bundles deployed on the osgi-intro
+  Platform, providing the necessary "wiring" of the components. In other words,
+  the modules, called osgi-intro Bundles, can export and/or import packages in the
+  level of a Java Class managed by the osgi-intro Framework;
+  \item \textbf{Service Layer}: resposible for the interoperability between 2
+  or more bundles, enabling the bundles to register services offered by
+  the its specification;
+  \item \textbf{Execution Layer}: executes the bundles, managing and changing
+  their life-cycle through the bundle execution.
+\end{itemize}
+
+\begin{figure}[!h]
+  \centering
+  \includegraphics[scale=0.5]{../diagrams/layering-osgi}
+  \caption{The osgi-intro Framework Layers}
+  \label{fig:layering-osgi-intro}
+\end{figure}
+
+NetBEAMS uses the concept of the modularization through the use of the the
+Producer-Consumer paradigm described in the previous section. Since the DSP
+Components are essentially osgi-intro bundles, the interoperability between them are
+given in the implementation and reuse existing bundles in the osgi-intro Framework
+and the DSP Platform, described by an artifact called MANIFEST.MF as seen in
+listing \cite{file:osgi-intro-manifest}. In general, an osgi-intro bundle must
+provide specifications that describes the module to be published into the osgi-intro
+Framework. In this way, the main properties of the osgi-intro MANIFEST.MF artifact can
+used by the DSP Data Persistence can be summarized as follows:
+
+\begin{itemize}
+  \item \textbf{Bundle-Activator}: The name of the instance of an osgi-intro Activator
+  class, responsible to manage the bundle. The implemented class
+  ``DataPersistenceActivator'' provides the activation mechanisms for the
+  services provided by the DSP Data Persistence Component;
+  \item \textbf{Bundle-ClassPath}: the necessary Java Jars list needed to run
+  the bundle. It lists the mongoDB Java driver as one of the required
+  dependencies deployed in the same package;
+  \item \textbf{Import-Package}: the Java Packages needed by the DSP Data
+  Persistence Component Bundle. These Java Packages must be provided by other
+  DSP components, forcing the execution to be dependent on those package.
+  They are provided through a ``Export-Package'' section in other osgi-intro
+  bundles and, as it is shown in listing \ref{file:osgi-intro-manifest}, they
+  are from different packages.
+\end{itemize}
+
+Once the osgi-intro bundle is installed into the osgi-intro Platform, it will be managed by
+the osgi-intro Execution layer and change the bundle state according to its
+lify-cycle, as described in the previous chapter. Therefore, the packaging of
+the osgi-intro source-code is done using a JAR\footnote{Java Archival Repository}
+artifact specification \cite{java-tutorial}. In this way, the DSP Data
+Platform osgi-intro bundle JAR is created by the Apache ANT build script
+\cite{apache-ant} in Listing \ref{file:dsp-build.xml}. The task
+``dsp-data-persistence.all'' compiles all the source-code created from the
+design of the previous chapter and packages everything into the artifact
+``DSPDataPersistence-x.x.jar'', where x.x is the version of the bundle. Figure
+\ref{fig:dsp-runtime-dir} shows the DSP Data Persistence file under the
+dirctory ``\textbf{RUNTIME-DIR}/deployment/''.
+
+\subsection{Adding DSP Data Persistence into DSP Platform }
+
+Once the new component was developed, the DSP Data Persistence component was
+deployed by editing two main descriptors, as well as the optional bootstrap
+message, as described in chapter 5:
+
+\begin{itemize}
+  \item \textbf{config.xml}: the DSP Data Persistence is added to the DSP
+  Platform adding an entry for the component;
+  \item \textbf{matcher\underline{ }config.xml}: the addition of the rules that
+  filters the messages to the DSP Data Persistence;
+  \item \textbf{DSPDataPersistence\underline{ }001.xml}: depicts the bootstrap
+  message for the DSP Data Persistence Component, as shown in Listing 
+  \ref{file:dsp-data-pers-bootstrap.xml}.
+\end{itemize}
+
+The first configuration artifact enables NetBEAMS to initialize contain the
+descriptor of the component, as shown in Listing \ref{file:dsp-config.xml}.
+The component is added with the highest priority, since it contains
+dependencies to other DSP Components. 
+\subsection{Starting the DSP Platform}
+
+Once the DSP Platform automatically installs all the components described by
+the configuration descriptor, the container verifies the dependencies and
+starts each of the DSP components of Figure \ref{fig:knopflerfish-execution},
+the osgi-intro Framework container Knopflerfish \cite{knopflerfish}.
+
+\begin{figure}[!t]
+  \centering
+  \includegraphics[scale=0.65]{../diagrams/knopflerfish-execution}
+  \caption{The execution of the Knopflerfish Container}
+  \label{fig:knopflerfish-execution}
+\end{figure}
+
+\subsection{Execution Logs}
+
+\section{mongoDB Deployment}
+
+Given that Document-Oriented Model makes a good candidate persist sensors'
+properties and the recently enumerated list technologies in the previous
+article DSPDataPersistence, the open-source project called mongoDB was chosen
+for the evaluation on our case study, the netBEAMS DSP Platform.
+
+mongoDB supports storage based on collections of data, stored using BSON, a
+binary representation the JSON data representation format, including dynamic
+queries and indexing support. As it's stated in their web site, mongoDB
+"bridges the gap between key/value stores (which are fast and highly scalable)
+and traditional RDBMS systems (which are deep in functionality)".
+
+\begin{itemize}
+  \item mongoDB implements a document-oriented structure, which is similar to
+  KVP;
+  \item mongoDB is written in C++, and therefore, can is available in any major
+  platform, as well as offers a broad range of API drivers written in different
+  languages such as Java, Python, Perl and Ruby; 
+  \item mongoDB is open-source, with good community support and availability
+  through mailing lists, freenode IRC channel, and commercial support through
+  10gen company; 
+  \item mongoDB has support to distributed systems properties such as
+  Master-Slave replication, and features like Database Shards with
+  auto-sharding based on shard keys.
+\end{itemize}
+
+The artifacts of the mongoDB are located in the thirdparty directory of the
+NetBEAMS resources. However, the build script from the DSP Data Persistence
+component can be executed to produce the persistence directory
+structure for NetBEAMS, as described in figure
+\ref{fig:dsp-persistence-system-dir}.
+
+\begin{figure}[!h]
+  \centering
+  \includegraphics[scale=0.5]{../diagrams/dsp-persistence-system-dir}
+  \caption{The osgi-intro Framework Layers}
+  \label{fig:dsp-persistence-system-dir}
+\end{figure}
+
+Installation of mongoDB
+- Explanation about mongoDB processes: mongod, mongos, mongo
+- Specification of the shards, shard key, etc
+
+    * The database instance is called "netbeams";
+    * The database "netbeams" may contain different collections, categorized by
+    the Sensor Content Type, that is, depending on how the DSP Component was
+    described;
+
+\subsection{Document-Oriented Data Model}
+
+An example of the data is as folows (using the JSON syntax). Listing
+\ref{file:mongodb-ysi-data-format} follows the strategy described on Section X,
+where the entity representation is denormalized, and each of the properties of
+the sensor is repeated in each instance of the data. The fact and transaction
+times are expressed in milliseconds. The data element contains the complete
+structure of the originating sensor such as the IP address. The collection of
+instances of this document represents the sensor data type.
+
+\subsection{Starting the Database}
+
+\subsection{Stopping the Database}
+
+\subsection{Execution Logs}
+
+\section{Experiment Execution}
+
+The experiments conducted using mongoDB can be summarized by the different
+data access methods provided by the technology. Each subsection describes these
+methods.
+
+\subsection{Data Access Using Database Shell}
+
+The mongoDB client can be started by using the following command. Make sure you
+have started the mongoDB server before executing the mongoDB client.
+
+mongo netbeams | tee output\underline{ }number\underline{ }date.log
+
+Here, the iterative mongo client shell offers users to verify and navigate on a
+given database and its collections. This first section shows the connection of
+the mongo client to the database netbeams. It also highlights the query for the
+collections available. During the experiment, the SondeDataContainer?
+collection was created as related to the type from the DSP Messages for the YSI
+Sonde.
+
+The shell references to the mongoDB system can be found at
+http://www.mongodb.org/display/DOCS/dbshell+Reference
+
+\lstset{label=cmd:mongo,caption=Execution of mongo client}
+\begin{lstlisting}
+MongoDB shell version: 1.1.0-
+url: netbeams
+connecting to: netbeams
+type "help" for help
+> show collections
+SondeDataContainer
+system.indexes
+>
+> db.SondeDataContainer.count()
+1000000
+\end{lstlisting}
+
+Then, the first verification of the data integrity is regarding the number of
+elements created. Here, the first count() function on the collection returned
+1000000.
+
+An example about retrieving the first element of the collection can be done
+using the findOne() function. It will return an element instance on the
+JSONnotation.
+
+\lstset{label=cmd:mongo-findone,caption=Querying the database: one item}
+\begin{lstlisting}
+> db.SondeDataContainer.findOne()
+{
+    "_id" : ObjectId("5d6f40078ec8074bba980900"),
+    "message_id" : "020d82e1-18b2-4fe2-800c-a0f68e22ea86",
+    "sensor" : {
+        "ip_address" : "192.168.0.91",
+        "location" : {
+            "latitude" : 37.89155,
+            "longitude" : -122.4464
+        }
+    },
+    "time" : {
+        "valid" : "Wed Nov 04 2009 02:30:12 GMT-0800 (PST)",
+        "transaction" : "Sat Nov 21 2009 03:01:33 GMT-0800 (PST)"
+    },
+    "observation" : {
+        "WaterTemperature" : 88.58,
+        "SpecificConductivity" : 180.5,
+        "Conductivity" : 167.3,
+        "Resistivity" : 491.89,
+        "Salinity" : 0.06,
+        "Pressure" : 1.109,
+        "Depth" : 2.642,
+        "pH" : 7.11,
+        "pHmV" : -42.9,
+        "Turbidity" : 0.2,
+        "ODOSaturation" : 72.4,
+        "ODO" : 10.85,
+        "Battery" : 2.7
+    }
+}
+\end{lstlisting}
+
+The query based on attributes can be done using the "dot" notation, as you
+navigate through the JSON documents. Additionally, you can use the functions as
+aggregated on the result of others. The example in listing
+\ref{cmd:mongo-find-list} counts the number of documents with the key
+"data.ph" equals to "5.64".
+
+\lstset{label=cmd:mongo-find-list,caption=Execution of mongo client}
+\begin{lstlisting}
+> db.SondeDataContainer.find({"data.ph":5.64)}).count()
+1226
+\end{lstlisting}
+
+The following example is the output of the first 2 documents from the same
+previous query using the limit() function, as shown in figure
+\ref{cmd:mongo-find-limit}.
+
+\lstset{label=cmd:mongo-find-limit,caption=Query Element with specific
+projection limiting the result set size}
+\begin{lstlisting}
+> db.SondeDataContainer.find({"data.ph":5.64}).limit(2)
+{"_id" :  ObjectId( "d36f4007b7e7ac4a03c60000")  , "sensor_ip_address" : "192.168.0.136" , "message_id" : "7b6624d6-0ca1-4cba-a343-f166e88da73b"
+, "transaction_time" : 1252845473412 , "fact_time" : 1252845346000 , "data" : {"temperature" : "45.01" , "sp_condition" : "37.6" ,
+"condition" : "145.8" , "resistence" : "159.77" , "salinitude" : "0.0" , "pressure" : "0.391" , "depth" : "0.46" , "ph" : "5.64" ,
+"pH_mv" : "-62.1" , "odo_sat" : "89.7" , "odo_condition" : "59.34" , "turbidity" : "0.0" , "battery" : "9.4"}}
+{"_id" :  ObjectId( "d36f4007b7e7ac4a1fc80000")  , "sensor_ip_address" : "192.168.0.136" , "message_id" : "7b6624d6-0ca1-4cba-a343-f166e88da73b" ,
+"transaction_time" : 1252845473412 , "fact_time" : 1252845346000 , "data" : {"temperature" : "46.71" , "sp_condition" : "60.8" ,
+"condition" : "160.6" , "resistence" : "1399.4" , "salinitude" : "0.01" , "pressure" : "1.057" , "depth" : "2.485" , "ph" : "5.64" ,
+"pH_mv" : "-16.3" , "odo_sat" : "58.8" , "odo_condition" : "19.29" , "turbidity" : "0.2" , "battery" : "9.2"}}
+>
+\end{lstlisting}
+
+Other logs are located at
+http://code.google.com/p/netbeams/downloads/list
+
+\subsection{Data Access Using APIs}
+
+The mongoDB server offers different drivers to access the data, as well as the
+Web Services.
+
+\begin{itemize}
+  \item The Java tutorial on the drivers is located at
+    http://www.mongodb.org/display/DOCS/Java+Tutorial. The driver is located on
+    the NETBEAMS/versions/v2/thirdparty/mongodb directory;  
+  \item The REST API can be called from any HTTP client. At the time of the
+  editing, this feature is still under alpha version. Check the documentation
+  athttp://www.mongodb.org/display/DOCS/Http+Interface for details 
+\end{itemize}
+
+The following HTTP GET Request method returns the first 5 documents in the collection:
+http://127.0.0.1:28017/netbeams/SondeDataContainer/?limit=-5, as shown in
+listing \ref{cmd:mongo-rest-request}.
+
+\lstset{label=cmd:mongo-rest-request,caption=REST HTTP GET Request Example}
+\begin{lstlisting}
+HTTP/1.0 200 OK
+x-action:
+x-ns: netbeams.SondeDataContainer
+Content-Type: text/plain;charset=utf-8
+
+{
+  "offset" : 0,
+  "rows": [
+    { "_id" : "156f4007e4c3b74a36ed3100", "sensor_ip_address" : "192.168.0.117", "message_id" : "08b02c08-9290-4517-9a28-c6ee7e16509a",
+"transaction_time" : 1253557219486, "fact_time" : 1253557217000, "data" : { "temperature" : "31.44", "sp_condition" : "99.8", "condition" : "53.5",
+"resistence" : "1157.08", "salinity" : "0.0", "pressure" : "1.066", "depth" : "0.161", "ph" : "1.08", "pH_mv" : "-82.0", "odo_sat" : "40.3",
+"odo_condition" : "56.85", "turbidity" : "0.2", "battery" : "8.2" } } ,
+    { "_id" : "156f4007e5c3b74a37ed3100", "sensor_ip_address" : "192.168.0.117", "message_id" : "08b02c08-9290-4517-9a28-c6ee7e16509a",
+"transaction_time" : 1253557219486, "fact_time" : 1253557217000, "data" : { "temperature" : "37.83", "sp_condition" : "176.3", "condition" : "2.6",
+"resistence" : "1324.97", "salinity" : "0.01", "pressure" : "1.36", "depth" : "1.564", "ph" : "0.12", "pH_mv" : "-23.5", "odo_sat" : "104.5",
+"odo_condition" : "19.44", "turbidity" : "0.1", "battery" : "5.0" } } ,
+    { "_id" : "156f4007e5c3b74a38ed3100", "sensor_ip_address" : "192.168.0.117", "message_id" : "08b02c08-9290-4517-9a28-c6ee7e16509a",
+"transaction_time" : 1253557219486, "fact_time" : 1253557217000, "data" : { "temperature" : "74.3", "sp_condition" : "84.0", "condition" : "104.7",
+"resistence" : "4089.13", "salinity" : "0.01", "pressure" : "1.222", "depth" : "2.788", "ph" : "6.56", "pH_mv" : "-78.1", "odo_sat" : "40.0",
+"odo_condition" : "6.02", "turbidity" : "0.3", "battery" : "3.2" } },
+  "total_rows" : 3 ,
+  "query" : {} ,
+  "millis" : 0
+}
+\end{lstlisting}
+
+Data visualisation tools for mongoDB is slowly being developed by open-source
+developers. The next picture shows the database "netbeams" and the collection
+"SondeDataContainer  ?" being rendered by futon4mongodb, one of the
+open-source tools developed to visualise mongoDB data.
+
+\subsection{Visualizing Data on The Browser}
+
+Futon4Mongo is an open-source software adapted from CouchDB \ref{couchdb}.
+First, the collections list is shown. Figure
+\ref{fig:view-collections-instance-browser-futondb} shows one single collection
+for the YSI data containing 1 million objects.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=0.3]{../diagrams/view-collections-instance-browser-futondb}
+  \caption{Viewing a partial list of data using the Futon for CouchDB/MongoDB}
+  \label{fig:view-collections-instance-browser-futondb}
+\end{figure}
+
+The collection is composed by instances of documents of representing the YSI
+Sonde type, being indexed by a document ID and the values being the keys
+defined in the previous chapter, as shown in figure
+\ref{fig:view-collected-data-list-browser-futondb}.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=0.3]{../diagrams/view-collected-data-list-browser-futondb}
+  \caption{Viewing a partial list of data using the Futon for CouchDB/MongoDB}
+  \label{fig:view-collected-data-list-browser-futondb}
+\end{figure}
+
+By clicking in one of the items, one can view the instance of the document, as
+it is shown in figure \ref{fig:view-collections-instance-browser-futondb}.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=0.3]{../diagrams/view-collected-data-instance-browser-futondb}
+  \caption{Viewing an instance of collected data using the Futon for
+  CouchDB/MongoDB}
+  \label{fig:view-collected-data-instance-browser-futondb}
+\end{figure}
+
+\subsection{Exporting Data to Spreadsheets}
+
+mongoDB has an export facility shell called mongoexport. It can export the data
+in JSON format or CSV. One may also write its own export tool in any of the
+languages such as Java, PHP, Python, Perl, Ruby, among others. A list of the
+existing drivers in different languages is provided
+athttp://www.mongodb.org/display/DOCS/Drivers. The following command can be
+executed to have the exported version of the data in CSV (read the help output
+of the command for details).
+
+An example to export the data in CSV format can be seen in listing
+\ref{cmd:mongoexport}, and an example containing 1 million objects can be
+downloaded at
+http://netbeams.googlecode.com/files/experiment-1000000-data-exported-20090913-053538.csv.tar.gz.
+
+\lstset{label=cmd:mongoexport,caption=Command to export data in CSV format}
+\begin{lstlisting}
+mongoexport -d netbeams -c SondeDataContainer --dbpath ./data/ --csv -f "_id,sensor_ip_address,transaction_time,fact_time,
+data.temperature,data.sp_condition,data.condition,data.resistence,data.salinitude,data.pressure,data.depth,data.ph,data.pH_mv,data.odo_sat,
+data.odo_condition,data.turbidity,data.battery" -o sonde-data-exported.csv
+\end{lstlisting}
+
+\subsection{Exporting Data to OPeNDAP Format}
+
+By using any programming language driver, one can export data in the OPeNDAP
+format.
+
 \section{Source-Code and Implemented Artifacts}
@@
 
+\lstinputlisting[language=Java,label=file:random-ysi-data-generator,caption=Experiment
+Implementation]{../../../../netbeams/versions/v2/apps/osgi-bundles/dsp/DSPSondeProducer/src/org/netbeams/dsp/ysi/SondeTestData.java}
+
 \lstinputlisting[language=Java,label=file:experiment-dsp-implementation,caption=Experiment
@@
 {
-             "_id" :  ObjectId( "d36f4007b7e7ac4a03c60000")  , 
-             "sensor_ip_address" : "192.168.0.136", 
-             "message_id" : "7b6624d6-0ca1-4cba-a343-f166e88da73b",
-             "transaction_time" : 1252845473412 , 
-             "fact_time" : 1252845346000,
-             "latitude" : 12.3450, 
-             "longitude" : -123.4456 , 
-             "data" : {
-                       "temperature" : 45.01, 
-                       "sp_condition" : 37.6, 
-                       "condition" : 145.8, 
-                       "resistence" : 159.77, 
-                       "salinitude" : 0.0, 
-                       "pressure" : 0.391, 
-                       "depth" : 0.46, 
-                       "ph" : 5.64, 
-                       "pH_mv" : -62.1, 
-                       "odo_sat" : 89.7, 
-                       "odo_condition" : 59.34, 
-                       "turbidity" : 0.0,
-                       "battery" : 9.4
+    "_id" : ObjectId("5d6f40078ec8074bba980900"),
+    "message_id" : "020d82e1-18b2-4fe2-800c-a0f68e22ea86",
+    "sensor" : {
+        "ip_address" : "192.168.0.91",
+        "location" : {
+            "latitude" : 37.89155,
+            "longitude" : -122.4464
+        }
+    },
+    "time" : {
+        "valid" : "Wed Nov 04 2009 02:30:12 GMT-0800 (PST)",
+        "transaction" : "Sat Nov 21 2009 03:01:33 GMT-0800 (PST)"
+    },
+    "observation" : {
+        "WaterTemperature" : 88.58,
+        "SpecificConductivity" : 180.5,
+        "Conductivity" : 167.3,
+        "Resistivity" : 491.89,
+        "Salinity" : 0.06,
+        "Pressure" : 1.109,
+        "Depth" : 2.642,
+        "pH" : 7.11,
+        "pHmV" : -42.9,
+        "Turbidity" : 0.2,
+        "ODOSaturation" : 72.4,
+        "ODO" : 10.85,
+        "Battery" : 2.7
     }

### 2009-11-27 docs/marcello-dissertation/thesis/chapter3-taxonomic-classification-of-persistence-storage-for-sensor-networks.tex
Major revision of the thesis structure, and review of chapters, moving chapter 7 to the appendix, and adding the first revision of the abstract in the preamble.
Chap 3. Reviewed the structure, misspelled words, sentences, etc;

Chap 4. Changes to the structure of the case study. The documentation regarding netbeams was pushed to the last chapter: appendix;

Chap 5. The selection of the technology follows the review with Prof. Puder. One section describes each of the technologies briefly, and details their characteristics during the analysis of each of the taxonomies;

Chap 6. Describes the imp

@@
 taxonomy\footnote{Taxonomy is the practice and science of classification, by
-using taxonomic units known as taxa (singular taxon)}, which will be used
+using taxonomic units known as taxa (singular taxon)}, which were used
 throughout this work to evaluate different existing database systems
@@
 
-No matter the type of resource management used, the first taxonomy presented is
-related to the purpose of the use of the collected sensor data. That is, when
-and how the collected data is going to be used as described in section
+No matter the resource management used, the first taxonomy presented relates to
+the purpose of the use of the collected sensor data. That is, when and how the
+collected data is going to be used as described in section
 \ref{sec:sn-data-purpose}.
@@
 
-Directly related to the purpose of data, the mechanism for data storage
-described in section \ref{sec:sn-infrastructure} depends on the purpose of
-the use of the collected data. The related taxon is depicted by figure
-\ref{fig:taxonomy-data-location}, and described as follows:
+Directly related to the purpose of data, the procedure for data storage
+described in section \ref{sec:sn-infrastructure} depends on the purpose of the
+use of the collected data. Figure \ref{fig:taxonomy-data-location}, shows the
+related taxon and can be described as follows:
 
@@
 \begin{itemize}
-  \item \textbf{Local Storage}: The data can be temporarily persisted
-   in-memory, or a secondary memory device in a local storage device. It is
-  directly related to the real-time data stream taxon, since it occurs at
-  the place of the sensor device;
-  \item \textbf{External Storage}: The data is usually located in an
-  external storage device with a data management system such as a database
-  system.
-  \item \textbf{Data-Centric Storage}: the collected data can arranged in a
-  categorized fashion as described by \cite{sn-storage03}. This taxon of
-  system allows the data to be partitioned by a specific property. Nowadays,
-  the strategy is often related to database partition \cite{db-partition}
-  and database sharding approaches \cite{db-shard01} \cite{db-shard02}.
+  \item \textbf{Local Storage}: The data can be persisted in a temporarily
+  memory address, or a secondary memory device in a local storage device. It is
+  directly related to the real-time data stream taxon, since it occurs at the
+  place of the sensor device;
+  \item \textbf{External Storage}: The data is usually located in an external
+  storage devicewith a data management system such as a database system;
+  \item \textbf{Data-Centric Storage}:the collected data can organized in a
+  categorizedmanner as described by \cite{sn-storage03}. This taxon regarding
+  the system allows the data to bepartitioned by a particular property.
+  Nowadays, this strategy is often related to database partition
+  \cite{db-partition}, and database sharding approaches \cite{db-shard-intro,
+  db-shard-discussion}.
 \end{itemize}
@@
 
-One of the major challenges in the development of a persistence storage for
-sensor network is related to the data model chosen to represent the collected
-data. The addition of new sensor devices may trigger refactoring of the
+One of the fundamental challenges in the development of a persistence storage
+in sensor network is related to the data model chosen to keep the collected
+data. The introduction of new sensor devices may cause refactoring of the
 existing model or not, depending on how the data model accommodates the data.
@@
 in section \ref{sec:data-models}. For this reason, taxonomy related to the
-data model depicts two different classes of data models, as shown in figure
+data model depicts two different classes of data models, as shown in Figure
 \ref{fig:taxonomy-data-model}.
@@
 \begin{itemize}
-  \item \textbf{Schema-Dependent Models}: this type of data model require
-  the definition of a master schema that describes the data through a regorous
-  data modeling process prior to the use of the data. In addition, an addition
- to a new sensor device, and consequently a new data entity, requires a
- re-evaluation of the existing data schema. For instance, relational model is
- governed by the relational algebra model, while the structured models such as
- XML needs a definition of an XML Schema.
-  \item \textbf{Schema-less Models}: contrary to the previous taxon, 
-  schema-less data models does not require the definition of a schema or master
-  table definition and relationships. Instances of such data model is the KVP
-  and the document-oriented models, whose usage in wireless sensor networks
-  have not been reported.
+  \item \textbf{Schema-Dependent Models}: this data model requires the
+  definitionof a master schema that describes the data through a rigorous data
+  modeling process prior to the use of the data. In addition, an addition to a
+  newsensor device, and hence a new data entity, requires a re-evaluation ofthe
+  existing data schema. For example, relational model is governed by
+  therelational algebra model \cite{relational-model}, while the structured
+  models such as XML \cite{xml} needs adefinition of an XML Schema
+  \cite{xml-schema};
+  \item \textbf{Schema-less Models}: contrary to the former taxon, schema-less
+  data models do not require the definition of a data schema or a master table
+  definition, along with its table relationships, if necessary. Instances of
+  such data model is the Key-Value Pair \cite{db-kvp} and its variation
+  document-oriented models \cite{db-is-rdbs-dommed}, whose use in wireless
+  sensor networks has not been reported.
 \end{itemize}
@@
 
-No matter the type of data model used, the support for Data Provenance as shown
-in section \ref{sec:sn-provenance} is as important as the data being collected.
+No matter the data model used, the support for Data Provenance as shown in
+section \ref{sec:sn-provenance} is as valuable as the data being collected.
 For this reason, the data model taxonomy must have support for the data types
 that support the use of provenance-related attributes in the collected sensor
-data. In this way, Figure \ref{fig:taxonomy-data-provenance} describes the data types
-for the metadata support.
+data. In this way, Figure \ref{fig:taxonomy-data-provenance} describes the
+data types for the metadata support.
 
@@
   \item \textbf{What: Data Identity}: the data type that uniquely identifies
-  the collected sensor dat;
+  the collected sensor data;
   \item \textbf{When: Time Dimension}: the data type that identifies the
@@
   \item \textbf{Where: Data Location}: an optional data type that identifies
-  where the data came from.
+  from where the data originated.
 \end{itemize}
@@
 different types of query processing mechanisms described in section
-\ref{sec:query-process}, and can be seen in figure
+\ref{sec:query-process}, and can be seen in Figure
 \ref{fig:taxonomy-query-mechanism}.
@@
   \item \textbf{In-Network Query Processing}: the sensor node receives a
-  request to the properties of the sensor, and returns the current values
-  related to them \cite{sn-intro01}.
-  \item \textbf{Centralized Query Processing}: as it is done in regular single
-  server relational databases, the data is collected to a centralized data
-  sink in order to be used.
+  request to theproperties of the sensor, and returns the current values
+  related to them \cite{sn-intro01};
+  \item \textbf{Centralized Query Processing}: as developed in single server
+  relational databases, the data is collected to a centralized data sink in
+  order to be used.
 \end{itemize}
@@
 the database system used, and how its architecture provides the data storage
-solution. For example, one important problem related to the centralized
-network sink in sensor networks is the presence of an unavoidable creation of a
-point of traffic concentration at the data collector or network nodes as
-described by \cite{sn-storage02}. In this way, in order to mitigate the
-problems associated with the so-called ``Funneling Effect", figure
-\ref{fig:taxonomy-database-architecture} summarizes the different capabilities
-of different types of database systems.
+solution. For example, one crucial problem related to the centralized network
+sink in sensor networks is the presence of an ineluctable develoment of a point
+of traffic concentration at this node \cite{sn-storage02}. In this way, in
+order to reduce the problems associated with the so-called ``Funneling Effect",
+Figure \ref{fig:taxonomy-database-architecture} summarizes the different
+capabilities of different types of database systems.
 
@@
 \begin{itemize}
-  \item \textbf{Centralized System}: this is a database system that runs in
-  one single centralized host, being the main point of data traffic as read
-  and writes \cite{sn-intro01}.
+  \item \textbf{Centralized System}: a database system that runs in one single
+  centralized host, being the focal purpose of data traffic as read and writes 
+  \cite{sn-intro01};
   \item \textbf{Distributed System}: a database system that can be composed by
-  more than one node. Usually, the use of the paradigm of Master-Slave
-  mechanisms can be used by implementing ``Replication". In addition to it,
-  ``Database Sharding" is used to portion the data related to a given shard
-  key.
+  more thanone node. Usually, the use of the pattern of Master-Slave
+  mechanisms can beused by implementing ``Replication". In addition to it,
+  ``Database Sharding" is used to portion the data related to a given shard key.
 \end{itemize}

### 2009-11-27 docs/marcello-dissertation/thesis/chapter5-taxonomic-persistence-storage-selection-for-netbeams.tex
Major revision of the thesis structure, and review of chapters, moving chapter 7 to the appendix, and adding the first revision of the abstract in the preamble.
Chap 3. Reviewed the structure, misspelled words, sentences, etc;

Chap 4. Changes to the structure of the case study. The documentation regarding netbeams was pushed to the last chapter: appendix;

Chap 5. The selection of the technology follows the review with Prof. Puder. One section describes each of the technologies briefly, and details their characteristics during the analysis of each of the taxonomies;

Chap 6. Describes the imp

@@
 This chapter analyzes each of the taxonomies defined in chapter
-\ref{chap:taxonomies} against the requirements defined by NetBEAMS as our case
-study in chapter \ref{sec:problem-requirements}, following the scope of a given
-solution also defined in that chapter. The contenders were drawn from a mixed
-list of technologies either related in the literature review of chapter 2 or
-from the Internet.
+\ref{chap:taxonomies} against different other variables. First, the
+requirements defined by NetBEAMS, as our case study in chapter
+\ref{sec:problem-requirements}, is be compared. Then, the scope of this work is
+compared to the technologies selected based on the readings of the
+literature review in persistence in sensor networks and other new trends in
+performance. Finally, a table summarizes the contenders strenths and discusses
+the selected technology to be integrated with the design of a Data
+Persistence component.
+
+\section{Scope of the use of the Techonolgy}
+
+The scope of the use of the database can cover different functional and
+non-functional requirements defined in the previous chapter. This work does
+not intend to implement an application based on the database system selected,
+but provide the foundation of a database management system that can be used to
+develop one focused in less refactoring and high scalability and performance.
+For this reason, the use of the database system will be restricted to the data
+persistence and use through native system or through the use of native
+programming languages or API. In this way, the scope of this technology
+selection is can be summarized as follows:
+
+\begin{itemize}
+  \item Provide data management over the collected sensor data, as a regular
+  database system;
+  \item Must be able to scale in the face of an increase to the data load of
+  the sensor network, as described in chapter 2;
+  \item Must be able to use single or multiple servers to improve performance
+  and optionally deal with Data-Centric approaches to sensor data partitioning;
+  \item Must provide a data model that scales, decreasing the constant data
+  schema changes.
+\end{itemize}
 
@@
 This section describes the database systems contenders considered to be used
-for the implementation of Data Persistence layer for NetBEAMS. As described in
-the previous chapter, the selection of the technologies must be based on the
-list of functional and non-functional requirements specified within the scope
-of implementation defined.
+together with for the implementation of the Data Persistence component for
+NetBEAMS. As described in the previous chapter, the selection of the
+technologies must be based on the list of functional and non-functional
+requirements specified within the scope this work, defined in the previous
+section.
 
 One of the primary questions related to data persistence is regarding the data
-model and type of the database, such as the use of relational model or
+model and type of the database, that is, the use of relational model or
 structured models. In addition to that, and given the current time of academic
@@
 systems. Among other intriguing questions, \cite{db-is-rdbs-dommed} reflects
-about the use of alternatives to the relational model for dynamic
-environments such as data persistence for the Web. Similarly, these
-trends are also related to the advances of Cloud Computing and its persistence
-layer as a solution to dynamic systems \cite{cloud-comp-architectures}. For
-this reason, this work also included the overview of alternative data models
-for the development of an experiment that brings data persistence to NetBEAMS.
+about the use of alternatives to the relational model for dynamic environments
+such as data persistence for the Web. Similarly, these trends are also related
+to the advances of Cloud Computing and its persistence layer as a solution to
+dynamic systems \cite{cloud-comp-architectures}. For this reason, this work
+also included the overview of alternative data models for the development of
+an experiment that brings data persistence to NetBEAMS. In such a fashion, the
+list of the systems considered for evaluation is the following:
 
 \begin{itemize}
-  \item \textbf{MySQL}: an open-source relational database system that is the
-  primary choice of the development of projects of different sizes. It was selected
+  \item \textbf{MySQL}\cite{mysql}: an open-source relational database system
+  that is the primary choice of the development of projects of different sizes. It was selected
   because of its popularity in the current community;
-  \item \textbf{Oracle}: is the most popular commercial database system, having
-  a campus wide academic licenses availabe;
+  \item \textbf{Oracle}\cite{oracle}: is the most popular commercial database
+  system, having a campus wide academic licenses availabe;
   \item \textbf{TinyDB}\cite{tinydb}: as referred in the literature review,
@@
 
-The following sections analyzes the taxonomies defined in chapter 3, along with
-the requirements for the Data Persistence Layer for NetBEAMS, in section 4.
+These technologies will be reviewed in contrast to their capabilities to adhire
+to the specifications of the taxonomies defined in chapter 3 and the
+requirements and the scope for a data persistence of NetBEAMS, as defined in
+chapter 4.
 
@@
 
-Considering the infrastructure of NetBEAMS, as described in the previous
-chapter, and the characteristics of the sensor network defined at the SF-BEAMS
-site \cite{sfbeams2006}, the data characteristics are as follows:
+The execution of the SF-BEAMS sensor network, together with the NetBEAMS
+automated infrastructure, can be summarized as follows:
 
@@
   for distribution using the OPEnDAP format;
-  \item Another approach for data collection of data generated by the SF-BEAMS
-  is the use of NetBEAMS \cite{netbeams2009}, as described in section
-  \ref{netbeams-architecture}.
+  \item An automated approach to the data collection procedures is done
+  using NetBEAMS \cite{netbeams2009}, using software components to extract and
+  send the data to the network sink. A complete description of this approach
+  is described in section \ref{netbeams-architecture}.
 \end{itemize}
@@
 As summarized in the previous section, the nature of data from NetBEAMS through
-SF-BEAMS is used for the purpose of \textbf{Data Archival}.
+SF-BEAMS is used for the purpose of \textbf{Data Archival}. After data is
+collected from the sensors, it needs to be processed and archived for reuse by
+different types of users, ranging from researchers, students, and the public
+over the Internet.
 
@@
 
-\begin{itemize}
-  \item \textbf{MySQL}: Supports
-  \item \textbf{Oracle}: Supports
-  \item \textbf{TinyDB}: Supports
-  \item \textbf{MongoDB}: Supports
-  \item \textbf{DB2}: Supports
-\end{itemize}
+Basically, all the selected database systems support the purpose of the data.
 
-\section{Analysis of the Location of Sensor Data Taxonomy}
+\section{Analysis of the Location of Sensor Data}
 \label{sec:sn-data-location}
 
-Some observations regarding the location of the collected sensor data for
-NetBEAMS were described in the previous section and summarized as follows:
+As seen in the previous section, NetBEAMS's architecture is based on a
+single-hop star infrastructure from the SF-BEAMS sensor network, having all
+the data collected from the sensors being directly sent to the network sink 
+\ref{sec:sn-infrastructure}.
+
+Based in the observations also described in the previous chapter, the simplest
+location to store the collected data is through the use of an \textbf{External
+Storage}. In addition to that, the use of a \textbf{Data-Centric Storage}
+approach is also considered, in situations where the data load in the
+network sink reaches its storage capacity.
+
+The Data-Centric storage separates the location of storage of the data based in
+any property of the data such as the value of a given key, etc. Such approach
+of data partitioning is provided by database system in two different flavors
+\ref{db-partitioning-relational}:
 
 \begin{itemize}
-  \item NetBEAMS's architecture is based on a single-hop star infrastructure 
-   with a direct network sink being the RTC server \ref{sec:sn-infrastructure}.
+  \item \textbf{Horizontal Partitioning}: the data segmentation is done based
+  on the values of the table rows in different groups allocated to different
+  datasets physically located in different servers. In this way, all the columns
+  defined for a given entity are found on each partition. For instance,
+  horizontal partitioning can be defined as a collection five years worth of
+  historical sensor data being partitioned into five distinct servers, where
+  each server has a single year's worth of sensor data;
+  \item \textbf{Vertical Partitioning}: data is partitioned vertically in a way
+  that the columns of an entity are divied to different database tables. In
+  this way, the particular dataset is broken down to different tables. For
+  example, if the database is used to store pictures of a given sensor camera 
+  in a BLOB column, it would be the most indicated to the scenario where 
+  the images are not constantly referenced when compared to the the textual
+  data.
 \end{itemize}
-
-Based in the observations also described in the previous chapter, the type used
-for storing data can be either the \textbf{External Storage}. However, the use
-of a \textbf{Data-Centric Storage} approach can be considered. The former is the
-current approach used by the SF-BEAMS sensor network, while the latter can be
-explored in the experiments.
 
@@
 The External Storage approach is the most common way to implement persistence
-for any type of systems, including sensor networks, due to its the simplistic
-data management. Furthermore, this type of setup is easier to use of
-distributed systems approaches such as Replication to help preventing scaling
-the system.
+for any type of systems since data management is usually offered by most
+database systems out-of-the-box. Furthermore, this approach can be accompained
+with the use of Distributed Systems approaches such as Replication to help a
+system to scaling in terms of performance.
 
@@
 
-The advantage of the Data-Centric approach lies on the way that data is
-distributed into different locations. By using techniques of distributed
-database systems, the approach improves overall performance of the use of
-massive data sets since the operations of the data are directed to specific
-locations. Besides, this approach helps the network managers scale the data
-storage just by adding new data nodes as required.
+The advantage of the Data-Centric approach can be directly associated with
+database partitioning mechanisms described earlier. Since data is distributed
+into different locations, this approach the decrease the size of datasets when
+the database server performs a query processing in a database table. For this
+reason, this approach helps the database administrators to scale their
+database infrastructure and improve the database performance.
 
-In order to support this strategy, approaches such as data partitioning based
-on the entities called Database Sharding
-\cite{db-shard02}, or the partitioning based on tables
-\cite{db-table-partition}. The former is a technique used in distributed
-database systems that uses denormalized data models to improve scalability in
-terms of data storage and data location, while the latter defines partitioning
-techniques in relational databases. It is a type of database system patitioning
-where de-normalized data is spread across different other hosts, or shard
-\cite{db-shard01}. Different models of database sharding can be developed
-\cite{db-shard-schemas}.
+The horizontal or vertical partitioning strategies are also referred to as
+Database Sharding \cite{db-shard-discussion}, or simply Database partitioning
+\cite{db-table-partition} \cite{db-partitioning-relational}. Both former and
+the latter are used in regular schema-dependent models or schema-less
+models, which uses denormalized data models. There are different approaches to
+partition data \cite{db-shard-schemas} \cite{db-partitioning-relational}
+\cite{db-partitioning-relational-oracle}.
 
 \begin{itemize}
-  \item Vertical Partitioning: each feature of the system, usually represented
-  by a table in the relational model, is placed in a different server host. For
-  example, each sensor in a given network can be divided by its device type;
-  \item Range Based Partitioning: after partitioning each feature on a
+  \item \textbf{Directory Based Partitioning}: a simple lookup service is used
+  to identify which database shard to use, where the lookup table can be stored
+  in a secondary area;
+  \item \textbf{Vertical Partitioning}: each feature of the system, usually
+  represented by a table in the relational model and placed in a different
+  server host. For example, each sensor in a given network can be divided by 
+  its device type;
+  \item \textbf{Key or Hash Based Partitioning}: the value of a given key
+  property can be used as the key of the shard, or used as an input to a hash
+  function, whose output determines which database shard location the data is
+  going to be stored;
+  \item \textbf{Range-Based Partitioning}: after partitioning each feature on a
   different server host, different multiple shards, or similar hosts, can be
   added by using the range of a given key of the entity of a function. For
-  example, all sensor types in a geographic location may contain its own shard
-  based on its IP address;
-  \item Key or Hash Based Partitioning: a given value for a given property of
-  an instance of data that is chosen to be used in a hash function, whose
-  output determines which shard to use;
-  \item Directory Based Partitioning: a simple lookup service is used to
-  identify which database shard to use, where the lookup table can be stored
-  in a secondary area. Each approach for database sharding used be used
-  together on a given function depending on the needs for performance.
+  example, different sensor data can be distributed by geographic origin of
+  the data.
 \end{itemize}
 
-For example, figure \ref{fig:database-sharding-by-region} is an example of
-data storage with shards \cite{db-shard01} related the location of data, where
-each shard is assumed to be placed in different server host. Since the data
-model is denormalized, the organization of data is entirely placed in each of
-these shards. As it is shown, the shards by location can be defined by a shard
-key, as described before. The horizontal scalability is given related to
-positions, while the vertical by the shards added.
+An example of a Data-Centric storage with different shards
+\cite{db-shard-intro} can be see in Figure
+\ref{fig:database-sharding-by-region}, where data is segregated by the
+location of data. Each shard is assumed to be placed in different server hosts.
+Since the data model is denormalized, the organization of data is entirely
+placed in each of these shards in a Horizontal partition of the entity. As it is
+shown, the shards by location can be defined by a shard key.
 
@@
 
-%This data organization architecture is currently being used to manage
-%structured data that scales to petabytes or more data. One example of such use
-%is on BigTable \cite{bigtable}, which provides support to different Google
-%products with requirements on high availability and dynamic data models and
-%uses automatic sharding to its data. The BigTable project uses MapReduce
-%[ds02], a programming model designed to search huge amounts of data in
-%paralell.
-
 \subsection{Disadvantages of the Data-Centric Approach}
@@
 The data partitioning is a very restricted technology, mostly available to
-advanced users of distributed database systems. The main users of wireless
-sensor networks do not have such expertise.
+advanced users of distributed database systems. Despite being a capability
+being offered common-off-the-shelf by some popular database systems, this
+strategy requires closer database management. Some of the problems related to
+this strategy are regarding its execution \cite{db-shard-discussion}:
 
 \begin{itemize}
-  \item Referential integrity: any cross-shard queries can represent problems 
-  since the application layer is responsible to enforce data integrity, like
-  verifying constraints of foreign keys;
-  \item Rebalancing Shards: If the schema model changes for the shards, the
-  strategy of rebalancing must be done. For instance, when one of the shards 
-  outgrows more than the others. As a consequence, the database shards must be
-  rebalanced by transferring data to new locations whenever necessary. This
-  technique is starting to appear as COTS implemetations such as
-  \cite{mongodb}.
+  \item \textbf{Rebalancing Shards}: If the schema model changes for the
+  shards, the strategy of rebalancing must be done. For instance, when one of
+  the shards outgrows more than the others. As a consequence, the database
+  shards must be rebalanced by transferring data to new locations whenever
+  necessary. This technique is starting to appear as COTS implemetations such
+  as mongoDB \cite{mongodb} or MySQL \cite{mysql};
+  \item \textbf{Referential integrity}: any cross-shard queries can represent
+  problems since the application layer is responsible to enforce data
+  integrity. Examples include verifying constraints of foreign keys when the
+  partition is done using Vertical table partitioning, considering that data
+  collections are organized in separate spaces, and physically placed in
+  different shard nodes \cite{db-table-partition}.
 \end{itemize}
@@
 \begin{itemize}
-  \item \textbf{MySQL}: Supports
-  \item \textbf{Oracle}: Supports
-  \item \textbf{Berkeley TinyDB}: Supports
-  \item \textbf{MongoDB}: Supports
-  \item \textbf{CouchDB}: Supports
-  \item \textbf{DB2}: Supports
+  \item \textbf{MySQL}: Supports External or Data-Centric Storage approaches.
+  When Data-Centric Storage is used, both vertical and horizontal partitions
+  \cite{db-partitioning-relational} can be chosen, using different strategies to
+  partition the data;
+  \item \textbf{Oracle}: Supports External or Data-Centric Storage
+  approaches. Only supports vertical partitioning with the definition
+  of different types \cite{db-table-partition};
+  \cite{db-partitioning-relational-oracle} when using Data-Centric storage;
+  \item \textbf{Berkeley TinyDB}: Only supports the External Storage;
+  \item \textbf{MongoDB}: Supports External or Data-Centric Storage.
+  Using the latter one, only the horizontal partition approach is available
+  using with automatic shards \cite{db-mongo-partition};
+  \item \textbf{DB2}: Supports External and Data-Centric storage. Only supports
+  vertical table partition \cite{db-db2-partition} for the data-centric storage;
 \end{itemize}
+
+All the selected database systems supports the External Storage approach.
+However, they all differ with the support to Data-Centric Storage support.
 
@@
 As a direct result from the previous section, the use of \textbf{Centralized Query
-Processing} is indicated for NetBEAMS, given the fact that NetBEAMS contains
-a single network sink, and therefore, a centralized location of the data.
+Processing} is indicated for NetBEAMS, also given the fact that SF-BEAMS data
+goes directly to the single network sink, and therefore, a centralized location
+of the data.
 
@@
 
-Centralized data management and query processing is simpler than in in-network.
-The centralized query processing can be
+Centralized data management and query processing is simpler than the in-network
+query method. Data is verified by a simple database management system without
+the risk of data being unavailable, when the data is spread among the network
+nodes.
 
@@
 
-The creation of the so-called Funneling Affect, since the point-of-traffic is
-concentrated in the centralized system. Any data exchange will be using the
-same channel of data exchange.
+When the Centralized Query processing is used, problems may occur. Depending
+on the size of the datasets and the database technology, performance may be a
+concern. For example, this approach creates of the so-called Funneling
+Affect \cite{sn-storage04}, since the point-of-traffic is concentrated in the
+centralized system.
+
+In order to mitigate the problems originated by this query processing
+mechanism, techniques such as Data Replication can be used.
 
@@
 
-\begin{itemize}
-  \item \textbf{MySQL}: Supports
-  \item \textbf{Oracle}: Supports
-  \item \textbf{Berkeley TinyDB}: Supports
-  \item \textbf{MongoDB}: Supports
-  \item \textbf{CouchDB}: Supports
-  \item \textbf{IBM DB2}: Supports
-\end{itemize}
+All the database systems technology support a centralized query system.
 
@@
 Modeling systems, and for this reason, one of the requirements of the system is
-the use of schema-less approaches. The \textbf{Key-Value-Pair Data Model} or the
-\textbf{Document-Oriented Data Model} seems to be the simplest choices of the
+try the use of schema-less approaches. The \textbf{Key-Value-Pair Data Model}
+or the \textbf{Document-Oriented Data Model} seems to be the simplest choices of the
 models.
@@
 Considering the inception of a relational data model \cite{relational-model} and
-the use of the YSI Sonde data as the main entity in the system, let figure 
+the use of the YSI Sonde data as the main entity in the system, let Figure 
 \ref{fig:Relational-Model-Original} represent a prototype of the relational
@@
 Supposing a new type is added into the system, let the refactored
-version of the relational model be depicted in figure
-\ref{fig:Relational-Model-Addition-Modified}.
+version of the relational model be depicted in Figure
+\ref{fig:Relational-Model-Addition-Modified}. Some observations are considered
+in this scenario:
 
@@
   \item Constant data schema changes require constant database normalization
-  process, changes to structure, database management, etc;
-  \item Some research have shown that the Relational Model does not fit the
-  nature of collected data from wired or wireless sensor networks. They usually
-  does not support time-series data nor provenance.
-  \item Even project aiming at updating the relational sensor database systems
-  and SQL clauses have been proposed. 
+  process, changes to structure, database management, etc. Schema-Dependent
+  databases always require refactoring of the schema to accommodate changes to
+  the entites properties;
+  \item The approach of Relational Model might does not fit the nature of
+  collected data from wired or wireless sensor networks, since they tend to
+  change over time \cite{db-is-rdbs-dommed}. They usually does not give support
+  time-series data nor provenance well \cite{sn-provenance};
+  \item Some research have suggested changes to the SQL clauses to better
+  support time-series \cite{sn-db-newop}.
 \end{itemize}
 
-Other projects have been using the XML data models. It falls into the same
-category as the Relational Model, since the XML documents \cite{xml}, when in
-the context of a database, must be complaint to an XML Schema \cite{xml-schema}. As shown in
-section 3, this model uses XPath \cite{xml-xpath} technology for querying
-documents, although some hibrid technologies may still use SQL \cite{sql} for
-that matter. In this way, an instance of such schema dependent types can be
-seen in figure \ref{fig:persistence-example-relational}.
+Other projects have been using the XML data models for persistence. It falls
+into the same category as the Relational Model using XML documents \cite{xml},
+being complaint to an XML Schema \cite{xml-schema}. This model uses XPath
+\cite{xml-xpath} technology for querying documents, although some hibrid
+technologies may still use SQL \cite{sql} for that matter \cite{db2}. In this
+way, a materialized version of the data from the defined previous schema can be
+seen in Figure \ref{fig:persistence-example-relational}.
 
@@
 \cite{db-kvp-in-relational02}. Based in these articles, consider the relational
-model depicted in figure \ref{fig:KVP-on-Relational-Model} as model for the
-persistence data for our case study using the key-value strategy. The problem
+model depicted in Figure \ref{fig:KVP-on-Relational-Model} as a data model for
+persisting data to our case study using the key-value strategy. The problem
 with it is that key repetition occurs, since the nature of time-series data
@@
 \begin{itemize}
-  \item entities from the same domains are placed into a bucket;
+  \item entities from the same domains are placed into a ``bucket'';
   \item entities have a set of attributes and relating values
@@
 
+Since the selected data model is the schema-less one, the following can be see
+from the technologies:
+
 \begin{itemize}
-  \item MySQL: NO SUPPORT
-  \item Oracle: NO SUPPORT
-  \item Berkeley TinyDB: NO SUPPORT
-  \item \textbf{MongoDB}: Supports Both
-  \item IBM DB2: NO SUPPORT
+  \item MySQL: Only supports the Relational Model;
+  \item Oracle: Only supports Relational Model;
+  \item Berkeley TinyDB: Only supports the Relational Model; 
+  \item \textbf{MongoDB}: Supports Document-Oriented Model;
+  \item IBM DB2: Only Supports the Relational Model.
 \end{itemize}
@@
 \begin{itemize}
-  \item \textbf{MySQL}: Supports
-  \item \textbf{Oracle}: Supports
-  \item \textbf{Berkeley TinyDB}: NO Supports
-  \item \textbf{MongoDB}: Supports
-  \item \textbf{CouchDB}: Supports
-  \item \textbf{IBM DB2}: Supports
+  \item \textbf{MySQL}: Supports Data Replication;
+  \item \textbf{Oracle}: Supports Data Replication;
+  \item \textbf{Berkeley TinyDB}: NO Supports to distributed data;
+  \item \textbf{MongoDB}: Supports Data Replication;
+  \item \textbf{IBM DB2}: Supports Data Replication.
 \end{itemize}
@@
 \begin{itemize}
-  \item Given the fact of the nature of the project, the technology to be used
-  must be \textbf{open-source} \cite{open-source}, that is, no costs involved
-  in the adoption of such technology;
-  \item A solution for persisting collected sensor data must be not only limited
-  to a technology that provides data access, such as SQL, but also by other
-  \textbf{data access mechanisms} such as through \textbf{native APIs}. The
-  reason is that part of the sensor network data does not have skills in
-  database models, but may have skills with programming languages;
-  \item Supports hot backup without service interruption.
+  \item \textbf{Open-source}: Given the fact of the nature of the project, the
+  technology to be used must be free of cost \cite{open-source}, that is, no
+  costs involved in the adoption of such technology; Together with it is the
+  support from the community;
+  \item \textbf{Native APIs}: In order to cover the scope of this work defined
+  in the previous chapter, the solution for persisting collected sensor data
+  must be not only limited to a technology that provides data access, such as
+  SQL, but also by other \textbf{data access mechanisms};
+  \item \textbf{Hot Backup}: Supports hot backup with less service interruption. 
 \end{itemize}
@@
 use of drivers in different programming and scripting languages such as Java,
-Python, Perl, etc.
+Python, Perl, etc. Moreover, only TinyDB does not support hot backup.
 
 \begin{itemize}
-  \item \textbf{MySQL}: Supports
-  \item \textbf{Oracle}: Supports, Not Open-Source
-  \item \textbf{Berkeley TinyDB}: Supports, Open-Source
-  \item \textbf{MongoDB}: Supports
-  \item \textbf{CouchDB}: Supports
-  \item \textbf{IBM DB2}: Supports, Open-Source
+  \item \textbf{MySQL}: Is open-source, supports hot back and contains
+  scripting APIs from the community;
+  \item \textbf{Oracle}: It is not Open-Source, but contains lots of support
+  from the community. Also APIS are available for the majority of languages;
+  \item \textbf{Berkeley TinyDB}: Not open-source;
+  \item \textbf{MongoDB}: An open-source database system with an increasing
+  support from the community, offering native APIS in different programming
+  languages and scripting languages;
+  \item \textbf{IBM DB2}: Not Open-Source, but offers support from its
+  community.
 \end{itemize}
+
+Only MySQL and MongoDB supports the non-functional requirements defined in this
+section.
 
@@
 types of taxonomies, and randomly selected by using the literature review's
-list, as well as the Internet.
+list, as well as the Internet. Since mongoDB is based on a schema-less database
+system, and supports most of the taxonomies defined in the previous chapter.
 

### 2009-11-27 docs/marcello-dissertation/thesis/chapter6-data-persistence-component-design-and-architecture.tex
Major revision of the thesis structure, and review of chapters, moving chapter 7 to the appendix, and adding the first revision of the abstract in the preamble.
Chap 3. Reviewed the structure, misspelled words, sentences, etc;

Chap 4. Changes to the structure of the case study. The documentation regarding netbeams was pushed to the last chapter: appendix;

Chap 5. The selection of the technology follows the review with Prof. Puder. One section describes each of the technologies briefly, and details their characteristics during the analysis of each of the taxonomies;

Chap 6. Describes the imp

@@
 
-The deployment of the persistence layer is comprised of two main
-components: the DSP Data Persistence and the MongoDB Database Server. Two
-different approaches can be taken for the deployment: a single server of a
-cluster of mongoDB instances for a sharded environment.
+The deployment of the persistence layer is comprised by the addition of two
+new components to the DSP Platform: the DSP Data Persistence and the MongoDB
+Database Server. The former is responsible for capturing any data
+produced by sensors in the data sink, as described in the previous sections of
+this chapter. On the other hand, the latter is where the data is going to be
+stored, as a result of the technology selection on the previous chapter.
+
+The deployment of the DSP Data Persistence follows the specifications of how to
+deploy a DSP Component on an OSGi Framework. For the development of this work,
+the Knopflerfish OSGi Container was used, as described in section
+\ref{sec:dsp-data-persistence-implementation}. In contrast, the
+deployment of the mongoDB can be accomplished in different approaches, as
+summarized:
+
+\begin{itemize}
+  \item \textbf{Single Server}: a regular deployment of a database system as it
+  is done in general applications, where data is managed by one single process;
+  \item \textbf{Sharded Server}: a deployment approach where data is sharded
+  into different servers.
+\end{itemize}
 

### 2009-11-27 docs/marcello-dissertation/thesis/chapter7-data-persistence-component-implementation.tex
Major revision of the thesis structure, and review of chapters, moving chapter 7 to the appendix, and adding the first revision of the abstract in the preamble.
Chap 3. Reviewed the structure, misspelled words, sentences, etc;

Chap 4. Changes to the structure of the case study. The documentation regarding netbeams was pushed to the last chapter: appendix;

Chap 5. The selection of the technology follows the review with Prof. Puder. One section describes each of the technologies briefly, and details their characteristics during the analysis of each of the taxonomies;

Chap 6. Describes the imp

--- 2009-11-15 docs/marcello-dissertation/thesis/chapter7-data-persistence-component-implementation.tex
@@
 %\bibliography{thesis}  % uses the references stored in Chapter1Radar.bib
-
-\chapter{DSP Data Persistence: Component Implementation}
-
-This section shows the implementation of the DSP Data Persistence, which
-follows the specifications of the DSP Components described on section 2.4,
-along with the setup of the mongoDB, the chosen database system that handles
-document-oriented instances. The implementation of the DSP Data Persistence
-component was developed using the DSP Subversion repository at
-http://code.google.com/p/netbeams.
-
-\section{DSP Deployment on OSGi Framework}
-
-    * Description of the OSGi MANIFEST.MF
-    * Description of the component on the config.xml
-    * Adding a matching rule on the matcher\underline{ }config.xml
-    * Adding a bootstrap message the component and database configuration
-
-    * Adding a new entry into the deployment configuration artifact config.xml;
-    * Adding a new entry into the matcher configuration artifact matcher\underline{ }config.xml;
-    * Adding an optional configuration message that sets up the new DSP Component;
-    * Adding new Java drivers that communicates with the Database system;
-    * Installing and configuring the proposed database system.
-
-\subsection{Starting the DSP Platform}
-
-    * command-line (Knopflerfish Framework)
-    * user interface (Knopflerfish Framework)
-
-\subsection{Stopping the the DSP Platform}
-
-    * command-line (Knopflerfish Framework)
-    * user interface (Knopflerfish Framework)
-
-\subsection{Execution Logs}
-
-\section{mongoDB Deployment}
-
-Given that Document-Oriented Model makes a good candidate persist sensors'
-properties and the recently enumerated list technologies in the previous
-article DSPDataPersistence, the open-source project called mongoDB was chosen
-for the evaluation on our case study, the netBEAMS DSP Platform.
-
-mongoDB supports storage based on collections of data, stored using BSON, a
-binary representation the JSON data representation format, including dynamic
-queries and indexing support. As it's stated in their web site, mongoDB
-"bridges the gap between key/value stores (which are fast and highly scalable)
-and traditional RDBMS systems (which are deep in functionality)".
-
-\begin{itemize}
-  \item mongoDB implements a document-oriented structure, which is similar to
-  KVP;
-  \item mongoDB is written in C++, and therefore, can is available in any major
-  platform, as well as offers a broad range of API drivers written in different
-  languages such as Java, Python, Perl and Ruby; 
-  \item mongoDB is open-source, with good community support and availability
-  through mailing lists, freenode IRC channel, and commercial support through
-  10gen company; 
-  \item mongoDB has support to distributed systems properties such as
-  Master-Slave replication, and features like Database Shards with
-  auto-sharding based on shard keys.
-\end{itemize}
-
-nstallation of mongoDB
-- Explanation about mongoDB processes: mongod, mongos, mongo
-- Specification of the shards, shard key, etc
-
-    * The database instance is called "netbeams";
-    * The database "netbeams" may contain different collections, categorized by
-    the Sensor Content Type, that is, depending on how the DSP Component was
-    described;
-
-\subsection{Document-Oriented Data Model}
-
-An example of the data is as folows (using the JSON syntax). Listing
-\ref{file:mongodb-ysi-data-format} follows the strategy described on Section X,
-where the entity representation is denormalized, and each of the properties of
-the sensor is repeated in each instance of the data. The fact and transaction
-times are expressed in milliseconds. The data element contains the complete
-structure of the originating sensor such as the IP address. The collection of
-instances of this document represents the sensor data type.
-
-\subsection{Starting the Database}
-
-\subsection{Stopping the Database}
-
-\subsection{Execution Logs}
-
-- Java API implementation
-- Python API implementation
-
-\section{Data Access Using Database Shell}
-
-The mongoDB client can be started by using the following command. Make sure you
-have started the mongoDB server before executing the mongoDB client.
-
-mongo netbeams | tee output\underline{ }number\underline{ }date.log
-
-Here, the iterative mongo client shell offers users to verify and navigate on a
-given database and its collections. This first section shows the connection of
-the mongo client to the database netbeams. It also highlights the query for the
-collections available. During the experiment, the SondeDataContainer?
-collection was created as related to the type from the DSP Messages for the YSI
-Sonde.
-
-The shell references to the mongoDB system can be found at
-http://www.mongodb.org/display/DOCS/dbshell+Reference

### 2009-11-27 docs/marcello-dissertation/thesis/chapter8-experimental-results-correct-behavior-and-performance.tex
Major revision of the thesis structure, and review of chapters, moving chapter 7 to the appendix, and adding the first revision of the abstract in the preamble.
Chap 3. Reviewed the structure, misspelled words, sentences, etc;

Chap 4. Changes to the structure of the case study. The documentation regarding netbeams was pushed to the last chapter: appendix;

Chap 5. The selection of the technology follows the review with Prof. Puder. One section describes each of the technologies briefly, and details their characteristics during the analysis of each of the taxonomies;

Chap 6. Describes the imp

@@
 
-\section{Design Model For Sensor Networks}
+In order to evaluate the mongoDB, the different taxonomies identified are
+verified from the logs and experiment setup. The implementation of the Data
+Persistence Component and the mongoDB infrastructure are described in section
+\ref{sec:dsp-data-persistence-implementation}.
 
-Sensor Networks are specialized types of network systems comprised of network 
-nodes with specific goals to observe and collect data. The following
-subsections describe classes of characteristis of these types of sensor
-networks, with applications on research, industrial or militery.
+The purpose of Sensor Data is used for Data Archival, since the data from the
+sensor data categorized by the date and time it was collected. The location
+of the collected data from the sensor devices are stored using the External
+Storage, or using the Data-Centric Strategy. In this way, one of the advantages
+of the denormalized data model provided by mongoDB is that the data can be
+distributed using Horizontal Partitioning.
 
-\subsection{Strengths as compared to literature}
+\section{Experiment Setup}
 
-\subsection{Weaknesses as compared to literature}
+Experiments were conducted to evaluate mongoDB against the taxonomies defined in
+chapter 3 by creating a general shell script
+\ref{file:experiment-setup-executor} that sets up the environment by
+initializing the mongoDB server and indirectly inserts the a desired number of
+randomly generated YSI Data types \ref{file:random-ysi-data-generator}
+implemented in Java as part of the NetBEAMS data collection mechanism as shown
+in section \ref{sec:dsp-payload-implementation}. First, the definition of a 
+document will be conducted according to the definition of the Data Provenance
+definition of chapter 3. Then, the infrastructure of the database will be
+evaluated using different scenarios.
 
-\section{Data-Centric Storage and Database Shards}
+\subsection{Key-Value definition}
 
-Horizontal scalability for data based on specific keys.
+In document-oriented databases, there is no definition of table. In constrast,
+the entities are persisted using any data arrangment, in any order, attributes
+and the associated values. The definition of the document for mongoDB takes
+into account the properties defined by the data provenance study defined by
+\cite{sn-provenance}. The following list are the definitions of the values of
+the defined keys and the provenance-aware document, as an example described by
+Listing \ref{file:mongodb-ysi-data-format}:
 
-\subsection{Scalability facing High Volume of Collected Data}
+\begin{itemize}
+  \item \textbf{What: Data Identity}: the identification of the data captured
+  from the sensor, as well as the DSP Platform message. The keys are comprised
+  by the attributes ``\underline{ }id'' as an unique identifier provided by
+  mongodb, the ``message\underline{ }id'';
+  \item \textbf{When: Time Dimension}: the time dimensions give information
+  regarding when the data was collected and transferred to the network sink.
+  They are the ``time.valid'', ``time.transaction'';
+  \item \textbf{Where: Data Location}: the location where 
+ the data originated. The ``sensor.ip\underline{ }address'',
+ ``sensor.location.latitude'', ``sensor.location.longitude'' give information
+ regarding the .
+\end{itemize}
 
-Adding new shards as new hardware into the cluster, facing increasing data load
-produced by the sensor network.
+\subsection{Workload}
 
-\subsection{Weakness}
+The workload selected for the experiments reflacts the current use of the data
+at the SF-BEAMS infrastructure. A number of data worth a year of collected data
+from one single YSI device is randomly generated and inserted at the database.
+Then, five additional run will be performed and investigate different search
+scenarios for performance.
 
-Technological, and skills-based.
+In order to verify the system capacity to scale, the data will be devided into
+different shards as part of the support to distribute system, that is, in a
+Data-Centric way.
+
+\begin{itemize}
+  \item First Round: 483,840 items worth one year of produced YSI data at the
+  rate of 1 observation per minute;
+  \item 483,840 * 5, or 2,419,200, representing 5 different YSI components
+  producing data every minute;
+\end{itemize}
+
+\subsection{Single Server Environment}
+
+The definition for the single server follows a regular specification of the
+mongoDB database system. All the collected data from sensors are stored in a
+given Single Server, which characterizes an External Storage device.
+
+\subsection{Data-Centric Environment}
+
+The definition of a Data-Centric deployment follows the specification of the
+mongoDB shard servers, where in contrast to the single server, all the
+collected data is spread out to different mongoDB clients that manage
+partitions of servers on a cluster.
+
+As described in the taxonomy reviews, the definition of the sharded system must
+include the definition of a shard key, which is responsible for partitioning
+the instances. In this way, the shard key is defined as the value of the
+attribute ``sensor.id\underline{ }address''.
+
+\subsection{Scenarios}
+
+In order to verify the feasibility of the system, the use of different search
+scenarios are used:
+
+\begin{itemize}
+  \item A) Find data by specific date ranges on either the valid or the
+  transaction time;
+  \item B) Find data by specific IP address, as it identifies a given sensor
+  device and any data produced;
+  \item C) Find data by specific values of the observed attributes of the sensor
+  device such as temperature, salinity, etc.
+\end{itemize}
+
+\section{Measurements Results}
+
+The result of the insertion of data in different scenarios are shown in the
+different logs captured during the experiments. First, the avarage insertion
+times gives the capacity of the system with and without the definition of
+indexes over the data. Then, the numbers regarding the scenarios described
+earlier will be shown:
+
+\subsection{Single Server Environment}
+
+The insersion time avarages are regarding the different approaches to model a
+document in the database system:
+
+\begin{itemize}
+  \item No Index : Insert Avarage = xyz 
+  \item Indexed : Insert Avarage = xyz
+\end{itemize}
+
+On the other hand, the avarages for each of the scenarios are devided by the
+amount of data produced by one single device or five different devices.
+
+For one single device, the values can be summarized as follows:
+\begin{itemize}
+  \item Not Index : A) = xyz
+  \item Indexed : A) = xzy
+  \item Not Index : B) = xyz
+  \item Indexed : B) = xzy
+  \item Not Index : C) = xyz
+  \item Indexed : C) = xzy
+\end{itemize}
+
+For data worth five different devices, the values can be summarized as follows:
+\begin{itemize}
+  \item Not Index : A) = xyz
+  \item Indexed : A) = xzy
+  \item Not Index : B) = xyz
+  \item Indexed : B) = xzy
+  \item Not Index : C) = xyz
+  \item Indexed : C) = xzy
+\end{itemize}
+
+\subsection{Data-Centric Cluster Environment}
+
+The insertions in a sharded environment are added the operation of finding
+which ``bucket'' the data must be placed. mongoDB gives support for automatic
+sharding after the definition of a key based on range of the values. The
+avarages of the insertions can be shown as follows:
+
+\begin{itemize}
+  \item NO Index : Insert Avarage = xyz
+  \item Indexed : Insert Avarage = xyz
+\end{itemize}
+
+For one single device, the values can be summarized as follows:
+
+\begin{itemize}
+  \item Not Index : A) = xyz
+  \item Indexed : A) = xzy
+  \item Not Index : B) = xyz
+  \item Indexed : B) = xzy
+  \item Not Index : C) = xyz
+  \item Indexed : C) = xzy
+\end{itemize}
+
+For data worth five different devices, the values can be summarized as follows:
+
+\begin{itemize}
+  \item Not Index : A) = xyz
+  \item Indexed : A) = xzy
+  \item Not Index : B) = xyz
+  \item Indexed : B) = xzy
+  \item Not Index : C) = xyz
+  \item Indexed : C) = xzy
+\end{itemize}
+
+\section{Discussion}
+
+mongoDB is that the data can be distributed using Horizontal Partitioning,
+which mitigates problems related to referencial integrity.
+
+\begin{itemize}
+  \item Strengths as compared to literature
+  \item Weaknesses as compared to literature
+\end{itemize}
+
+Data-Centric Storage and Database Shards = Horizontal scalability for data
+based on specific keys = Scalability facing High Volume of Collected Data
+
+Weakness, Technological, and skills-based.

### 2009-11-27 docs/marcello-dissertation/thesis/preamble.tex
Updating the thesis and added chapters 2-5. Still revisions are needed in section 4 and 5, adjusting images and tables. Images are in the eps format.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@642 559edde2-a4d4-44e0-96fc-4035e40d7396
Major revision of the thesis structure, and review of chapters, moving chapter 7 to the appendix, and adding the first revision of the abstract in the preamble.
Chap 3. Reviewed the structure, misspelled words, sentences, etc;

Chap 4. Changes to the structure of the case study. The documentation regarding netbeams was pushed to the last ch

@@
 
+%\definecolor{darkgreen}{named}{green}
+%\definecolor{darkblue}{named}{blue}
+%\definecolor{darkred}{named}{red}
+%\definecolor{grau}{named}{gray}
+%\let\Righttorque\relax
+%\lstset{
+%commentstyle=\itshape\color{darkgreen},
+%keywordstyle=\bfseries\color{darkblue},
+%stringstyle=\color{darkred},
+%extendedchars=true,
+%basicstyle=\scriptsize\ttfamily,
+%basicstyle=\tiny\ttfamily,
+%tabsize=2,
+% keywordstyle=\textbf,
+%commentstyle=\color{grau},
+% stringstyle=\textit,
+%numbers=left,
+%numberstyle=\tiny,
+% für schönen Zeilenumbruch
+%breakautoindent = true,
+%breakindent = 2em,
+%breaklines = true,
+%postbreak = ,
+%prebreak = \raisebox{-.8ex}[0ex][0ex]{\ensuremath{\lrcorner}},
+%prebreak = \raisebox{-.8ex}[0ex][0ex]{\Righttorque},
+%showspaces=false, % Keine Leerzeichensymbole
+%showtabs=false, % Keine Tabsymbole
+%showstringspaces=false,% Leerzeichen in Strings
+%}
+
 \usepackage{caption}
@@
 %Hifens
-
 \usepackage[latin1]{inputenc}
 \usepackage[USenglish]{babel}
+
+%Harvard citation style
+\usepackage{natbib}
+%enhanced items
+% http://dante.ctan.org/tex-archive/macros/latex/contrib/enumitem/enumitem.pdf
+\usepackage{enumitem}
 
@@
 \onehalfspacing{\noindent      
-Abstract goes here
+Sensor Networks are becoming more and more important to the different
+sciences communities bacause it provides the raw data for the different studies
+related to a given expertise area. In order to make use of data, sensors are
+directly or indirectly interrogated by using secondary storage devices, where
+they are stored using different data models divided by schema-dependent or
+schema-less models. In this way, this work summarizes the current state of the
+art in data persistence for sensor networks by using taxonomies, and conducts an
+empirical evaluation for a technology to be used in experiments. Moreover, this
+work shows a novice use of a Key-Value Pair database system by an
+implementation of a software component that is easily attached to an existing
+sensor network platform. When it comes scalability, this work suggests the use
+of Horizontal database partitioning in order to provide a Data-Centric
+storage approach to persist data collected from an existing real-world sensor
+network. Finally, it also contributes with the necessary background
+documentation of the software used, among others.
+
 }
-
 \vspace*{\fill}

### 2009-11-29 docs/marcello-dissertation/thesis/chapter4-netbeams-in-depth-requirements-analysis-and-specification.tex
Major revision of the thesis structure, and review of chapters, moving chapter 7 to the appendix, and adding the first revision of the abstract in the preamble.
Chap 3. Reviewed the structure, misspelled words, sentences, etc;

Chap 4. Changes to the structure of the case study. The documentation regarding netbeams was pushed to the last chapter: appendix;

Chap 5. The selection of the technology follows the review with Prof. Puder. One section describes each of the technologies briefly, and details their characteristics during the analysis of each of the taxonomies;

Chap 6. Describes the imp

@@
 
-This chapter presents the main motivation and scope of of this dissertation.
-The infrastructure of SF-BEAMS is presented as a real-world environmental
-sensor network, deployed by the department of Biology at the San Francisco State
-University. It serves as the foundation of the execution of NetBEAMS, a
-joint project between the department of Biology and the Computer Science.
-As an alternative approach for data collection, NetBEAMS is used to
-automate the data management and collection process of SF-BEAMS. However, it
-offered different open problems, including Data Persistence. Therefore, the
-scope of the solution presented in the following sections and chapters are
-based in the case study presented in this current chapter.
+This chapter presents the primary motivation and scope of this paper. The
+infrastructure of SF-BEAMS is presented as a real-world environmental sensor
+network, deployed by the department of Biology at the San Francisco State
+University. It serves as the foundation of the execution of NetBEAMS, a joint
+venture between the department of Biology and the Computer Science. As an
+alternative approach for data collection, NetBEAMS is used to automate the data
+management and collection process of SF-BEAMS. However, it offered different
+open problems, including Data Persistence. Therefore, the realm of the solution
+presented in the following sections and chapters are based in the case study
+presented in this current chapter.
 
@@
 The San Francisco Bay Environmental Assessment and Monitoring Station, or
-SF-BEAMS \cite{sfbeams2006}, is a marine sensor network operated by the Romberg
-Tiburon Center (RTC), whose main focus is the study of complex marine and
-estuarine environments using the SF-BEAMS sensors. The network is deployed
-offshore of its pier located on the banks of the San Francisco Bay at the
-Tiburon Island, California, being operated by the department of Biology from
-San Francisco State University.
+SFBEAMS \cite{sfbeams2006}, is a marine sensor network operated by the Romberg Tiburon
+Center (RTC), whose primary focus is the study of complex marine and estuarine
+environments using the SF-BEAMS sensors. The network is deployed offshore of
+its pier located on the banks of the San Francisco Bay at the Tiburon Island,
+California, being operated by the department of Biology from San Francisco
+State University.
 
@@
 monitoring, as well as weather and surface conditions. Figure
-\ref{fig:sf-beams} shows a picture taken from the SF-BEAMS web-cam placed
-in-site.
+\ref{fig:sf-beams} shows a picture taken from the SF-BEAMS web-camera.
 
@@
 
-In general, the SF-BEAMS network infrastructure contains heterogeneous wired and
-wireless devices standing on pilons docked in the pier, where each of them is
-responsible for observing different conditions of the area, as well as having
-its own mechanisms for internal storage for the collected data. In this way,
-data can be directly transferred to the labs via ethernet cables or collected
-manually by a laptop computer.
+In general, the SF-BEAMS network infrastructure contains varying wired and
+wireless devices standing on pylons docked at the pier. Each of them is
+responsible for observing different conditions of the area, having its own
+mechanisms for internal storage for the collected data. In this way, data can
+be directly transferred to the labs via Ethernet cables or collected manually
+by a laptop computer.
 
@@
 
-As described in section \ref{sec:sn-infrastructure}, sensor devices generate
+As described in section \ref{sec:sn-infrastructure}, sensor devices produce
 its observed data based on properties of measurements defined by its
 manufacturer. SF-BEAMS can be seen a single-hop star sensor network containing
-different instances of sensors, including the "YSI 6600 ESD V2", as seen on
+different instances of sensors, including the ``YSI 6600 ESD V2", as seen in
 figure \ref{fig:ysi-device}, which is a powerful water quality monitoring
@@
 
-In order to have access to the observed data, the network staff downloads
-the data using one of the device's connection such as the RS-232 serial
-connector \cite{rs232}. Then, the data is transferred to the RTC labs for
-archival, where it is described, indexed and distributed to its main users.
-According to one of its staff members, the infrastructure of the SF-BEAMS in
-May 2009 was comprised of the following:
+In order to have access to the observed data, the network staff downloads the
+data using one of the device's connection such as the RS-232 serial connector
+\cite{rs232}. Then, the RTC labs receives the collected data transferred from
+the sensor devices to be used for archival. The data is described, indexed and
+distributed to its main users. According to one of its staff members, the
+infrastructure of the SF-BEAMS in May 2009 was comprised of the following:
 
@@
   15 minutes, depending on the specification;
-  \item Data for final users are published with Provenance Data regarding time,
-  but no metadata regarding the origin of data.
+  \item Data for final users is published with Provenance Data regarding time,
+  without description of the origin of data.
 \end{itemize}
@@
 Considering the infrastructure previously described, the size of the data
-produced by the YSI sonde sensors at the RTC pier can estimated by its
+produced by the YSI sonde sensors at the RTC pier can be estimated by its
 measurement frequency as follows:
@@
 As it is shown, the data produced by this type of sensor runs in the order of
-at most hundreds of Megabytes a year for a maximun of \textbf{483,840} samples,
-what can be considered a very low data storage requirement as compared to
-other networks. Upon collecting the data from sensors, the RTC lab staff use
-automation scripts written in Matlab to process, index and distribute the raw
-data in different formats, including the OPeNDAP \cite{opendap}, a format
-widely used in the research institutions that promotes easier data exchange
-among them. The access to the archived data is done by using the OPeNDAP
-server is located at http://sfbeams.sfsu.edu:8080/opendap, as a request and
-the related output for one of the YSI devices can be see in Listing
-\ref{file:rtc-ysi-opendap}.
+at most hundreds of Megabytes a year for a maximun of \textbf{483,840}
+samples, what can be considered a exceptionally sparse data storage
+requirement as compared to other networks. Upon collecting data from sensors,
+the RTC staff use automation scripts written in Matlab to process, index and
+distribute the raw data in different formats. One of these formats is the
+OPeNDAP \cite{opendap}, a format widely used at research institutions that
+promote easier data exchange among them. The access to the archived data is
+done by using the OPeNDAP server is located at
+http://sfbeams.sfsu.edu:8080/opendap, as a request and the corresponding
+output for one of the YSI devices can be seen in Listing \ref{file:rtc-ysi-opendap}.
 
@@
 
-In order to distinct each of the properties of the SF-BEAMS sensor
-networks, consider the following list as a classification of the data
-persistence mechanisms from the sensor network used by the project:
+In order to distinct each of the properties of the SF-BEAMS sensor networks,
+consider the following list as a classification of the data persistence
+mechanisms from the sensor network used by the project:
 
 \begin{itemize}
-  \item \textbf{The Purpose of Sensor Data}: the data from SF-BEAMS is purely
-  used for \textbf{Data Archival} purposes, since they are stored in the RTC's
-  lab as part of the collaboration research project that monitors each part of
-  the California's coast. The data sets are distributed by specific date in a
-  yearly calendar;
+  \item \textbf{The Purpose of Sensor Data}: the data from SF-BEAMS is solely
+  used for Data Archival purposes, since they are stored in the RTC's lab as
+  part of the collaboration research project that monitors each part of the
+  California coast. The data sets are distributed by a specific date in an
+  annual calendar;
   \item \textbf{The Location of the Sensor Data}: it is clear that the purpose
-  of data is directly related to its location and, as consequence, the strategy
-  of \textbf{External Data Storage} is used to store its collected data at the
-  network sink at the RTC's labs;
+  of data is directly related to its location and, as a consequence, the
+  strategy of \textbf{External Data Storage} is used to store its collected
+  data at the network sink at the RTC's labs;
   \item \textbf{Data Model}: no data model is used by the RTC's staff to model
-  the data. A simple transformation of the data collected from the sensors
+  the data. A straightforward conversion of the data collected from the sensors
   directly to spreadsheets, using comma-delimited values, is used. That is, the
-  data mode used is \textbf{Schema-less};
+  data mode used \textbf{Schema-less};
   \item \textbf{Data Provenance}: data is stored in files with conventions that
-  takes the time when the data was collected. The contents of the files include
-  the \textbf{Data Identity} and \textbf{Time Dimensions}, as shown in
-  the samples collected and shown in Listing \ref{file:rtc-ysi-opendap};
+  take the time when the data was collected. The contents of the files contain
+  the \textbf{Data Identity} and \textbf{Time Dimensions}, as shown in the
+  samples collected and shown in Listing \ref{file:rtc-ysi-opendap};
   \item \textbf{Query Processing Mechanism}: data is queried in a
   \textbf{Centralized} fashion, due to the nature of its purpose and locality;
-  \item \textbf{Database System Organization}: there are no database in place.
-  The SF-BEAMS labs use the OPeNDAP Hyrax server as a middleware used to read
+  \item \textbf{Database System Organization}: there are no databases in use.
+  The SF-BEAMS staff use the OPeNDAP Hyrax server as a middleware used to read
   and interpret the collected data. 
@@
 management and distribution. By the use of COTS\footnote{Common-Off-The-Shelf
-designates a product that is produced and sold in bulk} embedded
-devices and open-source \cite{open-source} software, the research group 
-developed a second version of NetBEAMS, a component-based approach suggested
-to improve the operational activities from SF-BEAMS by using systems
-automation techniques. In brief, one of the open problems of NetBEAMS was
-regarding Data Persistence, the main motivation of this dissertation.
+designates a product that is produced and sold in bulk} embedded devices and
+open-source \cite{open-source} software, the research group developed a second
+version of NetBEAMS, a component-based approach suggested improving
+operational activities from SF-BEAMS using systems automation. In brief, one
+of the open problems of NetBEAMS was regarding Data Persistence, the primary
+motivation of this dissertation.
 
-The Networked Bay Environmental Assessment and Monitoring System, or NetBEAMS, 
+The Networked Bay Environmental Assessment and Monitoring System, or Net-BEAMS,
 is a work-in-progress joint project proposed by the department of Computer
 Science and the RTC at San Francisco State University. The main reason for the
-inception of NetBEAMS is the potential to decrease operational costs and speed
-the data gathering process. In the current state of SF-BEAMS, Biologists need
-to manually collect the data from the sensor devices in order to take them to
-the network sink (lab) for further processing. For this reason, the Data
+inception of NetBEAMS is the potential to reduce operational costs and
+facilitate the data gathering process. In the current state of SF-BEAMS,
+Biologists need to obtain data manually from the sensor devices in order to
+take them to the network sink for further processing. For this reason, the Data
 Sensor Platform (DSP) \cite{netbeams2009} was proposed as the system
-architecture that can address the operation of SF-BEAMS. Therefore, an in-depth
-detail of the the DSP Platform, its architecture and data gathering process
-can are provided in the appendix section \ref{sec:dsp-details}, and may be
-referred at any time.
+architecture that can address the operation of SF-BEAMS. In this way, in-depth
+documentation describing the DSP Platform, including its architecture and data
+gathering process, is provided in the appendix section \ref{sec:dsp-details}.
 
 The scope of this work then is defined as to provide a persistence layer for
-NetBEAMS, making a selection of a persistence layer that is complaint to the
+NetBEAMS, making a selection of a persistence layer that is a exception to the
 taxonomical classification of SF-BEAMS. Furthermore, the technology to be
@@
 
-\subsection{Requirements for NetBEAMS Data Use}
+\section{Requirements for NetBEAMS Data Use}
 
@@
   Database Systems;
-  \item Data Representation must be similar to the ones used by RTC,
-  with a more human-readable format;
-  \item Scalable System with low degree of maintanance, in a way that data
-  gathering service should be able to manage the system with less service
-  interruption;
-  \item Data must be searchable in near-real-time with good performance, as
-  compared to the time when data was collected to the time when they are available;
+  \item Data Representation must be similar to the ones used by RTC, with a
+  more human-readable format;
+  \item Scalable Systema low degree of maintenance, in a way that data
+  gathering service should be able to manage the system with less service interruption;
+  \item Data must be searchable in near-real-time with outstanding performance,
+  as compared to the time when data was collected to the time when they are available;
   \item Easy data integration and export capabilities must be also taken into

### 2009-12-01 docs/marcello-dissertation/thesis/chapter10-appendix.tex
Updates to the images of the experiments and revisions to different chapters from 4 to 10.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@664 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 and the DSP Platform, described by an artifact called MANIFEST.MF as seen in
-listing \cite{file:osgi-intro-manifest}. In general, an osgi-intro bundle must
+Listing \ref{file:osgi-intro-manifest}. In general, an osgi-intro bundle must
 provide specifications that describes the module to be published into the osgi-intro
@@
   They are provided through a ``Export-Package'' section in other osgi-intro
-  bundles and, as it is shown in listing \ref{file:osgi-intro-manifest}, they
+  bundles and, as it is shown in Listing \ref{file:osgi-intro-manifest}, they
   are from different packages.
@@
   message for the DSP Data Persistence Component, as shown in Listing 
-  \ref{file:dsp-data-pers-bootstrap.xml}.
+  \ref{file:dsp-data-persistence-bootstrap.xml}.
 \end{itemize}
@@
 
-Futon4Mongo is an open-source software adapted from CouchDB \ref{couchdb}.
+Futon4Mongo is an open-source software adapted from CouchDB \cite{couchdb}.
 First, the collections list is shown. Figure
@@
 
-\lstinputlisting[label=file:osgi-manifest,caption=OSGi Manifest
+\lstinputlisting[label=file:osgi-intro-manifest,caption=OSGi Manifest
 Descriptor]{../../../../netbeams/versions/v2/apps/osgi-bundles/dsp/DSPDataPersistence/META-INF/MANIFEST.MF}
@@
 \lstinputlisting[language=Bash,label=file:experiment-setup-executor,caption=Experiment
-Implementation]{../../../../netbeams/versions/v2/persistence/run-persistence-experiment}
-
-\lstinputlisting[language=Java,label=file:random-ysi-data-generator,caption=Experiment
-Implementation]{../../../../netbeams/versions/v2/apps/osgi-bundles/dsp/DSPSondeProducer/src/org/netbeams/dsp/ysi/SondeTestData.java}
-
-\lstinputlisting[language=Java,label=file:experiment-dsp-implementation,caption=Experiment
-Implementation]{../../../../netbeams/versions/v2/apps/osgi-bundles/dsp/DSPDataPersistence/src/org/netbeams/dsp/persistence/controller/DSPMessageToMongoDBExperiment.java}
+Main Execution Script]{../../../../netbeams/versions/v2/persistence/run-persistence-experiment}
+
+\lstinputlisting[language=Bash,label=file:experiment-query-scenarios,caption=Experiment
+Scenarios Implementation]{../../../../netbeams/versions/v2/persistence/experiment-client-scenarios.js}
+
+\lstinputlisting[language=Java,label=file:random-ysi-data-generator,caption=Random
+Test Data Generator]{../../../../netbeams/versions/v2/apps/osgi-bundles/dsp/DSPSondeProducer/src/org/netbeams/dsp/ysi/SondeTestData.java}
+
+\lstinputlisting[language=Java,label=file:experiment-dsp-implementation,caption=Inserting
+Random Data to MongoDB]{../../../../netbeams/versions/v2/apps/osgi-bundles/dsp/DSPDataPersistence/src/org/netbeams/dsp/persistence/controller/DSPMessageToMongoDBExperiment.java}
 
@@
 \end{lstlisting}
+
+\lstset{label=file:mongodb-export-command,caption=Export Command to CSV format}
+\begin{lstlisting}
+mongoexport -d netbeams -c SondeDataContainer --dbpath data --csv -f "_id,sensor_ip_address,transaction_time,fact_time,data.temperature,data.sp_condition,data.condition,data.resistence,data.salinitude,data.pressure,data.depth,data.ph,data.pH_mv,data.odo_sat,data.odo_condition,data.turbidity,data.battery" -o logs/experiment-483840-data-exported-20091129-190207.csv
+
+Sun Nov 29 19:07:04 query netbeams.SondeDataContainer ntoreturn:0 reslen:2258 nscanned:101 { query: {}, $snapshot: 1 }  nreturned:101 202ms
+Sun Nov 29 19:07:04 getmore netbeams.SondeDataContainer cid:7749363894863150628 ntoreturn:0 query: { query: {}, $snapshot: 1 }  bytes:1048562 nreturned:47661 1767ms
+...
+...
+...
+Sun Nov 29 19:10:18 getmore netbeams.SondeDataContainer cid:7749363894863150628 ntoreturn:0 query: { query: {}, $snapshot: 1 }  bytes:793098 nreturned:36049 1975ms
+exported 2419200 records
+\end{lstlisting}

### 2009-12-01 docs/marcello-dissertation/thesis/chapter4-netbeams-in-depth-requirements-analysis-and-specification.tex
Updates to the images of the experiments and revisions to different chapters from 4 to 10.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@664 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 manufacturer. SF-BEAMS can be seen a single-hop star sensor network containing
-different instances of sensors, including the ``YSI 6600 ESD V2", as seen in
-figure \ref{fig:ysi-device}, which is a powerful water quality monitoring
-device that produces around 52 bytes (13x4 Bytes) on a single real-time data
-stream as shown on table \ref{tab:ysi-data-stream}.
+different instances of sensors. One example of such sensors is the ``YSI 6600
+ESD V2", as seen in Figure \ref{fig:ysi-device}. Therefore, it can be seen as
+a powerful water quality monitoring device that produces around 52 bytes (13x4
+Bytes) on a single real-time data stream as shown on table
+\ref{tab:ysi-data-stream}.
 

### 2009-12-01 docs/marcello-dissertation/thesis/chapter5-taxonomic-persistence-storage-selection-for-netbeams.tex
Updates to the images of the experiments and revisions to different chapters from 4 to 10.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@664 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 requirements defined by NetBEAMS, as our case study in chapter
-\ref{sec:problem-requirements}, is be compared. Then, the scope of this work is
-compared to the technologies selected based on the readings of the
-literature review in persistence in sensor networks and other new trends in
-performance. Finally, a table summarizes the contenders strenths and discusses
-the selected technology to be integrated with the design of a Data
-Persistence component.
+\ref{sec:problem-requirements}, is to be compared. Then, the scope of this
+work is compared to the technologies selected based on the readings of the
+literature reviewing data persistence in sensor networks and other recent
+trends in performance. Finally, a table summarizes the contenders strengths
+and discusses the selected technology to be integrated with the design of a
+Data Persistence component.
 
@@
 
-The scope of the use of the database can cover different functional and
-non-functional requirements defined in the previous chapter. This work does
-not intend to implement an application based on the database system selected,
-but provide the foundation of a database management system that can be used to
+The scope of the use of the database can involve different functional and
+non-functional requirements defined in the last chapter. This work does not
+intend to implement an application based on the database system selected, but
+to provide the foundation of a database management system that can be used to
 develop one focused in less refactoring and high scalability and performance.
@@
 programming languages or API. In this way, the scope of this technology
-selection is can be summarized as follows:
+selection can be summarized as follows:
 
 \begin{itemize}
-  \item Provide data management over the collected sensor data, as a regular
-  database system;
+  \item Provide data management over the collected sensor data, as a
+  conventional database system;
   \item Must be able to scale in the face of an increase to the data load of
@@
 
-This section describes the database systems contenders considered to be used
+This section describes the database systems contenders considered being used
 together with for the implementation of the Data Persistence component for 
@@
 
-One of the primary questions related to data persistence is regarding the data
-model and type of the database, that is, the use of relational model or
-structured models. In addition to that, and given the current time of academic
-research and new trends distributed computing such as Cloud Computing and the
-Internet \cite{cloud-comp-survey}, the selection of a data persistence for
+One of the essential questions related to data persistence is regarding the
+data model and nature of the database, that is, the use of the relational or
+structured models. In addition to that, and given the present time of academic
+research and recent trends distributed computing such as Cloud Computing and
+the Internet \cite{cloud-comp-survey}, the selection of a data persistence for
 sensor networks should maintain a wider spectrum of the types of database 
@@
 also included the overview of alternative data models for the development of
-an experiment that brings data persistence to NetBEAMS. In such a fashion, the
-list of the systems considered for evaluation is the following:
+an experiment that brings data persistence to NetBEAMS. In such a manner, the
+list of the systems considered for assessment is the following:
 
@@
   \item \textbf{MySQL}\cite{mysql}: an open-source relational database system
-  that is the primary choice of the development of projects of different sizes. It was selected
-  because of its popularity in the current community;
+  that is the primary choice of the development of projects of different sizes.
+  It was selected because of its popularity in the contemporary community;
   \item \textbf{Oracle}\cite{oracle}: is the most popular commercial database
-  system, having a campus wide academic licenses availabe;
+  system, having a campus wide academic licenses available;
   \item \textbf{TinyDB}\cite{tinydb}: as referred in the literature review,
-  TinyDB is a database system used in many different sensor networks implementation
-  \cite{sn-db-tinydb};
-  \item \textbf{MongoDB}\cite{mongodb}: an open-source key-value database system
-  described in the article that questions the use of the relational model
-  \cite{db-is-rdbs-dommed}, as well as the Cloud Computing trends survey
-  \cite{cloud-comp-architectures}.
-  \item \textbf{DB2}\cite{db2}: in the search of an XML database system, DB2 is
-  listed as a hybrid database system that offers both XML and relational model
-  \cite{db-xml-enabled}. It was selected because NetBEAMS can represent the
-  collected data in the XML format \ref{sec:dsp-message}.
+  TinyDB is a database system used in many different sensor networks
+  implementation \cite{sn-db-tinydb};
+  \item \textbf{MongoDB}\cite{mongodb}: an open-source key-value database
+  system described in the article that questions the use of the relational
+  model \cite{db-is-rdbs-dommed}, as well as the Cloud Computing trends survey
+  \cite{cloud-comp-architectures};
+  \item \textbf{DB2}\cite{db2}: in the pursuit of an XML database system, DB2
+  is listed as a hybrid database system that offers both the relational and the
+  XML models \cite{db-xml-enabled}. It was selected because NetBEAMS can
+  represent the collected data in the XML format as shown in setion 
+  \ref{sec:dsp-message}.
 \end{itemize}
 
-These technologies will be reviewed in contrast to their capabilities to adhire
+These technologies will be reviewed in contrast to their capabilities to adhere
 to the specifications of the taxonomies defined in chapter 3 and the
@@
   sink at the RTC laboratories;
-  \item Upon data reception, the RTC staff index and archive the data for
-  for distribution using the OPEnDAP format;
-  \item An automated approach to the data collection procedures is done
-  using NetBEAMS \cite{netbeams2009}, using software components to extract and
-  send the data to the network sink. A complete description of this approach
-  is described in section \ref{netbeams-architecture}.
+  \item Upon data reception, the RTC staff index and archive the data for for
+  distribution using the OPEnDAP format;
+  \item An automated approach to the data collection procedures is done using 
+  NetBEAMS \cite{netbeams2009}, using software components to obtain and send
+  the data to the network sink. A full description of this approach is
+  described in section \ref{netbeams-architecture}.
 \end{itemize}
@@
 
-As seen in the previous section, NetBEAMS's architecture is based on a
-single-hop star infrastructure from the SF-BEAMS sensor network, having all
-the data collected from the sensors being directly sent to the network sink 
-\ref{sec:sn-infrastructure}.
+As shown in the previous section, NetBEAMS's architecture is based on a
+single-hop star infrastructure from the SF-BEAMS sensor network, having all the
+data collected from the sensors being immediately sent to the network sink
+(Section \ref{sec:sn-infrastructure}).
 
-Based in the observations also described in the previous chapter, the simplest
-location to store the collected data is through the use of an \textbf{External
-Storage}. In addition to that, the use of a \textbf{Data-Centric Storage}
-approach is also considered, in situations where the data load in the
-network sink reaches its storage capacity.
+Based on the observations also described in the last chapter, the simplest
+position to save the collected data is through the use of an \textbf{External
+Storage}. In addition to that, the use of a Data-Centric Storage strategy is
+also considered in situations where the data load in the network sink reaches
+its storage capacity. The \textbf{Data-Centric Storage} storage separates the
+location of storage of the data based in any property of the data such as the
+value of a given key. Such approach of data partitioning is provided by
+database system in two different flavors.
 
@@
 of data partitioning is provided by database system in two different flavors
-\ref{db-partitioning-relational}:
+\cite{db-partitioning-relational}:
 
@@
   on the values of the table rows in different groups allocated to different
-  datasets physically located in different servers. In this way, all the columns
-  defined for a given entity are found on each partition. For instance,
+  datasets physically located in different servers. In this way, all the
+  columns defined for a given entity are found on each partition. For instance,
   horizontal partitioning can be defined as a collection five years worth of
-  historical sensor data being partitioned into five distinct servers, where
+  historical sensor data being partitioned into five separate servers, where
   each server has a single year's worth of sensor data;
   \item \textbf{Vertical Partitioning}: data is partitioned vertically in a way
-  that the columns of an entity are divied to different database tables. In
+  that the  columns of an entity are divided into different database tables. In
   this way,  the particular dataset is broken down to different tables. For
   example, if the  database is used to store pictures of a given sensor camera
-  in a BLOB column, it would be the most indicated to the scenario where 
-  the images are not constantly referenced when compared to the the textual
-  data.
+  in a BLOB column, it would be the most preferable to the scenario where the
+  images are not  constantly referenced when compared to the textual data.
 \end{itemize}
@@
 The External Storage approach is the most common way to implement persistence
-for any type of systems since data management is usually offered by most
-database systems out-of-the-box. Furthermore, this approach can be accompained
+for any breed of systems since data management is usually offered by most
+database systems out-of-the-box. Furthermore, this approach can be accompanied
 with the use of Distributed Systems approaches such as Replication to help a
@@
 The most common problem related to a single External Data Storage is that 
-resources may run into data management problems such as full disk space or disk
-failures.
+resources may run into data management problems such as full-disk space or
+disk failures.
 
@@
 
-The advantage of the Data-Centric approach can be directly associated with
+The benefit of the Data-Centric approach can be directly associated with
 database partitioning mechanisms described earlier. Since data is distributed
-into different locations, this approach the decrease the size of datasets when
-the database server performs a query processing in a database table. For this
-reason, this approach helps the database administrators to scale their
-database infrastructure and improve the database performance.
+into different locations, this approach decreases the size of datasets when the
+database server performs a query processing in a database table. For this
+reason, this approach helps the database administrators to scale their database
+infrastructure and improve the database performance.
 
@@
 Database Sharding \cite{db-shard-discussion}, or simply Database partitioning
-\cite{db-table-partition} \cite{db-partitioning-relational}. Both former and
-the latter are used in regular schema-dependent models or schema-less
-models, which uses denormalized data models. There are different approaches to
-partition data \cite{db-shard-schemas} \cite{db-partitioning-relational}
-\cite{db-partitioning-relational-oracle}.
+\cite{db-table-partition, db-partitioning-relational}. Both former and the
+latter are used in regular schema-dependent models or schema-less models,
+which uses denormalized data models. There are different approaches to
+partition data \cite{db-shard-schemas, db-partitioning-relational,
+db-partitioning-relational-oracle}.
 
 \begin{itemize}
-  \item \textbf{Directory Based Partitioning}: a simple lookup service is used
-  to identify which database shard to use, where the lookup table can be stored
-  in a secondary area;
-  \item \textbf{Vertical Partitioning}: each feature of the system, usually
+  \item \textbf{Directory Based Partitioning}: an ordinary lookup service is
+  used to identify which database shard to use, where the lookup table can be
+  stored in a secondary area;
+  \item \textbf{Vertical Partitioning}: each feature of the system is usually
   represented by a table in the relational model and placed in a different
@@
   going to  be stored;
-  \item \textbf{Range-Based Partitioning}: after partitioning each feature on a
-  different server host, different multiple shards, or similar hosts, can be
+  \item \textbf{Range-Based Partitioning}: after partitioning each feature on
+  a different  server host, different multiple shards, or similar hosts, can be
   added by using  the range of a given key of the entity of a function. For
@@
 An example of a Data-Centric storage with different shards
-\cite{db-shard-intro} can be see in Figure
+\cite{db-shard-intro} ca n be see in Figure
 \ref{fig:database-sharding-by-region}, where data is segregated by the
-location of data. Each shard is assumed to be placed in different server hosts.
-Since the data model is denormalized, the organization of data is entirely
-placed in each of these shards in a Horizontal partition of the entity. As it is
-shown, the shards by location can be defined by a shard key.
+location of data. Each shard is assumed to be placed in different server
+hosts. Since the data model is denormalized, the organization of data is
+entirely placed in each of these shards in a Horizontal partition of the
+entity. As it is shown, the shards by location can be defined by a shard key.
 
@@
 
-The data partitioning is a very restricted technology, mostly available to
+The data partitioning is a highly restricted technology, normally available to
 advanced users of distributed database systems. Despite being a capability
@@
   shards, the strategy of rebalancing must be done. For instance, when one of
-  the shards outgrows more than the others. As a consequence, the database
+  the shards out-grows more than the others. As a consequence, the database
   shards must be rebalanced by transferring data to new locations whenever
-  necessary. This technique is starting to appear as COTS implemetations such
+  necessary. This technique is starting to appear as COTS implementations such
   as mongoDB \cite{mongodb} or MySQL \cite{mysql};
@@
   When Data-Centric Storage is used, both vertical and horizontal partitions
-  \cite{db-partitioning-relational} can be chosen, using different strategies to
-  partition the data;
-  \item \textbf{Oracle}: Supports External or Data-Centric Storage
-  approaches. Only supports vertical partitioning with the definition
-  of different types \cite{db-table-partition};
-  \cite{db-partitioning-relational-oracle} when using Data-Centric storage;
+  \cite{db-partitioning-relational} can be  chosen, using different strategies
+  to partition the data;
+  \item \textbf{Oracle}: Supports External or Data-Centric Storage approaches.
+  Only sup-  ports vertical partitioning with the definition of different types
+  \cite{db-table-partition, db-partitioning-relational-oracle} when using
+  Data-Centric storage;
   \item \textbf{Berkeley TinyDB}: Only supports the External Storage;
-  \item \textbf{MongoDB}: Supports External or Data-Centric Storage.
-  Using the latter one, only the horizontal partition approach is available
-  using with automatic shards \cite{db-mongo-partition};
+  \item \textbf{MongoDB}: Supports External or Data-Centric Storage. When the
+  latter approach is used, the horizontal partition strategy is available with
+  automatic shards \cite{db-mongo-partition};
   \item \textbf{DB2}: Supports External and Data-Centric storage. Only supports
-  vertical table partition \cite{db-db2-partition} for the data-centric storage;
+  vertical table partition \cite{db-db2-partition} for the data-centric storage.
 \end{itemize}
 
-All the selected database systems supports the External Storage approach.
-However, they all differ with the support to Data-Centric Storage support.
+All the selected database systems support the External Storage method. However,
+they all differ with the support to Data-Centric Storage support.
 
@@
 Processing} is indicated for NetBEAMS, also given the fact that SF-BEAMS data
-goes directly to the single network sink, and therefore, a centralized location
-of the data.
+goes directly to the individual network sink and, therefore, a centralized
+point of the data.
 
@@
 Centralized data management and query processing is simpler than the in-network
-query method. Data is verified by a simple database management system without
-the risk of data being unavailable, when the data is spread among the network
-nodes.
+query method. Data is verified by a straightforward database management system
+without the risk of data being unavailable, when the data is spread among the
+network nodes.
 
@@
 
-When the Centralized Query processing is used, problems may occur. Depending
-on the size of the datasets and the database technology, performance may be a
-concern. For example, this approach creates of the so-called Funneling
-Affect \cite{sn-storage04}, since the point-of-traffic is concentrated in the
+When the Centralized Query processing is used, problems may occur. Depending on
+the size of the datasets and the database technology, performance may be a
+concern. For example, this procedure creates of the alleged Funneling Affect
+\cite{sn-storage04}, since the point-of-traffic is concentrated in the
 centralized system.
@@
 Modeling systems, and for this reason, one of the requirements of the system is
-try the use of schema-less approaches. The \textbf{Key-Value-Pair Data Model}
-or the \textbf{Document-Oriented Data Model} seems to be the simplest choices of the
-models.
+trying the use of schema-less approaches. The \textbf{Key-Value-Pair Data
+Model} or the \textbf{Document-Oriented Data Model} seems to be the simplest
+choices of the models.
 
@@
 
-Considering the inception of a relational data model \cite{relational-model} and
-the use of the YSI Sonde data as the main entity in the system, let Figure 
-\ref{fig:Relational-Model-Original} represent a prototype of the relational
-model, after passing through the process of normalization
-\cite{db-normalization}.
+Considering the inception of a relational data model \cite{relational-model}
+and the use of the YSI Sonde data as the main entity in the system, let Figure
+\ref{fig:Relational-Model-Original} represent a prototype of the relational model, after passing through the process of
+normalization \cite{db-normalization}.
 
@@
 
-Supposing a new type is added into the system, let the refactored
-version of the relational model be depicted in Figure
-\ref{fig:Relational-Model-Addition-Modified}. Some observations are considered
-in this scenario:
+Supposing a new type is added into the system, let the refactored version of
+the relational model be depicted in Figure
+\ref{fig:Relational-Model-Addition-Modified}. Some observations are
+considered in this scenario:
 
@@
   \item The approach of Relational Model might does not fit the nature of
-  collected data from wired or wireless sensor networks, since they tend to
-  change over time \cite{db-is-rdbs-dommed}. They usually does not give support
-  time-series data nor provenance well \cite{sn-provenance};
+  collected data from wired or wireless sensor devices, since they tend to
+  change over time \cite{db-is-rdbs-dommed}. They usually does not give
+  support time-series data nor provenance well \cite{sn-provenance};
   \item Some research have suggested changes to the SQL clauses to better
-  support time-series \cite{sn-db-newop}.
+  support time-series when collecting data using the relational model 
+  \cite{sn-db-newop}.
 \end{itemize}
@@
 being complaint to an XML Schema \cite{xml-schema}. This model uses XPath
-\cite{xml-xpath} technology for querying documents, although some hibrid
+\cite{xml-xpath} technology for querying documents, although some hybrid
 technologies may still use SQL \cite{sql} for that matter \cite{db2}. In this
@@
 led developers to propose such approach using the relational model as
-documented in technical articles such as \cite{db-kvp-in-relational01} and
-\cite{db-kvp-in-relational02}. Based in these articles, consider the relational
-model depicted in Figure \ref{fig:KVP-on-Relational-Model} as a data model for
+documented in technical articles such as  and
+. Based in these articles, consider the relational
+model depicted in Figure  as a data model for
 persisting data to our case study using the key-value strategy. The problem
@@
 sensor networks.
+
+Although this data model has been used in different types of applications,
+developers have tried to use the Relational Model to model an application that
+could prevent any database change through the use of a technique that relates a
+key to a value, called Key-Value data model. With the emergence of Internet
+applications, the necessity of such a model that does not require constant
+refactoring led developers to propose such manner using the relational model as
+documented in technical articles such as \cite{db-kvp-in-relational01} and
+\cite{db-kvp-in-relational02}. Based on these articles, consider the
+relational model depicted in Figure \ref{fig:KVP-on-Relational-Model} as a
+data model for persisting data to our case study using the key-value strategy.
+The problem with it is that key repetition occurs, since the nature of
+time-series data requires the use of a timestamp key for each property of the
+sensor. Therefore, this strategy is not well-suited to provide persistence
+data from collected sensor networks.
 
@@
 infrastructure and how to represent models. An powerful up-coming approach is
-the use of database systems that does not use structured language for querying
+the use of database systems that do not use structured language for querying
 its data. One such example is called Key-Value Pair (KVP) data model
@@
 Computing \cite{cloud-comp-architectures}. Databases implementing this
-strategy is the mongoDB \cite{mongodb}.
+strategy is the mongoDB .
+
+There are no records of the use of this data model with Sensor Networks.
+Different variants of such data model is the document-oriented data model,
+where data is modeled using structured documents. However, these documents have
+a dynamic structure that can be freely described without the use of a schema
+validation mechanism as used in XML Schema. One example of such approach is the
+use of the JSON data format [Ste08] in the implementation of database systems
+which are becoming standard with the latest trend called Cloud Computing [LKN+
+09]. Databases implementing this strategy is the mongoDB \cite{mongodb},
+CouchDB \cite{couchdb}, Amazon Dynamo \cite{db-kvp}, among others.
 
@@
 For instance, all data needed for the YSI sonde data, as well as all necessary
-provenance data that describes the data, depicted in section
+provenance data that describe the data, depicted in section
 \ref{sec:sn-provenance}, are represented by means of key-value pairs
@@
     \begin{enumerate}
-      \item Real-world model by entities, classified in tables;
+      \item Real-world model represented by entities, classified in tables;
       \item Tables composed by columns and rows. Rows are comprised by column
@@
 
-Since the selected data model is the schema-less one, the following can be see
+Since the selected data model is the schema-less one, the following can be seen
 from the technologies:
@@
 \begin{itemize}
-  \item MySQL: Only supports the Relational Model;
-  \item Oracle: Only supports Relational Model;
-  \item Berkeley TinyDB: Only supports the Relational Model; 
-  \item \textbf{MongoDB}: Supports Document-Oriented Model;
-  \item IBM DB2: Only Supports the Relational Model.
+  \item MySQL: supports the Relational Model;
+  \item Oracle: supports Relational Model;
+  \item Berkeley TinyDB: supports the Relational Model; 
+  \item \textbf{MongoDB}: supports Document-Oriented Model;
+  \item IBM DB2: supports the Relational Model and the XML model.
 \end{itemize}
@@
   \item Very advanced topics in Database System;
-  \item May be easy of use, depending on the application.
+  \item Partitions need to be rebalanced in case the key of the selected 
+  partitioning strategy is changed.
 \end{itemize}
@@
   \item \textbf{Open-source}: Given the fact of the nature of the project, the
-  technology to be used must be free of cost \cite{open-source}, that is, no
-  costs involved in the adoption of such technology; Together with it is the
+  technology to be used must be free of charge \cite{open-source}, that is, no
+  costs involved in the adoption a such technology; Together with it is the
   support from the community;
-  \item \textbf{Native APIs}: In order to cover the scope of this work defined
-  in the previous chapter, the solution for persisting collected sensor data
-  must be not only limited to a technology that provides data access, such as
-  SQL, but also by other \textbf{data access mechanisms};
+  \item \textbf{Native APIs}: In order to consider the scope of this work
+  defined in the last chapter, the solution for persisting collected sensor
+  data must be not only limited to a technology that provides data access, such
+  as SQL, but also by other data access mechanisms \textbf{data access
+  mechanisms};
   \item \textbf{Hot Backup}: Supports hot backup with less service interruption. 
@@
 use of drivers in different programming and scripting languages such as Java,
-Python, Perl, etc. Moreover, only TinyDB does not support hot backup.
+Python, Perl, among others. Moreover, only TinyDB does not support hot backup.
 
 \begin{itemize}
-  \item \textbf{MySQL}: Is open-source, supports hot back and contains
+  \item \textbf{MySQL}: Is  open-source, supports hot backup and contains
   scripting APIs from  the community;
@@
 list, as well as the Internet. Since mongoDB is based on a schema-less database
-system, and supports most of the taxonomies defined in the previous chapter.
+system, has not been reported in use by the sensor network community, and
+supports most of the taxonomies defined in the previous chapter, it was
+selected for the implementation of the DSP Data Persistence component for
+NetBEAMS.
 

### 2009-12-01 docs/marcello-dissertation/thesis/chapter6-data-persistence-component-design-and-architecture.tex
Updates to the images of the experiments and revisions to different chapters from 4 to 10.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@664 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 Bay, composed by different types of sensors. The process of extracting the
-measurements from the sensors to be reused using the DSP Platform can be
+measurements from the sensors to be reused through the DSP Platform can be
 summarized following the UML Business Process Diagram \cite{uml} depicted in
@@
 \subsection{Functional Requirements}
+\label{sec:use-cases}
+
 As part of the user stories related to data access, figure
@@
 
-\begin{figure}[!b]
-  \centering
-  \includegraphics[scale=0.5]{../diagrams/DSP-Data-Persistence-UseCases-Diagram-System}
-  \caption{UML Use Case diagram for System Persistence Functions}
-  \label{fig:DSP-Data-Persistence-UseCases-Diagram-System}
-\end{figure}
+\begin{itemize}
+  \item \textbf{Search Measurements by Sensor Properties}: users of the
+  collected data must be able to search data through the properties of the
+  sensors. This is related to identity principle of what was collected
+  as provenance \cite{db-provenance};
+  \item \textbf{Search Measurements by Sensor Location}: users of the collected
+  data must be able to search data by a given sensor location, as referred to
+  the provenance from where the data was collected;
+  \item \textbf{Search Measurements by Data collection date}: users of the
+  collected data must be able to search data by a given date, or ranges of
+  start and end date;
+  \item \textbf{Access the data through API, Programming Languages}: users of the
+  collected data must be able to search data by the use of different
+  programming and scripting languages of their preference;
+  \item \textbf{Export Data to OPEnDAP format}: users of the collected data
+  must be able to export the collected data to the OPEnDAP format;
+  \item \textbf{Export Data to Spreadsheets}: users of the collected data must
+  be able to export the collected data to spreadsheets.
+\end{itemize}
 
@@
   \item The Persistence storage system must be able to scale without too much
-  architectural changes;
+  architectural changes, or schema changes;
   \item The Persistence storage system must not impose the knowledge of
   complicated database system languages such as SQL \cite{sql} or XML XPath
-  \cite{xml-xpath}, but use access through API calls, since it is closer to
-  researchers.
+  \cite{xml-xpath}, but use access through API calls, as it offers better
+  abstractions for non-computer science researchers
+  \cite{sn-programming-language}.
 \end{itemize}
@@
 \label{sec:dsp-persistence-data-model}
+
 As described in the previous chapter, the data model that might give better
@@
 Model. Beyond the data collected from the sensor devices, Data Provenance 
-techniques must also be considered to better describe the collected data. In
-general, the data model must include the following metadata besed on Data
-Provenance:
+techniques must also be considered to better describe the collected data
+\cite{sn-provenance}. In general, the data model must include the following
+metadata besed on Data Provenance:
 
@@
   UUID\footnote{Universally Unique Identifier};
-  \item \textbf{Time-Space}: when the data was ``sensed'' \cite{db-provenance}
-  and when the collection transaction happened \cite{sn-time-series}.
+  \item \textbf{Time-Dimensions}: when the data was ``sensed''
+  \cite{db-provenance} and when the collection transaction happened
+  \cite{sn-time-series}.
 \end{itemize}
@@
 \begin{itemize}
-  \item \textbf{sensor\underline{ }ip\underline{ }address}: it's extracted from
-  the DSP Message Produce and identifies which sensor generated the sampling;
+  \item \textbf{sensor.ip\underline{ }address}: it is extracted
+  from the DSP Message Produce and identifies which sensor generated the sampling;
+  \item \textbf{sensor.location.latitude, longitude}: the latitude and
+  longitude GPS coordinates of the position of the sensor device, captured from
+  references of the device. This is currently provided as a parameter in the
+  bootstrap message for the component;
   \item \textbf{message\underline{ }id}: it's extracted from each of the messages contained
   in the DSP Message Container;
-  \item \textbf{transaction\underline{ }time}: it's extracted from the DSP message container
-  creation time and it is used to identify when the transaction occurred;
-  \item \textbf{fact\underline{ }time}: it's extracted from the SondeDataContainer's date
-  and time, and identifies the time in which the collected data occurred;
-  \item \textbf{latitude}: the latitude coordinate of the position of
-  the sensor device, captured from references of the device (not supported by
-  NetBEAMS);
-  \item \textbf{longitude}: the longitude coordinate of the position of the
-  sensor device, captured from references of the device (not supported by
-  NetBEAMS).
+  \item \textbf{time.transaction}: it is extracted from the DSP
+  message container creation time and it is used to identify when the transaction occurred;
+  \item \textbf{time.fact}: it is extracted from the
+  SondeDataContainer's date and time, and identifies the time in which the
+  collected data occurred.
 \end{itemize}
@@
 \begin{itemize}
-  \item data: this key defines the values of the document, and will have every
-  different property of the sensor. For instance, the data key for the
+  \item \textbf{observation}: this key defines the values of the document, and
+  will have every different property of the sensor. For instance, the data key for the
   properties of the YSI device will contain the set of properties defined for
@@
 The new proposed Data Persistence component must follow the DSP requirements
-for a new DSP Component following the the design-pattern of Data
-Producer/Consumer. In a nutshell, the design of a DSP Component resulting
-maintains the dependency among 3 different major Java packages, as the addition
-of such a DSP Data Persistence component depends on 2 existing packages, namely
-the OSGi Framework and the DSP Platform, as well as and an additional package
-for the database access driver API, as shown in the UML Package Dependency
-Diagram \cite{uml} of figure
+for a new DSP Component, which specifies the design-pattern of Data
+Producer/Consumer. In a nutshell, the design of a DSP Component maintains the
+dependency among 3 different major Java packages: it depends on 2 existing
+packages, namely the OSGi Framework and the DSP Platform, as well as and an
+additional package for the database access driver API, as shown in the UML
+Package Dependency Diagram \cite{uml} of Figure
 \ref{fig:DSP-Data-Persistence-Packages-Dependency}.
@@
 
-The DSP Data Persistence component starts with the design of an OSGi Activator
-class, and the implementation of the DSP Component, resposible to provide
-services to the platform. Based on these specifications, the following set of
-classes in figure \ref{fig:DSP-DataPersistence-Activator-Class-Diagram} shows
-the UML Class Diagram \cite{uml} implemented for the component.
+The DSP Data Persistence component starts with an OSGi Activator class, and the
+implementation of the DSP Component, resposible to provide services to the
+platform. Based on these specifications, the following set of classes in
+figure \ref{fig:DSP-DataPersistence-Activator-Class-Diagram} shows the UML
+Class Diagram \cite{uml} implemented for the component.
 
@@
 The behavior of the activator class follows the specifications of the
-OSGi Framework, and for this reason, instances of the DSP Data Persistence
+OSGi Framework and, for this reason, instances of the DSP Data Persistence
 Activator can be in the states described on the UML State Diagram depicted in
-figure \ref{fig:DSPPlatform-Install-Usage-State-Diagram}, as it is an extension
-of the OSGI class BundleActivator, as well as being responsible for stopping
-or starting the the DSP Data Persistence component instance. Following the
-specifications of section \ref{netbeams-architecture}, the DSP Data Persistence
-Platform can be initialized as described by the UML Sequence Diagram of figure
+Figure \ref{fig:DSPPlatform-Install-Usage-State-Diagram}. It is important to
+note that the activator is an extension of the OSGI class BundleActivator, as
+well as being responsible for stopping or starting the the DSP Data Persistence
+component instance. In this way, following the specifications of section
+\ref{netbeams-architecture}, the DSP Data Persistence Platform can be
+initialized as described as it is described by the UML Sequence Diagram of
+Figure
 \ref{fig:From-OSGi-Framework-to-DSP-Data-PersistenceActivator-Sequence-Diagram}:
@@
 Any configuration parameter required by a DSP Component is sent during its
-bootstrap process through the use of an optional DSP Update Message, as
-described in section \ref{sec:dsp-message}. The initialization of the DSP
-Component Activator and the DSP Component registration are also shown in
-figure \ref{fig:From-OSGi-Framework-to-DSP-Data-PersistenceActivator-Sequence-Diagram}.
+bootstrap process through the use of an optional DSP Update Message, as shown
+in Listing \ref{file:dsp-data-persistence-bootstrap.xml}. The initialization of
+the DSP Component Activator and the DSP Component registration are also shown
+in figure \ref{fig:From-OSGi-Framework-to-DSP-Data-PersistenceActivator-Sequence-Diagram}.
 
@@
 
-The the gate "Start DSP Activator Gate" reaches the DSP Data Persistence, it
-first creates the unique instance of the class DSP Data Persisence and uses
-that instance to register it as the OSGi service through the call to the DSP
-Platform helper class ActivatorHelper. At this point, the DSP Data Persistence
-has been installed and is in the state "Active".
+As soon as the gate "Start DSP Activator Gate" reaches the DSP Data
+Persistence instance, it first creates the unique instance of the class DSP
+Data Persisence and uses that instance to register it as the OSGi service
+through the call to the DSP Platform helper class ActivatorHelper. At this
+point, the DSP Data Persistence has been installed and is in the state
+"Active".
 
@@
 \ref{fig:DSP-DataPersistence-Flusher-State-Diagram}: RUNNABLE and
-TIMED\underline{ }WAITING. In the former, the DSP Data Flusher will be
-contacting the Transient Persistence Layer to verify if there are any message
-waiting to be sent to the Database service, while in the latter the DSP Data
-flusher will be waiting for the next cycle defined by the bootstrap parameter
-TRANSIENT\underline{ }DATA\underline{ }FLUSHER\underline{ }DELAY.
+TIMED\underline{ }WAITING. In the former state, the DSP Data Flusher will be
+contacting the Transient Persistence Layer in order to verify if there are any
+message in the outbound queue ready to be transmitted sent to the Database
+service, while the DSP Data flusher will be waiting for the next cycle defined
+by the bootstrap parameter TRANSIENT\underline{ }DATA\underline{
+}FLUSHER\underline{ }DELAY in the latter state.
 
@@
 by the non-functional requirement in the beginning of this chapter. In this
-way, the DSP Data Persistence component maintains a dependency on the
-Singleton class TransientPersistenceLayer, which is responsible for keeping
-track of messages that are received by working component. However, in order to
-add the DSP message into its aggregated map, it first transforms the message
-into an instance of the class PersistentMessageUnit, which has references to
-the class regarding the location of the message, SensorLocation.
+way, the DSP Data Persistence component maintains a dependency on the Singleton
+class TransientPersistenceLayer, which is responsible for keeping track of
+messages that are received by working component. However, in order to add the
+DSP message into its aggregated map, it first transforms the message into an
+instance of the class PersistentMessageUnit, which has references to the class
+regarding the location of the message, SensorLocation.
 
@@
   flush temporary messages into the database at the rate of
-  TRANSIENT\underline{ }DATA\underline{ }FLUSHER\underline{ }DELAY in seconds;
+  TRANSIENT\underline{ }DATA\underline{ }FLUSHER\underline{ }DELAY in seconds,
+  among others. Listing \ref{file:dsp-data-persistence-bootstrap.xml} shows an
+  implementation of a bootstrap message for the DSP Data Persistence;
   \item Any DSP Measure Message can be persisted by the DSP Data Persistence
@@
   I/O on the database, as depicted by the gate "create PersistentMessageUnit
-  Gate". The simple procedure to add the message into the transient persistent
-  layer.
+  Gate". This last gate describes the simple procedure to add the message
+  into the transient persistent layer. Listing
+  \ref{file:dsp-message-serialized-ysi}.
 \end{itemize}
@@
 been saved into the Database or not as depicted by the UML State Diagram
-\cite{uml} in figure \ref{fig:PersistentMessageState-Diagram}.
+\cite{uml} in Figure \ref{fig:PersistentMessageState-Diagram}.
 
@@
 this reason, an instance of the class SensorLocation carries the information
-about the IP address from the component that originally produced the DSP
-Message. Then, an instance of the class PersistenceMessageUnit is created,
-having its state on TRANSIENT, as detailed on UML State Diagram \cite{uml} in
-figure \ref{fig:PersistentMessageState-Diagram}.
+about the IP address and and optional GPS location from the component that
+originally produced the DSP Message. Then, an instance of the class
+PersistenceMessageUnit is created at the state ``TRANSIENT'', as detailed
+on UML State Diagram \cite{uml} in Figure \ref{fig:PersistentMessageState-Diagram}.
 

### 2009-12-01 docs/marcello-dissertation/thesis/chapter8-experimental-results-correct-behavior-and-performance.tex
Updates to the images of the experiments and revisions to different chapters from 4 to 10.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@664 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 \subsection{Key-Value definition}
+\label{sec:keys-definition}
 
@@
 the entities are persisted using any data arrangment, in any order, attributes
-and the associated values. The definition of the document for mongoDB takes
-into account the properties defined by the data provenance study defined by
-\cite{sn-provenance}. The following list are the definitions of the values of
-the defined keys and the provenance-aware document, as an example described by
-Listing \ref{file:mongodb-ysi-data-format}:
+and the associated values, as shown in Listing
+\ref{file:mongodb-ysi-data-format}. The definition of the document for mongoDB
+takes into account the pairs of key and values related for the collected data
+from the sensors, as well as the properties defined by the data provenance
+study \cite{sn-provenance}. The collected data is stored in a key called
+``observation''. On the other hand, the provenance-aware keys are defined as
+follows:
 
@@
 \subsection{Workload}
+\label{sec:workload}
 
@@
 
-In order to verify the system capacity to scale, the data will be devided into
-different shards as part of the support to distribute system, that is, in a
-Data-Centric way.
-
 \begin{itemize}
-  \item First Round: 483,840 items worth one year of produced YSI data at the
-  rate of 1 observation per minute;
-  \item 483,840 * 5, or 2,419,200, representing 5 different YSI components
-  producing data every minute;
+  \item \textbf{First Round}: 483,840 items worth one year of produced YSI data
+  at the rate of 1 observation per minute;
+  \item \textbf{Next Rounds}: 483,840 * 5, or 2,419,200, representing 5
+  different YSI components producing data every minute.
 \end{itemize}
@@
 mongoDB database system. All the collected data from sensors are stored in a
-given Single Server, which characterizes an External Storage device.
+given single server, which characterizes an External Storage device.
 
@@
 \subsection{Scenarios}
+\label{sec:exp-scenarios}
 
 In order to verify the feasibility of the system, the use of different search
-scenarios are used:
+scenarios are used defined based on the use cases of section
+\ref{sec:use-cases}:
 
@@
 
+Moreover, the experiments must also export data in different formats such as 
+spreadsheets and use different programming languages APIs to access the data.
+
 \section{Measurements Results}
+\label{sec:exp-measurements}
 
@@
 indexes over the data. Then, the numbers regarding the scenarios described
-earlier will be shown:
+earlier will be shown.
 
@@
 
-The insersion time avarages are regarding the different approaches to model a
-document in the database system:
+The insersion time avarages an impressive ~169,508 instances per minute during
+the experiments using the Java driver to connect with the mongoDB server. The
+following lists the storage consumed by the generated data described in the
+section \ref{sec:workload}.
 
 \begin{itemize}
-  \item No Index : Insert Avarage = xyz 
-  \item Indexed : Insert Avarage = xyz
+  \item 1 YSI worth of data = 234.16 MB 
+  \item 5 YSI worth of data = 1.23 GB
 \end{itemize}
@@
 On the other hand, the avarages for each of the scenarios are devided by the
-amount of data produced by one single device or five different devices.
-
-For one single device, the values can be summarized as follows:
-\begin{itemize}
-  \item Not Index : A) = xyz
-  \item Indexed : A) = xzy
-  \item Not Index : B) = xyz
-  \item Indexed : B) = xzy
-  \item Not Index : C) = xyz
-  \item Indexed : C) = xzy
-\end{itemize}
-
-For data worth five different devices, the values can be summarized as follows:
-\begin{itemize}
-  \item Not Index : A) = xyz
-  \item Indexed : A) = xzy
-  \item Not Index : B) = xyz
-  \item Indexed : B) = xzy
-  \item Not Index : C) = xyz
-  \item Indexed : C) = xzy
-\end{itemize}
-
-\subsection{Data-Centric Cluster Environment}
-
-The insertions in a sharded environment are added the operation of finding
-which ``bucket'' the data must be placed. mongoDB gives support for automatic
-sharding after the definition of a key based on range of the values. The
-avarages of the insertions can be shown as follows:
+amount of data produced by one single device or five different devices. For
+data worth five different devices, without indexes in the database, the values
+can be summarized as follows:
 
 \begin{itemize}
-  \item NO Index : Insert Avarage = xyz
-  \item Indexed : Insert Avarage = xyz
-\end{itemize}
-
-For one single device, the values can be summarized as follows:
-
-\begin{itemize}
-  \item Not Index : A) = xyz
-  \item Indexed : A) = xzy
-  \item Not Index : B) = xyz
-  \item Indexed : B) = xzy
-  \item Not Index : C) = xyz
-  \item Indexed : C) = xzy
-\end{itemize}
-
-For data worth five different devices, the values can be summarized as follows:
-
-\begin{itemize}
-  \item Not Index : A) = xyz
-  \item Indexed : A) = xzy
-  \item Not Index : B) = xyz
-  \item Indexed : B) = xzy
-  \item Not Index : C) = xyz
-  \item Indexed : C) = xzy
+  \item A) = ``time.valid'' for week period, or 7 days = 3820ms
+  \item B) = a given IP address by ``sensor.ip\underline{ }address'' = 3721ms
+  \item C) = given values for the keys ``observation.Salinity'' and 
+  ``observation.WaterTemperature'' = 3167ms
 \end{itemize}
@@
 
-mongoDB is that the data can be distributed using Horizontal Partitioning,
-which mitigates problems related to referencial integrity.
+The implementation of a data persistence for any sensor network must take into
+the characteristics of the sensor network, the location of the collected data,
+etc. As described in chapter 4, the taxonomies classify different aspects of
+the persistence of data. Along with it, the design and implementation of a data
+persistence layer for NetBEAMS must meet the requirements and specifications
+described in chapter 5 and chapter 6, respectively. This section discusses the
+experiments scenarios described to evaluate a real-world experiment relating
+the implementation and results to each of the taxonomies. Furthermore, the
+experiments provided by this work cover a one year worth of data for one and
+five YSI devices and evaluates the use cases defined in section
+\ref{sec:exp-scenarios}.
 
-\begin{itemize}
-  \item Strengths as compared to literature
-  \item Weaknesses as compared to literature
-\end{itemize}
+\subsection{Taxonomic Evaluation}
 
-Data-Centric Storage and Database Shards = Horizontal scalability for data
-based on specific keys = Scalability facing High Volume of Collected Data
+Any database system covers the purpose of the sensor data for NetBEAMS, which
+is Data Archival, along with the type of storage for the data as an External
+Storage. After selecting mongoDB, the development of the persistence layer met
+the requirements of archiving data after the randomly generated data for the
+YSI sonde was persisted. As it is shown in section \ref{sec:exp-measurements},
+the collected data is archived in files using the BSON format \cite{bson}, the
+binary representation used to store instances of the document described in
+Listing \ref{file:mongodb-ysi-data-format}. Along with the purpose of data
+archival, the use of mongoDB also met the specification of an external storage
+since the data persisted in the file system.
 
-Weakness, Technological, and skills-based.
+One of the requirements for NetBEAMS is the use of a schema-less data model.
+mongoDB uses a variation of the key-value data representation based on the JSON
+\cite{json}, and provides APIs in different languages for the data creation
+process. In this way, the implementation of the persistence model captures the
+instances of any known data collected from NetBEAMS and saves it using the
+document pattern as shown in Listing \ref{file:mongodb-ysi-data-format}. One
+application of a schema-less data model is the addition of any new key-value
+pair to any instance of the document, as mongoDB supports update operations to
+any of the collections. For instance, annotation in sensor networks is
+widely used to describe the findings of a particular data set
+\cite{sn-annotation-ex1, sn-annotation-ex2}. In case of an oil spill in the San
+Francisco Bay occurs, researchers can update the data set of a given date range
+by adding tags to the relating documents. As a consequence of the use of a
+schema-less data model, these changes do not affect the other data instances,
+since there are no schema changes.
+
+As described in chapter 5, the schema-less data model can provide a better
+support to Data Provenance because the data model gives support to the change
+of records without affecting the others. In this way, the metadata used by the
+implementation described in the previous chapter clearly describes the three
+fundamental questions regarding the collected data from sensor devices, as
+described in section \ref{sec:keys-definition}. First, the identity of the data
+is given by the key ``message\underline{ }id'', in a way that it uniquely
+identifies the data produced by the sensor identified by the key
+``sensor.ip\underline{ }address''. In this way, the message can be tracked back
+to the sensor that produced the data for different purposes such as tracking
+the data producer. Second, time dimensions were used in order to track two
+different aspects of the data collection: the valid time can be used to
+correlate observations to the exact moment it occurred, while the transaction
+time can be used to verify when the data was collected, and make decisions such
+as deleting the instances of data were collected in the previous two days, etc.
+Last, but not least, the use of the key ``sensor.location'' can give the exact
+location of the data, since researchers can be interested in finding data based
+on a specific location. Finally, the most important aspect of the schema-less
+data model is that the observed data is added to the key ``observation'', as
+any sensor device carries its set of attributes of key-value pairs. A good
+example of the applicability of the schema-less data model is the scalability
+regarding changes to the data structure. For instance, if an existing sensor
+device is flushed with a new version of the firmware\footnote{The first set of
+machine instructions to run on the hardware after the application of power stored
+permanently in PROM or ROM or semi-permanently in EPROM}, changing the format
+of data types or adding new values, the following instances of collected data
+may be different to the previous one by simply adding the new additions of
+data. For this reason, the DSP Data Persistence gives administrators the
+ability to choose which fields to persist by declaring it using a bootstrap
+message for the component, as shown in the parameter
+``YSI\underline{ }DESIRED\underline{ }PROPERTIES'' of Listing
+\ref{file:dsp-data-persistence-bootstrap.xml}.
+
+The centralized query processing used by NetBEAMS is a result of the
+infrastructure used by NetBEAMS, as well as the purpose of data and the
+location of the collected data. As a direct result of a schema-less data
+model, the centralized query processing is easily developed using any of the
+available programming languages provided by mongoDB. Considering the fact that
+programming languages are easier abstractions to researchers without computer
+science background \cite{sn-programming-language}, mongoDB's approach to data
+access through API can be seen as an easier task to manage data. As it is shown
+in Listing \ref{file:experiment-query-scenarios}, the scenarios designed to
+evaluate the implementation are written in Javascript, since it is the script
+language by the mongoDB Shell. Therefore, it attends the requirement of not
+using the SQL language as the main query language of the data.
+
+One of the most important features of a data persistence for sensor
+networks is the replacement of the SQL by the use of programming
+languages, as they tend to be easier abstraction to users without the
+background in database systems \cite{sn-programming-language}. In this way,
+mongoDB covers the requirements of providing data access to the collected
+sensor data in different languages. For example, the data collected by the DSP
+Data Persistence component is written in Java, while the implementation of the
+experiments scenarios for querying the collected data is given in Javascript. In
+this way, researchers of the RTC may choose one of the enumarated languages
+APIs provided by the mongoDB to access the stored data from NetBEAMS.
+
+Finally, the implementation supported by mongoDB is either in a single or
+distributed servers. The former approach is done as a regular database server
+on a single host machine, where its main goal is to manage the collected sensor
+data. On the other hand, however, the latter approach can be used to improve
+the performance of the data selection in larger data sets by offering different
+capabilities such as Database Shards and Data Replication.
+
+\subsection{General Evaluation}
+
+In general, the implementation of the DSP Data Persistence component meets
+the specification of the use cases defined in section \ref{sec:exp-scenarios}.
+The implementation of mongoDB using a single server environment can give
+support to the current workload and infrastructure from the use of NetBEAMS to
+collect data from the sensors. From the perspective of disk-space utilization
+and given that ~1.5 Gigabyte of disk storage is necessary to store
+one-year-worth of collected data for five YSI devices, the implementation
+scales to 80 years of data in a server with 120 Gb of available disk space.
+However, if the database server runs out of space, the addition of a new
+database shard can be done.
+
+The performance to insert data into the database system using the Java driver
+is documented by mongoDB as the fastest approach, providing persistence to
+approximaly 169,508 instances of data. Considering this number of insertions to
+the database, it is clear to say the mongoDB implementation can virtually
+support around 169 thousands of sensors a minute. Similarly, the implementation
+of a data-centric scenario using mongoDB can improve the performance of the
+system, given that the schema-less data model gives the ability to partition
+data horizontaly. In this way, latency related to data constraints and data
+integrity checks are avoid because an entire instance of the collected data
+resides in the same ``chunck'' of data. This is the main advantage over the
+relational model, previously discussed in chapter 5.
+
+One possibility of improving performance on the given implementation is the use
+of indexes on each of the keys. As a direct consequence of this procedure,
+insertion time to this approach is approximaly decreased by a 10th of the
+insertion time experienced with the experiments without indexes. Similarly, it
+turns out that the use of indexes improves the query processing by 5 times when
+indexes are not used. Therefore, the use of indexes should be enforced in a
+real-world deployment, since both requirements for insertion and retrieval are
+met.
+
+When it comes to re-using the data collected by the DSP Data Persistence,
+different approaches are available. First, the use of the export capabilities
+to a common-delimited values was performed \ref{file:mongodb-export-command},
+taking approximaly 3.3 minutes to export one year worth of data for 5 YSI
+devices. Second, the use of the client shell gives a centralized access to
+either the single server of sharded server option. However, in order to better
+provide access of data to users without skills in database systems, web
+applications can be built on top of the REST \cite{http-rest} Web Services APIs
+provided by mongoDB. One such example of using the REST API is the adaptation
+of Apache Futon 4 mongoDB, as shown in Figure
+\ref{fig:view-collections-instance-browser-futondb}, where users can navigate
+through the collections of the collected sensor network data categorized by the
+collection. Therefore, RTC staff technicians could take advantage of this tool
+to view the collected data.
+
+The design and implementation of the guidelines designed by the taxonomies can
+be applied to any sensor network that produces data with the purpose of Data
+Archival in External or Data-Centric storage.

### 2009-12-01 docs/marcello-dissertation/thesis/chapter9-conclusions-and-future-works.tex
Updates to the images of the experiments and revisions to different chapters from 4 to 10.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@664 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 
-This work presented a solution to store data from sensor networks.....
-...the main result of this work was that the persistence is
+This work presented a collection of taxonomies related to data persistence for
+sensor networks using different approaches ranging from data models and server
+infrastructure organization. The main result of this work was that persistence
+for collected data from sensors are better maintained using schema-less data
+models, since it provides offers an approach to deal with unanticipated data
+types from the sensor devices.
 
-There are many important features covered in this work, including...
-  - Use of open-source UI for KVP Sensor Data
+There are many important features covered in this work, including the
+importance of the sensor networks rely in how users have access to the
+collected data. In order to provide access to the collected data, one must
+first assess the different types of infrastructural characteristics of the
+studied sensor network. In general, the selection of a technology that
+provides such persistence capabilities can be part of a process of analysis of
+taxonomies proposed by this work. However, this work proposes a data model not
+yet used in the sensor network community and, thus, represents an important
+contribution for the area.
 
-Other important features that COULD HAVE BEEN covered, includes\ldots
-  - Different Dababase system, use of cloud computing, etc
-  - Use of MapReduce technology and implementations, distributed Processing that scales the proposed KVP storage
+Taking into account the software architecture provided by NetBEAMS, as well as
+the requirements for data collection from the RTC research group, this work
+shows a novice way to persist data collected from sensor devices by using a
+schema-less data model, whose query capabilities are based upon programming
+languages, in contrast to the use of the Structured Query Language, SQL. The
+experiment results showed that the implementation of a persistence layer with
+mongoDB can be used to persist data at its maximum rate of one sample per
+minute. However, performance can be improved by using techniques such as
+database sharding, since it offers a direct implementation capability to
+Data-Centric Storage approach. This approach decreases the amount of time to
+search data since it distributes the data instances across different cluster
+nodes.
 
-... The main results from this dissertation were important because of the contributions on new technologies...
-  - An alternative Data Model for Data Storage in Sensor Networks
-  - An alternative apprach for Data-Centric Storage in Sensor Networks
+There are many different directions for future works. First, the Data-Centric
+implementation performance may be extremely improved by using the technique
+called MapReduce \cite{map-reduce}. This approach allows mongoDB to search
+different machines in parallel whenever the data requested is located in
+different database shard hosts. In addition to that, the use techniques to
+cluster collected data before they are inserted into the database may be
+important as shown in \cite{sn-time-series}. One of the key issues in the
+implemented solution is that any measurement of data is considered to be
+persisted without further analysis.
 
-\section{TODO list}
+All in all, the DSP Data Persistence component implemented for the NetBEAMS
+solution to SF-BEAMS can be used in any network server. The solution for data
+persistence is a novice way to save data into the database, provided by
+mongoDB's ability to cope with the uncertainty of any properties collected from
+any sensor specification. Furthermore, the solution also provides a good
+support to the provenance-related data.

### 2009-12-01 docs/marcello-dissertation/thesis/chapter10-appendix.tex
Updates to the abstract
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@665 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 \end{figure}
-
-Installation of mongoDB
-- Explanation about mongoDB processes: mongod, mongos, mongo
-- Specification of the shards, shard key, etc
-
-    * The database instance is called "netbeams";
-    * The database "netbeams" may contain different collections, categorized by
-    the Sensor Content Type, that is, depending on how the DSP Component was
-    described;
 

### 2009-12-01 docs/marcello-dissertation/thesis/preamble.tex
Updates to the abstract
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@665 559edde2-a4d4-44e0-96fc-4035e40d7396

--- 2009-11-11 docs/marcello-dissertation/thesis/preamble.tex
@@
 \onehalfspacing{\noindent      
-Abstract goes here
+Sensor Networks are becoming more and more important to different science
+communities bacause of what it produces: the raw data for different studies.
+In order to make use the collected data, researchers may have to dissect the
+characteristics of the sensor network in question, regarding different
+properties such as the purpose and location of the observed data, as well as
+how the data is described. In this way, this work first proposes a set of
+taxonomies that can be used to study and classify a given instance of sensor
+network and its produced data. Then, based on the classes of these taxonomies, 
+this work selected a Document-Oriented Database system as a result of an
+empirical evaluation of the requirements for a new persistence layer for 
+NetBEAMS, the case study used in this work. For this reason, the design and
+implementation of a new data persistence component for NetBEAMS show how
+schema-less data models help decreasing the complexity related to the database
+layer, not yet reported by the state of the art. Furthermore, this work also
+proposes the use of Database Shards in order support the implementation of
+Data-Centric sensor networks. 
 }
-
 \vspace*{\fill}

### 2009-12-02 docs/marcello-dissertation/thesis/chapter1-introduction.tex
Revision after kleber's feedback on first and part of the second chapters.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@666 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 Sensor networks are commonly used in different areas for different purposes
-such as accurate measurements and the general public use \cite{sn-intro01}
-\cite{sn-intro02}. In the scientific community, they serve as tools to watch
-the state of the environment by storing samples of data at particular periods
-of time. For example, the NASA Jet Propulsion Laboratory uses the Volcano
-SensorWeb \cite{sn-ex02}, a sensor network which aims at collecting data from
-individual volcanoes in Alaska and South Pacific to be used with heuristics
-regarding hazardous activities. Similarly, the National Data Buoy Center, a
-division of the National Oceanic and Atmospheric Administration (NOAA)
-\cite{sn-ex03}, provides online information from different buoys in different
-shores around the world regarding water quality, current information, etc,
-what characterizes sensor network for environmental monitoring \cite{sn-ex01}.
-Therefore, sensor networks can be seen as an extremely valuable scientific
-research and commercial sector with direct application and use of its produced
-data.
+such as accurate measurements during scientific research \cite{sn-intro01} and
+the general public access such as weather forecast \cite{sn-intro02}. In the
+scientific community, they serve as tools to watch the state of the environment
+by storing samples of data at particular periods of time. For example, the
+NASA Jet Propulsion Laboratory uses the Volcano SensorWeb \cite{sn-ex02}, a
+sensor network which aims at collecting data from individual volcanoes in
+Alaska and South Pacific to be used with heuristics regarding hazardous
+activities. Similarly, the National Data Buoy Center, a division of the
+National Oceanic and Atmospheric Administration (NOAA) \cite{sn-ex03},
+provides online information from different buoys in different shores around
+the world regarding water quality, current information, etc, what
+characterizes sensor network for environmental monitoring \cite{sn-ex01}.
+Therefore, sensor networks can be seen as an extremely valuable tool
+for the scientific and commercial sectors, makeing direct use of the sensor
+networks' produced data.
 
 In order to provide access to the collected sensor data, different approaches
-can be taken to interrogate sensor devices indirectly by accessing its data in
-places called network sink. Similarly, the data model and database system
-location may also influence in the process of reusing the collected data. In
-this way, the lack of a persistence layer for existing sensor networks requires
-a deep understanding about its infrastructure, the nature and location of the
+can be taken: interrogate sensor devices indirectly by accessing its data in
+places called network data sink or simply connect directly to the device. The 
+While the latter approach the data is only available in memory as a data
+stream, the former approach contains any sort of secondary memory such as a
+hard-disk or a flash memory, which are used to archive the collected data.
+Along with the location of the collected data from sensors, the the data model
+and may also influence in the process of reusing the collected data. In this
+way, the lack of a persistence layer for existing sensor networks requires a
+deep understanding about its infrastructure, the nature and location of the
 collected data, the chosen data model representation, among others. For
@@
 access may enforce the difficulties, technical issues may also increase while
-the implementation of a sensor network such as growth in terms of the number of
-sensor devices and, consequently, the number of data produced. Last, but not
-least, the way to read the data for reuse is the most striking features of the
-sensor network, taking into account the different data formats used by the
+the implementation of a sensor network such as growth in terms of the number
+of sensor devices and, consequently, the number of data produced. Last, but
+not least, the way to read the data for reuse is the most striking features of
+the sensor network, taking into account the different data formats used by the
 primary users of the system. How can individual collected sensor data be
-exported to other formats such as spreadsheets or archived files meant to be
-shared among research peers? For these reasons, the inception of a persistence
-storage can be an exceptionally challenging problem to solve.
+exported to other formats such as spreadsheets in order to be shared among
+research peers? For these reasons, the inception of a persistence storage can
+be an exceptionally challenging problem to solve.
 

### 2009-12-02 docs/marcello-dissertation/thesis/chapter10-appendix.tex
Revision after kleber's feedback on first and part of the second chapters.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@666 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 
-\chapter{DSP Data Persistence: Component Implementation}
+\section{DSP Data Persistence: Component Implementation}
 \label{sec:dsp-data-persistence-implementation}
@@
 
-\section{DSP Platform Deployment}
+\section{DSP Platform Deployment Details}
 
@@
   \centering
-  \includegraphics[scale=0.3]{../diagrams/view-collections-instance-browser-futondb}
+  \includegraphics[scale=0.5]{../diagrams/view-collections-instance-browser-futondb}
   \caption{Viewing a partial list of data using the Futon for CouchDB/MongoDB}
@@
   \centering
-  \includegraphics[scale=0.3]{../diagrams/view-collected-data-list-browser-futondb}
+  \includegraphics[scale=0.5]{../diagrams/view-collected-data-list-browser-futondb}
   \caption{Viewing a partial list of data using the Futon for CouchDB/MongoDB}
@@
   \centering
-  \includegraphics[scale=0.3]{../diagrams/view-collected-data-instance-browser-futondb}
+  \includegraphics[scale=0.5]{../diagrams/view-collected-data-instance-browser-futondb}
   \caption{Viewing an instance of collected data using the Futon for
@@
 
-\lstinputlisting[language=Java,label=file:experiment-dsp-implementation,caption=Inserting
+\lstinputlisting[language=Java,label=file:experiment-dsp-java-executor,caption=Inserting
 Random Data to MongoDB]{../../../../netbeams/versions/v2/apps/osgi-bundles/dsp/DSPDataPersistence/src/org/netbeams/dsp/persistence/controller/DSPMessageToMongoDBExperiment.java}

### 2009-12-02 docs/marcello-dissertation/thesis/chapter2-state-of-the-art.tex
Revision after kleber's feedback on first and part of the second chapters.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@666 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 
-Different aspects related to two fields of Computer Science have motivated to
-the commencement of this work. In the prospect of Database Systems and
-Distributed Systems, this work seeks a different approach for a persisting
-collected data from sensor networks into a storage system that can be seen as a
-cluster of distributed databases. Likewise, in the perspective of Software
-Engineering, this work adds a crucial capability to an existing sensor network
-framework, meeting the requirements and interfaces. In general, this chapter
-gives the foundation based on the motivations for the establishment of a
-different approach to persisting data using a different data model, covering
-topics that will be referred throughout the following other chapters.
+In general, this chapter reviews sensor networks characteristics and
+properties, necessary to the understanding of the basic properties of a sensor
+network. Moreover, it reviews the characteristics of data persistence for
+sensor networks, characterizing the different approaches used to store data
+such as data model used and how the collected data is described.
 
-This chapter will first give an overview on different related problems
-described in the literature that solved problems related to data modeling and
-representation in sensor networks.
-
-\section{Survey in Sensor Networks}
+\section{Survey in Sensor Networks and Its Data}
 
 Sensor Networks are specialized types of network systems comprised of nodes
-with specific goals to monitor and collect data. Data Persistence in Sensor
-Networks, however, depends on the approaches used to store the data, since data
-nodes can be located close to the devices in the network or in a centralized
-node. In this way, the query process can take place on the network nodes
-themselves or in the centralized data sink. The former approach is used when
-the data is collected at the sensor node on its local storage
-\cite{sn-storage01}\cite{sn-storage03}, strategy typically used to maximize
-the sensor's resources while minimizing energy consumption
-\cite{sn-storage04}. However, this strategy is used when data archival is not
-the main purpose, given the limited resources on the sensor devices. For this
-reason, the latter strategy has been used when the collected data is generally
-archived in a centralized database system for later reuse \cite{sn-storage02}.
+with specific goals such as to produce and collect data. The former is seen
+as a sensor device capable of producing data to the network based on an
+observation from the environment or just a raw value regarding a given
+propertiy, while the latter can be seen as the network sink responsible to
+receive the observed or ``sensed'' data. In this way, data persistence in sensor
+networks depends on the approaches used to store the data on the data sink,
+since these types of nodes can be located in the network close to the devices 
+or in a centralized location. As a consequence, the query used to query the
+data can take place in the network nodes themselves or in the centralized node.
+The former approach is used when the data is collected at the sensor node on
+its local storage \cite{sn-storage01, sn-storage03}, strategy typically
+used to maximize the sensor's resources, while minimizing energy consumption
+\cite{sn-storage04}.The latter strategy, however, has been used when the
+collected data is archived in a centralized database system for later reuse
+\cite{sn-storage02}.
 
@@
 maintaining the database schema normalized, the use of Data Provenance 
-approaches in sensor network data \cite{sn-provenance} usually addresses the
-problem related with the origin of data and its entire history. However, once
-the data model is chosen and implemented, one of the most prominent
-nonfunctional requirements in sensor networks is scalability of the persistence
-storage. Since an operational sensor network can generate enormous amounts of
-data, it must be able to cope with the increase and data load during data
-gathering, and yet provide users the same performance during data retrieval by
-using different strategies such a data-centric storage \cite{sn-storage03}.
+approaches in sensor network data usually addresses the problem related with
+the origin of data and its entire history \cite{sn-provenance}.
 
-The following subsections describe classes of characteristis of these types of
-sensor networks, with applications on research, industrial or militery.
-
-The following subsections describe classes of characteristics of these types of
-sensor networks, with applications on research, commercial or military.
+The following subsections describe the main characteristics of sensor networks
+described in the literature, and then others that interest data persistence in
+sensor networks.
 
@@
 
-Each of the nodes can be deployed in a arbitrary method
-such as being dropped from an airplane, or in a manual mode at carefully chosen
-geographical points \cite{sn-intro01}. Moreover, the deployment can be done like a
-one-time event or an iterative process.
+Each node of the sensor network can be deployed in an arbitrary method such as
+being dropped from an airplane, or in a manually placed at chosen geographic
+locations \cite{sn-intro01}. Moreover, the deployment can be done as a one-time
+event or as an iterative process, depending on the needs of where the sensor
+device needs to be placed.
 
-However, nodes can be able to move around the deployed locality or remain fixed
-in this place. In the case of the amount of the former behaviour, a sensor node
-can move occasionally or continuously, being a result of an accidental
-side-effect or a desired property of the device, where they are an automotive
-device or attached to another moving object. For this reason, \cite{sn-intro01}
-describes those types of mobility as active or passive.
+Nodes can be able to move around the deployed locality or remain fixed at a
+single place. In the case of a sensor networks comprised with different moving
+sensors, a given sensor node can move occasionally or continuously, as a result
+of an accidental side-effect or a desired property of the device. Usually,
+these sensors are automotive or they can be attached to another moving object.
+For this reason, \cite{sn-intro01} describes those types of mobility as active
+or passive.
 
@@
 being limited in size, they are also limited in resources and computational
-power, forcing their  conception limited to maintain its energy stored  either
-on batteries or scavenged from solar cells \cite{sn-intro01}.
+power, as they are designed to either stored energy on batteries or
+use energy scavenged from solar cells \cite{sn-intro01}.
 
-Since sensor devices range in size and capabilities, the cost for each of the
-devices very from hundreds to thousands of dollars, depending on its
-functionality, energy utilization, etc. For this reason, \cite{sn-intro01}
-explains that a sensor network may potentially contain similar sensor devices,
-as they can contain dissimilar ones from each other, which characterizes the
-heterogeneity of a sensor network by the hardware capabilities such as GPS
-enabled or having GSM network connectivity.
+Sensor devices range in size, capabilities and other properties, which are
+associated to their cost ranging from hundreds to thousands of dollars. All in
+all, the price depends on the device's functionality, energy utilization
+capabilities, among others. For this reason, \cite{sn-intro01} explains that a
+sensor network may potentially contain similar sensor devices, as they can
+contain dissimilar ones from each other, which characterizes the heterogeneity
+of a sensor network by the hardware capabilities such as GPS enabled or having
+GSM network connectivity.
 
-\subsection{Communication Modality, Network Infrastructure, Coverage and
+\subsection{Communication Mode, Network Infrastructure, Coverage and
 Connectivity}
@@
 
-Sensor nodes can communicate in different modalities such as radio, diffuse
+Sensor nodes can communicate in different modes such as radio, diffuse
 light, inductive or sound, which may influence on the type of protocols used to
@@
 
-When it comes to the sensor network different organizational infrastructure,
-they can differ in types of topology: single-hop, star, networked stars, tree
-and graph, where each of them may affect the characteristics of the network as
-it may also influence on data collection and routing mechanisms. While in a
-single-hop sensor network there are only two levels of sensor connectivity, the
-tree, graph or networked stars networks uses communications among the network
-hosts for data communication.
+When it comes to infrastructure, sensor networks differ in topology:
+single-hop, star, networked stars, tree and graph are examples of these
+topologies. Each of they topologies may influence on how data is collection
+and routed to their data sink. While there are only two levels of node
+connectivity in sensor networks deployed as single-hop and single start
+topologies, networks whose topology is either tree, graph or interconnected
+stars use communications among each of their nodes for data communication.
 
-Last, but not least, sensor networks may be influenced its coverage designed in
-its purpose, its size in terms of the number of nodes and its relating
-lifetime. As sensors are deployed in different areas, they can be distributed
-in a sparse or dense nature, yet providing essential coverage for the
-environment. However, considering that sensor devices are very likely to
-regular failures, projects have provided redundant nodes \cite{sn-intro01}.
-This level of coverage also influences on the connectivity of the sensor
-nodes, where they can be considered connected when a connection is available
-at all times. On the other hand, nodes can communicate with other nodes in a
-sporadic way. In the former, the sensor node is partitioned, while in the
-latter the nodes are isolated most of the time and communicates at certain
-moments in time.
+Last, but not least, the coverage of a sensor network is related to the
+physical space that is ``visible" to the sensors, which is directly related to
+the purpose of the network, its size in terms of the number of nodes that are
+deployed and the sensor network lifetime. In this way, sensor devices can be
+deployed in different areas and distributed in sparse or dense manners, where
+the sensor network they support is providing the necessary data that
+represents the coverage of the network. However, considering that some types of
+sensor devices are very likely to regular hardware failures, projects have
+provided redundant nodes \cite{sn-intro01} to provide better data coverage.
+
+No matter the type of coverage provided by a sensor network, different levels
+of coverage may also influence on the connectivity of the sensor nodes. In
+order to provide connectivity to redundant nodes, for example, a physical 
+connection must be available at all times. On the other hand, sensor nodes can
+also communicate with other nodes in a sporadic way when they are designed to
+be isolated with only communication activities to report the collected data to
+their immediate peer (other nodes and/or data sink).
 
@@
 
-Similar to the degree of coverage of the sensor network is the number of
-devices deployed in the network, where they can be in the order of dozens,
-hundreds or even thousand of nodes \cite{sn-intro01}.
+The network size is related to the degree of coverage of a sensor network, and
+it is measured by the number of devices deployed as nodes of the network.
+\cite{sn-intro01} reports examples of sensor networks with sensor nodes ranging
+from dozens, hundreds or even thousand of network nodes.
 
-The application of the sensor network determines how long the sensor network
+The application of the sensor network determines for how long the sensor network
 may be exist. For example, the existence of an environmental sensor networks
-\cite{sn-ex01} may be not determined since its use serves in a regular basis,
-whereas the lifetime of a sensor network to determine the number of people in
-an event is limited to the event.
+\cite{sn-ex01} may be not determined since its the data collected from the
+network are used for archival purposes, hence, without termination of the
+execution of its sensing tasks. On the other hand, however, the lifetime of a
+sensor network designed to count the number of people who entered in different
+rooms of a building during an event has its life-time related to the period of
+the event.
 
-Finally, the design of sensor networks can be
-related to the constraints related to quality-of-service (QoS) such as
-real-time delivery within a certain period of time or robustness such as
-maintaining data alive even though there a communication link is not available
-\cite{sn-intro02}.
+Finally, the design of sensor networks can be related to the constraints
+related to quality-of-service (QoS) such as real-time delivery within a
+certain period of time or robustness such as maintaining data alive even
+though there a communication link is not available \cite{sn-intro02}.
 
@@
 
-\subsection{The Purpose of the Collected Data}
+\subsection{How the Collected Data is Used}
 \label{sec:sn-data-purpose}
@@
 starting from its transmission from its producing device to the specified data
-sink as specified by \cite{sn-storage01}\cite{sn-storage02}\cite{sn-storage03}.
+sink as specified by \cite{sn-storage01, sn-storage02, sn-storage03}.
 
@@
 Given that collected data can be used for different purposes as described in
-section \ref{sec:sn-data-purpose}, the sensor node that produces the data is responsible to send
-the raw data to a specified data storage target. Upon producing data, a sensor
-network node is specified to deal with the produced data where, in most cases,
-depending on the sensor network infrastructure used \cite{sn-storage03} (see
-section \ref{sec:sn-infrastructure}). In such a way, different storage
-mechanisms are provided:
+section \ref{sec:sn-data-purpose}, the sensor node that produces the data is
+responsible to send the raw data to a specified data storage target. Upon
+producing data, a sensor network node is specified to deal with the produced
+data where, in most cases, depending on the sensor network infrastructure used
+\cite{sn-storage03} (see section \ref{sec:sn-infrastructure}). In such a way,
+different storage mechanisms are provided:
 
@@
 node to another until it reaches its final storage node as described by
-\cite{sn-storage01}\cite{sn-storage03}.
+\cite{sn-storage01, sn-storage03}.
 
@@
 proven alternative for the purpose of energy consumption
-\cite{sn-storage03}\cite{sn-storage01}.Its primary goal is to organize the
+\cite{sn-storage03, sn-storage01}.Its primary goal is to organize the
 collected data based on a particular property, say location or event, in order
@@
   it \cite{sn-xml-usage01}\cite{sn-xml-usage02}, implementing its query by
-  using a middleware \cite{sn-xml-middleware}\cite{sn-xml-query-engines}.
+  using a middleware \cite{sn-xml-middleware, sn-xml-query-engines}.
 \end{itemize}
@@
 
-\subsection{Relational Data Model and Query}
+\subsection{Relational Data Model and its Query Mechanism}
 
@@
 
-\subsection{Structured Data Model and Query}
+\subsection{Structured Data Model and its Query Mechanism}
 
@@
 
-\section{Data Provenance: How to describe collected data?}
+\section{Data Provenance: How Collected Data is Described?}
 \label{sec:sn-provenance}
@@
 as a noteworthy feature to reduce the traffic generated in a sensor network by
-clustering similar data when the sensor devices are placed in concentrated environment.
+clustering similar data when the sensor devices are placed in concentrated 
+environment.

### 2009-12-04 docs/marcello-dissertation/thesis/chapter1-introduction.tex
Print reviews of chapter 1 and 2, as suggested by Kleber. It definitely helped to see things we can't see when writing on the computer screen. :)
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@668 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 Therefore, sensor networks can be seen as an extremely valuable tool
-for the scientific and commercial sectors, makeing direct use of the sensor
-networks' produced data.
+for the scientific and commercial sectors, whose interest is the use of
+the sensor's produced data.
 
-In order to provide access to the collected sensor data, different approaches
-can be taken: interrogate sensor devices indirectly by accessing its data in
-places called network data sink or simply connect directly to the device. The 
-While the latter approach the data is only available in memory as a data
-stream, the former approach contains any sort of secondary memory such as a
-hard-disk or a flash memory, which are used to archive the collected data.
-Along with the location of the collected data from sensors, the the data model
-and may also influence in the process of reusing the collected data. In this
-way, the lack of a persistence layer for existing sensor networks requires a
-deep understanding about its infrastructure, the nature and location of the
-collected data, the chosen data model representation, among others. For
-instance, how can a data persistence layer safely represent the collected
-sensor devices' properties without the necessity of restructuring any existing
-data model? Should the end users' skills in database system be taken into
-account or the use of programming languages to access the data
-\cite{sn-programming-language}? This problem is currently faced by the
-department of Biology at San Francisco State University with a marine sensor
-network managed by researchers without Computer Science skills. While the data
-access may enforce the difficulties, technical issues may also increase while
-the implementation of a sensor network such as growth in terms of the number
-of sensor devices and, consequently, the number of data produced. Last, but
-not least, the way to read the data for reuse is the most striking features of
-the sensor network, taking into account the different data formats used by the
-primary users of the system. How can individual collected sensor data be
-exported to other formats such as spreadsheets in order to be shared among
-research peers? For these reasons, the inception of a persistence storage can
-be an exceptionally challenging problem to solve.
+The first motivation for the inception of this work was to provide a data
+storage system for NetBEAMS \cite{netbeams2009}, a component-based sensor
+network infrastructure that can be used to automate the process of data
+gathering in sensor networks. In this way, the review of the state of the
+art in rewiewed that different approaches can be taken to provide access to the
+collected sensor data: interrogate sensor devices indirectly by accessing its
+data in places called network data sink or simply connect directly to the
+device. While the collected data is only available in-memory as a data stream
+in the latter approach, it is stored in in any sort of secondary memory
+device such as a hard-disk or a flash memory for archival purposes in the
+former approach. Along with the location of the collected data, the data model 
+may also influence in the process of reusing the collected data. In this way,
+the introduction of a persistence layer for existing sensor networks requires
+an understanding about its infrastructure, the nature and the location of the
+collected data is temporalily located after it is produced from the sensor
+device. Similarly related to the infrastructure, the data model used to
+represent the collected data is another aspect of the data persistence layer
+implemented to manage the raw data values, and may rise questions. Considering
+that sensor networks devices may change over time, how can a data model take
+into account future additions of new entities to the database schema? Should 
+such a solution take into account the end users' skills in database system or
+provide access to the data through programming languages
+\cite{sn-programming-language}? These questions are faced when the solution
+considers the primary users of the collected data such as Biologists from the
+the department of Biology at San Francisco State University, which are
+responsible to maintain SF-BEAMS \cite{sfbeams2006}, a marine sensor network
+used by NetBEAMS. While access to the collected data may represent
+difficulties, other technical issues may also be faced such as growth in terms
+of the number of sensor devices and, consequently, the number of data
+produced. In this way, sensot networks engineered with the purpose of data
+archival has to answer questions related to how disk storage can be maximized
+when storing the collected data. Last, but not least, the way to read the data
+for reuse is the most striking features of a sensor network, taking into
+account the different data formats used by the primary users of the system.
+How can individual collected sensor data be exported to other formats such as
+spreadsheets in order to be shared among research colleagues? For these and
+other reasons, the inception of a persistence storage for NetBEAMS, and any
+other sensor network, can be an exceptionally challenging problem to solve.
 
 In view of the fact that the persistence storage of a sensor network imposes
-many different challenges, this work proposes the assessment of the current
-state of the art for sensor networks in order to provide a data persistence
-layer for an existing sensor network. As a matter of fact, this work's first
-contribution is the creation of a set of basic taxonomies for data persistence
-in sensor networks, proposed to drive an empirical analysis of different
-technologies used in the sensor network community. Furthermore, it also takes a
-look at different alternatives not yet explored through experiments for a
-selected sensor network as a case study. Similarly, this work makes another
-novice contribution by suggesting a ``lighter" data model and database
-infrastructure that simply captures sensor network devices' observations,
-taking into account the non-predictable scalability of its infrastructure, as
-well as aspects of the nature of the data such as identity, time and location
-of the collected data.
+many different challenges, the assessment of the current state of the art for
+data persistence in sensor networks revealed different alternatives to better
+evaluate database technologies for sensor networks and, specifically speaking,
+to NetBEAMS. As a matter of fact, this first contribution of this work is the
+creation of a set of basic taxonomies for data persistence in sensor networks,
+proposed to drive an empirical analysis of different database systems used by
+researchers. In this way, based on the findings of the review of current trends
+in database systems \cite{db-is-rdbs-dommed, db-shard-intro} and distributed
+systems such as Cloud Computing \cite{cloud-comp-survey}, this work proposes a
+data persistence solution based on mongoDB \cite{mongodb}, a schema-less
+document-oriented database system. This selection was based after an empirical
+analysis of the different characteristics of the relational
+\cite{relational-model} and the key-value pair \cite{db-kvp} data models.
+After designing and implementing a solution whose technology has not yet been
+used in the sensor network community, experiments were conducted to simulate
+the workload of real-world scenarios used by SF-BEAMS during one year of data
+produced by a specific type of sensor. In conclusion, the implementation
+attends most of the requirements and specifications defined by the taxonomies,
+giving an alternative data mode that solves the problem related to constant
+schema changes, providing an outstanding support for Data Provenance
+\cite{sn-provenance} for sensor networks. In addition, the solution provides
+different data access alternatives such as the use of programming languages or
+REST Web Services \cite{http-rest} instead of the Structure Query Language SQL
+\cite{sql} used in relational databases \cite{relational-model}, as well as
+native support to export data to formats used by the SF-BEAMS staff.
 
@@
 sensor networks in chapter 3. Then, chapter 4 examines the selected case study
-and its requirements for a persistence layer, whose technology properties are
-evaluated in chapter 5 against the developed taxonomies. The culmination of
+and its requirements for a persistence layer, whose technology's properties are
+evaluated against the developed taxonomies in chapter 5. The culmination of
 this work is chapter 6, which details the design and implementation of a data
 persistence solution for the selected case study. Likewise, chapter 7 discusses
-the solution by presenting the measurements and conducted experiments, and
-describes the findings and contributions of this work. Last, but not least,
-chapter 8 presents the conclusions of this work, together with the possible
-future works not addressed here.
+the solution by presenting the measurements and results of the conducted
+experiments. Last, but not least, chapter 8 presents the conclusions of this
+work, together with the different possible future works not addressed by this
+dissertation.

### 2009-12-04 docs/marcello-dissertation/thesis/chapter2-state-of-the-art.tex
Print reviews of chapter 1 and 2, as suggested by Kleber. It definitely helped to see things we can't see when writing on the computer screen. :)
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@668 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 
-In general, this chapter reviews sensor networks characteristics and
-properties, necessary to the understanding of the basic properties of a sensor
-network. Moreover, it reviews the characteristics of data persistence for
-sensor networks, characterizing the different approaches used to store data
-such as data model used and how the collected data is described.
+In general, this chapter reviews surveys and related papers in networks
+networks, providing a basic introduction about their characteristics to new 
+audience. Similarly, it reviews the characteristics of related works that
+discusses and provides data persistence solutions for sensor networks, 
+characterizing the different approaches used to described and model the
+gathered data from sensor devices.
 
-\section{Survey in Sensor Networks and Its Data}
+\section{Survey in Sensor Networks and Its Infrastructure}
 
-Sensor Networks are specialized types of network systems comprised of nodes
-with specific goals such as to produce and collect data. The former is seen
-as a sensor device capable of producing data to the network based on an
-observation from the environment or just a raw value regarding a given
-propertiy, while the latter can be seen as the network sink responsible to
-receive the observed or ``sensed'' data. In this way, data persistence in sensor
-networks depends on the approaches used to store the data on the data sink,
-since these types of nodes can be located in the network close to the devices 
-or in a centralized location. As a consequence, the query used to query the
-data can take place in the network nodes themselves or in the centralized node.
-The former approach is used when the data is collected at the sensor node on
-its local storage \cite{sn-storage01, sn-storage03}, strategy typically
-used to maximize the sensor's resources, while minimizing energy consumption
-\cite{sn-storage04}.The latter strategy, however, has been used when the
-collected data is archived in a centralized database system for later reuse
-\cite{sn-storage02}.
+According to \cite{sn-intro02}, sensor networks are specialized types of
+network systems comprised of nodes with specific goals such as to produce and
+collect data. The former is seen as a sensor device capable of producing data
+to the network based on an observation from the environment or simply a raw
+data. On the other hand, the latter is described as the final destination of
+the raw data in a sensor network and called the network data sink, or simply
+data sink, whose responsibility is to receive, categorize and store the
+``sensed'' data. In this way, the location of the the data sink is determined by
+the purpose of the collected data: data nodes are placed in-network close to
+the sensor devices when collected data has a life cycle, being discarded after
+being utilized. In contrast, data nodes are placed in in a centralized node
+when the collected data is used for historical purposes.
 
-When it comes to data representation, the relational data model
-\cite{relational-model} is one of the most used approaches to represent data in
-sensor network databases in any of the query processing strategies and, in most
-cases, with a modified version of the SQL language \cite{sn-db-newop}. However,
-the design of the database system must be done with prior knowledge of the
-sensor types when choosing the relational data model, as the database design
-must be normalized \cite{db-normalization} for the entities already identified
-for the system. For example, the introduction of a new sensor device might
-represent a potential change on the database structure by seeking a new
-normalized version of the database design. In addition to challenges of
-maintaining the database schema normalized, the use of Data Provenance 
-approaches in sensor network data usually addresses the problem related with
-the origin of data and its entire history \cite{sn-provenance}.
+The following sections describe general properties of sensor networks regarding
+infrastructure and components as described in the literature.
 
-The following subsections describe the main characteristics of sensor networks
-described in the literature, and then others that interest data persistence in
-sensor networks.
+\subsection{Properties of Sensor Device Nodes in Sensor Networks}
 
-\subsection{Deployment and Mobility}
+Each sensor node perform different tasks for the network and, therefore,
+influences on the method of deployment used. \cite{sn-intro01} identifies that
+the deployment of sensor devices can use arbitrary methods such as being
+dropped from an airplane, or manually placed at chosen geographic locations.
+Moreover, those authors also mentioned that the deployment can be done as a
+one-time event or as an iterative process, where the sensor devices are moved
+to different places as needed.
 
-Each node of the sensor network can be deployed in an arbitrary method such as
-being dropped from an airplane, or in a manually placed at chosen geographic
-locations \cite{sn-intro01}. Moreover, the deployment can be done as a one-time
-event or as an iterative process, depending on the needs of where the sensor
-device needs to be placed.
+Moving to a differnt point, \cite{sn-intro01} reports that sensor nodes can be
+able to move around the place where they were deployed or a stationary sensor
+at a given location. In the case of a sensor networks comprised with different
+moving sensors, a given sensor node can move occasionally or continuously, as
+a result of an accidental side-effect or a desired property of the device. As
+a matter of fact, these sensors are automotive or they can be attached to
+another moving object. For this reason, \cite{sn-intro01} classifies those
+types of mobility as active or passive.
 
-Nodes can be able to move around the deployed locality or remain fixed at a
-single place. In the case of a sensor networks comprised with different moving
-sensors, a given sensor node can move occasionally or continuously, as a result
-of an accidental side-effect or a desired property of the device. Usually,
-these sensors are automotive or they can be attached to another moving object.
-For this reason, \cite{sn-intro01} describes those types of mobility as active
-or passive.
+Considering the physical design and capabilities of sensor devices, they range
+from microscopically small to bigger devices of a few cubic feet
+\cite{sn-intro02}. In addition to their limited size, their computational
+resources and power availability are also scarse, when they are located in
+areas of limited access to energy. In this way, \cite{sn-intro01} also
+distinguishes how sensor devices are usually designed to either store energy on
+batteries or use energy scavenged from solar cells. Even though a sensor
+network is designed for a purpose, \cite{sn-intro01} also explains that it can
+be comprised of different sensor devices playing different roles on what they
+observe in the environment, characterizing the heterogeneity the sensor
+network by the hardware capabilities and what types of data each of them
+produce.
 
-\subsection{Device Size, Resources, Cost, Energy, Heterogeneity}
+Finally, \cite{sn-intro01} mentions that sensor device nodes can communicate
+with each other using different media such as radio, diffuse light or
+inductive sound, influencing on the type of protocols they exchange data. They
+can have knowledge of their physical location if they are integrated with a
+GPS\footnote{Global Positioning System} device, which gives the latitude or
+longitude coordinate values. Therefore, their cost of these devices may range
+from hundreds to thousands of dollars, depends on a given device's functionality
+and capabilities of its design.
 
-Such sensor nodes can be of different sizes, ranging from microscopically small
-particles, small devices to bigger ones with a few feet cubic. In addition to
-being limited in size, they are also limited in resources and computational
-power, as they are designed to either stored energy on batteries or
-use energy scavenged from solar cells \cite{sn-intro01}.
-
-Sensor devices range in size, capabilities and other properties, which are
-associated to their cost ranging from hundreds to thousands of dollars. All in
-all, the price depends on the device's functionality, energy utilization
-capabilities, among others. For this reason, \cite{sn-intro01} explains that a
-sensor network may potentially contain similar sensor devices, as they can
-contain dissimilar ones from each other, which characterizes the heterogeneity
-of a sensor network by the hardware capabilities such as GPS enabled or having
-GSM network connectivity.
-
-\subsection{Communication Mode, Network Infrastructure, Coverage and
-Connectivity}
+\subsection{Properties of the Infrastructure of Sensor Networks}
 \label{sec:sn-infrastructure}
 
-Sensor nodes can communicate in different modes such as radio, diffuse
-light, inductive or sound, which may influence on the type of protocols used to
-communicate among the nodes, where they can be said to be in an organized
-infrastructure or as an ad hoc way \cite{sn-intro01}.
+Last section discussed the different properties of the sensor devices when
+placed in a sensor network. Depending on how they communicate with each other,
+\cite{sn-intro02} discusses that sensor devices can be in an organized in a
+structured or ad hoc way. For this reason, the infrastructure, of sensor
+networks differ in topology such as single-hop, star, tree or  graph. Each of
+these topologies may influence on how the produced data from sensor nodes is
+collected and routed to their related data sink. While there are only two
+levels of node connectivity in sensor networks deployed as single-hop and start
+topologies, sensor networks whose topology is either tree or graph use
+multi-hop communication among each of their nodes for data communication from
+the devices to the data sink.
 
-When it comes to infrastructure, sensor networks differ in topology:
-single-hop, star, networked stars, tree and graph are examples of these
-topologies. Each of they topologies may influence on how data is collection
-and routed to their data sink. While there are only two levels of node
-connectivity in sensor networks deployed as single-hop and single start
-topologies, networks whose topology is either tree, graph or interconnected
-stars use communications among each of their nodes for data communication.
+The property of coverage on a sensor network is defined by
+\cite{sn-intro01} as the physical space ``visible" to the sensors devices, which
+is also directly related to the purpose of the network, its size in terms of
+the number of nodes deployed and the sensor network lifetime. Therefore,
+sensors are deployed in sparse or dense manners, according to the
+specification of what needs to be covered by the sensor devices. Furthermore,
+considering that some types of sensor devices are very likely to regular
+hardware failures, \cite{sn-intro2} described sensor network examples with 
+redundant nodes that provide better data coverage and levels of coverage.
+Finally, the coverage of sensor devices also influence on the connectivity
+property of the sensor nodes. In order communicate with redundant nodes, a
+physical connection must be available at all times. On the other hand, lower
+level of coverage is is related sporadic communication among the sensor nodes.
 
-Last, but not least, the coverage of a sensor network is related to the
-physical space that is ``visible" to the sensors, which is directly related to
-the purpose of the network, its size in terms of the number of nodes that are
-deployed and the sensor network lifetime. In this way, sensor devices can be
-deployed in different areas and distributed in sparse or dense manners, where
-the sensor network they support is providing the necessary data that
-represents the coverage of the network. However, considering that some types of
-sensor devices are very likely to regular hardware failures, projects have
-provided redundant nodes \cite{sn-intro01} to provide better data coverage.
+The sensor network size is related to the degree of coverage mentioned earlier,
+and it is measured by the number of devices deployed. \cite{sn-intro01}
+reports examples of sensor networks composed by different ranges of deployed
+sensor devices such as dozens, hundreds or even thousands of nodes. However,
+the life time of a sensor network determines how long these devices are going
+to be used, as well as how long the sensor network may exist. For example, an
+environmental sensor network \cite{sn-ex01} with the purpose of collecting data
+for historic purposes may be have their sensor devices continuously performing
+their data collection tasks. On the other hand, however, the lifetime of a
+sensor network designed to achieve a goal may not exist after the sensor
+devices finish their tasks. Finally, \cite{sn-intro02} argues that the
+performance related to the orchestration of sensor devices in sensor networks
+is related to the constraints defined for quality-of-service (QoS) such as
+real-time data delivery within a thredshold or for robustness such as
+maintaining data alive even though a communication link is not available.
 
-No matter the type of coverage provided by a sensor network, different levels
-of coverage may also influence on the connectivity of the sensor nodes. In
-order to provide connectivity to redundant nodes, for example, a physical 
-connection must be available at all times. On the other hand, sensor nodes can
-also communicate with other nodes in a sporadic way when they are designed to
-be isolated with only communication activities to report the collected data to
-their immediate peer (other nodes and/or data sink).
-
-\subsection{Network Size, Lifetime and Quality-Of-Service}
-
-The network size is related to the degree of coverage of a sensor network, and
-it is measured by the number of devices deployed as nodes of the network.
-\cite{sn-intro01} reports examples of sensor networks with sensor nodes ranging
-from dozens, hundreds or even thousand of network nodes.
-
-The application of the sensor network determines for how long the sensor network
-may be exist. For example, the existence of an environmental sensor networks
-\cite{sn-ex01} may be not determined since its the data collected from the
-network are used for archival purposes, hence, without termination of the
-execution of its sensing tasks. On the other hand, however, the lifetime of a
-sensor network designed to count the number of people who entered in different
-rooms of a building during an event has its life-time related to the period of
-the event.
-
-Finally, the design of sensor networks can be related to the constraints
-related to quality-of-service (QoS) such as real-time delivery within a
-certain period of time or robustness such as maintaining data alive even
-though there a communication link is not available \cite{sn-intro02}.
+All things considered, sensor nodes play an important role during the execution
+of a sensor network, whose infrastructure is characterized by how and where the
+sensor nodes are deployed and communicate with each other. In this way, next
+section covers the approaches to deal with the product of this type of system
+network.
 
@@
 
-The most valuable objective of a sensor network is related to its first
-product: the use of the raw data. Data collection in sensor networks is
-directly dependent on the specifications of the sensor network, its routing
-mechanisms and the final destination of the produced data. In a nutshell,
-\cite{sn-storage03} describes data as follows:
+As traced in the last few sections, sensor networks are comprised of
+different types of sensors nodes. However, an important sensor node type is the
+data sink because it is the place where the raw data produced by sensor
+nodes resides. Data collection and access in sensor networks is directly
+related to the characteristics of the sensor network, its routing mechanisms
+and where the data sink is placed within the network. In a nutshell,
+\cite{sn-storage03} describes collected data as can be described as Low-level
+property or an observation such as the level of the ocean, etc.
 
-\begin{itemize}
-  \item \textbf{low-level property} as the raw data collected from the sensor
-  device; For example, the sea level as a number;
-  \item \textbf{Observation or Event} as the processed data in the format of
-  the final observation. For instance, if the ocean is in its highlight or low
-  tight.
-\end{itemize}
+The type of query mechanism used to directly or indirectly interrogate sensor
+devices in a sensor network is directly related to the aforementioned locations
+of the collected data. The in-network query processing is used when the
+location of the collected data is a local storage \cite{sn-storage01,
+sn-storage03} located in the network, strategy typically used to maximize the
+sensor's resources, while minimizing energy consumption \cite{sn-storage04}. On
+the contrary, the a typical centralized query processing is used to search for
+values over the collected data when its location is a centralized data
+node \cite{sn-storage02}.
 
-\subsection{How the Collected Data is Used}
+When it comes to the data representation, the relational data model
+\cite{relational-model} is one of the most traditional approaches used to
+describe the collected data of sensor networks, as it is reported in database
+systems such as the TinyDB \cite{sn-db-tinydb}. For instance, \cite{sn-db-newop}
+proposes a modified version of the SQL language of TinyDB proposed to solve
+SQL Join problems, so that the query processing takes into account the location
+of the distributed data. Conversely, \cite{sn-provenance} proposes the use of
+Data Provenance as the primary guidelines to describe collected data from
+sensor devices, giving the fact that these devices only produce observations as
+stream of raw data, and therefore, described to address the problems related to
+the origin of the data.
+
+The following sections describe general properties of sensor networks regarding
+data persistence, as reviewed in the literature.
+
+\subsection{Data Use and Where Data is Stored}
 \label{sec:sn-data-purpose}
+\label{sec:sn-storage-locations}
 
 At first glance, the raw data serves as the main outcome of the sensor network.
-However, the first consideration that must be considered is regarding the
-purpose of the collected data, as \cite{sn-provenance} explains differentiates
-between two main benefit of the sensor data:
+\cite{sn-provenance} differentiates the use of the collected data as follows as 
+\textbf{real-time data stream}, when they are generated data from a sensor
+device whenever the data is requested. In this way, the data is still in its raw
+format specified by the sensor manufacturer; On the other hand, the authors
+describes the data as \textbf{archivable}, since they a have historical
+significance and will be processed to be reused by other systems in a latter
+time. While the real-time data streams are used and potentially discarded, the
+archivable data is stored into secondary memory such as a hard-disk. Under
+these circumstances, the collected data needs first to face its journey
+starting from its production, transmission from the sensor device, to finally
+reach a specified data sink in the network \cite{sn-storage01, sn-storage02,
+sn-storage03}.
 
-\begin{itemize}
-  \item \textbf{Real-time Data Stream}, as the generated data from the sensor
-  is not transmitted, but extracted from the device as it is requested. The
-  data is still in its standard form specified by the sensor manufacturer;
-  \item \textbf{Archival Data}, as the collected data include historical
-  significance or will be use by other applications at a later time.
-\end{itemize}
+After producing the ``sensed'' values for the observed parameters from the
+environment, the sensor node is responsible to send the raw data to a specific
+data node defined by the network infrastructure, as delineated in section
+\ref{sec:sn-infrastructure}. In such a way, \cite{sn-storage03}
+describes that the location for the collected sensor data provides two
+different types of storage devices: the \textbf{external storage} and the
+\textbf{local storage}. In the former type, the sensor device sends the
+produced data to another sensor node that contains an external storage, whereas
+the data only exists in memory or between its reading to its delivery as
+real-time data stream in the the latter type. On the whole, the sensor network
+infrastructure dictates where the movement performed by the sensor data, as
+\cite{sn-storage02} sees a wireless sensor networks as a many-to-one data
+collection pattern, when its organization is based on single-hop or star. On
+the other hand, data may also flow from one intermediate sensor node to another
+until it reaches its final data sink when the data organization is structured
+as a graph, tree \cite{sn-storage01, sn-storage03}.
 
-While real-time data stream are used and can potentially be discarded, the
-archival data needs to be stored in a secondary memory. Nonetheless, given the
-nature of the sensor device specification, the low-level data produced by a
-given sensor device might not include vital information regarding the identity
-of the data itself. The use of Data Provenance techniques is described by
-\cite{sn-provenance} as an essential step to provide the collected data
-specific metadata regarding its location and lineage.
+Last, but not least, the use of different external storage devices can be used
+to partition the collected data. This practice is called \textbf{Data-Centric
+Storage}, and aims at clustering the collected data from different sensor
+devices based on a given a property of the data such as location or time of
+creation.
 
-Under these circumstances, the collected data needs first to face its journey
-starting from its transmission from its producing device to the specified data
-sink as specified by \cite{sn-storage01, sn-storage02, sn-storage03}.
-
-\subsection{Storage Location for Collected Data}
-\label{sec:sn-storage-locations}
-
-Given that collected data can be used for different purposes as described in
-section \ref{sec:sn-data-purpose}, the sensor node that produces the data is
-responsible to send the raw data to a specified data storage target. Upon
-producing data, a sensor network node is specified to deal with the produced
-data where, in most cases, depending on the sensor network infrastructure used
-\cite{sn-storage03} (see section \ref{sec:sn-infrastructure}). In such a way,
-different storage mechanisms are provided:
-
-\begin{itemize}
-  \item \textbf{External Storage}: the sensor device sends the produced data to
-  another sensor node that contains an external storage.
-  \item \textbf{Local Storage}: as it is the case of real-time data stream,
-  where data is allocated on the sensor device's main read-only memory.
-\end{itemize}
-
-It is clear that the data flows from the sensor node to another depending on
-the definition of the network. On the whole, the sensor network infrastructure
-dictates where the data will flow to, as described in section
-\ref{sec:sn-infrastructure}. \cite{sn-storage02} sees a wireless sensor
-networks as a many-to-one data collection pattern, when its organization is
-based on single-hop or star. On the other hand, when the data organization is
-structured as a graph, tree, etc, the data flows from one intermediate sensor
-node to another until it reaches its final storage node as described by
-\cite{sn-storage01, sn-storage03}.
-
-Last, but not least, yet a different storage strategy can be used: Data-Centric
-Storage. This strategy aims at clustering the collected data based on a given
-nature of the data such as location. Regardless the data storage mechanism
-used, the query processing will depend on the sensor network infrastructure and
-data location.
-
-\subsection{Query Processing for Storage Locations}
+\subsection{Query Processing in Sensor Networks}
 \label{sec:query-process}
 
-Given the different locations where the collected data might be as described in
-section \ref{sec:sn-storage-locations}, different query processing for the
-collected sensor data can be used as described by \cite{sn-storage03}:
+The query processing in sensor networks is a direct or indirect method of
+accessing the collected data stored at data nodes, whose location is determined
+by different facts described in the previous section. \cite{sn-storage03}
+describes an example of an \textbf{in-network query processing} when sensor
+nodes, acting as data nodes, provides real-time data stream as a response
+to queries requesting the device's collected data. Likewise, if the collected
+data is in use for a specific period of time, the data is cached in an external
+storage device in a close-by data node. In this way, when a query is issued to
+the sensor network with these characteristics, the query processor considers 
+all data nodes return its final response to the query. In comparison, a sensor
+network whose topology is in form of star, the data sink is usually located in
+the center of structure, and thus, receives all the collected data from other
+sensor devices in the network. For this reason, it can be seen as a 
+\textbf{centralized query processing}. Factoring in that a centralized data
+sink receives all the collected data from the peer nodes, this query processing
+strategy is the solo responsible for responding to indirect queries about the
+produced data from the network.
 
-\begin{itemize}
-  \item \textbf{In-Network Query Processing}: sensor nodes that provide
-  real-time data stream usually replies to requests to properties.
-  Likewise, when the data is used for archival purpose, the data can be
-  located at a network node specialized in storing the neighbors
-  collected data. In this way, when a query is issued to the sensor
-  network, the query processor will collectively verify every storage
-  node for the collected data.
-  \item \textbf{Centralized Query Processing}: a leading network sink receives
-  the collected data from the peer nodes. In this case, the query processor
-  queries the solo centralized node for the global values of the
-  participating sensor devices.
-\end{itemize}
+Based on the literature review, there are advantages and disadvantages related
+each of the query processing approaches described with regards to the
+infrastructure of the sensor network. \cite{sn-storage03} argues that the 
+primary purpose of the in-network query processing is to minimize the number
+of data transmitted from sensor devices to a data sink, as sensor devices are
+generally limited by energy use. Considering that sensor devices are also used
+as data nodes, \cite{sn-storage04} delineate that this approach can potentially
+flood all sensor nodes in the network, and therefore, decrease its performance.
+For this reason, \cite{sn-storage01} reports the use of the data-centric
+storage by organizing the collected data based on its particular properties
+values. In this way, the query processor hits specific data nodes that contains
+the types of data in the range of the requested data and, therefore, the author
+concludes that this approach decreases the network congestion and maximizes the
+capacity on the sensor devices and the performance of the sensor network.
 
-There are pros and cons related to both of the approaches of query processing,
-which are also related to the infrastructure, sensor network nature and
-requirements. The primary purpose of the in-network query processing is to
-minimize the number of data transmitted from the sensor devices to the final
-data sink, as sensor devices are generally limited by constraints such as
-energy utilization \cite{sn-storage03}. Along with this strategy, the literature
-describes different proposals that aim at upgrading the Structured Query
-Language (SQL) adding location-sensitive clauses to the language \cite{sn-db-newop}.
-However, this crucial problem related to this method is that it may potentially
-flood all nodes with query messages \cite{sn-storage04}.
+Unlike the in-network, the centralized query processing is related to a regular
+database system, located in one of the network nodes. For this reason, this
+indirect type of query processing is used when the data purpose of the use of
+data is for data archival. However, \cite{sn-storage02} showes that as a
+consequence of the many-to-one communication between the centralized data sink
+and the other data nodes, this query process does not affect any of the sensor
+nodes in the network and, thus, maximizes the sensor device utilization. 
+Moreover, this strategy better exposes the collected data to the primary users
+raw data, which follows the process of data processing, compression,
+aggregation, and so on \cite{sn-db-modeling02}.
 
-In order to mitigate the problems related flooded nodes in in-network query
-processing, the use of a Data-Centric storage strategy has been shown to be a
-proven alternative for the purpose of energy consumption
-\cite{sn-storage03, sn-storage01}.Its primary goal is to organize the
-collected data based on a particular property, say location or event, in order
-to be sent to individual storage nodes that relate to the observation. As the
-main outcome of this strategy is to create network partitioning strategy, the
-query processor would hit specific sensor nodes with the needed data and,
-therefore, would decrease the load and maximize the capacity on the actual
-sensor devices.
-
-A centralized network query processing processes data in a network node
-specified as the network sink. The network sink usually is a network node with
-a database server with the collection of data delivered by all the sensor
-devices from the network. As a consequence of this many-to-one communication
-strategy \cite{sn-storage02}, the query process does not affect any of the
-sensor nodes in the network and, thus, maximizes the sensor device
-utilization. Moreover, the use of the data for historical purposes is easier
-to manage and provide access to the primary users of the data, following the
-process of data processing, compression, aggregation, etc
-\cite{sn-db-modeling02}.
-
-Although this approach seemed to be appropriate to the point of view of the
-sensor nodes and used as the tradeoff between achieving capacity and energy
-consumption, the centralized data collector may suffer from the unavoidable
-development of a point of traffic concentration \cite{sn-storage02}, also
-referred to as the phenomenon of congestion, or funneling effect
-\cite{sn-storage04}. For this reason, simple distributed system techniques
-such as the use of master-slave or data replication can be used to reduce the
-overhead.
+Although the approach of centralized query processing maximizes the sensor
+nodes capacity and minimize energy consumption, \cite{sn-storage02} explains
+that this trade off imposes the creation of the unavoidable development of a
+point of traffic concentration or the phenomenon of network congestion. For
+this reason, \cite{sn-storage04} describes that different techniques used in 
+distributed system such as the use of master-slave data replication or 
+database partitioning \cite{db-partitioning-relational} can be used to reduce
+or minigate the effects of this funneling effect.
 
@@
 use of basic data descriptors has been used in research groups which expertise
-is not database systems, while others use such technology and the way to
-represent data using more pragmatic approaches. \cite{sn-data-model-survey}
-have shown different data models and the advantages and disadvantages. The 
-following is a overview of these approaches used by the sensor network
-community.
+is not database systems, while others use such technology in pragmatic
+approaches. \cite{sn-programming-language} reports different applications using
+different types of languages to acces the data modeled in different data
+models. \cite{sn-provenance} describes the use of a \textbf{Tabular Data Model}
+\cite{tabular-model} using comma-separated values to model and organized the
+values related to the properties of observed parameters from a given sensor
+device. On the other hand, \cite{sn-db-tinydb} describes the use of the
+\textbf{Relational Data Model} \cite{relational-model} is used in the majority
+of the projects that involves data persistence due to its popularity in other
+different areas of Software Engineering and Distributed Systems. Finally,
+\cite{sn-xml-usage01, sn-xml-usage02} are examples of sensor networks
+application that makes use of XML \cite{xml} to represent the collected data
+on a \textbf{Structured Data Model}.
 
-\begin{itemize}
-  \item \textbf{Tabular Data Model}: comma-separated values are immensely
-  popular due to its simplicity. Although data is not index , it is an easy
-  approach for data exchange among researchers \cite{sn-provenance};
-  \item \textbf{Relational Data Model}: it is used in the majority of the
-  projects. Most of them due to the easy access to database systems, with
-  refactored versions that support the implementation of sensor networks and
-  its requirements \cite{sn-db-tinydb};
-  \item \textbf{Structured Data Model}: similar to the relational model,
-  structured data models are becoming popular due to the Internet. The use of
-  XML data \cite{xml} is not yet popular, but there are a few projects using
-  it \cite{sn-xml-usage01}\cite{sn-xml-usage02}, implementing its query by
-  using a middleware \cite{sn-xml-middleware, sn-xml-query-engines}.
-\end{itemize}
+The following sections describe the application of each of aforenamed data
+models in the context of sensor networks.
 
@@
 
-Simple use of Comma-Separated Values is an example of sensor network data. Data
-is structured by the top header, and the values are in each line of the file.
-The implementation of the SF-BEAMS \cite{sfbeams2006} uses this approach. In
-summary, in a tabular data model
+The most basic computer application to process data is through the use of
+spreadsheets, where data is organized into cells \cite{tabular-model}. When
+exported to the so-called comma-separated values (CSV) file format, the
+resulting file is comprised of lines of ascii characters, separated by a
+delimiter. The first line represents a header, with the identification of
+columns of specific data, where the following lines contains the values
+related to each top column. \cite{sn-provenance} reports that the tabular
+data model is commonly used to provide data persistence for sensor networks.
+As a matter of fact, the SF-BEAMS environmental sensor networks
+\cite{sfbeams2006} uses this approach to distribute the collected data over
+the Internet using the OPenDAP \cite{opendap} data format.
 
-\begin{itemize}
-  \item data is represented by columns in a flat file separated by a character
-  that separates the values. One example of such data in sensor networks is the
-  OPenDAP \cite{opendap} format, used by the SF-BEAMS project
-  \cite{sfbeams2006};
-  \item data is queried using a text editor file, without any indexing
-  associated with it.
-\end{itemize}
+This data model can be used in different context, as \cite{sn-provenance}
+describes that the versions of the same document can be easily managed and
+shared among research colleages using a version-control system such as
+Subversion \cite{subversion}.
 
@@
 
-Many different projects have used the relational model \cite{relational-model}
-as the representation of data for sensor networks. In general, the following
-properties apply to sensor networks:
+\cite{sn-db-newop} shows the use of the Relational Data Model
+\cite{relational-model} can be used in sensor networks to describe the
+collected data from sensor devices. As a matter of fact, the project described
+by the authors is based on the TinyDB database system \cite{sn-db-tinydb}, a
+database system specifically built for data nodes of sensor networks. It uses
+the same properties and constraints of regular database system, where the
+entities of the sensor network are required to be modeled and represented by
+tables.
 
-\begin{itemize}
-  \item data is structured in tables, and sensor properties are represented by
-  the columns of these tables. Data instances are represented by tuples or rows
-  of these tables. Finally, table relationships are implemented based on the
-  relational algebra theory, which is the mechanism used to describe the
-  relationship among different data types.
-  \item data is queried using the Structured Query Languange (SQL) and uses the
-  definitions of the tables and columns to extract data.
-\end{itemize}
-
-Before the creation of such a model, a database normalization process is
-necessary to describe the data in a convenient way. An example of such use of
-the relational model is the TinyDB database systems \cite{sn-db-tinydb}, which
-uses the relational model to represent the types of sensors. This is a popular
-choice because of the vast availability of research on relational database
-models.
-
-Although the implementation of a relational database seems to cover the needs
-of data representation, \cite{sn-data-model-survey} described the problems
-related to this approach. One such problem is that the relational model is not
-suited for modeling time series data and metadata related to identify, time
-and location. For problems related to it, came the implementation of different
-approaches that add semantics to data types and different query clauses to the
-Structured Query Language (SQL) to solve join operations \cite{sn-db-newop}.
+In order to query the data collected from sensor devices, the use of the 
+Structured Query Language (SQL) \cite{sql} is the main language used to extract
+the data managed by the system. As mentioned earlier, the default versions of
+this language have been modified to give support to the constraints of the
+location of the database such as the addition of new SQL operators for TinyDB
+\cite{sn-db-newop}.
 
@@
 
-The use of XML is used to represent data in sensor networks has become popular
-due the popularity of this data format. In general, the use of XML for
-persistence can be summarized as follows:
+XML is an increasing data format used popularized by the applications developed
+for the Internet. \cite{sn-xml-usage01} reports an efficient way to represent
+the collected data from sensors using XML, while the authors of
+\cite{sn-xml-usage02} describe a sensor network using XML for data model
+representation. In general, the collected data is structured into classes of
+relating elements and validated by an XML Schema \cite{xml-schema}.
 
-\begin{itemize}
-  \item data is structured in the collected data properties can be represented
-  through the XML \cite{xml} instances by using tags, having its model validated
-  by the use of XML Schema \cite{xml-schema}.
-  \item in order to extract data from XML-based databases, the use of XML XPath
-  \cite{xml-xpath} is used to query XML documents.
-\end{itemize}
+In order to extract data from XML-based databases, the XML XPath language
+\cite{xml-xpath} is often used to query XML data. However, with the inception
+of hybrid XML-relational database, \cite{sn-xml-usage03} reported the use XML
+as the data model and the use of SQL to query the stored data on the IBM's DB2
+database system.
 
-Some examples of the application of XML in sensor networks in
-\cite{sn-xml-usage02} some have investigated the effective use of XML in
-sensor networks to represent the collected data \cite{sn-xml-usage01}. An
-example of querying XML data in sensor networks described by
-\cite{sn-xml-query-engines}, while others use XML as a means of data exchange
-by implementing of middleware to query and extract query data
-\cite{sn-xml-middleware}. NetBEAMS is one of such sensor network that contains
-an XML-based communication middleware for data exchange among the sensor nodes
-\cite{netbeams2009}.
+Other examples of the use of the structured data model using XML is the through
+the use of middlewares, as the authors of \cite{sn-xml-query-engines} reports.
+Such approach is used for data exchange by the implementing of a middleware
+capable of query and extract the collected data from sensor devices
+\cite{sn-xml-middleware}. A practical example of the use of XML as a middleware
+to communicate between nodes is the communication mechanism used by NetBEAMS
+\cite{netbeams2009}. See section \ref{sec:dsp-serialization} for details.
 
@@
 
-Most of sensor networks implementations use the collected data from the sensor
-devices after they perceive the environment. As shown by
-\cite{sn-data-model-survey}, different data models contains their pros and
-cons to represent data. However, one crucial feature in sensor networks is the
-description of it collected data, as \cite{sn-provenance} discussed the
-properties and nature of sensor networks generated data. First, the device's
-sampling are usually in raw format or without obvious names. In order to
-provide search over the collected sensor data, the data needs first to be
-described and indexed, and this process can be related to Data Provenance
-\cite{db-provenance}. Needless to say, Data Provenance is the
-Metadata\footnote{Data about data} about the collected sensor data.
+As described in the previous section, the collected data from sensor networks
+can be represented by different data models, requiring the knowledge of the
+properties of the data. Not to mention that the low-level or raw data produced
+by a given sensor device might not include vital information regarding the
+origing and identity of the collected data. For this reason,
+\cite{sn-provenance} describes Data Provenance as a technique used to describe
+those essential properties of the collected data from sensors. One typical
+example of the application of Data Provenance is the so-called blackbox of an
+aircraft, which carries the collection of raw data collected from a group of
+sensors placed in different locations of an aircraft, as shown by
+\cite{sn-exemple-blackbox}. Last, but not least, the author describes the
+collected data as used in a process of data recovery and analysis in order to
+identify the probable reason of an aircraft accident, by using the time, and
+location of the aircraft, along with the independent variables from the
+sensors.
+
+In general, sensor devices produce data and either keep them onto its local
+storage or transmit them to data nodes as described earlier in this chapter. It
+is vital to understand that the majority of these devices outputs the data in
+its raw format, containing only a stream of values with any type of
+delimitation between each other. For this reason, the processes of parsing and
+description of each of the values located in the data streams regularly require
+the use of the device's specification documentation. Similarly, in order to
+have access to each of the properties of data collected by a given property
+related to time or space, \cite{sn-provenance} also mentions that 
+Metadata\footnote{Data about data} should be included together with the
+collected data.
 
@@
 over time to estimate the changing effects of the weather in a given region.
-For this reason, the collected data needs descriptions or annotations that can
-provide enough information to track changes, that is, the history of how and
-when a given data has changed. The following should be the essence of a
-provenance-aware collected data according to \cite{sn-provenance}.
+For this reason, the collected data from sensor devices need meaningful
+annotations to better describe the identity of the collected data, as 
+\cite{sn-provenance} reports. Interesting questions regarding the
+data can be asked such as \textbf{what was collected?}, \textbf{When was the
+data collected?} or \textbf{Where was the data collected?}. Similarly, while
+Data Provenance provides basic mechanisms to describe sensor networks data in
+general, others see the importance of the time space of sensor networks.
+\cite{sn-time-series} describes strategies used to cluster distributed sensor
+data as time series of data, as well as to decrease the size of data streams
+being transmitted in real-time sensor networks when these collected data does
+not need to be transferred \cite{sn-data-reduction}.
 
-\begin{itemize}
-  \item \textbf{what was collected?}: it can be merely one or more low-level
-  observations,  which is a set of values with or without a description;
-  \item \textbf{When was the data collected?}: is the time when the data was
-  collected,  as most sensor data must be used depending on its existence;
-  \item \textbf{Where was the data collected?}: it is typically the value of 
-  the of the location from where the data was collected. However, it is vital to
-  note that the use of this metadata depends on the applicability of the data.
-\end{itemize}
-
-While Data Provenance provides basic mechanisms to describe sensor networks
-data in general, others see the importance of the time space of sensor
-networks. \cite{sn-time-series} describes strategies to used to cluster
-distributed sensor data as time series of data, as well as to decrease the
-size of data streams being transmitted in real-time sensor networks when these
-collected data does not need to be transferred \cite{sn-data-reduction}.
-
-Finally, an indispensable application of collected sensor data is the so-called
-blackbox of an aircraft, which carries the collection of raw data collected
-from a group of sensors placed in different places of the aircraft, as
-described by \cite{sn-exemple-blackbox}. The author describes the data is used
-in a process of data recovery and analysis in order to identify the probable
-reason of an aircraft accident, by using the time, and location of the
-aircraft, along with the independent variables from the sensors.
+Finally, data provenance provides practical use of the collected data from
+different sensor devices and different sensor networks in the process called
+Information Fusison \cite{sn-info-fusion}. An example of the aggregation of
+different data from different devices located in different geographic sensor
+networks is \cite{sn-geo-metadata}, which is capable of process
+spatial-temporal queries of real-time or historical data. The following
+sections describe each of the metadata properties as defined by
+\cite{sn-provenance}, giving different examples of each of them.
 
@@
 
-Different sensor devices will produce different attribute sets regarding a
-particular observation. In general, these data sets does not have distinct
-names, which leads to difficulties when indices are required in order to make
-the data searchable. Specially for the end of data archival, as described
-earlier in this chapter, the collected data needs to receive and identity,
-which identifies the data by itself in a regular database
-\cite{relational-model}. Equally, \cite{sn-provenance} described that
-traditional sensor networks uses the collected data for analysis long after
-the attribute set has been collected, usually before it is categorized and
-indexed. Annotations and tags are also names given by \cite{sn-provenance} as
-means to describe collected sensor data.
+Different sensor devices produce different sets of attributes based on its
+capabilities and design. In general, these data sets do not have distinct
+names, which leads to difficulties when indices over the properties of the
+collected data are required to make them searchable. Specially for the purpose
+of data archival, the collected data needs to be identified, giving a meaning
+of independence on its own, the same idea behind the unique identification of
+entities on the relational data model \cite{relational-model}. Equally,
+\cite{sn-provenance} describes that traditional sensor networks uses the
+collected data for analysis long after the attributes have been collected,
+usually before it is categorized and indexed. Annotations and tags are also
+names given by \cite{sn-provenance} as means to describe collected sensor data.
 
-In a nutshell, data is described by the name of the observation, when that is
-not provided. For example,
-
-\begin{itemize}
-  \item \textbf{temperature} = 73.45: this is the value of the collected
-  observation, which in general, the attribute name ``temperature" is not
-  provided in the data stream;
-  \item \textbf{temperature-scale} = ``Fahrenheit": the metadata regarding the
-  temperature attribute, which describes the scale of the temperature;
-  \item \textbf{device-version} = 5: the metadata that describes the version of
-  the sensor device, since sensors can be replaced, and as a consequence, data
-  accuracy may be as well be changed.
-\end{itemize}
+In a nutshell, the collected data is described by the name of the observation
+property. For instance, \textbf{temperature} = 73.45 expresses the value of the
+observation of the temperature. In other words, the string ``temperature'' is
+the property describing the raw collected data 73.45. On the other hand,
+however, the example \textbf{temperature-scale} = ``Fahrenheit": describes
+the metadata regarding the temperature.
 
@@
 
-In order to use historical sensor data, the use of a point in time is the most
-prominent feature of data, as it is considered a pivotal metadata for different
-types of sensor networks such as an aircraft the blackbox
-\cite{sn-exemple-blackbox}. Other projects classify data in time-series
-\cite{sn-time-series-example}, while others provide approaches to decrease
-the creation of data based on clustering data before they are sent to the sink
-\cite{sn-time-series}.
+In order to use the collected data for the purpose of historical analysis, the
+use of a point in time is the most prominent feature of data, as it is
+considered a pivotal metadata for different types of sensor networks such as
+an aircraft the blackbox \cite{sn-exemple-blackbox} or any
+environmental sensor network \cite{sfbeams2006, sn-xml-middleware, sn-ex01}.
+In this way, Other projects classify data in time-series
+\cite{sn-time-series-example}, where data is related to all the
+observations a given sensor device performed over a period of time. The
+apparent implications is that all the produced data is necessary, but
+\cite{sn-time-series} argued that some sorts of sensor networks could decrease
+the creation of ``noisy data sets" based on the proper clustering of
+similar time-series data before before they are sent to the data sink.
 
-In different areas, temporal databases or time-series are the ones that take
-the real-world time into consideration. In general, data in a temporal database
-can be classified by its dimension, as surveyed by \cite{db-temporal}:
+Based on the guidelines defined by \cite{db-temporal}, the time properties of
+the collected data can be divided into two different dimensions: \textbf{valid
+time} and the \textbf{transaction time}. The former depicts the point in time
+when the observation took place with respect to the real-world space. An
+important aspect of this metadata is that since it captures the time when the
+observation occurred, its value cannot be changed during its existence, as well
+as placing restrictions on the update operations of this constant. Similarly,
+the latter type of is the point in time at which a given fact was transferred
+to the persistence layer.
 
-\begin{itemize}
-  \item \textbf{Valid time}: it is the point in time during which a fact
-  happened with respect to the real world. This value is only used when data is
-  captured, and cannot be changed because of its nature, as well as placing a
-  restriction over Update operations;
-  \item \textbf{Transaction time}: is the point in time at which a fact is
-  collected, transferred or processed in the database. This value is used
-  during Insertion operations, as well as in Update ones.
-\end{itemize}
+When both temporal dimensions are applied, the collected data is referred to as
+Bitemporal \cite{db-temporal}. Usually, the data type of this metadata is a
+regular timestamps, with any supported precision. To clarify, users may be
+interested not only when the temperature of 70 degrees happened, but also when
+it was collected to the data sink. Finally, \cite{sn-provenance} shows that
+the collected data for archival purpose is usually saved in self-described
+files described by file names, containing complex naming conventions using the 
+timestamp as a matter of data index.
 
-When both temporal dimensions are applied, the data is considered to be
-Bitemporal when \cite{db-temporal}. Usually, the data type of the valid and
-transaction time columns are straightforward timestamps, with any supported
-precision. For example, the following set of attribute and values shows a
-temperature of 70 degrees that was observed on ``November 15, 2009'', but
-collected only 3 days later.
-
-\begin{itemize}
-  \item \textbf{temperature} = 70;
-  \item \textbf{fact-time} = ``Nov 15, 2009 10:45 PM PST''
-  \item \textbf{transaction-time} = ``Nov 18, 2009 11:11 PM PST''
-\end{itemize}
-
-When it comes to Data Provenance in sensor networks, \cite{sn-provenance}
-shows that the collected data for archival purpose is usually saved in 
-self-described flat-files by file-names, containing complex naming conventions
-that contain the timestamp information about the data. For this reason, time
-is a central metadata needed in any data model used, since it helps
-identifying data uniquely for a given data set.
-
-\subsection{Where was the data collected from?}
+\subsection{From where was the data collected?}
 
 Depending on the application of the collected sensor data, the location in
-which the data was collected may be needed. For example, the geographic
-location of the collected data is critical to identify the specific
-GPS\footnote{Global Positioning System} coordinates of the sensor device
-\cite{sn-ex02}. Similarly, the position of any  event captured in a black-box
-of an aircraft carries is also used as metadata for the event. It will either
-depend on the sensor device provide such sampling attribute, or entered
-manually on the database.
+which the data was collected may be needed. For example, \cite{sn-geo-metadata}
+shows the geographic location of the collected data is critical to identify
+the locations of the events of the sensor device \cite{sn-ex02}. Similarly,
+the position of any event captured in a black-box of an aircraft carries is
+also used as metadata for the event. For this reason, engineers can open a
+real-world explanation regarding my family.
 
-A correlation between time and location is presented by \cite{sn-time-series}
-as a noteworthy feature to reduce the traffic generated in a sensor network by
-clustering similar data when the sensor devices are placed in concentrated 
-environment.
+All in all, Data Provenance enriches the value of the collected sensor,
+describing its identity creation, taking into account the specific location
+where the data was deployed. Such application ca be seen by having the problems
+and solutions Information Fusion \cite{sn-info-fusion}.

### 2009-12-04 docs/marcello-dissertation/thesis/chapter2-state-of-the-art.tex
Fixing typos on chapter 2, since the writing was very incorrect :)
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@669 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 considering that some types of sensor devices are very likely to regular
-hardware failures, \cite{sn-intro2} described sensor network examples with 
+hardware failures, \cite{sn-intro02} described sensor network examples with 
 redundant nodes that provide better data coverage and levels of coverage.
@@
 the locations of the events of the sensor device \cite{sn-ex02}. Similarly,
-the position of any event captured in a black-box of an aircraft carries is
-also used as metadata for the event. For this reason, engineers can open a
-real-world explanation regarding my family.
+the position of any event captured by a black-box of an aircraft carries 
+metadata for the event. Examples of this are the metadata event =
+``high-pressure" of 1304, at the altitude of 1,500 feet above the ocean on GPS
+coordinates of latitude 13.445 and longitude -23,003. This example clearly
+gives necessary identification of where the event of high-pressure occurred.
 
-All in all, Data Provenance enriches the value of the collected sensor,
-describing its identity creation, taking into account the specific location
-where the data was deployed. Such application ca be seen by having the problems
-and solutions Information Fusion \cite{sn-info-fusion}.
+All in all, Data Provenance enriches the value of the collected data from
+sensor devices by describing its identity creation, time of creation and
+specific location where the observation occurred. It is important to note,
+however, that those type of data do not necessarily need to be on the data
+model used, but the specifications of the use of the data determines which
+ones are needed. Finally, collected data that are provenance-aware has a direct
+use on technologies called Information Fusion \cite{sn-info-fusion}, which aims
+at using data from different sensor devices.

### 2009-12-04 docs/marcello-dissertation/thesis/chapter3-taxonomic-classification-of-persistence-storage-for-sensor-networks.tex
Revisions on chapter 3, after reading the entire chapter and aligning it with the previous ones.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@670 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 
-\chapter{Taxonomic Classification of Data Persistence in Sensor Networks}
+\chapter{Proposal of a Data Persistence in Sensor Networks Taxonomy}
 \label{chap:taxonomies}
 
-Different mechanisms are used for persisting data collected from sensor
-networks, as described by the section regarding the state of the art in sensor
-networks and different approaches for persistence. This section analyzes these
-approaches and presents the different approaches through
-taxonomy\footnote{Taxonomy is the practice and science of classification, by
-using taxonomic units known as taxa (singular taxon)}, which were used
-throughout this work to evaluate different existing database systems
-implementations, taking into account different requirements and specifications
-of different sensor networks.
+As revailed in the previous chapter, data persistence for sensor
+networks require an understanding of not only the properties of the
+infrastructure of a given sensor network, but also of the properties and nature
+of the collected data themselves. To put things differently, in order to
+identify these concepts more clearly, each section of this chapter proposes an
+individual taxonomy based on the aspects described in the literature
+review. Then, it briefly discusses the relationships among them.
+
+A taxonomy is defined as the practice and science of classification and uses
+the taxonomic units of ``taxa'', the plural form of ``taxon''. Similarly, they
+are represented by a hierarchy diagram that relates the main taxonomy to the
+different taxa and their own related ones, if any.
 
@@
 
-No matter the resource management used, the first taxonomy presented relates to
-the purpose of the use of the collected sensor data. That is, when and how the
-collected data is going to be used as described in section
-\ref{sec:sn-data-purpose}.
+As delineated in section \ref{sec:sn-data-purpose}, the purpose of the
+collected data from sensor devices determines how the collected data is
+retrieved and used in the network. In such a fashion, this taxonomy relates to
+how the collected data is used, and should be the primary question to be
+answered during the analysis of a sensor network. The taxa related to this
+taxonomy, depicted in Figure \ref{fig:taxonomy-data-purpose}, is described as
+follows:
 
@@
 \begin{itemize}
-  \item \textbf{Real-time Data Stream}: data is used as real-time data stream
-  during data query \cite{sn-intro01};
-  \item \textbf{Archival Data}: data is used for historical purposes such as
-  data analysis or data fusion \cite{sn-intro01, sn-intro02}.
+  \item \textbf{Real-time Data Stream}: the collected data is accessed
+  on-the-fly from the sensor device as a real-time data stream, and may have a
+  short life cycle as it temporarily resides in memory;
+  \item \textbf{Data Archival}: the collected data is used for historical
+  purposes and is related to the historical data analysis \cite{sn-intro01,
+  sn-intro02} or Information Fusion \ref{sn-info-fusion}. 
 \end{itemize}
@@
 
-Directly related to the purpose of data, the procedure for data storage
-described in section \ref{sec:sn-infrastructure} depends on the purpose of the
-use of the collected data. Figure \ref{fig:taxonomy-data-location}, shows the
-related taxon and can be described as follows:
+The location where the collected data is stored plays an important role
+in the different aspects of the access to the collected data from sensor
+devices, as shown in section \ref{sec:sn-storage-locations}. In this way,
+Figure \ref{fig:taxonomy-data-location} depicts this taxonomy, which can be
+described as follows:
 
@@
 \begin{itemize}
-  \item \textbf{Local Storage}: The data can be persisted in a temporarily
-  memory address, or a secondary memory device in a local storage device. It is
-  directly related to the real-time data stream taxon, since it occurs at the
-  place of the sensor device;
-  \item \textbf{External Storage}: The data is usually located in an external
-  storage devicewith a data management system such as a database system;
-  \item \textbf{Data-Centric Storage}:the collected data can organized in a
-  categorizedmanner as described by \cite{sn-storage03}. This taxon regarding
-  the system allows the data to bepartitioned by a particular property.
-  Nowadays, this strategy is often related to database partition
-  \cite{db-partition}, and database sharding approaches \cite{db-shard-intro,
-  db-shard-discussion}.
+  \item \textbf{Local Storage}: characterizes the placed of the collected data
+  temporarily located in-memory, or in a secondary storage device on a node
+  located in-network;
+  \item \textbf{External Storage}: The collected data is usually located in an
+  external storage device with a data management system such as a relational
+  database;
+  \item \textbf{Data-Centric Storage}: when the collected data occupies
+  different external storages, being organized by categories based on the
+  collected data keys and values. In other words, what characterizes this taxon
+  is the presence of a data partitioning strategy to store the collected data.
 \end{itemize}
@@
 One of the fundamental challenges in the development of a persistence storage
-in sensor network is related to the data model chosen to keep the collected
-data. The introduction of new sensor devices may cause refactoring of the
-existing model or not, depending on how the data model accommodates the data.
-Different data models are reported to be used in the literature, as described
-in section \ref{sec:data-models}. For this reason, taxonomy related to the
-data model depicts two different classes of data models, as shown in Figure
-\ref{fig:taxonomy-data-model}.
+in sensor networks is related to the data model chosen to represent the
+collected data. Different data models were reported in the review of related
+literature in section \ref{sec:data-models}. As a consequence, this taxonomy
+relates to the data model used to design the collected data, as drawn in Figure
+\ref{fig:taxonomy-data-model}, and described as follows.
 
@@
   \item \textbf{Schema-Dependent Models}: this data model requires the
-  definitionof a master schema that describes the data through a rigorous data
-  modeling process prior to the use of the data. In addition, an addition to a
-  newsensor device, and hence a new data entity, requires a re-evaluation ofthe
-  existing data schema. For example, relational model is governed by
-  therelational algebra model \cite{relational-model}, while the structured
-  models such as XML \cite{xml} needs adefinition of an XML Schema
-  \cite{xml-schema};
+  definition of a master schema that describes the data through a rigorous of 
+  data modeling process prior to the insertion of the data. Examples of such
+  model is the traditional Relational Data Model\cite{relational-model}, as
+ well as the Structured Data Models such as the XML \cite{xml};
   \item \textbf{Schema-less Models}: contrary to the former taxon, schema-less
-  data models do not require the definition of a data schema or a master table
-  definition, along with its table relationships, if necessary. Instances of
-  such data model is the Key-Value Pair \cite{db-kvp} and its variation
-  document-oriented models \cite{db-is-rdbs-dommed}, whose use in wireless
-  sensor networks has not been reported.
+  data models do not require the definition of a data schema or table
+  definition. Instances of such data model reviewed in the literature review
+  is the use of the Tabular Data model.
 \end{itemize}
@@
 
-No matter the data model used, the support for Data Provenance as shown in
-section \ref{sec:sn-provenance} is as valuable as the data being collected.
-For this reason, the data model taxonomy must have support for the data types
-that support the use of provenance-related attributes in the collected sensor
-data. In this way, Figure \ref{fig:taxonomy-data-provenance} describes the
-data types for the metadata support.
+No matter which data model used to represent the collected data, Data
+Provenance provides a valuable guideline on how to describe the collected data
+from sensors devices. As reported in section \ref{sec:sn-provenance}, the
+description of the properties should take into account different properties of
+the nature of the data as shown in Figure \ref{fig:taxonomy-data-provenance}.
+Those taxa are summarized as follows:
 
@@
   \item \textbf{What: Data Identity}: the data type that uniquely identifies
-  the collected sensor data;
-  \item \textbf{When: Time Dimension}: the data type that identifies the
-  collected data;
-  \item \textbf{Where: Data Location}: an optional data type that identifies
-  from where the data originated.
+  the collected data from sensors. First and foremost, the properties of the
+  sensor devices must be described according to the meaningful naming
+  convention for the property. For instance, ``Water Temperature'' better
+  describes the intent of the property as ``temp''. Second, the identification
+  of the collected data related to the sensor type should be given, that is,
+  which sensor device created the collected data, in case the data origin is
+  relevant;
+  \item \textbf{When: Time Dimension}: time is an important attribute of the
+  data because it identifies the age of the collected data. Depending on the 
+  use of the collected data, this is often taken into account. Two different
+  types of time dimensions were described: \textbf{fact} and
+  \textbf{transaction} times. The former is used to identify when the data
+  was collected, whereas the latter when the data was transmitted to the
+  data sink where the query processing acts;
+  \item \textbf{Where: Data Location}: the point where the collected data was
+  observed by the sensor device. Usually the latitude and longitude as the GPS
+  coordinates are provided by devices which are capable of sensing the
+  location. A descriptive metadata about the location could also be given such
+  as room-name = ``deep see''.
 \end{itemize}
@@
 The query processing mechanism in sensor networks is directly related to how
-and where the collected data is used and stored, as shown in the previous
-section. As a consequence, the following taxonomy taxon is related to the
-different types of query processing mechanisms described in section
-\ref{sec:query-process}, and can be seen in Figure
-\ref{fig:taxonomy-query-mechanism}.
+and where the collected data is used and stored, as shown in section
+\ref{sec:query-process}. As a consequence, the following taxa are related to the
+different types of query processing mechanisms and summarized in Figure
+\ref{fig:taxonomy-query-mechanism}:
 
@@
 \begin{itemize}
-  \item \textbf{In-Network Query Processing}: the sensor node receives a
-  request to theproperties of the sensor, and returns the current values
-  related to them \cite{sn-intro01};
-  \item \textbf{Centralized Query Processing}: as developed in single server
-  relational databases, the data is collected to a centralized data sink in
-  order to be used.
+  \item \textbf{In-Network Query Processing}: when the data node is the
+  sensor device itself, and it uses the local storage mechanism, then the
+  sensor node is able to reply to request related to the observations of the 
+  environment defined by the sensor device. In this way, the query takes 
+  place in the network itself, and usually returns values related to the current
+  state of the environment;
+  \item \textbf{Centralized Query Processing}: as developed by a single 
+  relational database server, the data queried from a centralized data
+  network sink in order to be reused. In general, this approach is used when
+  the purpose of data is archival data.
 \end{itemize}
@@
 
-Different approaches of data storage can also be related to the performance in
-the database system used, and how its architecture provides the data storage
-solution. For example, one crucial problem related to the centralized network
-sink in sensor networks is the presence of an ineluctable develoment of a point
-of traffic concentration at this node \cite{sn-storage02}. In this way, in
-order to reduce the problems associated with the so-called ``Funneling Effect",
-Figure \ref{fig:taxonomy-database-architecture} summarizes the different
-capabilities of different types of database systems.
+Different approaches of data systems were described in the literature review in
+the last chapter in different sections. This taxonomy is directly related to
+the use of external storage devices on a centralized network node so that it
+often uses a single database server. Paradoxically, the use of a data-centric
+storage determines the use of distributed database nodes around the network. 
+In this way, Figure \ref{fig:taxonomy-database-architecture} describes the
+taxonomy of the database system organization as follows:
 
@@
   \item \textbf{Distributed System}: a database system that can be composed by
-  more thanone node. Usually, the use of the pattern of Master-Slave
-  mechanisms can beused by implementing ``Replication". In addition to it,
-  ``Database Sharding" is used to portion the data related to a given shard key.
+  more than one node. Usually, the use of the pattern of Master-Slave
+  mechanism or data ``Replication" is used. Another mechanism is the use of
+  database paritioning, if the technology supports it.
 \end{itemize}
+
+\section{Relationships among the Taxonomies}
+
+The different taxonomies defined in the previous sections of this chapter are
+important to the understanding of a data persistence layer for sensor
+networks. As they are closely related to each other, it is important to
+briefly highlight these relationships in order to better analyze a problem
+related to data persistence.
+
+The taxonomy related to the purpose of the sensor data can be directly related
+to the taxonomy of the location of the sensor data. While the taxon of the 
+real-time data stream relates to the local data storage, the taxon related to
+data archival correlates the storage of the collected data into an external
+storage. In addition to that, the converse is also true.
+
+The taxonomy related to the data model can be related to any taxon from the
+taxonomy of the location of the sensor data, as well as the taxons related to
+Data Provenance. In the former, the data of a relational model is stored using
+SQL messages.
+
+In the light of these taxonomies, the classification of the data persistence
+properties of a sensor networks can be studied and reviewed. As described in
+the introduction of this dissertation, one the main motivations for this work
+was to provide means of data persistence of NetBEAMS, a case study which does
+not contain a data layer. Therefore, the following chapter seeks to describe
+that case study in order to make a section of a given technology for evaluation.

### 2009-12-05 docs/marcello-dissertation/thesis/chapter10-appendix.tex
Print revision of chapter 4, with corrections on chapter 10.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@671 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 will receive the DSP Message, and if its delivery is done by means of a gateway
-component. In this way, the DSP Matcher can be seen as a function that takes a
-DSP Message as an input and returns a list of DSP consumers, and the
-matching configuration artifact provided by the administrator. See listing
+component. In this way, the DSP Matcher can be seen as a function, as depicted
+in Table \ref{tab:matching-rule}, that takes a DSP Message as an input and
+returns a list of DSP consumers, and the matching configuration artifact provided by the administrator. See listing
 \ref{file:dsp-matcher-config.xml}) for details.
@@
     \end{center}
-    \label{tab:ysi-data-stream}
+    \label{tab:matching-rule}
 \end{table}
@@
 \subsection{Sending DSP Messages through the Wire}
+\label{sec:dsp-serialization}
 

### 2009-12-05 docs/marcello-dissertation/thesis/chapter4-netbeams-in-depth-requirements-analysis-and-specification.tex
Print revision of chapter 4, with corrections on chapter 10.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@671 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 
-This chapter presents the primary motivation and scope of this paper. The
-infrastructure of SF-BEAMS is presented as a real-world environmental sensor
-network, deployed by the department of Biology at the San Francisco State
-University. It serves as the foundation of the execution of NetBEAMS, a joint
-venture between the department of Biology and the Computer Science. As an
-alternative approach for data collection, NetBEAMS is used to automate the data
-management and collection process of SF-BEAMS. However, it offered different
-open problems, including Data Persistence. Therefore, the realm of the solution
-presented in the following sections and chapters are based in the case study
-presented in this current chapter.
+As discussed in chapter 2, different properties of a sensor network and the
+nature of the collected data must be taken into account in order to provide a
+data persistence layer for a given sensor netwokr. Similarly, in order to
+better assist one analysis of such functionality, chapter 3 proposed a set of
+data persistence taxonomies related to different properties of the collected
+data life cycle. In this way, the goal of this chapter is to describe the
+fundamental properties of the case study used for this work, that is, NetBEAMS,
+in order to ``set the stage'' for the selection of a database technology for it.
+In this way, given the fact that NetBEAMS provides an automated infrastructure
+solution for SF-BEAMS, this last component will be covered first. Then, this
+work proposes a categorization of NetBEAMS to the face of the taxonomies, and
+finally discusses the discusses the requirements for a data persistence.
 
@@
 
-The San Francisco Bay Environmental Assessment and Monitoring Station, or
-SFBEAMS \cite{sfbeams2006}, is a marine sensor network operated by the Romberg Tiburon
-Center (RTC), whose primary focus is the study of complex marine and estuarine
+As described by \cite{netbeams2009}, NetBEAMS is a joint venture between the
+department of Biology and Computer Science at San Francisco State University,
+whose goal was to automate the operational execution of SF-BEAMS. On the other
+hand, the San Francisco Bay Environmental Assessment and Monitoring Station, or
+SFBEAMS \cite{sfbeams2006}, is an instance of an environmental sensor
+network whose primary focus is the study of complex marine and estuarine
 environments using the SF-BEAMS sensors. The network is deployed offshore of
@@
 California, being operated by the department of Biology from San Francisco
-State University.
+State University. In this way, NetBEAMS is an experimental component that
+utilizes the SF-BEAMS infrastructure to operate. This section details SF-BEAMS
+in general, and presents the requirements for data persistence.
 
@@
 
-In general, the SF-BEAMS network infrastructure contains varying wired and
-wireless devices standing on pylons docked at the pier. Each of them is
-responsible for observing different conditions of the area, having its own
+In general, the SF-BEAMS network infrastructure contains varying a number of
+wired and wireless devices standing on pylons docked at the pier. Each of them
+is responsible for observing different conditions of the area, having its own
 mechanisms for internal storage for the collected data. In this way, data can
 be directly transferred to the labs via Ethernet cables or collected manually
-by a laptop computer.
+with a laptop computer.
 
@@
 its observed data based on properties of measurements defined by its
-manufacturer. SF-BEAMS can be seen a single-hop star sensor network containing
-different instances of sensors. One example of such sensors is the ``YSI 6600
-ESD V2", as seen in Figure \ref{fig:ysi-device}. Therefore, it can be seen as
-a powerful water quality monitoring device that produces around 52 bytes (13x4
-Bytes) on a single real-time data stream as shown on table
-\ref{tab:ysi-data-stream}.
+manufacturer. One example of such sensors is the ``YSI 6600 ESD V2", as seen in
+Figure \ref{fig:ysi-device}. It is a powerful water quality monitoring device,
+capable of producing around 52 bytes on a single real-time data stream reading,
+as shown on Table \ref{tab:ysi-data-stream}.
 
@@
 
-\begin{table}
+\begin{table}[!b]
     \caption{YSI Data Stream}
@@
 \cite{rs232}. Then, the RTC labs receives the collected data transferred from
-the sensor devices to be used for archival. The data is described, indexed and
-distributed to its main users. According to one of its staff members, the
-infrastructure of the SF-BEAMS in May 2009 was comprised of the following:
+the sensor devices to be used for archival purposes. According to one of its
+staff members, the infrastructure of the SF-BEAMS in was comprised of
+the following components (as of May 2009):
 
 \begin{itemize}
-  \item Among other devices, 5 YSI sondes in operation at the RTC pier;
-  \item Each YSI sonde produces 52 Bytes for each its observations;
-  \item The sampling frequency rate could be configured in ranges of 1, 6 or
+  \item 5 YSI Sonde devices, among others, in operation at the RTC pier;
+  \item The sampling frequency rate is configured in ranges of either 1, 6 or
   15 minutes, depending on the specification;
-  \item Data for final users is published with Provenance Data regarding time,
-  without description of the origin of data.
+  \item The processed data, used for distribution over the Internet, is
+   contains information regarding the time of the data collection.
 \end{itemize}
 
-Considering the infrastructure previously described, the size of the data
-produced by the YSI sonde sensors at the RTC pier can be estimated by its
-measurement frequency as follows:
+Considering the infrastructure aforementioned, the data load in the server-side
+of the RTC's laboratories produced by the YSI sonde sensors devices can be
+estimated by analyzing the data distribution over a period of one year, as
+shown in Table \ref{tab:ysi-data-distribution}:
 
-\begin{table}
+\begin{table}[!b]
     \label{tab:ysi-data-distribution}
@@
 
-As it is shown, the data produced by this type of sensor runs in the order of
-at most hundreds of Megabytes a year for a maximun of \textbf{483,840}
-samples, what can be considered a exceptionally sparse data storage
-requirement as compared to other networks. Upon collecting data from sensors,
-the RTC staff use automation scripts written in Matlab to process, index and
+By analyzing the maximum amount of produced data by the YSI sonde device is
+around hundreds of Megabytes a year, as \textbf{483,840} samples are generated
+by one single device. Upon collecting data from sensors, the RTC staff uses
+automation scripts written in Matlab \cite{matlab} to process, index and
 distribute the raw data in different formats. One of these formats is the
-OPeNDAP \cite{opendap}, a format widely used at research institutions that
-promote easier data exchange among them. The access to the archived data is
-done by using the OPeNDAP server is located at
-http://sfbeams.sfsu.edu:8080/opendap, as a request and the corresponding
-output for one of the YSI devices can be seen in Listing \ref{file:rtc-ysi-opendap}.
+OPeNDAP \cite{opendap}, which is widely used at research institutions that
+promote easier data exchange among them. The SF-BEAMS website provides access
+to the collected data through the Internet at
+http://sfbeams.sfsu.edu:8080/opendap. An example of an HTTP Request from
+accessing the data for a specific YSI is shown in Listing
+\ref{file:rtc-ysi-opendap}.
 
@@
 
-In order to distinct each of the properties of the SF-BEAMS sensor networks,
-consider the following list as a classification of the data persistence
-mechanisms from the sensor network used by the project:
+This section classifies the sensor network deployed by SF-BEAMS according to
+the taxonomies proposed in the previous chapter. This classification is
+necessary to better understand the requirements of a data persistence for
+NetBEAMS, which indirectly uses the infrastructure of SF-BEAMS.
 
@@
   \item \textbf{The Purpose of Sensor Data}: the data from SF-BEAMS is solely
-  used for Data Archival purposes, since they are stored in the RTC's lab as
-  part of the collaboration research project that monitors each part of the
-  California coast. The data sets are distributed by a specific date in an
-  annual calendar;
+  used for the purpose of Data Archival, since they are stored in the RTC's
+  lab;
   \item \textbf{The Location of the Sensor Data}: it is clear that the purpose
@@
   data at the network sink at the RTC's labs;
-  \item \textbf{Data Model}: no data model is used by the RTC's staff to model
-  the data. A straightforward conversion of the data collected from the sensors
-  directly to spreadsheets, using comma-delimited values, is used. That is, the
-  data mode used \textbf{Schema-less};
-  \item \textbf{Data Provenance}: data is stored in files with conventions that
-  take the time when the data was collected. The contents of the files contain
-  the \textbf{Data Identity} and \textbf{Time Dimensions}, as shown in the
-  samples collected and shown in Listing \ref{file:rtc-ysi-opendap};
-  \item \textbf{Query Processing Mechanism}: data is queried in a
-  \textbf{Centralized} fashion, due to the nature of its purpose and locality;
-  \item \textbf{Database System Organization}: there are no databases in use.
-  The SF-BEAMS staff use the OPeNDAP Hyrax server as a middleware used to read
-  and interpret the collected data. 
+  \item \textbf{Data Model}: since the format uses by the RTC staff to share
+  data among researcher is in the OPEnDAP format, it is clear that the data
+  model used is the tabular data model, since it uses the comma-delimited
+  files. In this way, the \textbf{Schema-less} is used;
+  \item \textbf{Data Provenance}: once the collected data reaches the RTC
+  lab, a quality assurance process is executed, and the data is transformed
+  into the format used by OPEnDAP. This conversion carries data regarding
+  the time of the data collection, and therefore, the use of one of the
+  \textbf{Time Dimensions} such as the valid time. Since the files are stored
+  with specific file name structure, using timestamps to describe the data, the
+  use of transaction time may be also considered. Finally, the contents of the
+  files contain the \textbf{Data Identity} as shown in the samples of
+  the collected in Listing \ref{file:rtc-ysi-opendap};
+  \item \textbf{Query Processing Mechanism}: similar to the data stored in a
+  centralized way, the query processing is also \textbf{Centralized}, as the
+  data is stored on file system;
+  \item \textbf{Database System Organization}: there are no databases in use,
+  but the use the OPeNDAP Hyrax server as a middleware that servers the
+  collected data using the Internet.
 \end{itemize}
@@
 Although the execution of the RTC's sensor network can be operated as it
-described in the previous sections, \cite{netbeams2009} described some
-operational challenges faced by the RTC staff during regular activities during
-the data collection process. It was clear to the research group that SF-BEAMS
+described in the previous sections, \cite{netbeams2009} described some of the
+operational challenges faced by the RTC staff during regular activities of
+data collection process. It was clear to the research group that SF-BEAMS
 could have not only its data gathering process automated, but also its data
@@
 open-source \cite{open-source} software, the research group developed a second
-version of NetBEAMS, a component-based approach suggested improving
-operational activities from SF-BEAMS using systems automation. In brief, one
-of the open problems of NetBEAMS was regarding Data Persistence, the primary
-motivation of this dissertation.
+version of NetBEAMS, a component-based approach suggested improving operational
+activities from SF-BEAMS using systems automation. In brief, one of the open
+problems of NetBEAMS was regarding Data Persistence, the primary motivation of
+this dissertation.
 
 The Networked Bay Environmental Assessment and Monitoring System, or Net-BEAMS,
-is a work-in-progress joint project proposed by the department of Computer
-Science and the RTC at San Francisco State University. The main reason for the
-inception of NetBEAMS is the potential to reduce operational costs and
-facilitate the data gathering process. In the current state of SF-BEAMS,
-Biologists need to obtain data manually from the sensor devices in order to
-take them to the network sink for further processing. For this reason, the Data
-Sensor Platform (DSP) \cite{netbeams2009} was proposed as the system
+offers the Data Sensor Platform (DSP) \cite{netbeams2009} as the system
 architecture that can address the operation of SF-BEAMS. In this way, in-depth
 documentation describing the DSP Platform, including its architecture and data
-gathering process, is provided in the appendix section \ref{sec:dsp-details}.
+gathering process, is provided startng in the appendix section
+\ref{sec:dsp-details}.
 
 The scope of this work then is defined as to provide a persistence layer for
-NetBEAMS, making a selection of a persistence layer that is a exception to the
+NetBEAMS, making a selection of a persistence layer that is an exception to the
 taxonomical classification of SF-BEAMS. Furthermore, the technology to be
@@
 NetBEAMS primary users such as Computer Science researchers, as well as end
-users such as Biologists from the RTC laboratories.
+users such as Biologists from the RTC laboratories. The requirements for the
+data persistece specifically for NetBEAMS is described in the following section.
 
@@
   more human-readable format;
-  \item Scalable Systema low degree of maintenance, in a way that data
-  gathering service should be able to manage the system with less service interruption;
-  \item Data must be searchable in near-real-time with outstanding performance,
-  as compared to the time when data was collected to the time when they are available;
-  \item Easy data integration and export capabilities must be also taken into
-  account;
+  \item The system must be scalable with a low degree of maintenance, in a
+  way that the data gathering process does not impose less service interruption;
+  \item Data must be searchable in near-real-time with outstanding performance;
+  \item The solution must also give support to export capabilities, which can
+  be possibly used to match the RCT's requirements of the OPeNDAP formatt;
   \item The database system must be free of charge, following the
-  implementation specifications of NetBEAMS.
+  implementation specifications of NetBEAMS for Open-source software.
 \end{itemize}
+
+After presenting NetBEAMS and its infrastructure, the the next step is the
+actual analysis of existing database systems that fits to the requirements
+and characteristics of NetBEAMS. Meanwhile, more details regarding the
+architecture of NetBEAMS, as well as development guidelines, can be reached in
+the appendix of chapter 10.

### 2009-12-07 docs/marcello-dissertation/thesis/chapter10-appendix.tex
Most of the revisions and updates... Changes to chapters 3, 4 after Kleber's chat online (adding missing taxonomy).
Revisions of print chapters until chapter 5.

Chapter 7 is reviewed until before discussion. 

Next revision should include all the necessary reviews for the
release draft.

git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@676 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 \section{mongoDB Deployment}
+\label{sec:mongodb-deployment}
 
@@
 
-\subsection{Starting the Database}
-
-\subsection{Stopping the Database}
-
-\subsection{Execution Logs}
+\subsection{MongoDB Database Administration}
+
+- starting the db
+- Stopping the Database
 
 \section{Experiment Execution}
+\label{sec:experiment-execution}
 
@@
 \subsection{Data Access Using Database Shell}
+\label{sec:mongodb-user-experience}
 
@@
 \begin{lstlisting}
+http://192.168.1.2:28017/netbeams/SondeDataContainer/?limit=-3&filter_observation.Conductivity=104.5
+
 HTTP/1.0 200 OK
@@
   "rows": [
-    { "_id" : "156f4007e4c3b74a36ed3100", "sensor_ip_address" : "192.168.0.117", "message_id" : "08b02c08-9290-4517-9a28-c6ee7e16509a",
-"transaction_time" : 1253557219486, "fact_time" : 1253557217000, "data" : { "temperature" : "31.44", "sp_condition" : "99.8", "condition" : "53.5",
-"resistence" : "1157.08", "salinity" : "0.0", "pressure" : "1.066", "depth" : "0.161", "ph" : "1.08", "pH_mv" : "-82.0", "odo_sat" : "40.3",
-"odo_condition" : "56.85", "turbidity" : "0.2", "battery" : "8.2" } } ,
-    { "_id" : "156f4007e5c3b74a37ed3100", "sensor_ip_address" : "192.168.0.117", "message_id" : "08b02c08-9290-4517-9a28-c6ee7e16509a",
-"transaction_time" : 1253557219486, "fact_time" : 1253557217000, "data" : { "temperature" : "37.83", "sp_condition" : "176.3", "condition" : "2.6",
-"resistence" : "1324.97", "salinity" : "0.01", "pressure" : "1.36", "depth" : "1.564", "ph" : "0.12", "pH_mv" : "-23.5", "odo_sat" : "104.5",
-"odo_condition" : "19.44", "turbidity" : "0.1", "battery" : "5.0" } } ,
-    { "_id" : "156f4007e5c3b74a38ed3100", "sensor_ip_address" : "192.168.0.117", "message_id" : "08b02c08-9290-4517-9a28-c6ee7e16509a",
-"transaction_time" : 1253557219486, "fact_time" : 1253557217000, "data" : { "temperature" : "74.3", "sp_condition" : "84.0", "condition" : "104.7",
-"resistence" : "4089.13", "salinity" : "0.01", "pressure" : "1.222", "depth" : "2.788", "ph" : "6.56", "pH_mv" : "-78.1", "odo_sat" : "40.0",
-"odo_condition" : "6.02", "turbidity" : "0.3", "battery" : "3.2" } },
+    { "_id" : { "$oid" : "616f400784ba194bc2dee200" }, "message_id" : "c76a9898-1eee-4463-bf5d-d40a7a8af6e3", "sensor" : { "ip_address" : "192.168.0.60", "location" : { "latitude" : 37.89155, "longitude" : -122.4464 } }, "time" : { "valid" : { "$date" : 1261229067000 }, "transaction" : { "$date" : 1259977271704 } }, "observation" : { "WaterTemperature" : 51.29, "SpecificConductivity" : 110.7, "Conductivity" : 104.5, "Resistivity" : 1710.07, "Salinity" : 0.01, "Pressure" : 1.094, "Depth" : 1.808, "pH" : 3.04, "pHmV" : -46.4, "Turbidity" : 0.2, "ODOSaturation" : 59, "ODO" : 10.66, "Battery" : 7.1 } } ,
+    { "_id" : { "$oid" : "616f400789ba194bdbe2e200" }, "message_id" : "2eb3f5ca-7a30-4b5d-8d16-c3e80a7b1b7a", "sensor" : { "ip_address" : "192.168.0.24", "location" : { "latitude" : 37.89155, "longitude" : -122.4464 } }, "time" : { "valid" : { "$date" : 1261740924000 }, "transaction" : { "$date" : 1259977271704 } }, "observation" : { "WaterTemperature" : 10.98, "SpecificConductivity" : 100.9, "Conductivity" : 104.5, "Resistivity" : 5448.21, "Salinity" : 0.08, "Pressure" : 1.152, "Depth" : 1.289, "pH" : 3.09, "pHmV" : -0.4, "Turbidity" : 0.1, "ODOSaturation" : 72.2, "ODO" : 59.65, "Battery" : 6.3 } } ,
+    { "_id" : { "$oid" : "616f400789ba194b13e5e200" }, "message_id" : "a7804d0a-44a8-4bd9-a9c7-51db84e91f08", "sensor" : { "ip_address" : "192.168.0.236", "location" : { "latitude" : 37.89155, "longitude" : -122.4464 } }, "time" : { "valid" : { "$date" : 1260969259000 }, "transaction" : { "$date" : 1259977271704 } }, "observation" : { "WaterTemperature" : 7.97, "SpecificConductivity" : 119.1, "Conductivity" : 104.5, "Resistivity" : 1971.59, "Salinity" : 0.05, "Pressure" : 1.516, "Depth" : 2.187, "pH" : 3.11, "pHmV" : -77.59999999999999, "Turbidity" : 0.2, "ODOSaturation" : 82.90000000000001, "ODO" : 41.99, "Battery" : 0.1 } }
   "total_rows" : 3 ,
@@
 An example to export the data in CSV format can be seen in listing
-\ref{cmd:mongoexport}, and an example containing 1 million objects can be
+\ref{file:mongodb-export-command}, and an example containing 1 million objects can be
 downloaded at
 http://netbeams.googlecode.com/files/experiment-1000000-data-exported-20090913-053538.csv.tar.gz.
-
-\lstset{label=cmd:mongoexport,caption=Command to export data in CSV format}
-\begin{lstlisting}
-mongoexport -d netbeams -c SondeDataContainer --dbpath ./data/ --csv -f "_id,sensor_ip_address,transaction_time,fact_time,
-data.temperature,data.sp_condition,data.condition,data.resistence,data.salinitude,data.pressure,data.depth,data.ph,data.pH_mv,data.odo_sat,
-data.odo_condition,data.turbidity,data.battery" -o sonde-data-exported.csv
-\end{lstlisting}
 
@@
 
-\lstinputlisting[language=Bash,label=file:experiment-query-scenarios,caption=Experiment
+\lstinputlisting[language=Python,label=file:experiment-setup-mongoshards,caption=Experiment
+Main Execution Script]{../../../../netbeams/versions/v2/persistence/setup-mongo-shards.js}
+
+\lstinputlisting[language=Python,label=file:experiment-export-python,caption=Data
+Export to from mongoDB to OPEnDAP format.]{../../../../netbeams/versions/v2/persistence/data-export-python.py}
+
+\lstinputlisting[language=Python,label=file:experiment-scenarios,caption=Experiment
 Scenarios Implementation]{../../../../netbeams/versions/v2/persistence/experiment-client-scenarios.js}
+
+\lstinputlisting[language=Java,label=file:main-ysi-data-handler,caption=The
+DSP YSI Sonde Data Handler]{../../../../netbeams/versions/v2/apps/xml/jaxb/sonde/java-src/org/netbeams/dsp/ysi/SondeDataType.java}
 
@@
 Random Data to MongoDB]{../../../../netbeams/versions/v2/apps/osgi-bundles/dsp/DSPDataPersistence/src/org/netbeams/dsp/persistence/controller/DSPMessageToMongoDBExperiment.java}
+
+\subsection{Experiment Execution Output}
 
@@
 ...
-...
 Sun Nov 29 19:10:18 getmore netbeams.SondeDataContainer cid:7749363894863150628 ntoreturn:0 query: { query: {}, $snapshot: 1 }  bytes:793098 nreturned:36049 1975ms
@@
 \end{lstlisting}
+
+\lstinputlisting[label=file:experiment-export-python-results,caption=Data
+Export to from mongoDB to OPEnDAP format.]{../../../../netbeams/versions/v2/persistence/export-mongodb-to-sfbeams.csv}

### 2009-12-07 docs/marcello-dissertation/thesis/chapter2-state-of-the-art.tex
Most of the revisions and updates... Changes to chapters 3, 4 after Kleber's chat online (adding missing taxonomy).
Revisions of print chapters until chapter 5.

Chapter 7 is reviewed until before discussion. 

Next revision should include all the necessary reviews for the
release draft.

git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@676 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 gathered data from sensor devices.
+
+It is important to note that the scope of this literature review is limited to
+the scope of the our case study described in chapter 4 and, therefore, it only
+covers a subset of disciplines in sensor networks and data persistence.
 
@@
 
-\section{Data Provenance: How Collected Data is Described?}
-\label{sec:sn-provenance}
+\section{How collected data is described?}
+\label{sec:sn-data-description}
 
@@
 properties of the data. Not to mention that the low-level or raw data produced
-by a given sensor device might not include vital information regarding the
-origing and identity of the collected data. For this reason,
+by a given sensor device might not include vital details regarding the
+origing and identity of the collected data. For this reason, the collected data
+is reported to be described in different ways such as Data Provenance or
+Annotations.
+
 \cite{sn-provenance} describes Data Provenance as a technique used to describe
@@
 example of the application of Data Provenance is the so-called blackbox of an
-aircraft, which carries the collection of raw data collected from a group of
-sensors placed in different locations of an aircraft, as shown by
-\cite{sn-exemple-blackbox}. Last, but not least, the author describes the
-collected data as used in a process of data recovery and analysis in order to
-identify the probable reason of an aircraft accident, by using the time, and
-location of the aircraft, along with the independent variables from the
-sensors.
+aircraft, which carries the collection of spacio-temporal data structures
+collected from a group of sensors placed in different locations of an
+aircraft, as shown by \cite{sn-exemple-blackbox}. Last, but not least, the
+author describes the collected data as used in a process of data recovery and
+analysis in order to identify the probable reason of an aircraft accident, by
+using the time, and location of the aircraft, along with the independent
+variables from the sensors.
 
@@
 delimitation between each other. For this reason, the processes of parsing and
-description of each of the values located in the data streams regularly require
-the use of the device's specification documentation. Similarly, in order to
-have access to each of the properties of data collected by a given property
-related to time or space, \cite{sn-provenance} also mentions that 
+description of each of the values located in the data streams regularly
+require the use of the device's specification documentation. Similarly, in
+order to have access to each of the properties of data collected by a given
+property related to time or space, \cite{sn-provenance} also mentions that
 Metadata\footnote{Data about data} should be included together with the
@@
 not need to be transferred \cite{sn-data-reduction}.
+
+Another different way to describe sensor devices is through the use of
+Annotations. \cite{sn-annotation} shows the use of tags to describe data
+collected from a physical area monitoring. The sensor network was
+designed to monitor a moving objects, capturing their location while in
+movement and their visual appearance using a video camera. In this way,
+different instants can be described by adding tags particular video frame, and
+therefore, providing an efficient way to search of a particular frame based on
+its visual observations. Similarly, this project also uses data provenance to
+describe its collected data, since each of the collected data contains the
+related GPS coordinates of the observed object.
 
@@
 In order to use the collected data for the purpose of historical analysis, the
-use of a point in time is the most prominent feature of data, as it is
-considered a pivotal metadata for different types of sensor networks such as
-an aircraft the blackbox \cite{sn-exemple-blackbox} or any
-environmental sensor network \cite{sfbeams2006, sn-xml-middleware, sn-ex01}.
-In this way, Other projects classify data in time-series
-\cite{sn-time-series-example}, where data is related to all the
-observations a given sensor device performed over a period of time. The
-apparent implications is that all the produced data is necessary, but
-\cite{sn-time-series} argued that some sorts of sensor networks could decrease
-the creation of ``noisy data sets" based on the proper clustering of
-similar time-series data before before they are sent to the data sink.
+use of a point in time is the most prominent feature of data. Also referred to
+as spatial-temporal data, it is considered a pivotal metadata used in different
+types of sensor networks such as an aircraft the blackbox
+\cite{sn-exemple-blackbox}, environmental sensor network \cite{sfbeams2006},
+and data communications management systems \cite{sn-dataware-house}. In this
+way, Other projects classify data in time-series \cite{sn-time-series-example},
+where data is related to all the observations a given sensor device performed
+over a period of time. The apparent implications is that all the produced data
+is necessary, but \cite{sn-time-series} argued that some sorts of sensor
+networks could decrease the creation of ``noisy data sets" based on the proper
+clustering of similar time-series data before before they are sent to the data
+sink.
 
@@
 the latter type of is the point in time at which a given fact was transferred
-to the persistence layer.
+to the persistence layer. Either types of time can be used depending on the
+requirements of the sensor network. For instance, \cite{sn-dataware-house}
+reports that all the data stored by its sensor network uses the time of
+collection, but do not use the transaction time for that purpose.
 
@@
 at using data from different sensor devices.
+
+\section{Data Load and Management Systems}
+\label{sec:data-load}
+
+Depending on the amount of data produced by a sensor network, different
+approaches are reported to be taken to support the data management depending on
+the amount of data collected. For example, \cite{sfbeams2006} is a an
+environmental sensor network that produces data regarding the marine life.
+Less than twenty sensor devices produce Megabytes of data in a daily bases,
+whose data is processed and stored in using flat-files directly in the
+operating system's file system with provenance information regarding the time
+of collection.
+
+When the amount of data collected grows to other scales, say, Gigabyes of data
+in a daily basis, data management tends to become more difficult in any type of
+database systems. Examples of sensor networks that produce data under this
+magnitude of data are reported in the literature. For instance, 
+\cite{sn-dataware-house} describes the creation of a Data Warehouse for the
+real-time sensor data management using a cluster of MySQL databases used to
+store data from different sensors such as optical sensors that collect metrics
+from physics experiments using eletromagnetic fields. As the author reported,
+requirements for data management evolved from storing flat files by a
+monolitic system to a robust distributed system using MySQL database.
+Moreover, the the system targed from 5 to 10MB of data per second, whose
+culmulative data reached 1 TB of data every year.
+
+Data management for large scales of data require different techniques to store
+and retrieve data, since it requires more computational power and data
+retrieval efficient algorithms. \cite{sn-data-center} describes the creation of
+a Sensor Data Center designed to store and analyze amounts of data on the range of
+Terabytes to a future Petabyes of data. Its sensor network deals with not only
+numerical sensor readings from transportation managemenet, but also radar
+images, used for historic purposes and data mining that performs traffic
+pattern analysis. In this way, the authors report the use of parallel data
+processing techniques in order to provide processing for the collected
+data, and load-balancing when storing data to minimize the sensor data
+processing delays. For this reason, they suggest the use of concurrent control
+such as snapshot isolation of the current data into separated replicated data.
+In this way, data access from users are not interrupted by the processing of
+the data that is processed.

### 2009-12-07 docs/marcello-dissertation/thesis/chapter3-taxonomic-classification-of-persistence-storage-for-sensor-networks.tex
Most of the revisions and updates... Changes to chapters 3, 4 after Kleber's chat online (adding missing taxonomy).
Revisions of print chapters until chapter 5.

Chapter 7 is reviewed until before discussion. 

Next revision should include all the necessary reviews for the
release draft.

git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@676 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
   purposes and is related to the historical data analysis \cite{sn-intro01,
-  sn-intro02} or Information Fusion \ref{sn-info-fusion}. 
+  sn-intro02} or Information Fusion \cite{sn-info-fusion}. 
 \end{itemize}
@@
   external storage device with a data management system such as a relational
-  database;
-  \item \textbf{Data-Centric Storage}: when the collected data occupies
-  different external storages, being organized by categories based on the
-  collected data keys and values. In other words, what characterizes this taxon
-  is the presence of a data partitioning strategy to store the collected data.
+  database. When the collected data occupies different external storages, the
+  strategy is called \textbf{Data-Centric Storage}, being organized by
+  categories based on the collected data keys and values. In other words, what
+  characterizes this taxon is the presence of a data partitioning strategy to
+  store the collected data.
 \end{itemize}
@@
   \centering
-  \includegraphics{../diagrams/taxonomy-data-model}
-  \caption{Data Model Taxonomy}
+  \includegraphics[scale=0.5]{../diagrams/taxonomy-data-model}
+  \caption{The Data Model Taxonomy}
   \label{fig:taxonomy-data-model}
@@
 
-\section{Data Provenance Taxonomy}
+\section{Data Description Taxonomy}
 
-No matter which data model used to represent the collected data, Data
-Provenance provides a valuable guideline on how to describe the collected data
-from sensors devices. As reported in section \ref{sec:sn-provenance}, the
-description of the properties should take into account different properties of
-the nature of the data as shown in Figure \ref{fig:taxonomy-data-provenance}.
-Those taxa are summarized as follows:
+Particulaly related to how the collected data is described, different projects
+provide data persistence describing the collected data by using its natural
+properties, or associating textual descriptions or keywords of the data. The
+way to describe data can follow different guidelines such as data provenance
+and annotations, discussed in section \ref{sec:sn-data-description}. As
+reported in section \ref{sec:sn-data-description}, the description of the
+properties should take into account different properties of the nature of the
+data and any other description needed. Therefore, the taxa that composes the
+Data Description taxonomy, shown in Figure
+\ref{fig:taxonomy-data-description}, are described as follows:
 
@@
   \centering
-  \includegraphics{../diagrams/taxonomy-data-provenance}
-  \caption{Data Provenance Taxonomy}
-  \label{fig:taxonomy-data-provenance}
+  \includegraphics[scale=0.5]{../diagrams/taxonomy-data-description}
+  \caption{The Data Description Taxonomy}
+  \label{fig:taxonomy-data-description}
 \end{figure}
@@
 \begin{itemize}
-  \item \textbf{What: Data Identity}: the data type that uniquely identifies
-  the collected data from sensors. First and foremost, the properties of the
-  sensor devices must be described according to the meaningful naming
-  convention for the property. For instance, ``Water Temperature'' better
-  describes the intent of the property as ``temp''. Second, the identification
-  of the collected data related to the sensor type should be given, that is,
-  which sensor device created the collected data, in case the data origin is
-  relevant;
-  \item \textbf{When: Time Dimension}: time is an important attribute of the
-  data because it identifies the age of the collected data. Depending on the 
-  use of the collected data, this is often taken into account. Two different
-  types of time dimensions were described: \textbf{fact} and
-  \textbf{transaction} times. The former is used to identify when the data
-  was collected, whereas the latter when the data was transmitted to the
-  data sink where the query processing acts;
-  \item \textbf{Where: Data Location}: the point where the collected data was
-  observed by the sensor device. Usually the latitude and longitude as the GPS
-  coordinates are provided by devices which are capable of sensing the
-  location. A descriptive metadata about the location could also be given such
-  as room-name = ``deep see''.
+  \item \textbf{Provenance-Aware}: when the collected data is categorized based
+  on the nature of the data, \textbf{What: Data Identity} is the data type
+  that uniquely identifies the collected data from sensors. Similarly, the
+  \textbf{When: Time Dimension}, is related instants of time related to the
+  collected data: \textbf{fact} and \textbf{transaction} times. The former is
+  used to identify when the data was collected, whereas the latter when the data was
+  transmitted to the data sink where the query processing acts. Finally, the
+  \textbf{Where: Data Location} identifies the point where the collected data
+  was observed by the sensor device. Usually the latitude and longitude as the
+  GPS coordinates;
+  \item \textbf{Annotations}: the annotation of the collected data, through the
+  use of keywords, full-text description, or any other method that helps
+  describing the collected data. It is important to note that this type of data
+  does not make part of the collected data from sensors. Rather, they
+  complement the observations of what was percieved by the sensor devices.
 \end{itemize}
@@
   \includegraphics{../diagrams/taxonomy-query-mechanism}
-  \caption{Query Processing Mechanism Taxonomy}
+  \caption{The Query Processing Mechanism Taxonomy}
   \label{fig:taxonomy-query-mechanism}
@@
 
-\section{Database System Organization Taxonomy}
+\section{Data Volume Taxonomy}
 
-Different approaches of data systems were described in the literature review in
-the last chapter in different sections. This taxonomy is directly related to
-the use of external storage devices on a centralized network node so that it
-often uses a single database server. Paradoxically, the use of a data-centric
-storage determines the use of distributed database nodes around the network. 
+When it comes to the data volume produced during a time frame, different ranges
+were reported from different types of sensor networks. Depending on the types
+of data produced such as numerical or images, they can be in the orders of
+Megabytes (1MB = 1000KB = ~10\superscript{3} bytes), Gigabytes (1GB =
+1000\superscript{3} = 10\superscript{9} bytes) and even Terabytes (1TB =
+1000\superscript{4} = 10\superscript{12}) to future Petabytes (1PB =
+1000\superscript{5} = 10\superscript{15} bytes) as shown in section
+\ref{sec:data-load}. As a comparison of what is processed, Google reports to
+process 20 Petabytes of data per day in their data centers
+\cite{map-reduce-load}. In this way, the following taxa are related to the
+different sizes of data produced by a sensor network, summarized in Figure
+\ref{fig:taxonomy-data-volume}:
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=0.5]{../diagrams/taxonomy-data-volume}
+  \caption{The Data Volume Taxonomy}
+  \label{fig:taxonomy-data-volume}
+\end{figure}
+
+\begin{itemize}
+  \item \textbf{Small}: sensor networks whose yearly produced data is in the
+  range of MBs or to 1GB of data, usually stored in flat files directly in the
+  operating system. They are usually collected from sensor devices that only
+  generate numerical data;
+  \item \textbf{Medium}: amount of data that ranges from GBs to 1TB data. In
+  order to manage such amount of data using disk arrays;
+  \item \textbf{Large}: when the amount of data produced by a sensor device
+  reaches amounts of data compatible to data centers, in the orders ranging
+  from TBs to 1PT of data. The construction of a data center specialized for
+  data produced by sensor devices, taking into account the addition of data
+  provenance, among others.
+\end{itemize}
+
+\section{System Organization Taxonomy}
+
+Different approaches of are developed computational systems to deal the data
+produced from sensor networks. As  
 In this way, Figure \ref{fig:taxonomy-database-architecture} describes the
@@
   \centering
-  \includegraphics{../diagrams/taxonomy-database-architecture}
-  \caption{Database Architecture Taxonomy}
+  \includegraphics[scale=0.5]{../diagrams/taxonomy-database-architecture}
+  \caption{The System Organization Taxonomy}
   \label{fig:taxonomy-database-architecture}
@@
   \item \textbf{Centralized System}: a database system that runs in one single
-  centralized host, being the focal purpose of data traffic as read and writes 
+  centralized host, being the focal purpose managing the collected data
   \cite{sn-intro01};
@@
   more than one node. Usually, the use of the pattern of Master-Slave
-  mechanism or data ``Replication" is used. Another mechanism is the use of
-  database paritioning, if the technology supports it.
+  mechanism or data ``Replication" is used to aliviate ``hot spots'' during
+  concurrent processes of data collection from the sensor devices and data
+  access from the final users. On the other hand, unique mechanisms can be
+  used to disagragate the produced data from a concentrated one: the use 
+  of database partioning, as related to the data-centric storage, where  
+  each partition is responsible to hand data produced by a given type of 
+  sensor, for instance.
 \end{itemize}
@@
 
-The different taxonomies defined in the previous sections of this chapter are
-important to the understanding of a data persistence layer for sensor
-networks. As they are closely related to each other, it is important to
-briefly highlight these relationships in order to better analyze a problem
-related to data persistence.
+On the whole, the taxonomies described in this chapter were based solely on the
+literature review of described in the previous chapter. They can be helpful in 
+understanding of requirements of a data persistence layer for sensor networks
+with different infrastructures. They are summarized on table
+\ref{tab:taxonomies-list-summary}, and briefly related highlight these
+relationships in order to better analyze a problem related to data persistence.
+
+\begin{table}[!b]
+    \label{tab:taxonomies-list-summary}
+    \begin{center}
+        \begin{tabular}{|p{170pt}|p{250pt}|}\hline 
+        \textbf{Taxonomy Name} & \textbf{Associated Taxas}\\\hline 
+        Purpose of the Sensor Data & Real-time Data Stream, Data Archival \\\hline 
+        The Location of the Sensor Data & Local Storage, External Storage\\\hline
+        Data Model & Schema-Dependent, Schema-less Models\\\hline 
+        Data Description & Provenance-Aware, Annotations\\\hline
+        Query Processing & In-Network Query, Centralized Query\\\hline 
+        Data Volume & Small, Medium, Large\\\hline
+        System Organization & Centralized System, Distributed System\\\hline
+        \end{tabular}
+    \end{center}
+    \caption{Summary of the Data Persistence Taxonomies}
+\end{table}
 
@@
 real-time data stream relates to the local data storage, the taxon related to
-data archival correlates the storage of the collected data into an external
-storage. In addition to that, the converse is also true.
+data archival correlates the storage of the collected data with the an external
+storage one. In addition to that, the converse is also true.
 
 The taxonomy related to the data model can be related to any taxon from the
-taxonomy of the location of the sensor data, as well as the taxons related to
-Data Provenance. In the former, the data of a relational model is stored using
-SQL messages.
+taxon described from the system organization, to data description and query
+processing and location of the collected data. In the former, the data of a
+relational model is stored using SQL messages.
+
+Similarly, the taxonomy related to the volume of data produced by a sensor
+network usually determines the system organization used to handle and process
+the collected data. When the volume is small, the use a single server is
+required, usually using an external storage device, which relates to the
+taxonomy for storage system. However, when the volume is medium to large, the
+use of distributed systems is required to cope with high volume of data
+processing. Techniques described for data replication is used to decrease the
+load in hot spots of the architecture. On the other hand, approaches to
+separate the data being written and read can be used such as Load Balancing or
+Database Partition. As a consequence, the use of a Data-Centric storage
+approach is used.
 

### 2009-12-07 docs/marcello-dissertation/thesis/chapter4-netbeams-in-depth-requirements-analysis-and-specification.tex
Most of the revisions and updates... Changes to chapters 3, 4 after Kleber's chat online (adding missing taxonomy).
Revisions of print chapters until chapter 5.

Chapter 7 is reviewed until before discussion. 

Next revision should include all the necessary reviews for the
release draft.

git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@676 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 \subsection{The SF-BEAMS Infrastructure}
+\label{sec:sfbeams}
 
@@
 be directly transferred to the labs via Ethernet cables or collected manually
-with a laptop computer.
+with a laptop computer. The current data collection process for SF-BEAMS is
+described in figure
+\ref{fig:SF-BEAMS-system-architecture}\cite{sfbeams-current-system}. Upon
+collecting data from sensors, the RTC staff uses automation scripts written in
+Matlab \cite{matlab} to process, index and distribute the raw data in
+different formats. One of these formats is the OPeNDAP \cite{opendap}, which
+is widely used at research institutions that promote easier data exchange
+among them. The SF-BEAMS website provides access to the collected data through
+the Internet at http://sfbeams.sfsu.edu:8080/opendap. An example of an HTTP
+Request from accessing the data for a specific YSI is shown in Listing
+\ref{file:rtc-ysi-opendap}.
 
-\subsection{SF-BEAMS produced data and collection process}
-\label{sec:sfbeams}
+\begin{figure}[!b]
+  \centering
+  \includegraphics[scale=0.5]{../diagrams/SF-BEAMS-system-architecture}
+  \caption{The Current SF-BEAMS Data Collection Process}
+  \label{fig:SF-BEAMS-system-architecture}
+\end{figure}
 
@@
 its observed data based on properties of measurements defined by its
-manufacturer. One example of such sensors is the ``YSI 6600 ESD V2", as seen in
-Figure \ref{fig:ysi-device}. It is a powerful water quality monitoring device,
-capable of producing around 52 bytes on a single real-time data stream reading,
-as shown on Table \ref{tab:ysi-data-stream}.
+manufacturer. NetBEAMS has used ``YSI 6600 ESD V2", as seen in Figure
+\ref{fig:ysi-device}, one of the sensors devices used by SF-BEAMS. It is a
+powerful water quality monitoring device, capable of producing around 52 bytes
+on a single real-time data stream reading, as shown on Table
+\ref{tab:ysi-data-stream}.
 
@@
 
-\begin{table}[!b]
-    \caption{YSI Data Stream}
+\begin{table}
     \begin{center}
-        \begin{tabular}{lr}
-  12.20    192    179 5588.40   0.09   0.084   0.059  7.98   -79.6   99.5  8.83 
-  0.4     8.7
+        \begin{tabular}{|l|}\hline
+  12.20    192    179 5588.40   0.09   0.084   0.059  7.98   -79.6   99.5  8.83  0.4     8.7\\\hline
         \end{tabular}
@@
     \label{tab:ysi-data-stream}
+    \caption{The Collected Data Stream from the YSI Sonde 6600ESDV2}
 \end{table}
@@
 around hundreds of Megabytes a year, as \textbf{483,840} samples are generated
-by one single device. Upon collecting data from sensors, the RTC staff uses
-automation scripts written in Matlab \cite{matlab} to process, index and
-distribute the raw data in different formats. One of these formats is the
-OPeNDAP \cite{opendap}, which is widely used at research institutions that
-promote easier data exchange among them. The SF-BEAMS website provides access
-to the collected data through the Internet at
-http://sfbeams.sfsu.edu:8080/opendap. An example of an HTTP Request from
-accessing the data for a specific YSI is shown in Listing
-\ref{file:rtc-ysi-opendap}.
+by one single device.
 
@@
   files. In this way, the \textbf{Schema-less} is used;
-  \item \textbf{Data Provenance}: once the collected data reaches the RTC
+  \item \textbf{Data Description}: once the collected data reaches the RTC
   lab, a quality assurance process is executed, and the data is transformed
@@
   data is stored on file system;
-  \item \textbf{Database System Organization}: there are no databases in use,
-  but the use the OPeNDAP Hyrax server as a middleware that servers the
-  collected data using the Internet.
+  \item \textbf{Data Volume}: the volume of data produced is in the range of
+  small. It is related to the volume of data produced by the sensor devices,
+  which is numerical data streams, and the total number of sensor devices;
+  \item \textbf{System Organization}: SF-BEAMS runs on a Single System, as
+  the collected data is stored directly in the file-system after previous 
+  processing. However, they provide the collected data to the Internet using
+  the OPeNDAP Hyrax system as a middleware.
 \end{itemize}
@@
 
-The scope of this work then is defined as to provide a persistence layer for
+The scope of this work is defined as to provide a persistence layer for
 NetBEAMS, making a selection of a persistence layer that is an exception to the

### 2009-12-07 docs/marcello-dissertation/thesis/chapter5-taxonomic-persistence-storage-selection-for-netbeams.tex
Most of the revisions and updates... Changes to chapters 3, 4 after Kleber's chat online (adding missing taxonomy).
Revisions of print chapters until chapter 5.

Chapter 7 is reviewed until before discussion. 

Next revision should include all the necessary reviews for the
release draft.

git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@676 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 
-\chapter{Data Persistence Technology for Sensor Networks: An Empirical-based
-Choice for NetBEAMS}
+\chapter{Empirical Analysis of Data Persistence Taxonomies for NetBEAMS}
 
-This chapter analyzes each of the taxonomies defined in chapter
-\ref{chap:taxonomies} against different other variables. First, the
-requirements defined by NetBEAMS, as our case study in chapter
-\ref{sec:problem-requirements}, is to be compared. Then, the scope of this
-work is compared to the technologies selected based on the readings of the
-literature reviewing data persistence in sensor networks and other recent
-trends in performance. Finally, a table summarizes the contenders strengths
-and discusses the selected technology to be integrated with the design of a
-Data Persistence component.
+Providing data persistence for an existing sensor networks require analyzes of
+the infrastructure and the nature of the produced data. Based on the taxonomies
+of chapter 3, and the requirements of a data persistence for NetBEAMS in
+chapter 4, this chapter aims at analyzing a group of database technologies
+selected from literature review topics covered in chapter 2. Similarly,
+supported by the scope of the required implementation presented in this
+chapter, the comparison of the technogies is summarized, highlighting the
+selected database system used for the design of a component for NetBEAMS and
+used for the design of the experiments.
 
@@
 \begin{itemize}
-  \item \textbf{MySQL}\cite{mysql}: an open-source relational database system
-  that is the primary choice of the development of projects of different sizes.
-  It was selected because of its popularity in the contemporary community;
-  \item \textbf{Oracle}\cite{oracle}: is the most popular commercial database
-  system, having a campus wide academic licenses available;
+  \item \textbf{MySQL}\cite{mysql}: system used in the implementation of a
+  dateware house with sensor network data \cite{sn-dataware-house};
   \item \textbf{TinyDB}\cite{tinydb}: as referred in the literature review,
@@
 which uses denormalized data models. There are different approaches to
-partition data \cite{db-shard-schemas, db-partitioning-relational,
-db-partitioning-relational-oracle}.
+partition data \cite{db-shard-schemas, db-partitioning-relational}.
 
@@
   \centering
-  \includegraphics[scale=0.65]{../diagrams/database-sharding-by-region}
+  \includegraphics[scale=0.5]{../diagrams/database-sharding-by-region}
   \caption{Example of Data-Centric Storage for Sensor Networks using Database
@@
   to partition the data;
-  \item \textbf{Oracle}: Supports External or Data-Centric Storage approaches.
-  Only sup-  ports vertical partitioning with the definition of different types
-  \cite{db-table-partition, db-partitioning-relational-oracle} when using
-  Data-Centric storage;
-  \item \textbf{Berkeley TinyDB}: Only supports the External Storage;
+  \item \textbf{TinyDB}: Only supports the External Storage;
   \item \textbf{MongoDB}: Supports External or Data-Centric Storage. When the
@@
 they all differ with the support to Data-Centric Storage support.
-
-\section{Analysis of the Query Processing Mechanism Taxonomy}
-
-As a direct result from the previous section, the use of \textbf{Centralized Query
-Processing} is indicated for NetBEAMS, also given the fact that SF-BEAMS data
-goes directly to the individual network sink and, therefore, a centralized
-point of the data.
-
-\subsection{Advantages of the Centralized Query Processing}
-
-Centralized data management and query processing is simpler than the in-network
-query method. Data is verified by a straightforward database management system
-without the risk of data being unavailable, when the data is spread among the
-network nodes.
-
-\subsection{Disadvantages of the Centralized Query Processing}
-
-When the Centralized Query processing is used, problems may occur. Depending on
-the size of the datasets and the database technology, performance may be a
-concern. For example, this procedure creates of the alleged Funneling Affect
-\cite{sn-storage04}, since the point-of-traffic is concentrated in the
-centralized system.
-
-In order to mitigate the problems originated by this query processing
-mechanism, techniques such as Data Replication can be used.
-
-\subsection{Technology Analysis}
-
-All the database systems technology support a centralized query system.
 
@@
 and the use of the YSI Sonde data as the main entity in the system, let Figure
-\ref{fig:Relational-Model-Original} represent a prototype of the relational model, after passing through the process of
-normalization \cite{db-normalization}.
+\ref{fig:Relational-Model-Original} represent a prototype of the relational 
+model, after passing through the process of normalization
+\cite{db-normalization}.
 
@@
   \centering
-  \includegraphics[scale=0.65]{../diagrams/Relational-Model-Original}
+  \includegraphics[scale=0.5]{../diagrams/Relational-Model-Original}
   \caption{Relational Data Model for NetBEAMS - A first prototype}
@@
   \centering
-  \includegraphics[scale=0.65]{../diagrams/Relational-Model-Addition-Modified}
+  \includegraphics[scale=0.5]{../diagrams/Relational-Model-Addition-Modified}
   \caption{Relational Data Model for NetBEAMS - Modified version with new
@@
 \end{figure}
-
-Although this data model have been used in different types of applications,
-developers have tried to use the Relational Model to model an application that
-could prevent any database change through the use of a technique that relates a
-key to a value, called Key-Value data model. With the advent of Internet
-applications, the need of such a model that does not need constant refactoring
-led developers to propose such approach using the relational model as
-documented in technical articles such as  and
-. Based in these articles, consider the relational
-model depicted in Figure  as a data model for
-persisting data to our case study using the key-value strategy. The problem
-with it is that key repetition occurs, since the nature of time-series data
-requires the use of a timestamp key for each property of the sensor. Therefore,
-this strategy is not well-suited to provide persistence data from collected
-sensor networks.
 
@@
 
-There are no records of the use of this data model with Sensor Networks.
-Different variants of such data model is the document-oriented data model,
-where data is modeled using structured documents. However, these documents have
-a dynamic structure that can be freely described without the use of a schema
-validation mechanism as used in XML Schema. One example of such approach is the
-use of the JSON data format [Ste08] in the implementation of database systems
-which are becoming standard with the latest trend called Cloud Computing [LKN+
-09]. Databases implementing this strategy is the mongoDB \cite{mongodb},
-CouchDB \cite{couchdb}, Amazon Dynamo \cite{db-kvp}, among others.
-
 \begin{itemize}
@@
 provenance data that describe the data, depicted in section
-\ref{sec:sn-provenance}, are represented by means of key-value pairs
+\ref{sec:sn-data-description}, are represented by means of key-value pairs
 \ref{fig:persistence-example-kvp}.
@@
   \centering
-  \includegraphics[scale=0.75]{../diagrams/persistence-example-kvp}
+  \includegraphics[scale=0.5]{../diagrams/persistence-example-kvp}
   \caption{KVP Database instance Prototype}
@@
 The relational and key-value data models were compared by
-\cite{db-is-rdbs-dommed} and can be summarized as follows:
+\cite{db-is-rdbs-dommed} and is compared on table
+\ref{tab:schema-vs-schemaless}:
 
 \begin{table}
-    \label{tab:ysi-data-distribution}
+    \label{tab:schema-vs-schemaless}
     \caption{Schema-Dependent X Schema-less Databases Compared: Properties}
@@
 \begin{table}
-    \label{tab:ysi-data-distribution}
+    \label{tab:schema-vs-schemaless-daccess}
     \caption{Schema-Dependent X Schema-less Databases: Data Access}
@@
 \begin{table}
-    \label{tab:ysi-data-distribution}
+    \label{tab:tab:schema-vs-schemaless-api-interface}
     \caption{Schema-Dependent X Schema-less Databases: Application Interface}
@@
 \begin{itemize}
-  \item MySQL: supports the Relational Model;
-  \item Oracle: supports Relational Model;
-  \item Berkeley TinyDB: supports the Relational Model; 
+  \item \textbf{MySQL}: supports the Relational Model;
+  \item \textbf{TinyDB}: supports the Relational Model; 
   \item \textbf{MongoDB}: supports Document-Oriented Model;
-  \item IBM DB2: supports the Relational Model and the XML model.
+  \item \textbf{IBM DB2}: supports the Relational Model and the XML model.
 \end{itemize}
 
-\section{Analysis of the Database System Organization Taxonomy}
+\section{Analysis of the Query Processing Mechanism Taxonomy}
+
+As a direct result from the previous section, the use of \textbf{Centralized Query
+Processing} is indicated for NetBEAMS, also given the fact that SF-BEAMS data
+goes directly to the individual network sink and, therefore, a centralized
+point of the data.
+
+\subsection{Advantages of the Centralized Query Processing}
+
+Centralized data management and query processing is simpler than the in-network
+query method. Data is verified by a straightforward database management system
+without the risk of data being unavailable, when the data is spread among the
+network nodes.
+
+\subsection{Disadvantages of the Centralized Query Processing}
+
+When the Centralized Query processing is used, problems may occur. Depending on
+the size of the datasets and the database technology, performance may be a
+concern. For example, this procedure creates of the alleged Funneling Affect
+\cite{sn-storage04}, since the point-of-traffic is concentrated in the
+centralized system.
+
+In order to mitigate the problems originated by this query processing
+mechanism, techniques such as Data Replication can be used.
+
+\subsection{Technology Analysis}
+
+All the database systems technology support a centralized query system.
+
+\section{Analysis of the System Organization Taxonomy}
 
@@
 
-\subsection{Advantages of Database Partitioning and Sharding}
+\subsection{Advantages of Database Partitioning}
 
@@
 
-\subsection{Disadvantages of Database Partitioning and Sharding}
+\subsection{Disadvantages of Database Partitioning}
 
@@
   \item \textbf{MySQL}: Supports Data Replication;
-  \item \textbf{Oracle}: Supports Data Replication;
-  \item \textbf{Berkeley TinyDB}: NO Supports to distributed data;
+  \item \textbf{TinyDB}: NO Supports to distributed data;
   \item \textbf{MongoDB}: Supports Data Replication;
@@
   scripting APIs from  the community;
-  \item \textbf{Oracle}: It is not Open-Source, but contains lots of support
   from the community. Also APIS are available for the majority of languages;
-  \item \textbf{Berkeley TinyDB}: Not open-source;
+  \item \textbf{TinyDB}: Not open-source;
   \item \textbf{MongoDB}: An open-source database system with an increasing
@@
     \label{tab:ysi-data-distribution}
-    \caption{Amount of data produced by the RTC's YSI sondes}
     \begin{center}
         \begin{tabular}{|c|c|c|c|c|c|c|}\hline 
-        \textbf{Database} & \textbf{MySQL} & \textbf{Oracle} &
-        \textbf{TinyDB} & \textbf{MongoDB} & \textbf{IBM DB2}\\\hline 
-        Centralized Query & + & + & + & + & + \\\hline 
-        Distributed System & ++ & ++ & - & ++ & +\\\hline 
-        Schema-less & - & - & - & + & +\\\hline 
-        Provenance Support & + & + & + & + & +\\\hline 
-        NO-SQL Query & - & - & - & + & +\\\hline 
-        Data Partition & + & + & - & + & +\\\hline 
-        Hot Backup & + & + & - & - & +\\\hline 
-        Export Capability & + & + & - & + & -\\\hline 
-        Programming Lang. & + & + & + & + & +\\\hline
-        Script Lang. & + & + & - & + & +\\\hline
-        Open-Source & + & - & + & + & -\\\hline
+        \textbf{Database} & \textbf{MySQL} & \textbf{TinyDB} & \textbf{MongoDB} & \textbf{IBM DB2}\\\hline
+        Centralized Query & + & + & + & + \\\hline 
+        Distributed System & ++ & - & ++ & +\\\hline 
+        Schema-less & - & - & + & -\\\hline 
+        Provenance Support & + & + & + & +\\\hline 
+        NO-SQL Query & - & - & + & -\\\hline 
+        Partition & + & - & + & +\\\hline 
+        Hot Backup & + & - & + & +\\\hline 
+        Export Capability & + & - & + & -\\\hline 
+        Programming Lang. & + & + & + & +\\\hline
+        Script Lang. & + & - & + & +\\\hline
+        Open-Source & + & + & + & -\\\hline
         \end{tabular}

### 2009-12-07 docs/marcello-dissertation/thesis/chapter6-data-persistence-component-design-and-architecture.tex
Most of the revisions and updates... Changes to chapters 3, 4 after Kleber's chat online (adding missing taxonomy).
Revisions of print chapters until chapter 5.

Chapter 7 is reviewed until before discussion. 

Next revision should include all the necessary reviews for the
release draft.

git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@676 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 
-In order to provide data to users, the RTC scientists use OPeNDAP \cite{opendap}
-server to distribute collected data over the Internet. As discussed in the
-previous chapter, this work proposes the use a Key-Value database system
-in order to provide an alternative way for data persistence in sensor
-networks, taking into account the taxonomies discussed earlier in chapter 
-\ref{chap:taxonomies}.
+In order to provide information to users, the RTC scientists use OPeNDAP
+\cite{opendap} server to distribute collected data over the Internet. As
+discussed in the previous chapter, this work proposes a data persistence
+application of a Key-Value database system in order to provide an alternative
+method for data persistence in sensor networks, taking into account the
+taxonomies discussed earlier in chapter \ref{chap:taxonomies}.
 
-Considering our case study being NetBEAMS, and its infrastructure, this chapter
-describes the design and architecture component designed specifically for the
-Data Sensor Platform, DSP, in order to provide a persistence layer for
-NetBEAMS. It is important to note that the technology chosen in the previous
-chapter can be adapted by any alternative that uses the data model and
-infrastructure options presented.
+Considering our case study being NetBEAMS, this chapter describes the design
+and architecture component designed specifically for the Data Sensor Platform,
+DSP, in order to provide a persistence layer for NetBEAMS. It is vital to note
+that the technology chosen in the last chapter can be adapted by any
+alternative that uses the data model and infrastructure options presented.
 
-The proposed DSP Data Persistence component for the DSP Platform was
-implemented by using the same approaches described in chapter
-\ref{chap:netbeams-overview}. The following sections describe the scenarios
-for persistence, analyzes the requirements from RTC, and describes the design
-and architecture of the solution.
+The proposed DSP Data Persistence component for the DSP
+Platform was implemented by using the same approaches described in chapter
+\ref{chap:netbeams-overview}. The following sections describe the scenarios for
+persistence, analyze the requirements from RTC, and describe the design and
+architecture of the solution.
 
@@
 \label{sec:business-process-analysis}
+
 As described in section \ref{chap:netbeams-overview}, the DSP Platform was
-designed to support an environmental sensor network based in San Francisco
-Bay, composed by different types of sensors. The process of extracting the
-measurements from the sensors to be reused through the DSP Platform can be
-summarized following the UML Business Process Diagram \cite{uml} depicted in
-figure \ref{fig:dsp-persistence-business-process}. It is important to note
-that this diagrams details remote communication between sensor devices and a
-centralized server, as part of the main scenario to solve the problem of data
-persistence.
+designed to support an environmental sensor network based in San Francisco Bay,
+composed by different types of sensors. The process of extracting the
+measurements from sensors can be summarized following the UML Business Process
+Diagram \cite{uml} depicted in figure
+\ref{fig:dsp-persistence-business-process}. It is important to note that this
+diagram details remote communication between sensor devices and a centralized
+server, as part of the main scenario to solve the problem of data persistence.
 
@@
 At the event of data collection, the DSP Client receives measurements collected
-from each of the sensor devices. Each of the co-located DSP Client processes
-the collected measurements and transforms them into the so-called DSP Message
+from each of the sensor devices. Each of the co-located DSP Client process the
+collected measurements and transform them into the alleged DSP Message
 (section \ref{sec:dsp-message}), which embodies the actual data values. After
 proper data processing, transformation and categorization, the DSP Message is
-added into temporary outbound queue. Then, as an asynchronous data
-transmission event occurs with an specified rate, transferring all the queued
-messages to be processed by a DSP Server host. As it is detailed in image
-\ref{fig:dsp-persistence-business-process}, the server is centralized and is
-able to concurrently process messages.
+added into temporary outbound queue. Then, as an asynchronous data transmission
+event occurs with a specified rate, transferring all the queued messages to be
+processed by a DSP Server host. As image
+\ref{fig:dsp-persistence-business-process}, the server is centralized and able
+to concurrently process messages.
 
@@
 the processed measurement data from sensors 1 through N must be collected in
-the same fashion and categorized as necessary. In this way, end users such as
-Biologists and/or Programmers can access the collected measurements directly
-from the Database. Furthermore, the same measurement data can be exported to
-different formats as the users find necessary.
+the same style and categorized as necessary. In this way, end users such as
+Biologists or Programmers can access the collected measurements directly
+from the Database. Furthermore, the same collected data can be exported to
+different formats as necessary.
 
 As a matter of fact, the data measurements collected from the sensor devices
-are simple properties with keys and associated values as described in section
+are basic properties with keys and associated values as described in section 
 \ref{sec:sfbeams}, after data has been described. According to the descriptions
@@
 version in XML \cite{xml}. Similarly, the collected data contains measurements
-that uniquely identifies them, together with other properties such as time
-stamps of data creation and collected, as well as the the IP
-address\footnote{Internet Protocol Address} of the DSP Client that produced
-the data. Therefore, these are important properties for the proposed solution
-of this work.
+that uniquely identifies them, together with other properties such as
+timestamps of data generation and collected and the IP address of the DSP
+Client that produced the data. Therefore, these are essential properties for
+the proposed solution of this work.
 
@@
 well as by collected requirements from the RTC staff, this section lists the
-main use cases that the system and the primary users require. One of the
-primary requirements of the system is to enable the use a persistence system
-that does not promote the use of specialized professionals to deal with data
-modeling, as it is required in most projects using the Structured Query
-Language (SQL) \cite{sql}, used to extract data from a Relational Databases.
+main use cases that the system and the primary users require.
 
@@
 \end{figure}
-
-The functional requirements listed in this section are in the format of UML
-User Cases Diagrams \cite{uml}. Figure
-\ref{fig:DSP-Data-Persistence-UseCases-Diagram-System} summarizes the use 
-cases that the system, as the main actor of the system, must perform:
 
@@
 
+Different constraints are required to execute the functional requirements
+described in the previous section. They are listed as non-functional
+requirements as follows:
+
 \begin{itemize}
   \item The DSP Data Persistence component must maintain the received DSP
-  Measurement Messages in-memory for a specific rate in order to decrease the
-  write load on the database;
+  Measurement Messages in-memory for a specific period of time in order to
+  decrease the write load on the database;
   \item The Persistence storage system must be able to cope with dozens of DSP
-  Messages;
+  Messages received from different sensors;
   \item The Persistence storage system must be able to save measurements quickly;
@@
   architectural changes, or schema changes;
-  \item The Persistence storage system must not impose the knowledge of
-  complicated database system languages such as SQL \cite{sql} or XML XPath
-  \cite{xml-xpath}, but use access through API calls, as it offers better
-  abstractions for non-computer science researchers
+  \item The Persistence storage system must not impose the use of
+  database languages such as SQL \cite{sql} or XML XPath \cite{xml-xpath}
+  related to the data being persisted. Instead, data must be handled through API
+  calls, as better abstractions for non-computer science researchers
   \cite{sn-programming-language}.
@@
 
-As described in the previous chapter, the data model that might give better
-support for collected data from sensor networks is the Key-Value-Pair Data
-Model. Beyond the data collected from the sensor devices, Data Provenance 
-techniques must also be considered to better describe the collected data
-\cite{sn-provenance}. In general, the data model must include the following
-metadata besed on Data Provenance:
+As described in the previous chapter, the data model chosen to support
+persistence of collected data from sensor networks is the Key-Value-Pair Data
+Model \cite{db-kvp}. Beyond the data collected from the sensor devices, Data
+Provenance techniques must also be considered to better describe the collected
+data \cite{sn-provenance}. In general, the data model must include the
+following metadata besed on Data Provenance:
 
@@
 location are related to the provenance support discussed in section
-\ref{sec:sn-provenance}:
+\ref{sec:sn-data-description}:
 
@@
   bootstrap message for the component;
-  \item \textbf{message\underline{ }id}: it's extracted from each of the messages contained
-  in the DSP Message Container;
+  \item \textbf{message\underline{ }id}: it's extracted from each of the
+  messages contained in the DSP Message Container;
   \item \textbf{time.transaction}: it is extracted from the DSP
-  message container creation time and it is used to identify when the transaction occurred;
+  message container creation time and it is used to identify when the
+  transaction occurred;
   \item \textbf{time.fact}: it is extracted from the
@@
   \item \textbf{observation}: this key defines the values of the document, and
-  will have every different property of the sensor. For instance, the data key for the
-  properties of the YSI device will contain the set of properties defined for
-  the device, such as battery, temperature, salinity, etc.
+  will have every different property of the sensor. For instance, the data key
+  for the properties of the YSI device will contain the set of properties 
+  defined for the device, such as battery, temperature, salinity, etc.
 \end{itemize}
@@
 \end{figure}
-
-More details about the implementation of the DSP Data Persistence can be see in
-the next chapter.

### 2009-12-07 docs/marcello-dissertation/thesis/chapter8-experimental-results-correct-behavior-and-performance.tex
Most of the revisions and updates... Changes to chapters 3, 4 after Kleber's chat online (adding missing taxonomy).
Revisions of print chapters until chapter 5.

Chapter 7 is reviewed until before discussion. 

Next revision should include all the necessary reviews for the
release draft.

git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@676 559edde2-a4d4-44e0-96fc-4035e40d7396

--- 2009-11-16 docs/marcello-dissertation/thesis/chapter8-experimental-results-correct-behavior-and-performance.tex
@@
 
-\section{Design Model For Sensor Networks}
+As discussed in the previous chapter, the implementation of a data persistence
+layer for NetBEAMS followed the requirements defined in chapter 4 having in
+mind the classification of the SF-BEAMS sensor network and the software
+infrastructure provided by the Data Sensor Platform. As a result, the
+implementation was developed using the infrastructure provided by the online
+version-control system repository from NetBEAMS, whose documentation is
+detailed in section \ref{sec:dsp-data-persistence-implementation}.
 
-Sensor Networks are specialized types of network systems comprised of network 
-nodes with specific goals to observe and collect data. The following
-subsections describe classes of characteristis of these types of sensor
-networks, with applications on research, industrial or militery.
+First, this chapter details the experiment setup designed to evaluate the
+proposed implementation, showing the measurement results collected from the
+simulation log and based on the user experience while reusing the collected
+data. Last, but not least, the pertinent discussion of this work is devided up
+into different sections, each of them analyzing different aspects related to
+data persistence for NetBEAMS, amongo others.
 
-\subsection{Strengths as compared to literature}
+\section{Experiment Setup}
 
-\subsection{Weaknesses as compared to literature}
+The experiment was designed to support different scenarios of a regular use of
+the mongoDB system infrastructure, simulating the activities of data collection
+from the SF-BEAMS network using the NetBEAMS infrastructure. First, the setup
+for the data collection of randomly generated data was implemented, which
+directly uses NetBEAMS already existing classes. Then, the implementation of
+the CRUD scenarios were implemented in different programming languages, from
+the data insertion scenario using the Java Programming Language, and then the
+remaining operations using the Javascript Scripting language \cite{javascript}
+and the Python Programming Language \cite{python}. Similarly, an open-source web
+application, developed in the Ruby Programming language \cite{ruby}, was also
+used. The experiment scripts were writen using Shell Bash Script Language
+\cite{bashshell} and is responsible for orchestrasting the execution of one
+single experiment ``run''.
 
-\section{Data-Centric Storage and Database Shards}
+\subsection{Experiments Artifacts}
 
-Horizontal scalability for data based on specific keys.
+Different artifacts were created to perform the different experiments
+operations. The Create function was developed in Java, without the interversion
+of the DSP or OSGi, isolating only the functionalities of provided by CRUD
+service class DSPMongoCRUDService (Listing \ref{file:dsp-mongo-service}),
+responsible for the connection handling with mongoDB, as well as the insertion
+operation from NetBEAMS. As described in section \ref{sec:dsp-details}, the
+current version of NetBEAMS supports the data collection for the YSI Sonde
+device, among others used for demo purposes such as a software sensor that
+captures the mouse activities over a window. The documentation of both data
+types is seen in section \ref{sec:dsp-payload-implementation}. In this way, the
+DSP Data Persistence, as described in the previous section, supports both data
+types to be persisted.
 
-\subsection{Scalability facing High Volume of Collected Data}
+In order to exercise mongoDB with the same data collected from NetBEAMS, the
+existing Java class TestSondeData (Listing
+\ref{file:random-ysi-data-generator}), was refactored to provide a random
+instance generator for the YSI Data handler implemented in the Java class
+SondeDataType (Listing \ref{file:main-ysi-data-handler}). When the Insert
+function is executed, the created instances are transferred from the JVM to
+the mongoDB, and stored in the collection ``SondeDataContainer''. Specific
+details about the execution of mongoDB are depicted in section
+\ref{sec:mongodb-deployment}. Finally, the implementation of the experiment
+in the Java class DSPMessageToMongoDBExperiment (Listing
+\ref{file:experiment-dsp-java-executor}) was responsible for managing the
+creation and insertion of any workload composed by instances of the class
+SondeDataType in a bulk manner.
 
-Adding new shards as new hardware into the cluster, facing increasing data load
-produced by the sensor network.
+Similarly, the support for the Retrieve, Update and Delete operations against
+the inserted workload were implemented using Javascript, as shown in Listing
+\ref{file:experiment-scenarios}, respectively by the function calls ``find'',
+``update'' and ``remove'' over the collection ``SondeDataContainer''.
+Similarly, in order to implement a customizable export system that converts
+from mongoDB into the format OpenDAP, a simpler prototype script was developed
+in Python, as shown in Listing \ref{file:experiment-export-python}. As a result
+of the execution of this script, the exported artifact has the same format as
+described in Listing \ref{file:rtc-ysi-opendap}.
 
-\subsection{Weakness}
+On the other hand, the execution of the main experiment shell script requires
+the use to provide the size of the workload, as shown in listing
+\ref{cmd:run-persistence-experiment}.
 
-Technological, and skills-based.
+\lstset{label=cmd:run-persistence-experiment,caption=Running main experiment
+shell script}
+\begin{lstlisting}
+marcello@netbeams-dev:~/workspaces/netbeams/versions/v2/persistence$ ./run-persistence-experiment 
+#######  NetBEAMS Experiments - Persistence on MongoDB  ########
+
+Usage: ./run-persistence-experiment X, where X is the size of the workload to be inserted from NetBEAMS YSI Data Handler to mongoDB
+\end{lstlisting}
+
+\subsection{Used Workload}
+\label{sec:workload}
+
+The workload selected for the experiments reflects the current use of the data
+at the SF-BEAMS infrastructure. In this way, the experiment runs for 5
+different times performed to investigate different scenarios to be described.
+In this way, this number is related to the total number of sensor devices
+reported to be in operation, as discussed in section \ref{sec:sfbeams}. In
+summary, the volume of the workload prepared after the experiment setup
+is as follows:
+
+\begin{itemize}
+  \item \textbf{First Round}: 483,840 documents, worth the production of the 
+  YSI Sonde device during one year, at the rate of 1 observation per minute;
+  \item \textbf{Next Rounds}: 483,840 * 4, or 2,419,200 documents, representing
+  5 different YSI devices similar to the first round.
+\end{itemize}
+
+\subsection{System Environment}
+
+Two different system environments were prepared: one for the External Storage
+environment, or Single Server; and another for the Data-Centric Storage, or
+Distributed Server. Since the single server is a regular centralized database
+system, mongoDB just requires the execution of the database process ``mongod'',
+which is started by specifying the directory where the inserted data will be
+managed. On the other hand, nonetheless, the distributed server setup requires
+a different list of processes (see section \ref{sec:mongodb-user-experience}):
+one ``mongod'' process for each mongoDB shard running on the different mongoDB
+cluster, one ``mongod'' process to manage the cluster metadata, and one
+``mongos'' process, which is the main one responsible to orchestrate the
+cluster. In conclusion, the single system environment is synmmetric to
+the External Storage used to persist the collected data, while the clustered one
+is simulates the Data-Centric Storage one.
+
+The simulation of both the single and distributed system used both a Mac OS 10
+Snow Leopard 64bit, with a total of 8GB of physical memory. However, in order
+to simulate different shards in different machines, the use of a VirtualBox
+\cite{virtualization} provided the use of guest operating systems to run and
+act as the mongoDB shards. In this way, two different versions of the Ubuntu
+Linux were used: one 8.04 32bits and another 9.04 64bits (see section
+\ref{sec:experiments-difficulties}).
+
+\subsection{Planned Scenarios}
+\label{sec:exp-scenarios}
+
+In order to verify the feasibility of the solution proposed, all CRUD
+operations were exercised through the implementation of the practical
+scenarios based on the use cases described in section \ref{sec:use-cases}. In
+brief, the use cases were chosen to run with random values such as follows:
+
+\begin{itemize}
+  \item (R0) Insert a volume of collected data from sensors using NetBEAMS
+  service compatible with 1 year of observations;
+  \item (R1) Find all documents whose observations were collected between two
+  different dates, say the days between December 8 and 15 of 2009;
+  \item (R2) Find all documents whose observations were collected from a given
+  known sensor device such as one whose IP address is equals to "192.168.0.102";
+  \item (R3) Find all documents whose observations contained specific values
+  read from the environment, say Salinity equals to 0.01 and the water temperature
+  equals to 46.47;
+  \item (U1) Update all the documents whose observations were collected during
+  the day of December 2nd, 2009, by adding a tag = ``oil spill'';
+  \item (D1) Remove all the documents whose observations were collected on
+  December 7th, 2009.
+  \item (R4) Export the produced data of the week to CSV files, using the same
+  tabular format sequence used in the files distributed by SF-BEAMS using the 
+  the OPEnDAP format, as in Listing \ref{file:rtc-ysi-opendap}.
+\end{itemize}
+
+\section{Measurement Results}
+\label{sec:exp-measurements}
+
+As described in the previous section, the design of the experiments were based
+on the specifications of a data persistence for NetBEAMS, using the mongoDB
+database system. In this way, after setting up the experiment environment,
+the execution of the workload was performed against the mongoDB database using
+its mechanisms used to manipulate the data generated. It uses the programming
+language abstraction to provide access and modifiy the data collected from the
+NetBEAMS component. For this reason, this section details these mechanisms uses
+to implement each of the scenarios defined in the previous section, associating
+numerical response time collected from the logs.
+
+\subsection{Insert Operation}
+
+The first set of measurements collected were the ones regarding the use case
+R1, shown in Listing \ref{file:dsp-mongo-service}, defined as the insertion of
+instances of the collected data from the sensor devices using Java. The use of
+the mongoDB Java driver, which provides the methods to access the database
+system and the collection. In this way, the construction of the keys and
+respective values were done using a Hash-like object ``BasicDBObject'' such as
+the ones shown in the method ``buildKeySegment''. Similarly, the creation of
+database indexes is performed using the same approach, as shown in the Java
+method ``setupMetadataIndexes''. Finally, the insertion of the documents is
+performed on the service method call ``insertPersistentUnitMessageContents'',
+which delimits the bulk insert using the transaction management mechanism over
+the collection. In this way, different insert times were collected from the
+logs of the different experiments as summarized on Table
+\ref{tab:experiment-insert-avarage}.
+
+\begin{table}
+    \begin{center}
+        \begin{tabular}{|p{100pt}|p{100pt}|p{100pt}|}\hline
+           x & \textbf{Linux VirtualBox Server} & \textbf{Mac OS Host Server}\\\hline 
+           \textbf{Single Server} & ~25K docs/sec & ~51K/sec \\\hline
+           \textbf{Clustered Server} & ~17K docs/sec & ~39K/sec \\\hline
+        \end{tabular}
+        \caption{Insertion averages on Virtual and Host in Single or Clustered
+        mongoDB Server}
+    \end{center}
+    \label{tab:experiment-insert-avarage}
+\end{table}
+
+After executing the experiments runs, the claimed disk space for
+the representation of the randomly generated data is separated into two
+different sizes: one related to the amount of data produced by one YSI device,
+and another produced by five different YSI devices, as summarized on Table
+\ref{tab:experiment-insert-diskspace}.
+
+\begin{table}
+    \begin{center}
+        \begin{tabular}{|p{100pt}|p{100pt}|p{100pt}|}\hline
+        x & \textbf{1 YSI} & \textbf{5 YSIs}\\\hline
+        \textbf{Indexed Keys} & 278.33 MB & 1.35 GB \\\hline
+        \end{tabular}
+        \caption{Amount of disk storage used}
+    \end{center}
+    \label{tab:experiment-insert-diskspace}
+\end{table}
+
+\subsection{Retrieve, Update and Delete Operations}
+
+When it comes to the simulation of the use case scenarios, the implementation
+used the mongoDB shell client, which uses Javascript, to implement most of the
+operations for ``retrieve'', ``update'' and ``delete'' shown in Listing
+\ref{file:experiment-scenarios}. On the other hand, the implementation of the
+use case to export the collected data to OPeDAP format used Python, depicted on
+Listing \ref{file:experiment-export-python}. The access pattern to use the
+collected data is ``db.collection\underline{ }name.operation", where
+``collection\underline{ }name'' is the name of the collection used for the
+collected data, and in this case, the name of the DSP component that represents
+the collected data. Therefore, mongoDB uses dynamic object binding for the
+collection object, and executes the CRUD operations as method call with a list
+of the needed parameters. Table \ref{tab:experiment-scenarios-mongo-methods}
+relates method calls used to implement each of the use cases defined, as
+implemented in Listing \ref{file:experiment-scenarios}.
+
+\begin{table}
+    \begin{center}
+        \begin{tabular}{|p{70pt}|p{100pt}|p{250pt}|p{100pt}|}\hline
+        \textbf{Scenario Type} & \textbf{Language} & \textbf{Method Call} & \textbf{Use Case}\\\hline 
+        \textbf{Insert} & Java & db.SondeDataContainer.insert() & R0\\\hline
+        \textbf{Retrive} & Javascript & db.SondeDataContainer.find() & R1, R2, R3\\\hline 
+        \textbf{Update} & Javascript & db.SondeDataContainer.update() & U1\\\hline 
+        \textbf{Delete} & Javascript & db.SondeDataContainer.remove() & D1\\\hline
+        \textbf{Export} & Python & db.SondeDataContainer.find() & R4\\\hline
+        \end{tabular}
+        \caption{mongoDB method calls used for the use cases}
+    \end{center}
+    \label{tab:experiment-scenarios-mongo-methods}
+\end{table}
+
+The execution time for each of the use cases were collected using the logs
+collected during the execution of the experiment script manager. The retrieval
+operations ran in avarage for ~300ms, while the updates took more than 10
+seconds when it involved larger data sets with more than 300 thousand
+documents. Similarly, the delete operation took less
+
+\subsection{User Experience}
+
+Different methods of data access were used after the randomly generated data
+was inserted into mongoDB: using the shell commands mongo and ``mongoexport'', 
+command-line execution developed in Python, through the REST Web Services API,
+and using a web application, written in the Ruby Programming Language
+\ref{ruby}, providing the view over the collected data using a web browser.
+More details about the execution of these commands during the experiments is
+shown in section \ref{sec:mongodb-user-experience}.
+
+First and foremost, the primary method to access to the database is using the
+shell command. Once the user requests an access to the database, he or she can
+issue an operation over the collection of documents using the regular the
+abstraction programming language, as the shown in Listing \ref{cmd:mongo}. The
+CRUD operations are available to the user and can be used at any time to any
+database collection. Aggregation operations such as to count the returned
+resultset is also provided via programming languages abstraction.
+
+In order to export the collected data to other data formats, the  command
+``mongoexport'' and the developed export script written in Python were used.
+The ``mongoexport'', exports the collected data from the database using the
+native support to comma-delimited values. As shon in Listing
+\ref{file:mongodb-export-command}, the actual execution, the remaining ones
+describe how mongo is querying the data from the database. As a result, the
+total number of documents exported is shown in on line 8. In addition to the
+regular command that exports the entire database, the user can provide a query
+string during the execution of the command in order to filter the dataset
+returned and, consequently, the number of documents exported. In contrast to
+the use of the default export capabilities of mongoDB, an Export script written
+in Python, as shown in Listing \ref{file:experiment-export-python}. It uses the
+package ``pymongo'' and creates a comma-delimited file with the same header
+format and lines as shown in the HTTP Response body of Listing
+\ref{file:rtc-ysi-opendap}. The results can be compared with the example of
+Listing \ref{file:experiment-export-python-results}.
+
+While the access to the database using the shell command gives full access to 
+the CRUD operations, the access to the data was also performed using the REST
+Web Services API, as shown in Listing \ref{cmd:mongo-rest-request}. As one can
+see, the IP address and the default port number of the mongoDB REST middleware
+are provided. Then, the REST operation uses the name of the database and
+the collection to correlate with the ones in the database system. In this way,
+querying the collections from mongoDB can be performed using the HTTP Request
+Parameter variables, by adding the prefix ``filter_'' before the name of the
+key of the document. The given example shows the query performed in the
+collected data for the key ``observation.Conductivity'' with the value
+``104.5''. Similarly, the HTTP Request variable ``limit'' is used to reduce
+the number of documents returned by the query. Finally, the HTTP Response from
+mongoDB returns the result using the JSON format, similar to the ones printed
+in the command-shell.
+
+Finally, the access of the collected data was performed using Futon4Mongo,
+an open-source web-based application written in Ruby that provides access to the
+mongoDB database collections and documents. The screenshots of the session used
+to browse through the collection of collected data is shown in Figure
+\ref{fig:view-collections-instance-browser-futondb}.
+
+In order to verify
+
+\section{Discussion}
+
+The implementation of the data persistence for NetBEAMS follows the
+specifications explained in chapter 4. The purpose of sensor network for
+NetBEAMS is to suport data archival from any collected data, and for this
+reason, mongoDB was selected after an empirical analysis described in chapter
+5. In this way, the requirements described in chapter 4 were expressed as
+examples of use cases in section \ref{sec:exp-scenarios}.
+
+All the collected data from sensors are stored in a
+given single server, which characterizes an External Storage device
+
+the data. The process to describe the data is done with prior knowledge of the
+sensor device data properties when choosing the relational data model, because
+of the process of database normalization \cite{db-normalization}, which takes
+place after the the entities have been identified. Consequently, the
+introduction of a new sensor device might potentially require changes on the
+database schema structure conceived before, as the process of database
+normalization must follow.
+
+\subsection{Taxonomic Evaluation}
+
+Any database system covers the purpose of the sensor data for NetBEAMS, which
+is Data Archival, along with the type of storage for the data as an External
+Storage. After selecting mongoDB, the development of the persistence layer met
+the requirements of archiving data after inserting the the randomly generated
+data that represents the data for the YSI sonde. As it is shown in section
+\ref{sec:exp-measurements}, the collected data is archived in files using the
+BSON format \cite{bson}, a binary representation used to store instances of
+the document described in Listing \ref{file:mongodb-ysi-data-format}. Along
+with the purpose of data archival, the use of mongoDB also met the
+specification of an external storage since the data persisted in the file
+system, as well as Data-Centric for the proposal to use mongoDB distributed
+shards.
+
+SF-BEAMS uses the Tabular Data model to describe the collected data from the
+sensor devices. As discussed in chapter 5, the decision to use mongoDB is the
+fact that the literature review does not mention the use of the Key-Value-Pair
+Data model. For this reason, since mongoDB implements a variation of the 
+key-value data model based on the JSON \cite{json} data format, the Data
+Model taxonomy should be updated, as shown in Figure
+\ref{fig:taxonomy-data-model-modified}.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=0.5]{../diagrams/taxonomy-data-model-modified}
+  \caption{Data Model Taxonomy}
+  \label{fig:taxonomy-data-model-modified}
+\end{figure}
+
+Another aspect of the implementation is that provides APIs in different
+languages for the data creation and extraction processes. In this way, the
+implementation of the persistence model captures the instances of any known
+data collected from NetBEAMS and saves it using the document pattern as shown
+in Listing \ref{file:mongodb-ysi-data-format}. One application of a
+schema-less data model is the addition of any new key-value pair to any
+instance of the document, as mongoDB supports update operations to any of the
+collections. It can use tags for the purpose of annotations as described in the
+taxonomy of data description. For example, in case of an oil spill in the San
+Francisco Bay occurs, researchers can update the data set of a given date
+range by adding tags to the relating documents. As a consequence of the use of
+a schema-less data model, these changes do not affect the other data
+instances, since there are no schema changes.
+
+As described in chapter 5, the schema-less data model can provide a better
+support to Data Provenance because the data model gives support to the change
+of records without affecting the others. In this way, the metadata used by the
+implementation clearly describes the three fundamental questions regarding the
+collected data from sensor devices, as described in section
+\ref{sec:keys-definition}. First, the identity of the data is given by the key
+``message\underline{ }id'', in a way that it uniquely identifies the data
+produced by the sensor identified by the key ``sensor.ip\underline{
+}address''. In this way, the message can be tracked back to the sensor that
+produced the data for different purposes such as tracking the data producer.
+Second, time dimensions were used in order to track two different aspects of
+the data collection: the valid time can be used to correlate observations to
+the exact moment it occurred, while the transaction time can be used to verify
+when the data was collected, and make decisions such as deleting the instances
+of data were collected in the previous two days, etc. Last, but not least, the
+use of the key ``sensor.location'' can give the exact location of the data,
+since researchers can be interested in finding data based on a specific
+location. Finally, the most important aspect of the schema-less data model is
+that the observed data is added to the key ``observation'', as any sensor
+device carries its set of attributes of key-value pairs. A good example of the
+applicability of the schema-less data model is the scalability regarding
+changes to the data structure. For instance, if an existing sensor device is
+flushed with a new version of the firmware\footnote{The first set of machine
+instructions to run on the hardware after the application of power stored
+permanently in PROM or ROM or semi-permanently in EPROM}, changing the format
+of data types or adding new values, the following instances of collected data
+may be different to the previous one by simply adding the new additions of
+data. For this reason, the DSP Data Persistence gives administrators the
+ability to choose which fields to persist by declaring it using a bootstrap
+message for the component, as shown in the parameter ``YSI\underline{
+}DESIRED\underline{ }PROPERTIES'' of Listing
+\ref{file:dsp-data-persistence-bootstrap.xml}.
+
+The centralized query processing used by NetBEAMS is a result of the
+infrastructure used by NetBEAMS, as well as the purpose of data and the
+location of the collected data. As a direct result of a schema-less data
+model, the centralized query processing is easily developed using any of the
+available programming languages provided by mongoDB. Considering the fact that
+programming languages are easier abstractions to researchers without computer
+science background \cite{sn-programming-language}, mongoDB's approach to data
+access through API can be seen as an easier task to manage data. As it is shown
+in Listing \ref{file:experiment-query-scenarios}, the scenarios designed to
+evaluate the implementation are written in Javascript, since it is the script
+language by the mongoDB Shell. Therefore, it attends the requirement of not
+using the SQL language as the main query language of the data, giving a better
+data access abstraction similar to programming languages for different users.
+
+One of the most important features of a data persistence for sensor
+networks is the replacement of the SQL by the use of programming
+languages, as they tend to be easier abstraction to users without the
+background in database systems \cite{sn-programming-language}. In this way,
+mongoDB covers the requirements of providing data access to the collected
+sensor data in different languages. For example, the data collected by the DSP
+Data Persistence component is written in Java, while the implementation of the
+experiments scenarios for querying the collected data is given in Javascript. In
+this way, researchers of the RTC may choose one of the enumarated languages
+APIs provided by the mongoDB to access the stored data from NetBEAMS.
+
+Finally, the implementation supported by mongoDB during the experiments was
+following a single dedicated server virtualization using VirtualBox 
+\cite{virtualization}.
+
+\subsection{Infrastructure Evaluation}
+
+Overall, the implementation of the DSP Data Persistence component meets
+the specification of the use cases defined in section \ref{sec:exp-scenarios}.
+The implementation of mongoDB using a single server environment can give
+support to the current workload and infrastructure from the use of NetBEAMS to
+collect data from the sensors devices from SF-BEAMS. From the perspective of
+disk-space utilization and given that aproximaly 1.3 Gigabytes of disk space
+were used to save one year-worth of collected data equivalent to five YSI
+devices. According to the mongoDB documentation \ref{mongodb}, the claimed
+amount of disk space is related to how mongoDB uses a binary representation
+of the JSON documents, which are composed only of strings, together with the
+definition of the indexes also represent in JSON. As a consequence, the size of
+the keys defined for the YSI sonde data defines the amount of disk space used.
+For this reason, different experiments were conducted with different
+definitions of the keys of the documents
+approach when compared the the RTC's OPEnDAP format shown in Listing
+\ref{file:rtc-ysi-opendap}. Then, the second specification used short names
+for the keys. In this way, the amount of storage space utilized by for the
+different workloads define in section \ref{sec:workload} can be summarized as
+follows
+
+The performance to insert data into the database system using the Java driver
+is documented by mongoDB as the fastest approach, providing persistence to
+approximaly 25,091 documents per minute on an indexed database. In contrast,
+when indexes are not used, more then 169,508 documents could be inserted per
+minute. Although the latter approach supports more ducuments than the former,
+the trade off between insertion and selection time is relevant due to the
+activities of retrieval for documents for specific periods of time. Therefore,
+the use of indexes might speed the retrieval time considerably. Considering
+this number of insertions to the database the number of sensors producing data
+and sending to the data sink, mongoDB and NetBEAMS implementation can
+virtually support thousands sensor devices.
+
+Disk utilization is one important characteristic of the selected database
+system. However, the retrieval of the collected data from sensor devices is
+a much more important requirement because most of the utilization of the data
+persistence of a sensor network is regarding the search for the properties on
+the collected data. After analyzing the response times for the scenarios
+described in section \ref{sec:exp-scenarios}, it is clear that the use of
+indexes improves the overall execution of the queries, specially because of the
+amount of data collected over the years. Furthermore, the response times
+described in this scenario is directly related to the use of the External
+Storage strategy. On the other hand, improvements over the response time can be
+achieved through the implementation of a Data-Centric approach on a Distributed
+Database environment, or clusters. In this approach, the inserted data is
+stored in different servers based on a partition key. Given the fact that less
+data is concentrated in each partition, the centralized query mechanism used is
+faster, and consequently improves the performance as described in the Data
+Location taxonomy in chapter 3.
+
+When it comes to re-using the data collected by the DSP Data Persistence,
+different approaches were available. First, the use of the export capabilities
+to a common-delimited values was performed \ref{file:mongodb-export-command},
+taking approximaly 3.3 minutes to export one year worth of data for 5 YSI
+devices without indexing. Second, the use of the client shell gives a
+centralized access to either the single server. However, in order to better
+provide access of data to users without skills in database systems, web
+applications can be built on top of the REST Web Services APIs \cite{http-rest}
+provided by mongoDB. One such example of using the REST API is the adaptation
+of Apache Futon 4 mongoDB, as shown in Figure
+\ref{fig:view-collections-instance-browser-futondb}, where users can navigate
+through the collections of the collected sensor network data categorized by the
+collection. Therefore, RTC staff technicians could take advantage of this tool
+to view and navigate over the collected data from the sensor devices.
+
+\subsection{Use Cases Evaluation}
+
+Supposing an environmental accident such as the San Francisco oil
+  spill happened again \cite{sfbay-oilspill2009}, changing the sampling
+  frequency drastically to different values.
+
+\section{Difficulties Experienced}
+\label{sec:experiments-difficulties}
+As the development of computing system, the inception of the experiments
+described in this chapter had different moments of Software Engineering.
+Particularly speaking about the learning curve, different situations
+contributed to the success of the first experiments. Since the technoogy is
+new, it brought new challenges regarding the usability and user experience.
+Since the technology is relatively new and still under development, many
+different bugs were could be reproduced by random during the execution of the
+experiments, and therefore, sent to the open-source community through use 
+different available communication channels. First, the fastest way to solve an
+issue was through the mongoDB's IRC channel, located at the Freenode server
+(www.freenode.org, channel mongodb). Second, the use of the users group, located
+at http://groups.google.com/group/mongodb-user also helped solving problems.
+
+Physical memory is limited, for this reason, it may not be enough to run the
+experiments for total workloads. In this case, the Java Virtual Memory was
+totally consumed by one of the revisions of the experiments setup. Huge data
+loads require different approach to manage data in memory such as creating
+object pools as temporary cache for objects whose state are similar to others.
+
+While performing the experiments, the 32bit machine reached its limitation of
+2GB of physical data. With know about that, the environment needed to be
+updagred for mongoDB requires a 64-bit operating system when the amount of
+stored data reaches 2 Gigabytes. For this reason, the scenarios with indexes
+used a dedicated virtualized server using Ubuntu Linux 9.04 64 bits with 1 Gb
+of memory.

### 2009-12-07 docs/marcello-dissertation/thesis/chapter9-conclusions-and-future-works.tex
Most of the revisions and updates... Changes to chapters 3, 4 after Kleber's chat online (adding missing taxonomy).
Revisions of print chapters until chapter 5.

Chapter 7 is reviewed until before discussion. 

Next revision should include all the necessary reviews for the
release draft.

git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@676 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 
-\chapter{Conclusions and Future Works}
+\chapter{Conclusions}
 
@@
 types from the sensor devices.
+
+As seen in the state of the art, different approaches can be done in order to
+evaluate data persistence for nsensor networks. T
 
@@
 mongoDB can be used to persist data at its maximum rate of one sample per
-minute. However, performance can be improved by using techniques such as
-database sharding, since it offers a direct implementation capability to
-Data-Centric Storage approach. This approach decreases the amount of time to
-search data since it distributes the data instances across different cluster
-nodes.
-
-There are many different directions for future works. First, the Data-Centric
-implementation performance may be extremely improved by using the technique
-called MapReduce \cite{map-reduce}. This approach allows mongoDB to search
-different machines in parallel whenever the data requested is located in
-different database shard hosts. In addition to that, the use techniques to
-cluster collected data before they are inserted into the database may be
-important as shown in \cite{sn-time-series}. One of the key issues in the
-implemented solution is that any measurement of data is considered to be
-persisted without further analysis.
+minute. However, the evaluation of mongoDB was not used to implement a
+Data-Centric Storage approach, which can decreases the amount of time to search
+data since the process uses parallel search.
 
@@
 any sensor specification. Furthermore, the solution also provides a good
-support to the provenance-related data.
+support to the provenance-related data as the schema-less data model gives the
+ability to .
+
+\chapter{Future Works}
+
+0. Todo list for this work
+1. Database System Problems
+2. Software Engineering Problems
+3. Applications
+ - plot graphs of collected,
+
+There are many different directions for future works. First, the
+implementation of a Data-Centric Storage approach can help the system scale
+horizontaly by using the mongoDB's database shards. In this way, this approach
+can reduce the retrieval time by focusing on centralized selection of data over
+a given database shard. Moreover, the overall search through the collected data
+attributes can be drastically decreased by parallel search over the database
+shards using the MapReduce \cite{map-reduce} tools. At this time of
+development, the MapReduce API offered by mongoDB is on its alpha version, and
+therefore, presenting lots of bugs. Figure
+\ref{fig:future-works-data-centric-map-reduce} shows an architectural idea of
+the deployment of a Data-Centric storage approach. When searching for specific
+document attributes, the process is sent in parallel to each of the database
+shards, and the result is collected by the cluster head. After calculating the
+final result, mongoDB server returns the relevant result to the user.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=0.5]{../diagrams/future-works-data-centric-map-reduce}
+  \caption{Data-Centric approach using MapReduce to search accross different
+  database shards}
+  \label{fig:future-works-data-centric-map-reduce}
+\end{figure}
+
+In addition to infrastructural, the use techniques for scheduling data
+gathering from the SF-BEAMS sensor networks can be used to decrease the data
+load into the system. While the period of times used by the RTC seems
+to cover the basic requirements of data, the data storage may be
+seen as overloaded with repeated data. Since environmental conditions
+does not change drastically over time, the use of data scheduling and
+data clustering approaches can be used. First, considering that SF-BEAMS
+contains a cluster of sensors, the data load could be decreased by using
+techniques such as clustering the data in the data sink before they are
+persisted \cite{sn-time-series, sn-schedule-minimal-aggregation-time}. Since
+the infrastructure of the SF-BEAMS sensor network is a one-hop start design,
+the knowledge about the data can only be achieved in the data sink. In this
+way, one suggestion is the implementation of a DSP Data Clustering component
+that is responsible for the data clustering before they are inserted into the
+database. Similarly, the use of schedulers on the sensors could also be used to
+decrease the amount of data to be sent to the data sink, as seen at the survey
+\cite{sn-scheduling-survey}.
+
+Finally, taking into account the use of the database system to complement
+NetBEAMS execution, the design of event-based applications could add different
+functionalities to the management capabilities of the sensor networks. For
+instance, the YSI sonde collected data carries the information regarding the
+battery life of the device, seen at the collected attribute 
+``observation.Battery''. A new DSP Component responsible for observing these
+attributes could define the threshold of the event of low-battery. With an
+event-based application in mind, events regarding the SF-BEAMS network could be
+better managed through NetBEAMS where scheduled sites visits would depend on
+such events to happen and, thus, avoiding unecessary operational costs.

### 2009-12-07 docs/marcello-dissertation/thesis/preamble.tex
Most of the revisions and updates... Changes to chapters 3, 4 after Kleber's chat online (adding missing taxonomy).
Revisions of print chapters until chapter 5.

Chapter 7 is reviewed until before discussion. 

Next revision should include all the necessary reviews for the
release draft.

git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@676 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 \begin{quote}
-I certify that I have read {\it \mytitle} by \myname, and that in my opinion this work meets the criteria for approving a thesis submitted in partial fulfillment of the requirements for the degree: Master of Science in Computer Science at San Francisco State University.
+I certify that I have read {\it \mytitle} by \myname, and that in my opinion
+this work meets the criteria for approving a thesis submitted in partial
+fulfillment of the requirements for the degree: Master of Science in Computer
+Science at San Francisco State University.
 \end{quote}
@@
 }
-
 \end{center}
@@
 Sensor Networks are becoming more and more important to different science
-communities bacause of what it produces: the raw data for different studies.
+communities bacause of what they produce: the raw data for different studies.
 In order to make use the collected data, researchers may have to dissect the
@@
 properties such as the purpose and location of the observed data, as well as
-how the data is described. In this way, this work first proposes a set of
-taxonomies that can be used to study and classify a given instance of sensor
-network and its produced data. Then, based on the classes of these taxonomies, 
-this work selected a Document-Oriented Database system as a result of an
-empirical evaluation of the requirements for a new persistence layer for 
-NetBEAMS, the case study used in this work. For this reason, the design and
-implementation of a new data persistence component for NetBEAMS show how
-schema-less data models help decreasing the complexity related to the database
-layer, not yet reported by the state of the art. Furthermore, this work also
-proposes the use of Database Shards in order support the implementation of
-Data-Centric sensor networks. 
-}
+how the data is described. For this reason, this work proposes a set of data
+persistence taxonomies based on the state of the art in data persistence for
+sensor networks, which can be used to classify the infrastructure of an
+existing sensor network and the nature of its produced raw data. In order to
+evaluate the proposed taxonomies, this work designs and implements a data
+persistence component for NetBEAMS, a modular sensor network infrastructure
+developed to improve the operation of the SF-BEAMS environmental Sensor
+Network. In this way, based on empirical analysis of the proposed taxonomies
+and the requirements of a persistence layer for the case study, this work
+selected mongoDB, a schema-less document-oriented database system intead of
+the traditional use of relational databases. As a result of the experiments
+conducted with mongoDB, this work concluded that schema-less data models can
+help decreasing the complexity related to the data persistence layer for
+sensor networks with external or data-centric storage mechanism, and whose
+purpose is data archival. Finally, this work proposes different future works
+such a data-centric approach using Database Shards to store the collected
+data and Map Reduce to execute parallel queries over the shards.}
 \vspace*{\fill}

### 2009-12-08 docs/marcello-dissertation/thesis/chapter10-appendix.tex
First Draft of the dissertation peer reviewed with Kleber.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@680 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 
-This section includes the source-code for the DSP Data Persistence. Each
-section details one single main artifact used on the development of the
-component. However, only the main ones will be added into this document, while
-the remainder can be downloaded directly from the Subversion repository.
+This section includes the source-code for the DSP Data Persistence, as well as
+the experiments artifacts used. Each section details one single main artifact
+used on the development of the component. However, only the main ones will be
+added into this document, while the remainder can be downloaded directly from
+the Subversion repository.
 
@@
 Gateway Embedded System, or on the system developed in the Gumstix device. As
-shown in figure \ref{fig:netbeams-software-stack}, the NetBEAMS Gateway Node is
+shown in Figure \ref{fig:netbeams-software-stack}, the NetBEAMS Gateway Node is
 an embedded system that contains a set of layer found in any computer.
@@
   \item \textbf{DSP Platform and other Bundles}: The plug-and-play DSP
-  Components are based on the OSGi bundles capabilities, which can reuse services registered in the OSGi framework.
+  Components are based on the OSGi bundles capabilities, which can reuse
+  services registered in the OSGi framework.
 \end{itemize}
@@
 
-As depicted in figure \ref{fig:netbeams-software-stack}, the DSP Sonde Data
+As depicted in Figure \ref{fig:netbeams-software-stack}, the DSP Sonde Data
 Producer component opens a serial connection with the device through the Java
@@
 \subsection{The DSP Platform Architecture}
-\label{netbeams-architecture}
+\label{sec:netbeams-architecture}
 
@@
 In order to show an example, consider the DSP Platform on the embedded system
-Gumstix in figure \ref{fig:sf-netbeams-node}. It is started by providing a
+Gumstix in Figure \ref{fig:sf-netbeams-node}. It is started by providing a
 deployment configuration artifact config.xml included in the appendix. It lists
@@
 
-These DSP Classes and other can be seen in figure the  UML Class
-Diagram \cite{uml} of figure \ref{fig:DSP-Messages-Classes}. Each DSP Message
+These DSP Classes and other can be seen in Figure the  UML Class
+Diagram \cite{uml} of Figure \ref{fig:DSP-Messages-Classes}. Each DSP Message
 contains an instance of a Header and Body. Whenever a DSP component is ready
@@
 are installed and bootstrapped by the DSP Platform, as the UML State Diagram
-\cite{uml} in figure \ref{fig:DSPPlatform-Install-Usage-State-Diagram}
+\cite{uml} in Figure \ref{fig:DSPPlatform-Install-Usage-State-Diagram}
 described.
@@
 forwarding copies of it to other entities in the system, abstraction in which
-is shown in figure \ref{fig:DSP-Message-Broker-Abstraction}. In this way, the
+is shown in Figure \ref{fig:DSP-Message-Broker-Abstraction}. In this way, the
 DSP Broker receives a DSP Message and makes necessary deliveries to DSP
@@
 in Table \ref{tab:matching-rule}, that takes a DSP Message as an input and
-returns a list of DSP consumers, and the matching configuration artifact provided by the administrator. See listing
-\ref{file:dsp-matcher-config.xml}) for details.
+returns a list of DSP consumers, and the matching configuration artifact
+provided by the administrator. See listing \ref{file:dsp-matcher-config.xml}) for details.
 
@@
 %This
-%process is depicted by the UML Sequence Diagram \cite{uml} in figure \ref{fig:}
+%process is depicted by the UML Sequence Diagram \cite{uml} in Figure \ref{fig:}
 
@@
 
-\subsection{Execution Logs}
-
 \section{mongoDB Deployment}
@@
 
-- starting the db
-- Stopping the Database
+In order to start the database, just type ``mongod --dbpath DIRECTORY/
 
@@
   \centering
-  \includegraphics[scale=0.5]{../diagrams/view-collections-instance-browser-futondb}
+  \includegraphics[scale=0.45]{../diagrams/view-collections-instance-browser-futondb}
   \caption{Viewing a partial list of data using the Futon for CouchDB/MongoDB}
@@
   \centering
-  \includegraphics[scale=0.5]{../diagrams/view-collected-data-list-browser-futondb}
+
+  \includegraphics[scale=0.45]{../diagrams/view-collected-data-list-browser-futondb}
   \caption{Viewing a partial list of data using the Futon for CouchDB/MongoDB}
@@
 By clicking in one of the items, one can view the instance of the document, as
-it is shown in figure \ref{fig:view-collections-instance-browser-futondb}.
+it is shown in Figure \ref{fig:view-collections-instance-browser-futondb}.
 
@@
   \centering
-  \includegraphics[scale=0.5]{../diagrams/view-collected-data-instance-browser-futondb}
-  \caption{Viewing an instance of collected data using the Futon for
-  CouchDB/MongoDB}
+  \includegraphics[scale=0.45]{../diagrams/view-collected-data-instance-browser-futondb} \caption{Viewing an instance of collected data using the Futon for CouchDB/MongoDB}
   \label{fig:view-collected-data-instance-browser-futondb}
@@
 Main Execution Script]{../../../../netbeams/versions/v2/persistence/setup-mongo-shards.js}
+
+\lstinputlisting[language=Bash,label=file:setup-os-shards-step1,caption=Experiment
+Main Execution Script]{../../../../netbeams/versions/v2/persistence/setup-os-shards-step1}
+
+\lstinputlisting[language=Bash,label=file:setup-os-shards-step2,caption=Experiment
+Main Execution Script]{../../../../netbeams/versions/v2/persistence/setup-os-shards-step2}
 
@@
 Export to from mongoDB to OPEnDAP format.]{../../../../netbeams/versions/v2/persistence/export-mongodb-to-sfbeams.csv}
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=0.4]{../diagrams/experiment-mouseactions-dsp}
+  \caption{Experiment showing the mongoDB persisting messages from the DSP
+  Mouse Actions Demo}
+  \label{fig:experiment-mouseactions-dsp}
+\end{figure}

### 2009-12-08 docs/marcello-dissertation/thesis/chapter2-state-of-the-art.tex
First Draft of the dissertation peer reviewed with Kleber.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@680 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 such as snapshot isolation of the current data into separated replicated data.
-In this way, data access from users are not interrupted by the processing of
+Therefore, data access from users are not interrupted by the processing of
 the data that is processed.
+
+All in all, this chapter described a subset of problems and solutions used by
+the sensor networks community. First, the survey in sensor networks was
+described to give a foundation of specific properties of sensor networks. Then,
+different surveys in data persistence in sensor networks were reviewed,
+followed by the data models and query engines. Last, but not least, different
+project described what is important to describe the collected to finally verify
+different properties of data load and management. The following chapter will
+clssify the properties found in this chapter in terms of taxonomies.

### 2009-12-08 docs/marcello-dissertation/thesis/chapter3-taxonomic-classification-of-persistence-storage-for-sensor-networks.tex
First Draft of the dissertation peer reviewed with Kleber.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@680 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 
-\section{The Location of the Sensor Data Taxonomy}
+\section{Location of the Sensor Data Taxonomy}
 
@@
         Purpose of the Sensor Data & Real-time Data Stream, Data Archival \\\hline 
-        The Location of the Sensor Data & Local Storage, External Storage\\\hline
+        Location of the Sensor Data & Local Storage, External Storage\\\hline
         Data Model & Schema-Dependent, Schema-less Models\\\hline 
@@
 the introduction of this dissertation, one the main motivations for this work
-was to provide means of data persistence of NetBEAMS, a case study which does
+was to provide means of data persistence for NetBEAMS, a case study which does
 not contain a data layer. Therefore, the following chapter seeks to describe
-that case study in order to make a section of a given technology for evaluation.
+NetBEAMS and its supporting infrastructure.

### 2009-12-08 docs/marcello-dissertation/thesis/chapter4-netbeams-in-depth-requirements-analysis-and-specification.tex
First Draft of the dissertation peer reviewed with Kleber.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@680 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 
-After presenting NetBEAMS and its infrastructure, the the next step is the
-actual analysis of existing database systems that fits to the requirements
-and characteristics of NetBEAMS. Meanwhile, more details regarding the
-architecture of NetBEAMS, as well as development guidelines, can be reached in
-the appendix of chapter 10.
+After presenting the SF-BEAMS sensor network, NetBEAMS was presented as the
+motivation and major study for this work. In this way, the infrastructure of
+NetBEAMS were described its infrastructure, the next chapter is the actual
+analysis of existing database systems mentioned on the review of the
+literature, that fits to the requirements and characteristics of NetBEAMS.
+Meanwhile, more details regarding the architecture of NetBEAMS, as well as
+development guidelines, can be found in the appendix of chapter 10.

### 2009-12-08 docs/marcello-dissertation/thesis/chapter5-taxonomic-persistence-storage-selection-for-netbeams.tex
First Draft of the dissertation peer reviewed with Kleber.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@680 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 
-\section{Database Systems Contenders}
+\section{Database System ``Contenders''}
 
-This section describes the database systems contenders considered being used
-together with for the implementation of the Data Persistence component for 
-NetBEAMS. As described in the previous chapter, the selection of the
-technologies must be based on the list of functional and non-functional
-requirements specified within the scope this work, defined in the previous
-section.
+This section describes the database systems ``contenders'' considered to be used
+as the persistence technology backend for NetBEAMS. As described in the
+previous chapter, the selection of the technologies must be based on the list
+of functional and non-functional requirements specified within the scope of
+this work.
 
 One of the essential questions related to data persistence is regarding the
-data model and nature of the database, that is, the use of the relational or
-structured models. In addition to that, and given the present time of academic
-research and recent trends distributed computing such as Cloud Computing and
-the Internet \cite{cloud-comp-survey}, the selection of a data persistence for
-sensor networks should maintain a wider spectrum of the types of database 
-systems. Among other intriguing questions, \cite{db-is-rdbs-dommed} reflects
-about the use of alternatives to the relational model for dynamic environments
-such as data persistence for the Web. Similarly, these trends are also related
-to the advances of Cloud Computing and its persistence layer as a solution to
-dynamic systems \cite{cloud-comp-architectures}. For this reason, this work
-also included the overview of alternative data models for the development of
-an experiment that brings data persistence to NetBEAMS. In such a manner, the
-list of the systems considered for assessment is the following:
+data model and breed of the database. That is, the use of the relational or
+any other model to describe data along with its management. Traditionally, the
+relational data model has been used to provide data persistence for different
+types of projects, inclusing sensor networks \cite{sn-dataware-house,
+db-xml-enabled, sn-db-tinydb}. However, tradition does not always get translated
+to efficiency in terms of problem solving. For this reason,
+\cite{db-is-rdbs-dommed} reflects about the use of alternatives to the
+relational model when the data persistence problem in consideration is to
+provide data persistence for dynamic environments as Internet applications. The
+author also suggests that the relational model is hard to maintain in
+ever-changing environments, and thus, proposes the use of the Key-Value-Pair
+(KVP) data model and a different number of technologies. Similarly,
+\cite{cloud-comp-survey} reports similar trends in academic and industrial
+research in distributed computing have used KVP databases in powerful systems
+such as Cloud Computing \cite{cloud-comp-architectures}. For this reason, in
+addition to the different database systems used by projects in the
+literature review, this work also includes the review of a KVP database system
+called mongoDB \cite{mongodb}, since it was among those cited in the surveys.
+In such a manner, the list of the systems considered for assessment is the
+following:
 
@@
   TinyDB is a database system used in many different sensor networks
-  implementation \cite{sn-db-tinydb};
-  \item \textbf{MongoDB}\cite{mongodb}: an open-source key-value database
-  system described in the article that questions the use of the relational
-  model \cite{db-is-rdbs-dommed}, as well as the Cloud Computing trends survey
-  \cite{cloud-comp-architectures};
+  implementation such as \cite{sn-db-tinydb};
+  \item \textbf{MongoDB}\cite{mongodb}: an open-source KVP database
+  system suggested by different surveys to be alternatives to relational
+  databases \cite{db-is-rdbs-dommed, cloud-comp-architectures};
   \item \textbf{DB2}\cite{db2}: in the pursuit of an XML database system, DB2
   is listed as a hybrid database system that offers both the relational and the
-  XML models \cite{db-xml-enabled}. It was selected because NetBEAMS can
-  represent the collected data in the XML format as shown in setion 
+  XML models \cite{db-xml-enabled}. Furthermore, it was selected because
+  NetBEAMS represents serialized objects in XML, as shown in section
   \ref{sec:dsp-message}.
@@
 
-These technologies will be reviewed in contrast to their capabilities to adhere
-to the specifications of the taxonomies defined in chapter 3 and the
+The following sections reviews these technologies' capabilities to adhere to
+the specifications of the taxonomies defined in chapter 3, as well as the
 requirements and the scope for a data persistence of NetBEAMS, as defined in
@@
   \item Data is generated by sensor devices, such as the YSI sonde
-  \cite{YSI-Sonde}, and manually collected by using a laptop to the network
-  sink at the RTC laboratories;
-  \item Upon data reception, the RTC staff index and archive the data for for
+  \cite{YSI-Sonde}, and manually collected by using a laptop to the data sink
+  at the RTC laboratories;
+  \item Upon data reception, the RTC staff index and archive the data for
   distribution using the OPEnDAP format;
-  \item An automated approach to the data collection procedures is done using 
-  NetBEAMS \cite{netbeams2009}, using software components to obtain and send
-  the data to the network sink. A full description of this approach is
-  described in section \ref{netbeams-architecture}.
+  \item An automated approach to the data collection procedures is performed 
+  using NetBEAMS \cite{netbeams2009}, using software components to obtain and
+  send the data to the data sink node. A complete description of this approach
+  is detailed in section \ref{sec:netbeams-architecture}.
 \end{itemize}
@@
 As summarized in the previous section, the nature of data from NetBEAMS through
-SF-BEAMS is used for the purpose of \textbf{Data Archival}. After data is
-collected from the sensors, it needs to be processed and archived for reuse by
-different types of users, ranging from researchers, students, and the public
+SF-BEAMS is used for the purpose of \textbf{Data Archival}. After the
+collected data captured from the sensors, it requires further processing
+in order to be archived for later reuse. Different types of users are expected
+to use the collected data, ranging from researchers, students, and the public 
 over the Internet.
@@
 
-As it is supposed to be, all the contenders are developed to support data
-persistence and management, what is the requirement for the Data Archival
-purpose of the sensor network data.
-
-Basically, all the selected database systems support the purpose of the data.
+Since all of developed database systems were implemented with the intension of
+persisting data in long-term storage devices, every single one in the list
+provides support for data persistence and management. Therefore, they are
+compliant to the purpose of Data Archival for the sensor networks.
 
@@
 
-As shown in the previous section, NetBEAMS's architecture is based on a
-single-hop star infrastructure from the SF-BEAMS sensor network, having all the
-data collected from the sensors being immediately sent to the network sink
-(Section \ref{sec:sn-infrastructure}).
-
-Based on the observations also described in the last chapter, the simplest
-position to save the collected data is through the use of an \textbf{External
-Storage}. In addition to that, the use of a Data-Centric Storage strategy is
-also considered in situations where the data load in the network sink reaches
-its storage capacity. The \textbf{Data-Centric Storage} storage separates the
-location of storage of the data based in any property of the data such as the
-value of a given key. Such approach of data partitioning is provided by
-database system in two different flavors.
-
-The Data-Centric storage separates the location of storage of the data based in
-any property of the data such as the value of a given key, etc. Such approach
-of data partitioning is provided by database system in two different flavors
-\cite{db-partitioning-relational}:
+As shown in the previous chapter, NetBEAMS's architecture is based on sensor
+network whose topology is a single-hop star with a centralized data sink. In
+this way, all the collected data is transmitted to that single location
+\ref{sec:sn-infrastructure}. In the same vein, the simplest location to save
+the collected data is through the use of at least an \textbf{External Storage}.
+Similarly, the use of a Data-Centric Storage strategy is considered in
+situations where the data sink is about to reach its storage capacity. As a
+consequence, the \textbf{Data-Centric Storage} separates the collected data
+into dissimilar locations, based on any property of the data. Therefore,
+there are two different approaches to data partitioningThe two different
+flavors:
 
 \begin{itemize}
-  \item \textbf{Horizontal Partitioning}: the data segmentation is done based
-  on the values of the table rows in different groups allocated to different
-  datasets physically located in different servers. In this way, all the
-  columns defined for a given entity are found on each partition. For instance,
-  horizontal partitioning can be defined as a collection five years worth of
-  historical sensor data being partitioned into five separate servers, where
-  each server has a single year's worth of sensor data;
+  \item \textbf{Horizontal Partitioning}: the data segmentation is implemented
+  based on the values of the table rows. That is, all the values defined by the
+  columns are used in a given table physically located in specific servers.
+  For instance, horizontal partitioning can be defined as a collection of five
+  years worth of historical sensor data, partitioned into five separate
+  servers. In this way, each server contains data related to each of the years;
   \item \textbf{Vertical Partitioning}: data is partitioned vertically in a way
@@
 for any breed of systems since data management is usually offered by most
-database systems out-of-the-box. Furthermore, this approach can be accompanied
-with the use of Distributed Systems approaches such as Replication to help a
-system to scaling in terms of performance.
+database systems out-of-the-box.
 
@@
 The most common problem related to a single External Data Storage is that 
-resources may run into data management problems such as full-disk space or
-disk failures.
+resources may run into data management problems such reaching the disk storage
+capacity or hardware failure.
 
@@
 The benefit of the Data-Centric approach can be directly associated with
-database partitioning mechanisms described earlier. Since data is distributed
-into different locations, this approach decreases the size of datasets when the
-database server performs a query processing in a database table. For this
-reason, this approach helps the database administrators to scale their database
-infrastructure and improve the database performance.
+database partitioning mechanisms described earlier. Since the collected data is
+distributed into different locations, this approach decreases the size of
+datasets when the database server performs a query processing in a database
+table. For this reason, this approach helps the database administrators to
+scale their database infrastructure and improve the database performance.
 
@@
 Database Sharding \cite{db-shard-discussion}, or simply Database partitioning
-\cite{db-table-partition, db-partitioning-relational}. Both former and the
+\cite{db-partitioning-relational}. Both the former and the
 latter are used in regular schema-dependent models or schema-less models,
@@
   stored in a secondary area;
-  \item \textbf{Vertical Partitioning}: each feature of the system is usually
-  represented by a table in the relational model and placed in a different
-  server host. For example, each sensor in a given network can be divided by
-  its device type;
   \item \textbf{Key or Hash Based Partitioning}: the value of a given key
-  property can  be used as the key of the shard, or used as an input to a hash
-  function, whose output determines which database shard location the data is
-  going to  be stored;
-  \item \textbf{Range-Based Partitioning}: after partitioning each feature on
-  a different  server host, different multiple shards, or similar hosts, can be
-  added by using  the range of a given key of the entity of a function. For
-  example, different  sensor data can be distributed by geographic origin of
-  the data.
+  property can  be used as the key of the shard. Moreover, it can be used as 
+  an input to a hash function, whose output determines at which database shard
+  the data needs to be stored.
 \end{itemize}
 
-An example of a Data-Centric storage with different shards
-\cite{db-shard-intro} ca n be see in Figure
-\ref{fig:database-sharding-by-region}, where data is segregated by the
-location of data. Each shard is assumed to be placed in different server
-hosts. Since the data model is denormalized, the organization of data is
-entirely placed in each of these shards in a Horizontal partition of the
-entity. As it is shown, the shards by location can be defined by a shard key.
+An example of a Data-Centric storage with different shards can be see in Figure
+\ref{fig:database-sharding-by-region}, where the collected data is segregated
+by its origin. Each shard is assumed to be placed in different server
+hosts, as well as each of them uses the same ``denormalized'' data models.
 
@@
 The data partitioning is a highly restricted technology, normally available to
-advanced users of distributed database systems. Despite being a capability
-being offered common-off-the-shelf by some popular database systems, this
-strategy requires closer database management. Some of the problems related to
-this strategy are regarding its execution \cite{db-shard-discussion}:
+advanced users of distributed database systems. Despite being offered as a
+common-off-the-shelf functionality by some popular database systems, this
+strategy requires closer database management and configuration. Some of the
+problems related to this strategy are regarding its lifecycle management
+\cite{db-shard-discussion}:
 
 \begin{itemize}
-  \item \textbf{Rebalancing Shards}: If the schema model changes for the
-  shards, the strategy of rebalancing must be done. For instance, when one of
-  the shards out-grows more than the others. As a consequence, the database
-  shards must be rebalanced by transferring data to new locations whenever
-  necessary. This technique is starting to appear as COTS implementations such
-  as mongoDB \cite{mongodb} or MySQL \cite{mysql};
-  \item \textbf{Referential integrity}: any cross-shard queries can represent
-  problems since the application layer is responsible to enforce data
-  integrity. Examples include verifying constraints of foreign keys when the
-  partition is done using Vertical table partitioning, considering that data
-  collections are organized in separate spaces, and physically placed in
-  different shard nodes \cite{db-table-partition}.
+  \item \textbf{Rebalancing Shards}: considering a schema-dependent database
+  system, changes on the schema of one shard requires rebalancing the
+  collected data on each of the shards. As a consequence, the database shards
+  transfers data to new locations whenever necessary. This technique is
+  starting to appear as COTS implementations such as MySQL \cite{mysql} and
+  mongoDB \cite{mongodb};
+  \item \textbf{Referential integrity}: also considering a schema-dependent
+  database system, any cross-shard query may fall into the problem where
+  references do no exist in other shards, since the application layer is
+  responsible for enforcing data integrity. In this way, examples of such
+  include verifying constraints of foreign keys when the partition is done
+  using Vertical table partitioning, considering that data collections are
+  organized in separate spaces, and physically placed in different shard nodes
+  \cite{db-partitioning-relational}.
 \end{itemize}
@@
   \item \textbf{DB2}: Supports External and Data-Centric storage. Only supports
-  vertical table partition \cite{db-db2-partition} for the data-centric storage.
+  vertical table partition for the data-centric storage \cite{db-db2-partition}.
 \end{itemize}
@@
 All the selected database systems support the External Storage method. However,
-they all differ with the support to Data-Centric Storage support.
+they all differ with the support to Data-Centric Storage support. Therefore,
+MySQL and mongoDB are promising choices, with the difference of the use of the
+data model.
 
-\section{Analysis of the Data Model Taxonomy}
+\section{Analysis of the Data Model}
 
@@
 not fit to solve the problem. In fact, the main users of sensor networks may
-not hold any expertise in database systems or data modeling, given they come
-from different science areas. Taking NetBEAMS as an example, we see a Sensor
-Network managed by Marine Biologists without expertise in Data Management,
-Modeling systems, and for this reason, one of the requirements of the system is
-trying the use of schema-less approaches. The \textbf{Key-Value-Pair Data
-Model} or the \textbf{Document-Oriented Data Model} seems to be the simplest
-choices of the models.
+not possess any expertise in database systems or data modeling, based on the
+fact that they are from different science areas. \cite{sn-programming-language}
+suggested the use of programming languages abstraction instead of regular
+database systems for researchers. For instance, SF-BEAMS is managed by Marine
+Biologists without expertise in Data Management, Modeling systems, and for
+this reason, one of the requirements of the system is trying the use of
+schema-less approaches. The \textbf{Key-Value-Pair Data Model} or the
+\textbf{Document-Oriented Data Model} seems to be the simplest choices of the
+models.
 
@@
 \begin{itemize}
-  \item Constant data schema changes require constant database normalization
-  process, changes to structure, database management, etc. Schema-Dependent 
-  databases always require refactoring of the schema to accommodate changes to
-  the entites properties;
-  \item The approach of Relational Model might does not fit the nature of
-  collected data from wired or wireless sensor devices, since they tend to
-  change over time \cite{db-is-rdbs-dommed}. They usually does not give
-  support time-series data nor provenance well \cite{sn-provenance};
+  \item Data schema that constantly changes require constant database
+  normalization processes, changes to structure, database management, and all
+  the code that uses the data. Schema-Dependent databases always require
+  refactoring of the schema to accommodate changes to the entites' properties;
+  \item The approach of Relational Model might not fit the nature of the
+  collected data from sensor devices, since they tend to change over time. 
+  Furthermore, the support to time-series or provenance are not better
+  supported \cite{sn-provenance};
   \item Some research have suggested changes to the SQL clauses to better
@@
 
-Other projects have been using the XML data models for persistence. It falls
+Other projects have used the XML data models for persistence. It falls
 into the same category as the Relational Model using XML documents \cite{xml},
-being complaint to an XML Schema \cite{xml-schema}. This model uses XPath
-\cite{xml-xpath} technology for querying documents, although some hybrid
-technologies may still use SQL \cite{sql} for that matter \cite{db2}. In this
-way, a materialized version of the data from the defined previous schema can be
-seen in Figure \ref{fig:persistence-example-relational}.
+since XML documents must fulfill the specifications defined by an XML Schema
+\cite{xml-schema}. This model uses XPath \cite{xml-xpath} technology for
+querying documents, although some hybrid \cite{db2} technologies may still use
+SQL \cite{sql} for that matter. In this way, a materialized version of the
+data from the defined previous schema can be seen in Figure
+\ref{fig:persistence-example-relational}.
 
@@
 Although this data model has been used in different types of applications,
-developers have tried to use the Relational Model to model an application that
+developers have tried to use the Relational Model to design an application that
 could prevent any database change through the use of a technique that relates a
-key to a value, called Key-Value data model. With the emergence of Internet
-applications, the necessity of such a model that does not require constant
-refactoring led developers to propose such manner using the relational model as
-documented in technical articles such as \cite{db-kvp-in-relational01} and
-\cite{db-kvp-in-relational02}. Based on these articles, consider the
-relational model depicted in Figure \ref{fig:KVP-on-Relational-Model} as a
-data model for persisting data to our case study using the key-value strategy.
-The problem with it is that key repetition occurs, since the nature of
-time-series data requires the use of a timestamp key for each property of the
-sensor. Therefore, this strategy is not well-suited to provide persistence
-data from collected sensor networks.
+key to a value, called Key-Value data model. As mentioned earlier, with the
+emergence of Internet applications, the necessity of such a model that does
+not require constant refactoring led developers to propose such manner using
+the relational model as documented in technical articles such as
+\cite{db-kvp-in-relational01} and \cite{db-kvp-in-relational02}. Based on
+these articles, consider the relational model depicted in Figure
+\ref{fig:KVP-on-Relational-Model} as a data model for persisting data to our
+case study using the key-value strategy. It is clear that this approach
+contains problems: repetition occurs, since the nature of
+time-series data requires the use of a timestamp key for each property of the sensor. Therefore, this strategy is not
+well-suited to provide persistence data from collected sensor networks.
 
@@
 of powerful database servers that can be organized in the context of
-infrastructure and how to represent models. An powerful up-coming approach is
+infrastructure and how to represent models. A powerful up-coming approach is
 the use of database systems that do not use structured language for querying
@@
 where data is modeled using structured documents. However, these documents
-have a dynamic structure that can be freely described without the use of a
+have a dynamic structure, freely described without the use of a
 schema validation mechanism as used in XML Schema. One example of such
-approach is the use of the JSON data format \cite{json} in the implementation
-of database systems which are becoming popular with the new trend called Cloud
-Computing \cite{cloud-comp-architectures}. Databases implementing this
-strategy is the mongoDB .
+approach is the use of the JSON data format \cite{json}. mongoDB implements the
+following:
 
@@
 For instance, all data needed for the YSI sonde data, as well as all necessary
-provenance data that describe the data, depicted in section
+provenance data that describes the data, depicted in section
 \ref{sec:sn-data-description}, are represented by means of key-value pairs
@@
       values, which have the same schema;
-      \item Data Model defined in advance with a schema, which contains
+      \item Data Model is defined in advance with a schema, which contains
       relationships and constraints to enforce data integrity;
@@
       duplication, as well as establishing data associations through table
-      relationships;
+      relationships \cite{db-normalization};
       \item Data Provenance can be provided through data types such as
@@
 \begin{itemize}
-  \item \textbf{MySQL}: supports the Relational Model;
-  \item \textbf{TinyDB}: supports the Relational Model; 
-  \item \textbf{MongoDB}: supports Document-Oriented Model;
-  \item \textbf{IBM DB2}: supports the Relational Model and the XML model.
+  \item \textbf{MySQL}: supports the Relational Data Model;
+  \item \textbf{TinyDB}: supports the Relational Data Model; 
+  \item \textbf{MongoDB}: supports Key-Value Pair Data Model;
+  \item \textbf{IBM DB2}: supports the Relational Model and the Structured
+  Data models.
 \end{itemize}
 
-\section{Analysis of the Query Processing Mechanism Taxonomy}
+\section{Analysis of the Data Description}
+
+\textbf{Data Description} is the methods used to describe the collected data
+from sensors, specially those produced by real-time data streams. Since those
+type of data does not include descriptions of its values, data provenance
+offers guidelines to improve the description of the collected data.
+
+An additiona to data provenance, annotations is another technique used to
+better describe the collected data from sensor devices. Its main purposes is to
+provide means of identifying observations, creating indexes based on simple
+keywords. For an environmental sensor networks, this is a valuable type of value
+used to better manage the collected data.
+
+\subsection{Advantages of the Provenance-Aware Data}
+
+The injection of provenance data into the collected data enriches description
+of a given event, and is always useful to determine different aspects of the
+data. As mentioned in chapter 3, data provenance helps identifying the nature
+of the data by the identification of the properties of a given sensor device,
+for example. Similarly, provenance adds spatio-time data in order to correlate
+the observed data with time a given location and time of occurrence.
+
+\subsection{Disadvantages of the Provenance-Aware Data}
+
+Additional data being inserted together with the raw data collected from the
+sensor devices. This additional data may contribute with any decreasing number
+of disk storage for the variable defined. Therefore, the decreasing amount
+of disk space may be one the reasons to not use annotations.
+
+\subsection{Advantages of the Annotations}
+
+Annotations are used to describe the collect data even more, since the
+additional ``notes'' of the data are being stored together with the data. For
+example, the use of tags to a given entity may help identifying and different
+observations regarding the data collected from sensor devices.
+
+\subsection{Disadvantages of the Annotations}
+
+This type of data does not make part of the collected data, nor provenance
+information.
+
+\subsection{Technology Analysis}
+
+All the database systems technology supports the use of Data Provenance or
+Annotations. The Key-Value Pair data model has the advantage of not maintaining
+a schema design, and therefore, can add any type of data, and therefore,
+mongoDB is the lead in this taxonomy.
+
+\section{Analysis of the Query Processing Mechanism}
 
@@
 
-All the database systems technology support a centralized query system.
+All the database systems technology support a centralized query system. In
+addition to that, they can be distributed into different distributed nodes when
+used in a partioned environement.
 
-\section{Analysis of the System Organization Taxonomy}
+\section{Analysis of the System Organization}
 
@@
 Each of the databases were scored according to its support to the different
-types of taxonomies, and randomly selected by using the literature review's
-list, as well as the Internet. Since mongoDB is based on a schema-less database
-system, has not been reported in use by the sensor network community, and
-supports most of the taxonomies defined in the previous chapter, it was
-selected for the implementation of the DSP Data Persistence component for
-NetBEAMS.
+types of taxonomies, as well as a different set of requirements. As defined
+in the scope of the database system, the database that can provide a
+programming language abstracton to its data access would be selected.
+Moreover, the database technology should offer the support to schema-less
+databases. As a result, mongoDB was selected because of different reasons.
+First, it is a document-oriented database system, which does require the use of
+SQL languages. Similarly, it had been reported in two different surveys as
+potential database systems. Finally, there were no references in the literature
+regarding the use o KVP databases in sensor networks, and for this reason, this
+proposal was performed, supporting most of the taxonomies defined in the
+previous chapter, it was selected for the implementation of the DSP Data
+Persistence component for NetBEAMS.
 
 \begin{table}
-    \label{tab:ysi-data-distribution}
+    \label{tab:technology-selection}
     \begin{center}
-        \begin{tabular}{|c|c|c|c|c|c|c|}\hline 
+        \begin{tabular}[!h]{|c|c|c|c|c|c|c|}\hline 
         \textbf{Database} & \textbf{MySQL} & \textbf{TinyDB} & \textbf{MongoDB} & \textbf{IBM DB2}\\\hline
@@
         Partition & + & - & + & +\\\hline 
-        Hot Backup & + & - & + & +\\\hline 
         Export Capability & + & - & + & -\\\hline 
-        Programming Lang. & + & + & + & +\\\hline
-        Script Lang. & + & - & + & +\\\hline
+        Programming Language. & + & + & + & +\\\hline
+        Script Language. & + & - & + & +\\\hline
         Open-Source & + & + & + & -\\\hline
         \end{tabular}
+        \caption{Technologies Selection Grading}
     \end{center}
 \end{table}
+
+To conclude, this chapter evaluated different database systems selected from
+the literature reviews of chapter 2. After presenting the database system
+``contenders'', the analysis of each of the taxonomies were conducted for each
+of the technologies, culminating with the technology selection comparison. In
+this way, the following chapter describes the design of the solution using
+mongoDB as the primary solution for the backend.

### 2009-12-08 docs/marcello-dissertation/thesis/chapter6-data-persistence-component-design-and-architecture.tex
First Draft of the dissertation peer reviewed with Kleber.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@680 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 
-In order to provide information to users, the RTC scientists use OPeNDAP
-\cite{opendap} server to distribute collected data over the Internet. As
-discussed in the previous chapter, this work proposes a data persistence
-application of a Key-Value database system in order to provide an alternative
-method for data persistence in sensor networks, taking into account the
-taxonomies discussed earlier in chapter \ref{chap:taxonomies}.
+In order to provide information to users, the RTC scientists use the
+\emph{OPeNDAP} \cite{opendap} server to distribute collected data com the
+\textit{SF-BEAMS} sensor network over the Internet. As discussed in the previous
+chapter, the database system \emph{mongoDB} was selected for the evaluation of a
+backend system for NetBEAMS. Unlike other projects reviewed in the literature, mongoDB has
+never been used as a selected database for data collected from sensor devices.
+In this way, the design of the proposed solution follows the requirements and
+specifications detailed in chapter 4. As described before, NetBEAMS provides a
+component-based infrastructure that automates the operation of SF-BEAMS sensor
+network. The proposed solution presented in this chapter is based on the
+NetBEAM's DSP Platform architecture, as described in chapter
+\ref{chap:netbeams-overview}. The following sections describe the scenarios for
+persistence, analyze the requirements captured from the RTC staff, and describe
+the design and architecture of the proposed solution.
 
-Considering our case study being NetBEAMS, this chapter describes the design
-and architecture component designed specifically for the Data Sensor Platform,
-DSP, in order to provide a persistence layer for NetBEAMS. It is vital to note
-that the technology chosen in the last chapter can be adapted by any
-alternative that uses the data model and infrastructure options presented.
-
-The proposed DSP Data Persistence component for the DSP
-Platform was implemented by using the same approaches described in chapter
-\ref{chap:netbeams-overview}. The following sections describe the scenarios for
-persistence, analyze the requirements from RTC, and describe the design and
-architecture of the solution.
+Technically speaking, this chapter includes detailed UML diagrams regarding the
+different point of views of the system, following the conventions defined at
+\cite{uml}.
 
@@
 As described in section \ref{chap:netbeams-overview}, the DSP Platform was
-designed to support an environmental sensor network based in San Francisco Bay,
-composed by different types of sensors. The process of extracting the
-measurements from sensors can be summarized following the UML Business Process
-Diagram \cite{uml} depicted in figure
-\ref{fig:dsp-persistence-business-process}. It is important to note that this
-diagram details remote communication between sensor devices and a centralized
-server, as part of the main scenario to solve the problem of data persistence.
+designed to support SF-BEAMS, an environmental sensor network based in San
+Francisco Bay, composed by different types of sensors. As shown in in Figure
+\ref{fig:dsp-persistence-business-process}, the event of data collection is
+started by the DSP Component deployed on each Gumstix system based on the
+sampling rate frequency value. When the DSP Client component finishes the data
+extraction process, a new process starts at the client side. First, each of the
+co-located DSP Client processses the collected measurements and transforms
+them into the alleged DSP Message, as shown in section \ref{sec:dsp-message}, 
+emboding the actual collected data from the sensor devices. In this way, an 
+instance of a DSP Measurement Message is added into the local temporary
+outbound queue that waits to be activated. Then, as an asynchronous data
+transmission event occurs within a specified rate, transferring all the queued
+messages to be processed by a DSP Server host. This server host is centralized
+and recives all the DSP Messages, to be forward to different messages.
 
@@
 
-At the event of data collection, the DSP Client receives measurements collected
-from each of the sensor devices. Each of the co-located DSP Client process the
-collected measurements and transform them into the alleged DSP Message
-(section \ref{sec:dsp-message}), which embodies the actual data values. After
-proper data processing, transformation and categorization, the DSP Message is
-added into temporary outbound queue. Then, as an asynchronous data transmission
-event occurs with a specified rate, transferring all the queued messages to be
-processed by a DSP Server host. As image
-\ref{fig:dsp-persistence-business-process}, the server is centralized and able
-to concurrently process messages.
-
-This work proposes the addition of a database system that is responsible for
-storing the gathered measurement data from sensor nodes of the SF-BEAMS through
-the use of NetBEAMS. Also in figure \ref{fig:dsp-persistence-business-process},
-the processed measurement data from sensors 1 through N must be collected in
-the same style and categorized as necessary. In this way, end users such as
-Biologists or Programmers can access the collected measurements directly
-from the Database. Furthermore, the same collected data can be exported to
-different formats as necessary.
-
-As a matter of fact, the data measurements collected from the sensor devices
-are basic properties with keys and associated values as described in section 
-\ref{sec:sfbeams}, after data has been described. According to the descriptions
-of the DSP Platform and the DSP Messages structure, the data is represented in
-programming language formats such as Java \cite{java} or in a serialized
-version in XML \cite{xml}. Similarly, the collected data contains measurements
-that uniquely identifies them, together with other properties such as
-timestamps of data generation and collected and the IP address of the DSP
-Client that produced the data. Therefore, these are essential properties for
-the proposed solution of this work.
+Once the DSP Server processes the received DSP Message, it uses the server
+configuration to transmit the collected data to the Measurements
+Database. As a matter of fact, the collected data from sensor devices are basic
+properties with keys and associated numerical values, described in section
+\ref{sec:sfbeams}. According to the descriptions of the DSP Platform and the
+DSP Messages structure, the collected data from the sensors is parsed and
+mapped to a POJO\footnote{Plain Old Java Object} or it can be serialized in
+XML \cite{xml}. Similarly, the collected data contains attributes that
+uniquely identifies them, along with provenance metadata as spatio-temporal
+attributes.
 
@@
 
-By analyzing the main scenario of collected data in the previous section, as
-well as by collected requirements from the RTC staff, this section lists the
-main use cases that the system and the primary users require.
+After analyzing the main data collection scenario and reviewing the
+requirements defined by the RTC staff, this section enumerates the main
+functions that the proposed solution must perform.
 
@@
 
-As part of the user stories related to data access, figure
-\ref{fig:DSP-Data-Persistence-UseCases-Diagram-Users} describes the use cases
-that the end users can perform with the system.
+The primary goal of system to be implemented is related to the collected data
+reuse. That is, allow users to have easy access to the collected data from
+\emph{NetBEAMS}. The main motivations of the proposed system is that it must
+give different types of users access to the collected data without requiring
+the use of SQL. Rather, it must provide a programming language abstraction
+that might make the access to those data easier \cite{sn-programming-language}.
+Furthermore, it must use the selected technology, mongoDB, using the
+properties defined by the system such as the use of a schema-less database
+systen. In this way, the designed functions for the solution, as depicted in
+the UML Use Cases Diagram of
+Figure \ref{fig:DSP-Data-Persistence-UseCases-Diagram-Users}, are summarized as
+follows:
 
@@
   start and end date;
+  \item \textbf{Annotate Existing Measurements}: users of the collected data
+  must be able to annotate any existing measurements, that is, they must be
+  able to add more metadata such ``tags'';
+  \item \textbf{Export Data to Spreadsheets}: users of the collected data must
+  be able to export the collected data to spreadsheets.
+  \item \textbf{Export Data to OPEnDAP format}: users of the collected data
+  must be able to export the collected data to the OPEnDAP format;
   \item \textbf{Access the data through API, Programming Languages}: users of the
@@
   programming and scripting languages of their preference;
-  \item \textbf{Export Data to OPEnDAP format}: users of the collected data
-  must be able to export the collected data to the OPEnDAP format;
-  \item \textbf{Export Data to Spreadsheets}: users of the collected data must
-  be able to export the collected data to spreadsheets.
+  \item \textbf{Remove Undesired Measurements}: users of
+  the collected data must be able remove any collected data.
+\end{itemize}
+
+On the other hand, the functions performed by the solution must be designed
+conformed the specifications of the Data Sensor Platform (DSP) architecture,
+described in section \ref{sec:netbeams-architecture}. In this way, the
+solution must be a Data Producer, implementing the following functionalities,
+as shown in Figure \ref{fig:DSP-Data-Persistence-UseCases-Diagram-System}:
+
+\begin{figure}[!b]
+  \centering
+  \includegraphics[scale=0.5]{../diagrams/DSP-Data-Persistence-UseCases-Diagram-System}
+  \caption{UML Use Case diagram for the Persistence Component}
+  \label{fig:DSP-Data-Persistence-UseCases-Diagram-System}
+\end{figure}
+
+\begin{itemize}
+  \item \textbf{Update Component Properties}: as described by the DSP Component
+  architecture, the new DSP component must be able to have its properties
+  updated. These properties must include the rate of how long it takes to the
+  component flush messages;
+  \item \textbf{Update Database Properties}: the component must be able to
+  update the database settings such as IP Address and Port of the service;
+  \item \textbf{Collect DSP Measurement Message}: as the component is designed
+  to be a Data Consumer, it must be able to receive and temporarily retain the
+  collected messages from the sensors;
+  \item \textbf{Categorize Measurement Message}: as the DSP Message is
+  delivered to the new component, it must categorize the message by its
+  properties such as transaction time, originating DSP Message ID, etc;
+  \item \textbf{Save Messages into the Database}: the main goal of the new
+  component is to send the payload of the DSP Messages to the database in use,
+  in this case, mongoDB.
 \end{itemize}
@@
 
-Different constraints are required to execute the functional requirements
-described in the previous section. They are listed as non-functional
-requirements as follows:
+Different constraints were required to execute the functional requirements
+described in the previous section, that is, non-functional requirements that
+the system must follow:
 
 \begin{itemize}
-  \item The DSP Data Persistence component must maintain the received DSP
-  Measurement Messages in-memory for a specific period of time in order to
-  decrease the write load on the database;
-  \item The Persistence storage system must be able to cope with dozens of DSP
-  Messages received from different sensors;
-  \item The Persistence storage system must be able to save measurements quickly;
+  \item The solution must maintain the received DSP Measurement Messages
+  in-memory for a specific period of time in order to decrease the write load
+  on the database (flusher rate);
+  \item The Persistence storage system must be able to cope with the Data
+  Volume produced by SF-BEAMS in a daily bases;
   \item The Persistence storage system must be able to scale without too much
   architectural changes, or schema changes;
-  \item The Persistence storage system must not impose the use of
-  database languages such as SQL \cite{sql} or XML XPath \cite{xml-xpath}
-  related to the data being persisted. Instead, data must be handled through API
-  calls, as better abstractions for non-computer science researchers
-  \cite{sn-programming-language}.
+  \item The Persistence storage system must be able to provide access to the
+  persisted data using an abstraction similar to programming languages, as
+  the authors of \cite{sn-programming-language} proposed;
+  \item The Persistence storage must be able to 
 \end{itemize}
@@
 
-As described in the previous chapter, the data model chosen to support
-persistence of collected data from sensor networks is the Key-Value-Pair Data
-Model \cite{db-kvp}. Beyond the data collected from the sensor devices, Data
-Provenance techniques must also be considered to better describe the collected
-data \cite{sn-provenance}. In general, the data model must include the
-following metadata besed on Data Provenance:
+MongoDB is defined as a Document-Oriented Database system, that uses the
+abstraction of the Key-Value-Pair data model. In this way, mongoDB does not
+require the creation of a schema. However, the definition of how to describe
+the collected data must be provided, as they define the keys to the system.
+Similarly, in order to describe the keys, the specifications details the use of
+Data Provenance as guidelines to the description of the collected data.
+Therefore, the key-value pairs of any instance must include the following
+properties:
 
@@
 After analyzing the DSP Platform and the structure of the DSP Messages in
-chapter \ref{netbeams-architecture}, the following set of keys are expected to
-be part of the design of the data model. Any key represented by time or
-location are related to the provenance support discussed in section
-\ref{sec:sn-data-description}:
+chapter \ref{sec:netbeams-architecture}, the following set of keys were designed
+as be part of the data model, defined as the ``key segment'':
 
 \begin{itemize}
-  \item \textbf{sensor.ip\underline{ }address}: it is extracted
-  from the DSP Message Produce and identifies which sensor generated the sampling;
-  \item \textbf{sensor.location.latitude, longitude}: the latitude and
-  longitude GPS coordinates of the position of the sensor device, captured from
-  references of the device. This is currently provided as a parameter in the
-  bootstrap message for the component;
-  \item \textbf{message\underline{ }id}: it's extracted from each of the
-  messages contained in the DSP Message Container;
-  \item \textbf{time.transaction}: it is extracted from the DSP
-  message container creation time and it is used to identify when the
+  \item \textbf{message\underline{ }id}: Defines the DSP Message IP that
+  collected the measurement message. This can be used to track which component generated, etc;
+  \item \textbf{sensor}: Defines the properties of the sensor device;
+    \begin{itemize}[label=\textbullet]
+        \item \textbf{ip\underline{ }address} as the IP Address of the sensor
+        device that produced the data;
+        \item \textbf{location}: defines the location of the sensor device;
+            \begin{itemize}[label=\textbullet]
+                \item \textbf{latitude} the GPS latitude coordinate of the
+                position of the sensor device, captured from references of the device;
+                \item \textbf{longitude}: the GPS longitude coordinate of the
+                position of the sensor device.
+            \end{itemize}
+    \end{itemize}
+  \item \textbf{time}: identifies the time dimensions about the collected data;
+    \begin{itemize}[label=\textbullet]
+      \item \textbf{valid}: it is extracted from the container's date and time
+      of creation, which was collected during the sensor device's reading;
+      \item \textbf{transaction}: it is extracted from the DSP message
+         container creation time and it is used to identify when the
          transaction occurred;
-  \item \textbf{time.fact}: it is extracted from the
-  SondeDataContainer's date and time, and identifies the time in which the
-  collected data occurred.
+    \end{itemize} 
 \end{itemize}
 
-It is important to note that this set of metadata must occur in every
-collected data from any sensor device. However, on the contrary of the
-metadata, a different set of properties will occur to describe the sensed
-properties of a given sensor, and for this reason, the following structure is
-defined to accommodate such data:
+While the ``Key Segment'' is the set of keys formed by the Data Provenance
+data, the ``Value Segment'' stores the observations from the sensor
+devices. In this sensor, the observation key is defined as follows::
 
 \begin{itemize}
-  \item \textbf{observation}: this key defines the values of the document, and
-  will have every different property of the sensor. For instance, the data key
-  for the properties of the YSI device will contain the set of properties 
-  defined for the device, such as battery, temperature, salinity, etc.
+  \item \textbf{observation}: this key stores the values of the document, and
+  will have every different property of a given sensor device. For instance, the
+  data key for the properties of the YSI device will contain the set of properties 
+  defined for the device, such as battery, temperature, salinity, among other,
+  as shown in Listing \ref{file:mongodb-ysi-data-format}.
 \end{itemize}
@@
 to the DSP Platform can be achieved by adding Data Consumer (DC) as described
-in section \ref{netbeams-architecture}. Based on the Requirements
+in section \ref{sec:netbeams-architecture}. Based on the Requirements
 specification described earlier, this component must be responsible for
@@
 an architectural view of the DSP Server node with the loosely-coupled
-components, highlighting the inclusion of the DSP Data Persistence.
+components, highlighting the inclusion of a new DSP Component called ``DSP Data
+Persistence''.
 
@@
 drivers are also added without any changes to the existing DSP infrastructure.
-First, as DSP Messages are delivered to the server through the DSP Wire
-Transport Server, as shown by flow 1, the DSP Platform is responsible to
-forward a copy of the Measurement messages directly to the DSP Data
-Persistence as shown by flow 2. Then, the sole responsibility of the DSP Data
-Persistence is to transfer the received measurement messages to the database
-system by using the connection drivers, as depicted by flow 3.
+In this way, a matching rule must be provided. First, as DSP Messages are
+delivered to the server through the DSP Wire Transport Server, as shown by
+flow 1, the DSP Platform is responsible to forward a copy of the Measurement
+messages directly to the DSP Data Persistence as shown by flow 2. Then, the
+sole responsibility of the DSP Data Persistence is to transfer the received
+measurement messages to the database system by using the connection drivers,
+as depicted by flow 3.
 
@@
 dependency among 3 different major Java packages: it depends on 2 existing
-packages, namely the OSGi Framework and the DSP Platform, as well as and an
+packages, namely the OSGi Framework and the DSP Platform, as well as an
 additional package for the database access driver API, as shown in the UML
-Package Dependency Diagram \cite{uml} of Figure
+Package Dependency Diagram of Figure
 \ref{fig:DSP-Data-Persistence-Packages-Dependency}.
@@
 figure \ref{fig:DSP-DataPersistence-Activator-Class-Diagram} shows the UML
-Class Diagram \cite{uml} implemented for the component.
+Class Diagram implemented for the component.
 
@@
 Persistence extends from the DSP Platform class DSPComponent. For this reason,
-this class inherits all the behavior described on \ref{netbeams-architecture}
+this class inherits all the behavior described on \ref{sec:netbeams-architecture}
 and can be in the roles of Data Consumer and a Data Producer.
@@
 component instance. In this way, following the specifications of section
-\ref{netbeams-architecture}, the DSP Data Persistence Platform can be
-initialized as described as it is described by the UML Sequence Diagram of
-Figure
+\ref{sec:netbeams-architecture}, the DSP Data Persistence Platform can be
+initialized as described by the UML Sequence Diagram of Figure
 \ref{fig:From-OSGi-Framework-to-DSP-Data-PersistenceActivator-Sequence-Diagram}:
@@
   \item After being selected based on the configuration priority, the DSP
-  Bundle artifact is identified and installed, by creating an instance of the
+  Bundle artifact is identified and installed by creating an instance of the
   class DSP DataPersistence Activator and making a call to the method start();
@@
 the DSP Component Activator and the DSP Component registration are also shown
-in figure \ref{fig:From-OSGi-Framework-to-DSP-Data-PersistenceActivator-Sequence-Diagram}.
+in Figure \ref{fig:From-OSGi-Framework-to-DSP-Data-PersistenceActivator-Sequence-Diagram}.
 
@@
 
-Although the DSP Data Persistence component has been instantiated as of image
-\ref{fig:From-OSGi-Framework-to-DSP-Data-PersistenceActivator-Sequence-Diagram}, 
-the process of delivering messages to the DSP Data Persistence has yet to be
+Although the DSP Data Persistence component has been instantiated as of Figure
+\ref{fig:From-OSGi-Framework-to-DSP-Data-PersistenceActivator-Sequence-Diagram}, the
+process of delivering messages to the DSP Data Persistence has yet to be
 completed. As defined by the requirements, the DSP Component must execute its
@@
 flushed into the Database with the class Transient Persistence Layer. Figure
-\ref{fig:DSP-DataPersistence-Flusher-Classes} describes the UML Class
-Diagram \cite{uml} for the main classes of this functionality.
+\ref{fig:DSP-DataPersistence-Flusher-Classes} describes the UML Class Diagram
+for the main classes of this functionality.
 
@@
 The DSP Data Flusher is a thread that is designed to be in two different
-states as depicted in figure
+states as depicted in Figure
 \ref{fig:DSP-DataPersistence-Flusher-State-Diagram}: RUNNABLE and
@@
 contacting the Transient Persistence Layer in order to verify if there are any
-message in the outbound queue ready to be transmitted sent to the Database
+message in the outbound queue ready to be transmitted to the Database
 service, while the DSP Data flusher will be waiting for the next cycle defined
@@
 way, the DSP Data Persistence component maintains a dependency on the Singleton
-class TransientPersistenceLayer, which is responsible for keeping track of
-messages that are received by working component. However, in order to add the
-DSP message into its aggregated map, it first transforms the message into an
-instance of the class PersistentMessageUnit, which has references to the class
-regarding the location of the message, SensorLocation.
+class TransientPersistenceLayer, which is responsible for keeping track of the
+messages that were received by the running component. Finally, in order to add
+the DSP message into its aggregated set, it first transforms the message into an
+instance of the class PersistentMessageUnit. This class carries references to
+the class \emph{SensorLocation}, responsible for handling the location of the
+sensor device, and consequently, the physical origin of the message.
 
-As depicted by the the UML Sequence Diagram \cite{uml} of figure
-\ref{fig:From-DSPWireTransport-Server-To-DSP-Broker}, a DSP Message
-is delivery to the DSP Broker for its delivery to Data Consumers configured to
-receive that specific message. For this reason, Figure
-\ref{fig:From-DSP-Broker-To-DSPDataPersistence-General-Sequence} continues the
+A DSP Message is deliveried to the DSP Broker with the intent of being
+forwarded to the data consumers, which are configured to receive that
+specific message. This process is described by Figure
+\ref{fig:From-DSPWireTransport-Server-To-DSP-Broker}. Similarly, continues the
 steps of the DSP Broker from the gate "From DSP WireTransport Server to Broker
@@
   \item Any DSP Measure Message can be persisted by the DSP Data Persistence
-  component. When the DSP Broker delivers a DSP MeasureMessage to this
-  component, it will add the message to a temporary memory location to decrease
-  I/O on the database, as depicted by the gate "create PersistentMessageUnit
-  Gate". This last gate describes the simple procedure to add the message
-  into the transient persistent layer. Listing
+  component. The component will add the message to a temporary memory location
+  to decrease I/O on the database, when the DSP Broker delivers a DSP
+  MeasureMessage to this component, as depicted by the gate "create
+  PersistentMessageUnit Gate". This last gate describes the simple procedure
+  to add the message into the transient persistent layer. Listing
   \ref{file:dsp-message-serialized-ysi}.
@@
 IP address, while the latter identifies whether the PersistenceMessageUnit has
-been saved into the Database or not as depicted by the UML State Diagram
-\cite{uml} in Figure \ref{fig:PersistentMessageState-Diagram}.
+been saved into the Database or not as depicted by the UML State Diagram in
+Figure \ref{fig:PersistentMessageState-Diagram}.
 
@@
 PersistenceMessageUnit is created at the state ``TRANSIENT'', as detailed
-on UML State Diagram \cite{uml} in Figure \ref{fig:PersistentMessageState-Diagram}.
+on UML State Diagram in Figure \ref{fig:PersistentMessageState-Diagram}.
 
@@
 
-In order to build the document key and value segments as described on section
-\ref{sec:dsp-persistence-data-model}. First, the value is built by extracting
-the DSP Message Content from the body of the DSP Message.
+In order to build the document key and value segments, as described on section
+\ref{sec:dsp-persistence-data-model}, the reference to the DSP message is
+used, as the value key is built by extracting the DSP Message Content from the
+body of the DSP Message. Similarly, the key is built by gathering information
+from the DSP Message and the PersistenceMessageUnit class.
 
@@
 
-\begin{figure}[!b]
+\begin{figure}[!h]
   \centering
@@
 \end{figure}
+
+All in all, the design of the DSP Data Persistence uses the DSP Component
+specification defined by the DSP Platform. First, the analysis of the business
+process helped understing the requirements of the system. Similarly, the data
+model design followed the specifications defined in chapter 5. In this way,
+each of the layers of the system were described using UML Diagrams, having
+partial details regarding the OSGi implementation. Next chapter entails the
+experiments conducted to evaluate the designs discussed in this
+chapter, walk-through the experiments conducted to verify the implementation
+using mongoDB.

### 2009-12-08 docs/marcello-dissertation/thesis/chapter8-experimental-results-correct-behavior-and-performance.tex
First Draft of the dissertation peer reviewed with Kleber.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@680 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 act as the mongoDB shards. In this way, two different versions of the Ubuntu
-Linux were used: one 8.04 32bits and another 9.04 64bits (see section
-\ref{sec:experiments-difficulties}).
+Linux were used: one 8.04 32bits and another 9.04 64bits. The configuration
+used for the single server only uses the process ``mongod'' to be started,
+while the clustered setup must be enabled in the database, as shown in
+Listing \ref{file:experiment-setup-mongoshards}, Listings
+\ref{file:setup-os-shards-step1} and \ref{file:setup-os-shards-step2}.
 
@@
 R1, shown in Listing \ref{file:dsp-mongo-service}, defined as the insertion of
-instances of the collected data from the sensor devices using Java. The use of
-the mongoDB Java driver, which provides the methods to access the database
-system and the collection. In this way, the construction of the keys and
-respective values were done using a Hash-like object ``BasicDBObject'' such as
-the ones shown in the method ``buildKeySegment''. Similarly, the creation of
-database indexes is performed using the same approach, as shown in the Java
-method ``setupMetadataIndexes''. Finally, the insertion of the documents is
-performed on the service method call ``insertPersistentUnitMessageContents'',
-which delimits the bulk insert using the transaction management mechanism over
-the collection. In this way, different insert times were collected from the
-logs of the different experiments as summarized on Table
-\ref{tab:experiment-insert-avarage}.
+instances of the collected data from the sensor devices using Java. The
+document follows the specification of the keys designed in section
+\ref{sec:dsp-persistence-data-model}. The use of the mongoDB Java driver, which
+provides the methods to access the database system and the collection. In this
+way, the construction of the keys and respective values were done using a
+Hash-like object ``BasicDBObject'' such as the ones shown in the method
+``buildKeySegment''. Similarly, the creation of database indexes is performed
+using the same approach, as shown in the Java method ``setupMetadataIndexes''.
+Finally, the insertion of the documents is performed on the service method
+call ``insertPersistentUnitMessageContents'', which delimits the bulk insert
+using the transaction management mechanism over the collection. In this way,
+different insert times were collected from the logs of the different
+experiments as summarized on Table \ref{tab:experiment-insert-avarage}.
 
@@
     \begin{center}
-        \begin{tabular}{|p{70pt}|p{100pt}|p{250pt}|p{100pt}|}\hline
+        \begin{tabular}{|c|c|c|c|}\hline
         \textbf{Scenario Type} & \textbf{Language} & \textbf{Method Call} & \textbf{Use Case}\\\hline 
@@
 \subsection{User Experience}
+\label{sec:experiments-user-experimence}
 
-Different methods of data access were used after the randomly generated data
-was inserted into mongoDB: using the shell commands mongo and ``mongoexport'', 
-command-line execution developed in Python, through the REST Web Services API,
-and using a web application, written in the Ruby Programming Language
-\ref{ruby}, providing the view over the collected data using a web browser.
-More details about the execution of these commands during the experiments is
-shown in section \ref{sec:mongodb-user-experience}.
+In general, the use experience of mongoDB is related to the use of programming
+languages, since most of the methods to create, access and modify the database
+is performed using the programming language abstraction of objects and method
+calls. The creation of data was the first contact with mongoDB in terms
+of data creation in Java. Later, Different methods of data access were used
+after the randomly generated data was inserted into mongoDB: using the shell
+commands mongo and ``mongoexport'', command-line execution developed in
+Python, through the REST Web Services API, and using a web application,
+written in the Ruby Programming Language \cite{ruby}, providing the view over
+the collected data using a web browser. More details about the execution of
+these commands during the experiments is shown in section
+\ref{sec:mongodb-user-experience}.
 
@@
 querying the collections from mongoDB can be performed using the HTTP Request
-Parameter variables, by adding the prefix ``filter_'' before the name of the
-key of the document. The given example shows the query performed in the
+Parameter variables, by adding the prefix ``filter\underline{ }'' before the
+name of the key of the document. The given example shows the query performed in the
 collected data for the key ``observation.Conductivity'' with the value
@@
 
-In order to verify
+The actual execution of the solution was performed using the demo Mouse
+Actions, which consists of a software sensor that perceives any mouse action in
+a window area. The experiment consisted of moving the mouse around a white
+window, where each action is captured and temporarily saved until a threshold is
+reached, making the component to send the messages to the persistence layer, so
+that it can save on mongoDB. Figure \ref{fig:experiment-mouseactions-dsp} shows
+a screenshot of the moment when the message is sent to mongoDB, being verified
+through the HTTP REST Interface using a web browser.
 
@@
 5. In this way, the requirements described in chapter 4 were expressed as
-examples of use cases in section \ref{sec:exp-scenarios}.
-
-All the collected data from sensors are stored in a
-given single server, which characterizes an External Storage device
-
-the data. The process to describe the data is done with prior knowledge of the
-sensor device data properties when choosing the relational data model, because
-of the process of database normalization \cite{db-normalization}, which takes
-place after the the entities have been identified. Consequently, the
-introduction of a new sensor device might potentially require changes on the
-database schema structure conceived before, as the process of database
-normalization must follow.
+examples of use cases in section \ref{sec:exp-scenarios}. This chapter
+discusses the experiments in different angles: analysis of the taxonomies, the
+infrastructure and the use cases evaluation.
 
@@
 Storage. After selecting mongoDB, the development of the persistence layer met
-the requirements of archiving data after inserting the the randomly generated
-data that represents the data for the YSI sonde. As it is shown in section
-\ref{sec:exp-measurements}, the collected data is archived in files using the
-BSON format \cite{bson}, a binary representation used to store instances of
-the document described in Listing \ref{file:mongodb-ysi-data-format}. Along
-with the purpose of data archival, the use of mongoDB also met the
-specification of an external storage since the data persisted in the file
-system, as well as Data-Centric for the proposal to use mongoDB distributed
-shards.
+the requirements of archiving data for different types of sensor devices from
+the DSP Platform such as the YSI Data Handler and the Mouse Actions demo data.
+As it is shown in section \ref{sec:exp-measurements}, the collected data is
+archived in files using the BSON format \cite{bson}, a binary representation
+used to store instances of the document described in Listing
+\ref{file:mongodb-ysi-data-format}. Along with the purpose of Data Archival,
+the use of mongoDB also met the specification of an External Storage, since the
+data persisted in the file system, as well as Data-Centric for running mongoDB
+with different  distributed mongo shards (cluster nodes).
 
-SF-BEAMS uses the Tabular Data model to describe the collected data from the
-sensor devices. As discussed in chapter 5, the decision to use mongoDB is the
-fact that the literature review does not mention the use of the Key-Value-Pair
-Data model. For this reason, since mongoDB implements a variation of the 
-key-value data model based on the JSON \cite{json} data format, the Data
-Model taxonomy should be updated, as shown in Figure
-\ref{fig:taxonomy-data-model-modified}.
+SF-BEAMS uses Tabular Data model, Schema-less data model that represents the
+collected data using comma-delimited files, without a database system to manage
+those files, just the OPEnDAP middleware. In contrast, the data persistence
+proposed by this work proposes mongoDB as a schema-less database system to
+manage the collected data collected using NetBEAMS. As described in chapter 5,
+the schema-less data model can provide a better support to Data Provenance
+because the data model gives support to the change of records without affecting
+the others. In this way, the metadata used by the implementation clearly
+describes the three fundamental questions regarding the collected data from
+sensor devices, as described in section \ref{sec:dsp-persistence-data-model}.
+First, the identity of the data is given by the key ``message\underline{
+}id'', in a way that it uniquely identifies the data produced by the sensor
+identified by the key ``sensor.ip\underline{ }address''. In this way, the
+message can be tracked back to the sensor that produced the data for different
+purposes such as tracking the data producer. Second, time dimensions were used
+in order to track two different aspects of the data collection: the valid time
+can be used to correlate observations to the exact moment it occurred, while
+the transaction time can be used to verify when the data was collected, and
+make decisions such as deleting the instances of data were collected in the
+previous two days, etc. Last, but not least, the use of the key
+``sensor.location'' can give the exact location of the data, since researchers
+can be interested in finding data based on a specific location. Finally, the
+most important aspect of the schema-less data model is that the observed data
+is added to the key ``observation'', as any sensor device carries its set of
+attributes of key-value pairs. A good example of the applicability of the
+schema-less model is the scalability regarding changes to the data structure.
+For instance, if an existing sensor device is flushed with a new version of
+the firmware\footnote{Primary machine instructions that run in the hardware},
+changing the format of data types or adding new values, the following instances
+of collected data may be different to the previous one by simply adding the
+new additions of data. For this reason, the DSP Data Persistence gives
+administrators the ability to choose which fields to persist by declaring it
+using a bootstrap message for the component, as shown in the parameter
+``YSI\underline{ }DESIRED\underline{ }PROPERTIES'' of Listing
+\ref{file:dsp-data-persistence-bootstrap.xml}.
 
-\begin{figure}[h]
+In order to access the collected data, the implementation used different APIs
+provided by the mongoDB community. The Centralized Query Mechanism was
+satisfied by the use of those APIs, which is also fultill the requirements
+defined by NetBEAMS. For example, the access of the database using Python or
+Java were based on the needs and the use of the languages for different
+purposes. In this way, the implementation of the persistence model captures
+the instances of any known data collected from NetBEAMS and saves it using the
+document pattern as shown in Listing \ref{file:mongodb-ysi-data-format}. Any
+programming language used was capable of accessing the same collected data by
+their own.
+
+In the end, the implementation of the data persistence for NetBEAMS was
+developed in both Single and Distributed Systems. The use of regular single
+server execution provided enough computational power for a compatible volume of
+data when compared to the one produced from the SF-BEAMS sensor network. In
+this way, the implementation suggested by this work fulfill the need of a
+persistence system for a Small volume of data being created at the same type
+in a Single Server. Furthermore, mongoDB also provided support for Distributed
+System, when the use of databse shards, in forms of different mongoDB nodes,
+were used as a clustered environment.
+
+\subsection{Infrastructure Evaluation}
+
+Overall, the implementation of the DSP Data Persistence component meets the
+specification of the use cases defined in section \ref{sec:exp-scenarios}.
+From the perspective of disk-space utilization for the compatible volume of
+data for 5 five YSI devices is that it requires a small amount of disk space.
+According to the mongoDB documentation \cite{mongodb}, the claimed amount of
+disk space is related to how mongoDB uses a binary representation of the JSON
+documents, which are composed only of strings, together with the definition of
+the indexes. As a consequence, the size of the keys defined for the YSI sonde
+data suggests additional disk space were used for the storage.
+
+The performance to insert data into the database system using the Java driver
+is documented by mongoDB as the fastest approach, providing persistence
+thousands of documents per second. This capacity is more than necessary for
+NetBEAMS, which can use the Single System environment. On the other hand,
+improvements over the response time can be achieved through the implementation
+of a Data-Centric approach on a Distributed Database environment, or clusters.
+In this approach, the inserted data is stored in different servers based on a
+partition key. Given the fact that less data is concentrated in each
+partition, the centralized query mechanism used is faster, and consequently
+improves the performance as described in the Data Location taxonomy in chapter
+3. In this way, the infrastructural performace during the inserts suggests that
+persisting data collected from the NetBEAMS sensors using a clustered
+environment is an option when the single server capacity is reached.
+
+Different problems and limitations were faced during the execution of
+mongoDB, which were relate to the infrastructure used: the limitations
+regarding disk storage in 32bit operating systems and very slow
+performance when using virtualized servers. During the execution of the
+experiments, mongoDB used memory accordingly, without taking control of the
+operating system's resources. However, when mongoDB is used in a 32bit server
+architecture, there are limitations on the disk space used to store
+collections. In this way, after reaching 2GB of collected data during many
+different runs, mongoDB crashed and did not allow the use of the collected
+data. For this reason, the setup of a new VirtualBox using Ubuntu Linux 64bits
+as a guest to a Mac OS X 64bits was prepared and used throughout the completion
+of this work's experiments. On the other hand, the performance was drastically
+decreased using mongoDB in virtualized servers due to the fact of dynamic
+partition chosen. The performance shell application ``iostat'' was used to
+verify the behavior, and the use of mongoDB was switched to the Host operating
+system.
+
+\subsection{Use Cases Evaluation}
+
+\cite{sn-programming-language} suggested that programming that programming
+languages are easier abstractions to researchers that use collected data from
+sensor devices without background in computer science or database systems. The
+authors have suggested the implementation of a programming language for this
+group of biologists, for instance. However, this suggestion has good
+intentions, but limits the users with a single mechanism to access the
+collected data from sensors. For this reason, this dissertation proposes the
+use of key-value pair databases, such as mongoDB, to provide transparent access
+the the data through the abstraction of any programming language that has a
+driver written for mongoDB. As one can see during the user experience section,
+different programming languages were used to access the collected data, a novel
+mechanism not yet reported in the sensor networks community.
+
+The development of the scenarios were clearly written in different ways, such
+as the ones defined in Listing \ref{file:experiment-scenarios}, which
+implements the use cases for data retrieval, update and deletion. The main
+advantage of this approach is that users do not use the SQL language to extract
+data from the database. Rather, the use of the method calls defined for each
+operation, as well as the concept of dynamic binding and objects, gives simple,
+yet powerful syntax to access data. For example, based on the description
+of the collected data by keys and values shown in Listing
+\ref{file:mongodb-ysi-data-format}, different use cases can evaluate documents
+in this structure. Usually researchers may search the collected data by a given
+range of dates, for their regular activities of data analysis. The scenario
+implemented by R1 in the abovecited listing shows it clearly. Similarly, the
+search for data in a given sensor device is expressed by the search for the IP
+address for them on the use cased defined by R2. Finally, the search for the
+raw data obtained by the sensor devices is used from the key ``observation'',
+as shown by the R3 use case implementation.
+
+Accompanying that non-predictable events might occur during the data
+collection process of sensor devices, and consequently, influence in their
+values, the support to annotations is a feature that can be used to describe
+the data. mongoDB indirectly supports the use of annotation because of its free
+method of updating documents and adding new keys to any of them without
+requiring design changes. Supposing an environmental accident happens, such as
+the San Francisco oil spill of October 2009 \cite{sfbay-oilspill2009}, the
+values of the collected data, and consequently data analysis, be influenced by
+different ranges of values. In this way, the implementation of the scenario U1
+provided the implementation of the addition of a new key called ``tag'', with
+the annotation ``oil spill''. This new key, after properly indexed, provides
+another option to meaningfully search the collected data from the sensor
+devices. Last, but not least, old data can be deleted as expressed by D1, as an
+operation supported by any database system.
+
+One common approach during research is data sharing. Researchers can choose
+the use of spreadsheets to analyze collected data with simpler tools than
+database systems or programming language. For this reason, OPEnDAP is used in
+the community to share data using a comma-delited format. In this way, the
+implementation of the use cases for R4 were developed in two different ways.
+First, the use of the export capabilities to a common-delimited values was
+performed \ref{file:mongodb-export-command}, taking approximaly 3.3 minutes to
+export one year worth of data for 5 YSI devices without indexing. Second, the
+implementation of a custom export utility using Python, as shown in Listing
+\ref{file:experiment-export-python} enabled the creation of the comma-delimited
+files with the same format as the ones used by OPEnDAP. The comparison can be
+done by analyzing the output of the OPEnDAP request from SF-BEAMS on Listing
+\ref{file:rtc-ysi-opendap}, and the output provided from the script written in
+Python, show in Listing \ref{file:experiment-export-python-results}.
+Furthermore, considering research teams with skills in Web Services, an easy
+ubiquos integration with mongoDB can be provided by using the REST Web Services
+APIs provided by mongoDB, still in alpha version. A practical application of
+this API was the created Apache Futon 4 mongoDB web application, used to give
+user friendlier access to the collected data as shown in 
+\ref{fig:view-collections-instance-browser-futondb}.
+
+As discussed in chapter 5, the differences between a traditional database
+system, which uses the relational data model, and a document-oriented database
+system were clearly noticed during the development of this work. As mentioned
+in the section \ref{sec:experiments-user-experimence}, the user experience was
+mostly related to the definition of the keys that describe the sensor device,
+as discussed during the design of the solution in chapter 6. On the contraty,
+the abstraction of objects and collections during the data access using
+different programming languages. Therefore, based on the use of a different
+type of data model in the scope of data persistence for sensor networks, the
+taxonomies defined in chapter 3 could be updated to the one shown in Figure
+\ref{fig:taxonomy-data-model-modified}, including the Key-Value Pair Data
+Model.
+
+\begin{figure}[!h]
   \centering
@@
 
-Another aspect of the implementation is that provides APIs in different
-languages for the data creation and extraction processes. In this way, the
-implementation of the persistence model captures the instances of any known
-data collected from NetBEAMS and saves it using the document pattern as shown
-in Listing \ref{file:mongodb-ysi-data-format}. One application of a
-schema-less data model is the addition of any new key-value pair to any
-instance of the document, as mongoDB supports update operations to any of the
-collections. It can use tags for the purpose of annotations as described in the
-taxonomy of data description. For example, in case of an oil spill in the San
-Francisco Bay occurs, researchers can update the data set of a given date
-range by adding tags to the relating documents. As a consequence of the use of
-a schema-less data model, these changes do not affect the other data
-instances, since there are no schema changes.
+\subsection{Difficulties Experienced}
+\label{sec:experiments-difficulties}
 
-As described in chapter 5, the schema-less data model can provide a better
-support to Data Provenance because the data model gives support to the change
-of records without affecting the others. In this way, the metadata used by the
-implementation clearly describes the three fundamental questions regarding the
-collected data from sensor devices, as described in section
-\ref{sec:keys-definition}. First, the identity of the data is given by the key
-``message\underline{ }id'', in a way that it uniquely identifies the data
-produced by the sensor identified by the key ``sensor.ip\underline{
-}address''. In this way, the message can be tracked back to the sensor that
-produced the data for different purposes such as tracking the data producer.
-Second, time dimensions were used in order to track two different aspects of
-the data collection: the valid time can be used to correlate observations to
-the exact moment it occurred, while the transaction time can be used to verify
-when the data was collected, and make decisions such as deleting the instances
-of data were collected in the previous two days, etc. Last, but not least, the
-use of the key ``sensor.location'' can give the exact location of the data,
-since researchers can be interested in finding data based on a specific
-location. Finally, the most important aspect of the schema-less data model is
-that the observed data is added to the key ``observation'', as any sensor
-device carries its set of attributes of key-value pairs. A good example of the
-applicability of the schema-less data model is the scalability regarding
-changes to the data structure. For instance, if an existing sensor device is
-flushed with a new version of the firmware\footnote{The first set of machine
-instructions to run on the hardware after the application of power stored
-permanently in PROM or ROM or semi-permanently in EPROM}, changing the format
-of data types or adding new values, the following instances of collected data
-may be different to the previous one by simply adding the new additions of
-data. For this reason, the DSP Data Persistence gives administrators the
-ability to choose which fields to persist by declaring it using a bootstrap
-message for the component, as shown in the parameter ``YSI\underline{
-}DESIRED\underline{ }PROPERTIES'' of Listing
-\ref{file:dsp-data-persistence-bootstrap.xml}.
+The inception of the experiments described in by this work moments of Software
+Engineering. Particularly speaking about the learning curve, different
+situations contributed to the success of the first experiments. Since the
+technoogy is new, it brought new challenges regarding the usability and user
+experience. The technology is relatively new and still under development,
+which many different bugs were be reproduced by random during the execution of
+the experiments, and therefore, reports were sent to the open-source community
+through use different available communication channels. First, the fastest way
+to solve an issue was through the mongoDB's IRC channel located at the Freenode
+server (www.freenode.org, channel mongodb). Second, the use of the users
+group, located at http://groups.google.com/group/mongodb-user also helped
+solving problems.
 
-The centralized query processing used by NetBEAMS is a result of the
-infrastructure used by NetBEAMS, as well as the purpose of data and the
-location of the collected data. As a direct result of a schema-less data
-model, the centralized query processing is easily developed using any of the
-available programming languages provided by mongoDB. Considering the fact that
-programming languages are easier abstractions to researchers without computer
-science background \cite{sn-programming-language}, mongoDB's approach to data
-access through API can be seen as an easier task to manage data. As it is shown
-in Listing \ref{file:experiment-query-scenarios}, the scenarios designed to
-evaluate the implementation are written in Javascript, since it is the script
-language by the mongoDB Shell. Therefore, it attends the requirement of not
-using the SQL language as the main query language of the data, giving a better
-data access abstraction similar to programming languages for different users.
+Physical memory is limited and, for this reason, it may not be enough to run
+the experiments for total workloads. In this case, the Java Virtual Memory was
+totally consumed by one of the revisions of the experiments setup. Therefore, 
+it was learned that large amount of data loads require different approach to
+manage data in memory accordingly, such as creating different object pools as
+temporary cache for objects whose state are similar to others.
 
-One of the most important features of a data persistence for sensor
-networks is the replacement of the SQL by the use of programming
-languages, as they tend to be easier abstraction to users without the
-background in database systems \cite{sn-programming-language}. In this way,
-mongoDB covers the requirements of providing data access to the collected
-sensor data in different languages. For example, the data collected by the DSP
-Data Persistence component is written in Java, while the implementation of the
-experiments scenarios for querying the collected data is given in Javascript. In
-this way, researchers of the RTC may choose one of the enumarated languages
-APIs provided by the mongoDB to access the stored data from NetBEAMS.
+Also related to the difficulties of this work, the fact that mongoDB is not yet
+stable makes the implementation a bit harder. The main mongod process is on its
+1.1.4 version as the time this line is writte, while the shards support is
+still in alpha version. For instance, different work-arounds were performed to
+have shards experimented as described, but the execution was not successful.
 
-Finally, the implementation supported by mongoDB during the experiments was
-following a single dedicated server virtualization using VirtualBox 
-\cite{virtualization}.
-
-\subsection{Infrastructure Evaluation}
-
-Overall, the implementation of the DSP Data Persistence component meets
-the specification of the use cases defined in section \ref{sec:exp-scenarios}.
-The implementation of mongoDB using a single server environment can give
-support to the current workload and infrastructure from the use of NetBEAMS to
-collect data from the sensors devices from SF-BEAMS. From the perspective of
-disk-space utilization and given that aproximaly 1.3 Gigabytes of disk space
-were used to save one year-worth of collected data equivalent to five YSI
-devices. According to the mongoDB documentation \ref{mongodb}, the claimed
-amount of disk space is related to how mongoDB uses a binary representation
-of the JSON documents, which are composed only of strings, together with the
-definition of the indexes also represent in JSON. As a consequence, the size of
-the keys defined for the YSI sonde data defines the amount of disk space used.
-For this reason, different experiments were conducted with different
-definitions of the keys of the documents
-approach when compared the the RTC's OPEnDAP format shown in Listing
-\ref{file:rtc-ysi-opendap}. Then, the second specification used short names
-for the keys. In this way, the amount of storage space utilized by for the
-different workloads define in section \ref{sec:workload} can be summarized as
-follows
-
-The performance to insert data into the database system using the Java driver
-is documented by mongoDB as the fastest approach, providing persistence to
-approximaly 25,091 documents per minute on an indexed database. In contrast,
-when indexes are not used, more then 169,508 documents could be inserted per
-minute. Although the latter approach supports more ducuments than the former,
-the trade off between insertion and selection time is relevant due to the
-activities of retrieval for documents for specific periods of time. Therefore,
-the use of indexes might speed the retrieval time considerably. Considering
-this number of insertions to the database the number of sensors producing data
-and sending to the data sink, mongoDB and NetBEAMS implementation can
-virtually support thousands sensor devices.
-
-Disk utilization is one important characteristic of the selected database
-system. However, the retrieval of the collected data from sensor devices is
-a much more important requirement because most of the utilization of the data
-persistence of a sensor network is regarding the search for the properties on
-the collected data. After analyzing the response times for the scenarios
-described in section \ref{sec:exp-scenarios}, it is clear that the use of
-indexes improves the overall execution of the queries, specially because of the
-amount of data collected over the years. Furthermore, the response times
-described in this scenario is directly related to the use of the External
-Storage strategy. On the other hand, improvements over the response time can be
-achieved through the implementation of a Data-Centric approach on a Distributed
-Database environment, or clusters. In this approach, the inserted data is
-stored in different servers based on a partition key. Given the fact that less
-data is concentrated in each partition, the centralized query mechanism used is
-faster, and consequently improves the performance as described in the Data
-Location taxonomy in chapter 3.
-
-When it comes to re-using the data collected by the DSP Data Persistence,
-different approaches were available. First, the use of the export capabilities
-to a common-delimited values was performed \ref{file:mongodb-export-command},
-taking approximaly 3.3 minutes to export one year worth of data for 5 YSI
-devices without indexing. Second, the use of the client shell gives a
-centralized access to either the single server. However, in order to better
-provide access of data to users without skills in database systems, web
-applications can be built on top of the REST Web Services APIs \cite{http-rest}
-provided by mongoDB. One such example of using the REST API is the adaptation
-of Apache Futon 4 mongoDB, as shown in Figure
-\ref{fig:view-collections-instance-browser-futondb}, where users can navigate
-through the collections of the collected sensor network data categorized by the
-collection. Therefore, RTC staff technicians could take advantage of this tool
-to view and navigate over the collected data from the sensor devices.
-
-\subsection{Use Cases Evaluation}
-
-Supposing an environmental accident such as the San Francisco oil
-  spill happened again \cite{sfbay-oilspill2009}, changing the sampling
-  frequency drastically to different values.
-
-\section{Difficulties Experienced}
-\label{sec:experiments-difficulties}
-As the development of computing system, the inception of the experiments
-described in this chapter had different moments of Software Engineering.
-Particularly speaking about the learning curve, different situations
-contributed to the success of the first experiments. Since the technoogy is
-new, it brought new challenges regarding the usability and user experience.
-Since the technology is relatively new and still under development, many
-different bugs were could be reproduced by random during the execution of the
-experiments, and therefore, sent to the open-source community through use 
-different available communication channels. First, the fastest way to solve an
-issue was through the mongoDB's IRC channel, located at the Freenode server
-(www.freenode.org, channel mongodb). Second, the use of the users group, located
-at http://groups.google.com/group/mongodb-user also helped solving problems.
-
-Physical memory is limited, for this reason, it may not be enough to run the
-experiments for total workloads. In this case, the Java Virtual Memory was
-totally consumed by one of the revisions of the experiments setup. Huge data
-loads require different approach to manage data in memory such as creating
-object pools as temporary cache for objects whose state are similar to others.
-
-While performing the experiments, the 32bit machine reached its limitation of
-2GB of physical data. With know about that, the environment needed to be
-updagred for mongoDB requires a 64-bit operating system when the amount of
-stored data reaches 2 Gigabytes. For this reason, the scenarios with indexes
-used a dedicated virtualized server using Ubuntu Linux 9.04 64 bits with 1 Gb
-of memory.
+To sum up, this chapter described the experimental results of the
+implementation of a data persistence component for NetBEAMS, showing the
+experiment results, the measurement results and finally the discussion about
+the solution in general. The following section brings the conclusions and
+future works for this dissertation.

### 2009-12-08 docs/marcello-dissertation/thesis/chapter9-conclusions-and-future-works.tex
First Draft of the dissertation peer reviewed with Kleber.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@680 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 
-\chapter{Conclusions}
+\chapter{Conclusions and Future Works}
 
-This work presented a collection of taxonomies related to data persistence for
-sensor networks using different approaches ranging from data models and server
-infrastructure organization. The main result of this work was that persistence
-for collected data from sensors are better maintained using schema-less data
-models, since it provides offers an approach to deal with unanticipated data
-types from the sensor devices.
+There were important important remarks covered in this work, including the
+importance that sensor networks are becoming ubiquitous, having their produced
+data available online and other types of media. In this way, in order to
+provide access to the collected data, one must first assess the different
+types of infrastructural characteristics of the studied sensor networks. In
+general, the selection of a technology that provides such a persistence
+solution capabilities can  part of a process of analysis of taxonomies
+proposed by this work, characterized as the first contribution of this
+disseration.
 
-As seen in the state of the art, different approaches can be done in order to
-evaluate data persistence for nsensor networks. T
-
-There are many important features covered in this work, including the
-importance of the sensor networks rely in how users have access to the
-collected data. In order to provide access to the collected data, one must
-first assess the different types of infrastructural characteristics of the
-studied sensor network. In general, the selection of a technology that
-provides such persistence capabilities can be part of a process of analysis of
-taxonomies proposed by this work. However, this work proposes a data model not
-yet used in the sensor network community and, thus, represents an important
-contribution for the area.
-
-Taking into account the software architecture provided by NetBEAMS, as well as
-the requirements for data collection from the RTC research group, this work
-shows a novice way to persist data collected from sensor devices by using a
-schema-less data model, whose query capabilities are based upon programming
-languages, in contrast to the use of the Structured Query Language, SQL. The
-experiment results showed that the implementation of a persistence layer with
-mongoDB can be used to persist data at its maximum rate of one sample per
-minute. However, the evaluation of mongoDB was not used to implement a
-Data-Centric Storage approach, which can decreases the amount of time to search
-data since the process uses parallel search.
+In my own opinion, based on the findings of the taxonomies, together with
+the empirical analysis of the technologies, this work revealed a novel approach
+to provide persistence and access to collected data from sensor devices using
+the Key-Value-Pair data model. Besides the contribution of a new component
+component for the Data Sensor Platform, the evidence that the system improves
+productivity was confirmed during the execution of the experiments, writing
+less source-code to obtain better functionalities when compared to other
+solutions. Similarly, improved work suggests that the abstraction of
+programming language should be used with research groups whose primary
+expertise is not in Computer Science or Database Systems.
 
@@
 mongoDB's ability to cope with the uncertainty of any properties collected from
-any sensor specification. Furthermore, the solution also provides a good
-support to the provenance-related data as the schema-less data model gives the
-ability to .
+any sensor specification.
 
-\chapter{Future Works}
-
-0. Todo list for this work
-1. Database System Problems
-2. Software Engineering Problems
-3. Applications
- - plot graphs of collected,
-
-There are many different directions for future works. First, the
-implementation of a Data-Centric Storage approach can help the system scale
-horizontaly by using the mongoDB's database shards. In this way, this approach
-can reduce the retrieval time by focusing on centralized selection of data over
-a given database shard. Moreover, the overall search through the collected data
+As far as future works, there are many different directions for future works.
+First, the first major release of mongoDB with complete support to shards is
+scheduled for 2010. An important mechanism of mongoDB is the fact that it can
+manage different mongoDB servers spread as a cluster, as it gives support to
+the Data-Centric Storage approach. In this way, this approach can reduce the
+retrieval time by focusing on centralized selection of data over a given
+database shard. Moreover, the overall search through the collected data
 attributes can be drastically decreased by parallel search over the database
-shards using the MapReduce \cite{map-reduce} tools. At this time of
-development, the MapReduce API offered by mongoDB is on its alpha version, and
-therefore, presenting lots of bugs. Figure
-\ref{fig:future-works-data-centric-map-reduce} shows an architectural idea of
-the deployment of a Data-Centric storage approach. When searching for specific
-document attributes, the process is sent in parallel to each of the database
-shards, and the result is collected by the cluster head. After calculating the
-final result, mongoDB server returns the relevant result to the user.
+shards using the MapReduce \cite{map-reduce} tools. At the time of development
+of the experiments, the MapReduce API from mongoDB was also in the alpha
+version, and therefore, presenting lots of bugs. one suggestion for the
+implementation of such functionality is based in Figure
+\ref{fig:future-works-data-centric-map-reduce}, which shows an architectural
+idea of the deployment of a Data-Centric storage approach. When searching for
+specific document attributes, the process is sent in parallel to each of the
+database shards, and the result is collected by the cluster head. After
+calculating the final result, mongoDB server returns the relevant result to
+the user. I was personally expecting to implement the scario of data retrieval
+of collected daa.
 

### 2009-12-08 docs/marcello-dissertation/thesis/preamble.tex
First Draft of the dissertation peer reviewed with Kleber.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@680 559edde2-a4d4-44e0-96fc-4035e40d7396

--- 2009-12-01 docs/marcello-dissertation/thesis/preamble.tex
@@
 properties such as the purpose and location of the observed data, as well as
-how the data is described. In this way, this work first proposes a set of
-taxonomies that can be used to study and classify a given instance of sensor
-network and its produced data. Then, based on the classes of these taxonomies, 
-this work selected a Document-Oriented Database system as a result of an
-empirical evaluation of the requirements for a new persistence layer for 
-NetBEAMS, the case study used in this work. For this reason, the design and
-implementation of a new data persistence component for NetBEAMS show how
-schema-less data models help decreasing the complexity related to the database
-layer, not yet reported by the state of the art. Furthermore, this work also
-proposes the use of Database Shards in order support the implementation of
-Data-Centric sensor networks. 
-}
-\vspace*{\fill}
+how the data is described. For this reason, this work proposes a set of data
+persistence taxonomies based on the state of the art, which can be used to
+classify the properties of the produced raw data. In order to evaluate the
+proposed taxonomies, this work designs and implements a data persistence
+component for NetBEAMS, our case study of modular sensor network infrastructure
+developed to improve the operation of the SF-BEAMS environmental sensor
+network. In this way, based on empirical analysis regarding proposed
+taxonomies for the case study, this work selected mongoDB, a schema-less
+document-oriented database instead of the traditional use of relational
+databases. As a result based on the experiments conducted, this work suggests
+that use of Key-Value Pair datasents in orde tor provide External of
+Data-Centric approach. Therefore, the use of a solution of the help decreasing
+the complexity related to the data persistence layer for sensor networks with
+external or data-centric storage mechanism, and whose purpose is data
+archival. Finally, this work proposes different future works such a
+data-centric approach using Databa se Shards to store te collected data and Map
+Reduce to execute parallel queries over the shards.} \vspace*{\fill}
 

### 2009-12-08 docs/marcello-dissertation/thesis/chapter1-introduction.tex
Most of the revisions and updates... Changes to chapters 3, 4 after Kleber's chat online (adding missing taxonomy).
Revisions of print chapters until chapter 5.

Chapter 7 is reviewed until before discussion. 

Next revision should include all the necessary reviews for the
release draft.

git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@676 559edde2-a4d4-44e0-96fc-4035e40d7396
Changes to chapter 1 after Niki's and Prof. Puder's audio feedback.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@681 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 such as accurate measurements during scientific research \cite{sn-intro01} and
-the general public access such as weather forecast \cite{sn-intro02}. In the
+the general public access such as weather forecasting \cite{sn-intro02}. In the
 scientific community, they serve as tools to watch the state of the environment
@@
 National Oceanic and Atmospheric Administration (NOAA) \cite{sn-ex03},
-provides online information from different buoys in different shores around
-the world regarding water quality, current information, etc, what
+provides online information from different buoys located in different shores around
+the world regarding water quality, current information, etc, that
 characterizes sensor network for environmental monitoring \cite{sn-ex01}.
@@
 network infrastructure that can be used to automate the process of data
-gathering in sensor networks. In this way, the review of the state of the
-art in rewiewed that different approaches can be taken to provide access to the
+gathering in sensor networks. In this way, state of the
+art of data persistence in sensor networks was reviewed for different approaches used to obtain access to the
 collected sensor data: interrogate sensor devices indirectly by accessing its
-data in places called network data sink or simply connect directly to the
+data in places called network data sinks or simply by connecting directly to the
 device. While the collected data is only available in-memory as a data stream
-in the latter approach, it is stored in in any sort of secondary memory
+in the latter approach, it is stored in any sort of secondary memory
 device such as a hard-disk or a flash memory for archival purposes in the
 former approach. Along with the location of the collected data, the data model 
-may also influence in the process of reusing the collected data. In this way,
+may also influence the process of reusing the collected data. In this way,
 the introduction of a persistence layer for existing sensor networks requires
-an understanding about its infrastructure, the nature and the location of the
-collected data is temporalily located after it is produced from the sensor
+an understanding about its infrastructure, the nature and the location of where the
+collected data is temporarily located after it is produced from the sensor
 device. Similarly related to the infrastructure, the data model used to
 represent the collected data is another aspect of the data persistence layer
-implemented to manage the raw data values, and may rise questions. Considering
+implemented to manage the raw data values, and may raise questions. Considering
 that sensor networks devices may change over time, how can a data model take
 into account future additions of new entities to the database schema? Should 
-such a solution take into account the end users' skills in database system or
+such a solution take into account the end users' skills in database systems or
 provide access to the data through programming languages
-\cite{sn-programming-language}? These questions are faced when the solution
-considers the primary users of the collected data such as Biologists from the
-the department of Biology at San Francisco State University, which are
-responsible to maintain SF-BEAMS \cite{sfbeams2006}, a marine sensor network
+\cite{sn-programming-language}? These questions are faced when 
+considering the primary users of the collected data such as Biologists from the
+the department of Biology at San Francisco State University, who are
+responsible for maintaining SF-BEAMS \cite{sfbeams2006}, a marine sensor network
 used by NetBEAMS. While access to the collected data may represent
 difficulties, other technical issues may also be faced such as growth in terms
-of the number of sensor devices and, consequently, the number of data
-produced. In this way, sensot networks engineered with the purpose of data
-archival has to answer questions related to how disk storage can be maximized
-when storing the collected data. Last, but not least, the way to read the data
-for reuse is the most striking features of a sensor network, taking into
+of the number of sensor devices deployed and, consequently, the amount of data
+produced. In this way, sensor networks engineered with the purpose of data
+archival must address questions related to how disk storage can be maximized
+when storing the collected data. Lastly, the way to read the data
+for reuse is the most striking feature of a sensor network, taking into
 account the different data formats used by the primary users of the system.
-How can individual collected sensor data be exported to other formats such as
+How can individually collected sensor data be exported to other formats such as
 spreadsheets in order to be shared among research colleagues? For these and
@@
 evaluate database technologies for sensor networks and, specifically speaking,
-to NetBEAMS. As a matter of fact, this first contribution of this work is the
+to NetBEAMS. As a matter of fact, the first contribution of this work is the
 creation of a set of basic taxonomies for data persistence in sensor networks,
@@
 produced by a specific type of sensor. In conclusion, the implementation
-attends most of the requirements and specifications defined by the taxonomies,
-giving an alternative data mode that solves the problem related to constant
-schema changes, providing an outstanding support for Data Provenance
+fulfills most of the requirements and specifications defined by the taxonomies,
+giving an alternative data model that solves the problem related to constant
+schema changes, providing a good support for Data Provenance
 \cite{sn-provenance} for sensor networks. In addition, the solution provides
 different data access alternatives such as the use of programming languages or
-REST Web Services \cite{http-rest} instead of the Structure Query Language SQL
+REST Web Services \cite{http-rest} instead of the Structure Query Language (SQL)
 \cite{sql} used in relational databases \cite{relational-model}, as well as
@@
 This paper is organized as follows: the next chapter describes the literature
-review as the state of art on sensor networks and data persistence, providing
+review of the state of art on sensor networks and data persistence, providing
 the underlying foundation for the design of taxonomies in data persistence for
-sensor networks in chapter 3. Then, chapter 4 examines the selected case study
-and its requirements for a persistence layer, whose technology's properties are
-evaluated against the developed taxonomies in chapter 5. The culmination of
-this work is chapter 6, which details the design and implementation of a data
-persistence solution for the selected case study. Likewise, chapter 7 discusses
+sensor networks in Chapter 3. Then, Chapter 4 examines the selected case study
+and its requirements for a persistence layer, whose technological properties are
+evaluated against the developed taxonomies in Chapter 5. The culmination of
+this work is Chapter 6, which details the design and implementation of a data
+persistence solution for the selected case study. Likewise, Chapter 7 discusses
 the solution by presenting the measurements and results of the conducted
-experiments. Last, but not least, chapter 8 presents the conclusions of this
-work, together with the different possible future works not addressed by this
-dissertation.
+experiments. Lastly, chapter 8 presents the conclusions of this
+project,together with the different possible future work not addressed by this
+report. Chapter 9 presents the appendix with source-code implemented, as well as
+logs collected from experiments.

### 2009-12-09 docs/marcello-dissertation/thesis/chapter2-state-of-the-art.tex
Changes to chapter 2, after Prof. Puder's and Niki's revision.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@682 559edde2-a4d4-44e0-96fc-4035e40d7396

--- 2009-12-04 docs/marcello-dissertation/thesis/chapter2-state-of-the-art.tex
@@
 
-\chapter{State of the Art}
+\chapter{State of the Art of Data Persistence in Sensor Networks}
 
-In general, this chapter reviews surveys and related papers in networks
-networks, providing a basic introduction about their characteristics to new 
-audience. Similarly, it reviews the characteristics of related works that
-discusses and provides data persistence solutions for sensor networks, 
-characterizing the different approaches used to described and model the
-gathered data from sensor devices.
+In general, this chapter reviews surveys and related papers on sensor networks,
+providing a basic introduction about their characteristics to a new audience.
+Similarly, it reviews the characteristics of related works that discuss and
+provide data persistence solutions for sensor networks, characterizing the
+different approaches used to described and model the gathered data from sensor
+devices.
+
+%It is important to note that the scope of this literature review is limited to
+%our case study described in Chapter 4 and, therefore, it only
+%covers a subset of disciplines in sensor networks and data persistence.
 
@@
 collect data. The former is seen as a sensor device capable of producing data
-to the network based on an observation from the environment or simply a raw
-data. On the other hand, the latter is described as the final destination of
-the raw data in a sensor network and called the network data sink, or simply
-data sink, whose responsibility is to receive, categorize and store the
-``sensed'' data. In this way, the location of the the data sink is determined by
-the purpose of the collected data: data nodes are placed in-network close to
-the sensor devices when collected data has a life cycle, being discarded after
-being utilized. In contrast, data nodes are placed in in a centralized node
-when the collected data is used for historical purposes.
+to the network based on an observation from the environment. On the other
+hand, the latter is described as the final destination of the raw data in a
+sensor network and called the network data sink, or simply data sink, whose
+responsibility is to receive, categorize and store the ``sensed'' data. In
+this way, the location of the data sink is determined by the purpose of
+the collected data: data nodes are placed in-network close to the sensor
+devices when collected data has a life cycle, being discarded after being
+utilized. In contrast, data nodes are placed in a centralized node when the
+collected data is used for historical purposes.
 
@@
 
-Each sensor node perform different tasks for the network and, therefore,
-influences on the method of deployment used. \cite{sn-intro01} identifies that
+Each sensor node performs different tasks for the network and, therefore,
+influences the method of deployment used. \cite{sn-intro01} identifies that
 the deployment of sensor devices can use arbitrary methods such as being
 dropped from an airplane, or manually placed at chosen geographic locations.
-Moreover, those authors also mentioned that the deployment can be done as a
+Moreover, the cited authors also mentioned that the deployment can be done as a
 one-time event or as an iterative process, where the sensor devices are moved
@@
 
-Moving to a differnt point, \cite{sn-intro01} reports that sensor nodes can be
-able to move around the place where they were deployed or a stationary sensor
-at a given location. In the case of a sensor networks comprised with different
+As noted by \cite{sn-intro01} sensor nodes can be
+moved around the deployment location, or remain in one place as a stationary sensor. 
+In the case of a sensor network comprised with different
 moving sensors, a given sensor node can move occasionally or continuously, as
@@
 Considering the physical design and capabilities of sensor devices, they range
-from microscopically small to bigger devices of a few cubic feet
+from microscopic to devices measuring a few cubic feet
 \cite{sn-intro02}. In addition to their limited size, their computational
-resources and power availability are also scarse, when they are located in
+resources and power availability can also be affected, when they are located in
 areas of limited access to energy. In this way, \cite{sn-intro01} also
-distinguishes how sensor devices are usually designed to either store energy on
+distinguishes how sensor devices are usually designed to either store energy in
 batteries or use energy scavenged from solar cells. Even though a sensor
 network is designed for a purpose, \cite{sn-intro01} also explains that it can
-be comprised of different sensor devices playing different roles on what they
-observe in the environment, characterizing the heterogeneity the sensor
+be comprised of different sensor devices playing different roles in what they
+observe in the environment, characterizing the heterogeneity of the sensor
 network by the hardware capabilities and what types of data each of them
@@
 with each other using different media such as radio, diffuse light or
-inductive sound, influencing on the type of protocols they exchange data. They
+inductive sound, influencing the type of protocols they use to exchange data. They
 can have knowledge of their physical location if they are integrated with a
-GPS\footnote{Global Positioning System} device, which gives the latitude or
-longitude coordinate values. Therefore, their cost of these devices may range
-from hundreds to thousands of dollars, depends on a given device's functionality
-and capabilities of its design.
+Global Positioning System (GPS) device, which gives the latitude or
+longitude coordinate values. Therefore, the cost of these devices may range
+from hundreds to thousands of dollars, depending on the design capabilities and functionality of a given device.
 
@@
 
-Last section discussed the different properties of the sensor devices when
+The last section discussed the different properties of the sensor devices when
 placed in a sensor network. Depending on how they communicate with each other,
-\cite{sn-intro02} discusses that sensor devices can be in an organized in a
-structured or ad hoc way. For this reason, the infrastructure, of sensor
+\cite{sn-intro02} states that sensor devices can be organized in a
+structured, or ad hoc way. For this reason, the infrastructure of sensor
 networks differ in topology such as single-hop, star, tree or  graph. Each of
-these topologies may influence on how the produced data from sensor nodes is
+these topologies may influence how the produced data from sensor nodes is
 collected and routed to their related data sink. While there are only two
-levels of node connectivity in sensor networks deployed as single-hop and start
-topologies, sensor networks whose topology is either tree or graph use
+levels of node connectivity in sensor networks deployed as single-hop and
+star topologies, sensor networks whose topology is either tree or graph use
 multi-hop communication among each of their nodes for data communication from
@@
 specification of what needs to be covered by the sensor devices. Furthermore,
-considering that some types of sensor devices are very likely to regular
+considering that some types of sensor devices are very likely to experience regular
 hardware failures, \cite{sn-intro02} described sensor network examples with 
-redundant nodes that provide better data coverage and levels of coverage.
-Finally, the coverage of sensor devices also influence on the connectivity
-property of the sensor nodes. In order communicate with redundant nodes, a
-physical connection must be available at all times. On the other hand, lower
-level of coverage is is related sporadic communication among the sensor nodes.
+redundant nodes that provide better data coverage.
+Finally, the coverage of sensor devices also have influence on the connectivity
+property of the sensor nodes. In order to communicate with redundant nodes, a
+physical connection must exist at all times. On the other hand, lower
+level of coverage is related to sporadic communication among the sensor nodes.
 
@@
 sensor devices such as dozens, hundreds or even thousands of nodes. However,
-the life time of a sensor network determines how long these devices are going
+the lifetime of a sensor network determines how long these devices are going
 to be used, as well as how long the sensor network may exist. For example, an
 environmental sensor network \cite{sn-ex01} with the purpose of collecting data
-for historic purposes may be have their sensor devices continuously performing
-their data collection tasks. On the other hand, however, the lifetime of a
-sensor network designed to achieve a goal may not exist after the sensor
+for historic purposes may have their sensor devices continuously performing
+their data collection tasks. On the other hand, the lifetime of a
+sensor network designed to achieve a specific goal will not exist after the sensor
 devices finish their tasks. Finally, \cite{sn-intro02} argues that the
 performance related to the orchestration of sensor devices in sensor networks
-is related to the constraints defined for quality-of-service (QoS) such as
-real-time data delivery within a thredshold or for robustness such as
-maintaining data alive even though a communication link is not available.
+is related to the constraints defined for quality of service (QoS).  This includes
+real-time data delivery within a threshold or for robustness such as
+maintaining live data even though a communication link is not available.
 
-All things considered, sensor nodes play an important role during the execution
+All things considered, sensor nodes play a vital role during the execution
 of a sensor network, whose infrastructure is characterized by how and where the
-sensor nodes are deployed and communicate with each other. In this way, next
-section covers the approaches to deal with the product of this type of system
+sensor nodes are deployed and communicate with each other. The next
+section covers the approaches dealing with the problem of this type of system
 network.
@@
 
-As traced in the last few sections, sensor networks are comprised of
-different types of sensors nodes. However, an important sensor node type is the
-data sink because it is the place where the raw data produced by sensor
+As outlined in the last few sections, sensor networks are comprised of
+different types of sensor nodes. An important sensor node type is the
+data sink because it is where the raw data produced by sensor
 nodes resides. Data collection and access in sensor networks is directly
 related to the characteristics of the sensor network, its routing mechanisms
-and where the data sink is placed within the network. In a nutshell,
-\cite{sn-storage03} describes collected data as can be described as Low-level
-property or an observation such as the level of the ocean, etc.
+and where the data sink is placed within the network.
 
@@
 of the collected data. The in-network query processing is used when the
-location of the collected data is a local storage \cite{sn-storage01,
-sn-storage03} located in the network, strategy typically used to maximize the
+location of the collected data is a local storage site \cite{sn-storage01,
+sn-storage03} within the network; a strategy typically used to maximize the
 sensor's resources, while minimizing energy consumption \cite{sn-storage04}. On
 the contrary, the a typical centralized query processing is used to search for
-values over the collected data when its location is a centralized data
+values over the collected data when the location is a centralized data
 node \cite{sn-storage02}.
@@
 of the distributed data. Conversely, \cite{sn-provenance} proposes the use of
-Data Provenance as the primary guidelines to describe collected data from
-sensor devices, giving the fact that these devices only produce observations as
-stream of raw data, and therefore, described to address the problems related to
-the origin of the data.
+Data Provenance as the primary guidelines used to describe collected  raw data from
+sensor devices.
 
@@
 
-At first glance, the raw data serves as the main outcome of the sensor network.
-\cite{sn-provenance} differentiates the use of the collected data as follows as 
-\textbf{real-time data stream}, when they are generated data from a sensor
+As presented in the previous section, the raw data is the one in-transit from
+the sensor to the storage device, where it is labeled to be
+identifiable. It is also referred to as the collected data. 
+Initially, the raw data serves as the main outcome of the sensor network.
+\cite{sn-provenance} differentiates the use of \textbf{real-time data stream}, when they are generated from a sensor
 device whenever the data is requested. In this way, the data is still in its raw
-format specified by the sensor manufacturer; On the other hand, the authors
-describes the data as \textbf{archivable}, since they a have historical
-significance and will be processed to be reused by other systems in a latter
-time. While the real-time data streams are used and potentially discarded, the
-archivable data is stored into secondary memory such as a hard-disk. Under
-these circumstances, the collected data needs first to face its journey
-starting from its production, transmission from the sensor device, to finally
-reach a specified data sink in the network \cite{sn-storage01, sn-storage02,
-sn-storage03}.
+format specified by the sensor manufacturer. However, it is not retained. 
+In contrast, the authors describe the data as \textbf{archivable}, when they a
+have historical significance and will be processed to be reused by other systems at
+a latter time. While the real-time data streams are used and potentially
+discarded, the archivable data is stored into secondary memory such as a
+hard-disk.
 
 After producing the ``sensed'' values for the observed parameters from the
-environment, the sensor node is responsible to send the raw data to a specific
+environment, the sensor node is responsible for sending the raw data to a specific
 data node defined by the network infrastructure, as delineated in section
-\ref{sec:sn-infrastructure}. In such a way, \cite{sn-storage03}
-describes that the location for the collected sensor data provides two
+\ref{sec:sn-infrastructure}. Thus, \cite{sn-storage03}
+describes that the destination for the collected sensor data will be one of two
 different types of storage devices: the \textbf{external storage} and the
-\textbf{local storage}. In the former type, the sensor device sends the
-produced data to another sensor node that contains an external storage, whereas
-the data only exists in memory or between its reading to its delivery as
-real-time data stream in the the latter type. On the whole, the sensor network
-infrastructure dictates where the movement performed by the sensor data, as
-\cite{sn-storage02} sees a wireless sensor networks as a many-to-one data
+\textbf{local storage}. In the former type, the sensor sends the
+produced data to another sensor node that contains an external storage device.
+In the latter the data is stored only within the memory of the device before its 
+reading as a real-time data stream. On the whole, the sensor network
+infrastructure dictates where the movement performed by the sensor data,
+for example, \cite{sn-storage02} sees a wireless sensor networks as a many-to-one data
 collection pattern, when its organization is based on single-hop or star. On
@@
 
-Last, but not least, the use of different external storage devices can be used
+Lastly, the use of different external storage devices can be used
 to partition the collected data. This practice is called \textbf{Data-Centric
@@
 the sensor network with these characteristics, the query processor considers 
-all data nodes return its final response to the query. In comparison, a sensor
-network whose topology is in form of star, the data sink is usually located in
-the center of structure, and thus, receives all the collected data from other
-sensor devices in the network. For this reason, it can be seen as a 
-\textbf{centralized query processing}. Factoring in that a centralized data
+all data nodes before it returns its final response to the query. In comparison, a sensor
+network whose topology is in the form of a star, the data sink is usually
+located in the center of the structure, and thus, receives all
+the collected data from other sensor devices within the network. For this reason, it 
+can be seen as a \textbf{centralized query processing}. Factoring in that a centralized data
 sink receives all the collected data from the peer nodes, this query processing
-strategy is the solo responsible for responding to indirect queries about the
+strategy is responsible for responding to indirect queries about the
 produced data from the network.
 
-Based on the literature review, there are advantages and disadvantages related
+Based on the literature review, there are advantages and disadvantages related to
 each of the query processing approaches described with regards to the
 infrastructure of the sensor network. \cite{sn-storage03} argues that the 
-primary purpose of the in-network query processing is to minimize the number
+primary purpose of the in-network query processing is to minimize the amount
 of data transmitted from sensor devices to a data sink, as sensor devices are
 generally limited by energy use. Considering that sensor devices are also used
-as data nodes, \cite{sn-storage04} delineate that this approach can potentially
-flood all sensor nodes in the network, and therefore, decrease its performance.
+as data nodes, \cite{sn-storage04} states that this approach can potentially
+flood all sensor nodes in the network, thereby decreasing its performance.
 For this reason, \cite{sn-storage01} reports the use of the data-centric
-storage by organizing the collected data based on its particular properties
-values. In this way, the query processor hits specific data nodes that contains
-the types of data in the range of the requested data and, therefore, the author
-concludes that this approach decreases the network congestion and maximizes the
+storage by organizing the collected data based on its particular property
+values. In this way, the query processor hits specific data nodes that contain
+the types of data in the range of the requested data concluding that this approach 
+decreases the network congestion and maximizes the
 capacity on the sensor devices and the performance of the sensor network.
@@
 database system, located in one of the network nodes. For this reason, this
-indirect type of query processing is used when the data purpose of the use of
-data is for data archival. However, \cite{sn-storage02} showes that as a
-consequence of the many-to-one communication between the centralized data sink
+indirect type of query processing is used when the data purpose of the data is for 
+data storage. As a consequence, \cite{sn-storage02} shows that the result of
+a many-to-one communication between the centralized data sink
 and the other data nodes, this query process does not affect any of the sensor
@@
 this reason, \cite{sn-storage04} describes that different techniques used in 
-distributed system such as the use of master-slave data replication or 
+distributed systems such as the use of master-slave data replication or 
 database partitioning \cite{db-partitioning-relational} can be used to reduce
-or minigate the effects of this funneling effect.
+or mitigate the effects of this funneling.
 
@@
 
-Different approaches are used to represent the collected data from sensors. The
-use of basic data descriptors has been used in research groups which expertise
-is not database systems, while others use such technology in pragmatic
-approaches. \cite{sn-programming-language} reports different applications using
-different types of languages to acces the data modeled in different data
-models. \cite{sn-provenance} describes the use of a \textbf{Tabular Data Model}
-\cite{tabular-model} using comma-separated values to model and organized the
-values related to the properties of observed parameters from a given sensor
-device. On the other hand, \cite{sn-db-tinydb} describes the use of the
-\textbf{Relational Data Model} \cite{relational-model} is used in the majority
-of the projects that involves data persistence due to its popularity in other
-different areas of Software Engineering and Distributed Systems. Finally,
-\cite{sn-xml-usage01, sn-xml-usage02} are examples of sensor networks
-application that makes use of XML \cite{xml} to represent the collected data
-on a \textbf{Structured Data Model}.
+Different approaches are used to represent the collected data from sensors.
+Some research groups lacking database experience arbitrarily assign data 
+descriptors, while others use database technology in a more pragmatic approach.
+\cite{sn-programming-language} describes several applications using
+various types of languages for accessing the data designed in different data
+models. \cite{sn-provenance} reports on the use of a \textbf{Tabular Data Model}
+\cite{tabular-model} using comma-separated values (CSV) to organize the
+measurements collected from sensor devices. In addition, \cite{sn-db-tinydb} 
+describes the use of the \textbf{Relational Data Model} \cite{relational-model} 
+as used in the majority of the projects that involves data persistence due to its
+popularity in areas such as Software Engineering and Distributed Systems. Finally,
+\cite{sn-xml-usage01, sn-xml-usage02} discusses examples of sensor network 
+applications that make use of XML \cite{xml} to represent the collected data on a 
+\textbf{Structured Data Model}.
 
@@
 exported to the so-called comma-separated values (CSV) file format, the
-resulting file is comprised of lines of ascii characters, separated by a
+resulting file is comprised of lines of ASCII characters, separated by a
 delimiter. The first line represents a header, with the identification of
@@
 data model is commonly used to provide data persistence for sensor networks.
-As a matter of fact, the SF-BEAMS environmental sensor networks
-\cite{sfbeams2006} uses this approach to distribute the collected data over
-the Internet using the OPenDAP \cite{opendap} data format.
+The SF-BEAMS environmental sensor networks \cite{sfbeams2006} uses this
+approach to distribute the collected data over the Internet using the OPenDAP
+\cite{opendap} data format.
 
-This data model can be used in different context, as \cite{sn-provenance}
+This data model can be used in different contexts, as \cite{sn-provenance}
 describes that the versions of the same document can be easily managed and
-shared among research colleages using a version-control system such as
+shared among research colleagues using a version-control system such as
 Subversion \cite{subversion}.
@@
 \cite{sn-db-newop} shows the use of the Relational Data Model
-\cite{relational-model} can be used in sensor networks to describe the
-collected data from sensor devices. As a matter of fact, the project described
+\cite{relational-model} as it is used in sensor networks to describe the
+collected data from sensor devices. The project described
 by the authors is based on the TinyDB database system \cite{sn-db-tinydb}, a
 database system specifically built for data nodes of sensor networks. It uses
-the same properties and constraints of regular database system, where the
+the same properties and constraints of regular database systems, where the
 entities of the sensor network are required to be modeled and represented by
@@
 
-XML is an increasing data format used popularized by the applications developed
+XML is an increasing popular data format used widely by the applications developed
 for the Internet. \cite{sn-xml-usage01} reports an efficient way to represent
@@
 
-Other examples of the use of the structured data model using XML is the through
+Other examples of the use of the structured data model using XML is found in
 the use of middlewares, as the authors of \cite{sn-xml-query-engines} reports.
-Such approach is used for data exchange by the implementing of a middleware
+Such an approach is used for data exchange by the implementing of a middleware
 capable of query and extract the collected data from sensor devices
@@
 
-\section{Data Provenance: How Collected Data is Described?}
-\label{sec:sn-provenance}
+\section{Description of the Collected Data}
+\label{sec:sn-data-description}
 
@@
 properties of the data. Not to mention that the low-level or raw data produced
-by a given sensor device might not include vital information regarding the
-origing and identity of the collected data. For this reason,
+by a given sensor device might not include vital details regarding the
+origin and identity of the collected data. For this reason, the collected data
+is reported to be described in different ways such as Data Provenance or
+Annotations.
+
 \cite{sn-provenance} describes Data Provenance as a technique used to describe
 those essential properties of the collected data from sensors. One typical
-example of the application of Data Provenance is the so-called blackbox of an
-aircraft, which carries the collection of raw data collected from a group of
-sensors placed in different locations of an aircraft, as shown by
-\cite{sn-exemple-blackbox}. Last, but not least, the author describes the
-collected data as used in a process of data recovery and analysis in order to
-identify the probable reason of an aircraft accident, by using the time, and
-location of the aircraft, along with the independent variables from the
-sensors.
+example of the application of Data Provenance is the so-called black-box of an
+aircraft, which carries the collection of spatial-temporal data structures
+collected from a group of sensors placed in different locations of an
+aircraft, as shown by \cite{sn-exemple-blackbox}. Finally, the author describes 
+the collected data as used in the process of data recovery and analysis in order to
+identify the probable reason of an aircraft accident, by analyzing the time, 
+location of the aircraft, and other independent variables from the sensors.
 
-In general, sensor devices produce data and either keep them onto its local
+In general, sensor devices produce and preserve data onto their local
 storage or transmit them to data nodes as described earlier in this chapter. It
-is vital to understand that the majority of these devices outputs the data in
-its raw format, containing only a stream of values with any type of
-delimitation between each other. For this reason, the processes of parsing and
-description of each of the values located in the data streams regularly require
-the use of the device's specification documentation. Similarly, in order to
-have access to each of the properties of data collected by a given property
-related to time or space, \cite{sn-provenance} also mentions that 
-Metadata\footnote{Data about data} should be included together with the
-collected data.
+is vital to understand that the majority of these devices outputs the raw data
+in streams of values with any type of delimitation between each other. For
+this reason, the processes of parsing and labeling each of the values located
+in the data stream regularly requires the use of the device's specification
+documentation. Similarly, in order to have access to each of the properties of
+collected data by a given property related to time or space,
+\cite{sn-provenance} also mentions that Metadata\footnote{Data about data}
+should be included together with the collected data.
 
-An example of Data Provenance application is the use of aggregation of data
+An example of a Data Provenance application is the use of the aggregation of data
 over time to estimate the changing effects of the weather in a given region.
@@
 annotations to better describe the identity of the collected data, as 
-\cite{sn-provenance} reports. Interesting questions regarding the
+\cite{sn-provenance} reports. Pertinent questions regarding the
 data can be asked such as \textbf{what was collected?}, \textbf{When was the
@@
 Data Provenance provides basic mechanisms to describe sensor networks data in
-general, others see the importance of the time space of sensor networks.
+general, others see the importance of the time and position of sensor networks.
 \cite{sn-time-series} describes strategies used to cluster distributed sensor
 data as time series of data, as well as to decrease the size of data streams
-being transmitted in real-time sensor networks when these collected data does
+being transmitted in real-time sensor networks when these collected data do
 not need to be transferred \cite{sn-data-reduction}.
+
+A different way to describe sensor devices is through the use of
+Annotations. \cite{sn-annotation} shows the use of tags to describe data
+collected from a physical area monitoring. The sensor network was
+designed to monitor moving objects, capturing their location while in
+movement and their visual appearance using a video camera. In this way,
+different instances can be described by adding tags to particular video frames, and
+therefore, providing an efficient way to search for a specific frame based on
+its visual observations. Similarly, the above referenced project also uses data
+provenance to describe its collected data, since each of the collected data 
+contains the related GPS coordinates of the observed object.
 
@@
 different sensor devices and different sensor networks in the process called
-Information Fusison \cite{sn-info-fusion}. An example of the aggregation of
-different data from different devices located in different geographic sensor
-networks is \cite{sn-geo-metadata}, which is capable of process
-spatial-temporal queries of real-time or historical data. The following
-sections describe each of the metadata properties as defined by
-\cite{sn-provenance}, giving different examples of each of them.
+Information Fusison \cite{sn-info-fusion}. \cite{sn-geo-metadata} is an example of
+the aggregation of different data from different devices located in different 
+geographic sensor networks which is capable of processing spatial-temporal 
+queries of real-time or historical data. The following sections describe each 
+of the metadata properties as defined by \cite{sn-provenance}, giving different
+examples of each of them.
 
-\subsection{What was collected?}
+\subsection{Data as an object}
 
-Different sensor devices produce different sets of attributes based on its
+Different sensor devices produce different sets of attributes based on their
 capabilities and design. In general, these data sets do not have distinct
@@
 entities on the relational data model \cite{relational-model}. Equally,
-\cite{sn-provenance} describes that traditional sensor networks uses the
+\cite{sn-provenance} describes that traditional sensor networks use the
 collected data for analysis long after the attributes have been collected,
@@
 
-In a nutshell, the collected data is described by the name of the observation
+Briefly, the collected data is described by the name of the observation
 property. For instance, \textbf{temperature} = 73.45 expresses the value of the
@@
 the property describing the raw collected data 73.45. On the other hand,
-however, the example \textbf{temperature-scale} = ``Fahrenheit": describes
-the metadata regarding the temperature.
+the example \textbf{temperature-scale} = ``Fahrenheit": describes the metadata
+regarding the temperature.
 
-\subsection{When was the data collected?}
+\subsection{Time of Data Creation}
 
 In order to use the collected data for the purpose of historical analysis, the
-use of a point in time is the most prominent feature of data, as it is
-considered a pivotal metadata for different types of sensor networks such as
-an aircraft the blackbox \cite{sn-exemple-blackbox} or any
-environmental sensor network \cite{sfbeams2006, sn-xml-middleware, sn-ex01}.
-In this way, Other projects classify data in time-series
-\cite{sn-time-series-example}, where data is related to all the
-observations a given sensor device performed over a period of time. The
-apparent implications is that all the produced data is necessary, but
-\cite{sn-time-series} argued that some sorts of sensor networks could decrease
-the creation of ``noisy data sets" based on the proper clustering of
-similar time-series data before before they are sent to the data sink.
+use of a point in time is the most prominent feature of the data. Also referred to
+as spatial-temporal data, it is considered a pivotal metadata used in different
+types of sensor networks such as an aircraft the black-box
+\cite{sn-exemple-blackbox}, environmental sensor network \cite{sfbeams2006},
+and data communications management systems \cite{sn-dataware-house}. In this
+way, Other projects classify data in time-series \cite{sn-time-series-example},
+where data is related to all the observations a given sensor device performed
+over a period of time. The apparent implications is that all the produced data
+is necessary, but \cite{sn-time-series} argued that some sorts of sensor
+networks could decrease the creation of ``noisy data sets" based on the proper
+clustering of similar time-series data before they are sent to the data
+sink.
 
@@
 it was collected to the data sink. Finally, \cite{sn-provenance} shows that
-the collected data for archival purpose is usually saved in self-described
+the collected data for archival purpose is usually saved in self-labeld
 files described by file names, containing complex naming conventions using the 
@@
 specific location where the observation occurred. It is important to note,
-however, that those type of data do not necessarily need to be on the data
+however, that those types of data do not necessarily need to be on the data
 model used, but the specifications of the use of the data determines which
@@
 at using data from different sensor devices.
+
+\section{Data Load and Management Systems}
+\label{sec:data-load}
+
+Depending on the amount of data produced by a sensor network, different
+approaches are taken to support the data management depending on
+the amount of data collected. For example, \cite{sfbeams2006} is an
+environmental sensor network that produces data regarding marine life.
+Less than twenty sensor devices produce Megabytes of data on a daily basis,
+whose data is processed and stored using flat-files directly in the
+operating system's file system with provenance information regarding the time
+of collection.
+
+When the amount of data collected grows to other scales, (i.e. Gigabytes) of data
+on a daily basis, data becomes more difficult to manage in any type of
+database systems. Examples of sensor networks that produce data under this
+magnitude are reported in the literature. For instance, 
+\cite{sn-dataware-house} describes the creation of a Data Warehouse for the
+real-time sensor data management using a cluster of MySQL databases used to
+store data from different optical sensors that collect metrics
+from physics experiments using eletromagnetic fields. As the author reported,
+requirements for data management evolved from storing flat files by a
+monolitic system to a robust distributed system using MySQL database.
+Moreover, the system targeted from 5 to 10MB of data per second, whose
+cumulative data reached 1 TB of data every year.
+
+Data management for large scales of data require different techniques to store
+and retrieve data, since it requires more computational power and data
+retrieval efficient algorithms. \cite{sn-data-center} describes the creation of
+a Sensor Data Center designed to store and analyze amounts of data on the range of
+Terabytes to a future Petabytes of data. Its sensor network deals with not only
+numerical sensor readings from transportation management, but also radar
+images, used for historic purposes and data mining that performs traffic
+pattern analysis. In this way, the authors report the use of parallel data
+processing techniques in order to provide processing for the collected
+data, and load-balancing when storing data to minimize the sensor data
+processing delays. For this reason, they suggest the use of concurrent control
+such as snapshot isolation of the current data into separated replicated data.
+Therefore, data access from users are not interrupted by the processing of
+the data that is processed.
+
+In summary, this chapter described a subset of problems and solutions used by
+the sensor networks community. First, the survey in sensor networks was
+reviewd to give a foundation of specific properties of sensor networks. Then,
+different surveys in data persistence in sensor networks were discussed,
+followed by the data models and query engines. In the end, the most important 
+discovery is that data provenance provides the best guidelines for describing 
+data. The following chapter will classify the properties found in this chapter
+in terms of taxonomies.

### 2009-12-09 docs/marcello-dissertation/thesis/chapter2-state-of-the-art.tex
This revision incorporates the revisions from Niki for chapter 3, and my own corrections to chapters 1 and 2 using MS word to find misspelled words. Not it is a clean version.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@687 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 redundant nodes that provide better data coverage.
-Finally, the coverage of sensor devices also have influence on the connectivity
+Finally, the coverage of sensor devices also has influence on the connectivity
 property of the sensor nodes. In order to communicate with redundant nodes, a
@@
 delimiter. The first line represents a header, with the identification of
-columns of specific data, where the following lines contains the values
+columns of specific data, where the following lines contain the values
 related to each top column. \cite{sn-provenance} reports that the tabular
@@
 Other examples of the use of the structured data model using XML is found in
-the use of middlewares, as the authors of \cite{sn-xml-query-engines} reports.
+the use of middleware, as the authors of \cite{sn-xml-query-engines} reports.
 Such an approach is used for data exchange by the implementing of a middleware
@@
 to communicate between nodes is the communication mechanism used by NetBEAMS
-\cite{netbeams2009}. See section \ref{sec:dsp-serialization} for details.
+\cite{netbeams2009}. See section the documentation at
+\cite{netbeams-dsp-architecture} for details.
 
@@
 storage or transmit them to data nodes as described earlier in this chapter. It
-is vital to understand that the majority of these devices outputs the raw data
+is vital to understand that the majority of these devices output the raw data
 in streams of values with any type of delimitation between each other. For
@@
 movement and their visual appearance using a video camera. In this way,
-different instances can be described by adding tags to particular video frames, and
+adding tags to particular video frames can describe different them, and
 therefore, providing an efficient way to search for a specific frame based on
@@
 capabilities and design. In general, these data sets do not have distinct
-names, which leads to difficulties when indices over the properties of the
+names, which lead to difficulties when indices over the properties of the
 collected data are required to make them searchable. Specially for the purpose
@@
 the property describing the raw collected data 73.45. On the other hand,
-the example \textbf{temperature-scale} = ``Fahrenheit": describes the metadata
+the example \textbf{temperature-scale} = ``Fahrenheit" describes the metadata
 regarding the temperature.
@@
 
-\subsection{From where was the data collected?}
+\subsection{Data Origin}
 
@@
 processing techniques in order to provide processing for the collected
-data, and load-balancing when storing data to minimize the sensor data
+data, and load balancing when storing data to minimize the sensor data
 processing delays. For this reason, they suggest the use of concurrent control
 such as snapshot isolation of the current data into separated replicated data.
-Therefore, data access from users are not interrupted by the processing of
+Therefore, data access from users is not interrupted by the processing of
 the data that is processed.

### 2009-12-09 docs/marcello-dissertation/thesis/chapter3-taxonomic-classification-of-persistence-storage-for-sensor-networks.tex
This revision incorporates the revisions from Niki for chapter 3, and my own corrections to chapters 1 and 2 using MS word to find misspelled words. Not it is a clean version.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@687 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 
-As revailed in the previous chapter, data persistence for sensor
+As reveled in the previous chapter, data persistence for sensor
 networks require an understanding of not only the properties of the
 infrastructure of a given sensor network, but also of the properties and nature
-of the collected data themselves. To put things differently, in order to
+of the collected data themselves. In order to
 identify these concepts more clearly, each section of this chapter proposes an
@@
 
-A taxonomy is defined as the practice and science of classification and uses
-the taxonomic units of ``taxa'', the plural form of ``taxon''. Similarly, they
-are represented by a hierarchy diagram that relates the main taxonomy to the
-different taxa and their own related ones, if any.
+Taxonomy is defined as the practice and science of classification and uses
+the taxonomic units of ``taxa'', the plural form of ``taxon'' \cite{taxonomy}.
+Similarly, they are represented by a hierarchical diagram that relates the
+main taxonomy to the different taxa and their own related ones, if any.
 
@@
 retrieved and used in the network. In such a fashion, this taxonomy relates to
-how the collected data is used, and should be the primary question to be
-answered during the analysis of a sensor network. The taxa related to this
+how the collected data is used, and should be the primary question answered 
+during the analysis of a sensor network. The taxa related to this
 taxonomy, depicted in Figure \ref{fig:taxonomy-data-purpose}, is described as
@@
 in the different aspects of the access to the collected data from sensor
-devices, as shown in section \ref{sec:sn-storage-locations}. In this way,
-Figure \ref{fig:taxonomy-data-location} depicts this taxonomy, which can be
+devices, as shown in section \ref{sec:sn-storage-locations}. Figure 
+\ref{fig:taxonomy-data-location} depicts this taxonomy, which can be
 described as follows:
@@
 \begin{itemize}
-  \item \textbf{Local Storage}: characterizes the placed of the collected data
+  \item \textbf{Local Storage}: characterizes the place the collected data is
   temporarily located in-memory, or in a secondary storage device on a node
@@
   external storage device with a data management system such as a relational
-  database. When the collected data occupies different external storages, the
+  database. When the collected data occupies different external storage devices, the
   strategy is called \textbf{Data-Centric Storage}, being organized by
@@
 in sensor networks is related to the data model chosen to represent the
-collected data. Different data models were reported in the review of related
-literature in section \ref{sec:data-models}. As a consequence, this taxonomy
-relates to the data model used to design the collected data, as drawn in Figure
-\ref{fig:taxonomy-data-model}, and described as follows.
+collected data. This taxonomy relates to the data model used to design the 
+collected data, as drawn in Figure \ref{fig:taxonomy-data-model}, and is 
+described as follows.
 
@@
   \item \textbf{Schema-Dependent Models}: this data model requires the
-  definition of a master schema that describes the data through a rigorous of 
-  data modeling process prior to the insertion of the data. Examples of such
+  definition of a master schema that describes the data through a rigorous
+  data modeling process prior to the insertion of the data. Examples of this
   model is the traditional Relational Data Model\cite{relational-model}, as
@@
   data models do not require the definition of a data schema or table
-  definition. Instances of such data model reviewed in the literature review
-  is the use of the Tabular Data model.
+  definition. Instances of such a data model is the use of the 
+  Tabular Data model.
 \end{itemize}
@@
 
-Particulaly related to how the collected data is described, different projects
+Particularly related to how the collected data is described, different projects
 provide data persistence describing the collected data by using its natural
-properties, or associating textual descriptions or keywords of the data. The
-way to describe data can follow different guidelines such as data provenance
-and annotations, discussed in section \ref{sec:sn-data-description}. As
-reported in section \ref{sec:sn-data-description}, the description of the
-properties should take into account different properties of the nature of the
+properties, or associating textual descriptions or keywords for the data. The
+way to describe or label data can follow different guidelines such as data provenance
+and annotations, discussed in section \ref{sec:sn-data-description}. The description of the
+properties should take into account the different values of the nature of the
 data and any other description needed. Therefore, the taxa that composes the
@@
   that uniquely identifies the collected data from sensors. Similarly, the
-  \textbf{When: Time Dimension}, is related instants of time related to the
+  \textbf{When: Time Dimension}, is related to instances of time relative to the
   collected data: \textbf{fact} and \textbf{transaction} times. The former is
@@
   use of keywords, full-text description, or any other method that helps
-  describing the collected data. It is important to note that this type of data
-  does not make part of the collected data from sensors. Rather, they
-  complement the observations of what was percieved by the sensor devices.
+  describe the collected data. It is important to note that this type of data
+  does not makeup part of the collected data from sensors. Rather, they
+  complement the observations of what was perceived by the sensor devices.
 \end{itemize}
@@
 and where the collected data is used and stored, as shown in section
-\ref{sec:query-process}. As a consequence, the following taxa are related to the
+\ref{sec:query-process}. The following taxa are related to the
 different types of query processing mechanisms and summarized in Figure
@@
   sensor device itself, and it uses the local storage mechanism, then the
-  sensor node is able to reply to request related to the observations of the 
+  sensor node is able to reply to the request related to the observations of the 
   environment defined by the sensor device. In this way, the query takes 
@@
 When it comes to the data volume produced during a time frame, different ranges
-were reported from different types of sensor networks. Depending on the types
+are reported from different types of sensor networks. Depending on the types
 of data produced such as numerical or images, they can be in the orders of
@@
 \begin{itemize}
-  \item \textbf{Small}: sensor networks whose yearly produced data is in the
+  \item \textbf{Small}: sensor networks whose annual production of data is in the
   range of MBs or to 1GB of data, usually stored in flat files directly in the
@@
   generate numerical data;
-  \item \textbf{Medium}: amount of data that ranges from GBs to 1TB data. In
-  order to manage such amount of data using disk arrays;
-  \item \textbf{Large}: when the amount of data produced by a sensor device
-  reaches amounts of data compatible to data centers, in the orders ranging
-  from TBs to 1PT of data. The construction of a data center specialized for
-  data produced by sensor devices, taking into account the addition of data
-  provenance, among others.
+  \item \textbf{Medium}: amount of data that ranges from GBs to 1TB data and is 
+  managed using disk arrays;
+  \item \textbf{Large}: the amount of data produced by a sensor device ranging 
+  from TBs to 1PT of data necessitates the construction of a data center 
+  specializing in the storage of data produced by sensor devices, taking into 
+  account the addition of data provenance, among others.
 \end{itemize}
@@
 
-Different approaches of are developed computational systems to deal the data
-produced from sensor networks. As  
-In this way, Figure \ref{fig:taxonomy-database-architecture} describes the
-taxonomy of the database system organization as follows:
+Different approaches have been developed to organize the storage of data
+produced from sensor networks. Figure \ref{fig:taxonomy-database-architecture} 
+describes the taxonomy of the database system organization as follows:
 
@@
   \item \textbf{Centralized System}: a database system that runs in one single
-  centralized host, being the focal purpose managing the collected data
+  centralized host, the focus being the managing of the collected data
   \cite{sn-intro01};
-  \item \textbf{Distributed System}: a database system that can be composed by
+  \item \textbf{Distributed System}: a database system that can be composed of
   more than one node. Usually, the use of the pattern of Master-Slave
-  mechanism or data ``Replication" is used to aliviate ``hot spots'' during
+  mechanism or data ``Replication" is used to alleviate ``hot spots'' during
   concurrent processes of data collection from the sensor devices and data
-  access from the final users. On the other hand, unique mechanisms can be
-  used to disagragate the produced data from a concentrated one: the use 
-  of database partioning, as related to the data-centric storage, where  
-  each partition is responsible to hand data produced by a given type of 
-  sensor, for instance.
+  access from the final users. Unique mechanisms can be used to disaggregate 
+  the produced data from a concentrated one: the use of database partitioning, 
+  as related to the data-centric storage, where each partition is responsible 
+  to handle data produced by a given type of sensor, for instance.
 \end{itemize}
@@
 On the whole, the taxonomies described in this chapter were based solely on the
-literature review of described in the previous chapter. They can be helpful in 
-understanding of requirements of a data persistence layer for sensor networks
-with different infrastructures. They are summarized on table
-\ref{tab:taxonomies-list-summary}, and briefly related highlight these
+literature review described in the previous chapter. They can be helpful in 
+understanding the requirements of a data persistence layer for sensor networks
+with different infrastructures. They are summarized on Table
+\ref{tab:taxonomies-list-summary}, and briefly highlight these
 relationships in order to better analyze a problem related to data persistence.
@@
 
-The taxonomy related to the purpose of the sensor data can be directly related
-to the taxonomy of the location of the sensor data. While the taxon of the 
+The taxonomy related to the purpose of the sensor data can be directly associated 
+with the taxonomy of the location of the sensor data. While the taxon of the 
 real-time data stream relates to the local data storage, the taxon related to
-data archival correlates the storage of the collected data with the an external
+data archival correlates the storage of the collected data with an external
 storage one. In addition to that, the converse is also true.
 
-The taxonomy related to the data model can be related to any taxon from the
+The taxonomy related to the data model can be linked to any taxon from the
 taxon described from the system organization, to data description and query
@@
 separate the data being written and read can be used such as Load Balancing or
-Database Partition. As a consequence, the use of a Data-Centric storage
-approach is used.
+Database Partition, and the use of a Data-Centric storage approach is used.
 
@@
 properties of a sensor networks can be studied and reviewed. As described in
-the introduction of this dissertation, one the main motivations for this work
+the introduction of this report, one the main motivations for this work
 was to provide means of data persistence for NetBEAMS, a case study which does

### 2009-12-09 docs/marcello-dissertation/thesis/chapter4-netbeams-in-depth-requirements-analysis-and-specification.tex
Update of chapter 5 after Niki's reviews and my verification using MS Word.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@688 559edde2-a4d4-44e0-96fc-4035e40d7396

--- 2009-12-07 docs/marcello-dissertation/thesis/chapter4-netbeams-in-depth-requirements-analysis-and-specification.tex
@@
 
-As discussed in chapter 2, different properties of a sensor network and the
+As discussed in Chapter 2, different properties of a sensor network and the
 nature of the collected data must be taken into account in order to provide a
-data persistence layer for a given sensor netwokr. Similarly, in order to
-better assist one analysis of such functionality, chapter 3 proposed a set of
+data persistence layer for a given sensor network. Similarly, in order to
+better assist one's analysis of such functionality, Chapter 3 proposed a set of
 data persistence taxonomies related to different properties of the collected
-data life cycle. In this way, the goal of this chapter is to describe the
-fundamental properties of the case study used for this work, that is, NetBEAMS,
-in order to ``set the stage'' for the selection of a database technology for it.
-In this way, given the fact that NetBEAMS provides an automated infrastructure
-solution for SF-BEAMS, this last component will be covered first. Then, this
-work proposes a categorization of NetBEAMS to the face of the taxonomies, and
-finally discusses the discusses the requirements for a data persistence.
+data life cycle. The goal of this chapter is to describe the fundamental properties 
+of the case study used for this work, that is, NetBEAMS, in order to lay the 
+groundwork as it relates to the selection of the database technology. NetBEAMS 
+provides an automated infrastructure solution for SF-BEAMS, and this last component 
+will be covered first. Then, this work proposes a categorization of NetBEAMS to 
+the face of the taxonomies.
 
@@
 department of Biology and Computer Science at San Francisco State University,
-whose goal was to automate the operational execution of SF-BEAMS. On the other
-hand, the San Francisco Bay Environmental Assessment and Monitoring Station, or
-SFBEAMS \cite{sfbeams2006}, is an instance of an environmental sensor
-network whose primary focus is the study of complex marine and estuarine
-environments using the SF-BEAMS sensors. The network is deployed offshore of
-its pier located on the banks of the San Francisco Bay at the Tiburon Island,
-California, being operated by the department of Biology from San Francisco
-State University. In this way, NetBEAMS is an experimental component that
-utilizes the SF-BEAMS infrastructure to operate. This section details SF-BEAMS
-in general, and presents the requirements for data persistence.
+whose goal is to automate the operational execution of SF-BEAMS. The San Francisco
+Bay Environmental Assessment and Monitoring Station, or SFBEAMS \cite{sfbeams2006}, 
+is an environmental sensor network system whose primary focus is the study of complex 
+marine and estuarine environments using the SF-BEAMS sensors. The sensors are deployed off a
+pier located on the San Francisco Bay, in Tiburon, California; and is operated by 
+the Romberg Tiburon Center for Environmental Studies (RTC) a part of San Francisco
+State University. NetBEAMS is an experimental component that utilizes the SF-BEAMS 
+infrastructure to operate. This section details SF-BEAMS in general, and presents 
+the requirements for data persistence.
 
@@
 
-In general, the SF-BEAMS network infrastructure contains varying a number of
-wired and wireless devices standing on pylons docked at the pier. Each of them
+In general, the SF-BEAMS network infrastructure contains a varying number of
+wired and wireless devices attached to pylons at the pier. Each of them
 is responsible for observing different conditions of the area, having its own
@@
 with a laptop computer. The current data collection process for SF-BEAMS is
-described in figure
+described in Figure
 \ref{fig:SF-BEAMS-system-architecture}\cite{sfbeams-current-system}. Upon
@@
 different formats. One of these formats is the OPeNDAP \cite{opendap}, which
-is widely used at research institutions that promote easier data exchange
-among them. The SF-BEAMS website provides access to the collected data through
+is widely used at research institutions to promote easier data exchange
+amongst them. The SF-BEAMS website provides access to the collected data through
 the Internet at http://sfbeams.sfsu.edu:8080/opendap. An example of an HTTP
-Request from accessing the data for a specific YSI is shown in Listing
-\ref{file:rtc-ysi-opendap}.
+Request from accessing the data for a specific YSI \cite{YSI-Sonde} is shown in
+Listing \ref{file:rtc-ysi-opendap}.
 
@@
 As described in section \ref{sec:sn-infrastructure}, sensor devices produce
-its observed data based on properties of measurements defined by its
-manufacturer. NetBEAMS has used ``YSI 6600 ESD V2", as seen in Figure
-\ref{fig:ysi-device}, one of the sensors devices used by SF-BEAMS. It is a
-powerful water quality monitoring device, capable of producing around 52 bytes
+the observed data based on properties of measurements defined by its
+manufacturer. NetBEAMS has used ``YSI 6600 ESD V2" \cite{YSI-Sonde}, as seen in
+Figure \ref{fig:ysi-device}, one of the sensors devices used by SF-BEAMS. It is a
+powerful water quality-monitoring device, capable of producing around 52 bytes
 on a single real-time data stream reading, as shown on Table
@@
   \includegraphics[scale=0.7]{../diagrams/ysi-device}
-  \caption{Picture of the case study sonde device: YSI 6600 ESD V2}
+  \caption{Picture of the case study Sonde device: YSI 6600 ESD V2}
   \label{fig:ysi-device}
@@
 In order to have access to the observed data, the network staff downloads the
-data using one of the device's connection such as the RS-232 serial connector
-\cite{rs232}. Then, the RTC labs receives the collected data transferred from
-the sensor devices to be used for archival purposes. According to one of its
-staff members, the infrastructure of the SF-BEAMS in was comprised of
-the following components (as of May 2009):
+data using one of the device's connections such as the RS-232 serial connector
+\cite{rs232}. Then, they upload the data into their data servers where they are
+processed and archived for historical analysis. According to one
+staff member, the infrastructure of the SF-BEAMS is comprised of the 
+following components (as of May 2009):
 
@@
   15 minutes, depending on the specification;
-  \item The processed data, used for distribution over the Internet, is
+  \item The processed data, used for distribution over the Internet,
    contains information regarding the time of the data collection.
@@
 
-Considering the infrastructure aforementioned, the data load in the server-side
-of the RTC's laboratories produced by the YSI sonde sensors devices can be
+Considering the infrastructure, the data load in the server-side
+of RTC's laboratories produced by the YSI Sonde sensors devices can be
 estimated by analyzing the data distribution over a period of one year, as
@@
     \label{tab:ysi-data-distribution}
-    \caption{Amount of data produced by the RTC's YSI sondes}
      \begin{center}
@@
         \textbf{YSIs} & \textbf{Rate} & \textbf{Hourly} & \textbf{Daily} &
-        \textbf{Weeky} & \textbf{Monthly} & \textbf{Yearly}\\\hline 
+        \textbf{Weekly} & \textbf{Monthly} & \textbf{Yearly}\\\hline 
         1 & 1 min & 3.04 Kb & 73.12 Kb & 511.87 Kb & 1.99 Mb & 23.99 Mb\\\hline 
@@
       \end{center}
+    \caption{Amount of data produced by the RTC's YSI sondes}
 \end{table}
 
-By analyzing the maximum amount of produced data by the YSI sonde device is
-around hundreds of Megabytes a year, as \textbf{483,840} samples are generated
-by one single device.
+The maximum amount of data produced by the YSI Sonde device is hundreds of Megabytes
+a year, as \textbf{483,840} samples are generated by one single device.
 
@@
 \begin{itemize}
-  \item \textbf{The Purpose of Sensor Data}: the data from SF-BEAMS is solely
-  used for the purpose of Data Archival, since they are stored in the RTC's
-  lab;
+  \item \textbf{The Purpose of Sensor Data}: the data from SF-BEAMS is exclusively
+  used for the purpose of Data Archival, since they are stored in RTC's lab;
   \item \textbf{The Location of the Sensor Data}: it is clear that the purpose
-  of data is directly related to its location and, as a consequence, the
+  of the data is directly related to the sensors location and, thus, the
   strategy of \textbf{External Data Storage} is used to store its collected
-  data at the network sink at the RTC's labs;
-  \item \textbf{Data Model}: since the format uses by the RTC staff to share
-  data among researcher is in the OPEnDAP format, it is clear that the data
+  data at the network sink at the RTC site;
+  \item \textbf{Data Model}: since the format used by the RTC staff to share
+  data with researchers is in the OPeNDAP format, the data
   model used is the tabular data model, since it uses the comma-delimited
-  files. In this way, the \textbf{Schema-less} is used;
+  files. In this way, the \textbf{Schema-less} model is used;
   \item \textbf{Data Description}: once the collected data reaches the RTC
   lab, a quality assurance process is executed, and the data is transformed
-  into the format used by OPEnDAP. This conversion carries data regarding
-  the time of the data collection, and therefore, the use of one of the
-  \textbf{Time Dimensions} such as the valid time. Since the files are stored
+  into the format used by OPeNDAP. This conversion carries data regarding
+  the time of the data collection, therefore, one of the
+  \textbf{Time Dimensions} such as the valid time is used. Since the files are stored
   with specific file name structure, using timestamps to describe the data, the
@@
   data is stored on file system;
-  \item \textbf{Data Volume}: the volume of data produced is in the range of
-  small. It is related to the volume of data produced by the sensor devices,
+  \item \textbf{Data Volume}: the volume of data produced is characterized by small.
+ It is related to the volume of data produced by the sensor devices,
   which is numerical data streams, and the total number of sensor devices;
@@
   the collected data is stored directly in the file-system after previous 
-  processing. However, they provide the collected data to the Internet using
+  processing. However, the collected data is provided to the Internet using
   the OPeNDAP Hyrax system as a middleware.
@@
 
-Although the execution of the RTC's sensor network can be operated as it
-described in the previous sections, \cite{netbeams2009} described some of the
-operational challenges faced by the RTC staff during regular activities of
+Although the execution of RTC's sensor network can be operated as it
+was mentioned in the previous sections, \cite{netbeams2009} described some of the
+operational challenges faced by the RTC staff during regular activities of the
 data collection process. It was clear to the research group that SF-BEAMS
@@
 open-source \cite{open-source} software, the research group developed a second
-version of NetBEAMS, a component-based approach suggested improving operational
-activities from SF-BEAMS using systems automation. In brief, one of the open
-problems of NetBEAMS was regarding Data Persistence, the primary motivation of
-this dissertation.
+version of NetBEAMS, a component-based approach which suggested the improvement of 
+operational activities from SF-BEAMS using systems automation. In brief, one of the 
+ongoing problems of NetBEAMS was regarding Data Persistence, the primary motivation of
+this report.
 
@@
 offers the Data Sensor Platform (DSP) \cite{netbeams2009} as the system
-architecture that can address the operation of SF-BEAMS. In this way, in-depth
+architecture that can address the operation of SF-BEAMS. The in-depth
 documentation describing the DSP Platform, including its architecture and data
-gathering process, is provided startng in the appendix section
-\ref{sec:dsp-details}.
+gathering process is provided at the online documentation 
+\cite{netbeams-dsp-architecture}.
 
@@
 users such as Biologists from the RTC laboratories. The requirements for the
-data persistece specifically for NetBEAMS is described in the following section.
+data persistence specifically for NetBEAMS are described in the following
+section.
 
@@
 approach is that it can potentially reduce operational costs and improve the
-data collection process. In review of this, the scope of a persistence layer
+data collection process. In view of this, the scope of a persistence layer
 can be summarized as a group of functional and non-functional requirements in
@@
   \item The data model used to describe the data must not impose restrictions
-  to the users of the system: Biologists and students without expertise in
-  Database Systems;
-  \item Data Representation must be similar to the ones used by RTC, with a
+  to the users of the system (i.e. Biologists and students without expertise in
+  Database Systems);
+  \item Data Representation must be similar to those used by RTC, with a
   more human-readable format;
   \item The system must be scalable with a low degree of maintenance, in a
-  way that the data gathering process does not impose less service interruption;
-  \item Data must be searchable in near-real-time with outstanding performance;
-  \item The solution must also give support to export capabilities, which can
-  be possibly used to match the RCT's requirements of the OPeNDAP formatt;
-  \item The database system must be free of charge, following the
+  way that the interruptions to the data gathering process are infrequent;
+  \item Data must be searchable in near-real-time with good performance;
+  \item The system must also support export capabilities, which may
+  be used to match RCT's requirements of the OPeNDAP format;
+  \item The database must be free of charge, following the
   implementation specifications of NetBEAMS for Open-source software.
@@
 
-After presenting NetBEAMS and its infrastructure, the the next step is the
-actual analysis of existing database systems that fits to the requirements
-and characteristics of NetBEAMS. Meanwhile, more details regarding the
-architecture of NetBEAMS, as well as development guidelines, can be reached in
-the appendix of chapter 10.
+In conclusion, after detailing the SF-BEAMS sensor network, and describing the 
+infrastructure of NetBEAMS, its the motivation and goals for this project were 
+presented. The next chapter is the actual analysis of existing database systems 
+mentioned in Chapter 2, which aligns with the requirements and characteristics of 
+NetBEAMS. Meanwhile, more details regarding the architecture of NetBEAMS, as well as
+the development guidelines can be found in the appendix of chapter 10.

### 2009-12-10 docs/marcello-dissertation/thesis/chapter10-appendix.tex
Reviews of chapter 4 of the report... Adding a label on chapter 2 and adding references to non-cited sections.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@689 559edde2-a4d4-44e0-96fc-4035e40d7396

--- 2009-12-07 docs/marcello-dissertation/thesis/chapter10-appendix.tex
@@
-\chapter{Appendix}
+\chapter{Appendix: Implementation and Experiments Artifacts}
 
-This section includes the source-code for the DSP Data Persistence. Each
-section details one single main artifact used on the development of the
-component. However, only the main ones will be added into this document, while
-the remainder can be downloaded directly from the Subversion repository.
+This chapter includes references to the implementation Data Data Persistence
+Implementation covering its design and architecture. Also, some of the
+implementation details of the DSP Data Persistence component designed in
+chapter 6 are presented, showing the generated experiment artifacts. While
+this section does not include all the related source-codes, the remainder can
+be downloaded directly from the Subversion repository branch at
+http://code.google.com/p/netbeams/source/browse/\#svn/branches/marcello/persistence/.
 
-\section{The Data Sensor Platform (DSP)}
-\label{sec:dsp-details}
-
-The Data Sensor Platform, in short DSP Platform, is based on a micro-kernel
-architecture developed on top of the Java modular framework called OSGi
-\cite{osgi}. The platform represents the foundation of the execution of
-NetBEAMS, since it is executed from an embedded hardware called Gumstix
-\cite{gumstix} using a Linux operating system and on top of the Java Virtual
-Machine (JVM). In order to transport data from sensors hosts to the network
-sink, the use of a cellular connection is used to transport the collected data
-from the senso devices using the HTTP Protocol, as described by
-\citet{netbeams2009}. Figure \ref{fig:sf-netbeams-node} shows the
-architecture of the inhanced node designed by the NetBEAMS research group.
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[scale=0.5]{../diagrams/DSP-Gateway-Node}
-  \caption{The NetBEAMS Node enhancement with the Gumstix and the GSM Modem}
-  \label{fig:sf-netbeams-node}
-\end{figure}
-
-The following list describes each of the devices in the NetBEAMS Node.
-
-\begin{itemize}
-  \item \textbf{YSI 6600EDS V2}: the YSI device is the regular sonde
-  responsible for observing the environment. It is connected to the Gumstix
-  embedded system, where the DSP is executed;
-  \item \textbf{Gumstix console-vx}: it's a COTS ARM-based hardware that
-  provides a computer-on-module environment for the development of small
-  embedded systems. This is the main environment of data extraction and remote
-  processing is accomplished. It runs a cross-compiled version of Linux Kernel
-  2.6;
-  \item \textbf{Huawei E220 USB Modem}: it is used to transfer the data from the
-  Gumstix to the RTC Labs Data Center.
-\end{itemize}
-
-The DSP runs inside of the embedded system Gumstix. The following sections
-details its design and process for extracting data.
-
-\subsection{The DSP System Design}
-
-This section focus on the development of a Software Platform for the NetBEAMS
-Gateway Embedded System, or on the system developed in the Gumstix device. As
-shown in figure \ref{fig:netbeams-software-stack}, the NetBEAMS Gateway Node is
-an embedded system that contains a set of layer found in any computer.
-However, each of the resources described from the system were customized for
-the NetBEAMS system. The main components of such system can be summarized as
-follows:
-
-\begin{itemize}
-  \item \textbf{Operating System}: it uses a cross-compiled Gentoo Linux, which
-  supports a wide variety of development tools, including the Java Virtual Machine
-  (JVM), the underlying platform system;
-  \item \textbf{Java Virtual Machine}: it uses the JamVM, a ~200 KB version of
-  the Sun Microsystems Java Virtual Machine, version 2.0;
-  \item \textbf{OSGi Framework}: it uses the Knopflerfish implementation of the
-  OSGi 4.1 specification. More details in the following sections.
-  \item \textbf{DSP Platform and other Bundles}: The plug-and-play DSP
-  Components are based on the OSGi bundles capabilities, which can reuse services registered in the OSGi framework.
-\end{itemize}
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[scale=0.6]{../diagrams/NetBEAMS-Client-Node-Components}
-  \caption{The NetBEAMS's Gumstix Software Stack}
-  \label{fig:netbeams-software-stack}
-\end{figure}
-
-As depicted in figure \ref{fig:netbeams-software-stack}, the DSP Sonde Data
-Producer component opens a serial connection with the device through the Java
-Native Interface (JNI) driver and produces the data flow 1. The data stream
-format at this stage is as described on table \ref{tab:ysi-data-stream}.
-Later, by parsing and transforming the data to DSP internal representation, it
-sends the data to the DSP Platform as depicted on flow 2. Finally, the DSP
-Platform decides to send the data to the DSP Wire Transport Client, as shown
-on flow 3, by following rules defined by the system administrator. At this
-point, the data is ready to be transmitted to the server, which is covered on
-the following sections.
-
-\subsection{The DSP Platform Architecture}
-\label{netbeams-architecture}
-
-The architecture of the DSP, aligned with the OSGi capabilities, reuses the
-Service-Oriented Architecture that can easily enable and disable plug-and-play
-components. By following a publish-subscriber design-pattern \cite{gof}, the
-service producer registers into the service broker while the consumer looks
-for a given registered service in the Service Broker. For this reason, this
-approach allows decoupled interoperability among modules as it is shown in
-figure \ref{fig:publish-subscriber-pattern}.
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[scale=0.6]{../diagrams/publish-subscriber-pattern}
-  \caption{The publish-subscriber design-pattern for services}
-  \label{fig:publish-subscriber-pattern}
-\end{figure}
-
-The DSP infrastructure is based on a loosely-coupled architecture hosted in a
-given host computer with an OSGi container, where the entire system is
-composed by a set of plug-n-play OSGi modules. Each DSP component, as an
-instance of an OSGi Module, can perform specialized tasks according to its
-specification. In general, each DSP component can be categorized as follows:
-
-\begin{itemize}
-  \item \textbf{Data Producer (DP)}, whose main function is to produce data for
-  the system;
-  \item \textbf{Data Consumer (DC)}, whose main function is to consume the
-  received data.
-\end{itemize}
-
-\begin{figure}[!b]
-  \centering
-  \includegraphics[scale=0.7]{../diagrams/DSP-Producers-Consumers-Components-Interactions}
-  \caption{The basic architecture of a Data Sensor Platform (DSP)}
-  \label{fig:DSP-Producers-Consumers-Components-Interactions}
-\end{figure}
-
-The Data Producers and Data Consumers exchange data through the DSP Message
-Broker using a message unit called DSP Message. As it is depicted in figure
-\ref{fig:DSP-Producers-Consumers-Components-Interactions}, a Data Producer
-creates a set of one or more DSP Messages wrapped up in a DSP
-MessagesContainer in order to be sent to the DSP Message Broker. When the
-broker receives the set of DSP Messages, it requests a list of Data Consumers
-to each of the DSP Messages based on rules managed by the DSP Matcher.
-Whenever the broker acquires the list of Data Consumers, it delivers a copy of
-the DSP Message to each of the Data Consumers in the receiving list. Upon the
-DSP Message arrival, each Data Consumer can extract the actual data from the
-DSPMessage. Finally, although the components can be distinguished between Data
-Producers and Consumers, any of them are entitled to send and receive
-administrative DSP Messages, which is covered in the following sections.
-
-When the DSP Platform is installed into the OSGi framework, it follows the
-OSGi specifications without any changes to remain it in the installed state.
-In this way, it is important to note that the DSP Platform and its
-plug-and-play components are directly mapped with the OSGi Framework as
-follows:
-
-\begin{itemize}
-  \item \textbf{OSGi Bundle} = DSP Component Artifact in Jar format
-  \begin{itemize}
-    \item SondeDSPProducer.jar is the component developed for the YSI sonde
-  \end{itemize}  
-  \item \textbf{OSGi Activator} = DSP Component Activator
-  \begin{itemize}
-    \item Class specified in the MANIFEST.MF artifact (next sections)
-  \end{itemize} 
-  \item \textbf{OSGi Service Reference} = DSP Component Class
-  \begin{itemize}
-    \item  Class specified by the component author.
-  \end{itemize}
-\end{itemize}
-
-The DSP Platform, along with the DSP Framework, are the only DSP Components
-needed to be installed into the OSGi container. For this reason,the DSP
-Platform is responsible to install and activate the additional DSP components
-that makes part of the platform during the start up process, which leverages
-its bootstrap process. Nevertheless, in order to automate the process of
-automatic install and removal of the plug-and-play components, the DSP Platform
-provides a deployment configuration file that describes which DSP components
-must be installed with a given priority number. The name of this configuration
-artifact is called config.xml (see listing \ref{file:dsp-config.xml}), and is
-depicted by the XML Schema \cite{xml-schema} in figure
-\ref{fig:dsp-config-schema}.
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[scale=0.8]{../diagrams/config-schema}
-  \caption{XML Schema for the main DSP deployment configuration artifact}
-  \label{fig:dsp-config-schema}
-\end{figure}
-
-In order to show an example, consider the DSP Platform on the embedded system
-Gumstix in figure \ref{fig:sf-netbeams-node}. It is started by providing a
-deployment configuration artifact config.xml included in the appendix. It lists
-the components that the DSP Platform needs to install and load for its
-operation. As it is shown, the DSPWireTransportClient and the DSPSondeProducer
-are the identification of the components, followed by the name of the physical
-file and an assigned loading priority. While the DSPWireTransportClient is
-responsible for the serialization and deserialization of the DSP Messages
-exchanged by the DSPSondeProducer and any remote DSP component. In this case,
-the DSPSondeProducer is the DSP component responsible to collect measurement
-data from the YSI Sonde and temporarily store them before being transmitted.
-
-Therefore, in order to orchestrate the DSP Components together, the DSP
-Platform component is responsible for installing and uninstalling during its
-state "Starting" and "Stopping", respectively, as it is shown in figure
-\ref{fig:DSPPlatform-Install-Usage-State-Diagram}. When the system
-administrator starts the DSP Platform, it performs the regular OSGi procedures
-to start the bundle. Up until that state, it will get the list of DSP
-Components to be installed as described in the config.xml artifact, trying to
-install each of them in an optimistic way. After that, it will continue on the
-state "Active" until the system administrator decides to stop it.
-
-\begin{figure}[!b]
-  \centering
-  \includegraphics[scale=0.6]{../diagrams/DSPPlatform-Install-Usage-State-Diagram}
-  \caption{The DSP Platform State Diagram}
-  \label{fig:DSPPlatform-Install-Usage-State-Diagram}
-\end{figure}
-
-When the DSP Platform is on the Active state, it is ready to manage the DSP
-Messages exchanges as described through figure
-\ref{fig:DSP-Producers-Consumers-Components-Interactions}. The following
-sections describe the participating classes in starting the DSP Platform and
-its execution.
-
-In a nutshell, the DSP Platform is an extension of the OSGi Platform, as it can
-be seen in the UML Class Diagram \cite{uml} in figure
-\ref{fig:DSP-Platform-Class-Diagram-Simple}. The class Activator contains a
-unique instance of the Platform class, which contains instances of the classes
-ComponentManager, Matcher, the MessageBroker and DSPContextImpl. The
-DSPBundleController is the main instance that manages the plug-and-play DSP
-components, and therefore, is tightly-coupled with the OSGi Framework classes.
-Additionally, this class is part of the publish-subscriber design-pattern, as
-it it implements the interfaces Bundle Listener and Service Listener, whose
-responsibility is to track listens to the changes of the state of
-ServicesReferences on the DSP Platform.
-
-\begin{figure}[!t]
-  \centering
-  \includegraphics[scale=0.7]{../diagrams/DSP-Platform-Class-Diagram-Simple}
-  \caption{The DSP Platform UML Class Diagram}
-  \label{fig:DSP-Platform-Class-Diagram-Simple}
-\end{figure}
-
-\section{DSP Message and DSP Messages Container}
-\label{sec:dsp-message}
-
-As shown in the previous section, a DSP Message is the main unit of
-communication among DSP Components. In a nutshell, a DSP Message can be seen
-as an abstraction of an envelop containing the sections: a header and a body.
-The former is used for identification and routing purposes, while the latter is
-used to carry the payload of the message, that is, the collected data from
-a sensor. Figure \ref{fig:DSP-Message-Representation} depicts the abstraction
-of a DSP Message and its main components.
-
-\begin{figure}[!b]
-  \centering
-  \includegraphics[scale=0.8]{../diagrams/DSP-Message-Representation}
-  \caption{An Abstraction of a DSP Message and its components}
-  \label{fig:DSP-Message-Representation}
-\end{figure}
-
-\begin{itemize}
-  \item \textbf{Header} is composed by 2 blocks that identify the data
-  producer and consumer of the message, as well as the time of creation, UUID,
-  etc. It is an important piece of information for the data delivery and
-  routing purposes.
-  \item \textbf{Body} contains the payload of the message. This section carries
-  the collected data from a sensor.
-\end{itemize}
-
-In general, the DSP Platform offers a veriety of types of DSP Messages for
-different purposes. For example, any measurement data must be
-wrapped up in a Measurement Message, while a Query Message is used to exchange
-messages among the components for the purpose of management. In this way, the
-main DSP Messages can be summarized as follows:
-
-\begin{itemize}
-  \item \textbf{Measurement Message} is used to transport any sensor collected
-  data;
-  \item \textbf{Query Message} is used to query a DSP component about its
-  configuration properties;
-  \item \textbf{Update Message} used to update a DSP component's configuration
-  properties;
-  \item \textbf{Acknowledgement Message} is used for the transport communication
-  protocol, by acknowledging the reception of a DSP Message. More details in the
-  Remote Data delivery section.
-\end{itemize}
-
-These DSP Classes and other can be seen in figure the  UML Class
-Diagram \cite{uml} of figure \ref{fig:DSP-Messages-Classes}. Each DSP Message
-contains an instance of a Header and Body. Whenever a DSP component is ready
-to transmit messages, it wraps up the set of DSP Messages into the instance of
-the Messages Container, which contains information about the collection of
-messages being transmitted with its own identification. In this fashion, the
-Messages Container is the main communication unit between 2 different DSP
-Components.
-
-\begin{figure}[!t]
-  \centering
-  \includegraphics[scale=0.6]{../diagrams/DSP-Messages-Classes}
-  \caption{The DSP Messages UML Class Diagram}
-  \label{fig:DSP-Messages-Classes}
-\end{figure}
-
-Another use of the DSP Messages is during the DSP Platform activation process,
-where bootstrap messages are used to configure each DSP component. In order to
-send configuration parameters, the DSP Platform uses Update Messages using a
-unique type of Message Content called DSP Properties, which is described in the
-next section.
-
-\section{Collected Sensor Data as the DSP Message Content}
-\label{sec:dsp-payload-implementation}
-
-As mentioned in the previsou section, a DSP Message can carry any data on its
-body, also called the payload. In such a way, any data representation must
-just extend from the Message Content abstract class as it is shown
-\ref{fig:Sonde-MessageContainer-Class-Diagram}, as defines the data
-representation of the YSI Sonde data.
-
-\begin{figure}[!h]
-  \centering
-  \includegraphics[scale=0.6]{../diagrams/Sonde-MessageContainer-Class-Diagram}
-  \caption{The YSI Payload - Message Content UML Class Diagram}
-  \label{fig:Sonde-MessageContainer-Class-Diagram}
-\end{figure}
-
-The properties measured from the YSI Sonde can be captured by the class
-SondeDataType as a direct render of the measurements data depicted in table
-\ref{tab:ysi-data-stream}. In addition to the regular data types, the class
-also provides additional properties such as the attribute "samplingTimeStamp"
-that carries the time when the data was collected from the YSI Sonde. However,
-in order to capture a set of measurements at once, the SondeDataContainer
-class used to carry a set of SondeDataType classes, as it is composed by one
-or more instances of SondeDataType. In order to illustrate other examples,
-consider the UML Class Diagram \cite{uml} of image
-\ref{fig:Mouse-Actions--MessageContainer-Class-Diagram} as a payload of
-observations made from a mouse over a screen, which captures the name of the
-event, the button name and the x/y coordinates.
-
-\begin{figure}[!h]
-  \centering
-  \includegraphics[scale=0.5]{../diagrams/Mouse-Actions--MessageContainer-Class-Diagram}
-  \caption{The Mouse Actions Payload - Message Content UML Class Diagram}
-  \label{fig:Mouse-Actions--MessageContainer-Class-Diagram}
-\end{figure}
-
-As described in the previous section, the DSP Platform uses DSP Update
-Messages to configure internal components' initial configuration values. In
-order to do so, the DSP Platform uses an instance of the DSP Properties, which
-is composed of a list of initial DSP Property and their relating Values
-instance depicted on the UML class diagram \cite{uml} of figure
-\ref{fig:DSP-Property-Class-Diagram}.
-
-\begin{itemize}
-  \item \textbf{Property Instance}: initial-rate
-  \item \textbf{Value Instance}: 40
-\end{itemize}
-
-\begin{figure}[!t]
-  \centering
-  \includegraphics[scale=0.6]{../diagrams/DSP-Property-Class-Diagram}
-  \caption{The DSP Property Payload - Message Content UML Class Diagram}
-  \label{fig:DSP-Property-Class-Diagram}
-\end{figure}
-
-\subsection{DSP Platform Initialization Process}
-
-All in all, the DSP Platform requires the knowledge of the aforementioned DSP
-classes in order to initialize. The UML Sequence Diagram \cite{uml} in figure
-\ref{fig:From-OSGi-DSP-Platform-Start-to-All-Components-Sequence-Diagram}
-depicts the exact moments when the additional the plug-and-play DSP components
-are installed and bootstrapped by the DSP Platform, as the UML State Diagram
-\cite{uml} in figure \ref{fig:DSPPlatform-Install-Usage-State-Diagram}
-described.
-
-\begin{itemize}
-  \item \textbf{Start DSP Activator Gate}: after installing the Activator
-  indicated by the DSP Bundle's MANIFEST.MF artifact, the OSGi Framework makes
-  a call to the method Activator.start();
-  \item \textbf{Bootstrap DSP Component Gate}: after the component has been
-  initialized (by the call the method init() from the class ComponentManager),
-  the DSP Component is ready to be bootstrapped, and the optional bootstrap
-  message is delivered.
-\end{itemize}
-
-\begin{figure}[!b]
-  \centering
-  \includegraphics[scale=0.5]{../diagrams/From-OSGi-DSP-Platform-Start-to-All-Components-Sequence-Diagram}
-  \caption{The DSP Platform Initialization UML Sequence Diagram}
-  \label{fig:From-OSGi-DSP-Platform-Start-to-All-Components-Sequence-Diagram}
-\end{figure}
-
-\newpage
-
-\subsection{Data Delivery Mechanism}
-
-In general, when a DSP Component finishes preparing the DSP Messages Container,
-it contacts an entity responsible for sending and proxying message to other
-components. The DSP Broker acts like a mailman, receiving messages and
-forwarding copies of it to other entities in the system, abstraction in which
-is shown in figure \ref{fig:DSP-Message-Broker-Abstraction}. In this way, the
-DSP Broker receives a DSP Message and makes necessary deliveries to DSP
-components that it has knowledge of.
-
-\begin{figure}[!h]
-  \centering
-  \includegraphics[scale=0.8]{../diagrams/DSP-Message-Broker-Abstraction}
-  \caption{A DSP Message Broker Abstraction}
-  \label{fig:DSP-Message-Broker-Abstraction}
-\end{figure}
-
-The use of messaging exchange paradigm is used in NetBEAMS because it decouples
-the system by using messages take into account their structure to filter the
-messages. In this way, the DSP Broker uses a filter component called DSP
-Matcher, which provides a list of DSP Components that needs to receive a copy
-of the DSP Message. It turns out that the DSP Matcher bases its decisions on
-structured rules provided by the system administrator for an specific DSP
-Platform instance. The UML Class diagram for the DSP Broker, Matcher and
-particiaping classes is depicted in figure
-\ref{fig:DSPBroker-Matcher-Class-Diagram}, and describes how the DSP Broker
-implementation is tightly coupled with the DSP Matcher, as well as to the
-lists of DSP Components sorted by name and type.
-
-\begin{figure}[!b]
-  \centering
-  \includegraphics[scale=0.6]{../diagrams/DSPBroker-Matcher-Class-Diagram}
-  \caption{The DSP Broker-Matcher UML Class Diagram}
-  \label{fig:DSPBroker-Matcher-Class-Diagram}
-\end{figure}
-
-In general, the DSP Matcher is composed by a list of MatchRule, which contains
-a reference to a MatchCriteria and a MatchTarget. The former identifies the
-rule for matching a given DSP Message based on the producer and consumer's
-message type and location, while the latter describes the DSP Component that
-will receive the DSP Message, and if its delivery is done by means of a gateway
-component. In this way, the DSP Matcher can be seen as a function, as depicted
-in Table \ref{tab:matching-rule}, that takes a DSP Message as an input and
-returns a list of DSP consumers, and the matching configuration artifact provided by the administrator. See listing
-\ref{file:dsp-matcher-config.xml}) for details.
-
-\begin{table}
-    \caption{DSP Matcher Rule Algorithm to filter Data Consumer components to
-    receive a given Message}
-    \begin{center}
-        \begin{tabular}{lr}
-          \textit{Data Consumers ( DSP Message )} := \\ 
-          Verify the DSP Message's Header + \\
-          Verify the matcher rules, which contains the list of consumers for
-          the given message
-        \end{tabular}
-    \end{center}
-    \label{tab:matching-rule}
-\end{table}
-
-One example of such matching rule used by the DSP installed at the Gumstix can
-be seen in listing \ref{file:dsp-matcher-config-gumstix.xml}. the Gumstix
-system to configure the data delivery of the messages produced by the DSP
-Component DSP Sonde Producer. It describes a list of maching rules that
-reseambles a filtering system. This configuration contains 2 match rules: one
-related to internal DSP Configuration and another for a remote destination.
-The match criteria describes what the DSP Message Header properties must be,
-while the match target is the destination of the message. One important
-definition is the gateway Component Type, which defines the name of a remote
-DSP Component that is capable of receiving a remote DSP Message.
-
-Upon receiving all the matching rules and analyzing the DSP Message against
-the rules, the DSP Broker selects a set of the running DSP components to receive
-a copy of the DSP Message object in two different ways:
-
-\begin{itemize}
-  \item \textbf{In-memory local message delivery}: if the receiving DSP
-  Component is located in the current local host, a deep copy of the instance
-  of the DSP message is delivered;
-    \begin{itemize}[label=\textbullet]
-        \item IP Addresses are correctly resolved by using the Ethernet card in
-        the device: localhost, 127.0.0.1, or the same IP Address for Producer
-        and Receiver resolves into a local device. An example for such type of
-        delivery can be seen in the match ruleID
-        all messages sent to dsp manager of listing 
-        \ref{file:dsp-matcher-config-gumstix.xml}.
-    \end{itemize}
-  \item \textbf{Serialized remote message delivery}: if the receiving DSP
-  Component is located in a foreign/remote host, the message is serialized in a
-  format defined by the transport protocol chosen the DSP Component responsible
-  for the transport. The following section describes the existing DSP Data
-  Transport component;     
-  \begin{itemize}[label=\textbullet]
-        \item IP Address from the Producer and Consumer are different, and are
-        not resolved to be in the same host. An instance of such match rule can
-        be seen in the match rule ID send-message-to-remote url dsp of the code
-        snippet \ref{file:dsp-matcher-config-gumstix.xml}.
-    \end{itemize}
-\end{itemize}
-
-\subsection{Sending DSP Messages through the Wire}
-\label{sec:dsp-serialization}
-
-In order to send messages through the wire, the DSP Platform uses two symmetric
-components that serialize and deserialize DSP Messages to XML and back to Java
-POJO, respectively, and vice-versa. Named DSP Wire Transport Client and DSP
-Wire Transport Server \cite{netbeams2009}, these components use the HTTP
-\cite{http} protocol to transport the serialized version of the DSP Messages.
-Figure \ref{fig:DSP-to-DSP-Remote-Communication} depicts the scenario where a
-DSP Sensor Node transmits a DSP Message through an HTTP POST request.
-
-\begin{figure}[!b]
-  \centering
-  \includegraphics[scale=0.6]{../diagrams/DSP-to-DSP-Remote-Communication}
-  \caption{The remote communication between two DSP Components using the HTTP
-  Protocol}
-  \label{fig:DSP-to-DSP-Remote-Communication}
-\end{figure}
-
-\begin{itemize}
-  \item \textbf{DSP Wire Transport Client}: responsible for making HTTP POST
-  Requests to the service provided by the DSP Wire Transport Server component.
-  The body of the HTTP POST request contains a serialized version of an
-  instance of a DSP MessagesContainer in XML format;
-  \item \textbf{DSP Wire Transport Server}: exposes a Web Server listening to
-  port 8080 that receives the HTTP POST request from the Client. Upon receiving
-  the request, the server deserializes the DSP MessagesContainer back to the
-  in-memory POJO and proceeds with the delivery to the DSP Broker. At this
-  situation, the component packages acknowledgement messages and any queued
-  message to the client and piggybacks \cite{xml-piggybacking} it to the HTTP
-  response body.
-\end{itemize}
-
-As described by \cite{netbeams2009}, NetBEAMS is a single-hop sensor network
-whose sensors transmits data from the sensors to the server. Upon receiving the
-data stream, the DSP Wire Transport Server component desirializes the DSP
-MessagesContainer and its enclosing DSP Messages back to a Java class instance
-and delivers each of the DSP Messages to the DSP Broker, which delivers a
-copy of the message to each DSP Component listed by the maching rule. 
-%This
-%process is depicted by the UML Sequence Diagram \cite{uml} in figure \ref{fig:}
-
-Whenever a DSP bundle producer is ready to transmit messages, they are added
-into a DSP Messages Container (see UML Class Diagram in figure
-\ref{fig:DSP-Messages-Classes}), which is serialized in XML to be transmitted.
-Listing \ref{file:dsp-message-serialized-ysi} shows an instance of a DSP
-MessagesContainer in serialized in XML, ready for transmission. In fact, the
-listing is an example of the transmitted data collected from a sensor located
-at the 192.168.0.103 to be transmitted to the host 192.168.0.106, and the
-process of transmission in the server-side can be followed in figure
-\ref{fig:From-DSPSondeComponent-to-DSPWireTransportClient-Sequence}.
-
-\begin{figure}[!h]
-  \centering
-  \includegraphics[scale=0.5]{../diagrams/From-DSPSondeComponent-to-DSPWireTransportClient-Sequence}
-  \caption{The UML Sequence Diagram for the production of a YSI Message}
-  \label{fig:From-DSPSondeComponent-to-DSPWireTransportClient-Sequence}
-\end{figure}
-
-The Sonde DSP Component is responsible for starting a worker thread called
-SondeProducer, which is responsible for constantly read from the RS-232 serial
-port by using an instance of the SerialReader class. This process depicts the
-production of the data flow "flow 1" of figure
-\ref{fig:netbeams-software-stack}, which is a collection of SondeDataType for
-each of the measurement data read on a SondeDataContainer. As depicted in
-figure \ref{fig:From-DSPSondeComponent-to-DSPWireTransportClient-Sequence}, it
-first read the serial connection using the class SerialReader, which parses 
-the data stream. After parsing the data, the reader creates an instance of the
-class SondeDataType, the POJO responsible for carrying out the measurement
-data. As described earlier, the data is added into the SondeDataContainer to
-dispatched to the DSP Broker, whose responsibility is to get the list of DSP
-Components that is setup to receive a copy of the DSP message (see previous
-code snippet). As for the setup done for the Gumstix, the measurement message
-is supposed to be sent to a remote destination of host 192.168.0.7. At this
-point, the DSP Broker makes a decision to send the DSP Message to the default
-gateway component DSPWireTransportClient.
-
-After the message arrives in the DSPWireTransportClient, the message is
-maintained in a temporary repository of outbound DSP Messages called
-MessagesQueues after the message is hashed by the destination IP address and
-the DSP Client sender using an instance of QueueMessageData. As shown in figure
-\ref{fig:From-DSPWireTransport-Client-to-DSPWireTransportServer-Sequence}, the
-DSP TransportSender is a worker thread that is started by the DSP Wire
-Transport Client and is constantly transmits the data in the outbound queues
-after a configurable rate. It makes a transport of the message and receives
-the response, updates the state of the messages to transmitted, to finally
-send the response messages to the DSP Message Broker.
-
-\begin{figure}[!t]
-  \centering
-  \includegraphics[scale=0.5]{../diagrams/From-DSPWireTransport-Client-to-DSPWireTransportServer-Sequence}
-  \caption{The UML Sequence Diagram describing the Wire Transport Server
-  receiving a DSP Message}
-  \label{fig:From-DSPWireTransport-Client-to-DSPWireTransportServer-Sequence}
-\end{figure}
-
-While the DSPWireTransportClient makes an HTTP Request with the DSP Message,
-the DSPWireTransportServer's servlet receives the request. It first
-deserializes the MessagesContainer and delivers each of its messages to the DSP
-Broker. At this point, the server verifies if there are messages in the
-outbound queue that are addressed for the host that requested the messages
-delivery. By using a technique called piggybacking, the server adds to the HTTP
-Response a serialized version of the MessagesContainer with any DSP Messages.
-In addition to any DSP Message, the server prepares and sends acknowledgement
-messages back to the requesting host to the host.
-
-\begin{figure}[!t]
-  \centering
-  \includegraphics[scale=0.5]{../diagrams/From-DSPWireTransport-Server-To-DSP-Broker}
-  \caption{The UML Sequence diagram describing the DSP
-Message delivery on the server-side and the response}
-  \label{fig:From-DSPWireTransport-Server-To-DSP-Broker}
-\end{figure}
-
-The gate "From DSP WireTransport Server to Broker Gate" defines the actual
-delivery of the DSP Message to the DSP Broker. As this section is a tutorial
-about the DSP Platform, the remaining steps will be explained on Section 3,
-while showing the design and implementation of the component DSP Data
-Persistence.
-
-\section{DSP Data Persistence: Component Implementation}
+\section{DSP Platform Deployment Details}
 \label{sec:dsp-data-persistence-implementation}
 
-This section shows the implementation of the DSP Data Persistence, which
-follows the specifications of the DSP Components in the previous chapter,
-along with the setup of the mongoDB, the chosen database system that models
-data in a type of key-value pair model. Moreover, this chapter also describes
-the experiments conducted for the persistence.
-
-The implementation of the DSP Data Persistence component, as well as the
-implementation of the experiments, were developed and managed using the DSP a
-reserved development branch at the Subversion \cite{subversion} repository
-located at the Google Code website. In this way, anyone can checkout the entire
-source-code from the branch
-\textit{https://netbeams.googlecode.com/svn/branches/marcello/persistence},
-as shown in Figure \ref{fig:dsp-data-persistence-dir}. In this way, the
-recommended checkout branch directory will give the remaining directory
-structure as summarized:
-
-\begin{itemize}
-  \item \textbf{NEBEAMS-DIR/}: the directories under the subdirectory
-  ``persistence'';
-  \item \textbf{DEV-VERSION}: the directory of the current development
-  version for the DSP; That is, ``\textbf{NEBEAMS-DIR}/versions/v2/''
-  \item \textbf{PERSISTENCE-DIR}: the directory of the DSP component, that is,
-  ``\textbf{DEV-VERSION}/apps/osgi-intro-bundles/dsp/DSPDataPersistence''.
-\end{itemize}
-
-\begin{figure}[!b]
-  \centering
-  \includegraphics[scale=0.6]{../diagrams/dsp-data-persistence-dir}
-  \caption{The DSP Data Persistence Directory Structure}
-  \label{fig:dsp-data-persistence-dir}
-\end{figure}
-
-The DSP Platform defines its own run-time directory, which organizes the
-artifacts used by the osgi-intro Framework, as well as the its own configuration
-descriptor and matcher, as described in the previous section.
-
-\begin{itemize}
-  \item \textbf{RUNTIME-DIR}: the main directory for the execution of the DSP
-Platform, with all the necessary osgi-intro-related artifacts are placed, as shown in
-Figure \ref{fig:dsp-runtime-dir}.
-\end{itemize}
-
-\begin{figure}[!t]
-  \centering
-  \includegraphics[scale=0.5]{../diagrams/dsp-runtime-dir}
-  \caption{The DSP Data Persistence Directory Structure}
-  \label{fig:dsp-runtime-dir}
-\end{figure}
-
-\section{DSP Platform Deployment Details}
-
 As described in the previous chapter, the DSP Data Persistence Component was
-implemented using the Java Programming language, on top of the osgi-intro Framework
-API. This section describes the setup process of the implementation, as well as
-the artifacts use during the implementation, which will be listed in the
+implemented using the Java Programming language, on top of the OSGi Framework
+API. This section describes the setup process of the implementation, as well
+as the artifacts use during the implementation, which will be listed in the
 appendix section.
@@
 The structure of the artifacts implemented for the DSP Data Persistence
-Component follow the conventions of the NetBEAMS implementation, where the main
-directory \textbf{PERSISTENCE-DIR}, depicted by Figure
+Component follow the conventions of the NetBEAMS implementation
+\cite{netbeams-dsp-architecture}, where the main directory
+\textbf{PERSISTENCE-DIR}, depicted by Figure
 \ref{fig:dsp-data-persistence-dir-checkedout}, holds each of the development
@@
   \item \textbf{META-INF}: this directory contains the descriptor file for
-  the osgi-intro Framework, this this case the MANIFEST.MF;
+  the OSGi Framework, this this case the MANIFEST.MF;
   \item \textbf{src}: the main directory structure for the source-code
@@
   therefore, lists the package org.netbeams.dsp.persistence as the main
-  package, including the layers controller, model and osgi-intro.
+  package, including the layers controller, model and OSGi.
 \end{itemize}
@@
 
-\subsection{osgi-intro Deployment Process}
+\subsection{OSGi Deployment Process}
 
-Since each NetBEAMS component is managed by an osgi-intro component and its
-infrastructure, this section describes the basic functionality of the osgi-intro
-platform \cite{osgi-intro}, a framework was conceived to support modularity in
-terms resources-limited environments such mobile devices and vehicles, but it
-was first widely deployed on Eclipe IDE\footnote{Integrated Development
-Environment}, because it promotes the use of reusable loosely-coupled modules
-using the Java Programming Language \cite{java}. The most basic layer of the
-osgi-intro Framework depicted in Figure \ref{fig:layering-osgi-intro} are as follows:
+Since each NetBEAMS component is managed by an OSGi component and its
+infrastructure, this section describes the basic functionality of the
+OSGi platform \cite{osgi}, a framework was conceived to support
+modularity in terms resources-limited environments such mobile devices and
+vehicles, but it was first widely deployed on Eclipe IDE\footnote{Integrated
+Development Environment}, because it promotes the use of reusable
+loosely-coupled modules using the Java Programming Language \cite{java}. The
+most basic layer of the OSGi Framework depicted in Figure
+\ref{fig:layering-osgi} are as follows:
 
 \begin{itemize}
-  \item \textbf{Module Layer}: manages the osgi-intro bundles deployed on the osgi-intro
-  Platform, providing the necessary "wiring" of the components. In other words,
-  the modules, called osgi-intro Bundles, can export and/or import packages in the
-  level of a Java Class managed by the osgi-intro Framework;
+  \item \textbf{Module Layer}: manages the OSGi bundles deployed on the
+  OSGi Platform, providing the necessary "wiring" of the components. In other words,
+  the modules, called OSGi Bundles, can export and/or import packages in the
+  level of a Java Class managed by the OSGi Framework;
   \item \textbf{Service Layer}: resposible for the interoperability between 2
@@
   \item \textbf{Execution Layer}: executes the bundles, managing and changing
-  their life-cycle through the bundle execution.
+  their life cycle through the bundle execution.
 \end{itemize}
@@
   \includegraphics[scale=0.5]{../diagrams/layering-osgi}
-  \caption{The osgi-intro Framework Layers}
-  \label{fig:layering-osgi-intro}
+  \caption{The OSGi Framework Layers}
+  \label{fig:layering-osgi}
 \end{figure}
 
-NetBEAMS uses the concept of the modularization through the use of the the
+NetBEAMS uses the concept of the modularization through the use of the
 Producer-Consumer paradigm described in the previous section. Since the DSP
-Components are essentially osgi-intro bundles, the interoperability between them are
-given in the implementation and reuse existing bundles in the osgi-intro Framework
-and the DSP Platform, described by an artifact called MANIFEST.MF as seen in
-Listing \ref{file:osgi-intro-manifest}. In general, an osgi-intro bundle must
-provide specifications that describes the module to be published into the osgi-intro
-Framework. In this way, the main properties of the osgi-intro MANIFEST.MF artifact can
-used by the DSP Data Persistence can be summarized as follows:
+Components are essentially osgi bundles, the interoperability between
+them are given in the implementation and reuse existing bundles in the
+OSGi Framework and the DSP Platform, described by an artifact called
+MANIFEST.MF as seen in Listing \ref{file:OSGi-manifest}. In general, an
+OSGi bundle must provide specifications that describes the module to be
+published into the OSGi Framework. In this way, the main properties of
+the OSGi MANIFEST.MF artifact can used by the DSP Data Persistence can
+be summarized as follows:
 
 \begin{itemize}
-  \item \textbf{Bundle-Activator}: The name of the instance of an osgi-intro Activator
-  class, responsible to manage the bundle. The implemented class
+  \item \textbf{Bundle-Activator}: The name of the instance of an OSGi
+  Activator class, responsible to manage the bundle. The implemented class
   ``DataPersistenceActivator'' provides the activation mechanisms for the
@@
   DSP components, forcing the execution to be dependent on those package.
-  They are provided through a ``Export-Package'' section in other osgi-intro
-  bundles and, as it is shown in Listing \ref{file:osgi-intro-manifest}, they
+  They are provided through a ``Export-Package'' section in other OSGi
+  bundles and, as it is shown in Listing \ref{file:OSGi-manifest}, they
   are from different packages.
@@
 
-Once the osgi-intro bundle is installed into the osgi-intro Platform, it will be managed by
-the osgi-intro Execution layer and change the bundle state according to its
-lify-cycle, as described in the previous chapter. Therefore, the packaging of
-the osgi-intro source-code is done using a JAR\footnote{Java Archival Repository}
+Once the OSGi bundle is installed into the OSGi Platform, it will be managed by
+the OSGi Execution layer and change the bundle state according to its lify
+cycle, as described in the previous chapter. Therefore, the packaging of the
+osgi into source-code is done using a JAR\footnote{Java Archival Repository}
 artifact specification \cite{java-tutorial}. In this way, the DSP Data
-Platform osgi-intro bundle JAR is created by the Apache ANT build script
+Platform osgi into bundle JAR is created by the Apache ANT build script
 \cite{apache-ant} in Listing \ref{file:dsp-build.xml}. The task 
@@
 \ref{fig:dsp-runtime-dir} shows the DSP Data Persistence file under the
-dirctory ``\textbf{RUNTIME-DIR}/deployment/''.
+directory ``\textbf{RUNTIME-DIR}/deployment/''.
 
-\subsection{Adding DSP Data Persistence into DSP Platform }
+\begin{figure}[!h]
+  \centering
+  \includegraphics[scale=0.5]{../diagrams/dsp-runtime-dir}
+  \caption{The DSP Runtime Dir}
+  \label{fig:dsp-runtime-dir}
+\end{figure}
+
+\subsection{DSP Data Persistence into the DSP Platform}
 
@@
 dependencies to other DSP Components.
-\subsection{Starting the DSP Platform}
 
-Once the DSP Platform automatically installs all the components described by
-the configuration descriptor, the container verifies the dependencies and
-starts each of the DSP components of Figure \ref{fig:knopflerfish-execution},
-the osgi-intro Framework container Knopflerfish \cite{knopflerfish}.
-
-\begin{figure}[!t]
-  \centering
-  \includegraphics[scale=0.65]{../diagrams/knopflerfish-execution}
-  \caption{The execution of the Knopflerfish Container}
-  \label{fig:knopflerfish-execution}
-\end{figure}
-
-\subsection{Execution Logs}
-
-\section{mongoDB Deployment}
+\section{mongoDB Deployment Details}
 \label{sec:mongodb-deployment}
 
-Given that Document-Oriented Model makes a good candidate persist sensors'
-properties and the recently enumerated list technologies in the previous
-article DSPDataPersistence, the open-source project called mongoDB was chosen
-for the evaluation on our case study, the netBEAMS DSP Platform.
-
-mongoDB supports storage based on collections of data, stored using BSON, a
-binary representation the JSON data representation format, including dynamic
-queries and indexing support. As it's stated in their web site, mongoDB
-"bridges the gap between key/value stores (which are fast and highly scalable)
-and traditional RDBMS systems (which are deep in functionality)".
+MongoDB was the selected technology described in chapter 5 because of its
+features. It provides data persistence based on collections of data, stored
+using BSON \cite{bson}, a binary representation the JSON data representation
+format. mongoDB uses the abstraction of programming language to
+represent the basic CRUD operators, as well as the regular functionalities
+provided by any type of database system such as indexing, users management,
+etc. As it's stated in their web site, mongoDB "bridges the gap between
+key/value stores, which are fast and highly scalable, traditional RDBMS
+systems (deep in functionality)".
 
@@
   \item mongoDB is written in C++, and therefore, can is available in any major
-  platform, as well as offers a broad range of API drivers written in different
-  languages such as Java, Python, Perl and Ruby; 
-  \item mongoDB is open-source, with good community support and availability
-  through mailing lists, freenode IRC channel, and commercial support through
-  10gen company; 
+  platform, as well as offers a broad range of API drivers written in
+  different languages such as Java, Python, Perl and Ruby;
   \item mongoDB has support to distributed systems properties such as
-  Master-Slave replication, and features like Database Shards with
-  auto-sharding based on shard keys.
+  Master-Slave replication, and database partitioning, referred to as Database
+  Shards.
 \end{itemize}
@@
 NetBEAMS resources. However, the build script from the DSP Data Persistence
-component can be executed to produce the persistence directory
-structure for NetBEAMS, as described in figure
-\ref{fig:dsp-persistence-system-dir}.
+component can be executed to produce the persistence directory structure for
+NetBEAMS, as described in Figure \ref{fig:dsp-persistence-system-dir}.
 
@@
   \includegraphics[scale=0.5]{../diagrams/dsp-persistence-system-dir}
-  \caption{The osgi-intro Framework Layers}
+  \caption{DSP Data Persistence directory}
   \label{fig:dsp-persistence-system-dir}
@@
 
-\subsection{Document-Oriented Data Model}
+\subsection{Using the Client and Server}
 
-An example of the data is as folows (using the JSON syntax). Listing
-\ref{file:mongodb-ysi-data-format} follows the strategy described on Section X,
-where the entity representation is denormalized, and each of the properties of
-the sensor is repeated in each instance of the data. The fact and transaction
-times are expressed in milliseconds. The data element contains the complete
-structure of the originating sensor such as the IP address. The collection of
-instances of this document represents the sensor data type.
+In order to start the server, the command-line shown in Listing
+\ref{cmd:mongodb-server-start} is used, where the data is configured to be
+stored at the directory  ``./data/'', relative to the directory shown in the
+shell. Also, the port number where the server listens to the client
+connections is give as port ``20000". Once the server is running, any client
+application can access to the server. For example, in order to connect the DSP
+Data Persistence component, the port 20000 is an optional value be provided in
+the DSP Data Persistence bootstrap message, as shown in Listing
+\ref{file:dsp-data-persistence-bootstrap.xml}. Although mongoDB offers the
+option to run with its specific default port number, it is recommended to
+assign different ports, once the use of shards requires different port numbers
+for each of the shard servers. An example of accessing mongoDB using a shell
+client is shown in Listing \ref{cmd:mongodb-client-start}.
 
-\subsection{MongoDB Database Administration}
+\lstset{label=cmd:mongodb-server-start,caption=Starting the Server}
+\begin{lstlisting}
+marcello-de-saless-macbook-pro:persistence marcello$ mongod --dbpath data/ --port 20000 &
 
-- starting the db
-- Stopping the Database
+Wed Dec  9 01:39:51 Mongo DB : starting : pid = 10096 port = 20000 dbpath = data/ master = 0 slave = 0  64-bit 
+Wed Dec  9 01:39:51 db version v1.1.4, pdfile version 4.5
+Wed Dec  9 01:39:51 git version: ad7fc8908396a3351a0f07bb1e2fc4baba2802ad
+Wed Dec  9 01:39:51 sys info: Darwin erh2.10gen.cc 9.6.0 Darwin Kernel Version 9.6.0: Mon Nov 24 17:37:00 PST 2008; root:xnu-1228.9.59~1/RELEASE_I386 i386 BOOST_LIB_VERSION=1_37
+Wed Dec  9 01:39:51 waiting for connections on port 20000
+\end{lstlisting}
 
-\section{Experiment Execution}
+\lstset{label=cmd:mongodb-client-start,caption=Starting the Server}
+\begin{lstlisting}
+marcello@netbeams-dev:~/workspaces $ mongo 192.168.1.2:20000/netbeams
+MongoDB shell version: 1.1.4
+url: 192.168.1.2:20000/netbeams
+connecting to: 192.168.1.2:20000/netbeams
+type "help" for help
+\end{lstlisting}
+
+\section{Experiments Execution}
 \label{sec:experiment-execution}
 
-The experiments conducted using mongoDB can be summarized by the different
-data access methods provided by the technology. Each subsection describes these
-methods.
+The experiments conducted using mongoDB can be summarized by the different data
+access methods provided by the technology. Each subsection describes these 
+methods. The server could be started using the commands specified in the
+previous section.
+
+\lstset{label=cmd:run-persistence-experiment,caption=Running main experiment
+shell script}
+\begin{lstlisting}
+marcello@netbeams-dev:~/workspaces/netbeams/versions/v2/persistence$ ./run-persistence-experiment 
+#######  NetBEAMS Experiments - Persistence on MongoDB  ########
+
+Usage: ./run-persistence-experiment X, where X is the size of the workload to be inserted from NetBEAMS YSI Data Handler to mongoDB
+\end{lstlisting}
 
@@
 
-The mongoDB client can be started by using the following command. Make sure you
-have started the mongoDB server before executing the mongoDB client.
+The iterative mongo client shell offers users to verify and navigate on a
+given database and its collections. This first section shows the connection of
+the mongo client to the database netbeams as on Listing
+\ref{cmd:mongodb-server-start}. The execution of the operation to count the
+existing number of documents in the mongoDB collection ``netbeams'' is shown in
+Listing \ref{cmd:mongodb-count}. It highlights the query for the collections
+available \cite{mongodb} in the netbeams database.
 
-mongo netbeams | tee output\underline{ }number\underline{ }date.log
-
-Here, the iterative mongo client shell offers users to verify and navigate on a
-given database and its collections. This first section shows the connection of
-the mongo client to the database netbeams. It also highlights the query for the
-collections available. During the experiment, the SondeDataContainer?
-collection was created as related to the type from the DSP Messages for the YSI
-Sonde.
-
-The shell references to the mongoDB system can be found at
-http://www.mongodb.org/display/DOCS/dbshell+Reference
-
-\lstset{label=cmd:mongo,caption=Execution of mongo client}
+\lstset{label=cmd:mongodb-count,caption=Starting the Server}
 \begin{lstlisting}
-MongoDB shell version: 1.1.0-
-url: netbeams
-connecting to: netbeams
+...
+...
+connecting to: 192.168.1.2:20000/netbeams
 type "help" for help
@@
 system.indexes
->
 > db.SondeDataContainer.count()
-1000000
+2419200
 \end{lstlisting}
 
-Then, the first verification of the data integrity is regarding the number of
-elements created. Here, the first count() function on the collection returned
-1000000.
-
-An example about retrieving the first element of the collection can be done
-using the findOne() function. It will return an element instance on the
-JSONnotation.
+An example about retrieving a random document from the collection ``netbeams'',
+just type the ``findOne()'' in the shell shown in Listing
+\ref{cmd:mongo-findone}. The output is the actual instance of a document,
+showing all the collected data in the key ``observation'', and the provenance
+metadata in the keys ``sensor.location'' and ``sensor.time''.
 
@@
 The query based on attributes can be done using the "dot" notation, as you
-navigate through the JSON documents. Additionally, you can use the functions as
-aggregated on the result of others. The example in listing
+navigate through the JSON documents. Additionally, you can use the functions
+as aggregated on the result of others. The example in listing
 \ref{cmd:mongo-find-list} counts the number of documents with the key
@@
 The following example is the output of the first 2 documents from the same
-previous query using the limit() function, as shown in figure
+previous query using the limit() function, as shown in Figure
 \ref{cmd:mongo-find-limit}.
@@
 
-Other logs are located at
-http://code.google.com/p/netbeams/downloads/list
-
 \subsection{Data Access Using APIs}
+\label{sec:dsp-mongodb-rest-ws}
 
 The mongoDB server offers different drivers to access the data, as well as the
-Web Services.
+REST Web Services.
 
-\begin{itemize}
-  \item The Java tutorial on the drivers is located at
-    http://www.mongodb.org/display/DOCS/Java+Tutorial. The driver is located on
-    the NETBEAMS/versions/v2/thirdparty/mongodb directory;  
-  \item The REST API can be called from any HTTP client. At the time of the
-  editing, this feature is still under alpha version. Check the documentation
-  athttp://www.mongodb.org/display/DOCS/Http+Interface for details 
-\end{itemize}
+In order to get the results using the REST Web Services interdace, the request
+can be issued using a web browser. For instance, the request to the documents
+using the HTTP Request Variables to filter by the fields as shown in Listing 
+\ref{cmd:mongo-rest-request}.
 
-The following HTTP GET Request method returns the first 5 documents in the collection:
-http://127.0.0.1:28017/netbeams/SondeDataContainer/?limit=-5, as shown in
-listing \ref{cmd:mongo-rest-request}.
-
-\lstset{label=cmd:mongo-rest-request,caption=REST HTTP GET Request Example}
+\lstset{label=cmd:mongo-rest-request,caption=REST HTTP Interface Example}
 \begin{lstlisting}
@@
   "rows": [
-    { "_id" : { "$oid" : "616f400784ba194bc2dee200" }, "message_id" : "c76a9898-1eee-4463-bf5d-d40a7a8af6e3", "sensor" : { "ip_address" : "192.168.0.60", "location" : { "latitude" : 37.89155, "longitude" : -122.4464 } }, "time" : { "valid" : { "$date" : 1261229067000 }, "transaction" : { "$date" : 1259977271704 } }, "observation" : { "WaterTemperature" : 51.29, "SpecificConductivity" : 110.7, "Conductivity" : 104.5, "Resistivity" : 1710.07, "Salinity" : 0.01, "Pressure" : 1.094, "Depth" : 1.808, "pH" : 3.04, "pHmV" : -46.4, "Turbidity" : 0.2, "ODOSaturation" : 59, "ODO" : 10.66, "Battery" : 7.1 } } ,
-    { "_id" : { "$oid" : "616f400789ba194bdbe2e200" }, "message_id" : "2eb3f5ca-7a30-4b5d-8d16-c3e80a7b1b7a", "sensor" : { "ip_address" : "192.168.0.24", "location" : { "latitude" : 37.89155, "longitude" : -122.4464 } }, "time" : { "valid" : { "$date" : 1261740924000 }, "transaction" : { "$date" : 1259977271704 } }, "observation" : { "WaterTemperature" : 10.98, "SpecificConductivity" : 100.9, "Conductivity" : 104.5, "Resistivity" : 5448.21, "Salinity" : 0.08, "Pressure" : 1.152, "Depth" : 1.289, "pH" : 3.09, "pHmV" : -0.4, "Turbidity" : 0.1, "ODOSaturation" : 72.2, "ODO" : 59.65, "Battery" : 6.3 } } ,
-    { "_id" : { "$oid" : "616f400789ba194b13e5e200" }, "message_id" : "a7804d0a-44a8-4bd9-a9c7-51db84e91f08", "sensor" : { "ip_address" : "192.168.0.236", "location" : { "latitude" : 37.89155, "longitude" : -122.4464 } }, "time" : { "valid" : { "$date" : 1260969259000 }, "transaction" : { "$date" : 1259977271704 } }, "observation" : { "WaterTemperature" : 7.97, "SpecificConductivity" : 119.1, "Conductivity" : 104.5, "Resistivity" : 1971.59, "Salinity" : 0.05, "Pressure" : 1.516, "Depth" : 2.187, "pH" : 3.11, "pHmV" : -77.59999999999999, "Turbidity" : 0.2, "ODOSaturation" : 82.90000000000001, "ODO" : 41.99, "Battery" : 0.1 } }
-  "total_rows" : 3 ,
-  "query" : {} ,
-  "millis" : 0
+    { "_id" : { "$oid" : "616f400784ba194bc2dee200" }, 
+      "message_id" : "c76a9898-1eee-4463-bf5d-d40a7a8af6e3", 
+      "sensor" : { "ip_address" : "192.168.0.60", 
+                   "location" : {
+                       "latitude" : 37.89155, "longitude" : -122.4464} 
+                 }, 
+      "time" : { 
+               "valid" : { "$date" : 1261229067000 },
+               "transaction" : { "$date" : 1259977271704 } 
+               },
+      "observation" : { 
+                "WaterTemperature" : 51.29, "SpecificConductivity" :110.7,
+                "Conductivity" : 104.5, "Resistivity" : 1710.07, "Salinity" :0.01, 
+                "Pressure" : 1.094, "Depth" : 1.808, "pH" : 3.04,
+                "pHmV" : -46.4, "Turbidity" : 0.2, "ODOSaturation" :59, 
+                "ODO" : 10.66, "Battery" : 7.1 
                 } 
+                ......
+                ...... More records....
+   } 
+   "total_rows" : 3 , "query" : {} , "millis" : 91 }
 \end{lstlisting}
+
+\subsection{Visualizing Data on The Browser}
 
@@
 "SondeDataContainer  ?" being rendered by futon4mongodb, one of the
-open-source tools developed to visualise mongoDB data.
-
-\subsection{Visualizing Data on The Browser}
-
-Futon4Mongo is an open-source software adapted from CouchDB \cite{couchdb}.
-First, the collections list is shown. Figure
-\ref{fig:view-collections-instance-browser-futondb} shows one single collection
-for the YSI data containing 1 million objects.
+open-source tools developed to visualise mongoDB data. Futon4Mongo is an
+adaptation of another software developed for CouchDB \cite{couchdb}. First, the
+collections list is shown. Figure
+\ref{fig:view-collections-instance-browser-futondb} shows one single
+collection for the YSI data containing 1 million objects.
 
@@
 Sonde type, being indexed by a document ID and the values being the keys
-defined in the previous chapter, as shown in figure
+defined in the previous chapter, as shown in Figure
 \ref{fig:view-collected-data-list-browser-futondb}.
@@
   \centering
+
   \includegraphics[scale=0.5]{../diagrams/view-collected-data-list-browser-futondb}
-  \caption{Viewing a partial list of data using the Futon for CouchDB/MongoDB}
+  \caption{Viewing a partial list of data using the Futon4CouchDB/MongoDB}
   \label{fig:view-collected-data-list-browser-futondb}
@@
   \includegraphics[scale=0.5]{../diagrams/view-collected-data-instance-browser-futondb} 
-  \caption{Viewing an instance of collected data using the Futon for
-  CouchDB/MongoDB}
+  \caption{Viewing an instance of collected data using the Futon4CouchDB/MongoDB}
   \label{fig:view-collected-data-instance-browser-futondb}
@@
 An example to export the data in CSV format can be seen in listing
-\ref{file:mongodb-export-command}, and an example containing 1 million objects can be
-downloaded at
+\ref{file:mongodb-export-command}, and an example containing 1 million objects
+can be downloaded at
 http://netbeams.googlecode.com/files/experiment-1000000-data-exported-20090913-053538.csv.tar.gz.
@@
 
-By using any programming language driver, one can export data in the OPeNDAP
-format.
+During the experiments of this work, the implementation of the export utility
+to OPeNDAP was written in Python, as shown in Listing
+\ref{file:experiment-export-python}. In order to execute the
+command, the python programming language \cite{python} must be installed in the
+system. In this way, some limited documents are exported. The following command
+is an example of the execution of the export utility
+\ref{cmd:mongo-python-export}:
+
+\lstset{label=cmd:mongo-python-export,caption=Query Element with specific
+projection limiting the result set size}
+\begin{lstlisting}
+marcello@netbeams-dev:~/workspaces/netbeams/versions/v2/persistence$ python data-export-python.py 192.168.1.2:20000
+\end{lstlisting}
 
@@
 
-\lstinputlisting[label=file:osgi-intro-manifest,caption=OSGi Manifest
+\lstinputlisting[label=file:OSGi-manifest,caption=OSGi Manifest
 Descriptor]{../../../../netbeams/versions/v2/apps/osgi-bundles/dsp/DSPDataPersistence/META-INF/MANIFEST.MF}
@@
 
+\lstinputlisting[language=Bash,label=file:setup-os-shards-step1,caption=Experiment
+Main Execution Script]{../../../../netbeams/versions/v2/persistence/setup-os-shards-step1}
+
+\lstinputlisting[language=Bash,label=file:setup-os-shards-step2,caption=Experiment
+Main Execution Script]{../../../../netbeams/versions/v2/persistence/setup-os-shards-step2}
+
 \lstinputlisting[language=Python,label=file:experiment-export-python,caption=Data
-Export to from mongoDB to OPEnDAP format.]{../../../../netbeams/versions/v2/persistence/data-export-python.py}
+Export to from mongoDB to OPeNDAP format.]{../../../../netbeams/versions/v2/persistence/data-export-python.py}
 
@@
 \lstinputlisting[label=file:experiment-export-python-results,caption=Data
-Export to from mongoDB to OPEnDAP format.]{../../../../netbeams/versions/v2/persistence/export-mongodb-to-sfbeams.csv}
+Export to from mongoDB to OPeNDAP format.]{../../../../netbeams/versions/v2/persistence/export-mongodb-to-sfbeams.csv}

### 2009-12-10 docs/marcello-dissertation/thesis/chapter5-taxonomic-persistence-storage-selection-for-netbeams.tex
Revisions of chapter 5 from Niki... This should have been committed on the last revision.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@690 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 
-Providing data persistence for an existing sensor networks require analyzes of
-the infrastructure and the nature of the produced data. Based on the taxonomies
-of chapter 3, and the requirements of a data persistence for NetBEAMS in
-chapter 4, this chapter aims at analyzing a group of database technologies
-selected from literature review topics covered in chapter 2. Similarly,
-supported by the scope of the required implementation presented in this
-chapter, the comparison of the technogies is summarized, highlighting the
-selected database system used for the design of a component for NetBEAMS and
-used for the design of the experiments.
+Providing data persistence for an existing sensor network requires analysis of
+the infrastructure and the nature of the data produced. Based on the
+taxonomies of Chapter 3, and the requirements of a data persistence for
+NetBEAMS in Chapter 4, this chapter aims at analyzing a group of database
+technologies selected from the topics covered in Chapter 2.
+Similarly, supported by the scope of the required implementation presented in
+this chapter, the comparison of the technologies is summarized, highlighting
+the selected database system used for the design of a component for NetBEAMS
+and used for the design of the experiments.
 
-\section{Scope of the use of the Techonolgy}
+\section{Scope of the use of the Technology}
 
@@
 to provide the foundation of a database management system that can be used to
-develop one focused in less refactoring and high scalability and performance.
+develop one focused less in refactoring and high scalability and performance.
 For this reason, the use of the database system will be restricted to the data
-persistence and use through native system or through the use of native
+persistence and use through a native system or through the use of native
 programming languages or API. In this way, the scope of this technology
@@
   \item Provide data management over the collected sensor data, as a
-  conventional database system;
-  \item Must be able to scale in the face of an increase to the data load of
+  conventional database systems;
+  \item Must be able to scale in any increases to the data load of
   the sensor network, as described in chapter 2;
@@
 One of the essential questions related to data persistence is regarding the
-data model and breed of the database. That is, the use of the relational or
+data model and types of the database. That is, the use of the relational or
 any other model to describe data along with its management. Traditionally, the
 relational data model has been used to provide data persistence for different
-types of projects, inclusing sensor networks \cite{sn-dataware-house,
-db-xml-enabled, sn-db-tinydb}. However, tradition does not always get translated
-to efficiency in terms of problem solving. For this reason,
+types of projects, including sensor networks \cite{sn-dataware-house,
+db-xml-enabled, sn-db-tinydb}. However, tradition does not always get
+translated into efficiency in terms of problem solving. For this reason,
 \cite{db-is-rdbs-dommed} reflects about the use of alternatives to the
 relational model when the data persistence problem in consideration is to
-provide data persistence for dynamic environments as Internet applications. The
-author also suggests that the relational model is hard to maintain in
+provide data persistence for dynamic environments such as Internet applications.
+The author also suggests that the relational model is hard to maintain in
 ever-changing environments, and thus, proposes the use of the Key-Value-Pair
@@
 \cite{cloud-comp-survey} reports similar trends in academic and industrial
-research in distributed computing have used KVP databases in powerful systems
+research in distributed computing using KVP databases in powerful systems
 such as Cloud Computing \cite{cloud-comp-architectures}. For this reason, in
@@
   \item \textbf{MySQL}\cite{mysql}: system used in the implementation of a
-  dateware house with sensor network data \cite{sn-dataware-house};
-  \item \textbf{TinyDB}\cite{tinydb}: as referred in the literature review,
-  TinyDB is a database system used in many different sensor networks
-  implementation such as \cite{sn-db-tinydb};
+  data warehouse with sensor network data \cite{sn-dataware-house};
+  \item \textbf{TinyDB}\cite{tinydb}: as referred in Section 
+  \ref{sec:sn-persitence-storage}, TinyDB is a database system used in many
+  different sensor networks implementation such as \cite{sn-db-tinydb};
   \item \textbf{MongoDB}\cite{mongodb}: an open-source KVP database
@@
   XML models \cite{db-xml-enabled}. Furthermore, it was selected because
-  NetBEAMS represents serialized objects in XML, as shown in section
-  \ref{sec:dsp-message}.
+  NetBEAMS represents serialized objects in XML, as shown
+  \cite{netbeams-dsp-architecture}.
 \end{itemize}
 
-The following sections reviews these technologies' capabilities to adhere to
+The following sections review these technologies' capabilities to adhere to
 the specifications of the taxonomies defined in chapter 3, as well as the
@@
 \begin{itemize}
-  \item Data is generated by sensor devices, such as the YSI sonde
+  \item Data is generated by sensor devices, such as the YSI Sonde
   \cite{YSI-Sonde}, and manually collected by using a laptop to the data sink
@@
   \item Upon data reception, the RTC staff index and archive the data for
-  distribution using the OPEnDAP format;
+  distribution using the OPeNDAP format;
   \item An automated approach to the data collection procedures is performed 
@@
   send the data to the data sink node. A complete description of this approach
-  is detailed in section \ref{sec:netbeams-architecture}.
+  is \cite{netbeams-dsp-architecture}.
 \end{itemize}
 
-As summarized in the previous section, the nature of data from NetBEAMS through
-SF-BEAMS is used for the purpose of \textbf{Data Archival}. After the
-collected data captured from the sensors, it requires further processing
-in order to be archived for later reuse. Different types of users are expected
-to use the collected data, ranging from researchers, students, and the public 
-over the Internet.
+The nature of data from NetBEAMS through SF-BEAMS is used for the purpose of
+\textbf{Data Archival}. After the collected data captured from the sensors, it
+requires further processing in order to be archived for later reuse. Different
+types of users are expected to use the collected data, ranging from
+researchers, students, and the public over the Internet.
 
@@
 
-Since all of developed database systems were implemented with the intension of
-persisting data in long-term storage devices, every single one in the list
-provides support for data persistence and management. Therefore, they are
-compliant to the purpose of Data Archival for the sensor networks.
+Since all of developed database systems were implemented with the intention of
+persisting data in long-term storage devices, every one listed provides
+support for data persistence and management. Therefore, they are compliant to
+the purpose of Data Archival for the sensor networks.
 
@@
 this way, all the collected data is transmitted to that single location
-\ref{sec:sn-infrastructure}. In the same vein, the simplest location to save
-the collected data is through the use of at least an \textbf{External Storage}.
-Similarly, the use of a Data-Centric Storage strategy is considered in
-situations where the data sink is about to reach its storage capacity. As a
-consequence, the \textbf{Data-Centric Storage} separates the collected data
-into dissimilar locations, based on any property of the data. Therefore,
-there are two different approaches to data partitioningThe two different
-flavors:
+\ref{sec:sn-infrastructure}. The simplest location to save the collected data
+is through the use of at least an \textbf{External Storage}. Similarly, the
+use of a Data-Centric Storage strategy is considered in situations where the
+data sink is about to reach its storage capacity. The \textbf{Data-Centric
+Storage} separates the collected data into dissimilar locations, based on any
+property of the data. Therefore, there are two different approaches to data
+partitioning:
 
@@
 The External Storage approach is the most common way to implement persistence
-for any breed of systems since data management is usually offered by most
+for any type of systems since data management is usually offered by most
 database systems out-of-the-box.
@@
 
-The most common problem related to a single External Data Storage is that 
-resources may run into data management problems such reaching the disk storage
-capacity or hardware failure.
+The most common problems related to a single External Data Storage approach are
+related to the capacity of the device and operational problems. If the
+external storage device is situated in the network, then the capacity of the
+device may be limited to receive collected data from the sensors in the
+network, since it may run out of space. Similarly, storage devices are very
+error prone and its hardware part may fail.
 
@@
 datasets when the database server performs a query processing in a database
-table. For this reason, this approach helps the database administrators to
-scale their database infrastructure and improve the database performance.
+table. For this reason, this approach helps the database administrators scale
+their database infrastructure and improve the database performance.
 
@@
 Database Sharding \cite{db-shard-discussion}, or simply Database partitioning
-\cite{db-partitioning-relational}. Both the former and the
-latter are used in regular schema-dependent models or schema-less models,
-which uses denormalized data models. There are different approaches to
-partition data \cite{db-shard-schemas, db-partitioning-relational}.
+\cite{db-partitioning-relational}. Both the former and the latter are used in
+regular schema-dependent models or schema-less models. Different approaches
+worth mentioning:
 
 \begin{itemize}
+  \item \textbf{Hash Partitioning}: a hash key is computed based on
+  different table columns of an entity \cite{db-shard-schemas,
+  db-partitioning-relational}, providing an even distribution of the data;
+  \item \textbf{Key Partitioning}: is based on the hash partitioning over the
+  values of a specific key \cite{db-mongo-partition};
+  \item \textbf{Range Partitioning}: the partition is based on specific ranges
+  of the values of given a given column of a table. For instance, the collected
+  data from the sensors could be separated by the year of its collection. For
+  example, data collected in the year of 2009 would be stored in a different
+  partition than the data collected in the year 2010
+  \cite{db-partitioning-relational};
   \item \textbf{Directory Based Partitioning}: an ordinary lookup service is
-  used to identify which database shard to use, where the lookup table can be
-  stored in a secondary area;
-  \item \textbf{Key or Hash Based Partitioning}: the value of a given key
-  property can  be used as the key of the shard. Moreover, it can be used as 
-  an input to a hash function, whose output determines at which database shard
-  the data needs to be stored.
+  used to identify which database partition to use \cite{db-shard-schemas}.
 \end{itemize}
 
-An example of a Data-Centric storage with different shards can be see in Figure
-\ref{fig:database-sharding-by-region}, where the collected data is segregated
-by its origin. Each shard is assumed to be placed in different server
-hosts, as well as each of them uses the same ``denormalized'' data models.
+An example of a data-centric storage with different partitions can be seen in
+Figure \ref{fig:database-sharding-by-region}, where the collected data is
+partitioned by its origin. Each shard is assumed to be placed in different
+server hosts, as well as each of them uses the same ``denormalized'' data
+models.
 
@@
 \begin{itemize}
-  \item \textbf{Rebalancing Shards}: considering a schema-dependent database
+  \item \textbf{Rebalancing Partitions}: considering a schema-dependent database
   system, changes on the schema of one shard requires rebalancing the
@@
   \item \textbf{Referential integrity}: also considering a schema-dependent
-  database system, any cross-shard query may fall into the problem where
+  database system, any cross-shard query may run into the problem where
   references do no exist in other shards, since the application layer is
@@
 All the selected database systems support the External Storage method. However,
-they all differ with the support to Data-Centric Storage support. Therefore,
-MySQL and mongoDB are promising choices, with the difference of the use of the
-data model.
+they all differ somewhat in terms of Data-Centric Storage support. Therefore,
+MySQL and mongoDB are promising choices, with the difference of the use of the data model.
 
@@
 One of the most common practices in the area of database system is to use the
-relational model to persist data, although the application of the system may
-not fit to solve the problem. In fact, the main users of sensor networks may
+relational model for data persistence, although the application of the system
+may not fit to solve the problem. In fact, the main users of sensor networks may
 not possess any expertise in database systems or data modeling, based on the
 fact that they are from different science areas. \cite{sn-programming-language}
-suggested the use of programming languages abstraction instead of regular
-database systems for researchers. For instance, SF-BEAMS is managed by Marine
-Biologists without expertise in Data Management, Modeling systems, and for
-this reason, one of the requirements of the system is trying the use of
-schema-less approaches. The \textbf{Key-Value-Pair Data Model} or the
-\textbf{Document-Oriented Data Model} seems to be the simplest choices of the
-models.
+suggested the use of programming languages abstraction, instead of regular
+database systems, to provide access to the collected data for both
+technical and non-technical users. For instance, marine biologists manage
+SF-BEAMS without expertise in data management or modeling systems. For this
+reason, one of the requirements of the system is the use of schema-less
+approach, which describes the collected data without data modeling process
+used in schema-dependent models. The \textbf{Key-Value-Pair Data Model} seem
+to be an optional candidate that provides such a method of data design.
 
@@
 Considering the inception of a relational data model \cite{relational-model}
-and the use of the YSI Sonde data as the main entity in the system, let Figure
-\ref{fig:Relational-Model-Original} represent a prototype of the relational 
+and the use of the YSI Sonde data as the main entity in the system, Figure
+\ref{fig:Relational-Model-Original} represents a prototype of the relational 
 model, after passing through the process of normalization
@@
 
-Supposing a new type is added into the system, let the refactored version of
-the relational model be depicted in Figure
-\ref{fig:Relational-Model-Addition-Modified}. Some observations are
-considered in this scenario:
+When a new type is added into the system, the refactored version of the
+relational model be depicted in Figure
+\ref{fig:Relational-Model-Addition-Modified}. Some observations are considered
+in this scenario:
 
@@
 \begin{itemize}
-  \item Data schema that constantly changes require constant database
+  \item Data schema that constantly change requires continuous database
   normalization processes, changes to structure, database management, and all
   the code that uses the data. Schema-Dependent databases always require
-  refactoring of the schema to accommodate changes to the entites' properties;
-  \item The approach of Relational Model might not fit the nature of the
-  collected data from sensor devices, since they tend to change over time. 
-  Furthermore, the support to time-series or provenance are not better
-  supported \cite{sn-provenance};
-  \item Some research have suggested changes to the SQL clauses to better
-  support time-series when collecting data using the relational model 
-  \cite{sn-db-newop}.
+  refactoring of the schema to accommodate changes to the entities' properties;
+  \item \cite{sn-provenance} describes the relational model as problematic 
+  when describing collected data from sensor devices;
+  \item \cite{sn-db-newop} suggests changes to the SQL clauses of TinyDB to
+  better support time-series, an important property when describing.
 \end{itemize}
@@
 emergence of Internet applications, the necessity of such a model that does
-not require constant refactoring led developers to propose such manner using
-the relational model as documented in technical articles such as
-\cite{db-kvp-in-relational01} and \cite{db-kvp-in-relational02}. Based on
-these articles, consider the relational model depicted in Figure
-\ref{fig:KVP-on-Relational-Model} as a data model for persisting data to our
-case study using the key-value strategy. It is clear that this approach
-contains problems: repetition occurs, since the nature of
-time-series data requires the use of a timestamp key for each property of the sensor. Therefore, this strategy is not
-well-suited to provide persistence data from collected sensor networks.
+not require constant refactoring led developers to propose the relational model
+as documented in technical articles such as \cite{db-kvp-in-relational01} and
+\cite{db-kvp-in-relational02}. Based on these articles, consider the
+relational model depicted in Figure \ref{fig:KVP-on-Relational-Model} as a
+data model for persisting data to our case study using the key-value
+pair strategy. It is clear that this approach contains problems: repetition
+occurs, since the nature of time-series data requires the use of a timestamp
+key for each property of the sensor. Therefore, this strategy is not well
+suited to provide persistence data from collected sensor networks.
 
@@
 There are no records of the use of this data model with Sensor Networks.
-Different variants of such data model is the document-oriented data model,
+A different variant of such data model is the document-oriented data model,
 where data is modeled using structured documents. However, these documents
@@
 
-For instance, all data needed for the YSI sonde data, as well as all necessary
-provenance data that describes the data, depicted in section
+For instance, all data needed for the YSI Sonde Data, as well as all necessary
+provenance data that describes the data depicted in section
 \ref{sec:sn-data-description}, are represented by means of key-value pairs
@@
 The relational and key-value data models were compared by
-\cite{db-is-rdbs-dommed} and is compared on table
+\cite{db-is-rdbs-dommed} and is compared on Table
 \ref{tab:schema-vs-schemaless}:
@@
       \item Data Provenance can be provided through data types such as
-      ``Timestamp'' for time or float for the GPS float-based coordinates; 
+      ``timestamp'' for time or float for the GPS float-based coordinates; 
     \end{enumerate} 
@@
       \item Domains are similar to tables, but like buckets that contains items
-      without a pre-defined schema, what enable them to have different schemas;
+      without a pre-defined schema, what enables them to have different schemas;
       \item Items are identified by keys, as well as have a dynamic set of
@@
     \begin{enumerate}
-      \item The basic operations CRUD\footnote{Create-Retrieve-Update-Update
-      database operations} data are performed using a structured language such
+      \item The basic CRUD (Create-Retrieve-Update-Delete)
+      database operations are performed using a structured language such
       as the SQL or XPath;
       \item Query languages can access data from different tables through
-      joins, contains functions for aggregation and complex filter.
+      joins, containing aggregation functions to apply on the returned values.
     \end{enumerate} 
@@
     \begin{enumerate}
-      \item The CRUD operations are performed via API\footnote{Application
-      Programming Interface} through programming languages;
+      \item The CRUD operations are performed via an API (Application
+      Programming Interface) through programming languages;
       \item Some technologies provide basic SQL-like syntax for filter criteria
-      with some predicates like =, !=, <, > that ca be applied;
+      with some predicates like equals to (=), different than (!=), less than
+      (<), greater than (>), among others;
       \item The data and application integrity logic is placed in the
@@
     \begin{enumerate}
-      \item Have their own specific API or through ODBC\footnote{Open Database
-      Connectivity};
+      \item Have their own specific API or through ODBC (Open Database
+      Connectivity);
       \item Data is stored in a format that represents its natural structure,
@@
     \begin{enumerate}
-      \item Systems tend to provide SOAP/REST \cite{http-rest};
+      \item Tend to provide Web Services Interfaces such as the REST
+      \cite{http-rest};
       \item \cite{db-is-rdbs-dommed} claims that data is stored in a more
@@
 
-Since the selected data model is the schema-less one, the following can be seen
-from the technologies:
+The following summarizes which data model each of the candidate database
+systems use.
 
 \begin{itemize}
-  \item \textbf{MySQL}: supports the Relational Data Model;
-  \item \textbf{TinyDB}: supports the Relational Data Model; 
-  \item \textbf{MongoDB}: supports Key-Value Pair Data Model;
-  \item \textbf{IBM DB2}: supports the Relational Model and the Structured
-  Data models.
+  \item \textbf{MySQL}: uses the Relational Data Model with the use of SQL;
+  \item \textbf{TinyDB}: uses the Relational Data Mode with the use of SQL; 
+  \item \textbf{MongoDB}: uses Key-Value Pair Data Model with an Non-Structure
+  Query language in forms of programming language abstraction;
+  \item \textbf{IBM DB2}: uses the Relational and Structured models with SQL
+  and XPath, respectively.
 \end{itemize}
@@
 
-\textbf{Data Description} is the methods used to describe the collected data
-from sensors, specially those produced by real-time data streams. Since those
+\textbf{Data Description} is the method used to describe the collected data
+from sensors, especially those produced by real-time data streams. Since that
 type of data does not include descriptions of its values, data provenance
@@
 
-An additiona to data provenance, annotations is another technique used to
-better describe the collected data from sensor devices. Its main purposes is to
-provide means of identifying observations, creating indexes based on simple
-keywords. For an environmental sensor networks, this is a valuable type of value
-used to better manage the collected data.
+In addition to data provenance, annotation is another technique used to
+better describe the collected data from sensor devices. Its main purpose is to
+provide a means of identifying observations, creating indexes based on simple
+keywords. For an environmental sensor network, this is invaluable used to
+better manage the collected data.
 
@@
 
-The injection of provenance data into the collected data enriches description
-of a given event, and is always useful to determine different aspects of the
-data. As mentioned in chapter 3, data provenance helps identifying the nature
-of the data by the identification of the properties of a given sensor device,
-for example. Similarly, provenance adds spatio-time data in order to correlate
-the observed data with time a given location and time of occurrence.
+The inclusion of provenance data into the collected data enriches the
+description of a given event, and is always useful to determine different
+aspects of the data. As mentioned in chapter 3, data provenance helps identify
+the nature of the data by the identification of the properties of a given
+sensor device, for example. Similarly, provenance adds spatial-time data in
+order to correlate the observed data with time a given location and time of
+occurrence.
 
@@
 Additional data being inserted together with the raw data collected from the
-sensor devices. This additional data may contribute with any decreasing number
-of disk storage for the variable defined. Therefore, the decreasing amount
-of disk space may be one the reasons to not use annotations.
+sensor devices may contribute with a decreasing amount of disk storage. 
+Therefore, the use of provenance-away data increases the required disk space
+used to save the collected data from sensor devices.
 
@@
 
-This type of data does not make part of the collected data, nor provenance
+This type of data does not make up part of the collected data, nor provenance
 information.
@@
 As a direct result from the previous section, the use of \textbf{Centralized Query
-Processing} is indicated for NetBEAMS, also given the fact that SF-BEAMS data
-goes directly to the individual network sink and, therefore, a centralized
-point of the data.
+Processing} is indicated for NetBEAMS, because the SF-BEAMS data goes directly
+to the individual network sink and, therefore, a centralized point of the data.
 
@@
 the size of the datasets and the database technology, performance may be a
-concern. For example, this procedure creates of the alleged Funneling Affect
-\cite{sn-storage04}, since the point-of-traffic is concentrated in the
-centralized system.
+concern due to the alleged Funneling Affect \cite{sn-storage04}, since the
+point-of-traffic is concentrated in the centralized system.
 
@@
 All the database systems technology support a centralized query system. In
-addition to that, they can be distributed into different distributed nodes when
-used in a partioned environement.
+addition to that, they can be sorted into different distributed nodes when
+used in a partitioned environment.
 
@@
 
-Sensor Network data can be saved in either Centralized or Distributed Systems.
-While Centralized Database Systems are easier to manage, it may
-face challenges regarding its data. In order to implement a Data-Centric
-solution, a distribute database system must be in place in order to manage
-the different data by a single node.
+Sensor Network data can be saved into database system that are organized
+by either a Centralized or Distributed System. While a centralized database
+system may be considered to be easier to manage, it may fail to adjust to
+changes of requirements to manage data. On the other hand, a distribute
+database system can be used to implement a data-centric solution. Therefore, a
+database system that can be organized in both approaches is a good candidate to
+be used a data persistence technology for sensor networks.
 
-\subsection{Advantages of Database Partitioning}
+\subsection{Advantages of Distributed Systems}
 
-\begin{itemize}
-  \item Data-Centric queries and data use;
-  \item Solves bottleneck problems related to reads/writes;
-  \item Decrease the funneling effect by directing queries to given data
-  partition;
-\end{itemize}
+Distributed Systems can improve the performance of data management by providing
+techniques for parallel data processing such as Map Reduce \cite{map-reduce,
+bigtable}. It also helps decreasing the system load and bottleneck problems
+related data creation and retrieval. It is also seen as an optional mechanism
+of reducing the funneling affect of a centralized system by directing queries
+to a given data partition \cite{sn-storage04}. Data Replication
+\cite{data-replication} provides a ways to decentralize data access by
+providing duplicated data that is intended to be used on reading operations.
 
-\subsection{Disadvantages of Database Partitioning}
+\subsection{Disadvantages of Distributed Systems}
 
-\begin{itemize}
-  \item Very advanced topics in Database System;
-  \item Partitions need to be rebalanced in case the key of the selected 
-  partitioning strategy is changed.
-\end{itemize}
+Requires more technical knowledge expertise to manage the different systems
+that composed a distributed database.
 
@@
 \begin{itemize}
-  \item \textbf{MySQL}: Supports Data Replication;
-  \item \textbf{TinyDB}: NO Supports to distributed data;
-  \item \textbf{MongoDB}: Supports Data Replication;
-  \item \textbf{IBM DB2}: Supports Data Replication.
+  \item \textbf{MySQL}: provides support to data replication;
+  \item \textbf{TinyDB}: it does not provide support to data replication;
+  \item \textbf{MongoDB}: supports data replication;
+  \item \textbf{IBM DB2}: supports data replication.
 \end{itemize}
@@
 
+Other non-functional requirements to be considered are related to the
+functionalities and type of database systems.
+
 \begin{itemize}
-  \item \textbf{Open-source}: Given the fact of the nature of the project, the
+  \item \textbf{Open-Source}: Given the nature of the project, the
   technology to be used must be free of charge \cite{open-source}, that is, no
-  costs involved in the adoption a such technology; Together with it is the
-  support from the community;
+  costs involved in the adoption a such technology; along with the support from
+  the community;
   \item \textbf{Native APIs}: In order to consider the scope of this work
@@
   mechanisms};
-  \item \textbf{Hot Backup}: Supports hot backup with less service interruption. 
+  \item \textbf{Export Feature}: the technology must support export
+  capabilities.
 \end{itemize}
@@
   scripting APIs from  the community;
-  from the community. Also APIS are available for the majority of languages;
+  from the community. Also APIs are available for the majority of languages;
   \item \textbf{TinyDB}: Not open-source;
   \item \textbf{MongoDB}: An open-source database system with an increasing
-  support from  the community, offering native APIS in different programming
+  support from  the community, offering native APIs in different programming
   languages and scripting languages;
@@
 
-Only MySQL and MongoDB supports the non-functional requirements defined in this
+Only MySQL and mongoDB supports the non-functional requirements defined in this
 section.
@@
 
-Each of the databases were scored according to its support to the different
-types of taxonomies, as well as a different set of requirements. As defined
-in the scope of the database system, the database that can provide a
-programming language abstracton to its data access would be selected.
-Moreover, the database technology should offer the support to schema-less
-databases. As a result, mongoDB was selected because of different reasons.
-First, it is a document-oriented database system, which does require the use of
-SQL languages. Similarly, it had been reported in two different surveys as
-potential database systems. Finally, there were no references in the literature
-regarding the use o KVP databases in sensor networks, and for this reason, this
-proposal was performed, supporting most of the taxonomies defined in the
-previous chapter, it was selected for the implementation of the DSP Data
-Persistence component for NetBEAMS.
+Each of the databases selected was scored according to the support provided to
+the different types of taxonomies and requirements defined for a selection of a
+persistence technology for NetBEAMS. Table \ref{tab:technology-selection}
+summarizes these scores using the (+) and (-) notation. Since they all provide
+support to data archival for a small volume of data, these parameters are not
+included in the table.
 
@@
     \begin{center}
-        \begin{tabular}[!h]{|c|c|c|c|c|c|c|}\hline 
+        \begin{tabular}[!t]{|c|c|c|c|c|c|c|}\hline 
         \textbf{Database} & \textbf{MySQL} & \textbf{TinyDB} & \textbf{MongoDB} & \textbf{IBM DB2}\\\hline
         Centralized Query & + & + & + & + \\\hline 
-        Distributed System & ++ & - & ++ & +\\\hline 
         Schema-less & - & - & + & -\\\hline 
+        Non-Structure Query & - & - & + & -\\\hline 
         Provenance Support & + & + & + & +\\\hline 
-        NO-SQL Query & - & - & + & -\\\hline 
-        Partition & + & - & + & +\\\hline 
-        Export Capability & + & - & + & -\\\hline 
-        Programming Language. & + & + & + & +\\\hline
-        Script Language. & + & - & + & +\\\hline
+        Distributed System & + & - & + & +\\\hline 
+        Data Partition & + & - & + & +\\\hline 
+        Export Capabilities & + & - & + & -\\\hline 
+        Native APIs & + & + & + & +\\\hline
         Open-Source & + & + & + & -\\\hline
         \end{tabular}
-        \caption{Technologies Selection Grading}
+        \caption{Technologies Selection Scoring}
     \end{center}
@@
 
-To conclude, this chapter evaluated different database systems selected from
-the literature reviews of chapter 2. After presenting the database system
-``contenders'', the analysis of each of the taxonomies were conducted for each
-of the technologies, culminating with the technology selection comparison. In
-this way, the following chapter describes the design of the solution using
-mongoDB as the primary solution for the backend.
+In general, all the selected databases can be used for the purpose of data
+storage, also associated with the location of the collected data being in an
+external storage. Since the data volume produced by NetBEAMS is small, a
+typical centralized server can be used to provide management of a database
+management system, and consequently a centralized query mechanism. Additionally,
+a distributed system can also be used in the case of future system expansion.
+All technologies, but TinyDB, fit to these requirements.
+
+All the database systems give support to data provenance as a feature or
+data types. Nonetheless, when it comes to the Data Model used, the majority of
+the technologies use schema-dependent data model. The abstraction of
+programming languages is reported to give users, without skills in database
+systems, a better method of data access to the collected data from sensor
+networks. In this way, although all the discussed tools provide access to data
+through an additional API, mongoDB was the only one to use non-structured
+query, as a consequence of using a schema-less data model, which is the type
+used by SF-BEAMS staff members to describe the collected data. For this reason,
+mongoDB was considered as a good candidate to be explored as a data persistence
+solution for NetBEAMS.
+
+To conclude, this chapter evaluated different database systems filtered
+from Chapter 2. After presenting the database system ``contenders'', the
+analysis of each of the taxonomies was conducted for each of the technologies,
+culminating with the technology selection comparison. The following chapter
+describes the design of the solution using mongoDB as the primary solution for
+the backend.

### 2009-12-10 docs/marcello-dissertation/thesis/chapter2-state-of-the-art.tex
Reviews of chapter 4 of the report... Adding a label on chapter 2 and adding references to non-cited sections.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@689 559edde2-a4d4-44e0-96fc-4035e40d7396
Revision on the feedback, chapter 2. -	Fixing typos -	Changing from temperature-scale to temperature unit -	Giving a practical example of transaction-time -   Giving a practical example about specific data location (precision) -   No additional information added for radar data produced (no references)
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@691 559edde

@@
 \section{Persistence Storage for Sensor Networks}
+\label{sec:sn-persitence-storage}
 
@@
 in streams of values with any type of delimitation between each other. For
-this reason, the processes of parsing and labeling each of the values located
+this reason, the process of parsing and labeling each of the values located
 in the data stream regularly requires the use of the device's specification
@@
 movement and their visual appearance using a video camera. In this way,
-adding tags to particular video frames can describe different them, and
+adding tags to particular video frames can describe each of them, and
 therefore, providing an efficient way to search for a specific frame based on
@@
 the property describing the raw collected data 73.45. On the other hand,
-the example \textbf{temperature-scale} = ``Fahrenheit" describes the metadata
+the example \textbf{temperature-unit} = ``Fahrenheit" describes the metadata
 regarding the temperature.
@@
 reports that all the data stored by its sensor network uses the time of
-collection, but do not use the transaction time for that purpose.
+collection, but do not use the transaction time for that purpose. Similarly,
+\cite{sn-provenance} shows a practical examples of critical queries performed
+over the transaction time, when doctors need all the collected data regarding
+the heart beat rate from the moment the patient entered the ambulance until
+the moment of arrival at the hospital.
 
@@
 gives necessary identification of where the event of high-pressure occurred.
+The precision of the location can be important for the collected data because
+it helps researchers correlate different measurements based on exact locations
+of the data. For instance, \ref{sn-provenance} depicts an example where two
+different sensor networks covering the same physical area: one covering the
+reginal traffic and another snowfall. Although each of them has individual
+functions, collectively they could allow real-time navigation service to answer
+queries regarding expected traffic patterns based on the current snowfall and
+advise drivers accordingly by reusing collected data from both sensor networks.
 

### 2009-12-10 docs/marcello-dissertation/thesis/chapter2-state-of-the-art.tex
Revision of chapter 3, after the audio feedback from Prof. Puder. Also includes changes to chapter 2 (lower case Section)
- Capitalizing all the section to Section in all documents;
- Data-Centric storage is the term used in the community. Added reference;
- Changing drawn to figure
- Fact changed to Valid... (Old reference)
- Updating the Data Description Taxonomy Diagram, showing Annotation;
- Adding the description of relating system organization and type of query used on each of them.
- Fixed word agreement
- Additional description about the relationships.


git-svn-id: svn://localhost/net

@@
 location of the collected data is a local storage site \cite{sn-storage01,
-sn-storage03} within the network; a strategy typically used to maximize the
+sn-data-centric-storage} within the network; a strategy typically used to maximize the
 sensor's resources, while minimizing energy consumption \cite{sn-storage04}. On
@@
 environment, the sensor node is responsible for sending the raw data to a specific
-data node defined by the network infrastructure, as delineated in section
-\ref{sec:sn-infrastructure}. Thus, \cite{sn-storage03}
+data node defined by the network infrastructure, as delineated in Section
+\ref{sec:sn-infrastructure}. Thus, \cite{sn-data-centric-storage}
 describes that the destination for the collected sensor data will be one of two
@@
 until it reaches its final data sink when the data organization is structured
-as a graph, tree \cite{sn-storage01, sn-storage03}.
+as a graph, tree \cite{sn-storage01, sn-data-centric-storage}.
 
@@
 accessing the collected data stored at data nodes, whose location is determined
-by different facts described in the previous section. \cite{sn-storage03}
+by different facts described in the previous section. \cite{sn-data-centric-storage}
 describes an example of an \textbf{in-network query processing} when sensor
@@
 each of the query processing approaches described with regards to the
-infrastructure of the sensor network. \cite{sn-storage03} argues that the 
+infrastructure of the sensor network. \cite{sn-data-centric-storage} argues that the 
 primary purpose of the in-network query processing is to minimize the amount
@@
 this reason, \cite{sn-storage04} describes that different techniques used in 
-distributed systems such as the use of master-slave data replication or 
-database partitioning \cite{db-partitioning-relational} can be used to reduce
-or mitigate the effects of this funneling.
+distributed systems such as the use of master-slave data replication
+\cite{sn-data-center-replication-load-balance} or database partitioning
+\cite{db-partitioning-relational} can be used to reduce or mitigate the
+effects of this funneling.
 
@@
 and retrieve data, since it requires more computational power and data
-retrieval efficient algorithms. \cite{sn-data-center} describes the creation of
+retrieval efficient algorithms. \cite{sn-data-center-replication-load-balance} describes the creation of
 a Sensor Data Center designed to store and analyze amounts of data on the range of

### 2009-12-10 docs/marcello-dissertation/thesis/chapter3-taxonomic-classification-of-persistence-storage-for-sensor-networks.tex
Revision of chapter 3, after the audio feedback from Prof. Puder. Also includes changes to chapter 2 (lower case Section)
- Capitalizing all the section to Section in all documents;
- Data-Centric storage is the term used in the community. Added reference;
- Changing drawn to figure
- Fact changed to Valid... (Old reference)
- Updating the Data Description Taxonomy Diagram, showing Annotation;
- Adding the description of relating system organization and type of query used on each of them.
- Fixed word agreement
- Additional description about the relationships.


git-svn-id: svn://localhost/net

@@
 
-Taxonomy is defined as the practice and science of classification and uses
-the taxonomic units of ``taxa'', the plural form of ``taxon'' \cite{taxonomy}.
-Similarly, they are represented by a hierarchical diagram that relates the
-main taxonomy to the different taxa and their own related ones, if any.
+The root of word ``Taxonomy'' is the Greek taxis, meaning 'order' or
+'arrangement', and, nomos in Latin, meaning 'law' or 'science'). In
+short, taxonomy is defined as the practice and science of classification and
+uses the taxonomic units of ``taxa'', the plural form of ``taxon''
+\cite{taxonomy}. Similarly, hierarchical diagram is used to relate each of the
+taxa to their main taxon.
 
@@
 
-As delineated in section \ref{sec:sn-data-purpose}, the purpose of the
+As delineated in Section \ref{sec:sn-data-purpose}, the purpose of the
 collected data from sensor devices determines how the collected data is
@@
 in the different aspects of the access to the collected data from sensor
-devices, as shown in section \ref{sec:sn-storage-locations}. Figure 
+devices, as shown in Section \ref{sec:sn-storage-locations}. Figure 
 \ref{fig:taxonomy-data-location} depicts this taxonomy, which can be
@@
   database. When the collected data occupies different external storage devices, the
-  strategy is called \textbf{Data-Centric Storage}, being organized by
-  categories based on the collected data keys and values. In other words, what
-  characterizes this taxon is the presence of a data partitioning strategy to
-  store the collected data.
+  strategy is called \textbf{Data-Centric Storage}
+  \cite{sn-data-centric-storage}, being organized by categories based on the
+  collected data keys and values. In other words, what characterizes this
+  taxon is the presence of a data partitioning strategy to store the collected
+  data.
 \end{itemize}
@@
 collected data. This taxonomy relates to the data model used to design the 
-collected data, as drawn in Figure \ref{fig:taxonomy-data-model}, and is 
+collected data, as shown in Figure \ref{fig:taxonomy-data-model}, and is 
 described as follows.
@@
 way to describe or label data can follow different guidelines such as data provenance
-and annotations, discussed in section \ref{sec:sn-data-description}. The description of the
+and annotations, discussed in Section \ref{sec:sn-data-description}. The description of the
 properties should take into account the different values of the nature of the
-data and any other description needed. Therefore, the taxa that composes the
+data and any other description needed. Therefore, the taxa that compose the
 Data Description taxonomy, shown in Figure
@@
   \textbf{When: Time Dimension}, is related to instances of time relative to the
-  collected data: \textbf{fact} and \textbf{transaction} times. The former is
+  collected data: \textbf{valid} and \textbf{transaction} times. The former is
   used to identify when the data was collected, whereas the latter when the data was
@@
 The query processing mechanism in sensor networks is directly related to how
-and where the collected data is used and stored, as shown in section
+and where the collected data is used and stored, as shown in Section
 \ref{sec:query-process}. The following taxa are related to the
@@
 1000\superscript{4} = 10\superscript{12}) to future Petabytes (1PB =
-1000\superscript{5} = 10\superscript{15} bytes) as shown in section
+1000\superscript{5} = 10\superscript{15} bytes) as shown in Section
 \ref{sec:data-load}. As a comparison of what is processed, Google reports to
@@
   centralized host, the focus being the managing of the collected data
-  \cite{sn-intro01};
+  \cite{sn-intro01}. When the system is organized in a centralized system,
+  usually a centralized query processing is used to process any request to
+  operations on the database management system;
   \item \textbf{Distributed System}: a database system that can be composed of
   more than one node. Usually, the use of the pattern of Master-Slave
-  mechanism or data ``Replication" is used to alleviate ``hot spots'' during
-  concurrent processes of data collection from the sensor devices and data
-  access from the final users. Unique mechanisms can be used to disaggregate 
-  the produced data from a concentrated one: the use of database partitioning, 
-  as related to the data-centric storage, where each partition is responsible 
-  to handle data produced by a given type of sensor, for instance.
+  mechanism or data replication \cite{sn-data-center-replication-load-balance} 
+  is used to alleviate ``hot spots'' during concurrent processes of data
+  collection from the sensor devices and data access from the final users.
+  Unique mechanisms can be used to disaggregate the produced data from a
+  concentrated one: the use of database partitioning, as related to the
+  data-centric storage, where each partition is responsible to handle data
+ produced by a given type of sensor, for instance. Finally, it is the type of
+ system used to support in-network query processing.
 \end{itemize}
@@
 The taxonomy related to the data model can be linked to any taxon from the 
-taxon described from the system organization, to data description and query
-processing and location of the collected data. In the former, the data of a
-relational model is stored using SQL messages.
+system organization because it is used in both centralized and distributed
+systems, also associated with the location of the data and, consequently, the
+query processing method used. Furthermore, the data description taxon can be
+used by any of the data models.
+
+The query processing is directly related to the location of the collected data
+and with the system organization taxa. While centralized systems use centralized
+query processing methods, in-network query processing is when the collected
+data is located in different external storage devices organized in a
+data-centric way, usually supported by a distributed system.
 
@@
 network usually determines the system organization used to handle and process
-the collected data. When the volume is small, the use a single server is
+the collected data. When the volume is small, the use of a single server is
 required, usually using an external storage device, which relates to the
-taxonomy for storage system. However, when the volume is medium to large, the
+taxonomy for storage systems. However, when the volume is medium to large, the
 use of distributed systems is required to cope with high volume of data

### 2009-12-11 docs/marcello-dissertation/thesis/chapter4-netbeams-in-depth-requirements-analysis-and-specification.tex
Reviews of chapter 4 of the report... Adding a label on chapter 2 and adding references to non-cited sections.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@689 559edde2-a4d4-44e0-96fc-4035e40d7396
Review of the audio feedback from Prof. Puder.
- Niki suggested the use of in, as Prof. Puder asked to verify;
- Adding the name of the YSI Sonde, referring to the manufacturer YSI Incorporated
- Adding the dimensions information of the device, as well as the name some properties collected
- Fixing the table reference, actually changing the data stream from table to Listing
- Ad

@@
 provides an automated infrastructure solution for SF-BEAMS, and this last component 
-will be covered first. Then, this work proposes a categorization of NetBEAMS to 
-the face of the taxonomies.
+will be covered first. Then, this work proposes a categorization of NetBEAMS
+according to the defined taxonomies.
 
@@
 marine and estuarine environments using the SF-BEAMS sensors. The sensors are deployed off a
-pier located on the San Francisco Bay, in Tiburon, California; and is operated by 
+pier located in the San Francisco Bay, in Tiburon, California; and is operated by 
 the Romberg Tiburon Center for Environmental Studies (RTC) a part of San Francisco
 State University. NetBEAMS is an experimental component that utilizes the SF-BEAMS 
-infrastructure to operate. This section details SF-BEAMS in general, and presents 
-the requirements for data persistence.
+infrastructure to operate. This section details SF-BEAMS in general, and
+presents the requirements for data persistence after classifying NetBEAMS based
+on the taxonomies defined in the previous chapter.
 
@@
 with a laptop computer. The current data collection process for SF-BEAMS is
-described in Figure
-\ref{fig:SF-BEAMS-system-architecture}\cite{sfbeams-current-system}. Upon
-collecting data from sensors, the RTC staff uses automation scripts written in
-Matlab \cite{matlab} to process, index and distribute the raw data in
-different formats. One of these formats is the OPeNDAP \cite{opendap}, which
-is widely used at research institutions to promote easier data exchange
-amongst them. The SF-BEAMS website provides access to the collected data through
-the Internet at http://sfbeams.sfsu.edu:8080/opendap. An example of an HTTP
-Request from accessing the data for a specific YSI \cite{YSI-Sonde} is shown in
-Listing \ref{file:rtc-ysi-opendap}.
+described in Figure \ref{fig:SF-BEAMS-system-architecture} and documented at
+\cite{sfbeams-current-system}. Upon collecting data from sensors, the RTC
+staff uses automation scripts written in Matlab \cite{matlab} to process,
+index and distribute the raw data in different formats. One of these formats
+is the OPeNDAP \cite{opendap}, which is widely used at research institutions
+to promote easier data exchange amongst them. The SF-BEAMS website provides
+access to the collected data through the Internet at
+http://sfbeams.sfsu.edu:8080/opendap. An example of an HTTP Request from
+accessing the data for a specific device is shown in Listing
+\ref{file:rtc-ysi-opendap} on page \pageref{file:rtc-ysi-opendap}.
 
@@
 
-As described in section \ref{sec:sn-infrastructure}, sensor devices produce
+As described in Section \ref{sec:sn-infrastructure}, sensor devices produce
 the observed data based on properties of measurements defined by its
-manufacturer. NetBEAMS has used ``YSI 6600 ESD V2" \cite{YSI-Sonde}, as seen in
-Figure \ref{fig:ysi-device}, one of the sensors devices used by SF-BEAMS. It is a
-powerful water quality-monitoring device, capable of producing around 52 bytes
-on a single real-time data stream reading, as shown on Table
-\ref{tab:ysi-data-stream}.
+manufacturer. SF-BEAMS has used, among other devices, the ``6600 ESD V2" Sonde
+\cite{YSI-Sonde}, manufactored by YSI Incorporated, as seen in Figure
+\ref{fig:ysi-device}. Although it is a small device of under 55 cm in length,
+8.9 cm in diameter, and weighting approximately 3.18 kg, it is a powerful water
+quality-monitoring device, capable of producing around 52 bytes on a single
+real-time data stream reading regarding 12 different parameters such as water
+temperature, salinity, turbiny, among others. An example of the data stream is
+as shown in Listing \ref{data:ysi-stream} on page \pageref{data:ysi-stream}.
 
@@
 
-\begin{table}
-    \begin{center}
-        \begin{tabular}{|l|}\hline
-  12.20    192    179 5588.40   0.09   0.084   0.059  7.98   -79.6   99.5  8.83  0.4     8.7\\\hline
-        \end{tabular}
-    \end{center}
-    \label{tab:ysi-data-stream}
-    \caption{The Collected Data Stream from the YSI Sonde 6600ESDV2}
-\end{table}
+\lstset{label=data:ysi-stream,caption=The Collected Data Stream from the
+YSI Sonde 6600ESDV2}
+\begin{lstlisting}
+12.20    192    179 5588.40   0.09   0.084   0.059  7.98   -79.6   99.5  8.83  0.4     8.7
+\end{lstlisting}
 
-In order to have access to the observed data, the network staff downloads the
-data using one of the device's connections such as the RS-232 serial connector
+In order to have access to the observed data, the RTC network staff downloads
+the data using one of the device's connections such as the RS-232 serial connector
 \cite{rs232}. Then, they upload the data into their data servers where they are
@@
 estimated by analyzing the data distribution over a period of one year, as
-shown in Table \ref{tab:ysi-data-distribution}:
+shown in Table \ref{tab:year-data-distribution}.
 
 \begin{table}[!b]
-    \label{tab:ysi-data-distribution}
+    \label{tab:year-data-distribution}
      \begin{center}
       \begin{tabular}{|c|c|c|c|c|c|c|}\hline 
-        \textbf{YSIs} & \textbf{Rate} & \textbf{Hourly} & \textbf{Daily} &
-        \textbf{Weekly} & \textbf{Monthly} & \textbf{Yearly}\\\hline 
+        \textbf{\# of YSIs} & \textbf{Rate} & \textbf{Hourly} &
+        \textbf{Daily} & \textbf{Weekly} & \textbf{Monthly} & \textbf{Yearly}\\\hline 
         1 & 1 min & 3.04 Kb & 73.12 Kb & 511.87 Kb & 1.99 Mb & 23.99 Mb\\\hline 
@@
         \end{tabular}
+        \caption{Estimated data volume produced by the RTC's YSI sondes}
       \end{center}
-    \caption{Amount of data produced by the RTC's YSI sondes}
 \end{table}
@@
   lab, a quality assurance process is executed, and the data is transformed
-  into the format used by OPeNDAP. This conversion carries data regarding
-  the time of the data collection, therefore, one of the
-  \textbf{Time Dimensions} such as the valid time is used. Since the files are stored
-  with specific file name structure, using timestamps to describe the data, the
-  use of transaction time may be also considered. Finally, the contents of the
-  files contain the \textbf{Data Identity} as shown in the samples of
-  the collected in Listing \ref{file:rtc-ysi-opendap};
+  into the format used by OPeNDAP. The process to describe the collected data
+  includes the time of the data collection, therefore, one of the \textbf{Time
+  Dimensions} such as the valid time is used. Since the files are stored with
+  specific file name structure, using timestamps to describe the data, the use
+  of transaction time may be also considered. Finally, the contents of the
+  files contain the \textbf{Data Identity} as shown in the samples of the
+  collected data in Listing \ref{file:rtc-ysi-opendap} on page
+  \pageref{file:rtc-ysi-opendap};
   \item \textbf{Query Processing Mechanism}: similar to the data stored in a
@@
 designates a product that is produced and sold in bulk} embedded devices and
-open-source \cite{open-source} software, the research group developed a second
-version of NetBEAMS, a component-based approach which suggested the improvement of 
-operational activities from SF-BEAMS using systems automation. In brief, one of the 
-ongoing problems of NetBEAMS was regarding Data Persistence, the primary motivation of
-this report.
+open-source \cite{open-source} software, the research group developed NetBEAMS,
+component-based approach which suggested the improvement of operational
+activities from SF-BEAMS using systems automation. In brief, one of the
+ongoing problems of NetBEAMS was regarding Data Persistence, the primary
+motivation of this report.
 
@@
 The scope of this work is defined as to provide a persistence layer for
-NetBEAMS, making a selection of a persistence layer that is an exception to the
-taxonomical classification of SF-BEAMS. Furthermore, the technology to be
-selected for the persistence layer must take into account the requirements of
-NetBEAMS primary users such as Computer Science researchers, as well as end
-users such as Biologists from the RTC laboratories. The requirements for the
-data persistence specifically for NetBEAMS are described in the following
-section.
+NetBEAMS, making a selection of a persistence layer based on the taxonomical
+classification of SF-BEAMS. Furthermore, the technology to be selected for the
+persistence layer must take into account the requirements of NetBEAMS primary
+users such as Computer Science researchers, as well as end users such as
+Biologists from the RTC laboratories. The requirements for the data
+persistence specifically for NetBEAMS are described in the following section.
 

### 2009-12-11 docs/marcello-dissertation/thesis/chapter4-netbeams-in-depth-requirements-analysis-and-specification.tex
Revision of Prof. Puder's audio feedback of chapter 5, as well as entire restructure of the document, such as using more page width and height, making it more visual. Table references not working automatically, as well as page numbers (will review later). Table references are entered manually in the text.
The structure of chapter 5 changed according to the fact that some
sections, subsections were very short, with 1 paragraph. So, I decided
to use two different options:

* If there are more to be discussed, a new advantages/disadvantages
subsection is created; or else, one section includes two

@@
 estimated by analyzing the data distribution over a period of one year, as
-shown in Table \ref{tab:year-data-distribution}.
+shown in Table 4.1.
 
-\begin{table}[!b]
+\begin{table}[!h]
     \label{tab:year-data-distribution}
@@
 NetBEAMS. Meanwhile, more details regarding the architecture of NetBEAMS, as well as
-the development guidelines can be found in the appendix of chapter 10.
+the development guidelines can be found in the appendix of Chapter 10.

### 2009-12-11 docs/marcello-dissertation/thesis/chapter5-taxonomic-persistence-storage-selection-for-netbeams.tex
Revision of Prof. Puder's audio feedback of chapter 5, as well as entire restructure of the document, such as using more page width and height, making it more visual. Table references not working automatically, as well as page numbers (will review later). Table references are entered manually in the text.
The structure of chapter 5 changed according to the fact that some
sections, subsections were very short, with 1 paragraph. So, I decided
to use two different options:

* If there are more to be discussed, a new advantages/disadvantages
subsection is created; or else, one section includes two

@@
 %\bibliography{thesis}  % uses the references stored in Chapter1Radar.bib
-
+\setcounter{table}{1}
 \chapter{Empirical Analysis of Data Persistence Taxonomies for NetBEAMS}
@@
   \item Must be able to scale in any increases to the data load of
-  the sensor network, as described in chapter 2;
+  the sensor network, as described in Chapter 2;
   \item  Must be able to use single or multiple servers to improve performance
@@
 \begin{itemize}
-  \item \textbf{MySQL}\cite{mysql}: system used in the implementation of a
-  data warehouse with sensor network data \cite{sn-dataware-house};
-  \item \textbf{TinyDB}\cite{tinydb}: as referred in Section 
-  \ref{sec:sn-persitence-storage}, TinyDB is a database system used in many
-  different sensor networks implementation such as \cite{sn-db-tinydb};
+  \item \textbf{MySQL}\cite{mysql}: an open-source relational database
+  system popularized by the development of Internet applications, maintained
+  by Sun Microsystems. \cite{sn-dataware-house} reports its use to develop a
+  data warehouse for data collected from sensor devices using this technology;
+  \item \textbf{TinyDB}\cite{tinydb}: developed and maintained by the Computer
+  Science department of University of California, Berkeley, TinyDB is a
+  relational database system specially developed for sensor devices. As
+  referred in Section \ref{sec:sn-persitence-storage}, it is system used in
+  many different sensor networks implementation such as \cite{sn-db-tinydb,
+  sn-db-newop};
   \item \textbf{MongoDB}\cite{mongodb}: an open-source KVP database
-  system suggested by different surveys to be alternatives to relational
-  databases \cite{db-is-rdbs-dommed, cloud-comp-architectures};
+  system suggested by different surveys to be an alternative to relational
+  databases \cite{db-is-rdbs-dommed, cloud-comp-architectures}. It is maintained
+  by 10Gen Incorporated;
   \item \textbf{DB2}\cite{db2}: in the pursuit of an XML database system, DB2
   is listed as a hybrid database system that offers both the relational and the
-  XML models \cite{db-xml-enabled}. Furthermore, it was selected because
-  NetBEAMS represents serialized objects in XML, as shown
-  \cite{netbeams-dsp-architecture}.
+  XML models \cite{db-xml-enabled}. It is privately owned by IBM.
 \end{itemize}
@@
 The following sections review these technologies' capabilities to adhere to
-the specifications of the taxonomies defined in chapter 3, as well as the
+the specifications of the taxonomies defined in Chapter 3, as well as the
 requirements and the scope for a data persistence of NetBEAMS, as defined in
-chapter 4.
+Chapter 4.
 
@@
 
+The advantages of the External Storage approach is the most common way to
+implement persistence for any type of systems since data management is usually
+offered by most database systems out-of-the-box. It is basically used to store
+small volumes of data. On the other hand, the most common problems related to a
+single External Data Storage approach are related to the capacity of the
+device and operational problems. If the external storage device is situated in
+the network, then the capacity of the device may be limited to receive
+collected data from the sensors in the network, since it may run out of space.
+Similarly, storage devices are very error prone and its hardware part may fail.
+
 As shown in the previous chapter, NetBEAMS's architecture is based on sensor
 network whose topology is a single-hop star with a centralized data sink. In
-this way, all the collected data is transmitted to that single location
-\ref{sec:sn-infrastructure}. The simplest location to save the collected data
-is through the use of at least an \textbf{External Storage}. Similarly, the
-use of a Data-Centric Storage strategy is considered in situations where the
-data sink is about to reach its storage capacity. The \textbf{Data-Centric
-Storage} separates the collected data into dissimilar locations, based on any
-property of the data. Therefore, there are two different approaches to data
-partitioning:
+this way, all the collected data is transmitted to that single location as
+described in Section \ref{sec:sn-infrastructure}. The simplest location to save
+the collected data is through the use of at least an \textbf{External
+Storage}. Similarly, the use of a Data-Centric Storage strategy is considered
+in situations where the data sink is about to reach its storage capacity. The
+\textbf{Data-Centric Storage} separates the collected data into dissimilar
+locations, based on any property of the data. Therefore,
+there are two different approaches to data partitioning:
 
@@
   \item \textbf{Horizontal Partitioning}: the data segmentation is implemented
-  based on the values of the table rows. That is, all the values defined by the
-  columns are used in a given table physically located in specific servers.
-  For instance, horizontal partitioning can be defined as a collection of five
-  years worth of historical sensor data, partitioned into five separate
-  servers. In this way, each server contains data related to each of the years;
+  based on the values of the table rows, in the relational model point of view,
+  or the collections elements, on KVP databases. That is, all the values defined
+  by the columns are used in a given table physically located in specific
+  servers. For instance, horizontal partitioning can be defined as a
+  collection of five years worth of historical sensor data, partitioned into
+  five separate servers. In this way, each server contains data related to
+  each of the years;
   \item \textbf{Vertical Partitioning}: data is partitioned vertically in a way
-  that the columns of an entity are divided into different database tables. In
-  this way, the particular dataset is broken down to different tables. For
-  example, if the database is used to store pictures of a given sensor camera
-  in a BLOB column, it would be the most preferable to the scenario where the
-  images are not constantly referenced when compared to the textual data.
+  that the columns of an entity on the relational model are divided into
+  different database tables. In this way, the particular dataset is broken
+  down to different tables. For example, if the database is used to store
+  pictures of a given sensor camera in a BLOB column, it would be the most
+  preferable to the scenario where the images are not constantly referenced
+  when compared to the textual data. Since KVP uses denormalized data, this
+  partitioning strategy is not used.
 \end{itemize}
-
-\subsection{Advantages of the External Storage Approach}
-
-The External Storage approach is the most common way to implement persistence
-for any type of systems since data management is usually offered by most
-database systems out-of-the-box.
-
-\subsection{Disadvantages of the External Storage Approach}
-
-The most common problems related to a single External Data Storage approach are
-related to the capacity of the device and operational problems. If the
-external storage device is situated in the network, then the capacity of the
-device may be limited to receive collected data from the sensors in the
-network, since it may run out of space. Similarly, storage devices are very
-error prone and its hardware part may fail.
 
@@
   partition than the data collected in the year 2010
-  \cite{db-partitioning-relational};
-  \item \textbf{Directory Based Partitioning}: an ordinary lookup service is
-  used to identify which database partition to use \cite{db-shard-schemas}.
+  \cite{db-partitioning-relational}.
 \end{itemize}
@@
 
-\begin{figure}
+\begin{figure}[!h]
   \centering
@@
   database system, any cross-shard query may run into the problem where
-  references do no exist in other shards, since the application layer is
+  references do not exist in other shards, since the application layer is
   responsible for enforcing data integrity. In this way, examples of such
@@
 SF-BEAMS without expertise in data management or modeling systems. For this
-reason, one of the requirements of the system is the use of schema-less
+reason, one of the requirements of the system is the use of a schema-less
 approach, which describes the collected data without data modeling process
-used in schema-dependent models. The \textbf{Key-Value-Pair Data Model} seem
+used in schema-dependent models. The \textbf{Key-Value-Pair Data Model} seems
 to be an optional candidate that provides such a method of data design.
@@
 
-\begin{figure}
+\begin{figure}[!t]
   \centering
-  \includegraphics[scale=0.5]{../diagrams/Relational-Model-Original}
+  \includegraphics[scale=0.65]{../diagrams/Relational-Model-Original}
   \caption{Relational Data Model for NetBEAMS - A first prototype}
@@
 
-\begin{figure}
+\begin{figure}[!b]
   \centering
-  \includegraphics[scale=0.5]{../diagrams/Relational-Model-Addition-Modified}
+  \includegraphics[scale=0.6]{../diagrams/Relational-Model-Addition-Modified}
   \caption{Relational Data Model for NetBEAMS - Modified version with new
@@
   \item \cite{sn-db-newop} suggests changes to the SQL clauses of TinyDB to
-  better support time-series, an important property when describing.
+  better support time-series, an important property when describing
+  collected data from with spatial-temporal properties.
 \end{itemize}
@@
 querying documents, although some hybrid \cite{db2} technologies may still use
-SQL \cite{sql} for that matter. In this way, a materialized version of the
-data from the defined previous schema can be seen in Figure
+SQL \cite{sql} for that matter. In this way, the view of an example of database
+state for the defined previous schema can be seen in Figure
 \ref{fig:persistence-example-relational}.
@@
 
-Although this data model has been used in different types of applications,
-developers have tried to use the Relational Model to design an application that
-could prevent any database change through the use of a technique that relates a
-key to a value, called Key-Value data model. As mentioned earlier, with the
-emergence of Internet applications, the necessity of such a model that does
-not require constant refactoring led developers to propose the relational model
-as documented in technical articles such as \cite{db-kvp-in-relational01} and
-\cite{db-kvp-in-relational02}. Based on these articles, consider the
-relational model depicted in Figure \ref{fig:KVP-on-Relational-Model} as a
-data model for persisting data to our case study using the key-value
-pair strategy. It is clear that this approach contains problems: repetition
-occurs, since the nature of time-series data requires the use of a timestamp
-key for each property of the sensor. Therefore, this strategy is not well
-suited to provide persistence data from collected sensor networks.
+This data model has been used to solve data models for different types of
+application. One such a problem is the one to represent entities by a list of
+properties with keys and relating values, called Key-Value data model
+\cite{db-kvp-in-relational01, db-kvp-in-relational02}. The author of
+\cite{db-kvp-in-relational01} suggests that this is used as a technique used
+when the addition of number of properties of an entity constantly changes, and
+thus, requires constant refactoring of the relational data model. Considering
+this data model for sensor network devices, whose properties are lists of keys
+and relating values, an ``one-fits-all'' attempt to use the KVP data model
+using the relational data model is depicted in Figure
+\ref{fig:KVP-on-Relational-Model}, and a given example of the state of a
+database using this schema in Figure
+\ref{fig:persistence-example-relational-kvp}.
 
@@
 
+\newpage
+
 \subsection{Analysis of the Schema-less Models}
@@
 its data. One such example is called Key-Value Pair (KVP) data model
-\cite{db-kvp}, which is also known as Document-Oriented, Distributed Hash
-Table. Such a model has the following properties:
+\cite{db-kvp}, and it is characterized as follows:
 
@@
 
-There are no records of the use of this data model with Sensor Networks.
-A different variant of such data model is the document-oriented data model,
-where data is modeled using structured documents. However, these documents
-have a dynamic structure, freely described without the use of a
-schema validation mechanism as used in XML Schema. One example of such
-approach is the use of the JSON data format \cite{json}. mongoDB implements the
-following:
+There are no records of the use of this data model with sensor networks. Data
+is generally aggregated by given entities and any needed property applied to
+it. Key may be composed by embedded keys having a dynamic structure and freely
+used without a schema validation mechanism as used schema-dependent data
+models, having the following organization:
 
@@
   \item entities from the same domains are placed into a ``bucket'';
-  \item entities have a set of attributes and relating values
+  \item entities have a set of attributes and relating values;
   \item entities with different set of attributes may be contained in the same
@@
 For instance, all data needed for the YSI Sonde Data, as well as all necessary
-provenance data that describes the data depicted in section
-\ref{sec:sn-data-description}, are represented by means of key-value pairs
-\ref{fig:persistence-example-kvp}.
+provenance data and annotation as defined in Section
+\ref{sec:sn-data-description}, are represented by means of key-value pairs as
+seen in Figure \ref{fig:persistence-example-kvp}.
 
@@
 
-The relational and key-value data models were compared by
-\cite{db-is-rdbs-dommed} and is compared on Table
-\ref{tab:schema-vs-schemaless}:
+The properties of the relational and key-value data models were compared by
+\cite{db-is-rdbs-dommed} and summarized in Table 5.1:
 
-\begin{table}
-    \label{tab:schema-vs-schemaless}
+\begin{table}[!b]
+    \label{tab:properties-schema-vs-schemaless}
     \caption{Schema-Dependent X Schema-less Databases Compared: Properties}
@@
 
-\begin{table}
-    \label{tab:schema-vs-schemaless-daccess}
+\newpage
+
+In order to have access to the persisted data, both types of data models
+support methods: one via structured languages, and the other via abstractions
+of programming languages. \cite{db-is-rdbs-dommed} summarizes the differences
+shown in Table 5.2:
+
+\begin{table}[!h]
+    \label{tab:daccess-schema-vs-schemaless}
     \caption{Schema-Dependent X Schema-less Databases: Data Access}
@@
     \begin{enumerate}
-      \item The CRUD operations are performed via an API (Application
-      Programming Interface) through programming languages;
+      \item The CRUD operations are performed via an abstraction of
+      programming language. (E.g. ``db.collection.find()'',
+      ``db.collection.insert()'');
       \item Some technologies provide basic SQL-like syntax for filter criteria
       with some predicates like equals to (=), different than (!=), less than
-      (<), greater than (>), among others;
+      or equals ($\leq$), greater than or equals ($\geq$), among others;
       \item The data and application integrity logic is placed in the
@@
 
-\begin{table}
-    \label{tab:tab:schema-vs-schemaless-api-interface}
+Finally, each data model provides its own APIs to the integration with external
+systems. \cite{db-is-rdbs-dommed} differentiates each of the application interfaces
+provided by each of them, as summarized on table 5.3.
+
+\begin{table}[!h]
+    \label{tab:api-interface-schema-vs-schemaless}
     \caption{Schema-Dependent X Schema-less Databases: Application Interface}
@@
 
-The following summarizes which data model each of the candidate database
-systems use.
+The data model used by each of the selected database systems can be summarized
+as follows:
 
@@
   \item \textbf{TinyDB}: uses the Relational Data Mode with the use of SQL; 
-  \item \textbf{MongoDB}: uses Key-Value Pair Data Model with an Non-Structure
-  Query language in forms of programming language abstraction;
+  \item \textbf{MongoDB}: uses Key-Value Pair Data Model with an abstraction of 
+  programming language functions calls;
   \item \textbf{IBM DB2}: uses the Relational and Structured models with SQL
@@
 
-\subsection{Advantages of the Provenance-Aware Data}
+\subsection{Use of Provenance-Aware Data}
 
@@
 description of a given event, and is always useful to determine different
-aspects of the data. As mentioned in chapter 3, data provenance helps identify
+aspects of the data. As mentioned in Chapter 3, data provenance helps identify
 the nature of the data by the identification of the properties of a given
@@
 
-\subsection{Disadvantages of the Provenance-Aware Data}
+Although metadata gives more information about the data, it just
+represents additional data inserted together with the raw data collected
+from the sensor devices may contribute with a decreasing amount of disk
+storage. Therefore, the use of provenance-away data increases the required
+disk space used to save the collected data from sensor devices.
 
-Additional data being inserted together with the raw data collected from the
-sensor devices may contribute with a decreasing amount of disk storage. 
-Therefore, the use of provenance-away data increases the required disk space
-used to save the collected data from sensor devices.
-
-\subsection{Advantages of the Annotations}
+\subsection{Use of Annotations}
 
 Annotations are used to describe the collect data even more, since the
-additional ``notes'' of the data are being stored together with the data. For
-example, the use of tags to a given entity may help identifying and different
-observations regarding the data collected from sensor devices.
+additional ``observation'' of the data are being stored together with the data.
+For example, the use of tags to a given entity may help identifying and different
+observations regarding the data collected from sensor devices. As shown in
+Figure \ref{fig:persistence-example-kvp}, the addition of the key ``tag'' was
+used to describe the observations that happened when the event of ``oil spill''
+occurred. In this way, tags can be indexed for better search of the collected
+data.
 
-\subsection{Disadvantages of the Annotations}
-
-This type of data does not make up part of the collected data, nor provenance
-information.
+On the other hand, the use of annotations behaves the same as the use of
+Provenance data. It will require additional disk space to store the
+collected data from sensors.
 
@@
 Annotations. The Key-Value Pair data model has the advantage of not maintaining
-a schema design, and therefore, can add any type of data, and therefore,
-mongoDB is the lead in this taxonomy.
+a schema design, and therefore, additional keys of descriptions can be added at
+any time without requiring schema changes. On the other hand, the relational
+data model requires schema changes to support the addition a new table column.
+
+\begin{itemize}
+  \item \textbf{MySQL}: supports data provenance and annotations via column
+  types on the entities, requiring schema changes;
+  \item \textbf{TinyDB}: supports data provenance and annotations via column
+  types on the entities, requiring schema changes. Different projects have
+  suggested changes to the SQL language used to give better support to
+  spatial-temporal data \cite{sn-db-newop}.
+  \item \textbf{MongoDB}: supports both data provenance and annotations with
+  the definition of a new key on a given collection. No changes are required to
+  the addition of the keys;
+  \item \textbf{IBM DB2}: supports data provenance and annotations via column
+  types on the entities, requiring schema changes.
+\end{itemize}
 
@@
 
-\subsection{Advantages of the Centralized Query Processing}
-
-Centralized data management and query processing is simpler than the in-network
-query method. Data is verified by a straightforward database management system
-without the risk of data being unavailable, when the data is spread among the
-network nodes.
-
-\subsection{Disadvantages of the Centralized Query Processing}
-
-When the Centralized Query processing is used, problems may occur. Depending on
-the size of the datasets and the database technology, performance may be a
-concern due to the alleged Funneling Affect \cite{sn-storage04}, since the
+An advantage of centralized data management and query processing is simpler
+than the in-network query method. Data is verified by a straightforward
+database management system without the risk of data being unavailable, when
+the data is spread among the network nodes. On the other hand, when the
+Centralized Query processing is used, problems may occur. Depending on the
+size of the datasets and the database technology, performance may be a concern
+due to the alleged Funneling Affect \cite{sn-storage04}, since the
 point-of-traffic is concentrated in the centralized system.
@@
 system may be considered to be easier to manage, it may fail to adjust to
-changes of requirements to manage data. On the other hand, a distribute
+changes of requirements to manage data. On the other hand, a distributed
 database system can be used to implement a data-centric solution. Therefore, a
@@
 be used a data persistence technology for sensor networks.
-
-\subsection{Advantages of Distributed Systems}
 
@@
 related data creation and retrieval. It is also seen as an optional mechanism
-of reducing the funneling affect of a centralized system by directing queries
+of reducing the funneling effect of a centralized system by directing queries
 to a given data partition \cite{sn-storage04}. Data Replication
-\cite{data-replication} provides a ways to decentralize data access by
+\cite{data-replication} provides a way to decentralize data access by
 providing duplicated data that is intended to be used on reading operations.
 
-\subsection{Disadvantages of Distributed Systems}
-
-Requires more technical knowledge expertise to manage the different systems
-that composed a distributed database.
+Although a distributed system may provide better support to the problem being
+solved, it may be a difficult solution to propose. Since it is designed as set
+of cooperating subsystems, multiple points of failure exist in this type of
+system, making it difficult to be operated. Furthermore, it requires more
+technical knowledge expertise to manage the different systems that composed a
+distributed database management system.
 
@@
   \item \textbf{Export Feature}: the technology must support export
-  capabilities.
+  capabilities during its execution such as hot backps.
 \end{itemize}
@@
 use of drivers in different programming and scripting languages such as Java,
-Python, Perl, among others. Moreover, only TinyDB does not support hot backup.
+Python, Perl, among others.
 
@@
   from the community. Also APIs are available for the majority of languages;
-  \item \textbf{TinyDB}: Not open-source;
+  \item \textbf{TinyDB}: Not open-source nor does it support hot backup;
   \item \textbf{MongoDB}: An open-source database system with an increasing
@@
   languages and scripting languages;
-  \item \textbf{IBM DB2}: Not Open-Source, but offers support from its
+  \item \textbf{IBM DB2}: Not open-source, but offers support from its
   community.
@@
 the different types of taxonomies and requirements defined for a selection of a
-persistence technology for NetBEAMS. Table \ref{tab:technology-selection}
-summarizes these scores using the (+) and (-) notation. Since they all provide
-support to data archival for a small volume of data, these parameters are not
-included in the table.
+persistence technology for NetBEAMS. Table 5.4 summarizes these scores using
+the (+) and (-) notation. Since they all provide support to data archival for
+a small volume of data, these parameters are not included in the table.
 
-\begin{table}
+\begin{table}[!b] 
     \label{tab:technology-selection}
+    \caption{Technologies Selection Scoring}
     \begin{center}
-        \begin{tabular}[!t]{|c|c|c|c|c|c|c|}\hline 
+        \begin{tabular}{|c|c|c|c|c|c|c|}\hline 
         \textbf{Database} & \textbf{MySQL} & \textbf{TinyDB} & \textbf{MongoDB} & \textbf{IBM DB2}\\\hline
@@
         \end{tabular}
-        \caption{Technologies Selection Scoring}
     \end{center}
@@
 
-All the database systems give support to data provenance as a feature or
-data types. Nonetheless, when it comes to the Data Model used, the majority of
-the technologies use schema-dependent data model. The abstraction of
-programming languages is reported to give users, without skills in database
-systems, a better method of data access to the collected data from sensor
-networks. In this way, although all the discussed tools provide access to data
-through an additional API, mongoDB was the only one to use non-structured
-query, as a consequence of using a schema-less data model, which is the type
-used by SF-BEAMS staff members to describe the collected data. For this reason,
-mongoDB was considered as a good candidate to be explored as a data persistence
-solution for NetBEAMS.
+As used in different projects, schema-dependent models such as the relational
+data model can be used to design the collected data entities. As shown
+in Figure \ref{fig:persistence-example-relational}, the specification of a
+schema for the YSI Sonde Data is achieved. However, what if a new sensor device
+needs to be included into the model? A refactoring would be necessary to
+accommodate this new requirement. All database presented technologies, but
+mongoDB, use the relational model.
+
+Considering the nature of sensor devices as a list of properties, where the
+collected data values are labeled and represented by keys, the Key-Value Pair
+data model is proposed in Figure \ref{fig:persistence-example-relational-kvp}.
+Considering that this technique can support any different device with
+different properties, this would be a good candidate to be used as a
+persistence model to support any properties list of any sensor device.
+However, this approach results in different problems. Taking into account the
+support to data provenance, the use of time-series for the solution results in
+data repetition, as shown in the values of the column ``time''. Thus, it
+represents an insertion anomaly on the First Normal Form of the relational model
+\cite{relational-model}, where the collected data metadata must be repeated.
+Furthermore, \cite{db-kvp-in-relational02} describes that the use of this
+strategy in regular relational databases as one that presents poor retrieval
+performance, due to the concentration of data in a single database table.
+Therefore, this strategy is not well suited to provide persistence for data
+collected from sensor networks, described by the guidelines of Data Provenance.
+
+Although the use of schema-less database systems in sensor networks has not been
+reported by the literature review, the schema-less data model can be used as
+opposed to the schema-dependent because it does not require the definition of a
+static schema. As shown in Figure \ref{fig:persistence-example-kvp}, different
+instances of the YSI Sensor are represented not only by the properties of the
+sensor device, but also by data provenance and annotations. Data provenance is
+prepresented by the keys ``time, ``latitude'' and ``longitude'', while
+Annotations by the key ``tag''. Furthermore, since the the method of data
+access of the KVP databases is through the use of an abstraction of programming
+languages, it may also offer a better support to non-skilled users such as the
+RTC marine biologists. mongoDB is the only technology that provides such
+support.
+
+All the discussed tools provide access to data through an additional API, but
+mongoDB also provides it as a query method. Without no reason to further
+discussions, mongoDB was considered as a good candidate to be explored as a
+data persistence solution for NetBEAMS. Moreover, it also provides all the
+other requirements as a technology as such being an open-source application and
+that it can be used as a single server or a distributed system, requirement
+necessary to develop a Data-Centric.
 

### 2009-12-11 docs/marcello-dissertation/thesis/preamble.tex
Revision of Prof. Puder's audio feedback of chapter 5, as well as entire restructure of the document, such as using more page width and height, making it more visual. Table references not working automatically, as well as page numbers (will review later). Table references are entered manually in the text.
The structure of chapter 5 changed according to the fact that some
sections, subsections were very short, with 1 paragraph. So, I decided
to use two different options:

* If there are more to be discussed, a new advantages/disadvantages
subsection is created; or else, one section includes two

@@
 
-\hoffset=.5in 
-\oddsidemargin=0in
-\evensidemargin=0in
-\topmargin=0.1in 
-\headheight=0in
-\headsep=1in
-\footskip=.8in
-\textwidth=5.9in
-\textheight=7in
+\pdfpagewidth 8.5in
+\pdfpageheight 11in 
+\setlength\topmargin{0in}
+\setlength\headheight{0in}
+\setlength\headsep{0in}
+\setlength\textheight{7.7in}
+\setlength\textwidth{6.5in}
+\setlength\oddsidemargin{0in}
+\setlength\evensidemargin{0in}
+\setlength\headheight{77pt}
+\setlength\headsep{0.25in}
+
+%\hoffset=.5in 
+%\oddsidemargin=0in
+%\evensidemargin=0in
+%\topmargin=0.0in 
+%\headheight=0in
+%\headsep=1in
+%\footskip=.8in
+%\textwidth=6.5in
+%\textheight=7in
 
@@
     Copyright \copyright{} \thisyear by \myname
-
-    \bigskip
-
-    All rights reserved. No part of the material protected by this
-    copyright notice may be reproduced or utilized in any form or by any
-    means, electronic or mechanical, including photocopying, recording or
-    by any information storage and retrieval system, without the prior
-    permission of the author.
+%    \bigskip
+%    All rights reserved. No part of the material protected by this
+%    copyright notice may be reproduced or utilized in any form or by any
+%    means, electronic or mechanical, including photocopying, recording or
+%    by any information storage and retrieval system, without the prior
+%    permission of the author.
 }
@@
 Sensor Networks are becoming more and more important to different science
-communities bacause of what they produce: the raw data for different studies.
+communities because of what they produce: the raw data for different domains.
 In order to make use the collected data, researchers may have to dissect the
@@
 proposed taxonomies, this work designs and implements a data persistence
-component for NetBEAMS, our case study of modular sensor network infrastructure
+component for NetBEAMS, a modular-based sensor network infrastructure
 developed to improve the operation of the SF-BEAMS environmental sensor
-network. In this way, based on empirical analysis regarding proposed
-taxonomies for the case study, this work selected mongoDB, a schema-less
-document-oriented database instead of the traditional use of relational
-databases. As a result based on the experiments conducted, this work suggests
-that use of Key-Value Pair datasents in orde tor provide External of
-Data-Centric approach. Therefore, the use of a solution of the help decreasing
-the complexity related to the data persistence layer for sensor networks with
-external or data-centric storage mechanism, and whose purpose is data
-archival. Finally, this work proposes different future works such a
-data-centric approach using Databa se Shards to store te collected data and Map
-Reduce to execute parallel queries over the shards.} \vspace*{\fill}
+network. SF-BEAMS is deployed at the Tiburon Island, California, and managed by
+the Department of Biology from SFSU. In this way, based on empirical analysis
+regarding proposed taxonomies for the case study, this work selected mongoDB,
+a schema-less document-oriented database instead of the traditional use of
+relational databases. As a result based of the experiments conducted, this
+work proposes a novel use of the Key-Value Pair databases in order tor provide
+External of Data-Centric approaches. Similarly, this work also agrees to
+related literature that the use of programming languages are better abstraction
+than the use of traditional query languages used to extract data from Relational
+databases, and therefore, it should be the proposed approach to non-technical
+researchers such as the Biologists. Finally, this work proposes different
+future works such a data-centric approach using Databasese Shards to store the
+collected data and Map Reduce to execute parallel queries over distributed
+database nodes.} \vspace*{\fill}
 
@@
 Date
-\doublespacing
+\onehalfspacing
 \newpage

### 2009-12-12 docs/marcello-dissertation/thesis/chapter6-data-persistence-component-design-and-architecture.tex
Revision containing the English review from Ivan, and rework in most of the sections.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@697 559edde2-a4d4-44e0-96fc-4035e40d7396

--- 2009-12-07 docs/marcello-dissertation/thesis/chapter6-data-persistence-component-design-and-architecture.tex
@@
 
-In order to provide information to users, the RTC scientists use OPeNDAP
-\cite{opendap} server to distribute collected data over the Internet. As
-discussed in the previous chapter, this work proposes a data persistence
-application of a Key-Value database system in order to provide an alternative
-method for data persistence in sensor networks, taking into account the
-taxonomies discussed earlier in chapter \ref{chap:taxonomies}.
-
-Considering our case study being NetBEAMS, this chapter describes the design
-and architecture component designed specifically for the Data Sensor Platform,
-DSP, in order to provide a persistence layer for NetBEAMS. It is vital to note
-that the technology chosen in the last chapter can be adapted by any
-alternative that uses the data model and infrastructure options presented.
-
-The proposed DSP Data Persistence component for the DSP
-Platform was implemented by using the same approaches described in chapter
-\ref{chap:netbeams-overview}. The following sections describe the scenarios for
-persistence, analyze the requirements from RTC, and describe the design and
-architecture of the solution.
+The mongoDB database system was selected, as stated in chapter 5, to store 
+the collected data from the SF-BEAMS sensor network. NetBEAMS is based on a 
+component-based architecture, where each of the components can play different
+roles during their lifecycle. It uses the abstraction of Data Producers and 
+Data Consumers: while a component produces data to the system, the latter 
+consumes \cite{netbeams2009}. In this way, a DSP Data Component was designed
+and implemented to provide the integration between NetBEAM and the mongoDB 
+database. This integration is based on the Data Sensor Platform (DSP), the
+kernel software stack provided by NetBEAMS, as documented in
+\cite{netbeams-dsp-architecture}. Finally, details about the implementation of
+this design are located in the Appendix Section
+\ref{sec:dsp-data-persistence-implementation}.
 
@@
 
-As described in section \ref{chap:netbeams-overview}, the DSP Platform was
-designed to support an environmental sensor network based in San Francisco Bay,
-composed by different types of sensors. The process of extracting the
-measurements from sensors can be summarized following the UML Business Process
-Diagram \cite{uml} depicted in figure
-\ref{fig:dsp-persistence-business-process}. It is important to note that this
-diagram details remote communication between sensor devices and a centralized
-server, as part of the main scenario to solve the problem of data persistence.
+As described in \cite{netbeams-dsp-architecture}, the DSP Platform was designed
+to support SF-BEAMS. As shown in Figure \ref{fig:dsp-persistence-business-process}, 
+the event of data collection starts by the DSP Component deployed on each 
+DSP Gateway Node \cite{netbeams-dsp-architecture}, by the event ``Measurements
+X Collected''. When this DSP Client component finishes the data extraction
+process, a new one starts. First, each of the co-located DSP Client processes 
+the collected measurements and creates a DSP Message, embodying the collected 
+data from the associated sensor device. Each DSP Message is added into a local
+temporary outbound queue that waits to be transmitted. Then, an asynchronous 
+data transmission event occurs at a specified rate, packaging all DSP
+Messages into the body of a DSP Messages Container in order to be serialized
+into XML (shown in Listing \ref{stream:dsp-message-serialized-ysi}) and
+transferred to the DSP Server host, a centralized server that receives all the
+collected data embodied into each of the DSP Messages.
 
@@
   \centering
-  \includegraphics[scale=0.5]{../diagrams/DSP-DataPersistence-Business-Diagram}
+  \includegraphics[scale=0.6]{../diagrams/DSP-DataPersistence-Business-Diagram}
   \caption{UML Business Process Model diagram - Adding Persistence for NetBEAMS}
@@
 
-At the event of data collection, the DSP Client receives measurements collected
-from each of the sensor devices. Each of the co-located DSP Client process the
-collected measurements and transform them into the alleged DSP Message
-(section \ref{sec:dsp-message}), which embodies the actual data values. After
-proper data processing, transformation and categorization, the DSP Message is
-added into temporary outbound queue. Then, as an asynchronous data transmission
-event occurs with a specified rate, transferring all the queued messages to be
-processed by a DSP Server host. As image
-\ref{fig:dsp-persistence-business-process}, the server is centralized and able
-to concurrently process messages.
-
-This work proposes the addition of a database system that is responsible for
-storing the gathered measurement data from sensor nodes of the SF-BEAMS through
-the use of NetBEAMS. Also in figure \ref{fig:dsp-persistence-business-process},
-the processed measurement data from sensors 1 through N must be collected in
-the same style and categorized as necessary. In this way, end users such as
-Biologists or Programmers can access the collected measurements directly
-from the Database. Furthermore, the same collected data can be exported to
-different formats as necessary.
-
-As a matter of fact, the data measurements collected from the sensor devices
-are basic properties with keys and associated values as described in section 
-\ref{sec:sfbeams}, after data has been described. According to the descriptions
-of the DSP Platform and the DSP Messages structure, the data is represented in
-programming language formats such as Java \cite{java} or in a serialized
-version in XML \cite{xml}. Similarly, the collected data contains measurements
-that uniquely identifies them, together with other properties such as
-timestamps of data generation and collected and the IP address of the DSP
-Client that produced the data. Therefore, these are essential properties for
-the proposed solution of this work.
+The goal of the new DSP Data Persistence component is to proxy the payload 
+from the DSP Messages to the new suggested Measurements Database. Once the DSP
+Server receives all the messages in a container, it processes the messages and
+sends their payloads to the Measurements Database. This process requires a new
+Database Configuration to be provided to the DSP Server host, as well as the
+installation and setup of the new Measurements Database. In this way, different
+users of the collected data, such as Programmers and Biologists, can access the
+collected data directly from the Measurements Database and reuse them by
+Exporting specific observation parameters to different formats of their
+interest.
 
@@
 
-By analyzing the main scenario of collected data in the previous section, as
-well as by collected requirements from the RTC staff, this section lists the
-main use cases that the system and the primary users require.
+This section outlines the functionalities and constraints that the developed
+solution must perform and follow, respectively, obtained by analyzing the main
+data collection scenario described in the previous section and reviewing the
+requirements defined by the RTC staff members.
 
@@
 
-As part of the user stories related to data access, figure
-\ref{fig:DSP-Data-Persistence-UseCases-Diagram-Users} describes the use cases
-that the end users can perform with the system.
+The primary goal of the system is to allow users to have easy access to the 
+collected data from \emph{NetBEAMS} using a schema-less database technology.
+mongoDB provides access to the stored data without requiring the use of 
+a Structured Query Language. Instead, that technology provides a programming
+language abstraction that makes the access to the stored data easier 
+\cite{sn-programming-language}. Using Agile User Stories to describe Use 
+Cases Diagrams \cite{use-cases-user-stories}, Figure 
+\ref{fig:DSP-Data-Persistence-UseCases-Diagram-Users} enumerates these
+functionalities as customers conversations, starting with a fictitious
+persona, the action it desires to perform and the outcome of the action, a
+requirements gathering pattern defined in Scrum \cite{agile-scrum}:
 
@@
   \centering
-  \includegraphics[scale=0.5]{../diagrams/DSP-Data-Persistence-UseCases-Diagram-Users}
+  \includegraphics[scale=0.65]{../diagrams/DSP-Data-Persistence-UseCases-Diagram-Users}
   \caption{UML Use Case diagram for Persistence Functions}
@@
 \begin{itemize}
-  \item \textbf{Search Measurements by Sensor Properties}: users of the
-  collected data must be able to search data through the properties of the
-  sensors. This is related to identity principle of what was collected
-  as provenance \cite{db-provenance};
-  \item \textbf{Search Measurements by Sensor Location}: users of the collected
-  data must be able to search data by a given sensor location, as referred to
-  the provenance from where the data was collected;
-  \item \textbf{Search Measurements by Data collection date}: users of the
-  collected data must be able to search data by a given date, or ranges of
-  start and end date;
-  \item \textbf{Access the data through API, Programming Languages}: users of the
-  collected data must be able to search data by the use of different
-  programming and scripting languages of their preference;
-  \item \textbf{Export Data to OPEnDAP format}: users of the collected data
-  must be able to export the collected data to the OPEnDAP format;
-  \item \textbf{Export Data to Spreadsheets}: users of the collected data must
-  be able to export the collected data to spreadsheets.
+  \item \textbf{Search Measurements by Sensor Properties}: "As a marine
+  biologist, I would like to search observations by filtering values of the 
+  sensor device's properties such as temperature or salinity, so that I can 
+  find specific associated values to the observation.";
+  \item \textbf{Search Measurements by Sensor Location}: "As an oceanographer,
+  I would like to search observations that took place at the geographic
+  coordinates (x latitude, y longitude), so that I can assess the area around
+  the given coordinates.";
+  \item \textbf{Search Measurements by Data collection date}: "As a marine 
+  biologist, I would like to search observations that took place last week, 
+  so that I can assess past environmental conditions";
+  \item \textbf{Annotate Existing Measurements}: "As a estuarine ecologist,
+  I would like to annotate observations from the time the oil spill occurred
+  in the San Francisco Bay, so that I can maintain historical evidence of 
+  the impact of such event.";
+  \item \textbf{Export Data to Spreadsheets}: "As a scientist from RTC, I 
+  would like to analyze of the observed data from yesterday using a 
+  spreadsheet, so that I can verify measurements using Microsoft Excel.";
+  \item \textbf{Export Data to OPeNDAP format}: "As a biologist, I would 
+  like to export the collected data produced during this month using the
+  OPeNDAP data format, so that I can collaborate with other research groups
+  that uses this data format.";
+  \item \textbf{Access the data through API, Programming Languages}: "As a
+  marine biologist who learned the Python scripting language, I would like to
+  write a software that connects directly to the database, so that I can
+  implement a customized analysis system for NetBEAMS.";
+  \item \textbf{Remove Undesired Measurements}: "As an oceanographer, I
+  would like to remove specific observations collected yesterday, 
+  so that the research group does not use 'junk data'.".
+\end{itemize}
+
+As one can see, the basic users' functionalities are related to the access to
+the collected data using different parameters related to the each observation's
+properties or implied metadata such as time of collection or location. In order
+to do so, the functions to be performed were developed following the
+specifications of the Data Sensor Platform (DSP) architecture, as described in
+\cite{netbeams-dsp-architecture}. In order to collect data sent to the NetBEAMS
+server, the new DSP component must act as a Data Consumer. The functionalities
+provided by this component can be seen in the UML Use Cases diagram \cite{uml}
+in Figure \ref{fig:DSP-Data-Persistence-UseCases-Diagram-System}, and
+summarized as follows:
+
+\begin{figure}[!h]
+  \centering
+  \includegraphics[scale=0.65]{../diagrams/DSP-Data-Persistence-UseCases-Diagram-System}
+  \caption{UML Use Case diagram for the Persistence Component}
+  \label{fig:DSP-Data-Persistence-UseCases-Diagram-System}
+\end{figure}
+
+\begin{itemize}
+  \item \textbf{Update Component Properties}: the component must be
+  configurable, with its own set of properties, such as the flush rate used 
+  to send data to the database system;
+  \item \textbf{Update Database Properties}: the component must be able to
+  update the mongoDB database settings such as IP Address and  communication 
+  port;
+  \item \textbf{Collect DSP Measurement Message}: the component must be
+  able to receive and temporarily retain the collected messages from
+  different the sensors devices;
+  \item \textbf{Categorize Measurement Message}: the component must be able
+  to categorize the DSP Message, as it arrives, by its properties such as 
+  transaction time, and originating DSP Message ID;
+  \item \textbf{Save Messages into the Database}: the component must be able
+  to send the payload of the DSP Messages to the database.
 \end{itemize}
@@
 
-Different constraints are required to execute the functional requirements
-described in the previous section. They are listed as non-functional
-requirements as follows:
+Different constraints were required to execute the functional requirements
+described in the previous section. Those constraints are listed as different
+non-functional requirements as follows:
 
 \begin{itemize}
-  \item The DSP Data Persistence component must maintain the received DSP
-  Measurement Messages in-memory for a specific period of time in order to
-  decrease the write load on the database;
-  \item The Persistence storage system must be able to cope with dozens of DSP
-  Messages received from different sensors;
-  \item The Persistence storage system must be able to save measurements quickly;
-  \item The Persistence storage system must be able to scale without too much
-  architectural changes, or schema changes;
-  \item The Persistence storage system must not impose the use of
-  database languages such as SQL \cite{sql} or XML XPath \cite{xml-xpath}
-  related to the data being persisted. Instead, data must be handled through API
-  calls, as better abstractions for non-computer science researchers
-  \cite{sn-programming-language}.
+  \item The solution must maintain the received DSP Measurement Messages
+  in-memory for a specific period of time in order to decrease the write load
+  on the database (flush rate);
+  \item The persistence storage system must be able to cope with the Data
+  Volume produced by SF-BEAMS on daily basis;
+  \item The persistence storage system must be able to scale without minimal
+  architectural and schema changes;
+  \item The persistence storage system must be able to provide access to the
+  persisted data using an programmatic abstraction interface, as the authors
+  of \cite{sn-programming-language} proposed.
 \end{itemize}
+
+Driven by the requirements defined, the foundation of the integration between
+the DSP Data Persistence component and the mongoDB database system is the
+specification of the Keys that comprise a set of observations collected from
+the sensor devices.
 
@@
 
-As described in the previous chapter, the data model chosen to support
-persistence of collected data from sensor networks is the Key-Value-Pair Data
-Model \cite{db-kvp}. Beyond the data collected from the sensor devices, Data
-Provenance techniques must also be considered to better describe the collected
-data \cite{sn-provenance}. In general, the data model must include the
-following metadata besed on Data Provenance:
+MongoDB is defined as a Document-Oriented Database system, which is based on 
+the abstraction of the Key-Value-Pair data model, using a binary version of JSON
+\cite{json} as its main data representation. MongoDB does not require the creation 
+of a database schema, rather the definition of key descriptions that comprises
+a document on the database organized into separate collections of relating
+documents. In order to describe the keys of a document, the specifications
+detail the use of Data Provenance as guidelines to the description of the
+collected data. Therefore, the key-value pairs of any instance of the
+collected data must include the following key properties:
 
 \begin{itemize}
-  \item \textbf{Locality}: properties that describe the location of the data.
-  For example, the host machine name or IP address, the GPS coordinate
-  \cite{gps};
-  \item \textbf{Identity}: what uniquely identities the data among the others
-  produced by other sensor devices in the network, as for instance a
-  UUID\footnote{Universally Unique Identifier};
-  \item \textbf{Time-Dimensions}: when the data was ``sensed''
-  \cite{db-provenance} and when the collection transaction happened
-  \cite{sn-time-series}.
+  \item \textbf{Location}: properties that describe the location of the data.
+  For example, the host machine name, IP address, or the GPS coordinates
+  \cite{gps} from the sensor device;
+  \item \textbf{Identity}: what uniquely identifies the data among 
+  others, produced by other sensor devices in the network such as the DSP
+  Message ID using UUID\footnote{Universally Unique Identifier};
+  \item \textbf{Time-Dimensions}: when the data was recorded 
+  \cite{db-provenance} and when the transaction occurred \cite{sn-time-series}.
 \end{itemize}
@@
 After analyzing the DSP Platform and the structure of the DSP Messages in
-chapter \ref{netbeams-architecture}, the following set of keys are expected to
-be part of the design of the data model. Any key represented by time or
-location are related to the provenance support discussed in section
-\ref{sec:sn-data-description}:
+\cite{netbeams-dsp-architecture}, the values of the collected data must be
+categorized into two different sets of data: the ``Key Segment'' and the
+``Value Segment''. The key segment provides the necessary metadata 
+used to describe the collected data. An instance of the design of the 
+keys is shown in Listing \ref{file:mongodb-ysi-data-format}:
 
 \begin{itemize}
-  \item \textbf{sensor.ip\underline{ }address}: it is extracted
-  from the DSP Message Produce and identifies which sensor generated the sampling;
-  \item \textbf{sensor.location.latitude, longitude}: the latitude and
-  longitude GPS coordinates of the position of the sensor device, captured from
-  references of the device. This is currently provided as a parameter in the
-  bootstrap message for the component;
-  \item \textbf{message\underline{ }id}: it's extracted from each of the
-  messages contained in the DSP Message Container;
-  \item \textbf{time.transaction}: it is extracted from the DSP
-  message container creation time and it is used to identify when the
-  transaction occurred;
-  \item \textbf{time.fact}: it is extracted from the
-  SondeDataContainer's date and time, and identifies the time in which the
-  collected data occurred.
+  \item \textbf{message\underline{ }id}: Defines the DSP Message ID that
+  carried the observation from the sensor device to the DSP Server. This ID
+  can be used during network and database inspections in order to track messages
+  for the DSP Platform;
+  \item \textbf{sensor}: Defines the properties of the sensor device;
+    \begin{itemize}[label=\textbullet]
+        \item \textbf{ip\underline{ }address} Defines the IP Address of the 
+        sensor device that produced the data;
+        \item \textbf{location}: Defines the geographical coordinates of the
+        sensor device, using the Decimals Degree format \cite{decimal-degrees};
+            \begin{itemize}[label=\textbullet]
+                \item \textbf{latitude} the GPS latitude coordinate of the
+                position of the sensor device;
+                \item \textbf{longitude}: the GPS longitude coordinate of the
+                position of the sensor device.
+            \end{itemize}
+    \end{itemize}
+  \item \textbf{time}: Defines different time dimensions according to Data 
+   Provenance;
+    \begin{itemize}[label=\textbullet]
+      \item \textbf{valid}: it is extracted from the container's date and time
+      of creation, which was collected during the sensor device's reading;
+      \item \textbf{transaction}: it is extracted from the DSP message
+         container's creation time and it is used to identify when the
+         transaction, which transferred the message from the DSP Gateway Client
+         to the DSP Server, occurred;
+    \end{itemize} 
 \end{itemize}
 
-It is important to note that this set of metadata must occur in every
-collected data from any sensor device. However, on the contrary of the
-metadata, a different set of properties will occur to describe the sensed
-properties of a given sensor, and for this reason, the following structure is
-defined to accommodate such data:
+An associated ``Value Segment'' for the aforementioned ``Key Segment'' carries
+the observations of a given sensor device. The key \textbf{observation} stores
+the values of the collected data. For instance, the observation key of
+for the YSI device \cite{YSI-Sonde} may contains all the subsets of 
+the properties defined for the device, such as battery, temperature and
+salinity, among others. An example of such use of the segments is shown in
+Listing \ref{file:mongodb-ysi-data-format}.
 
-\begin{itemize}
-  \item \textbf{observation}: this key defines the values of the document, and
-  will have every different property of the sensor. For instance, the data key
-  for the properties of the YSI device will contain the set of properties 
-  defined for the device, such as battery, temperature, salinity, etc.
-\end{itemize}
+Based on this foundation, the design of the DSP Data Persistence component
+evolved from a top-down approach.
 
@@
 
-Considering the business analysis described in section
+Considering the business analysis described in Section
 \ref{sec:business-process-analysis}, the addition of a data persistence layer
-to the DSP Platform can be achieved by adding Data Consumer (DC) as described
-in section \ref{netbeams-architecture}. Based on the Requirements
-specification described earlier, this component must be responsible for
-providing the processing of the DSP Measure Messages, extracting the payload of
-the body of the message and converting it to the data model defined in the
-previous section. Finally, the data can be transmitted to the chosen database
-system. Figure \ref{fig:NetBEAMS-Persistence-Server-Node-Components} provides
-an architectural view of the DSP Server node with the loosely-coupled
-components, highlighting the inclusion of the DSP Data Persistence.
+to the DSP Platform can be achieved by adding a Data Consumer (DC) as 
+described by \cite{netbeams-dsp-architecture}. Based on the 
+Requirements specification, this component must be responsible for the 
+extraction process that takes the payload of the body of a filtered DSP
+Measurement Message and converts it to the key-value data model defined
+in the previous section. This process guarantees that the collected data is
+inserted to a running mongoDB database system and gives covers the use cases
+defined. Figure \ref{fig:NetBEAMS-Persistence-Server-Node-Components}
+provides an architectural overview of the DSP Server node with the
+loosely-coupled components, highlighting the inclusion of a new DSP Component
+called ``DSP Data Persistence''.
 
-\begin{figure}[!b]
+\begin{figure}[!h]
   \centering
-  \includegraphics[scale=0.5]{../diagrams/NetBEAMS-Persistence-Server-Node-Components}
+  \includegraphics[scale=0.65]{../diagrams/NetBEAMS-Persistence-Server-Node-Components}
   \caption{UML Components diagram of the DSP Data Collector}
@@
 
-As it is depicted in figure
+As it is depicted the diagram 
 \ref{fig:NetBEAMS-Persistence-Server-Node-Components}, the DSP Data
-Persistence component can be added into the system as an independent
-plug-and-play component. In addition, the database system and the necessary
-drivers are also added without any changes to the existing DSP infrastructure.
-First, as DSP Messages are delivered to the server through the DSP Wire
-Transport Server, as shown by flow 1, the DSP Platform is responsible to
-forward a copy of the Measurement messages directly to the DSP Data
-Persistence as shown by flow 2. Then, the sole responsibility of the DSP Data
-Persistence is to transfer the received measurement messages to the database
-system by using the connection drivers, as depicted by flow 3.
+Persistence can be added into the existing architecture as an 
+independent plug-and-play component. The database system and the necessary
+drivers are also added without any changes to the existing DSP 
+infrastructure. First, the DSP Messages are delivered to the DSP server 
+through the DSP Wire Transport Server, as shown by flow 1. Next, the DSP 
+Platform is responsible to forward a copy of the Measurement messages 
+directly to the DSP Data Persistence as shown by flow 2. Then, the sole 
+responsibility of the DSP Data Persistence is to transfer the received 
+measurement messages to the database system by using the connection 
+drivers, as shown by flow 3.
 
-The following sections focus on the DSP Data Persistence component design in
-the context of an OSGi bundle \cite{osgi}.
+Any DSP Component is essentially an OSGi Bundle \cite{osgi-intro}, a unit of
+deployment in a Java JAR format \cite{java-tutorial} containing the Java
+classes and a descriptor manifest artifact that identifies the major components
+of the package, which services the component requires to use and which ones it
+will optionally publish during its activation and lifecycle. Details about the
+deployment of the DSP Data Persistence component, as well as about the OSGi
+environment used, are given in the Appendix section
+\ref{sec:dsp-component-osgi-deployment}.
+
+The following sections focus on the DSP Data Persistence component 
+design in the context of an OSGi bundle \cite{osgi}.
 
@@
 
-The new proposed Data Persistence component must follow the DSP requirements
-for a new DSP Component, which specifies the design-pattern of Data
-Producer/Consumer. In a nutshell, the design of a DSP Component maintains the
-dependency among 3 different major Java packages: it depends on 2 existing
-packages, namely the OSGi Framework and the DSP Platform, as well as and an
-additional package for the database access driver API, as shown in the UML
-Package Dependency Diagram \cite{uml} of Figure
-\ref{fig:DSP-Data-Persistence-Packages-Dependency}.
+The proposed Data Persistence component must follow the DSP Component
+requirements, which specifies the design-pattern of Data Producer and Consumer
+mentioned earlier in this chapter. As seen in the UML Package  Dependency
+Diagram in Figure \ref{fig:DSP-Data-Persistence-Packages-Dependency}, a new
+DSP Component has dependency on three different packages: the OSGi Framework
+and the DSP Platform responsible to integrate a new component into the
+micro-kernel, and the mongoDB Java driver API that connects to the mongoDB
+server.
 
@@
   \centering
-  \includegraphics[scale=0.5]{../diagrams/DSP-Data-Persistence-Packages-Dependency}
+  \includegraphics[scale=0.65]{../diagrams/DSP-Data-Persistence-Packages-Dependency}
   \caption{UML Components diagram of the DSP Data Collector}
@@
 
-The following sections details each step of the DSP Data Persistence activation
-and parallel execution of the contained classes.
+In order to use a DSP Component, one must first activate it into the OSGi
+Framework the DSP Platform runs on. Both the activation and the design of how
+the new component acts as a Data Consumer is explained in the following
+sections.
 
@@
 
-The DSP Data Persistence component starts with an OSGi Activator class, and the
-implementation of the DSP Component, resposible to provide services to the
-platform. Based on these specifications, the following set of classes in
-figure \ref{fig:DSP-DataPersistence-Activator-Class-Diagram} shows the UML
-Class Diagram \cite{uml} implemented for the component.
+The DSP Data Persistence component starts with an OSGi Activator class, and 
+the implementation of the DSP Component, responsible for providing services 
+to the platform \cite{netbeams-dsp-architecture}. Based on these 
+specifications, the following set of classes, shown in Figure 
+\ref{fig:DSP-DataPersistence-Activator-Class-Diagram}, depicts the implemented
+classes for the new DSP Data Persistence component.
 
@@
   \centering
-  \includegraphics[scale=0.5]{../diagrams/DSP-DataPersistence-Activator-Class-Diagram}
+  \includegraphics[scale=0.65]{../diagrams/DSP-DataPersistence-Activator-Class-Diagram}
   \caption{UML Class diagram for the DSP Data Persistence Activator and Component}
@@
 
-The DSP Data Persistence Activator is the main class responsible for the
-activation of the DSP Data Persistence component, as it extends the OSGi class
-BundleActivator and maintains a reference of the DSP Data Persistence class.
-Furthermore, the activator maintains references to the classes BundleContext
-and ServiceReference both from the OSGi Framework. The former is responsible
-for maintaining the link with the OSGi Framework, while the latter is used to
-register the DSP Component as a service. Therefore, it is clear that the class
-DSP Data Persistence Activator is the main communication interface between the
-DSP Platform and the OSGi Framework.
+The class DSPDataPersistenceActivator is the main class responsible for the
+activation of the class DSPDataPersistence, as it extends from the OSGi
+class BundleActivator. The BundleActivator maintains references to the classes
+BundleContext and ServiceReference, both from the OSGi Framework, in order to
+maintain the component managed by the OSGi Framework and register the class
+DSPDataPersistece as a service to the OSGi environment. Therefore, the class
+DSPDataPersistenceActivator is the main communication interface between the 
+DSP Platform and the OSGi Framework, are responsible for the orchestration of
+the application lifecycle.
 
-While the DSP Data Persistence Activator is responsible for the orchestration
-of the classes during the execution on the OSGi Framework, the class DSP Data
-Persistence extends from the DSP Platform class DSPComponent. For this reason,
-this class inherits all the behavior described on \ref{netbeams-architecture}
-and can be in the roles of Data Consumer and a Data Producer.
+The class DSPDataPersistence implements the interface DSPComponent, responsible
+to provide the contract between the component and the DSP platform. As a
+consequence, this class inherits all the behavior described in
+\cite{netbeams-dsp-architecture} necessary to be executed in the DSP Platform
+in a given context (through the instance of class DSPContext). The different
+roles played by the class DSPDataPersistence are defined as follows:
 
 \begin{itemize}
-  \item \textbf{DSP Data Persistence as Data Consumer}: receives any measurement
-  message from remote sensors;
-  \item \textbf{DSP Data Persistence as Data Producer}: transforms any
-  measurement message received into a format that must be ready to be saved on the
-  database, based on the description of the data model in section
-  \ref{sec:dsp-persistence-data-model}.
+  \item \textbf{Data Consumer}: receives any measurement message from remote
+  sensors;
+  \item \textbf{Data Producer}: transforms any received measurement message
+  into the format defined in Section \ref{sec:dsp-persistence-data-model}, in
+  order to be sent to the database.
 \end{itemize}
 
-The behavior of the activator class follows the specifications of the
-OSGi Framework and, for this reason, instances of the DSP Data Persistence
-Activator can be in the states described on the UML State Diagram depicted in
-Figure \ref{fig:DSPPlatform-Install-Usage-State-Diagram}. It is important to
-note that the activator is an extension of the OSGI class BundleActivator, as
-well as being responsible for stopping or starting the the DSP Data Persistence
-component instance. In this way, following the specifications of section
-\ref{netbeams-architecture}, the DSP Data Persistence Platform can be
-initialized as described as it is described by the UML Sequence Diagram of
+The class DSPDataPersistenceActivator is responsible for starting and stopping,
+the services provided by the DSPDataPersistence component, as this behavior is
+inherited from the OSGi class BundleActivator from OSGi Framework. Following
+the specification of the DSP Architecture \cite{netbeams-dsp-architecture},
 Figure
-\ref{fig:From-OSGi-Framework-to-DSP-Data-PersistenceActivator-Sequence-Diagram}:
+\ref{fig:From-OSGi-Framework-to-DSP-Data-PersistenceActivator-Sequence-Diagram} 
+shows a UML Sequence Diagram that depicts the steps of the activation of the
+DSP component, and summarized as follows:
+
+\begin{figure}[!h]
+  \centering
+  \includegraphics[scale=0.65]{../diagrams/From-OSGi-Framework-to-DSP-Data-PersistenceActivator-Sequence-Diagram}
+  \caption{UML Sequence diagram describing the DSP Data Persistence bundle activation}
+  \label{fig:From-OSGi-Framework-to-DSP-Data-PersistenceActivator-Sequence-Diagram}
+\end{figure}
 
 \begin{enumerate}
-  \item During the DSP Platform activation, it will get the name of the
-  persistence component from the configuration artifact config.xml as seen in
-  listing \ref{file:dsp-config.xml};
-  \item After being selected based on the configuration priority, the DSP
-  Bundle artifact is identified and installed, by creating an instance of the
-  class DSP DataPersistence Activator and making a call to the method start();
-  \item During the activation, an instance of the DSP Data Persistence
-  component is created and registered as an OSGi Service;
+  \item During its activation, the DSP Platform will get the name of the
+  DSP Data Persistence component from the configuration artifact config.xml as
+  seen in Listing \ref{file:dsp-config.xml};
+  \item After being selected based on the configuration priority, the 
+  DSPBundle artifact is identified and installed by creating an instance of the
+  class DSPDataPersistenceActivator, making a call to the method start();
+  \item During the activation, an instance of the class DSPDataPersistence
+  is created and registered as an OSGi Service;
   \item Upon registering the DSP Data Persistence component, the DSP Platform
-  "listens" the event serviceUpdate() and, and triggers the last operations to
+  listens for the event serviceUpdate() and triggers the last operations to
   bootstrap the DSP Component:
    \begin{enumerate}
-      \item Initializes the component by calling the method initComponent();
-      \item Starts the component by calling the method startComponent();
-      \item Bootstrap messages are delivered to the component.
+      \item Initialize the component by calling the method initComponent();
+      \item Start the component by calling the method startComponent();
+      \item Bootstrap messages by calling the method delivery(). This is where
+      both the DSP Measurements Messages and Update Messages are delivered by
+      the DSP Broker. This concept is based on the listener of Observer
+      Design-Pattern \cite{gof}.
    \end{enumerate}
@@
 the DSP Component Activator and the DSP Component registration are also shown
-in figure \ref{fig:From-OSGi-Framework-to-DSP-Data-PersistenceActivator-Sequence-Diagram}.
+in Figure \ref{fig:From-OSGi-Framework-to-DSP-Data-PersistenceActivator-Sequence-Diagram}.
 
-\begin{figure}[!b]
-  \centering
-  \includegraphics[scale=0.5]{../diagrams/From-OSGi-Framework-to-DSP-Data-PersistenceActivator-Sequence-Diagram}
-  \caption{UML Sequence diagram describing the DSP Data Persistence bundle activation}
-  \label{fig:From-OSGi-Framework-to-DSP-Data-PersistenceActivator-Sequence-Diagram}
-\end{figure}
+As soon as the process "Start DSP Activator Gate" is executed, an instance of
+the class DSPDataPersistence component is created to be registered into the
+OSGi Framework using the DSP Platform's helper class ActivationHelper. It is
+responsible to proxy the method call to the class BundleContext, which makes
+the actual registration of the DSPDataPersistence component as an OSGi
+Service. At this point, the DSPDataPersistence is installed and on the state
+"Active", as defined by the OSGi Framework (Refer to
+\cite{netbeams-dsp-architecture} for details).
 
-As soon as the gate "Start DSP Activator Gate" reaches the DSP Data
-Persistence instance, it first creates the unique instance of the class DSP
-Data Persisence and uses that instance to register it as the OSGi service
-through the call to the DSP Platform helper class ActivatorHelper. At this
-point, the DSP Data Persistence has been installed and is in the state
-"Active".
+Once the DSP Data Persistence component is concurrently running with other
+components from both the DSP Platform and the OSGi Framework, it is able to
+receive DSP Messages. The following section will explain how to bootstrap the
+DSP Data Component in order to receive Measurement Messages.
 
@@
 
-Although the DSP Data Persistence component has been instantiated as of image
+Although the DSP Data Persistence component has been instantiated as of Figure
 \ref{fig:From-OSGi-Framework-to-DSP-Data-PersistenceActivator-Sequence-Diagram}, 
-the process of delivering messages to the DSP Data Persistence has yet to be
-completed. As defined by the requirements, the DSP Component must execute its
-function in a concurrent fashion, using configuration parameters during its
-bootstrap process. The class DSP Data Flusher is responsible for running as a
-"worker thread", whose function is to verify if there are messages to be
-flushed into the Database with the class Transient Persistence Layer. Figure 
-\ref{fig:DSP-DataPersistence-Flusher-Classes} describes the UML Class
-Diagram \cite{uml} for the main classes of this functionality.
+the process of delivering messages to the DSP Data Persistence component has yet
+to be completed. As defined by the requirements, the DSP Component must execute its
+function concurrently, using the configuration parameters during its
+bootstrap process. The DSP Platform delivers any bootstrap message defined in
+the DSP Platform runtime environment, as explained in the Appendix section
+\ref{sec:dsp-persistence-boostrap}. However, to deliver DSP Measurement
+Messages, a new filter must be provided. As explained in
+\cite{netbeams-dsp-architecture}, the DSP Broker uses the DSP Matcher's rules
+to select to Data Consumers configured in a given DSP. For this reason, a rule
+that filters any remote measurement message received by running DSP Server is
+used, as shown in Listing \ref{file:dsp-matcher-config.xml}. The rule defined
+in line 63 express that all messages created by any producer on any DSP Host,
+either client or sever, has as delivery target the DSP Data Persistence
+component. In this way, when DSP Measurement Messages are delivered, the DSP
+Broker delivers the DSP Message to the DSP Data Persistence Component.
 
-\begin{figure}[!b]
+The UML Class Diagram for the participating classes of the DSP Data
+Persistence component is shown in Figure
+\ref{fig:DSP-DataPersistence-Flusher-Classes}. The class DSPDataFlusher is
+responsible for running as a worker thread, whose function is to verify if
+there are messages to be flushed to the Database by contacting the Singleton
+class \cite{gof} TransientPersistenceLayer, responsible to keep track of the
+DSP Messages received when the DSP Broker delivered them to the DSP Data
+Persistence component.
+
+\begin{figure}[!h]
   \centering
-  \includegraphics[scale=0.5]{../diagrams/DSP-DataPersistence-Flusher-Classes}
-  \caption{UML Class diagram with the DSP Data Flusher worker thread}
+  \includegraphics[scale=0.65]{../diagrams/DSP-DataPersistence-Flusher-Classes}
+  \caption{UML Class diagram with DSP Data Persistence and Participating classes}
   \label{fig:DSP-DataPersistence-Flusher-Classes}
@@
 
-\begin{figure}[!b]
-  \centering
-  \includegraphics[scale=0.5]{../diagrams/DSP-Data-Persistence-Mongo-Classes}
-  \caption{UML Class diagram with Persistence Service for MongoDB}
-  \label{fig:DSP-Data-Persistence-Mongo-Classes}
-\end{figure}
+Upon receiving a new DSP measurement message, the DSP Data Persistence
+component keeps any received message in a temporary local cache, as described 
+by the non-functional requirements. In this way, the DSP Data Persistence 
+component maintains a dependency on the Singleton class TransientPersistenceLayer, 
+responsible for encapsulating the DSP Message into an instance of the class 
+PersistentMessageUnit, to be stored in the defined local cache.
 
-The DSP Data Flusher is a thread that is designed to be in two different
-states as depicted in figure
-\ref{fig:DSP-DataPersistence-Flusher-State-Diagram}: RUNNABLE and
-TIMED\underline{ }WAITING. In the former state, the DSP Data Flusher will be
-contacting the Transient Persistence Layer in order to verify if there are any
-message in the outbound queue ready to be transmitted sent to the Database
-service, while the DSP Data flusher will be waiting for the next cycle defined
-by the bootstrap parameter TRANSIENT\underline{ }DATA\underline{
-}FLUSHER\underline{ }DELAY in the latter state.
-
-\begin{figure}[!b]
-  \centering
-  \includegraphics[scale=0.5]{../diagrams/DSP-DataPersistence-Flusher-State-Diagram}
-  \caption{UML State diagram for the DSP Data Flusher Worker Thread}
-  \label{fig:DSP-DataPersistence-Flusher-State-Diagram}
-\end{figure}
-
-Upon receiving a new DSP measurement message, the DSP Data Persistence
-component must send its reference to the temporary memory space, as described
-by the non-functional requirement in the beginning of this chapter. In this
-way, the DSP Data Persistence component maintains a dependency on the Singleton
-class TransientPersistenceLayer, which is responsible for keeping track of
-messages that are received by working component. However, in order to add the
-DSP message into its aggregated map, it first transforms the message into an
-instance of the class PersistentMessageUnit, which has references to the class
-regarding the location of the message, SensorLocation.
-
-As depicted by the the UML Sequence Diagram \cite{uml} of figure
-\ref{fig:From-DSPWireTransport-Server-To-DSP-Broker}, a DSP Message
-is delivery to the DSP Broker for its delivery to Data Consumers configured to
-receive that specific message. For this reason, Figure
-\ref{fig:From-DSP-Broker-To-DSPDataPersistence-General-Sequence} continues the
-steps of the DSP Broker from the gate "From DSP WireTransport Server to Broker
-Gate", showing the steps taken to transfer a DSP Message to the Singleton
-class TransientPersistenceLayer, responsible for maintaining a queue of
-messages waiting to be sent to the database system.
-
-According to the DSP Broker model described in section \ref{sec:dsp-message}, 
+According to the DSP Broker model described in \cite{netbeams-dsp-architecture}, 
 the DSP Matcher needs to be updated with the addition of a matching rule that
-filters a copy of any Measurement Message to the DSP Data Persistence component
-designed. As it is shown in figure
+filters a copy of any Measurement Message to the DSP Data Persistence component. 
+As it is shown in Figure
 \ref{fig:From-DSP-Broker-To-DSPDataPersistence-General-Sequence}, the worker
@@
 layers, as it depends on the references from the classes
-TransientPersistenceLayer and the DSPMongoCRUDService. The former maintains
-DSP Messages in memory wrapped by an instance of PersistenceMessageUnit and
-the latter is responsible for sending the MessageContent from the DSPMessage
-body to the Database.
+TransientPersistenceLayer and the DSPMongoCRUDService.The class
+DSPMongoCRUDService is responsible for extracting the collected data from the
+body of the DSP Message and creating the keys designed to be persisted
+in the mongoDB Database.
 
-\begin{figure}[!b]
+The DSP Broker retrieves the list of matching rules by contacting the Matcher.
+As the DSP Platform has already loaded all the Matching rules, a copy of an
+analyzed DSP Message is delivered to the DSP Data Persistence component specified
+by a matcher rule. However, the UML Sequence Diagram depicted in Figure 
+\ref{fig:From-DSP-Broker-To-DSPDataPersistence-General-Sequence} shows two 
+different moments of message delivery:
+
+\begin{figure}[!h]
   \centering
-  \includegraphics[scale=0.5]{../diagrams/From-DSP-Broker-To-DSPDataPersistence-General-Sequence}
-  \caption{UML Sequence Diagram - Flow 1: From the DSP Broker to the Data
-  Persistence Component}
+  \includegraphics[scale=0.6]{../diagrams/From-DSP-Broker-To-DSPDataPersistence-General-Sequence} 
+  \caption{UML Sequence Diagram - Flow 1: From the DSP Broker to the Data Persistence Component}
   \label{fig:From-DSP-Broker-To-DSPDataPersistence-General-Sequence}
@@
 
-The DSP Broker retrieves the list of matching rules by contacting the Matcher.
-As the DSP Platform have already loaded all the Matching rules, a copy of a
-given DSP message is delivered to the DSP Data Persistence component. However,
-the sequence diagram shows two different stages of reception:
-
 \begin{itemize}
   \item A bootstrap DSP Update Message is delivered to the DSP Data Persistence
-  component when the DSP Platform is initializing the component. In this way,
-  the DSP Data Persistence can initialize the DspDataFlusher worker thread to
+  component when the DSP Platform is initializing the component. For example, the
+  class DSPDataPersistence can initialize the class DspDataFlusher to
   flush temporary messages into the database at the rate of
-  TRANSIENT\underline{ }DATA\underline{ }FLUSHER\underline{ }DELAY in seconds,
-  among others. Listing \ref{file:dsp-data-persistence-bootstrap.xml} shows an
+  TRANSIENT\underline{ }DATA\underline{ }FLUSHER\underline{ }DELAY, defined in 
+  seconds. 
+  Listing \ref{file:dsp-data-persistence-bootstrap.xml} shows an
   implementation of a bootstrap message for the DSP Data Persistence;
   \item Any DSP Measure Message can be persisted by the DSP Data Persistence
-  component. When the DSP Broker delivers a DSP MeasureMessage to this
-  component, it will add the message to a temporary memory location to decrease
-  I/O on the database, as depicted by the gate "create PersistentMessageUnit
-  Gate". This last gate describes the simple procedure to add the message
-  into the transient persistent layer. Listing
-  \ref{file:dsp-message-serialized-ysi}.
+  component. The component will add the message to a temporary memory location
+  to decrease I/O on the database, when the DSP Broker delivers a DSP
+  MeasureMessage to this component, as depicted by the gate "create
+  PersistentMessageUnit Gate". This last gate describes the simple procedure
+  to add the message into the transient persistent layer. Listing
+  \ref{stream:dsp-message-serialized-ysi}.
 \end{itemize}
@@
 carries references regarding the originating DSP Message that was collected by
-the DSP Data Persistence component. It is composed by a SensorLocation and
-contains a reference to the PersistentMessageState enumaration. The former
-identifies which sensor produced the collected data from the DSP Message by the
-IP address, while the latter identifies whether the PersistenceMessageUnit has
-been saved into the Database or not as depicted by the UML State Diagram
-\cite{uml} in Figure \ref{fig:PersistentMessageState-Diagram}.
+the DSP Data Persistence component. It is composed by an instance of the class 
+SensorLocation and contains a reference to the PersistentMessageState
+enumeration. The former identifies which sensor produced the collected data
+from the DSP Message by the IP address, while the latter identifies whether
+the PersistenceMessageUnit has been saved into the Database or not as depicted
+by the UML State Diagram in Figure \ref{fig:PersistentMessageState-Diagram}. 
+The class SensorLocation was developed as a to facilitate the identification of
+sensor devices that are not capable of sampling their geographical location by
+giving their references during the bootstrap process.
 
-\begin{figure}[!b]
+\begin{figure}[!h]
   \centering
-  \includegraphics[scale=0.5]{../diagrams/PersistentMessageState-Diagram}
+  \includegraphics[scale=0.65]{../diagrams/PersistentMessageState-Diagram}
   \caption{UML State Diagram for the instance of the class
@@
 
+When a DSP Message is delivered to the DSP Data Persistence component, the
+message is directly transferred to the Singleton class
+TransientPersistenceLayer, as described before. As described in the UML
+Sequence Diagram in Figure
+\ref{fig:From-Create-PersistentMessageUnit-to-TransientPersistence-Layer-Sequence},
+the first point of execution is the ``create PersistentMessageUnit Gate'',
+where an instance of the SensorLocation is acquired for any DSP Message
+received, in order to be added into the local cache. Before that occurs, the
+PersistentMessageUnit is placed in the state ``TRANSIENT''.
+
 \begin{figure}[!b]
   \centering
-  \includegraphics[scale=0.5]{../diagrams/From-Create-PersistentMessageUnit-to-TransientPersistence-Layer-Sequence}
+  \includegraphics[scale=0.65]{../diagrams/From-Create-PersistentMessageUnit-to-TransientPersistence-Layer-Sequence}
   \caption{UML Sequence diagram - Adding a DSP Message into the Transient Persistence layer}
@@
 
-According to the data model specifications defined in section
-\ref{sec:dsp-persistence-data-model}, the persistence model must carry
-information regarding which sensor device produced the collected data. For
-this reason, an instance of the class SensorLocation carries the information
-about the IP address and and optional GPS location from the component that
-originally produced the DSP Message. Then, an instance of the class
-PersistenceMessageUnit is created at the state ``TRANSIENT'', as detailed
-on UML State Diagram \cite{uml} in Figure \ref{fig:PersistentMessageState-Diagram}.
+At this point, the DSP Data Persistence component maintains a set of DSP
+Messages in a local cache during the specific ``flush rate'' specified during
+the bootstrap message. When the class DSPDataFlusher finds any number of
+instances of PersistentMessageUnit in cache, it starts the process of flushing
+them into the mongoDB database, described in the following section.
 
@@
 
-The last stage sequence described in the UML Sequence diagram of figure
-\ref{fig:From-DSP-Broker-To-DSPDataPersistence-General-Sequence} depicts the
-DspDataFlusher worker thread continuously waking up at the rate of
-TRANSIENT\underline{ }DATA\underline{ }FLUSHER\underline{ }DELAY. Its purpose
-is as simple as to iterate over the list of PersitenceMessageUnit on the state
-of TRANSIENT, and send each of them to be saved on the persistence storage.
-Figure \ref{fig:DSP-Data-Persistence-Classes} shows the participating classes
-of the component.
+The class DSPDataFlusher is a thread designed to be in two different states, as
+depicted in Figure \ref{fig:DSP-DataPersistence-Flusher-State-Diagram}:
 
-\begin{figure}[!b]
+\begin{figure}[!h]
   \centering
-  \includegraphics[scale=0.5]{../diagrams/DSP-Data-Persistence-Classes}
-  \caption{UML Class diagram for the DSP Data Persistence component.}
-  \label{fig:DSP-Data-Persistence-Classes}
+  \includegraphics[scale=0.65]{../diagrams/DSP-DataPersistence-Flusher-State-Diagram}
+  \caption{UML State diagram for the DSP Data Flusher Worker Thread}
+  \label{fig:DSP-DataPersistence-Flusher-State-Diagram}
 \end{figure}
 
-Finally, this sequence of events is continued by the "insert
-PersistenceMessageUnit Gate", which triggers the persistence service from the
-chosen database system as shown in figure
+\begin{itemize}
+  \item \textbf{TIMED\underline{ }WAITING}: the initial state of a Java thread
+  after its initialization, waiting to be on the ``runnable'' state. This
+  period of time is defined by the ``flush rate''  specified in the bootstrap
+  message of the DSP Data Persistence component;
+  \item \textbf{RUNNABLE}: the class DSPDataFlusher contacts the class
+  TransientPersistenceLayer, in order to verify if there are any instance of
+  the class PersistentMessageUnit waiting to be transmitted from the temporary
+  cache to the Database service. If the cache is not empty, the DSPDataFlusher 
+  acquires the lock over the TransientPersistenceLayer and retrieves the
+  existing set of PersistentMessageUnit to be sent to mongoDB. Then, it sends
+  all of the PersistentMessageUnit classes to the Singleton Service class 
+  DSPMongoCRUDService, which is responsible to extract the collected data from
+  the DSP Message's Body and convert it to the Key-Value model specified in
+  Section \ref{sec:dsp-persistence-data-model}.
+\end{itemize}
+
+The main participating classes from the DSPMongoCRUDService are shown in the
+UML Class Diagram \ref{fig:DSP-Data-Persistence-Mongo-Classes}. It keeps references
+to all types of sensor devices such as the SondeDataContainer and the
+MouseActionsContainer. Furthermore, it is responsible to implement the CRUD
+methods to modify the mongoDB database state. The class Mongo is used to acquire
+an instance of the connection with mongoDB server and provide Factory Methods 
+\cite{gof} to create instances of the classes DB, DBCollection and
+BasicDBObject. The class DB is used as a reference to a database namespace, 
+such as ``netbeams'', the one suggested and implemented by this work. Similarly,
+class DBCollection offers an API to manipulate similar documents via the
+abstraction of a collection of items. Finally, the class BasicDBObject is used
+to represent the keys and values to be persisted in the database.
+
+\begin{figure}[!h]
+  \centering
+  \includegraphics[scale=0.65]{../diagrams/DSP-Data-Persistence-Mongo-Classes}
+  \caption{UML Class diagram with Persistence Service for MongoDB}
+  \label{fig:DSP-Data-Persistence-Mongo-Classes}
+\end{figure}
+
+The last method call of the UML Sequence Diagram, depicted in Figure
+\ref{fig:From-DSP-Broker-To-DSPDataPersistence-General-Sequence}, shows the
+method call to the DSPMongoCRUDService method call to persist each of the
+PersistentMessageUnit (see execution point ``addPersistentMessageUnit Gate'').
+In short, the class DspDataFlusher continuously wakes up at the rate of
+``TRANSIENT\underline{ }DATA\underline{ }FLUSHER\underline{ }DELAY''. Its
+purpose is as simple as to iterate over the list of PersitenceMessageUnit on
+the state of TRANSIENT, and send each of them to be saved on the persistence
+storage.
+
+Finally, this sequence of events is continued by "inserting''
+``PersistenceMessageUnit Gate", which triggers the persistence service from the
+chosen database system as shown in Figure
 \ref{fig:From-Insert-PersistentMessageUnit-to-mongoDB}.
 
-\begin{figure}[!b]
+\begin{figure}[!h]
   \centering
-  \includegraphics[scale=0.5]{../diagrams/From-Insert-PersistentMessageUnit-to-mongoDB}
+  \includegraphics[scale=0.6]{../diagrams/From-Insert-PersistentMessageUnit-to-mongoDB} 
   \caption{UML Sequence Diagram - MessageContent instance saved by the class
@@
 
-In order to build the document key and value segments as described on section
-\ref{sec:dsp-persistence-data-model}. First, the value is built by extracting
-the DSP Message Content from the body of the DSP Message.
+\subsubsection{Creating mongoDB Document instance of Key-Values}
 
-\begin{figure}[!b]
+In order to build the document key and value segments, as described in
+section \ref{sec:dsp-persistence-data-model}, the reference to the DSP message
+is used, as the value key is built by extracting the DSP Message Content from
+the body of the DSP Message. Similarly, the key is built by gathering
+information from the DSP Message and the PersistenceMessageUnit class. Each of
+the properties collected from the YSI Sonde sensor device is collected and
+associated with the keys defined for the Value Segment.
+
+\begin{figure}[!h]
   \centering
-  \includegraphics[scale=0.5]{../diagrams/From-Creating-Value-Segment-Sequence}
+  \includegraphics[scale=0.6]{../diagrams/From-Creating-Value-Segment-Sequence}
   \caption{UML Sequence diagram showing the creation of the value document segment}
@@
 
-Finally, the document key segment is prepared as follows in figure
+\newpage
+
+Once the Value Segment is created, the Key Segment is prepared as described in
+the UML Sequence Diagram in Figure
 \ref{fig:From-Creating-Key-Segment-Sequence}.
 
-\begin{figure}[!b]
+\begin{figure}[!h]
   \centering
-  \includegraphics[scale=0.5]{../diagrams/From-Creating-Key-Segment-Sequence}
+  \includegraphics[scale=0.65]{../diagrams/From-Creating-Key-Segment-Sequence}
   \caption{UML Sequence diagram showing the creation of the key document segment}
@@
 
+To Summarize, the DSP Data Persistence Component is added into the DSP Platform
+by introducing a new component that provides service to the transfer any
+collected data from sensor devices to mongoDB, the Key-Value Pair database
+technology. The general dependencies of the system can be seen in UML
+Class Diagram depicted by Figure \ref{fig:DSP-Data-Persistence-Classes}, that
+expands the UML Package Dependency Diagram shown in Figure
+\ref{fig:DSP-Data-Persistence-Packages-Dependency}.
+
+\begin{figure}[!t]
+  \centering
+  \includegraphics[scale=0.65]{../diagrams/DSP-Data-Persistence-Classes}
+  \caption{UML Class diagram for the DSP Data Persistence component.}
+  \label{fig:DSP-Data-Persistence-Classes}
+\end{figure}
+
+Next section analyzes the deployment details for the DSP Platform and the
+mongoDB integration.
+
 \section{DSP Data Persistence Deployment}
 
-The deployment of the persistence layer is comprised by the addition of two
-new components to the DSP Platform: the DSP Data Persistence and the MongoDB
-Database Server. The former is responsible for capturing any data
-produced by sensors in the data sink, as described in the previous sections of
-this chapter. On the other hand, the latter is where the data is going to be
-stored, as a result of the technology selection on the previous chapter.
+The deployment of the persistence layer is accomplished by the addition of two
+new components to the DSP Platform: the DSP Data Persistence component and MongoDB
+Database Server. The DSP Data Persistence component is responsible for capturing 
+any data produced by sensors in the data sink, while mongoDB is where the data is 
+going to be stored, as a result of the technology selection on the previous chapter.
 
-The deployment of the DSP Data Persistence follows the specifications of how to
-deploy a DSP Component on an OSGi Framework. For the development of this work,
-the Knopflerfish OSGi Container was used, as described in section
-\ref{sec:dsp-data-persistence-implementation}. In contrast, the
-deployment of the mongoDB can be accomplished in different approaches, as
-summarized:
+The deployment of the DSP Data Persistence follows the deployment specifications
+of a DSP Component over an OSGi Framework. For the development of this work,
+the Knopflerfish OSGi Container was used. The details of implementation are described
+in section \ref{sec:dsp-data-persistence-implementation}. The deployment of the 
+mongoDB can be accomplished in different approaches, summarized as follows:
 
@@
 
-However, given that mongoDB provides support to Shards, then the implementation
-of a Data-Centric approach can be done by deploying different mongoDB instances
-in a distributed cluster, where each node can hold a given type of data. The
-specification of what types of data should be stored in each of the cluster
-nodes is defined by a shard-key.
+Given that mongoDB provides support for Database Partitioning by using
+the concept of Shards, the implementation of a Data-Centric approach can be done
+by deploying different mongoDB instances in a distributed cluster, where each
+node can hold different collections of data. mongoDB provides Horizontal
+Partition with automatic distribution based on the ranges of a given key of
+the document. For NetBEAMS collected data, a good candidate for partitioning is
+the time when the data was created, and separate each shard in groups of month
+of the year.
 
-\begin{figure}[!b]
+\begin{figure}[!h]
   \centering
@@
 \end{figure}
+
+The design of the DSP Data Persistence uses the DSP Component specification
+defined by the DSP Platform. First, the analysis of the business process
+helped to understand the requirements of the system. The data model design
+followed the specifications defined in chapter 5. Each of the layers of the
+system was described using UML Diagrams, having partial details regarding the
+OSGi implementation. The next chapter entails the experiments conducted to
+evaluate the designs discussed in this chapter, and walks through the
+experiments conducted to verify the implementation of using mongoDB.

### 2009-12-12 docs/marcello-dissertation/thesis/chapter8-experimental-results-correct-behavior-and-performance.tex
Revisions from Niki on Chapters 7 and 8... (artifacts 8 and 9 because 7 was deleted)
Grammatical errors, and addition of a section to chapter 7.

git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@698 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 As discussed in the previous chapter, the implementation of a data persistence
-layer for NetBEAMS followed the requirements defined in chapter 4 having in
-mind the classification of the SF-BEAMS sensor network and the software
-infrastructure provided by the Data Sensor Platform. As a result, the
-implementation was developed using the infrastructure provided by the online
-version-control system repository from NetBEAMS, whose documentation is
-detailed in section \ref{sec:dsp-data-persistence-implementation}.
+layer for NetBEAMS followed the requirements defined in Chapter 4 in regards
+to the classification of the SF-BEAMS sensor network and the software
+infrastructure provided by the Data Sensor Platform. The implementation was
+developed using the infrastructure provided by the online version-control
+system repository from NetBEAMS, whose documentation is detailed in Section
+\ref{sec:dsp-data-persistence-implementation}.
 
-First, this chapter details the experiment setup designed to evaluate the
+This chapter details the experiment setup designed to evaluate the
 proposed implementation, showing the measurement results collected from the
 simulation log and based on the user experience while reusing the collected
-data. Last, but not least, the pertinent discussion of this work is devided up
-into different sections, each of them analyzing different aspects related to
-data persistence for NetBEAMS, amongo others.
+data. The pertinent discussion of this work is divided up into different
+sections, each of them analyzing different aspects related to data persistence
+for NetBEAMS, among others.
 
@@
 
-The experiment was designed to support different scenarios of a regular use of
+The experiment was designed to support different scenarios of regular use of
 the mongoDB system infrastructure, simulating the activities of data collection
@@
 for the data collection of randomly generated data was implemented, which
-directly uses NetBEAMS already existing classes. Then, the implementation of
-the CRUD scenarios were implemented in different programming languages, from
-the data insertion scenario using the Java Programming Language, and then the
-remaining operations using the Javascript Scripting language \cite{javascript}
-and the Python Programming Language \cite{python}. Similarly, an open-source web
-application, developed in the Ruby Programming language \cite{ruby}, was also
-used. The experiment scripts were writen using Shell Bash Script Language
-\cite{bashshell} and is responsible for orchestrasting the execution of one
-single experiment ``run''.
+directly uses already existing NetBEAMS classes. The CRUD scenarios were
+implemented in different programming languages, from the data insertion
+scenario using the Java Programming Language, to the remaining operations using
+the Javascript Scripting language \cite{javascript} and the Python Programming
+Language \cite{python}. Similarly, an open-source web application, developed
+in the Ruby Programming language \cite{ruby}, was also used. The experiment
+scripts were written using Shell Bash Script Language \cite{bashshell} and is
+responsible for orchestrating the execution of one single experimental ``run''.
+
+In order to evaluate the DSP Data Persistence component, experiments were
+conducted using the DSP Mouse Actions component, a software sensor simulator
+that captures the actions performed by the mouse pointer.
 
@@
 
-Different artifacts were created to perform the different experiments
-operations. The Create function was developed in Java, without the interversion
-of the DSP or OSGi, isolating only the functionalities of provided by CRUD
-service class DSPMongoCRUDService (Listing \ref{file:dsp-mongo-service}),
-responsible for the connection handling with mongoDB, as well as the insertion
-operation from NetBEAMS. As described in section \ref{sec:dsp-details}, the
-current version of NetBEAMS supports the data collection for the YSI Sonde
-device, among others used for demo purposes such as a software sensor that
-captures the mouse activities over a window. The documentation of both data
-types is seen in section \ref{sec:dsp-payload-implementation}. In this way, the
-DSP Data Persistence, as described in the previous section, supports both data
-types to be persisted.
+Different artifacts were created to perform the different experiments regarding
+the CRUD operations. The ``Create'' operation' was developed in Java, isolating
+the services provided by the class DSPMongoCRUDService (Listing
+\ref{file:dsp-mongo-service}). It is responsible for the connection handling
+with mongoDB, as well as the insertion operation from NetBEAMS. As described
+\cite{netbeams-dsp-architecture}, the current version of NetBEAMS supports the
+data collection for the YSI Sonde device \cite{YSI-Sonde}, among others used
+for demo purposes. After evaluating the database with the YSI Sonde data
+handler, the use of the demo component DSP Mouse Actions was used. It captures
+the mouse pointer activities performed over a window, for example, mouse clicks
+and mouse grag over coordinates x and y.
 
@@
 the mongoDB, and stored in the collection ``SondeDataContainer''. Specific
-details about the execution of mongoDB are depicted in section
+details about the execution of mongoDB are depicted in Section
 \ref{sec:mongodb-deployment}. Finally, the implementation of the experiment
@@
 On the other hand, the execution of the main experiment shell script requires
-the use to provide the size of the workload, as shown in listing
+the user to provide the size of the workload, as shown in listing
 \ref{cmd:run-persistence-experiment}.
-
-\lstset{label=cmd:run-persistence-experiment,caption=Running main experiment
-shell script}
-\begin{lstlisting}
-marcello@netbeams-dev:~/workspaces/netbeams/versions/v2/persistence$ ./run-persistence-experiment 
-#######  NetBEAMS Experiments - Persistence on MongoDB  ########
-
-Usage: ./run-persistence-experiment X, where X is the size of the workload to be inserted from NetBEAMS YSI Data Handler to mongoDB
-\end{lstlisting}
 
@@
 
-The workload selected for the experiments reflects the current use of the data
-at the SF-BEAMS infrastructure. In this way, the experiment runs for 5
-different times performed to investigate different scenarios to be described.
-In this way, this number is related to the total number of sensor devices
-reported to be in operation, as discussed in section \ref{sec:sfbeams}. In
-summary, the volume of the workload prepared after the experiment setup
-is as follows:
+The workload selected for the experiments reflects the current use of the
+collected data at the SF-BEAMS infrastructure. In this way, the experiment
+runs for 5 different times to investigate different scenarios. This number is
+related to the total number of sensor devices reported to be in operation, as
+discussed in Section \ref{sec:sfbeams}. The volume of the workload prepared
+after the experiment setup is as follows:
 
 \begin{itemize}
-  \item \textbf{First Round}: 483,840 documents, worth the production of the 
+  \item \textbf{First Round}: 483,840 documents, worth of production of the 
   YSI Sonde device during one year, at the rate of 1 observation per minute;
@@
 which is started by specifying the directory where the inserted data will be
-managed. On the other hand, nonetheless, the distributed server setup requires
-a different list of processes (see section \ref{sec:mongodb-user-experience}):
-one ``mongod'' process for each mongoDB shard running on the different mongoDB
-cluster, one ``mongod'' process to manage the cluster metadata, and one
-``mongos'' process, which is the main one responsible to orchestrate the
-cluster. In conclusion, the single system environment is synmmetric to
-the External Storage used to persist the collected data, while the clustered one
-is simulates the Data-Centric Storage one.
+managed. The distributed server setup requires a different list of processes
+(see Section \ref{sec:mongodb-user-experience}): one ``mongod'' process for
+each mongoDB shard running on the different mongoDB cluster, one ``mongod''
+process to manage the cluster metadata, and one ``mongos'' process, which is
+the main one responsible to orchestrate the cluster. The single system
+environment is symmetric to the External Storage used to persist the collected
+data, while the clustered one simulates the Data-Centric Storage one.
 
@@
 operations were exercised through the implementation of the practical
-scenarios based on the use cases described in section \ref{sec:use-cases}. In
+scenarios based on the use cases described in Section \ref{sec:use-cases}. In
 brief, the use cases were chosen to run with random values such as follows:
@@
   tabular format sequence used in the files distributed by SF-BEAMS using the 
-  the OPEnDAP format, as in Listing \ref{file:rtc-ysi-opendap}.
+  the OPeNDAP format, as in Listing \ref{file:rtc-ysi-opendap}.
 \end{itemize}
@@
 on the specifications of a data persistence for NetBEAMS, using the mongoDB
-database system. In this way, after setting up the experiment environment,
-the execution of the workload was performed against the mongoDB database using
-its mechanisms used to manipulate the data generated. It uses the programming
-language abstraction to provide access and modifiy the data collected from the
-NetBEAMS component. For this reason, this section details these mechanisms uses
-to implement each of the scenarios defined in the previous section, associating
+database system. After setting up the experiment environment, the execution of
+the workload was performed against the mongoDB database using its mechanisms
+used to manipulate the data generated. It uses the programming language
+abstraction to provide access and modifiy the data collected from the NetBEAMS
+component. For this reason, this section details these mechanisms used to
+implement each of the scenarios defined in the previous section, associating
 numerical response time collected from the logs.
@@
 instances of the collected data from the sensor devices using Java. The
-document follows the specification of the keys designed in section
-\ref{sec:dsp-persistence-data-model}. The use of the mongoDB Java driver, which
-provides the methods to access the database system and the collection. In this
-way, the construction of the keys and respective values were done using a
-Hash-like object ``BasicDBObject'' such as the ones shown in the method
-``buildKeySegment''. Similarly, the creation of database indexes is performed
-using the same approach, as shown in the Java method ``setupMetadataIndexes''.
-Finally, the insertion of the documents is performed on the service method
-call ``insertPersistentUnitMessageContents'', which delimits the bulk insert
-using the transaction management mechanism over the collection. In this way,
+document follows the specification of the keys designed in Section
+\ref{sec:dsp-persistence-data-model}. The construction of the keys
+and respective values were done using a Hash-like object ``BasicDBObject''
+such as the ones shown in the method ``buildKeySegment''. Similarly, the
+creation of database indexes was performed using the same approach, as shown in
+the Java method ``setupMetadataIndexes''. Then, the insertion of the
+documents was performed on the service method call
+``insertPersistentUnitMessageContents'', which delimits the bulk creation using
+the transaction management mechanism over the collection. In this way,
 different insert times were collected from the logs of the different
-experiments as summarized on Table \ref{tab:experiment-insert-avarage}.
+experiments as summarized on Table 7.1.
 
-\begin{table}
+\begin{table}[!h]
     \begin{center}
@@
 
-After executing the experiments runs, the claimed disk space for
-the representation of the randomly generated data is separated into two
-different sizes: one related to the amount of data produced by one YSI device,
-and another produced by five different YSI devices, as summarized on Table
-\ref{tab:experiment-insert-diskspace}.
+After executing the experiment, the claimed disk space for the representation
+of the randomly generated data was separated into two different sizes: one
+related to the amount of data produced by one YSI device, and another produced
+by five different YSI devices, as summarized on Table 7.2.
 
-\begin{table}
+\begin{table}[!h]
     \begin{center}
@@
 
-When it comes to the simulation of the use case scenarios, the implementation
-used the mongoDB shell client, which uses Javascript, to implement most of the
-operations for ``retrieve'', ``update'' and ``delete'' shown in Listing
-\ref{file:experiment-scenarios}. On the other hand, the implementation of the
-use case to export the collected data to OPeDAP format used Python, depicted on
-Listing \ref{file:experiment-export-python}. The access pattern to use the
-collected data is ``db.collection\underline{ }name.operation", where
+The simulation of the use case scenarios for the implementation used the mongoDB
+shell client, which uses Javascript, for most of the operations for
+``retrieve'', ``update'' and ``delete'' shown in Listing
+\ref{file:experiment-scenarios}. The implementation of the use case to export
+the collected data to OPeDAP format used Python, depicted on Listing
+\ref{file:experiment-export-python}. The access pattern to use the collected
+data is ``db.collection\underline{ }name.operation", where
 ``collection\underline{ }name'' is the name of the collection used for the
-collected data, and in this case, the name of the DSP component that represents
-the collected data. Therefore, mongoDB uses dynamic object binding for the
-collection object, and executes the CRUD operations as method call with a list
-of the needed parameters. Table \ref{tab:experiment-scenarios-mongo-methods}
-relates method calls used to implement each of the use cases defined, as
-implemented in Listing \ref{file:experiment-scenarios}.
+collected data, and in this case, the name of the DSP component that
+represents the collected data. Therefore, mongoDB uses dynamic object binding
+for the collection object, and executes the CRUD operations as method calls
+with a list of the needed parameters. Table 7.3 relates method calls used to
+implement each of the use cases defined, as implemented in Listing
+\ref{file:experiment-scenarios}.
 
-\begin{table}
+\begin{table}[!h]
     \begin{center}
@@
 collected during the execution of the experiment script manager. The retrieval
-operations ran in avarage for ~300ms, while the updates took more than 10
-seconds when it involved larger data sets with more than 300 thousand
-documents. Similarly, the delete operation took less
+operations ran in average for 300ms, while the updates took more than 10
+seconds when it involved larger data sets with more than 300,000 documents.
+Similarly, the delete operation took less than 13 seconds to perform.
 
@@
 
-In general, the use experience of mongoDB is related to the use of programming
-languages, since most of the methods to create, access and modify the database
-is performed using the programming language abstraction of objects and method
-calls. The creation of data was the first contact with mongoDB in terms
-of data creation in Java. Later, Different methods of data access were used
-after the randomly generated data was inserted into mongoDB: using the shell
-commands mongo and ``mongoexport'', command-line execution developed in
-Python, through the REST Web Services API, and using a web application,
-written in the Ruby Programming Language \cite{ruby}, providing the view over
-the collected data using a web browser. More details about the execution of
-these commands during the experiments is shown in section
-\ref{sec:mongodb-user-experience}.
+In general, the user experience of mongoDB is related the methods used to
+access and modify the state of the database using a different approach when
+compared to traditional Relational databases. Since most of the methods to
+create, access and modify the database were performed using the programming
+language abstraction of objects and method calls, the transition from the use
+of a relational data model is very smooth. The creation of data was the first
+contact with mongoDB in terms of data creation in Java. Later, different
+methods of data access were used after the randomly generated data was
+inserted into mongoDB: using the shell commands mongo and ``mongoexport'',
+command-line execution developed in Python, through the REST Web Services API,
+and using a web application, written in the Ruby Programming Language
+\cite{ruby}, providing the view over the collected data using a web browser.
+More details about the execution of these commands during the experiments are
+shown in Section \ref{sec:mongodb-user-experience}.
 
-First and foremost, the primary method to access to the database is using the
-shell command. Once the user requests an access to the database, he or she can
-issue an operation over the collection of documents using the regular the
-abstraction programming language, as the shown in Listing \ref{cmd:mongo}. The
-CRUD operations are available to the user and can be used at any time to any
-database collection. Aggregation operations such as to count the returned
-resultset is also provided via programming languages abstraction.
+The primary method used to access the mongoDB database is using the shell
+command. Once the user requests access to the database, he or she can issue an
+operation over the collection of documents using regular the abstraction
+programming language, as the shown in Listing \ref{file:experiment-scenarios}.
+The CRUD operations are available to the user and can be used at any time to
+any database collection. Aggregation operations such as to count the returned
+result set is also provided via programming languages abstraction.
 
@@
 The ``mongoexport'', exports the collected data from the database using the
-native support to comma-delimited values. As shon in Listing
-\ref{file:mongodb-export-command}, the actual execution, the remaining ones
-describe how mongo is querying the data from the database. As a result, the
-total number of documents exported is shown in on line 8. In addition to the
-regular command that exports the entire database, the user can provide a query
-string during the execution of the command in order to filter the dataset
-returned and, consequently, the number of documents exported. In contrast to
-the use of the default export capabilities of mongoDB, an Export script written
-in Python, as shown in Listing \ref{file:experiment-export-python}. It uses the
-package ``pymongo'' and creates a comma-delimited file with the same header
-format and lines as shown in the HTTP Response body of Listing
-\ref{file:rtc-ysi-opendap}. The results can be compared with the example of
-Listing \ref{file:experiment-export-python-results}.
+native support to comma-delimited values. As shown in Listing
+\ref{file:mongodb-export-command}, the actual execution describe how mongo is
+querying the data from the database. As a result, the total number of documents
+exported is shown in on line 8. In addition to the regular command that
+exports the entire database, the user can provide a query string during the
+execution of the command in order to filter the dataset returned and,
+consequently, the number of documents exported. Similar to the use of the
+default export capabilities of mongoDB, a script written in Python to export
+data from mongoDB was written as shown in Listing
+\ref{file:experiment-export-python}. It uses the package ``pymongo'' to create
+a comma-delimited file with the same header and line format as shown in Listing
+\ref{file:experiment-export-python-results}. The results can be compared with
+the example the output of the HTTP Response body of Listing
+\ref{file:rtc-ysi-opendap}, requested directly to the OPeNDAP server of RTC.
+Therefore, the experiment is capable of reproducing theat the OPeNDAP can.
 
@@
 are provided. Then, the REST operation uses the name of the database and
-the collection to correlate with the ones in the database system. In this way,
+the collection to correlate with the ones in the database system. Thus,
 querying the collections from mongoDB can be performed using the HTTP Request
 Parameter variables, by adding the prefix ``filter\underline{ }'' before the
-name of the key of the document. The given example shows the query performed in the
-collected data for the key ``observation.Conductivity'' with the value
+name of the key of the document. The given example shows the query performed
+in the collected data for the key ``observation.Conductivity'' with the value
 ``104.5''. Similarly, the HTTP Request variable ``limit'' is used to reduce
-the number of documents returned by the query. Finally, the HTTP Response from
-mongoDB returns the result using the JSON format, similar to the ones printed
-in the command-shell.
+the number of documents returned by the query. The HTTP Response from mongoDB
+returns the result using the JSON format, similar to the ones printed in the
+command-shell.
 
@@
 reached, making the component to send the messages to the persistence layer, so
-that it can save on mongoDB. Figure \ref{fig:experiment-mouseactions-dsp} shows
-a screenshot of the moment when the message is sent to mongoDB, being verified
-through the HTTP REST Interface using a web browser.
+that it can be saved on mongoDB. Figure \ref{fig:experiment-mouseactions-dsp}
+shows a screenshot of the moment when the message is sent to mongoDB, being
+verified through the HTTP REST Interface using a web browser
+\ref{sec:dsp-mongodb-rest-ws}.
+
+\begin{figure}[h]
+  \centering
+  \includegraphics[scale=0.4]{../diagrams/experiment-mouseactions-dsp}
+  \caption{Experiment showing the mongoDB persisting messages from the DSP
+  Mouse Actions Demo}
+  \label{fig:experiment-mouseactions-dsp}
+\end{figure}
 
@@
 The implementation of the data persistence for NetBEAMS follows the
-specifications explained in chapter 4. The purpose of sensor network for
-NetBEAMS is to suport data archival from any collected data, and for this
-reason, mongoDB was selected after an empirical analysis described in chapter
-5. In this way, the requirements described in chapter 4 were expressed as
-examples of use cases in section \ref{sec:exp-scenarios}. This chapter
-discusses the experiments in different angles: analysis of the taxonomies, the
-infrastructure and the use cases evaluation.
+specifications explained in Chapter 4. The purpose of the sensor network for
+NetBEAMS is to support data archival from any collected data, and for this
+reason, mongoDB was selected after an empirical analysis described in Chapter
+5. The requirements described in Chapter 4 were expressed as examples of use
+cases in Section \ref{sec:exp-scenarios}. This chapter discusses the
+experiments from different angles: analysis of the taxonomies, the
+infrastructure, the use cases evaluation and applicability in other projects.
 
@@
 the DSP Platform such as the YSI Data Handler and the Mouse Actions demo data.
-As it is shown in section \ref{sec:exp-measurements}, the collected data is
+As it is shown in Section \ref{sec:exp-measurements}, the collected data is
 archived in files using the BSON format \cite{bson}, a binary representation
@@
 \ref{file:mongodb-ysi-data-format}. Along with the purpose of Data Archival,
-the use of mongoDB also met the specification of an External Storage, since the
+the use of mongoDB also met the specification of External Storage, since the
 data persisted in the file system, as well as Data-Centric for running mongoDB
@@
 
-SF-BEAMS uses Tabular Data model, Schema-less data model that represents the
+SF-BEAMS uses Tabular Data model, a schema-less data model that represents the
 collected data using comma-delimited files, without a database system to manage
-those files, just the OPEnDAP middleware. In contrast, the data persistence
-proposed by this work proposes mongoDB as a schema-less database system to
-manage the collected data collected using NetBEAMS. As described in chapter 5,
-the schema-less data model can provide a better support to Data Provenance
-because the data model gives support to the change of records without affecting
-the others. In this way, the metadata used by the implementation clearly
-describes the three fundamental questions regarding the collected data from
-sensor devices, as described in section \ref{sec:dsp-persistence-data-model}.
-First, the identity of the data is given by the key ``message\underline{
-}id'', in a way that it uniquely identifies the data produced by the sensor
-identified by the key ``sensor.ip\underline{ }address''. In this way, the
-message can be tracked back to the sensor that produced the data for different
-purposes such as tracking the data producer. Second, time dimensions were used
-in order to track two different aspects of the data collection: the valid time
-can be used to correlate observations to the exact moment it occurred, while
-the transaction time can be used to verify when the data was collected, and
-make decisions such as deleting the instances of data were collected in the
-previous two days, etc. Last, but not least, the use of the key
-``sensor.location'' can give the exact location of the data, since researchers
-can be interested in finding data based on a specific location. Finally, the
-most important aspect of the schema-less data model is that the observed data
-is added to the key ``observation'', as any sensor device carries its set of
-attributes of key-value pairs. A good example of the applicability of the
-schema-less model is the scalability regarding changes to the data structure.
-For instance, if an existing sensor device is flushed with a new version of
-the firmware\footnote{Primary machine instructions that run in the hardware},
-changing the format of data types or adding new values, the following instances
-of collected data may be different to the previous one by simply adding the
-new additions of data. For this reason, the DSP Data Persistence gives
-administrators the ability to choose which fields to persist by declaring it
-using a bootstrap message for the component, as shown in the parameter
-``YSI\underline{ }DESIRED\underline{ }PROPERTIES'' of Listing
-\ref{file:dsp-data-persistence-bootstrap.xml}.
+those files, just the OPeNDAP middleware. In contrast, the data persistence
+method proposed by this work suggests mongoDB as a schema-less database system
+to manage the data collected using NetBEAMS. As described in Chapter 5, the
+schema-less data model can provide better support to Data Provenance because
+the data model allows the change of records without affecting the
+others. In this way, the metadata used by the implementation clearly describes
+the three fundamental questions regarding the collected data from sensor
+devices, as described in Section \ref{sec:dsp-persistence-data-model}. The 
+identity of the data is given by the key ``message\underline{ }id'', in a way
+that it uniquely identifies the data produced by the sensor identified by the
+key ``sensor.ip\underline{ }address''. Thus, the message can be tracked
+back to the sensor that produced the data for different purposes such as
+tracking the data producer. Then, time dimensions were used in order to
+track two different aspects of the data collection: the valid time can be used
+to correlate observations to the exact moment it occurred, while the
+transaction time can be used to verify when the data was collected, and make
+decisions such as deleting the instances of data collected in the previous two
+days, etc. Similarly, the use of the key ``sensor.location'' can give the exact
+location of the data, since researchers can be interested in finding data
+based on a specific location. The most important aspect of the schema-less data
+model is that the observed data is added to the key ``observation'', as any
+sensor device carries its set of attributes of key-value pairs. A good example
+of the applicability of the schema-less model is the scalability regarding
+changes to the data structure. For example, if an existing sensor device is
+flushed with a new version of the firmware\footnote{Primary machine
+instructions that run in the hardware}, changing the format of data types or
+adding new values, the following instances of collected data may very
+from the previous one by simply adding the new additions of data. For this
+reason, the DSP Data Persistence gives administrators the ability to choose
+which fields to persist by declaring it using a bootstrap message for the
+component, as shown in the parameter ``YSI\underline{ }DESIRED\underline{
+}PROPERTIES'' of Listing \ref{file:dsp-data-persistence-bootstrap.xml}.
 
@@
 provided by the mongoDB community. The Centralized Query Mechanism was
-satisfied by the use of those APIs, which is also fultill the requirements
-defined by NetBEAMS. For example, the access of the database using Python or
+satisfied by the use of those APIs, which also fulfills the requirements
+defined by NetBEAMS. For example, the accesses of the database using Python or
 Java were based on the needs and the use of the languages for different
@@
 server execution provided enough computational power for a compatible volume of
-data when compared to the one produced from the SF-BEAMS sensor network. In
-this way, the implementation suggested by this work fulfill the need of a
-persistence system for a Small volume of data being created at the same type
-in a Single Server. Furthermore, mongoDB also provided support for Distributed
-System, when the use of databse shards, in forms of different mongoDB nodes,
-were used as a clustered environment.
+data when compared to the one produced from the SF-BEAMS sensor network. The
+implementation suggested by this work fulfill the needs of a persistence system
+for a Small volume of data being created at the same time in a Single Server.
+Furthermore, mongoDB also provided support for a Distributed System, when the
+use of database shards, in forms of different mongoDB nodes, was used as a 
+clustered environment.
 
@@
 Overall, the implementation of the DSP Data Persistence component meets the
-specification of the use cases defined in section \ref{sec:exp-scenarios}.
+specification of the use cases defined in Section \ref{sec:exp-scenarios}.
 From the perspective of disk-space utilization for the compatible volume of
-data for 5 five YSI devices is that it requires a small amount of disk space.
+data for five YSI devices is that it requires a small amount of disk space.
 According to the mongoDB documentation \cite{mongodb}, the claimed amount of
@@
 documents, which are composed only of strings, together with the definition of
-the indexes. As a consequence, the size of the keys defined for the YSI sonde
-data suggests additional disk space were used for the storage.
+the indexes. The size of the keys defined for the YSI Sonde data suggests
+additional disk space was used for the storage.
 
@@
 thousands of documents per second. This capacity is more than necessary for
-NetBEAMS, which can use the Single System environment. On the other hand,
-improvements over the response time can be achieved through the implementation
-of a Data-Centric approach on a Distributed Database environment, or clusters.
-In this approach, the inserted data is stored in different servers based on a
-partition key. Given the fact that less data is concentrated in each
-partition, the centralized query mechanism used is faster, and consequently
-improves the performance as described in the Data Location taxonomy in chapter
-3. In this way, the infrastructural performace during the inserts suggests that
-persisting data collected from the NetBEAMS sensors using a clustered
-environment is an option when the single server capacity is reached.
+NetBEAMS, which can use the Single System environment. Improvements over the
+response time can be achieved through the implementation of a Data-Centric
+approach on a Distributed Database environment, or clusters. In this approach,
+the inserted data is stored in different servers based on a partition key.
+Given that less data is concentrated in each partition, the centralized query
+mechanism used is faster, and consequently improves the performance as
+described in the Data Location taxonomy in Chapter 3. In this way, the
+infrastructural performance during the inserts suggests that persisting data
+collected from the NetBEAMS sensors using a clustered environment is an option
+when the single server capacity is reached.
 
 Different problems and limitations were faced during the execution of
-mongoDB, which were relate to the infrastructure used: the limitations
+mongoDB, which were related to the infrastructure used: the limitations
 regarding disk storage in 32bit operating systems and very slow
@@
 architecture, there are limitations on the disk space used to store
-collections. In this way, after reaching 2GB of collected data during many
-different runs, mongoDB crashed and did not allow the use of the collected
-data. For this reason, the setup of a new VirtualBox using Ubuntu Linux 64bits
-as a guest to a Mac OS X 64bits was prepared and used throughout the completion
-of this work's experiments. On the other hand, the performance was drastically
-decreased using mongoDB in virtualized servers due to the fact of dynamic
-partition chosen. The performance shell application ``iostat'' was used to
-verify the behavior, and the use of mongoDB was switched to the Host operating
-system.
+collections. After reaching 2GB of collected data during many different runs,
+mongoDB crashed and did not allow the use of the collected data. For this
+reason, the setup of a new VirtualBox using Ubuntu Linux 64bits as a guest to
+a Mac OS X 64bits was prepared and used throughout the completion of these
+experiments. On the other hand, the performance was drastically decreased 
+using mongoDB in virtualized servers due to the fact of dynamic partition
+chosen. The performance shell application ``iostat'' was used to verify the
+behavior, and the use of mongoDB was switched to the host operating system.
 
@@
 
-\cite{sn-programming-language} suggested that programming that programming
-languages are easier abstractions to researchers that use collected data from
-sensor devices without background in computer science or database systems. The
-authors have suggested the implementation of a programming language for this
-group of biologists, for instance. However, this suggestion has good
-intentions, but limits the users with a single mechanism to access the
-collected data from sensors. For this reason, this dissertation proposes the
-use of key-value pair databases, such as mongoDB, to provide transparent access
-the the data through the abstraction of any programming language that has a
-driver written for mongoDB. As one can see during the user experience section,
-different programming languages were used to access the collected data, a novel
-mechanism not yet reported in the sensor networks community.
+\cite{sn-programming-language} suggested that programming languages are easier
+abstractions to researchers that use collected data from sensor devices
+without a background in computer science or database systems. The authors have
+suggested the implementation of a programming language for this group of
+biologists, for instance. This suggestion has good intentions, but
+limits the users with a single mechanism to access the collected data from
+sensors. For this reason, this report proposes the use of key-value pair
+databases, such as mongoDB, to provide transparent access the data through
+the abstraction of any programming language that has a driver written for
+mongoDB. During the user experience section, different programming languages
+were used to access the collected data, a novel mechanism not yet reported in
+the sensor networks community.
 
@@
 as the ones defined in Listing \ref{file:experiment-scenarios}, which
-implements the use cases for data retrieval, update and deletion. The main
-advantage of this approach is that users do not use the SQL language to extract
-data from the database. Rather, the use of the method calls defined for each
-operation, as well as the concept of dynamic binding and objects, gives simple,
-yet powerful syntax to access data. For example, based on the description
-of the collected data by keys and values shown in Listing
+implements the use cases for data Retrieval, Update and Deletion operations. The
+main advantage of this approach is that users do not use the SQL language to extract
+data from the database, but use method calls defined for each the
+abovementioned operations. For example, based on the description of the
+collected data by keys and values shown in Listing
 \ref{file:mongodb-ysi-data-format}, different use cases can evaluate documents
-in this structure. Usually researchers may search the collected data by a given
-range of dates, for their regular activities of data analysis. The scenario
-implemented by R1 in the abovecited listing shows it clearly. Similarly, the
-search for data in a given sensor device is expressed by the search for the IP
-address for them on the use cased defined by R2. Finally, the search for the
-raw data obtained by the sensor devices is used from the key ``observation'',
-as shown by the R3 use case implementation.
+in this structure. Usually researchers may search the collected data by a
+given range of dates, for their regular activities of data analysis. The
+scenario implemented by R1 in the above cited listing shows it clearly.
+Similarly, the search for data in a given sensor device is expressed by the
+search for the IP address for them on the use cased defined by R2. Finally,
+the search for the raw data obtained by the sensor devices is used from the
+key ``observation'', as shown by the R3 use case implementation.
 
-Accompanying that non-predictable events might occur during the data
-collection process of sensor devices, and consequently, influence in their
-values, the support to annotations is a feature that can be used to describe
-the data. mongoDB indirectly supports the use of annotation because of its free
-method of updating documents and adding new keys to any of them without
-requiring design changes. Supposing an environmental accident happens, such as
-the San Francisco oil spill of October 2009 \cite{sfbay-oilspill2009}, the
-values of the collected data, and consequently data analysis, be influenced by
-different ranges of values. In this way, the implementation of the scenario U1
-provided the implementation of the addition of a new key called ``tag'', with
-the annotation ``oil spill''. This new key, after properly indexed, provides
-another option to meaningfully search the collected data from the sensor
-devices. Last, but not least, old data can be deleted as expressed by D1, as an
-operation supported by any database system.
+Allowing for non-predictable events that might occur during the data
+collection process of sensor devices, which will influence their values, the
+support to annotations is a feature that can be used to describe the data.
+mongoDB indirectly supports the use of annotation because of its method
+of updating documents and adding new keys to any of them without requiring
+design changes. Supposing an environmental event occurs, such as the recent oil
+spill in the San Francisco bay in October 2009 \cite{sfbay-oilspill2009}, the
+values of the collected data, and data analysis, was influenced by different
+presence of different objects in the water. The implementation of the scenario
+U1 provided the addition of a new key called ``tag'', with the annotation ``oil
+spill''. This new key, after properly indexed, provides another option to
+meaningfully search the collected data from the sensor devices. Old data then
+can be deleted as expressed by the scenario D1, as a regular operation
+supported by any database system.
 
-One common approach during research is data sharing. Researchers can choose
+Commonly during scientific research, data is shared. Researchers can choose
 the use of spreadsheets to analyze collected data with simpler tools than
-database systems or programming language. For this reason, OPEnDAP is used in
-the community to share data using a comma-delited format. In this way, the
-implementation of the use cases for R4 were developed in two different ways.
+database systems or programming language. For this reason, OPeNDAP is used in
+the community to share data using a comma-delimited format. In this way, the
+implementation of the use cases for R4 was developed in two different ways.
 First, the use of the export capabilities to a common-delimited values was
-performed \ref{file:mongodb-export-command}, taking approximaly 3.3 minutes to
-export one year worth of data for 5 YSI devices without indexing. Second, the
+performed \ref{file:mongodb-export-command}, taking proximally 3.3 minutes to
+export one year's worth of data for 5 YSI devices without indexing. Second, the
 implementation of a custom export utility using Python, as shown in Listing
 \ref{file:experiment-export-python} enabled the creation of the comma-delimited
-files with the same format as the ones used by OPEnDAP. The comparison can be
-done by analyzing the output of the OPEnDAP request from SF-BEAMS on Listing
+files with the same format as the ones used by OPeNDAP. The comparison can be
+done by analyzing the output of the OPeNDAP request from SF-BEAMS on Listing
 \ref{file:rtc-ysi-opendap}, and the output provided from the script written in
@@
 Furthermore, considering research teams with skills in Web Services, an easy
-ubiquos integration with mongoDB can be provided by using the REST Web Services
-APIs provided by mongoDB, still in alpha version. A practical application of
-this API was the created Apache Futon 4 mongoDB web application, used to give
+ubiquitous integration with mongoDB can be achieved by using the REST Web
+Services APIs provided by mongoDB, still in alpha version. A practical application of
+this API was the created Apache Futon4mongoDB web application, used to give
 user friendlier access to the collected data as shown in 
@@
 
-As discussed in chapter 5, the differences between a traditional database
-system, which uses the relational data model, and a document-oriented database
-system were clearly noticed during the development of this work. As mentioned
-in the section \ref{sec:experiments-user-experimence}, the user experience was
+As discussed in Chapter 5, the differences between a traditional database
+system, which uses the relational data model, and document-oriented database
+systems were clearly noticed during the development of this work. As mentioned
+in Section \ref{sec:experiments-user-experimence}, the user experience was
 mostly related to the definition of the keys that describe the sensor device,
-as discussed during the design of the solution in chapter 6. On the contraty,
-the abstraction of objects and collections during the data access using
-different programming languages. Therefore, based on the use of a different
-type of data model in the scope of data persistence for sensor networks, the
-taxonomies defined in chapter 3 could be updated to the one shown in Figure
+developed as discussed during the design of the solution in Chapter 6. 
+Therefore, based on the use of a different type of data model in the scope of
+data persistence for sensor networks, the taxonomies defined in Chapter 3
+could be updated to the one shown in Figure
 \ref{fig:taxonomy-data-model-modified}, including the Key-Value Pair Data
@@
 
+\subsection{Applicability and Recommendations}
+
+This work can be used to evaluate data persisten in sensor networks in
+different ways. The use of the data persistence taxonomies through this work
+and guided both the evaluation and implementation of the data model using
+mongoDB. In order to use the guidelines proposed by this work, one may first
+determine the properties of the sensor network in question. Considering that
+sensor devices produce streams of non-described data, the use of key-value
+pairs data model did not require changes throughout different application
+layers.
+
+Focusing on the different layers of an application, consider the different
+multi-tier application approaches where each layer implements a functionality
+of the system such as the Model-View-Controler (MVC) \cite{sw-architectures}.
+Usually, the use of traditional database systems and object-programming 
+languages require ``data plumbing'' from the table relations from
+the database system to the object instances. In order to implement all the
+CRUD operations to a relational database without an ORM \cite{orm} offering,
+the implementation of the data model potentially uses the Structured-Query
+Language (SQL) for the problem of data plumbing, what results in a
+tightly-coupled implementation. On the other hand, mongoDB eliminates the use
+of an additional query language because it uses declarative approach on top of
+the programming language drivers it support, decreasing the complexity of the
+application.
+
+Sensor networks with medium or large volumes of data are implemented nowadays
+using clusters and parallel programming. Google's BigTable \cite{bigtable} uses
+the principle of structured data in distributed tabular format, processing large
+volumes of data using the MapReduce parallel approach \cite{map-reduce}. The
+current version of mongoDB offers its first version of the MapReduce approach,
+but it is still imature. However, it can be useful to process large amount of
+data collected from sensor devices, or process historical data for any type of
+analysis.
+
 \subsection{Difficulties Experienced}
@@
 
-The inception of the experiments described in by this work moments of Software
-Engineering. Particularly speaking about the learning curve, different
-situations contributed to the success of the first experiments. Since the
-technoogy is new, it brought new challenges regarding the usability and user
-experience. The technology is relatively new and still under development,
-which many different bugs were be reproduced by random during the execution of
+Particularly speaking about the learning curve, different situations
+contributed to the success of the first experiments. The first problem to be
+solved was the selection of the data model to be used during the technology
+selection. Since relational database systems are the ``traditional'' way to
+provide data persitence, the investigations started by creating a relational
+model described in Chapter 5. Then, the technique using KVP data models in
+relational databases were attempted without success. Only after reviewing the
+technical article \cite{db-is-rdbs-dommed}, suggested the advisor of this 
+work, the search for a database system that uses the KVP data model was
+considered, and finally successfully implemented.
+
+Since the technology is new, it brought new challenges regarding the usability
+and user experience. The technology is relatively new and still under development,
+many different bugs were be reproduced at random during the execution of
 the experiments, and therefore, reports were sent to the open-source community
-through use different available communication channels. First, the fastest way
-to solve an issue was through the mongoDB's IRC channel located at the Freenode
-server (www.freenode.org, channel mongodb). Second, the use of the users
-group, located at http://groups.google.com/group/mongodb-user also helped
-solving problems.
+through various communication channels. To address this, the fastest way to
+solve an issue was through the mongoDB's IRC channel located at the Freenode
+server (www.freenode.org, channel mongodb). Second, the users group, located at
+http://groups.google.com/group/mongodb-user also helped troubleshooting
+problems.
 
@@
 totally consumed by one of the revisions of the experiments setup. Therefore, 
-it was learned that large amount of data loads require different approach to
+it was learned that large amount of data loads require different techniques to
 manage data in memory accordingly, such as creating different object pools as
-temporary cache for objects whose state are similar to others.
+temporary caches for objects whose internal state are similar to others in the
+Java Virtual Machine.
 
-Also related to the difficulties of this work, the fact that mongoDB is not yet
-stable makes the implementation a bit harder. The main mongod process is on its
-1.1.4 version as the time this line is writte, while the shards support is
-still in alpha version. For instance, different work-arounds were performed to
-have shards experimented as described, but the execution was not successful.
+Also related to the difficulties of this work, mongoDB is not yet stable, and
+therefore, harder to implement. The main production version of mongoDB is
+is on version 1.1.4 at this time, while the shards support is still in alpha
+version. For instance, different work-arounds were performed to have shards
+experimented as described, but the execution was not successful.
 
@@
 the solution in general. The following section brings the conclusions and
-future works for this dissertation.
+future works for this report.

### 2009-12-12 docs/marcello-dissertation/thesis/chapter9-conclusions-and-future-works.tex
Revisions from Niki on Chapters 7 and 8... (artifacts 8 and 9 because 7 was deleted)
Grammatical errors, and addition of a section to chapter 7.

git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@698 559edde2-a4d4-44e0-96fc-4035e40d7396

--- 2009-12-01 docs/marcello-dissertation/thesis/chapter9-conclusions-and-future-works.tex
@@
 
-This work presented a collection of taxonomies related to data persistence for
-sensor networks using different approaches ranging from data models and server
-infrastructure organization. The main result of this work was that persistence
-for collected data from sensors are better maintained using schema-less data
-models, since it provides offers an approach to deal with unanticipated data
-types from the sensor devices.
+There were significant remarks covered in this work, including the
+importance that sensor networks are becoming ubiquitous, having their produced
+data available online and other types of media. In order to provide access to
+the collected data, one must first assess the different types of
+infrastructural characteristics of the studied sensor networks. The selection
+of a technology that provides such a persistence solution capabilities can be 
+part of a process of analysis of taxonomies proposed by this work,
+characterized as the first contribution of this report.
 
-There are many important features covered in this work, including the
-importance of the sensor networks rely in how users have access to the
-collected data. In order to provide access to the collected data, one must
-first assess the different types of infrastructural characteristics of the
-studied sensor network. In general, the selection of a technology that
-provides such persistence capabilities can be part of a process of analysis of
-taxonomies proposed by this work. However, this work proposes a data model not
-yet used in the sensor network community and, thus, represents an important
-contribution for the area.
+Based on the findings of the taxonomies, together with the empirical analysis
+of the technologies, this work revealed a novel approach to provide
+persistence and access to collected data from sensor devices using the
+Key-Value-Pair data model. Besides the contribution of a new component for the
+Data Sensor Platform, the evidence that the system improves productivity was
+confirmed during the execution of the experiments, writing less source-code to
+obtain better functionalities when compared to other solutions. Similarly,
+improved work suggests that the abstraction of programming language should be
+used with research groups whose primary expertise is not in Computer Science
+or Database Systems.
 
-Taking into account the software architecture provided by NetBEAMS, as well as
-the requirements for data collection from the RTC research group, this work
-shows a novice way to persist data collected from sensor devices by using a
-schema-less data model, whose query capabilities are based upon programming
-languages, in contrast to the use of the Structured Query Language, SQL. The
-experiment results showed that the implementation of a persistence layer with
-mongoDB can be used to persist data at its maximum rate of one sample per
-minute. However, performance can be improved by using techniques such as
-database sharding, since it offers a direct implementation capability to
-Data-Centric Storage approach. This approach decreases the amount of time to
-search data since it distributes the data instances across different cluster
-nodes.
+The DSP Data Persistence component improves NetBEAMS's support to SF-BEAMS,
+incorporating a persistence layer for the sensor network that can be used in
+different ways. The solution for data persistence using Key-Value Pair
+databases is a novel way to save collected data from sensor devices into a
+single or multiple servers of a database technology. mongoDB's ability to cope
+with the uncertainty of any properties collected from any sensor specification
+is the most compeling reason to use the Key-value pair data model for collected
+data from sensor devices.
 
-There are many different directions for future works. First, the Data-Centric
-implementation performance may be extremely improved by using the technique
-called MapReduce \cite{map-reduce}. This approach allows mongoDB to search
-different machines in parallel whenever the data requested is located in
-different database shard hosts. In addition to that, the use techniques to
-cluster collected data before they are inserted into the database may be
-important as shown in \cite{sn-time-series}. One of the key issues in the
-implemented solution is that any measurement of data is considered to be
-persisted without further analysis.
+As far as future works, there are many different directions for exploration.
+The first major release of mongoDB with complete support to shards is
+scheduled for 2010. An important mechanism of mongoDB is that it can
+manage different mongoDB servers spread out in a cluster. This feature will give
+better support to the Data-Centric Storage approach. This approach can reduce
+the retrieval time by focusing on centralized selection of data over a given
+database shard. Moreover, the overall search through the collected data
+attributes can be drastically decreased by a parallel search over the database
+shards using the MapReduce \cite{map-reduce} tools. At the time of development
+of the experiments, the MapReduce API from mongoDB was also in the alpha
+version, and therefore, presented many bugs. One suggestion for the
+implementation of such functionality is based in Figure
+\ref{fig:future-works-data-centric-map-reduce}, which shows an architectural
+idea of the deployment of a Data-Centric storage approach. When searching for
+specific document attributes, the process is sent in parallel to each of the
+database shards, and the cluster head collects the result. After calculating
+the final result, mongoDB server returns the relevant result to the user.
 
-All in all, the DSP Data Persistence component implemented for the NetBEAMS
-solution to SF-BEAMS can be used in any network server. The solution for data
-persistence is a novice way to save data into the database, provided by
-mongoDB's ability to cope with the uncertainty of any properties collected from
-any sensor specification. Furthermore, the solution also provides a good
-support to the provenance-related data.
+\begin{figure}[!h]
+  \centering
+  \includegraphics[scale=0.65]{../diagrams/future-works-data-centric-map-reduce}
+  \caption{Data-Centric approach using MapReduce to search across different
+  database shards}
+  \label{fig:future-works-data-centric-map-reduce}
+\end{figure}
+
+In addition to the infrastructure, the use techniques for scheduling data
+gathering from the SF-BEAMS sensor networks can be used to decrease the data
+load into the system. While the period of times used by RTC seems
+to cover the basic requirements of data, the data storage may be
+seen as overloaded with repeated data. Since environmental conditions
+might not change drastically over time, the use of data scheduling and
+data clustering approaches can be used. Considering that SF-BEAMS
+contains a cluster of sensors, the data load could be decreased by using
+techniques such as clustering the data in the data sink before they are
+persisted \cite{sn-time-series, sn-schedule-minimal-aggregation-time}. Since
+the infrastructure of the SF-BEAMS sensor network is a one-hop start design,
+the knowledge about the data can only be achieved in the data sink. One
+suggestion is the implementation of a DSP Data Clustering component that is
+responsible for the data clustering before they are inserted into the
+database. Similarly, the use of schedulers on the sensors could also be used
+to decrease the amount of data to be sent to the data sink, as seen at the
+survey \cite{sn-scheduling-survey}.
+
+Finally, taking into account the use of the database system to complement
+NetBEAMS execution, the design of event-based applications could add different
+functionalities to the management capabilities of the sensor networks. For
+instance, the collected data from a YSI device \cite{YSI-Sonde} carries the
+information regarding the battery life of the device, seen at the collected
+attribute ``observation.Battery''. A new DSP Component responsible for
+observing these attributes could define the threshold of the event of
+low-battery. With an event-based application in mind, events regarding the
+SF-BEAMS network could be better managed through NetBEAMS where scheduled
+sites visits would depend on such events to happen and, thus, avoiding
+unnecessary operational costs.

### 2009-12-12 docs/marcello-dissertation/thesis/chapter10-appendix.tex
Reviews of Grammatical errors of Chapter 9.
Chapter 3 updated minor things.

THESIS CANDIDATE!!!!!!!

git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@699 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 
-This chapter includes references to the implementation Data Data Persistence
+This chapter includes references to the implementation Data Persistence
 Implementation covering its design and architecture. Also, some of the
 implementation details of the DSP Data Persistence component designed in
-chapter 6 are presented, showing the generated experiment artifacts. While
+Chapter 6 are presented, showing the generated experiment artifacts. While
 this section does not include all the related source-codes, the remainder can
@@
 
-As described in the previous chapter, the DSP Data Persistence Component was
-implemented using the Java Programming language, on top of the OSGi Framework
-API. This section describes the setup process of the implementation, as well
-as the artifacts use during the implementation, which will be listed in the
-appendix section.
+The DSP Data Persistence component was implemented using the Java Programming
+language, on top of the OSGi Framework. This section describes the setup
+process of the implementation, as well as the artifacts use during the
+implementation, which will be listed in the appendix section.
 
@@
   \ref{file:dsp-build.xml}) and other Eclipse-related artifacts. The other
-  major directories are also listed:
+  major directories are also listed;
   \item \textbf{META-INF}: this directory contains the descriptor file for
-  the OSGi Framework, this this case the MANIFEST.MF;
+  the OSGi Framework, in this case, the MANIFEST.MF;
   \item \textbf{src}: the main directory structure for the source-code
@@
   \centering
-  \includegraphics[scale=0.5]{../diagrams/dsp-data-persistence-dir-checkedout}
+  \includegraphics[scale=0.65]{../diagrams/dsp-data-persistence-dir-checkedout}
   \caption{The DSP Data Persistence Directory Structure}
@@
 \subsection{OSGi Deployment Process}
+\label{sec:dsp-component-osgi-deployment}
 
@@
 modularity in terms resources-limited environments such mobile devices and
-vehicles, but it was first widely deployed on Eclipe IDE\footnote{Integrated
-Development Environment}, because it promotes the use of reusable
-loosely-coupled modules using the Java Programming Language \cite{java}. The
-most basic layer of the OSGi Framework depicted in Figure
-\ref{fig:layering-osgi} are as follows:
+vehicles, but it was first widely deployed on Eclipse IDE\footnote{Integrated
+Development Environment}. It promotes reusable loosely-coupled modules to be
+integrated using the Java Programming Language \cite{java}. The most basic
+layers of the OSGi Framework depicted in Figure \ref{fig:layering-osgi} are as
+follows:
 
@@
   level of a Java Class managed by the OSGi Framework;
-  \item \textbf{Service Layer}: resposible for the interoperability between 2
+  \item \textbf{Service Layer}: responsible for the interoperability between two
   or more bundles, enabling the bundles to register services offered by
@@
   \item \textbf{Execution Layer}: executes the bundles, managing and changing
-  their life cycle through the bundle execution.
+  their lifecycle through the bundle execution.
 \end{itemize}
@@
   \centering
-  \includegraphics[scale=0.5]{../diagrams/layering-osgi}
+  \includegraphics[scale=0.65]{../diagrams/layering-osgi}
   \caption{The OSGi Framework Layers}
@@
 Producer-Consumer paradigm described in the previous section. Since the DSP
-Components are essentially osgi bundles, the interoperability between
-them are given in the implementation and reuse existing bundles in the
-OSGi Framework and the DSP Platform, described by an artifact called
-MANIFEST.MF as seen in Listing \ref{file:OSGi-manifest}. In general, an
-OSGi bundle must provide specifications that describes the module to be
-published into the OSGi Framework. In this way, the main properties of
-the OSGi MANIFEST.MF artifact can used by the DSP Data Persistence can
-be summarized as follows:
+Components are essentially OSGi bundles, the interoperability between
+them is given by the OSGi Framework, described by an artifact called
+MANIFEST.MF (Listing \ref{file:OSGi-manifest}). In general, an OSGi bundle must
+provide specifications that describe the module to be published into the OSGi
+Framework. In this way, the main properties of the OSGi MANIFEST.MF artifact
+can used by the DSP Data Persistence can be summarized as follows:
 
@@
   ``DataPersistenceActivator'' provides the activation mechanisms for the
-  services provided by the DSP Data Persistence Component;
+  services of the DSP Data Persistence Component;
   \item \textbf{Bundle-ClassPath}: the necessary Java Jars list needed to run
@@
   Persistence Component Bundle. These Java Packages must be provided by other
-  DSP components, forcing the execution to be dependent on those package.
-  They are provided through a ``Export-Package'' section in other OSGi
-  bundles and, as it is shown in Listing \ref{file:OSGi-manifest}, they
-  are from different packages.
+  DSP components, forcing the execution to be dependent on those packages.
+  They are provided through a section called ``Export-Package'', in other
+  OSGi bundles and as it is shown in Listing \ref{file:OSGi-manifest}, they
+  are can be from different packages.
 \end{itemize}
@@
 Once the OSGi bundle is installed into the OSGi Platform, it will be managed by
-the OSGi Execution layer and change the bundle state according to its lify
-cycle, as described in the previous chapter. Therefore, the packaging of the
-osgi into source-code is done using a JAR\footnote{Java Archival Repository}
-artifact specification \cite{java-tutorial}. In this way, the DSP Data
-Platform osgi into bundle JAR is created by the Apache ANT build script
-\cite{apache-ant} in Listing \ref{file:dsp-build.xml}. The task 
-``dsp-data-persistence.all'' compiles all the source-code created from the
-design of the previous chapter and packages everything into the artifact
-``DSPDataPersistence-x.x.jar'', where x.x is the version of the bundle. Figure
-\ref{fig:dsp-runtime-dir} shows the DSP Data Persistence file under the
-directory ``\textbf{RUNTIME-DIR}/deployment/''.
+the OSGi Execution layer and change the bundle state according to its
+lifecycle. Therefore, the packaging of the OSGi source-code is done using a
+JAR\footnote{Java Archival Repository} artifact specification
+\cite{java-tutorial}. The DSP Platform is bundled into a JAR created by the
+Apache ANT build script \cite{apache-ant}, as shown in Listing
+\ref{file:dsp-build.xml}. The task ``dsp-data-persistence.all'' compiles all
+the source-code created from the design of Chapter 6 and packages
+everything into an artifact name ``DSPDataPersistence-x.x.jar'', where x.x is
+the version of the bundle. Figure \ref{fig:dsp-runtime-dir} shows the DSP Data
+Persistence file under the directory ``\textbf{RUNTIME-DIR}/deployment/''.
 
@@
   \centering
-  \includegraphics[scale=0.5]{../diagrams/dsp-runtime-dir}
+  \includegraphics[scale=0.65]{../diagrams/dsp-runtime-dir}
   \caption{The DSP Runtime Dir}
@@
 \subsection{DSP Data Persistence into the DSP Platform}
+\label{sec:dsp-persistence-boostrap}
 
-Once the new component was developed, the DSP Data Persistence component was
-deployed by editing two main descriptors, as well as the optional bootstrap
-message, as described in chapter 5:
+When the new component was developed and deployed by editing two main
+configuration descriptors:
 
@@
   \item \textbf{config.xml}: the DSP Data Persistence is added to the DSP
-  Platform adding an entry for the component;
+  Platform by adding an entry for the component's OSGi Bundle artifact. This
+  artifact can be seen in Listing \ref{file:dsp-config.xml};
   \item \textbf{matcher\underline{ }config.xml}: the addition of the rules that
-  filters the messages to the DSP Data Persistence;
+  filters DSP Messages to the DSP Data Persistence component;
   \item \textbf{DSPDataPersistence\underline{ }001.xml}: depicts the bootstrap
@@
 
-The first configuration artifact enables NetBEAMS to initialize contain the
+The first configuration artifact enables NetBEAMS to initialize the
 descriptor of the component, as shown in Listing \ref{file:dsp-config.xml}.
@@
 
-MongoDB was the selected technology described in chapter 5 because of its
+MongoDB was the selected technology described in Chapter 5 because of its
 features. It provides data persistence based on collections of data, stored
 using BSON \cite{bson}, a binary representation the JSON data representation
-format. mongoDB uses the abstraction of programming language to
-represent the basic CRUD operators, as well as the regular functionalities
-provided by any type of database system such as indexing, users management,
-etc. As it's stated in their web site, mongoDB "bridges the gap between
-key/value stores, which are fast and highly scalable, traditional RDBMS
-systems (deep in functionality)".
+format. mongoDB uses the abstraction of programming language to represent the
+basic CRUD operators, as well as the regular functionalities provided by any
+type of database system such as indexing and users management. As it's stated
+in their web site, mongoDB "bridges the gap between key/value stores, which
+are fast and highly scalable, traditional RDBMS systems (deep in
+functionality)".
 
@@
   \item mongoDB implements a document-oriented structure, which is similar to
-  KVP;
+  Key-Value Pairs;
   \item mongoDB is written in C++, and therefore, can is available in any major
@@
   different languages such as Java, Python, Perl and Ruby;
-  \item mongoDB has support to distributed systems properties such as
-  Master-Slave replication, and database partitioning, referred to as Database
+  \item mongoDB has support to distributed systems such as Master-Slave
+  Replication, and horizontal database partitioning, referred to as Database
   Shards.
@@
 
-The artifacts of the mongoDB are located in the thirdparty directory of the
-NetBEAMS resources. However, the build script from the DSP Data Persistence
+The artifacts of the mongoDB are located in the ``thirdparty'' directory of the
+NetBEAMS resources directory. The build script from the DSP Data Persistence
 component can be executed to produce the persistence directory structure for
@@
   \centering
-  \includegraphics[scale=0.5]{../diagrams/dsp-persistence-system-dir}
+  \includegraphics[scale=0.65]{../diagrams/dsp-persistence-system-dir}
   \caption{DSP Data Persistence directory}
@@
 given database and its collections. This first section shows the connection of
-the mongo client to the database netbeams as on Listing
+the mongo client to the database ``netbeams'' as on Listing
 \ref{cmd:mongodb-server-start}. The execution of the operation to count the
@@
 
-An example about retrieving a random document from the collection ``netbeams'',
-just type the ``findOne()'' in the shell shown in Listing
-\ref{cmd:mongo-findone}. The output is the actual instance of a document,
-showing all the collected data in the key ``observation'', and the provenance
-metadata in the keys ``sensor.location'' and ``sensor.time''.
+An example about data retrieving of a random document from the collection
+``netbeams'', ``findOne()'' is documented to provide this support as shown in
+Listing \ref{cmd:mongo-findone}. The output is the actual instance of a
+document, showing all the collected data in the key ``observation'', and the
+provenance metadata in the keys ``sensor.location'' and ``sensor.time''.
 
@@
 
-The query based on attributes can be done using the "dot" notation, as you
-navigate through the JSON documents. Additionally, you can use the functions
-as aggregated on the result of others. The example in listing
-\ref{cmd:mongo-find-list} counts the number of documents with the key
+The query based on attributes can be done using the "dot" notation, as the user
+can navigate through the JSON documents. Additionally, the use of internal
+functions is provided by mongoDB on the result of others. The example in
+Listing \ref{cmd:mongo-find-list} counts the number of documents with the key
 "data.ph" equals to "5.64".
@@
 The following example is the output of the first 2 documents from the same
-previous query using the limit() function, as shown in Figure
-\ref{cmd:mongo-find-limit}.
+previous query using the function ``limit()``, as shown in Figure
+\ref{cmd:mongo-find-limit}. The function ``limit()'' is used to decrease the
+number of items returns by the given number.
 
@@
 The mongoDB server offers different drivers to access the data, as well as the
-REST Web Services.
-
-In order to get the results using the REST Web Services interdace, the request
-can be issued using a web browser. For instance, the request to the documents
-using the HTTP Request Variables to filter by the fields as shown in Listing 
-\ref{cmd:mongo-rest-request}.
+REST Web Services. In order to get the results using the REST Web Services
+interdace, the request can be issued using a web browser. For instance, the
+request to the documents using the HTTP Request Variables to filter by the
+fields as shown in Listing \ref{cmd:mongo-rest-request}.
 
@@
 
-\subsection{Visualizing Data on The Browser}
+\subsection{Visualizing Data on the Browser}
 
 Data visualisation tools for mongoDB is slowly being developed by open-source
-developers. The next picture shows the database "netbeams" and the collection
-"SondeDataContainer  ?" being rendered by futon4mongodb, one of the
-open-source tools developed to visualise mongoDB data. Futon4Mongo is an
-adaptation of another software developed for CouchDB \cite{couchdb}. First, the
-collections list is shown. Figure
-\ref{fig:view-collections-instance-browser-futondb} shows one single
-collection for the YSI data containing 1 million objects.
+developers. Futon4mongodb, one of the open-source tools developed to visualise
+mongoDB data, is an adaptation of another software developed for CouchDB
+\cite{couchdb}. Figure \ref{fig:view-collections-instance-browser-futondb}
+shows one single collection for the YSI data containing of 1 million objects.
 
@@
   \centering
-  \includegraphics[scale=0.5]{../diagrams/view-collections-instance-browser-futondb}
+  \includegraphics[scale=0.6]{../diagrams/view-collections-instance-browser-futondb}
   \caption{Viewing a partial list of data using the Futon for CouchDB/MongoDB}
@@
 
-The collection is composed by instances of documents of representing the YSI
+The collection is composed by instances of documents that represents the YSI
 Sonde type, being indexed by a document ID and the values being the keys
-defined in the previous chapter, as shown in Figure
-\ref{fig:view-collected-data-list-browser-futondb}.
+defined, as shown in Figure \ref{fig:view-collected-data-list-browser-futondb}.
 
-\begin{figure}[h]
+\begin{figure}[!h]
   \centering
-
-  \includegraphics[scale=0.5]{../diagrams/view-collected-data-list-browser-futondb}
+  \includegraphics[scale=0.65]{../diagrams/view-collected-data-list-browser-futondb}
   \caption{Viewing a partial list of data using the Futon4CouchDB/MongoDB}
@@
 
-By clicking in one of the items, one can view the instance of the document, as
-it is shown in Figure \ref{fig:view-collections-instance-browser-futondb}.
+By clicking in one of the items, the document properties are shown for
+navigation purposes as shown in Figure
+\ref{fig:view-collections-instance-browser-futondb}.
 
-\begin{figure}[h]
+\begin{figure}[!h]
   \centering
-  \includegraphics[scale=0.5]{../diagrams/view-collected-data-instance-browser-futondb} 
+  \includegraphics[scale=0.6]{../diagrams/view-collected-data-instance-browser-futondb}
   \caption{Viewing an instance of collected data using the Futon4CouchDB/MongoDB}
@@
 
-mongoDB has an export facility shell called mongoexport. It can export the data
-in JSON format or CSV. One may also write its own export tool in any of the
+mongoDB has an export facility shell called ``mongoexport''. It can export the
+data in JSON format or CSV. One may also write its own export tool in any of the
 languages such as Java, PHP, Python, Perl, Ruby, among others. A list of the
@@
 An example to export the data in CSV format can be seen in listing
-\ref{file:mongodb-export-command}, and an example containing 1 million objects
-can be downloaded at
+\ref{file:mongodb-export-command}, and the resulting example containing, 1
+million objects can be downloaded at
 http://netbeams.googlecode.com/files/experiment-1000000-data-exported-20090913-053538.csv.tar.gz.
@@
 
-During the experiments of this work, the implementation of the export utility
-to OPeNDAP was written in Python, as shown in Listing
-\ref{file:experiment-export-python}. In order to execute the
-command, the python programming language \cite{python} must be installed in the
-system. In this way, some limited documents are exported. The following command
-is an example of the execution of the export utility
-\ref{cmd:mongo-python-export}:
+During the experiments conducted for this work, the implementation of the
+export utility to OPeNDAP was written in Python, as shown in Listing
+\ref{file:experiment-export-python}. In order to execute the command, the
+python programming language \cite{python} must be installed in the system. In
+this way, some limited documents are exported. The following command is an
+example of the execution of the export utility \ref{cmd:mongo-python-export}:
 
@@
 \end{lstlisting}
+
+Complete source-code listing can be reviewed in the next sections.
 
@@
 
-\lstset{language=XML,morecomment=[s]{!--}{--},label=file:dsp-message-serialized-ysi,caption=DSP
+\lstset{language=XML,morecomment=[s]{!--}{--},label=stream:dsp-message-serialized-ysi,caption=DSP
 Message with a YSI sonde data payload}

### 2009-12-13 versions/v2/persistence/data-export-python.py
This revision contains different logs and scripts created for the experiments. They were documented in the Appendix of my MS report.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@700 559edde2-a4d4-44e0-96fc-4035e40d7396

@@ -1,0 +1,85 @@
+# Copyright 2009 Marcello de Sales (marcello.sales@gmail.com).
+# Converts the collected sensor data from the mongoDB to the OPEnDAP format
+
+import pymongo
+import sys
+
+EXPORT_FILE_NAME="exported-by-python"
+
+# Prints the header similar to the OPEnDAP format in the given file f, 
+# with the first line identifying the dataset, and the second one the
+# columns of the measurements.
+def printHeader(f):
+	f.write('Dataset: ntb_YSI_PUF.dsp\n')
+	f.write('YSI_REALTIME_CSV.Month,YSI_REALTIME_CSV.Day,YSI_REALTIME_CSV.Year,YSI_REALTIME_CSV.Hour,YSI_REALTIME_CSV.Min,YSI_REALTIME_CSV.Sec,YSI_REALTIME_CSV.WaterTemperature,YSI_REALTIME_CSV.SpecificConductivity,YSI_REALTIME_CSV.Conductivity,YSI_REALTIME_CSV.Resistivity,YSI_REALTIME_CSV.Salinity,YSI_REALTIME_CSV.Pressure,YSI_REALTIME_CSV.Depth,YSI_REALTIME_CSV.pH,YSI_REALTIME_CSV.pHmV,YSI_REALTIME_CSV.Turbidity,YSI_REALTIME_CSV.ODOSaturation,YSI_REALTIME_CSV.ODO,YSI_REALTIME_CSV.Battery\n')
+
+# Prints given ysiData returned from mongoDB, in a dict format, in to the file
+# f, following the format of comma-separated values. 
+def printRows(f, ysiData):
+	f.write(ysiData["time"]["valid"].strftime("%m,%d,%Y,%H,%M,%S,"))
+	f.write(str(ysiData["observation"]["WaterTemperature"]))
+	f.write(",")
+	f.write(str(ysiData["observation"]["SpecificConductivity"]))
+	f.write(",")
+	f.write(str(ysiData["observation"]["Conductivity"]))
+	f.write(",")
+	f.write(str(ysiData["observation"]["Resistivity"]))
+	f.write(",")
+	f.write(str(ysiData["observation"]["Salinity"]))
+	f.write(",")
+	f.write(str(ysiData["observation"]["Pressure"]))
+	f.write(",")
+	f.write(str(ysiData["observation"]["Depth"]))
+	f.write(",")
+	f.write(str(ysiData["observation"]["pH"]))
[…]

### 2009-12-13 versions/v2/persistence/experiment-client-scenarios.js

@@ -1,0 +1,57 @@
+//Select the database NetBEAMS to be used
+use netbeams
+
+//Each collected data from the sersor devices are
+//stored on separate collections. Show all of them 
+print("Show all the collections of collected data (sensor devices' data)")
+show collections
+
+print("Show the statistics of the collection for the collected data for the YSI Sonde device")
+db.SondeDataContainer.stats()
+
+print("Verify some of the values using command abstraction shortcuts")
+db.SondeDataContainer.count()
+db.SondeDataContainer.dataSize()
+
+print("An example of a random Show one random sample from the database using the findOne method
+db.SondeDataContainer.findOne()
+
+//db.SondeDataContainern.find({'$where' => 'this.time.valid.getTime() != this.time.transaction.getTime()'},{:limit => 50}).sort(:time.valid, Mongo::DESCENDING)
+
+//the month variable of date class in javascript is defined as [0-11]. Therefore, 10 is November.
+
+print("Execute Scenario R1) search all the documents in a given date range. Count and Limit retrieval result")
+db.SondeDataContainer.find( {"time.valid" : { $gte:new Date(2009,11,8) , $lt:new Date(2009,11,15) }} ).count()
+db.SondeDataContainer.find( {"time.valid" : { $gte:new Date(2009,11,8) , $lt:new Date(2009,11,15) }} ).limit(3)
+print ("Explain the search over the shards for R1)")
+db.SondeDataContainer.find( {"time.valid" : { $gte:new Date(2009,11,8) , $lt:new Date(2009,11,15) }} ).explain()
+
+print("Execute Scenario R2) search all the documents for a given sensor device by IP address. Count and Limit retrieval result")
+db.SondeDataContainer.find( {"sensor.ip_address":"192.168.0.102"} ).count()
+db.SondeDataContainer.find( {"sensor.ip_address":"192.168.0.102"} ).limit(3)
+print("Explain the search over the shards for R2")
+db.SondeDataContainer.find( {"sensor.ip_address":"192.168.0.102"} ).explain()
+
+print("Execute Scenario R3) search all the documents that has some observation values. Salinity = .01, Water Temperature = 46.47")
+db.SondeDataContainer.find( {"observation.Salinity":0.01, "observation.WaterTemperature":46.47} ).count()
[…]

### 2009-12-13 versions/v2/persistence/setup-experiment-shards.sh

@@ -1,0 +1,34 @@
+#!/bin/bash
+# Using "seq"
+
+echo "############## Setting up shards for experiment..."
+if [ -z "$1" ]; then 
+    echo "Usage: $0 x, where x is the number of shards to be created"
+    exit
+fi
+
+NUMBER_SHARDS=$1
+SHARDS_DIR=shards
+SHARD_DIR_PREFIX=shard
+
+rm -rf $SHARDS_DIR
+mkdir $SHARDS_DIR
+
+mkdir data/$SHARDS_DIR
+for shard_number in $(seq $NUMBER_SHARDS)
+do
+	mkdir data/$SHARDS_DIR/$SHARD_DIR_PREFIX-$shard_number
+	mongod --dbpath data/$SHARDS_DIR/$SHARD_DIR_PREFIX-$shard_number/ --port 2000$shard_number | tee logs/shard-$shard_number.log &
+done
+
+mkdir data/$SHARDS_DIR/config
+mongod --dbpath data/$SHARDS_DIR/config --port 10000 | tee logs/shard-config.log &
+
+ps aux | grep mongod
+
+mongos -vvv --configdb localhost:10000 | tee logs/mongos-cluster-head.log & 
+
+sleep 1
+
+mongo < setup-mongo-shards.js | tee logs/setup-mongo-shards-for-netbeams.log &
+

### 2009-12-13 versions/v2/persistence/setup-mongo-shards.js

@@ -1,12 +1,35 @@
 admin = db.getSisterDB("admin")
 config = db.getSisterDB("config")
 
-admin.runCommand({addshard:"192.168.1.10:10000"})
-admin.runCommand({addshard:"192.168.1.11:10001"})
+admin.runCommand( { addshard: "192.168.1.2:20001" } )
+admin.runCommand( { addshard: "192.168.1.2:20002" } )
+admin.runCommand( { addshard: "localhost:20001", allowLocal: true } )
 
-admin.runCommand({listshards:1})
+admin.runCommand( { listshards:1 } )
 
-admin.runCommand({enablesharding:"netbeams"})
-admin.runCommand({shardcollection:"netbeams.SondeDataContainer", key:{latitude:1}})
+admin.runCommand( { enablesharding: "netbeams" } )
+admin.runCommand( { shardcollection: "netbeams.SondeDataContainer", key: { "observation.pH" : 1} } )
+
+netbeams = db.getSisterDB("netbeams")
+netbeams.SondeDataContainer.ensureIndex( { "message_id":1 } )
+netbeams.SondeDataContainer.ensureIndex( { "sensor.ip_address":1 } )
+netbeams.SondeDataContainer.ensureIndex( { "sensor.location.latitude":1 } )
+netbeams.SondeDataContainer.ensureIndex( { "sensor.location.longitude":1 } )
+netbeams.SondeDataContainer.ensureIndex( { "time.valid":1 } )
+netbeams.SondeDataContainer.ensureIndex( { "time.transaction":1 } )
+netbeams.SondeDataContainer.ensureIndex( { "observation.WaterTemperature":1 } )
+netbeams.SondeDataContainer.ensureIndex( { "observation.SpecificConductivity":1 } )
+netbeams.SondeDataContainer.ensureIndex( { "observation.Conductivity":1 } )
+netbeams.SondeDataContainer.ensureIndex( { "observation.Resistivity":1 } )
+netbeams.SondeDataContainer.ensureIndex( { "observation.Salinity":1 } )
+netbeams.SondeDataContainer.ensureIndex( { "observation.Pressure":1 } )
+netbeams.SondeDataContainer.ensureIndex( { "observation.Depth":1 } )
+netbeams.SondeDataContainer.ensureIndex( { "observation.pH":1 } )
+netbeams.SondeDataContainer.ensureIndex( { "observation.pHmV":1 } )
[…]

### 2009-12-13 docs/marcello-dissertation/thesis/chapter10-appendix.tex
Reviews on the Appendix section
- Adding references to the execution logs and developed scripts
- Adding more description about the infrastructure setup
- Updating reference on chapter 2  (wrong)
- Updating the reference to the setup scripts on chapter 8.

git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@701 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 
+The description of a distributed system can be seen next section.
+
 \section{Experiments Execution}
@@
 previous section.
+
+\subsection{Setting Up the Single and the Distributed Cluster Servers}
+\label{sec:experiments-setup-single-distributed}
+mongoDB can be used in two different ways: as a single server, using the single
+process ``mongod'', or as a distributed cluster where each cluster node runs on
+``mongod'' managed by a cluster head ``mongos''.
+
+Before starting the main cluster head, all cluster nodes must be started.
+Listing \ref{file:shard-setup-192.168.1.2.log} shows the manual setup process
+performed in the Mac OS X 64bits Host. It starts two different shards, with
+their own respective directory, and starts the process ``mongod'' for each of
+them, forwarding the execution output to their respective log artifacts. Note
+that the IP address is capture in the beginning of the process because it will
+be used while setting up the cluster head.
+
+In order to automate the manual process, different automation scripts were
+written using the Linux shell bash execution, as well as the ones used to setup
+mongoDB using Javascript, the language used by mongoDB shell client. First, the
+scripts to setup the mongoDB distributed cluster is shown in Listing
+\ref{file:setup-mongo-shards.js}. It enables shard support for the different
+shards that must be already running, and then enable it on the collection
+``SondeDataContainer'' for the database ``netbeams''. As it is shown in the
+script, the declaration of the shards are executed in the ``admin'' database,
+while shards are in the database ``config'', since it is the main cluster
+metadata server.
+
+The first script created was the automated shell bash script to setup local
+shards directories and processes, as well as the cluster head metadata server
+(config) and the start of the cluster head ``mongos''. Each local shard can be
+candidates to store the collected data, depending on the decisions made by the
+mongos cluster head. Listing \ref{file:setup-experiment-shards.sh} expects to be
+executed by passing the number of local shards as a parameter to create the
+local directories and start the processes described. Listing
+\ref{file:shards-and-mongos-cluster-head-setup-execution.log} shows the output
+of the exeuction of the setup script, incorporating the output of the Listing
+\ref{file:setup-mongo-shards.js}. Note that the same IP address, shown in the
+setup of the cluster nodes, is also shown in the list of the shards.
+
+\subsection{Running the Experiments}
+
+After setting up the infrastructure, the main experiment script can be executed
+in the client-side. Considering the distributed server approach, the process
+``mongos'' must be running in the server host. The infrastructure for the
+execution of the experiments used Sun Microsystem's VirtualBox to run different
+operating systems at the same time. The cluster head ``mongos'' was running on
+a Ubuntu Linux 9.04 64 bits together with the cluster metadata server
+``mongod'' referred to as ``config'' server. The cluster nodes were running on
+the Mac OS X 64 bits on the host server, running two different shards. Finally,
+the execution of the experiments used Ubuntu Linux 8.04 32 Bits.
+
+As developed for the infrastructure setup, the execution of the experiments
+were performed using automation scripts written in different languages.
+First, the experiments for the ``Create" operations was written in Java
+as shown in Listing \ref{file:experiment-dsp-java-executor}. Similarly,
+remaining operations ``Retrieve, Update and Delete" were written in
+Javascript, as shown in Listing \ref{file:experiment-scenarios}. Then, the
+execution of this Java class is performed by the main experiment shell script
+displayed in Listing \ref{file:experiment-setup-executor}. It requires the
+number of randomly created YSI Sonde data to be executed, as shown in Listing
+\ref{cmd:run-persistence-experiment}.
 
@@
 \end{lstlisting}
+
+The first experiment conducted was running the main experiment script of Listing
+\ref{file:experiment-setup-executor}, as one of the execution runs shown in
+Listing \ref{file:experiment-run.log}. The total number of inserted documents
+measured every 20 seconds is displayed. Then, the execution of the use cases
+was performed, as shown in Listing \ref{file:URD-scenarios-execution.log}.
+Additional execution logs are shown in Section
+\ref{sec:experiment-execution-logs}.
 
@@
 
+\newpage
+
 \subsection{Exporting Data to Spreadsheets}
@@
 
+\newpage
+
 \section{Source-Code and Implemented Artifacts}
@@
 {../../../../netbeams/versions/v2/apps/xml/samples/DSPDataPersistence_001.xml}
-
-%TODO: add bootstrap message in case it is implemented
 
@@
 
-\lstinputlisting[language=Bash,label=file:experiment-setup-executor,caption=Experiment
-Main Execution Script]{../../../../netbeams/versions/v2/persistence/run-persistence-experiment}
+\lstinputlisting[language=Bash,label=file:shard-setup-192.168.1.2.log,caption=Remote
+Shards Setup Execution Log]{../../../../netbeams/versions/v2/persistence/logs/shard-setup-192.168.1.2.log}
 
-\lstinputlisting[language=Python,label=file:experiment-setup-mongoshards,caption=Experiment
-Main Execution Script]{../../../../netbeams/versions/v2/persistence/setup-mongo-shards.js}
+\lstinputlisting[language=Bash,label=file:setup-experiment-shards.sh,caption=Remote
+Shards Cluster Head and Metadata Setup Bash Script]{../../../../netbeams/versions/v2/persistence/setup-experiment-shards.sh}
 
-\lstinputlisting[language=Bash,label=file:setup-os-shards-step1,caption=Experiment
-Main Execution Script]{../../../../netbeams/versions/v2/persistence/setup-os-shards-step1}
+\lstinputlisting[label=file:setup-mongo-shards.js,caption=Shards
+Setup in Javascript]{../../../../netbeams/versions/v2/persistence/setup-mongo-shards.js}
 
-\lstinputlisting[language=Bash,label=file:setup-os-shards-step2,caption=Experiment
-Main Execution Script]{../../../../netbeams/versions/v2/persistence/setup-os-shards-step2}
+\lstinputlisting[language=Bash,label=file:experiment-setup-executor,caption=Main Client
+Experiment Execution Bash Script]{../../../../netbeams/versions/v2/persistence/run-persistence-experiment}
+
+\lstinputlisting[label=file:experiment-scenarios,caption=Experiment
+Scenarios Implementation in Javascript]{../../../../netbeams/versions/v2/persistence/experiment-client-scenarios.js}
 
 \lstinputlisting[language=Python,label=file:experiment-export-python,caption=Data
-Export to from mongoDB to OPeNDAP format.]{../../../../netbeams/versions/v2/persistence/data-export-python.py}
-
-\lstinputlisting[language=Python,label=file:experiment-scenarios,caption=Experiment
-Scenarios Implementation]{../../../../netbeams/versions/v2/persistence/experiment-client-scenarios.js}
+Export to from mongoDB to OPeNDAP Format Python Script]{../../../../netbeams/versions/v2/persistence/data-export-python.py}
 
 \lstinputlisting[language=Java,label=file:main-ysi-data-handler,caption=The
-DSP YSI Sonde Data Handler]{../../../../netbeams/versions/v2/apps/xml/jaxb/sonde/java-src/org/netbeams/dsp/ysi/SondeDataType.java}
+DSP YSI Sonde Data Java Handler]{../../../../netbeams/versions/v2/apps/xml/jaxb/sonde/java-src/org/netbeams/dsp/ysi/SondeDataType.java}
 
 \lstinputlisting[language=Java,label=file:random-ysi-data-generator,caption=Random
-Test Data Generator]{../../../../netbeams/versions/v2/apps/osgi-bundles/dsp/DSPSondeProducer/src/org/netbeams/dsp/ysi/SondeTestData.java}
+Test Data Java Generator]{../../../../netbeams/versions/v2/apps/osgi-bundles/dsp/DSPSondeProducer/src/org/netbeams/dsp/ysi/SondeTestData.java}
 
 \lstinputlisting[language=Java,label=file:experiment-dsp-java-executor,caption=Inserting
-Random Data to MongoDB]{../../../../netbeams/versions/v2/apps/osgi-bundles/dsp/DSPDataPersistence/src/org/netbeams/dsp/persistence/controller/DSPMessageToMongoDBExperiment.java}
+Random Data to MongoDB Java Experiment]{../../../../netbeams/versions/v2/apps/osgi-bundles/dsp/DSPDataPersistence/src/org/netbeams/dsp/persistence/controller/DSPMessageToMongoDBExperiment.java}
 
-\subsection{Experiment Execution Output}
+\subsection{Experiment Execution Output Logs}
+\label{sec:experiment-execution-logs}
+
+\lstinputlisting[language=Bash,label=file:shards-and-mongos-cluster-head-setup-execution.log,caption=Execution
+of Shards Setup Log]{../../../../netbeams/versions/v2/persistence/logs/shards-and-mongos-cluster-head-setup-execution.log}
+
+\lstinputlisting[language=Bash,label=file:experiment-run.log,caption=Main Client
+Experiment Execution Log]{../../../../netbeams/versions/v2/persistence/experiment-run.log}
+
+\lstinputlisting[language=Bash,label=file:URD-scenarios-execution.log,caption=RUD
+Scenarios Execution Log]{../../../../netbeams/versions/v2/persistence/logs/URD-scenarios-execution.log}
+
+\lstinputlisting[language=Bash,label=file:mongos-shards-cluster-head-metadata-config-server-192.168.1.10:10000.log,caption=Remote
+Shards Cluster Head Execution Log]{../../../../netbeams/versions/v2/persistence/logs/mongos-shards-cluster-head-metadata-config-server-192.168.1.10:10000.log}
+
+\lstinputlisting[language=Bash,label=file:mongos-shards-cluster-head-192.168.1.10.log,caption=Remote
+Shards Cluster Head Execution]{../../../../netbeams/versions/v2/persistence/logs/mongos-shards-cluster-head-192.168.1.10.log}
+
+\lstinputlisting[language=Bash,label=file:shard-192.168.1.2:20002.log,caption=Remote
+Shard Execution Log]{../../../../netbeams/versions/v2/persistence/logs/shard-192.168.1.2:20002.log}
+
+\lstinputlisting[label=file:experiment-export-python-results,caption=Data
+Export to from mongoDB to OPeNDAP format.]{../../../../netbeams/versions/v2/persistence/export-mongodb-to-sfbeams.csv}
 
@@
 \end{lstlisting}
-
-\lstinputlisting[label=file:experiment-export-python-results,caption=Data
-Export to from mongoDB to OPeNDAP format.]{../../../../netbeams/versions/v2/persistence/export-mongodb-to-sfbeams.csv}

### 2009-12-13 docs/marcello-dissertation/thesis/chapter2-state-of-the-art.tex
Reviews on the Appendix section
- Adding references to the execution logs and developed scripts
- Adding more description about the infrastructure setup
- Updating reference on chapter 2  (wrong)
- Updating the reference to the setup scripts on chapter 8.

git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@701 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 it helps researchers correlate different measurements based on exact locations
-of the data. For instance, \ref{sn-provenance} depicts an example where two
+of the data. For instance, \cite{sn-provenance} depicts an example where two
 different sensor networks covering the same physical area: one covering the

### 2009-12-13 docs/marcello-dissertation/thesis/chapter8-experimental-results-correct-behavior-and-performance.tex
Reviews on the Appendix section
- Adding references to the execution logs and developed scripts
- Adding more description about the infrastructure setup
- Updating reference on chapter 2  (wrong)
- Updating the reference to the setup scripts on chapter 8.

git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@701 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 used for the single server only uses the process ``mongod'' to be started,
-while the clustered setup must be enabled in the database, as shown in
-Listing \ref{file:experiment-setup-mongoshards}, Listings
-\ref{file:setup-os-shards-step1} and \ref{file:setup-os-shards-step2}.
+while the clustered setup must be enabled in the database, as in Section
+\ref{sec:experiments-setup-single-distributed}.
 
@@
   \centering
-  \includegraphics[scale=0.5]{../diagrams/taxonomy-data-model-modified}
+  \includegraphics[scale=0.65]{../diagrams/taxonomy-data-model-modified}
   \caption{Data Model Taxonomy}
@@
 temporary caches for objects whose internal state are similar to others in the
-Java Virtual Machine.
+JVM. Also, operations that involve writing in disk, such as during the bulk
+``Create'' operations, may lead to different database performance. First, the
+experiment seemed to stop inserting objects in one shard while it ran for over
+2 minutes without inserting any documents. The reason was that the partition
+was ``/home'', completely full, leading to the problem. The mongos cluster
+header did not realize that it was the time to switch to another shard.
+Similarly, any global ``Retreave'' operation on a cluster blocks all the
+shards. For this reason, dividing the write from the read load was necessary in
+order to decouple these two operations.
 

### 2009-12-14 docs/marcello-dissertation/thesis/chapter6-data-persistence-component-design-and-architecture.tex
This revision includes the revisions of the Feedback of Chapter 6, with some changes of the Appendix (references).
- Reference to the architecture documented after I removed from my thesis reference. The page brings my name on the top of the page as the author, so just put my name for that reference. It is an online documentation, and therefore, I think it is a better contribution for the DSP future users. This documentation might also be prepared as a Technical paper suggested by Prof. Puder. In this way, this reference can be changed later.

- I had just missed these 2 references about the c

@@
 
-The mongoDB database system was selected, as stated in chapter 5, to store 
+The mongoDB database system was selected, as stated in Chapter 5, to store 
 the collected data from the SF-BEAMS sensor network. NetBEAMS is based on a 
@@
 consumes \cite{netbeams2009}. In this way, a DSP Data Component was designed
-and implemented to provide the integration between NetBEAM and the mongoDB 
+and implemented to provide the integration between NetBEAMS and the mongoDB 
 database. This integration is based on the Data Sensor Platform (DSP), the
 kernel software stack provided by NetBEAMS, as documented in
-\cite{netbeams-dsp-architecture}. Finally, details about the implementation of
-this design are located in the Appendix Section
-\ref{sec:dsp-data-persistence-implementation}.
+\cite{netbeams-dsp-architecture}.
+
+This chapter follows the steps taken in Software Engineering guidelines to
+analyze and design a software component using Object-Oriented Analysis and 
+Design \cite{oop} using the techniques from the Agile and XP Scrum
+\cite{agile-scrum} processes. First, Section \ref{sec:business-process-analysis} gives an
+overview of the problem by analyzing a basic scenario of the data collection process used in NetBEAMS.
+Then, the Section \ref{sec:requirements-specs} describes the requirements for
+the design of the new component, followed by the specifications of the data
+model used for the mongoDB database in Section
+\ref{sec:dsp-persistence-data-model}. Lastly, the high-level architecture
+of Section \ref{sec:system-architecture} gives the foundation of the
+DSP Data Persistence component design of Section \ref{sec:dsp-data-comp-design}.
+
+Details about the implementation of this design are located in the
+Appendix Section \ref{sec:dsp-data-persistence-implementation}.
 
@@
 Messages into the body of a DSP Messages Container in order to be serialized
-into XML (shown in Listing \ref{stream:dsp-message-serialized-ysi}) and
-transferred to the DSP Server host, a centralized server that receives all the
-collected data embodied into each of the DSP Messages.
+into XML (shown in Listing \ref{stream:dsp-message-serialized-ysi} on page
+\pageref{stream:dsp-message-serialized-ysi}) and transferred to the DSP Server
+host, a centralized server that receives all the collected data embodied into each of the DSP Messages.
 
@@
 \section{Requirements Specification}
-
+\label{sec:requirements-specs}
 This section outlines the functionalities and constraints that the developed
@@
   OPeNDAP data format, so that I can collaborate with other research groups
-  that uses this data format.";
+  that use this data format.";
   \item \textbf{Access the data through API, Programming Languages}: "As a
@@
 
-Driven by the requirements defined, the foundation of the integration between
-the DSP Data Persistence component and the mongoDB database system is the
-specification of the Keys that comprise a set of observations collected from
-the sensor devices.
+Driven by the requirements defined, the definition of the keys of the key-value
+pairs model, used to describe the collected data from the sensor devices and
+their associated metadata, is the foundation for the integration between the
+DSP Data Persistence component and the mongoDB database. Next section details
+the data model design that specifies those keys.
 
@@
 \cite{json} as its main data representation. MongoDB does not require the creation 
-of a database schema, rather the definition of key descriptions that comprises
-a document on the database organized into separate collections of relating
-documents. In order to describe the keys of a document, the specifications
-detail the use of Data Provenance as guidelines to the description of the
-collected data. Therefore, the key-value pairs of any instance of the
+of a database schema, rather the definition of keys that comprise a document on
+the database organized into separate collections of relating documents. In
+order to describe the keys of a document, the guidelines provided by the Data
+Provenance taxonomy was used. The key-value pairs of any instance of the
 collected data must include the following key properties:
@@
 
-After analyzing the DSP Platform and the structure of the DSP Messages in
-\cite{netbeams-dsp-architecture}, the values of the collected data must be
-categorized into two different sets of data: the ``Key Segment'' and the
-``Value Segment''. The key segment provides the necessary metadata 
-used to describe the collected data. An instance of the design of the 
-keys is shown in Listing \ref{file:mongodb-ysi-data-format}:
+After analyzing the DSP Message Structure, as well as the YSI Sonde Data
+Handler of the DSP Platform architecture documentation
+\cite{netbeams-dsp-architecture}, the design of the document developed to
+identify both the data and metadata of the collected data is organized as
+follows:
 
@@
     \end{itemize} 
+    \item \textbf{observation}: carries the observations of a given sensor
+    device. For example, the observation key of for the YSI device
+    \cite{YSI-Sonde} may contains all the subsets of the properties defined
+    for the device, such as battery, temperature and salinity, among others.
 \end{itemize}
 
-An associated ``Value Segment'' for the aforementioned ``Key Segment'' carries
-the observations of a given sensor device. The key \textbf{observation} stores
-the values of the collected data. For instance, the observation key of
-for the YSI device \cite{YSI-Sonde} may contains all the subsets of 
-the properties defined for the device, such as battery, temperature and
-salinity, among others. An example of such use of the segments is shown in
-Listing \ref{file:mongodb-ysi-data-format}.
+MongoDB uses JSON data structure \cite{json} to represent a document, as the
+example shown in Figure \ref{fig:kvp-json-viewer} on page
+\pageref{fig:kvp-json-viewer}. Any number of keys can be in the root branch of
+the tree, as well as in sub branches of the document structure. The values of
+the tree are concentrated in the leaves of the tree. Listing
+\ref{file:mongodb-ysi-data-format}, on page
+\pageref{file:mongodb-ysi-data-format}, shows the instance of this design on
+mongoDB database. Based on this foundation, the design of the DSP Data
+Persistence component evolved from a top-down approach to meet this data model
+design.
 
-Based on this foundation, the design of the DSP Data Persistence component
-evolved from a top-down approach.
+\begin{figure}[!h]
+  \centering
+  \includegraphics[scale=0.65]{../diagrams/kvp-json-viewer}
+  \caption{Data Design using JSON Data Structure}
+  \label{fig:kvp-json-viewer}
+\end{figure}
+
+It is important to note that mongoDB does not use any structure language, like
+the use of the SQL in relational databases, to access the persisted data. In
+order to access the values of the documents, the ``dot notation''
+\cite{db-mongo-data-access} is used to reach each of the tree nodes. For
+example, the use of ``observation.WaterTemperature'' and
+``sensor.location.latitude'' give the values of the data ``water temperature'' and
+the metadata ``latitude'' of the collected data, respectively.
 
 \section{High-Level System Architecture}
+\label{sec:system-architecture}
 
@@
 in the previous section. This process guarantees that the collected data is
-inserted to a running mongoDB database system and gives covers the use cases
+inserted to a running mongoDB database system and covers the use cases
 defined. Figure \ref{fig:NetBEAMS-Persistence-Server-Node-Components}
@@
 
-\section{DSP Data Persistence: OSGi-DSP Bundle Design}
+\section{DSP Data Persistence Component Design}
+\label{sec:dsp-data-comp-design}
 
@@
 Diagram in Figure \ref{fig:DSP-Data-Persistence-Packages-Dependency}, a new
-DSP Component has dependency on three different packages: the OSGi Framework
+DSP Component has dependencies on three different packages: the OSGi Framework
 and the DSP Platform responsible to integrate a new component into the
@@
 maintain the component managed by the OSGi Framework and register the class
-DSPDataPersistece as a service to the OSGi environment. Therefore, the class
+DSPDataPersistence as a service to the OSGi environment. Therefore, the class
 DSPDataPersistenceActivator is the main communication interface between the 
@@
 the services provided by the DSPDataPersistence component, as this behavior is
-inherited from the OSGi class BundleActivator from OSGi Framework. Following
+inherited from the OSGi class BundleActivator from the OSGi Framework. Following
 the specification of the DSP Architecture \cite{netbeams-dsp-architecture},
@@
   DSP Data Persistence component from the configuration artifact config.xml as
-  seen in Listing \ref{file:dsp-config.xml};
+  seen in Listing \ref{file:dsp-config.xml} on page
+  \pageref{file:dsp-config.xml};
   \item After being selected based on the configuration priority, the 
@@
 bootstrap process through the use of an optional DSP Update Message, as shown
-in Listing \ref{file:dsp-data-persistence-bootstrap.xml}. The initialization of
-the DSP Component Activator and the DSP Component registration are also shown
-in Figure \ref{fig:From-OSGi-Framework-to-DSP-Data-PersistenceActivator-Sequence-Diagram}.
+in Listing \ref{file:dsp-data-persistence-bootstrap.xml} on page
+\pageref{file:dsp-data-persistence-bootstrap.xml}. The initialization of the
+DSP Component Activator and the DSP Component registration are also shown in
+Figure
+\ref{fig:From-OSGi-Framework-to-DSP-Data-PersistenceActivator-Sequence-Diagram}.
 
@@
 that filters any remote measurement message received by running DSP Server is
-used, as shown in Listing \ref{file:dsp-matcher-config.xml}. The rule defined
-in line 63 express that all messages created by any producer on any DSP Host,
-either client or sever, has as delivery target the DSP Data Persistence
-component. In this way, when DSP Measurement Messages are delivered, the DSP
-Broker delivers the DSP Message to the DSP Data Persistence Component.
+used, as shown in Listing \ref{file:dsp-matcher-config.xml} on page
+\pageref{file:dsp-matcher-config.xml}. The rule defined in line 63 express
+that all messages created by any producer on any DSP Host, either client or
+sever, has as delivery target the DSP Data Persistence component. In this way,
+when DSP Measurement Messages are delivered, the DSP Broker delivers the DSP
+Message to the DSP Data Persistence Component.
 
@@
   seconds. 
-  Listing \ref{file:dsp-data-persistence-bootstrap.xml} shows an
-  implementation of a bootstrap message for the DSP Data Persistence;
+  Listing \ref{file:dsp-data-persistence-bootstrap.xml}, on page
+  \pageref{file:dsp-data-persistence-bootstrap.xml}, shows an implementation of
+  a bootstrap message for the DSP Data Persistence;
   \item Any DSP Measure Message can be persisted by the DSP Data Persistence
@@
   to decrease I/O on the database, when the DSP Broker delivers a DSP
-  MeasureMessage to this component, as depicted by the gate "create
-  PersistentMessageUnit Gate". This last gate describes the simple procedure
-  to add the message into the transient persistent layer. Listing
-  \ref{stream:dsp-message-serialized-ysi}.
+  MeasureMessage to this component, as depicted by the UML Gate element
+  \cite{uml-seq-gate} "create PersistentMessageUnit Gate". This last Gate
+  describes the simple procedure to add the message into the transient
+  persistent layer, as shown in Listing
+  \ref{stream:dsp-message-serialized-ysi} on page
+  \pageref{stream:dsp-message-serialized-ysi}.
 \end{itemize}
@@
 At this point, the DSP Data Persistence component maintains a set of DSP
-Messages in a local cache during the specific ``flush rate'' specified during
-the bootstrap message. When the class DSPDataFlusher finds any number of
-instances of PersistentMessageUnit in cache, it starts the process of flushing
-them into the mongoDB database, described in the following section.
+Messages in a local cache during the specific ``flush rate'' specified in
+the bootstrap message parameter. When the class DSPDataFlusher finds any number
+of instances of PersistentMessageUnit in cache, it starts the process of
+flushing them into the mongoDB database, described in the following section.
 
@@
 \begin{itemize}
-  \item \textbf{TIMED\underline{ }WAITING}: the initial state of a Java thread
+  \item \textbf{Waiting}: the initial state of a Java thread
   after its initialization, waiting to be on the ``runnable'' state. This
@@
   message of the DSP Data Persistence component;
-  \item \textbf{RUNNABLE}: the class DSPDataFlusher contacts the class
+  \item \textbf{Running}: the class DSPDataFlusher contacts the class
   TransientPersistenceLayer, in order to verify if there are any instance of
@@
 UML Class Diagram \ref{fig:DSP-Data-Persistence-Mongo-Classes}. It keeps references
-to all types of sensor devices such as the SondeDataContainer and the
-MouseActionsContainer. Furthermore, it is responsible to implement the CRUD
-methods to modify the mongoDB database state. The class Mongo is used to acquire
+to all types of sensor devices. Furthermore, it is responsible to implement the
+CRUD methods to modify the mongoDB database state. The class Mongo is used to acquire
 an instance of the connection with mongoDB server and provide Factory Methods 
@@
 
-In order to build the document key and value segments, as described in
-section \ref{sec:dsp-persistence-data-model}, the reference to the DSP message
-is used, as the value key is built by extracting the DSP Message Content from
-the body of the DSP Message. Similarly, the key is built by gathering
-information from the DSP Message and the PersistenceMessageUnit class. Each of
-the properties collected from the YSI Sonde sensor device is collected and
-associated with the keys defined for the Value Segment.
+In order to build the mongoDB document described in section
+\ref{sec:dsp-persistence-data-model}, the reference to the DSP Message is
+used to extracted the metadata, while the reference to the DSP Message Content 
+from the body of the DSP Message is used to extract the collected data from
+sensors.
 
@@
   \includegraphics[scale=0.6]{../diagrams/From-Creating-Value-Segment-Sequence}
-  \caption{UML Sequence diagram showing the creation of the value document segment}
+  \caption{UML Sequence diagram showing the creation of the keys for the
+  collected data}
   \label{fig:From-Creating-Value-Segment-Sequence}
@@
 
-Once the Value Segment is created, the Key Segment is prepared as described in
-the UML Sequence Diagram in Figure
+After the keys for the collected data are created, the keys for the metadata
+are captured as shown in the UML Sequence Diagram in Figure
 \ref{fig:From-Creating-Key-Segment-Sequence}.
@@
   \includegraphics[scale=0.65]{../diagrams/From-Creating-Key-Segment-Sequence}
-  \caption{UML Sequence diagram showing the creation of the key document segment}
+  \caption{UML Sequence diagram showing the creation of the keys for the
+  metadata}
   \label{fig:From-Creating-Key-Segment-Sequence}
@@
 
-To Summarize, the DSP Data Persistence Component is added into the DSP Platform
+To summarize, the DSP Data Persistence Component is added into the DSP Platform
 by introducing a new component that provides service to the transfer any
@@
 
-\section{DSP Data Persistence Deployment}
+\subsection{DSP Data Persistence Deployment}
 
@@
 of a DSP Component over an OSGi Framework. For the development of this work,
-the Knopflerfish OSGi Container was used. The details of implementation are described
-in section \ref{sec:dsp-data-persistence-implementation}. The deployment of the 
-mongoDB can be accomplished in different approaches, summarized as follows:
+the Knopflerfish OSGi Container \cite{knopflerfish} was used. The details of the
+implementation are described in Section
+\ref{sec:dsp-data-persistence-implementation}. The deployment of the mongoDB
+can be accomplished in different approaches, summarized as follows:
 
@@
 the document. For NetBEAMS collected data, a good candidate for partitioning is
-the time when the data was created, and separate each shard in groups of month
+the time when the data was created, and separate each shard in groups of months
 of the year.
@@
 helped to understand the requirements of the system. The data model design
-followed the specifications defined in chapter 5. Each of the layers of the
+followed the specifications defined in Chapter 5. Each of the layers of the
 system was described using UML Diagrams, having partial details regarding the

### 2009-12-14 docs/marcello-dissertation/thesis/chapter10-appendix.tex
Final Revision After Prof. Puder's Audio Feedback. Chapters 7 and Chapter 8 were included together. Chapter 10 also added as the appendix, instead of chapter.
* Chapter 7 

- The mouse experiment was used as demo purpose and to show the DSP use of the persistence. The data used was the for the workload used the YSI Generator from Teresa, which generated the SAME values each time… In this way, it was NOT random… I added the proper random functions in everything, including the parameter passed to the method to generate a given number of generated values. 

- The workload uses the amount of data

@@
 them is given by the OSGi Framework, described by an artifact called
-MANIFEST.MF (Listing \ref{file:OSGi-manifest}). In general, an OSGi bundle must
-provide specifications that describe the module to be published into the OSGi
+MANIFEST.MF (Listing \ref{file:OSGi-manifest} on page
+\pageref{file:OSGi-manifest}). In general, an OSGi
+bundle must provide specifications that describe the module to be published into the OSGi
 Framework. In this way, the main properties of the OSGi MANIFEST.MF artifact
@@
   They are provided through a section called ``Export-Package'', in other
-  OSGi bundles and as it is shown in Listing \ref{file:OSGi-manifest}, they
-  are can be from different packages.
+  OSGi bundles and as it is shown in Listing \ref{file:OSGi-manifest}, page
+  \pageref{file:OSGi-manifest}, they are can be from different packages.
 \end{itemize}
@@
   Platform by adding an entry for the component's OSGi Bundle artifact. This
-  artifact can be seen in Listing \ref{file:dsp-config.xml};
+  artifact can be seen in Listing \ref{file:dsp-config.xml}, page
+  \pageref{file:dsp-config.xml};
   \item \textbf{matcher\underline{ }config.xml}: the addition of the rules that
@@
 The first configuration artifact enables NetBEAMS to initialize the
-descriptor of the component, as shown in Listing \ref{file:dsp-config.xml}.
-The component is added with the highest priority, since it contains
-dependencies to other DSP Components.
+descriptor of the component, as shown in Listing \ref{file:dsp-config.xml},
+page \pageref{file:dsp-config.xml}. The component is added with the highest
+priority, since it contains dependencies to other DSP Components.
 
@@
 for each of the shard servers. An example of accessing mongoDB using a shell
-client is shown in Listing \ref{cmd:mongodb-client-start}.
+client is shown in Listing \ref{cmd:mongodb-client-start} page
+\pageref{cmd:mongodb-client-start}.
 
@@
 existing number of documents in the mongoDB collection ``netbeams'' is shown in
-Listing \ref{cmd:mongodb-count}. It highlights the query for the collections
-available \cite{mongodb} in the netbeams database.
+Listing \ref{cmd:mongodb-count}, page \pageref{cmd:mongodb-count}. It
+highlights the query for the collections available in the netbeams database.
 
@@
 ``netbeams'', ``findOne()'' is documented to provide this support as shown in
-Listing \ref{cmd:mongo-findone}. The output is the actual instance of a
-document, showing all the collected data in the key ``observation'', and the
-provenance metadata in the keys ``sensor.location'' and ``sensor.time''.
+Listing \ref{cmd:mongo-findone}, page \pageref{cmd:mongo-findone}. The output
+is the actual instance of a document, showing all the collected data in the
+key ``observation'', and the provenance metadata in the keys
+``sensor.location'' and ``sensor.time''.
 
@@
 functions is provided by mongoDB on the result of others. The example in
-Listing \ref{cmd:mongo-find-list} counts the number of documents with the key
-"data.ph" equals to "5.64".
+Listing \ref{cmd:mongo-find-list}, page \pageref{cmd:mongo-find-list} counts
+the number of documents with the key "data.ph" equals to "5.64".
 
@@
 request to the documents using the HTTP Request Variables to filter by the
-fields as shown in Listing \ref{cmd:mongo-rest-request}.
+fields as shown in Listing \ref{cmd:mongo-rest-request}, page
+\pageref{cmd:mongo-rest-request}.
 

### 2009-12-14 docs/marcello-dissertation/thesis/chapter9-conclusions-and-future-works.tex
Final Revision After Prof. Puder's Audio Feedback. Chapters 7 and Chapter 8 were included together. Chapter 10 also added as the appendix, instead of chapter.
* Chapter 7 

- The mouse experiment was used as demo purpose and to show the DSP use of the persistence. The data used was the for the workload used the YSI Generator from Teresa, which generated the SAME values each time… In this way, it was NOT random… I added the proper random functions in everything, including the parameter passed to the method to generate a given number of generated values. 

- The workload uses the amount of data

@@
 
-There were significant remarks covered in this work, including the
-importance that sensor networks are becoming ubiquitous, having their produced
-data available online and other types of media. In order to provide access to
-the collected data, one must first assess the different types of
+There were significant trends covered in this work, including the importance
+that sensor networks are becoming ubiquitous, having their produced data
+available online and other types of media. In the scope of Environmental Sensor
+Networks, the collected data is the most important parameter to real-world
+applications based on sensor networks. For example, the data collected in the
+water may determine if a river is polluted, while the data collected during a
+heavy snowfall can indicate hazard alerts the population. However, in order to
+provide access to the collected data, one must first assess the different types of
 infrastructural characteristics of the studied sensor networks. The selection
@@
 obtain better functionalities when compared to other solutions. Similarly,
-improved work suggests that the abstraction of programming language should be
+this work suggests that the abstraction of programming language should be
 used with research groups whose primary expertise is not in Computer Science
@@
 with the uncertainty of any properties collected from any sensor specification
-is the most compeling reason to use the Key-value pair data model for collected
+is the most compelling reason to use the Key-value pair data model for collected
 data from sensor devices.
@@
 implementation of such functionality is based in Figure
-\ref{fig:future-works-data-centric-map-reduce}, which shows an architectural
-idea of the deployment of a Data-Centric storage approach. When searching for
-specific document attributes, the process is sent in parallel to each of the
-database shards, and the cluster head collects the result. After calculating
-the final result, mongoDB server returns the relevant result to the user.
+\ref{fig:future-works-data-centric-map-reduce} (\cite{map-reduce-notes}), which
+shows an architectural idea of the deployment of a Data-Centric storage approach. One scenario of
+data retrieval is when a given processing in the search results are
+required to be executed in parallel. The task of querying broken down into
+two phases: the definition of the reduced problem as the parts, and the
+mapping of the final result calculated with the response from each part. If
+``Task" is considered to be a query (arrow 1) tor the server, then the
+execution of the same query occurs in all different partitions (arrow 2), or
+mongoDB shards. After processing the request, each node returns the result of
+the query to the server (arrow 3), whose responsibility is to process each
+individual result and return the final Result to the client application (arrow
+4).
 
@@
   \includegraphics[scale=0.65]{../diagrams/future-works-data-centric-map-reduce}
-  \caption{Data-Centric approach using MapReduce to search across different
-  database shards}
+  \caption{Data-Centric approach using MapReduce (\cite{map-reduce-notes})}
   \label{fig:future-works-data-centric-map-reduce}
@@
 
-In addition to the infrastructure, the use techniques for scheduling data
-gathering from the SF-BEAMS sensor networks can be used to decrease the data
-load into the system. While the period of times used by RTC seems
-to cover the basic requirements of data, the data storage may be
-seen as overloaded with repeated data. Since environmental conditions
-might not change drastically over time, the use of data scheduling and
-data clustering approaches can be used. Considering that SF-BEAMS
-contains a cluster of sensors, the data load could be decreased by using
-techniques such as clustering the data in the data sink before they are
-persisted \cite{sn-time-series, sn-schedule-minimal-aggregation-time}. Since
-the infrastructure of the SF-BEAMS sensor network is a one-hop start design,
-the knowledge about the data can only be achieved in the data sink. One
-suggestion is the implementation of a DSP Data Clustering component that is
-responsible for the data clustering before they are inserted into the
-database. Similarly, the use of schedulers on the sensors could also be used
-to decrease the amount of data to be sent to the data sink, as seen at the
-survey \cite{sn-scheduling-survey}.
+In addition to the infrastructure, the use  scheduling techniques for
+data gathering in sensor networks can help the SF-BEAMS sensor networks
+decrease the data load into the system. While the period of times used by RTC
+seems to cover the basic requirements of data, the data storage may be seen as
+overloaded with repeated data. Since environmental conditions might not change
+drastically over time, the use of data scheduling and data clustering
+approaches can be used. Considering that SF-BEAMS contains a cluster of
+sensors, the data load could be decreased by using techniques such as
+clustering the data in the data sink before they are persisted
+\cite{sn-time-series, sn-schedule-minimal-aggregation-time}. Since the
+infrastructure of the SF-BEAMS sensor network is a one-hop start design, the
+knowledge about the data can only be achieved in the data sink. One suggestion
+is the implementation of a DSP Data Clustering component that is responsible
+for the data clustering before they are inserted into the database. Similarly,
+the use of schedulers on the sensors could also be used to decrease the amount
+of data to be sent to the data sink, as seen at the survey
+\cite{sn-scheduling-survey}.
 

### 2009-12-14 docs/marcello-dissertation/thesis/chapter10-appendix.tex
Reviews of misspelling and references to pages, Section, etc.
- Chapter 7: Fixes on "Chapter" to chapter
- Chapter 8: Subversion Conflict line removal, and other fixes
- Appendix: Removing "on page x" for the image on the same page.

git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@706 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 The first experiment conducted was running the main experiment script of Listing
-\ref{file:experiment-setup-executor}, as one of the execution runs shown in
-Listing \ref{file:experiment-run.log}. The total number of inserted documents
+\ref{file:experiment-setup-executor} (page
+\pageref{file:experiment-setup-executor}), as one of the execution runs shows
+the results in Listing \ref{file:experiment-run.log}, page
+\pageref{file:experiment-run.log}. The total number of inserted documents
 measured every 20 seconds is displayed. Then, the execution of the use cases
-was performed, as shown in Listing \ref{file:URD-scenarios-execution.log}.
-Additional execution logs are shown in Section
-\ref{sec:experiment-execution-logs}.
+was performed, as shown in Listing \ref{file:URD-scenarios-execution.log} page
+\pageref{file:URD-scenarios-execution.log}. Additional execution logs are
+shown in Section \ref{sec:experiment-execution-logs}.
 
@@
 the mongo client to the database ``netbeams'' as on Listing
-\ref{cmd:mongodb-server-start}. The execution of the operation to count the
-existing number of documents in the mongoDB collection ``netbeams'' is shown in
-Listing \ref{cmd:mongodb-count}, page \pageref{cmd:mongodb-count}. It
-highlights the query for the collections available in the netbeams database.
+\ref{cmd:mongodb-server-start} on page \pageref{cmd:mongodb-server-start}. The
+execution of the operation to count the existing number of documents in the
+mongoDB collection ``netbeams'' is shown in Listing \ref{cmd:mongodb-count},
+(page \pageref{cmd:mongodb-count}). It highlights the query for the collections
+available in the netbeams database.
 

### 2009-12-14 docs/marcello-dissertation/thesis/chapter8-experimental-results-correct-behavior-and-performance.tex
Reviews of misspelling and references to pages, Section, etc.
- Chapter 7: Fixes on "Chapter" to chapter
- Chapter 8: Subversion Conflict line removal, and other fixes
- Appendix: Removing "on page x" for the image on the same page.

git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@706 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 
-This chapter details the experiment setup designed to evaluate the
-proposed implementation, showing the measurement results collected from the
-simulation log and based on the user experience while reusing the collected
-data. The pertinent discussion of this work is divided up into different
-sections, each of them analyzing different aspects related to data persistence
-for NetBEAMS, among others.
+This chapter details the experiment setup designed to evaluate the proposed
+implementation, showing the measurement results collected from the simulation
+log and based on the user experience while reusing the collected data. The
+pertinent discussion of this work is divided up into different sections, each
+of them analyzing different aspects related to data persistence for NetBEAMS.
+Section \ref{sec:exp-setup} introduces the experiments setup and methods used
+for the evaluation such as the artifacts used, the definition of the workload
+to be used during the evaluation of the planned scenarios of evaluation. Then,
+Section \ref{sec:exp-measurements} shows the measurement results of the
+experiment detailing the operations performed by each of the scenarios and
+user experience, to finally proceed with the final analysis and discussions.
 
 \section{Experiment Setup}
+\label{sec:exp-setup}
 
@@
 
-In order to evaluate the DSP Data Persistence component, experiments were
-conducted using the DSP Mouse Actions component, a software sensor simulator
-that captures the actions performed by the mouse pointer.
-
 \subsection{Experiments Artifacts}
@@
 the services provided by the class DSPMongoCRUDService (Listing
-\ref{file:dsp-mongo-service}). It is responsible for the connection handling
-with mongoDB, as well as the insertion operation from NetBEAMS. As described
-\cite{netbeams-dsp-architecture}, the current version of NetBEAMS supports the
-data collection for the YSI Sonde device \cite{YSI-Sonde}, among others used
-for demo purposes. After evaluating the database with the YSI Sonde data
-handler, the use of the demo component DSP Mouse Actions was used. It captures
-the mouse pointer activities performed over a window, for example, mouse clicks
-and mouse grag over coordinates x and y.
+\ref{file:dsp-mongo-service}, page \pageref{file:dsp-mongo-service}). It is
+responsible for the connection handling with mongoDB, as well as the insertion
+operation from NetBEAMS. As described \cite{netbeams-dsp-architecture}, the
+current version of NetBEAMS supports the data collection for the YSI Sonde
+device \cite{YSI-Sonde}, among others used for demo purposes.
 
@@
 existing Java class TestSondeData (Listing
-\ref{file:random-ysi-data-generator}), was refactored to provide a random
+\ref{file:random-ysi-data-generator}, page
+\pageref{file:random-ysi-data-generator}), was refactored to provide a random
 instance generator for the YSI Data handler implemented in the Java class
-SondeDataType (Listing \ref{file:main-ysi-data-handler}). When the Insert
-function is executed, the created instances are transferred from the JVM to
-the mongoDB, and stored in the collection ``SondeDataContainer''. Specific
-details about the execution of mongoDB are depicted in Section
-\ref{sec:mongodb-deployment}. Finally, the implementation of the experiment
-in the Java class DSPMessageToMongoDBExperiment (Listing
-\ref{file:experiment-dsp-java-executor}) was responsible for managing the
+SondeDataType (Listing \ref{file:main-ysi-data-handler} on page
+\pageref{file:main-ysi-data-handler}). When the Insert function is executed,
+the created instances are transferred from the JVM to the mongoDB, and stored
+in the collection ``SondeDataContainer''. Specific details about the execution
+of mongoDB are depicted in Section \ref{sec:mongodb-deployment}. Finally, the
+implementation of the experiment in the Java class
+DSPMessageToMongoDBExperiment (Listing
+\ref{file:experiment-dsp-java-executor}, page
+\pageref{file:experiment-dsp-java-executor}) was responsible for managing the
 creation and insertion of any workload composed by instances of the class
@@
 the inserted workload were implemented using Javascript, as shown in Listing
-\ref{file:experiment-scenarios}, respectively by the function calls ``find'',
-``update'' and ``remove'' over the collection ``SondeDataContainer''.
-Similarly, in order to implement a customizable export system that converts
-from mongoDB into the format OpenDAP, a simpler prototype script was developed
-in Python, as shown in Listing \ref{file:experiment-export-python}. As a result
-of the execution of this script, the exported artifact has the same format as
-described in Listing \ref{file:rtc-ysi-opendap}.
+\ref{file:experiment-scenarios} on page \pageref{file:experiment-scenarios},
+respectively by the function calls ``find'', ``update'' and ``remove'' over
+the collection ``SondeDataContainer''. Similarly, in order to implement a
+customizable export system that converts from mongoDB into the format OpenDAP,
+a simpler prototype script was developed in Python, as shown in Listing
+\ref{file:experiment-export-python} on page
+\pageref{file:experiment-export-python}. As a result of the execution of this
+script, the exported artifact has the same format as described in Listing
+\ref{file:rtc-ysi-opendap}, page \pageref{file:rtc-ysi-opendap}.
 
@@
   tabular format sequence used in the files distributed by SF-BEAMS using the 
-  the OPeNDAP format, as in Listing \ref{file:rtc-ysi-opendap}.
+  the OPeNDAP format, as in Listing \ref{file:rtc-ysi-opendap} on page
+  \pageref{file:rtc-ysi-opendap}.
 \end{itemize}
@@
 The first set of measurements collected were the ones regarding the use case
-R1, shown in Listing \ref{file:dsp-mongo-service}, defined as the insertion of
-instances of the collected data from the sensor devices using Java. The
-document follows the specification of the keys designed in Section
-\ref{sec:dsp-persistence-data-model}. The construction of the keys
-and respective values were done using a Hash-like object ``BasicDBObject''
-such as the ones shown in the method ``buildKeySegment''. Similarly, the
-creation of database indexes was performed using the same approach, as shown in
-the Java method ``setupMetadataIndexes''. Then, the insertion of the
-documents was performed on the service method call
-``insertPersistentUnitMessageContents'', which delimits the bulk creation using
-the transaction management mechanism over the collection. In this way,
-different insert times were collected from the logs of the different
-experiments as summarized on Table 7.1.
+R1, shown in Listing \ref{file:dsp-mongo-service} page
+\pageref{file:dsp-mongo-service}, defined as the insertion of instances of the
+collected data from the sensor devices using Java. The document follows the
+specification of the keys designed in Section
+\ref{sec:dsp-persistence-data-model}. The construction of the keys and
+respective values were done using a Hash-like object ``BasicDBObject'' such as
+the ones shown in the method ``buildKeySegment''. Similarly, the creation of
+database indexes was performed using the same approach, as shown in the Java
+method ``setupMetadataIndexes''. Then, the insertion of the documents was
+performed on the service method call ``insertPersistentUnitMessageContents'',
+which delimits the bulk creation using the transaction management mechanism
+over the collection. In this way, different insert times were collected from
+the logs of the different experiments as summarized on Table 7.1.
 
@@
 
-In general, the user experience of mongoDB is related the methods used to
+In general, the user experience of mongoDB is related to the methods used to
 access and modify the state of the database using a different approach when
@@
 language abstraction of objects and method calls, the transition from the use
-of a relational data model is very smooth. The creation of data was the first
+of a relational data model is very easy. The creation of data was the first
 contact with mongoDB in terms of data creation in Java. Later, different
@@
 operation over the collection of documents using regular the abstraction
-programming language, as the shown in Listing \ref{file:experiment-scenarios}.
-The CRUD operations are available to the user and can be used at any time to
-any database collection. Aggregation operations such as to count the returned
-result set is also provided via programming languages abstraction.
+programming language, as the shown in Listing \ref{file:experiment-scenarios}
+on page \pageref{file:experiment-scenarios}. The CRUD operations are available
+to the user and can be used at any time to any database collection.
+Aggregation operations such as to count the returned result set is also
+provided via programming languages abstraction.
 
@@
 native support to comma-delimited values. As shown in Listing
-\ref{file:mongodb-export-command}, the actual execution describe how mongo is
-querying the data from the database. As a result, the total number of documents
-exported is shown in on line 8. In addition to the regular command that
-exports the entire database, the user can provide a query string during the
-execution of the command in order to filter the dataset returned and,
-consequently, the number of documents exported. Similar to the use of the
-default export capabilities of mongoDB, a script written in Python to export
-data from mongoDB was written as shown in Listing
-\ref{file:experiment-export-python}. It uses the package ``pymongo'' to create
-a comma-delimited file with the same header and line format as shown in Listing
-\ref{file:experiment-export-python-results}. The results can be compared with
-the example the output of the HTTP Response body of Listing
-\ref{file:rtc-ysi-opendap}, requested directly to the OPeNDAP server of RTC.
-Therefore, the experiment is capable of reproducing theat the OPeNDAP can.
+\ref{file:mongodb-export-command}, page \pageref{file:mongodb-export-command},
+the actual execution describe how mongo is querying the data from the
+database. As a result, the total number of documents exported is shown in on
+line 8. In addition to the regular command that exports the entire database,
+the user can provide a query string during the execution of the command in
+order to filter the dataset returned and, consequently, the number of
+documents exported. Similar to the use of the default export capabilities of
+mongoDB, a script written in Python to export data from mongoDB was written as
+shown in Listing \ref{file:experiment-export-python}, page
+\pageref{file:experiment-export-python}. It uses the package ``pymongo'' to
+create a comma-delimited file with the same header and line format as shown in
+Listing \ref{file:experiment-export-python-results}, page
+\pageref{file:experiment-export-python-results}. The results can be compared
+with the example the output of the HTTP Response body of Listing
+\ref{file:rtc-ysi-opendap} on page \pageref{file:rtc-ysi-opendap}, requested
+directly to the OPeNDAP server of RTC. Therefore, the experiment is capable of
+reproducing the targeted OPeNDAP format used by the RTC server.
 
@@
 the CRUD operations, the access to the data was also performed using the REST
-Web Services API, as shown in Listing \ref{cmd:mongo-rest-request}. As one can
-see, the IP address and the default port number of the mongoDB REST middleware
-are provided. Then, the REST operation uses the name of the database and
-the collection to correlate with the ones in the database system. Thus,
-querying the collections from mongoDB can be performed using the HTTP Request
-Parameter variables, by adding the prefix ``filter\underline{ }'' before the
-name of the key of the document. The given example shows the query performed
-in the collected data for the key ``observation.Conductivity'' with the value
-``104.5''. Similarly, the HTTP Request variable ``limit'' is used to reduce
-the number of documents returned by the query. The HTTP Response from mongoDB
-returns the result using the JSON format, similar to the ones printed in the
-command-shell.
+Web Services API, as shown in Listing \ref{cmd:mongo-rest-request} on page
+\pageref{cmd:mongo-rest-request}. As one can see, the IP address and the
+default port number of the mongoDB REST middleware are provided. Then, the
+REST operation uses the name of the database and the collection to correlate
+with the ones in the database system. Thus, querying the collections from
+mongoDB can be performed using the HTTP Request Parameter variables, by adding
+the prefix ``filter\underline{ }'' before the name of the key of the document.
+The given example shows the query performed in the collected data for the key
+``observation.Conductivity'' with the value ``104.5''. Similarly, the HTTP
+Request variable ``limit'' is used to reduce the number of documents returned
+by the query. The HTTP Response from mongoDB returns the result using the JSON
+format, similar to the ones printed in the command-shell.
 
@@
 \ref{fig:view-collections-instance-browser-futondb}.
-
-The actual execution of the solution was performed using the demo Mouse
-Actions, which consists of a software sensor that perceives any mouse action in
-a window area. The experiment consisted of moving the mouse around a white
-window, where each action is captured and temporarily saved until a threshold is
-reached, making the component to send the messages to the persistence layer, so
-that it can be saved on mongoDB. Figure \ref{fig:experiment-mouseactions-dsp}
-shows a screenshot of the moment when the message is sent to mongoDB, being
-verified through the HTTP REST Interface using a web browser
-\ref{sec:dsp-mongodb-rest-ws}.
-
-\begin{figure}[h]
-  \centering
-  \includegraphics[scale=0.4]{../diagrams/experiment-mouseactions-dsp}
-  \caption{Experiment showing the mongoDB persisting messages from the DSP
-  Mouse Actions Demo}
-  \label{fig:experiment-mouseactions-dsp}
-\end{figure}
 
@@
 
-Any database system covers the purpose of the sensor data for NetBEAMS, which
-is Data Archival, along with the type of storage for the data as an External
-Storage. After selecting mongoDB, the development of the persistence layer met
-the requirements of archiving data for different types of sensor devices from
-the DSP Platform such as the YSI Data Handler and the Mouse Actions demo data.
-As it is shown in Section \ref{sec:exp-measurements}, the collected data is
-archived in files using the BSON format \cite{bson}, a binary representation
-used to store instances of the document described in Listing
-\ref{file:mongodb-ysi-data-format}. Along with the purpose of Data Archival,
-the use of mongoDB also met the specification of External Storage, since the
-data persisted in the file system, as well as Data-Centric for running mongoDB
-with different  distributed mongo shards (cluster nodes).
+After selecting mongoDB as the best candidate for the persistence
+infrastructure for NetBEAMS, the development of the persistence layer met the
+requirements of archiving data for different types of sensor devices from the
+DSP Platform such as the YSI Data Handler. As it is shown in Section
+\ref{sec:exp-measurements}, the collected data is archived in files using the
+BSON format \cite{bson}, a binary representation used to store instances of
+the document described in Listing \ref{file:mongodb-ysi-data-format} on page
+\pageref{file:mongodb-ysi-data-format}. Along with the purpose of Data
+Archival, the use of mongoDB also met the specification of External Storage,
+since the data persisted in the file system, as well as Data-Centric for
+running mongoDB with different distributed mongo shards (cluster nodes).
 
@@
 identity of the data is given by the key ``message\underline{ }id'', in a way
-that it uniquely identifies the data produced by the sensor identified by the
-key ``sensor.ip\underline{ }address''. Thus, the message can be tracked
+that it uniquely identifies the data produced. Thus, the message can be tracked
 back to the sensor that produced the data for different purposes such as
@@
 component, as shown in the parameter ``YSI\underline{ }DESIRED\underline{
-}PROPERTIES'' of Listing \ref{file:dsp-data-persistence-bootstrap.xml}.
+}PROPERTIES'' of Listing \ref{file:dsp-data-persistence-bootstrap.xml}, page
+\pageref{file:dsp-data-persistence-bootstrap.xml}.
 
@@
 the instances of any known data collected from NetBEAMS and saves it using the
-document pattern as shown in Listing \ref{file:mongodb-ysi-data-format}. Any
-programming language used was capable of accessing the same collected data by
-their own.
+document pattern as shown in Listing \ref{file:mongodb-ysi-data-format} on page
+\pageref{file:mongodb-ysi-data-format}. Any programming language used was
+capable of accessing the same collected data by their own.
 
 In the end, the implementation of the data persistence for NetBEAMS was
-developed in both Single and Distributed Systems. The use of regular single
+developed in both Centralized and Distributed Systems. The use of regular single
 server execution provided enough computational power for a compatible volume of
 data when compared to the one produced from the SF-BEAMS sensor network. The
-implementation suggested by this work fulfill the needs of a persistence system
+implementation suggested by this work fulfills the needs of a persistence system
 for a Small volume of data being created at the same time in a Single Server.
@@
 The performance to insert data into the database system using the Java driver
-is documented by mongoDB as the fastest approach, providing persistence
-thousands of documents per second. This capacity is more than necessary for
+is documented by mongoDB as the fastest approach, providing persistence for
+thousands of documents per second. This capacity is more than adequate for
 NetBEAMS, which can use the Single System environment. Improvements over the
@@
 The development of the scenarios were clearly written in different ways, such
-as the ones defined in Listing \ref{file:experiment-scenarios}, which
-implements the use cases for data Retrieval, Update and Deletion operations. The
-main advantage of this approach is that users do not use the SQL language to extract
-data from the database, but use method calls defined for each the
-abovementioned operations. For example, based on the description of the
-collected data by keys and values shown in Listing
-\ref{file:mongodb-ysi-data-format}, different use cases can evaluate documents
-in this structure. Usually researchers may search the collected data by a
-given range of dates, for their regular activities of data analysis. The
+as the ones defined in Listing \ref{file:experiment-scenarios} on page
+\pageref{file:experiment-scenarios}, which implements the use cases for data
+Retrieval, Update and Deletion operations. The main advantage of this approach
+is that users do not use the SQL language to extract data from the database,
+but use method calls defined for each the abovementioned operations. For
+example, based on the description of the collected data by keys and values
+shown in Listing \ref{file:mongodb-ysi-data-format}, page
+\pageref{file:mongodb-ysi-data-format}, different use cases can evaluate
+documents in this structure. Usually researchers may search the collected data
+by a given range of dates, for their regular activities of data analysis. The
 scenario implemented by R1 in the above cited listing shows it clearly.
@@
 First, the use of the export capabilities to a common-delimited values was
-performed \ref{file:mongodb-export-command}, taking proximally 3.3 minutes to
-export one year's worth of data for 5 YSI devices without indexing. Second, the
+performed (Listing \ref{file:mongodb-export-command}, page
+\pageref{file:mongodb-export-command}), taking proximally 3.3 minutes to export
+one year's worth of data for 5 YSI devices without indexing. Second, the
 implementation of a custom export utility using Python, as shown in Listing
-\ref{file:experiment-export-python} enabled the creation of the comma-delimited
-files with the same format as the ones used by OPeNDAP. The comparison can be
-done by analyzing the output of the OPeNDAP request from SF-BEAMS on Listing
-\ref{file:rtc-ysi-opendap}, and the output provided from the script written in
-Python, show in Listing \ref{file:experiment-export-python-results}.
-Furthermore, considering research teams with skills in Web Services, an easy
-ubiquitous integration with mongoDB can be achieved by using the REST Web
-Services APIs provided by mongoDB, still in alpha version. A practical application of
-this API was the created Apache Futon4mongoDB web application, used to give
-user friendlier access to the collected data as shown in 
-\ref{fig:view-collections-instance-browser-futondb}.
+\ref{file:experiment-export-python}, page
+\pageref{file:experiment-export-python}, enabled the creation of the
+comma-delimited files with the same format as the ones used by OPeNDAP. The
+comparison can be done by analyzing the output of the OPeNDAP request from
+SF-BEAMS on Listing \ref{file:rtc-ysi-opendap}, page
+\pageref{file:rtc-ysi-opendap}, and the output provided from the script
+written in Python, show in Listing \ref{file:experiment-export-python-results}
+page \pageref{file:experiment-export-python-results}. Furthermore, considering
+research teams with skills in Web Services, an easy ubiquitous integration
+with mongoDB can be achieved by using the REST Web Services APIs provided by
+mongoDB, still in alpha version. A practical application of this API was the
+created Apache Futon4mongoDB web application, used to give user friendlier
+access to the collected data as shown in Figure
+\ref{fig:view-collections-instance-browser-futondb}, on page
+\pageref{fig:view-collections-instance-browser-futondb}.
 
@@
   \includegraphics[scale=0.65]{../diagrams/taxonomy-data-model-modified}
-  \caption{Data Model Taxonomy}
+  \caption{Updated Data Model Taxonomy, adding the KVP Data Model}
   \label{fig:taxonomy-data-model-modified}
@@
 
-This work can be used to evaluate data persisten in sensor networks in
+This work can be used to evaluate data persistence in sensor networks in
 different ways. The use of the data persistence taxonomies through this work
-and guided both the evaluation and implementation of the data model using
-mongoDB. In order to use the guidelines proposed by this work, one may first
-determine the properties of the sensor network in question. Considering that
-sensor devices produce streams of non-described data, the use of key-value
-pairs data model did not require changes throughout different application
-layers.
+guided both the evaluation and implementation of the data model using mongoDB.
+In order to use the guidelines proposed by this work, one may first determine
+the properties of the sensor network in question. Considering that sensor
+devices produce streams of non-described data, the use of key-value pairs data
+model did not require changes throughout different application layers.
 
@@
 the implementation of the data model potentially uses the Structured-Query
-Language (SQL) for the problem of data plumbing, what results in a
+Language (SQL) for the problem of data plumbing, which results in a
 tightly-coupled implementation. On the other hand, mongoDB eliminates the use
@@
 selection. Since relational database systems are the ``traditional'' way to
-provide data persitence, the investigations started by creating a relational
+provide data persistence, the investigations started by creating a relational
 model described in Chapter 5. Then, the technique using KVP data models in
@@
 Similarly, any global ``Retreave'' operation on a cluster blocks all the
-shards. For this reason, dividing the write from the read load was necessary in
-order to decouple these two operations.
+shards. For this reason, dividing the write from the read load was necessary
+in order to decouple these two operation.
 
 Also related to the difficulties of this work, mongoDB is not yet stable, and
-therefore, harder to implement. The main production version of mongoDB is
-is on version 1.1.4 at this time, while the shards support is still in alpha
+therefore, harder to implement. The main production version of mongoDB
+is on version 1.2 at this time, while the shards support is still in alpha
 version. For instance, different work-arounds were performed to have shards
@@
 experiment results, the measurement results and finally the discussion about
-the solution in general. The following section brings the conclusions and
+the solution in general. The following chapter brings the conclusions and
 future works for this report.

### 2009-12-14 docs/marcello-dissertation/thesis/chapter9-conclusions-and-future-works.tex
Reviews of misspelling and references to pages, Section, etc.
- Chapter 7: Fixes on "Chapter" to chapter
- Chapter 8: Subversion Conflict line removal, and other fixes
- Appendix: Removing "on page x" for the image on the same page.

git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@706 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 water may determine if a river is polluted, while the data collected during a
-heavy snowfall can indicate hazard alerts the population. However, in order to
-provide access to the collected data, one must first assess the different types of
+heavy snowfall can indicate hazard alerts to the population. However, in order
+to provide access to the collected data, one must first assess the different types of
 infrastructural characteristics of the studied sensor networks. The selection
@@
 \ref{fig:future-works-data-centric-map-reduce} (\cite{map-reduce-notes}), which
-shows an architectural idea of the deployment of a Data-Centric storage approach. One scenario of
-data retrieval is when a given processing in the search results are
+shows an architectural idea of the deployment of a Data-Centric storage
+approach. One scenario of data retrieval is to process search results
 required to be executed in parallel. The task of querying broken down into
 two phases: the definition of the reduced problem as the parts, and the
-mapping of the final result calculated with the response from each part. If
-``Task" is considered to be a query (arrow 1) tor the server, then the
+map of the final result calculated with the response from each part. If
+``Task" is considered to be a query (arrow 1) to the server, then the
 execution of the same query occurs in all different partitions (arrow 2), or
@@
 the query to the server (arrow 3), whose responsibility is to process each
-individual result and return the final Result to the client application (arrow
+individual result and return the final one to the client application (arrow
 4).

### 2009-12-15 docs/marcello-dissertation/thesis/preamble.tex
Revision replacing "thesis" with "report" or "culminating experience".
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@708 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
 
-\singlespace{A thesis presented to the faculty of\\
+\singlespace{A report presented to the faculty of\\
 San Francisco State University\\
@@
 I certify that I have read {\it \mytitle} by \myname, and that in my opinion
-this work meets the criteria for approving a thesis submitted in partial
-fulfillment of the requirements for the degree: Master of Science in Computer
+this work meets the criteria for approving a culminating experience submitted in
+partial fulfillment of the requirements for the degree: Master of Science in Computer
 Science at San Francisco State University.
@@
 \onehalfspacing{\noindent
-Sensor Networks are becoming more and more important to different science
-communities because of what they produce: the raw data for different domains.
-In order to make use the collected data, researchers may have to dissect the
+Sensor Networks are becoming important to different scientific and industrial
+communities because of what they produce: the raw data of diverse domains. In
+order to make use of the collected data, researchers may have to dissect the
 characteristics of the sensor network in question, regarding different
 properties such as the purpose and location of the observed data, as well as
-how the data is described. For this reason, this work proposes a set of data
-persistence taxonomies based on the state of the art, which can be used to
-classify the properties of the produced raw data. In order to evaluate the
-proposed taxonomies, this work designs and implements a data persistence
-component for NetBEAMS, a modular-based sensor network infrastructure
-developed to improve the operation of the SF-BEAMS environmental sensor
-network. SF-BEAMS is deployed at the Tiburon Island, California, and managed by
-the Department of Biology from SFSU. In this way, based on empirical analysis
-regarding proposed taxonomies for the case study, this work selected mongoDB,
-a schema-less document-oriented database instead of the traditional use of
-relational databases. As a result based of the experiments conducted, this
-work proposes a novel use of the Key-Value Pair databases in order tor provide
-External of Data-Centric approaches. Similarly, this work also agrees to
-related literature that the use of programming languages are better abstraction
-than the use of traditional query languages used to extract data from Relational
-databases, and therefore, it should be the proposed approach to non-technical
-researchers such as the Biologists. Finally, this work proposes different
-future works such a data-centric approach using Databasese Shards to store the
-collected data and Map Reduce to execute parallel queries over distributed
-database nodes.} \vspace*{\fill}
+how the data is described. For this reason, this work's first contribution is
+a set of data persistence taxonomies based on the state of the art of Data
+Persistence for Sensor Networks, which can be used to classify the properties
+of the produced raw data. In order to evaluate the proposed taxonomies, a data
+persistence component for NetBEAMS was designed and implemented. NetBEAMS is a
+component-based sensor network infrastructure developed to improve the
+operation of the SF-BEAMS environmental sensor network. SF-BEAMS is deployed
+in Tiburon, California, and managed by the Romberg Tiburon Center (RTC). Based
+on an empirical analysis regarding proposed taxonomies, a Key-Value Data Model
+is proposed as an alternative to the Relational Data Model traditionally used.
+Furthermore, the mongoDB database, a schema-less document-oriented database,
+was selected for evaluation. Results based on the experiments suggest a novel
+approach to provide External or Data-Centric persistence for networks.
+Similarly, the literature supports programming languages as a better
+abstraction when it comes to data access and modification by non-technical
+users such as biologists. Finally, this report lends itself to future work in
+the area of data persistence in sensor networks.} \vspace*{\fill}
 
+\textbf{Keywords:} Environmental Sensor Network, Data Persistence, Taxonomies,
+KVP Data Model, Document-Oriented Databases, KVP Databases, SF-BEAMS, NetBEAMS
 \hspace*{\fill}
-
 \noindent
-I certify that the Abstract is a correct representation of the content of this thesis.
+I certify that the Abstract is a correct representation of the content of this
+report.
 
@@
 \vspace{.5in}
+
+\noindent
+\textit{``The brick walls are not there to keep us out. The brick walls are there
+to give us a chance to show how badly we want something. Because the brick walls
+are there to stop the people who don't want it badly enough.''} Dr. Randy
+Pausch
+
 \begin{quote}
 \noindent
-Acknowledgments
+
+As Dr. Randy Pausch's last lecture reminded me after the first quarter of the
+program, there are purposes for the difficult things in life as the brick walls
+will always be standing one after the other in course of your life. As my
+childhood dream was to become a scientist and study in America, I can say that
+I might have crossed a good percentage of those brick walls. This program gave
+me the wings necessary to keep dreaming and believing that anything is
+possible. You've just got to dream.
+
+I have been fortunate in these twenty-four months to meet many people who have
+given me more of their personal and professional time, their companionship,
+their help, and above all, their patience. They all contributed, directly or
+indirectly, for the success of this work, which I kindly refer to as one of my
+childhood dreams.
+
+The academic professors were fundamental in different moments during this
+program. First, I would like to thank my adviser, Dr. Arno Puder, who not only
+had to put up with the times my laptop PC broke and I delayed the delivery of
+components, but also the times I was just late for the meetings. He was the 
+first one who believed on my work in the program and accepted me in the
+team. A big thanks to him, someone whose expertise and professionalism I have
+learned to admired. Dr. Dragutin Petkovic also played an important role
+throughout my life during the program, advising me about work schedule,
+interviews, etc. Finally, Prof. Murphy offered the hardest classes of the
+program, and I must say that my writing skills were improved because of her
+class.
+
+There are four fellow researchers whom I would like to specifically thank:
+Kleber Sales, who have developed the core of NetBEAMS, and for many times
+explained and helped me during development. Many were the afternoons and weeks
+of revisions to have a working demo. Teresa Johnson was also important for the
+development of NetBEAMS and provided her feedbacks. William Murad helped me a
+lot during the classes of Biology and Algorithms, while Ivan Weiss was a great
+co-worker during the Operating Systems class developing TOS as one of the best
+teammates during school time.
+
+A special word of gratitude goes to the ones who made everything possible.
+My co-workers from CollabNet, in special for Richard, Jack, and Adam because
+their great recommendations, and to Connie, Edgar, Yiping, Norah, Kathryn and
+Andres (Facebook) who I had more contact with during the first and second
+times I worked over there as a full-time and an Intern, respectively.
+Subversion is part of my life today! Also, Eliot from 10gen for the
+innumerous days and nights on the #mongodb IRC channel and mailing list,
+helping me figure out the value of mongoDB. This work was only presented
+because of the help of three VERY important people I believe are important to
+me: Niki is the angel that appears from nowhere to salve you! I decided to
+start the program because I was very welcomed by her. During the entire
+program she was always there for each of us, I am sure, and now towards the
+end she stepped up and started reading and fixing the entire chapters. Niki,
+there are no words to thank you. In addition, Ivan also kindly offered his
+help to review the technical chapters for the gramatical errors. Man, I'm very
+lucky to have you around. And finally Kleber, who suffered with the first
+drafts of this work by reading so many errors and gramatical mistakes. I thank
+him a lot for the help and I will definitely have a better version for him from
+now on!
+
+Moving towards to more personal acknowledgements, I would like to execute a big
+of annotated thanks towards all my family and friends. They are roommates,
+friends, landlords, distant friends, that I have lived with, had been with,
+etc. These showed they cared about me (at least I felt it) in
+what I wanted to accomplish. My gratitude to people here in the California
+goes to Rhea, Sonia, Demian and Maria, Eduardo, Paula, Amanda and Eric,
+Leanna, Ivan, Julie, Diane and in special to Michelle, who believed on me
+since even before being in the program. The ones in Seattle who are also worth
+mentioning are Eduardo and Renata, for receiving me during the attempts to
+interview with Microsoft, and Wade and Fran for their friendship and support.
+When it comes to the West Coast, a lot of important people are in New York and New
+Jersey. Patricia is always available to talk about family, all the guys from
+the Focolare Movement, who I haven't been physically in contact, but online.
+Ricardo for being the best roommate and close friend, and finally Regina from
+Boston, who have belived in my potential since I first arrived in this
+country. My second-degree cousin Thamires with whom I have enjoyed hanging out
+in the night life of San Francisco!
+
+\medskip
+
+I am, for sure, particularly indebted to my parents and my brothers for their
+monumental, unwavering support and encouragement during these months of
+dedication. For all my anties, uncles, cousins (Kaline, Nicinho and
+Fernanda), they all have truly always been there for me since even before the
+program started. Without my mom and dad, or my brothers Leandro and Thiago,
+none of this dream would have been even possible remotely. I love you all and
+I will always remember your support during these two years of less phone calls
+and visits. Some brick walls were hard to climb, some of them were even more
+difficult, but none of them were impossible, because I have faith in the God
+of the impossible, who helped me stand up when I fell so many times. That's
+why I thank God Jesus Christ for bringing another of my oldest childhood
+dreams alive.
 \end{quote}
+
+\textit{``It's not about how to achieve your dreams. It's about how to lead
+your life. If you lead your life the right way, the karma will take care of
+itself. The dreams will come to you."} Dr. Randy Pausch
 

### 2009-12-17 docs/marcello-dissertation/thesis/chapter8-experimental-results-correct-behavior-and-performance.tex
Fixing the header of the table of the operations. Replacing "Insert" by "Create" to follow the pattern CRUD (Create, ...)
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@712 559edde2-a4d4-44e0-96fc-4035e40d7396
Feedback from Prof. Murphy after the presentation, adding the hardware specifications.
git-svn-id: svn://localhost/netbeams/branches/marcello/persistence@718 559edde2-a4d4-44e0-96fc-4035e40d7396

@@
         \textbf{Scenario Type} & \textbf{Language} & \textbf{Method Call} & \textbf{Use Case}\\\hline 
-        \textbf{Insert} & Java & db.SondeDataContainer.insert() & R0\\\hline
+        \textbf{Create} & Java & db.SondeDataContainer.insert() & R0\\\hline
         \textbf{Retrive} & Javascript & db.SondeDataContainer.find() & R1, R2, R3\\\hline 
@@
 chosen. The performance shell application ``iostat'' was used to verify the
-behavior, and the use of mongoDB was switched to the host operating system.
+behavior, and the use of mongoDB was switched to the host operating system,
+whose hardware specifications are summarized on Table \ref{tab:mac-os-setup}.
+
+\begin{table}[!h] 
+    \label{tab:mac-os-setup}
+    \caption{Experiment Host Machine Hardware Specifications}
+    \begin{center}
+        \begin{tabular}{|c|c|}\hline 
+        \textbf{Model Name} &  MacBook Pro\\\hline
+        \textbf{Model Identifier} & MacBookPro5,3\\\hline 
+        \textbf{Processor Name} & Intel Core 2 Duo\\\hline 
+        \textbf{Processor Speed} & 2.66 GHz\\\hline 
+        \textbf{Number Of Processors} & 1\\\hline 
+        \textbf{Total Number Of Cores} & 2\\\hline 
+        \textbf{L2 Cache} & 3 MB\\\hline 
+        \textbf{Memory} & 8 GB\\\hline 
+        \textbf{Bus Speed} & 1.07 GHz\\\hline
+        \end{tabular}
+    \end{center}
+\end{table}
 

### 2020-09-06 docs/marcello-dissertation/thesis/gen.sh
Generate thesis.pdf from docker-compose: latex
It's very hard to generate latex pdf in a host.
Using docker, we can do it quickly.
A couple of missing logs are hard to recover from
the PDF from http://citeseerx.ist.psu.edu/viewdoc/download?doi=10.1.1.469.7337&rep=rep1&type=pdf
Added the architecture diagram from taking a screenshot.

docker-compose up generates the pdf

@@ -1,0 +1,4 @@
+make clean
+make
+cp thesis.pdf thesis-review.pdf
+make clean
